### 2018-02-26 2lorem.tex
Replace old ugly latex template with new one.

\chapter{Lorem}
\label{section:lorem}

Cras tincidunt bibendum erat, vel tincidunt diam porttitor aliquam. Donec sit
amet urna non felis placerat pharetra. Aenean ultrices facilisis nulla vitae
semper. Nullam non libero quis dui fermentum aliquam id vel eros. Praesent
elementum tortor quis sem congue iaculis sit amet eget nisl. Quisque erat
tortor, condimentum eu volutpat et, blandit et augue. Phasellus erat turpis,
pretium non feugiat id, posuere id velit. Vestibulum ut sapien felis, quis
convallis dui.

\section{Etiam vitae}
\label{section:lorem:etiam}

In id fringilla velit. Maecenas sed ante sit amet nisi iaculis bibendum sed vel
elit. Quisque eleifend lacus nec ipsum lobortis ornare. Nam lectus diam,
facilisis eget porttitor ac, fringilla quis massa. Phasellus ac dolor sem, eget
varius lacus. Sed sit amet ipsum eget arcu tristique aliquam. Integer aliquam
velit sit amet odio tempus commodo. Quisque commodo lacus in leo sagittis vel
dignissim quam vestibulum. Cras fringilla velit et diam dictum faucibus.
Pellentesque at eros non mauris auctor euismod. Nullam convallis arcu vel lectus
sollicitudin rutrum. Praesent consequat, nisl at pretium posuere, neque arcu
dapibus lacus, ut sollicitudin elit velit ultricies libero~\cite{Author:06}.
Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia
Curae;

Nulla semper hendrerit molestie. Pellentesque blandit velit sit amet est
vestibulum faucibus. Nullam massa turpis, venenatis non mollis fringilla, mattis
et diam. Fusce molestie convallis elementum. Morbi nec lacus dapibus arcu mollis
gravida. Aliquam erat volutpat. Nam vitae magna nunc. Nunc ut ipsum at massa
porttitor vestibulum. Praesent diam lorem, ultrices nec vestibulum id, volutpat
nec lacus.

\begin{figure}[ht]
	\centering
	\input{images/test1.tex}
	\caption{Another figure}
	\label{figure:another}
\end{figure}

\section{Quisque arcu orci}
\label{section:lorem:quisque}

Pellentesque hendrerit aliquet placerat. Sed tincidunt lobortis eros, ut laoreet
lorem facilisis sit amet. Suspendisse ultrices, orci nec rhoncus hendrerit,
turpis augue luctus dolor, eget dignissim urna urna ac velit. Vivamus ligula
lorem, consequat vel iaculis et, lacinia in arcu. Nam imperdiet faucibus
lobortis. Ut quis libero sapien. Praesent odio est, sodales ut commodo vitae,
ultrices vel nibh. Etiam quis orci tincidunt sapien mattis aliquam. Maecenas
consequat, odio quis convallis ullamcorper, nibh justo elementum nulla, eu
molestie enim felis id neque. Sed rutrum lobortis velit, non facilisis neque
sodales et. Nunc nibh purus, gravida non ultricies in, suscipit et erat. Mauris
dui ante, fringilla vitae auctor eu, mattis a quam. Quisque at lectus
orci~\ref{table:sometable}.

\begin{table}[ht]
	\begin{center}
		\caption{Some table}
		\begin{tabular}{@{}lllr@{}}
			\toprule
			AA & BB & CC & Size \\% (\si{\kilo\byte})  \\
			\midrule
			baz & bar & OFF & 128 \\
			aaa & dd & ON & 4 \\
			bbb & efg & OFF & 64 \\
			ccc & hij & ON & 96 \\
			\bottomrule
		\end{tabular}
		\label{table:sometable}
	\end{center}
\end{table}

Mauris risus sem, sodales quis lobortis ac, vehicula sit amet leo. Sed interdum
scelerisque tellus vitae porta. Morbi eros urna, luctus ac adipiscing ut, luctus
eleifend turpis. Sed in ligula risus, ac posuere dui. Aenean bibendum est felis.
Suspendisse a pretium quam. Sed quam lectus, semper sed pulvinar eu, condimentum
id augue.~\cite{Doe:2009}

### 2018-02-27 0_intro.tex
Add first draft thesis structure.

\chapter{Introduction}
\label{section:introduction}

problems of concurrency

necessity of Weak Memory Models
\begin{itemize}
\item hardware wmms
\item wmms of programming languages
\item specific wmms like for kernel
\end{itemize}

wmm as a formal way to define guaranties that a hw, programming language, execution environment provide for programmers.

considering wmm as a set of allowed behaviours, the latter wmms are the supersets

wmm allows and disallows optimisations: partial sync of memory buffers, out-of-order execution (reordering), <more> => more behaviours that are unallowed in SC.

question possible to answer with wmm: which behaviours (in addition to SC) are allowed? which new states are allowed? Consequently, correctness, absence of data races, deadlocks or portability issues, etc.

existing tools: herd, diy7 -- exhaustive search approach for exploring the state space.

another approach: using smt solver, e.g. for answering questions like

base paper: aims to investigate portability of small programs written in a C-like pseudocode and provides the proof-of-concept tool PORTHOS [link]. As input, it takes a program and two memory models in CAT language. Then it encodes programs and memory models into an smt formula and tries to solve it via z3. Current thesis aims to extend this tool functionality to process the real C code, therefore it proposes different modula program architecture and multiple optimisations.

\section{Thesis structure}
\label{section:introduction:structure}

...

### 2018-02-27 1_wmm.tex
Add first draft thesis structure.

\chapter{Weak Memory Models}
\label{section:wmm}

briefly what it is, as in Intro

Some examples of what wmms allow to do

Lamport's sequentially consistent memory model -- global order "a global order, even if they are actually executed in different threads running in parallel in different processors. In this model, each read from some memory location is guaranteed to see a value written by the last write to this location." ([c11 by natasha]). Produces same result as sequential program.

\section{The CAT language}

the CAT language [ref to Jade's paper] (hard part: to decipher the Alglave's paper).

the event representation

\section{Examples of WMM}

briefly known hw memory models: X86-TSO, Alpha, POWER, -- ref to Jade;
language memory models: Java, C++;
library-level kernel memory model, ref to github with tests

Relationship between different models %http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks

\section{The CAT language}
\label{section:wmm:cat}

### 2018-02-27 2_impl.tex
Add first draft thesis structure.

\chapter{Implementation}
\label{section:impl}

This chapter describes the architecture of the tool \texttt{mousquitaires} ...

language: java

\section{Program Requirements}
\label{section:impl:requirements}

- stability (tests)
- scalability (new features of language, new models, new tasks for a program -- e.g.!)
- transparency
- efficiency

\section{Program Components}
\label{section:impl:comp}

Big view

\subsection{C11 to YTree parser}
\label{section:impl:comp:ytree}

- The language-dependent syntax tree:
        - for now it's the C subset language which I called 'Cmin'; as a base, I used the C11 grammar from ANTLR github repository, then I simplified it a lot, cutting off many unnecessary C syntax features and making it more convenient for parsing. When developing the Cmin language, I kept in mind C elements that are necessary for processing the linux kernel code, though for now not the whole grammar element described in file 'Cmin.g4' are being implemented;
        - later I am going to add the litmus grammar as well;
        - in future, it will be not a problem to add any new C-like language;

- The language-independent abstract syntax tree (aliased 'Ytree', where 'Y' resembles branching of the tree):
        - all tree nodes in my code are prefixed with 'Y', see tentative (yet almost complete) class hierarchy in picture 'YEntity.png';
        - this AST contains very basic language elements according to the C execution model (statements and expressions);
        - converting the language-dependent syntax tree to the language-independent syntax tree is performed by Visitor pattern (e.g., for Cmin->Ytree conversion is made by 'CminToYtreeConverterVisitor')
        - minor changes are performed by converting to ytree representation: desugaring the target code, etc.

\subsection{YTree to XGraph event converter}
\label{section:impl:comp:xgraph}

- Then, the AST is being interpreted and converted to event-based representation (aliased 'Xrepr' for eXecution representation):
        - more low-level code representation (or high-level assembly);
        - I try to keep this representation close to the one you described in your papers: basic load \& store events, branching events, fence events;
        - this representation is being implementing these days, I've just started doing it (see current class hierarchy in the picture 'XEntity.png');

- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them):
        - converting to SSA form as one of necessary steps before encoding;
        - (more? -- I'm not thinking about it yet);

\subsection{XGraph to ZFormula (SMT) encoder}
\label{section:impl:comp:zformula}

- Then, this modified event-representation is being encoded to SMT formula and sent to the solver.

\section{Optimisations}

... in each stage

### 2018-03-06 2_impl.tex
add chapter on loop unrolling

@@
 - stability (tests)
-- scalability (new features of language, new models, new tasks for a program -- e.g.!)
+
+- scalability (new features of language, new models, new tasks for a program )
+
 - transparency
+
 - efficiency
@@
 \label{section:impl:comp}
-
 Big view
 
-\subsection{C11 to YTree parser}
-\label{section:impl:comp:ytree}
+\section{C11 to YTree parser}
+\label{section:impl:ytree}
+
+below: mostly mock text.
 
@@
 
-\subsection{YTree to XGraph event converter}
+\section{YTree to XGraph event converter}
 \label{section:impl:comp:xgraph}
@@
 
-\subsection{XGraph to ZFormula (SMT) encoder}
+\subsection{Loop unrolling}
+The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}).
+The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).
+However, the cyclic flow graph cannot be encoded into SMT formula since ...
+//TODO:REFERENCE.%TODO
+
+Although there are multiple techniques for converting cyclic directed graphs into DAGs (Directed Acyclic Graphs), most of them are either not applicable to the problem of encoding graph into SMT formula, or work only for special cases (e.g., only for single-nested loops).
+// TODO: examples%TODO
+
+In the work~\cite{Porthos17}, the authors propose using the \textit{loop unrolling} technique to solve this problem, however the loops in flow graph are restricted syntactically by using only \texttt{while} statement without loop-breaking statements.
+The general case (that uses jumps to arbitrary instructions, which may be produced by \texttt{C} instruction \texttt{goto}) requires more thorough control-flow analysis.
+%The following paragraphs will explain in the details of unrolling algorithm used in current work.
+
+\subsubsection{Identifying the loops}
+
+In order define the unrolling algorithm, we need to introduce some graph-theoretical terminology used in compiler theory.
+All edges of a flow graph may be divided into three categories: \textit{advancing edges} that go from a node to its proper descendant in the tree constructed by the depth-first traverse starting at the source node; \textit{retreating edges} that go from a node to its ancestor in the tree, and \textit{cross edges} (edges that lead from a node to another node that is neither its descendant nor ancestor)~\cite{aho2007compilers::loop_unrolling}.
+%The node \textit{n} of a flow graph \textit{dominates} the node \textit{d} iff every path from the entry node of the flow graph to \textit{n} goes through \textit{d}.
+In a control flow graph, branching statements (\texttt{if-then-else}, \texttt{switch}) introduce cross edges in the graph, while other control flow statements (\texttt{while}, \texttt{do-while}, \texttt{for}, \texttt{continue}, \texttt{break}, \texttt{goto}, etc.) introduce retreating edges.
+
+If the head of retreating edge dominates its tail (i.e., \textit{every} path from head to tail goes through head), such edge is called \textit{back edge}.
+If all retreating edges of a flow graph are back edges, such graph is called \textit{reducible} (otherwise \textit{irreducible}).
+As it was noted in ~\cite{aho2007compilers::loop_unrolling}, the flow graphs that does not use jumps to arbitrary instructions are reducible.
+However, it is not hard to transform the irreducible graph to the reducible equivalent by cloning the node incident to two back edges to two nodes as it is shown in Figure~\ref{fig:irr-to-r}.
+Therefore, the loop unrolling algorithm can assume that the flow graph is reducible.
+%All edges of a flow graph may be divided into four categories: \textit{tree edges} (edges of the tree constructed by the depth-first traverse starting at the source node), \textit{forward edges} (edges that lead from a node to its proper descendant that is not a tree edge), \textit{back edges} (edges that lead from a node to its ancestor), and \textit{cross edges} (edges that lead from a node to another node that is neither its descendant nor ancestor)~\cite{Dasgupta:2006:ALG:1177299::dfs}.
+Obviously, every back edge forms the loop in the flow graph -- the path ending in its start node, that can be represented as set of nodes forming this path.
+
+% todo: move Aho's chapters names to the name of reference
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[draw=none] (invis) {};
+\node[c] (1) [above=0.5cm of invis] {$1$};
+\node[c] (2) [below left of=invis] {$2$};
+\node[c] (3) [below right of=invis] {$3$};
+\path[->]
+(1) edge [bend right=10] node {} (2)
+(1) edge [bend left=10] node {} (3)
+(2) edge [bend left=15] node {} (3)
+(3) edge [bend left=15] node {} (2)
+;
+
+\node[draw=none] (invis) [right=3cm of invis] {$\longmapsto$};
+
+\node[draw=none] (invis) [right=3cm of invis] {};
+\node[c] (1) [above=0.5cm of invis] {$1$};
+\node[c] (2) [below left of=invis] {$2$};
+\node[c] (3) [below right of=invis] {$3$};
+\node[c] (2') [right=1.5cm of 3] {$2'$};
+\path[->]
+(1) edge [bend right=10] node {} (2)
+(1) edge [bend left=10] node {} (3)
+(2) edge [bend right=10] node {} (3)
+(3) edge [bend left=15] node {} (2')
+(2') edge [bend left=15] node {} (3)
+;
+\end{tikzpicture}
+\caption{Transforming irreducible flow graph into reducible graph~\cite{aho2007compilers::handling_irreducible}} \label{fig:irr-to-r}
+\end{figure}
+
+The \textit{loop} (or \textit{natural loop}) in a program is the set of nodes that form a path in a flow graph with a single single-entry node (\textit{loop head}) which dominates all other loop nodes, and single back edge that goes to the head.
+The more precise definition of loop allows merging some cycles, caused by the set of back edges $E_i = \{(v_i, w_i) \ | \ (v_i, w_i) \in G.E \land w_i = h \}$ that point to the same head and having non mutually-nested set of nodes, into single loop:
+
+%The loop in a program is the set of nodes reachable in the reversed graph from the origin of the back edge in the original flow graph.
+
+\begin{definition}
+The \textit{loop} in a flow graph $G = (V, E)$ caused by some set of back edges $E_i \subset G.E$ is the tuple $l_i = (h, N, T, R)$ where
+\begin{itemize}
+\item the loop \textit{head} $h \in G.N$,
+\item the set of \textit{loop nodes} $N = \{ u \ | \ u$ is reachable in reversed graph $G^{-1}$ from origins of back edges $\{w \ | \ (w, h) \in E_i\}$ such that the reachability search does not go through the loop head $h \}$,
+\item the set of \textit{tails} of the loop $T \subseteq G.N$ such that $T = \{w \ | \ (w, h) \in G.E\}$
+\item the set of \textit{reference nodes} $R \subseteq G.N$ such that $R = \{(u,v) \ | \ u \in l_i.N \land v~\notin~l_i.N\}$
+\end{itemize}
+
+\end{definition}
+
+This constructive definition of the loop directly leads to the loop searching algorithm~\cite{aho2007compilers::loop_unrolling}:
+
+\begin{algorithm}
+\caption{Algorithm for finding loops in the flow graph}\label{alg:find-loops}
+\algorithmicrequire{The flow graph $G$ with the source node $v_s$}
+\algorithmicensure{Set $Loops$ of discovered loops $loop = (h, T, R, N)$ where $h$ is the head of loop, $T$ is the set of tails of the loop, $N$ is the set of all nodes of the loop, $R=\{(u,v)| u \in N \land v \notin N \}$ is the set of edges (references) outside the loop}
+\begin{algorithmic}[1]
+\Function{FindLoops}{G}
+\State $E_b \gets$ all back edges of the graph $G$
+\ForAll {back edge $(v_i, h_i) \in E_b$}
+  \State mark the head $h_i$ as visited \Comment{do not go outside the loop head}
+  \ForAll {edge $(u, w^{-1}) \in G^{-1}.E$ explored while traversing the reversed graph $G^{-1}$ in depth-first order starting from node $v_i$}
+    \State remember visited node $u$ in the set $N_i$
+    \State $merged \gets false$
+    \ForAll {already collected loop $l \in Loops$}
+      \If {$(l.h = h_i) \land (l.N \setminus N_i \neq \emptyset \land N_i \setminus l.N \neq \emptyset)$} 
+        \newline \Comment{merge these two loops into one:}
+        \State $l.N \gets l.N \cup N_i$ \label{alg:find-loops:merge-loops-1}
+        \State $l.T \gets l.T \cup T_i$ \label{alg:find-loops:merge-loops-2}
+        \State $merged \gets true$
+        \State \textbf{break}
+      \EndIf
+    \EndFor
+     \If {not $merged$}
+        \State add $l_i = (h_i, T_i, R_i, N_i)$ to $Loops$
+      \EndIf
+  \EndFor
+  \newline \Comment{fill in the set $l_i.R$ of references outside the loop:}
+  \ForAll {loop node $v \in l_i.N$}
+    \ForAll {edge $(v, w) \in G.E$}
+      \If {$w \notin l_i.N$}
+        \State add $v$ to the set $l_i.R$ of references outside the loop
+      \EndIf
+    \EndFor
+  \EndFor
+\EndFor
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+Examples of possible mutual arrangements of two nested loops in a flow graph are shown in Figure~\ref{fig:nested-loops} (the back edges are pictured with dotted lines).
+In the first example~\ref{fig:nested-loops-1}, the loop $l_{2 \rightarrow 3}$ caused by back edge $2 \rightarrow 3$ has the set of nodes $\{2, 3\}$, while the loop $l_{4 \rightarrow 1}$ has the set of nodes $\{1,2,3,4\}$, which is a superset of the set of nodes of $l_{2 \rightarrow 3}$, therefore the latter loop is the outer loop of the former one, and it should be unrolled firstly.
+The same reasoning applies to all other examples~\ref{fig:nested-loops-2} -- \ref{fig:nested-loops-4}.
+
+Figure~\ref{fig:merged-loop} illustrates two cycles $1\rightarrow2\rightarrow3$ and $1\rightarrow2\rightarrow4$ with the same head $1$ that can be safely merged into one loop $\{1,2,3,4\}$ with head $1$ (lines~\ref{alg:find-loops:merge-loops-1} and~\ref{alg:find-loops:merge-loops-2} of Algorithm~\ref{alg:find-loops}) and interpreted as a single loop with branching expression at the head node.
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.18\textwidth}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below of=1] {$2$};
+\node[c] (3) [below of=2] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [] node {} (3)
+(3) edge [] node {} (4)
+(3) edge [bend left=40,dotted] node {} (2)
+(4) edge [bend left=50,dotted] node {} (1)
+;
+\end{tikzpicture}
+\caption{} \label{fig:nested-loops-1}
+\end{subfigure}
+\hspace*{\fill}
+\begin{subfigure}{0.18\textwidth}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below of=1] {$2$};
+\node[c] (3) [below of=2] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [] node {} (3)
+(3) edge [] node {} (4)
+(3) edge [bend right=50,dotted] node {} (1)
+(4) edge [bend left=50,dotted] node {} (1)
+;
+\end{tikzpicture}
+\caption{} \label{fig:nested-loops-2}
+\end{subfigure}
+\hspace*{\fill}
+\begin{subfigure}{0.18\textwidth}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below of=1] {$2$};
+\node[c] (3) [below of=2] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=40,dotted] node {} (2)
+(4) edge [bend left=50,dotted] node {} (1)
+;
+\end{tikzpicture}
+\caption{} \label{fig:nested-loops-3}
+\end{subfigure}
+\hspace*{\fill}
+\begin{subfigure}{0.18\textwidth}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below of=1] {$2$};
+\node[c] (3) [below of=2] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=40,dotted] node {} (2)
+(3) edge [bend left=40,dotted] node {} (1)
+;
+\end{tikzpicture}
+\caption{} \label{fig:nested-loops-4}
+\end{subfigure}
+\hspace*{\fill}
+\begin{subfigure}{0.18\textwidth}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below of=1] {$2$};
+\node[c] (3) [below of=2] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [] node {} (3)
+(3) edge [] node {} (4)
+(2) edge [bend left=40,dotted] node {} (1)
+(3) edge [bend left=40,dotted] node {} (2)
+(4) edge [bend left=40,dotted] node {} (3)
+;
+\end{tikzpicture}
+\caption{} \label{fig:nested-loops-5}
+\end{subfigure}
+\caption{Examples of possible mutual arrangements of two nested loops in a flow graph} \label{fig:nested-loops}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below left of=1] {$2$};
+\node[c] (3) [below right of=1] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [bend left=50,dotted] node {} (1)
+
+(1) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=80,dotted] node {} (1)
+;
+\end{tikzpicture}
+\caption{Example of flow graph with multiple cycles merged into a single loop} \label{fig:merged-loop}
+\end{figure}
+
+\subsubsection{Binding unrolled loops}
+
+Once the set of loops of the flow graph is defined, they need to be unrolled and bounded into single graph.
+Unrolling is being performed up to the user-defined bound $k$, this means every loop is being unwound to the acyclic subgraph of size at most $k$.
+This bound is defined separately for each collected loop, so that the $l$-nested loop has at most $k^l$ nodes.
+Note that the loop unrolling does not require any data-flow information: the reachability analysis of a particular node in the unrolled graph will be performed later by the SMT solver.
+This is rather rough underapproximation, which is to be handled by user by specifying the unrolling bound $k$.
+
+The algorithm for unrolling and binding nested loops is as following:
+
+\begin{algorithm}
+\caption{Algorithm for unrolling and binding nested loops}\label{alg:bind-loops}
+\algorithmicrequire{The ordered list $Loops$ of nested loops $l_i$ such that $l_1.N \subset l_2.N \subset ... l_t.N$}
+\algorithmicensure{Set $Loops$ of discovered loops $loop = (h, N)$ where $d$ is the head of loop, $L$ is the set of nodes of the loop}
+\begin{algorithmic}[1]
+\Function{UnrollAndBindLoops}{$Loops$}
+\State //to be done
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\begin{figure}1
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below left of=1] {$2$};
+\node[c] (3) [below right of=1] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [bend left=50,dotted] node {} (1)
+(1) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=80,dotted] node {} (1)
+;
+\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
+;
+\node[c] (21) [right=3cm of impl] {$2_1$};
+\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
+\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
+\node[c] (41) [below of=31] {$4_1$};
+\node[c] (12) [below of=21] {$1_2$};
+\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
+\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
+\node[c] (13) [below of=41] {$1_3$};
+\node[c] (14) [below of=22] {$1_4$};
+\node[c] (43) [below of=32] {$4_3$};
+\node[c] (23) [below of=13] {$2_3$};
+\node[c] (33) [below of=13] {$3_3$};
+\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
+\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
+\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
+\node[c] (44) [below of=33] {$4_4$};
+\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
+\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
+\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
+\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
+\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
+\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
+\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
+\path[->]
+(11) edge [] node {} (21)
+(11) edge [] node {} (31)
+(31) edge [] node {} (41)
+(21) edge [] node {} (12)
+(12) edge [bend right=10] node {} (22)
+(12) edge [bend left=10] node {} (32)
+(41) edge [] node {} (13)
+(22) edge [] node {} (14)
+(32) edge [] node {} (43)
+(13) edge [] node {} (23)
+(13) edge [] node {} (33)
+(14) edge [bend right=10] node {} (24)
+(14) edge [bend left=10] node {} (34)
+(43) edge [] node {} (15)
+(33) edge [bend right=10] node {} (15)
+(33) edge [bend left=10] node {} (44)
+(24) edge [bend right=20] node {} (6)
+(34) edge [] node {} (6)
+(15) edge [] node {} (6)
+(44) edge [bend left=20] node {} (6)
+;
+\end{tikzpicture}
+\label{fig:loop-unwind}
+\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop}, unwinded up to the bound $k = 6$} 
+\end{figure}
+
+\section{XGraph to ZFormula (SMT) encoder}
 \label{section:impl:comp:zformula}
@@
 
-... in each stage
+... performed on each stage

### 2018-03-06 test.tex
add chapter on loop unrolling

--- 2018-02-27 2_impl.tex
@@
-\chapter{Implementation}
-\label{section:impl}
+\documentclass[12pt,a4paper,oneside,pdftex]{report}
 
-This chapter describes the architecture of the tool \texttt{mousquitaires} ...
+\usepackage{subcaption}
+\usepackage{graphicx}
+\usepackage{amsmath}
 
-language: java
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+\usetikzlibrary{calc}
+\usetikzlibrary{arrows}
+\usetikzlibrary{decorations.pathmorphing,decorations.markings}
+\usetikzlibrary{shapes}
+\usetikzlibrary{patterns}
+\usetikzlibrary{automata}
 
-\section{Program Requirements}
-\label{section:impl:requirements}
+\tikzset{
+c/.style={
+  circle,
+  inner sep=0pt,
+  text width=6mm,
+  align=center,
+  draw=black,
+  fill=white
+  }
+}
 
-- stability (tests)
-- scalability (new features of language, new models, new tasks for a program -- e.g.!)
-- transparency
-- efficiency
+\begin{document}
 
-\section{Program Components}
-\label{section:impl:comp}
+\begin{figure}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below left of=1] {$2$};
+\node[c] (3) [below right of=1] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [bend left=50,dotted] node {} (1)
+(1) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=80,dotted] node {} (1)
+;
+\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
+;
+\node[c] (21) [right=3cm of impl] {$2_1$};
+\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
+\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
+\node[c] (41) [below of=31] {$4_1$};
+\node[c] (12) [below of=21] {$1_2$};
+\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
+\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
+\node[c] (13) [below of=41] {$1_3$};
+\node[c] (14) [below of=22] {$1_4$};
+\node[c] (43) [below of=32] {$4_3$};
+\node[c] (23) [below of=13] {$2_3$};
+\node[c] (33) [below of=13] {$3_3$};
+\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
+\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
+\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
+\node[c] (44) [below of=33] {$4_4$};
+\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
+\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
+\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
+\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
+\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
+\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
+\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
+\path[->]
+(11) edge [] node {} (21)
+(11) edge [] node {} (31)
+(31) edge [] node {} (41)
+(21) edge [] node {} (12)
+(12) edge [bend right=10] node {} (22)
+(12) edge [bend left=10] node {} (32)
+(41) edge [] node {} (13)
+(22) edge [] node {} (14)
+(32) edge [] node {} (43)
+(13) edge [] node {} (23)
+(13) edge [] node {} (33)
+(14) edge [bend right=10] node {} (24)
+(14) edge [bend left=10] node {} (34)
+(43) edge [] node {} (15)
+(33) edge [bend right=10] node {} (15)
+(33) edge [bend left=10] node {} (44)
+(24) edge [bend right=20] node {} (6)
+(34) edge [] node {} (6)
+(15) edge [] node {} (6)
+(44) edge [bend left=20] node {} (6)
+;
+\end{tikzpicture}
+\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop}, unwinded up to the bound $k = 6$} \label{fig:merged-loop}
+\end{figure}
 
-Big view
-
-\subsection{C11 to YTree parser}
-\label{section:impl:comp:ytree}
-
-- The language-dependent syntax tree:
-        - for now it's the C subset language which I called 'Cmin'; as a base, I used the C11 grammar from ANTLR github repository, then I simplified it a lot, cutting off many unnecessary C syntax features and making it more convenient for parsing.
-When developing the Cmin language, I kept in mind C elements that are necessary for processing the linux kernel code, though for now not the whole grammar element described in file 'Cmin.g4' are being implemented;
-        - later I am going to add the litmus grammar as well;
-        - in future, it will be not a problem to add any new C-like language;
-
-- The language-independent abstract syntax tree (aliased 'Ytree', where 'Y' resembles branching of the tree):
-        - all tree nodes in my code are prefixed with 'Y', see tentative (yet almost complete) class hierarchy in picture 'YEntity.png';
-        - this AST contains very basic language elements according to the C execution model (statements and expressions);
-        - converting the language-dependent syntax tree to the language-independent syntax tree is performed by Visitor pattern (e.g., for Cmin->Ytree conversion is made by 'CminToYtreeConverterVisitor')
-        - minor changes are performed by converting to ytree representation: desugaring the target code, etc.
-
-\subsection{YTree to XGraph event converter}
-\label{section:impl:comp:xgraph}
-
-- Then, the AST is being interpreted and converted to event-based representation (aliased 'Xrepr' for eXecution representation):
-        - more low-level code representation (or high-level assembly);
-        - I try to keep this representation close to the one you described in your papers: basic load \& store events, branching events, fence events;
-        - this representation is being implementing these days, I've just started doing it (see current class hierarchy in the picture 'XEntity.png');
-
-- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them):
-        - converting to SSA form as one of necessary steps before encoding;
-        - (more? -- I'm not thinking about it yet);
-
-\subsection{XGraph to ZFormula (SMT) encoder}
-\label{section:impl:comp:zformula}
-
-- Then, this modified event-representation is being encoded to SMT formula and sent to the solver.
-
-\section{Optimisations}
-
-... in each stage
+\end{document}

### 2018-03-07 0_intro.tex
remove written loop unrolling chapter. add firstly-first intro.

@@
 
-problems of concurrency
+Most modern computer systems contain large parts working concurrently.
+Though system parallelising can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult areas of programming~\cite{mckenney2017parallel}.
+% Most articles, presentations and books on concurrent programming start with words how hard it is.
 
-necessity of Weak Memory Models
-\begin{itemize}
-\item hardware wmms
-\item wmms of programming languages
-\item specific wmms like for kernel
-\end{itemize}
+Traditionally, studies related to concurrent programming concern mostly classical cases of implementing race-free and lock-free parallel programs, asynchronous data structures and synchronisation primitives of a programming language.
+Unfortunately,
+when it comes to 
+%in cases of 
+analysis of the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
+The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
+For instance, some processor architecture performs reordering of memory access instructions in a thread, while it should not be considered as a valid code transformation with respect to the logic of program (//TODO: example).
+It is said that this kind of processor exploits \textit{relaxed}, or \textit{weak memory model~(WMM)}, that allow certain amount of memory operations reordering while preserving consistency.
+Weak memory models serve as set of guarantees made by designers of hardware (programming language, compiler, operation environment, etc.) for programmers on which behaviours of their concurrent code they should expect.
+
+In last decades, the problem of formalising WMMs of different 
+%hardware
+architectures has been developed significantly.
+The research is aimed, firstly, at formalising the weak memory models, providing systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to these models.
+Secondly, the researchers work on extracting hardware memory models from the public specifications written in natural language and thus suffering from ambiguities and incompletenesses.
+Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to weak memory models.
+
+while the first two have been developed .... 
+The latter approach still suffers from the exponent.... (need this?)
+Kernel.
 

### 2018-03-07 2_impl.tex
remove written loop unrolling chapter. add firstly-first intro.

--- 2018-02-27 2_impl.tex
@@
 
-\subsection{XGraph to ZFormula (SMT) encoder}
+\subsection{Loop unrolling}
+The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}).
+The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).
+However, the cyclic flow graph cannot be encoded into SMT formula since ...
+//TODO:REFERENCE.%TODO
+
+\begin{figure}
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below left of=1] {$2$};
+\node[c] (3) [below right of=1] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [bend left=50,dotted] node {} (1)
+(1) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=80,dotted] node {} (1)
+;
+\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
+;
+\node[c] (21) [right=3cm of impl] {$2_1$};
+\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
+\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
+\node[c] (41) [below of=31] {$4_1$};
+\node[c] (12) [below of=21] {$1_2$};
+\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
+\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
+\node[c] (13) [below of=41] {$1_3$};
+\node[c] (14) [below of=22] {$1_4$};
+\node[c] (43) [below of=32] {$4_3$};
+\node[c] (23) [below of=13] {$2_3$};
+\node[c] (33) [below of=13] {$3_3$};
+\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
+\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
+\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
+\node[c] (44) [below of=33] {$4_4$};
+\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
+\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
+\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
+\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
+\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
+\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
+\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
+\path[->]
+(11) edge [] node {} (21)
+(11) edge [] node {} (31)
+(31) edge [] node {} (41)
+(21) edge [] node {} (12)
+(12) edge [bend right=10] node {} (22)
+(12) edge [bend left=10] node {} (32)
+(41) edge [] node {} (13)
+(22) edge [] node {} (14)
+(32) edge [] node {} (43)
+(13) edge [] node {} (23)
+(13) edge [] node {} (33)
+(14) edge [bend right=10] node {} (24)
+(14) edge [bend left=10] node {} (34)
+(43) edge [] node {} (15)
+(33) edge [bend right=10] node {} (15)
+(33) edge [bend left=10] node {} (44)
+(24) edge [bend right=20] node {} (6)
+(34) edge [] node {} (6)
+(15) edge [] node {} (6)
+(44) edge [bend left=20] node {} (6)
+;
+\end{tikzpicture}
+\label{fig:loop-unwind}
+\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop}, unwinded up to the bound $k = 6$} 
+\end{figure}
+
+\section{XGraph to ZFormula (SMT) encoder}
 \label{section:impl:comp:zformula}

### 2018-03-18 0_intro.tex
Forgotten changes for release 14.03.2018.

--- 2018-02-27 0_intro.tex
@@
 \chapter{Introduction}
-\label{section:introduction}
+\label{ch:introduction}
 
-problems of concurrency
+Most modern computer systems contain large parts that operate concurrently.
+Though system parallelising can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
+% Most articles, presentations and books on concurrent programming start with words how hard it is.
 
-necessity of Weak Memory Models
-\begin{itemize}
-\item hardware wmms
-\item wmms of programming languages
-\item specific wmms like for kernel
-\end{itemize}
+Traditionally, studies related to concurrent programming concern on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language.
+Unfortunately,
+when it comes to 
+%in cases of 
+the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
+The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
+For instance, Figure~\ref{example:simple_wmm} provides simple example of
+unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable in x86 machines (such little examples that illustrate specific behaviours of a WMM are called \textit{litmus tests}).
+%reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
+This behaviour is caused by write buffers used by all processors in x86 architecture.
+These buffers cache writes to shared variables, so that the writes to shared memory does not become visible by other processes immediately.
+In the example, the write `\texttt{MOV~[x],1}' performed in the process \texttt{P0} stores value~\texttt{1} into the shared variable~\texttt{[x]} in the write buffer of process~\texttt{P0}.
+Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread.
+More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
 
-wmm as a formal way to define guaranties that a hw, programming language, execution environment provide for programmers.
+\begin{figure}
+\small \ttfamily
+\begin{tabular}{ |l|l| }
+\hline
+\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P0 & P1 \\ \hline
+MOV [x],1 & MOV [y],1 \\
+MOV EAX,[y] & MOV EBX,[x] \\
+\hline
+%\multicolumn{2}{|l|}{locations [x;y;]} \tabularnewline
+\multicolumn{2}{|l|}{exists (0:EAX=0~/\textbackslash~1:EAX=0)} \tabularnewline
+\hline
+\multicolumn{2}{|l|}{x86-TSO: allow} \tabularnewline
+\hline
 
-considering wmm as a set of allowed behaviours, the latter wmms are the supersets
+\end{tabular}
+\caption{Litmus test of memory operations reordering allowed by the x86-TSO weak memory model}
+\label{example:simple_wmm}
+\end{figure}
 
-wmm allows and disallows optimisations: partial sync of memory buffers, out-of-order execution (reordering), <more> => more behaviours that are unallowed in SC.
+The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
+This memory model, called the \textit{sequential consistency (SC) model}, allows only those executions (interleavings) that produce the same result as if the operations had been executed by single process.
+This means that the order of operations executed by a process is strictly defined by the program it executes.
+The SC model does requires the write to a shared variable performed in one process to become visible by all other processes not instantly, but simultaneously.
+This means each process communicates to the shared memory directly, without local buffering.
+Another important requirement of SC memory model is that it forbids memory operations reordering within single process (the order is strictly defined by the program).
 
-question possible to answer with wmm: which behaviours (in addition to SC) are allowed?
-which new states are allowed?
-Consequently, correctness, absence of data races, deadlocks or portability issues, etc.
+The SC model is considered to be the strong memory model in the sence that it provides strong guarantees regarding the ordering and caused effect of memory operations.
+Different relaxations of this model lead to the class of \textit{weak memory models~(WMM)}.
+%while preserving consistency
+WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may expect.
 
-existing tools: herd, diy7 -- exhaustive search approach for exploring the state space.
+Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them.
+The state-of-the-art tool is \texttt{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK, and firstly released back in 2010.
+It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \texttt{diy7}, \texttt{diycross7} and \texttt{diyone7}, the litmus tests concrete executor \texttt{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \texttt{herd7} that implements reachability analysis for capturing states allowed by the WMM.
+The latter tool uses systematic exhaustive search in a state space, which may be inefficient when analysing real-world programs, therefore more sophisticated search techniques should be applied for this problem.
 
-another approach: using smt solver, e.g. for answering questions like
+One possible approach is to use an efficient implementation of an SMT-solver (a SAT-solver extended by satisfiability modulo theories)~\cite{Porthos17}. % <-- TODO
+This approach allows to capture symbolically the semantics of both the program and the weak memory model while encoding it into a single SMT-formula.
+Most modern SMT-solvers are efficient enough to be able to operate the state space of size~??? [CITATION-?].
 
-base paper: aims to investigate portability of small programs written in a C-like pseudocode and provides the proof-of-concept tool PORTHOS [link].
-As input, it takes a program and two memory models in CAT language.
-Then it encodes programs and memory models into an smt formula and tries to solve it via z3.
-Current thesis aims to extend this tool functionality to process the real C code, therefore it proposes different modula program architecture and multiple optimisations.
+%In the work of Finnish and German researchers ... names ...
+In the work~\cite{Porthos17}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
+Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a trivial problem of symbolic execution, encoding of the weak memory model is more tedious.
+The reason is that some relations of WMMs are defined as mutually-recursive and need to be linearised in order to be encoded into an equivalent logical formula.
+%"a bounded model checking method that reduces the portability analysis problem to satisfiability modulo theorie" [Porthos17]
+%"A portability bug is an execution that is consistent with the target but inconsistent with the source memory model.
+We capture this alternation with a single existential query.
+Consistency is specified in terms of acyclicity (and irreflexivity) of relations.
+Hence, an execution is inconsistent if a derived relation of the (source) memory model contains a cycle (or is not irreflexive)."
+
+Current thesis aims to improve the proof-of-concept tool \texttt{PORTHOS} firstly introduced in April 2017 in the work~\cite{Porthos17} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
 
 \section{Thesis structure}
-\label{section:introduction:structure}
+\label{ch:introduction:structure}
 
-...
+The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory models analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
+Chapter~...

### 2018-03-18 1_wmm.tex
add chapter on loop unrolling
Forgotten changes for release 14.03.2018.

@@
 \chapter{Weak Memory Models}
-\label{section:wmm}
+\label{ch:wmm}
 
-briefly what it is, as in Intro
+Over the last decades, the problem of formalisation of the weak memory models
+%for different software systems and hardware architectures
+has been developed significantly.
+Research of WMM aimes, firstly, to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
+//the cat language here.
 
-Some examples of what wmms allow to do
+Secondly, researchers work on extracting the formal hardware memory models from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
+Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, SPARC-TSO for Sparc architecture~\cite{?!?!?!}, much more relaxed memory model ???
+for Power and ARM architectures defined in ???~\cite{sarkar2011understanding}, Alpha (???).
+Moreover, in 2005 <who?Milner> started the work on developing the weak memory model for C++ language, which was introduced in C++11 (?) standard~\cite{???} (//todo: C11 MM ~\cite{batty2011mathematizing}).
+The memory model for Java that is based on the \textit{happens-before} principle was introduced in JVM???
+in ??<year>.
 
-Lamport's sequentially consistent memory model -- global order "a global order, even if they are actually executed in different threads running in parallel in different processors.
-In this model, each read from some memory location is guaranteed to see a value written by the last write to this location." ([c11 by natasha]).
-Produces same result as sequential program.
+Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to weak memory models.
+In this <area> the notable works <are on> defining the Linux kernel memory model <that is being actively developing these days~\cite{kernel_wmm_1}>.
+Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
+
+%wmm as a formal way to define guaranties that a hw, programming language, execution environment provide for programmers.
+%considering wmm as a set of allowed behaviours, the latter wmms are the supersets
+%wmm allows and disallows optimisations: partial sync of memory buffers, out-of-order execution (reordering), <more> => more behaviours that are unallowed in SC.
+%question possible to answer with wmm: which behaviours (in addition to SC) are allowed?
+which new states are allowed?
+Consequently, correctness, absence of data races, deadlocks or portability issues, etc.
+
+One important question that rises naturally when talking about weak memory models is whether a program is portable from one processor architecture to another. This can be solved ...<ref to PORTHOS>
+
+% more examples of what wmms allow to do?
+
+\section{The Litmus tests}
+
+Test Execution Diagrams
 
@@
 
-\section{Examples of WMM}
+\section{Some known WMM}
+
+\subsection{x86-TSO}
+\label{ch:wmm:x86}
+// exmaple with reordering
+// ex. with 
+// Rev-29 Example 7-6. Stores Are Transitively Visible. %see http://www.cl.cam.ac.uk/~pes20/weakmemory/x86tso-paper.pdf
+
+There is a barrier instruction \texttt{mfence} that may be used for flushing the buffers into the main memory.
 
@@
 
-Relationship between different models %http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks
-
-\section{The CAT language}
-\label{section:wmm:cat}
+Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}

### 2018-03-18 0_intro.tex
Add description of wmm.

@@
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
-For instance, Figure~\ref{example:simple_wmm} provides simple example of
+For instance, Figure~\ref{simple_wmm_x86} provides simple example of
 unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable in x86 machines (such little examples that illustrate specific behaviours of a WMM are called \textit{litmus tests}).
@@
 MOV [x],1 & MOV [y],1 \\
-MOV EAX,[y] & MOV EBX,[x] \\
+MOV EAX,[y] & MOV EAX,[x] \\
 \hline
@@
 \caption{Litmus test of memory operations reordering allowed by the x86-TSO weak memory model}
-\label{example:simple_wmm}
+\label{simple_wmm_x86}
 \end{figure}

### 2018-03-18 1_wmm.tex
Add description of wmm.

@@
 
-One important question that rises naturally when talking about weak memory models is whether a program is portable from one processor architecture to another. This can be solved ...<ref to PORTHOS>
+%One important question that rises naturally when talking about weak memory models is whether a program is portable from one processor architecture to another. This can be solved ...<ref to PORTHOS>
 
@@
 
-\section{The Litmus tests}
+\section{The event-based program representation}
 
-Test Execution Diagrams
+The classical approach to model the concurrent programs is to use the \textit{global time}, a single order of interleavings of all actions happened in different threads.
+However these models are easy to understand, it may be hard to consider \textit{all} possible states, number of which is exponentially large.
+Another way to do this is to to use non-deterministic computation-centric models defined in~\cite{fri97}, one of which represents the program as the graph of \textit{memory events}.
+The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independant from the order of local computation events.
+These models may be further restricted by constraints of a weak memory model, adding \textit{relations} to the memory events.
+%Perhaps, the most convenient way to model the non-deterministic properties of concurrent programs is to use the program representation based on \textit{memory events}.
+%In order to model the non-deterministic properties of concurrent programs
+
+The event-based program model represents the directed graph (\textit{event-graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events.
+An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program.
+An \textit{execution} (trace, run) of a given program is an order of events.
+%associated with the instructions of the program
+%the path in the event-graph. 
+An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
+An execution to be checked on validity is called the \textit{candidate execution}.
+
+Below we describe some basic types of events and relations.
+
+\subsection{Events}
+\label{ch:wmm:model:events}
+
+A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory.
+Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events.
+The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}).
+A memory event is specified by its direction with respect to the shared variable, its location~$\texttt{loc}(e_m)$, its processor label~$\texttt{proc}(e_m)$, and a unique event label~$\texttt{id}(e_m)$~\cite{alglave2010shared}.
+%\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
+The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
+We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\texttt{load}(r_1 \leftarrow l_2); \ e_m''~=~\texttt{store}(l_1 \leftarrow r_1)$.
+
+A \textit{computation event} $e_c \in \mathbb{C} \subseteq \mathbb{E}$, represents a low-level assembly computation operation performed solely on local-memory arguments.
+An example of computation event may be the event $e_c = r_3 \leftarrow add(r_1, r_2)$ that writes the sum of values stored in registers $r_1$ and $r_2$ to the register $r_3$.
+The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded to the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
+
+The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}).
+Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the model that restrict the set of allowed behaviours.
+Technically, a fence may either serve as a synchronisation barrier, or flush local memory caches, etc.
+
+\subsection{Relations}
+\label{ch:wmm:model:relations}
+
+The basic relation in the event-based program model is the \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of memory events \textit{within single process}, which never relates events from different processes.
+Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\texttt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
+This relation encodes the control-flow of the program into the event-graph.
+%Some new relations may be acquired : dp, po-loc
+
+The data-flow of a program is encoded by the \textit{communication relations}: the \texttt{rf}-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \texttt{co}-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \texttt{fr}-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event.
+%TODO: perephrase last sentence! alglave thesis, p. 36
+
+Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model (the picture is generated by the \texttt{herd7} tool, version 7.47).
+%The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
+%the event $d$:\texttt{(Rx=0)} reads value \texttt{0} at the shared location \texttt{x} from the initial write event $f$\texttt{(Wx=0)} (the red edge of \texttt{rf}-relation).
+
+\includegraphics[width=0.4\textwidth]{img/my/simple_wmm_x86.png}
+\label{simple_wmm_x86_pic}
+% TODO: re-generate with new name, not 'SO'
 

### 2018-03-22 0_intro.tex
started Chapter 3 on describing the portability in general and the encoding into SMT.

@@
 \chapter{Introduction}
-\label{ch:introduction}
+\label{ch:intro}
 
@@
 
+%TODO: BMC
 One possible approach is to use an efficient implementation of an SMT-solver (a SAT-solver extended by satisfiability modulo theories)~\cite{Porthos17}. % <-- TODO
 This approach allows to capture symbolically the semantics of both the program and the weak memory model while encoding it into a single SMT-formula.
-Most modern SMT-solvers are efficient enough to be able to operate the state space of size~??? [CITATION-?].
+Most modern SMT-solvers are efficient enough to be able to operate the state space of size~[CITATION-?].
 
@@
 
-Current thesis aims to improve the proof-of-concept tool \texttt{PORTHOS} firstly introduced in April 2017 in the work~\cite{Porthos17} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
+Current work aims to improve the proof-of-concept tool \texttt{PORTHOS} firstly introduced in April 2017 in the work~\cite{Porthos17} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
 
 \section{Thesis structure}
-\label{ch:introduction:structure}
+\label{ch:intro:structure}
 
@@
 Chapter~...
+
+% TODO: Add intro to the bounded reachability analysis using SAT -- ?

### 2018-03-22 1_2_encod.tex
started Chapter 3 on describing the portability in general and the encoding into SMT.

\chapter{Portability of Concurrent Software}
%\chapter{The Concurrent Software Portability Analysis as a Bounded Model Checking Problem}
\label{ch:port}

As it was discussed in Chapter~\ref{ch:intro}, the program may behave differently when compiled for different parallel hardware architectures. This can cause the portability bugs, the behaviour allowed under one architecture and forbidden under another. 
The concurrent software portability analysis may be stated as a Bounded Model Checking~(BMC) problem, which in turn can be reduced to the problem of satisfiability modulo theories (SMT)~\cite{Porthos17}.

\section{The model checking problem}
\label{ch:port:mc}

The classical model checking algorithms explore the state space of an abstract automata or a transition system in order to find states that violate the specification. The general schema of model checking is the following: firstly, the analysing system is being represented as a transition system, a finite directed graph with labeled nodes representing states of the system such that each state corresponds to the unique subset of atomic propositions, that characterise the behavioral properties of each state. 
Then, the system constraints are being defined in terms of a modal temporal logic with respect to the atomic propositions. Commonly, the Linear Temporal Logic~(LTL) or Computational Tree Logic~(CTL), along with their extensions, are used as a specification language due to the expressiveness and verifiability of their statements. 
In the described schema, the model checking problem is reducible to the reachability analysis, an iterative process of a systematic exhaustive search in the state space. This approach is called \textit{unbounded model checking (UMC)}.

However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables of the system. In case of modeling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
Therefore, the research in model checking over past 40 years was aimed at tackling the state explosion problem, mostly by optimising search space, search strategy or basic data structures of existing algorithms.

%[-- TODO: REPHRASE
One of the first major optimising technique was symbolic model checking with binary decision diagrams (BDDs). In this approach, a set of states is represented by a BDD instead of by listing each state individually~\cite{clarke2012model}.
%--]
The BDD representation can be linear of size of variables it encodes if the ordering of variables is optimal, otherwise the size of BDD is exponential. The problem of finding such an optimal ordering is known as NP-complete problem, which makes this approach inapplicable in some cases.

The other idea is to use satisfiability solvers for symbolic exploration of state space~\cite{clarke2001bounded}. In this approach, the state space exploration consists of sequence of queries to the SAT-solver, represented as boolean formulas that encode the constraints of the model and the finite path to a state in the corresponding transition system.  
%This approach uses an iterative process of constructing queries to the SAT-solver as a boolean formula which encodes the constraints of the model and the finite path to a state in the corresponding transition system. 
Due to the SAT-solver. This technique is called \textit{bounded model checking (BMC)}, because the search process is being repeated up to user-defined bound $k$, which may result to incomplete analysis in general case. However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).

%For instance, the idea of grouping states with similar properties into equivalence classes lead to the concept of traces in concurrent systems proposed by A.~Mazurkiewicz in 1986~\cite{mazurkiewicz1986trace}.

\section{The portability as a BMC problem}
\label{ch:port:enc}

A program $P$ is called portable from the source weak memory model $\mathcal{M_S}$ to the target memory model $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:

\begin{definition}[Portability]
Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
$cons_{\mathcal{M_T}}(P) \subseteq cons{\mathcal{M_S}}(P)$
\end{definition}

Note, that the formulation of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.

It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the BMC problem. The full SMT formula $\phi$ should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$. The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.
%$The \phi_{\mathcal{M_T}}$ is an encoding of the derived relations and all assertions of
%, and φ ¬M S

\subsection{Encoding of the control-flow constraints}
\label{ch:port:enc:cf}

\subsection{Encoding of the data-flow constraints}
\label{ch:port:enc:cf}

\subsection{Encoding of the memory model constraints}
\label{ch:port:enc:wmm}

### 2018-03-22 1_wmm.tex
started Chapter 3 on describing the portability in general and the encoding into SMT.

@@
 
-Over the last decades, the problem of formalisation of the weak memory models
-%for different software systems and hardware architectures
-has been developed significantly.
+%"Semantics and verification under weak memory models have been the subject of study at least since 2007." [PORTHOS]
+The \textit{weak memory model} is a set of predicative constraints on possible executions of a concurrent program.
+The study of formalisation the weak memory models for different architectures is being rapidly developed over last decade.
 Research of WMM aimes, firstly, to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
-//the cat language here.
+
+%The generic framework of program analysis under weak memory models presented in~\ref{alglave2010shared} is discussed in Section~\ref{ch:wmm:event}.
 
@@
 \section{The event-based program representation}
+\label{ch:wmm:event}
+
+%"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf
 
@@
 An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program.
-An \textit{execution} (trace, run) of a given program is an order of events.
+An \textit{execution} (trace, run) of a given program is an ordered set of events.
 %associated with the instructions of the program
@@
 
+\subsection{Executions}
+\label{ch:wmm:model:executions}
+
+The semantics of a concurrent program is represented by the set of allowed executions.
+An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$ (\textit{data-flow} of a program)~\ref{alglave2010shared}.
+As it was shown in~\ref{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
+
 \section{The CAT language}
 
-the CAT language [ref to Jade's paper] (hard part: to decipher the Alglave's paper).
+The Porthos tool uses the language \texttt{CAT}~\ref{alglave2016syntax} for defining weak memory models.
 

### 2018-03-22 0_intro.tex
Cleanup introduction.

@@
 Most modern computer systems contain large parts that operate concurrently.
-Though system parallelising can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
+Though parallelisation of the system can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
 % Most articles, presentations and books on concurrent programming start with words how hard it is.
@@
 \end{tabular}
-\caption{Litmus test of memory operations reordering allowed by the x86-TSO weak memory model}
+\caption{Litmus test on memory operations reordering allowed by the x86-TSO weak memory model}
 \label{simple_wmm_x86}
@@
 The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
-This memory model, called the \textit{sequential consistency (SC) model}, allows only those executions (interleavings) that produce the same result as if the operations had been executed by single process.
+This memory model, called the \textit{sequential consistency (SC)}, allows only those executions (interleavings) that produce the same result as if the operations had been executed by single process.
 This means that the order of operations executed by a process is strictly defined by the program it executes.
@@
 Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them.
-The state-of-the-art tool is \texttt{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK, and firstly released back in 2010.
-It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \texttt{diy7}, \texttt{diycross7} and \texttt{diyone7}, the litmus tests concrete executor \texttt{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \texttt{herd7} that implements reachability analysis for capturing states allowed by the WMM.
-The latter tool uses systematic exhaustive search in a state space, which may be inefficient when analysing real-world programs, therefore more sophisticated search techniques should be applied for this problem.
+The state-of-the-art tool is \texttt{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010.
+It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \texttt{diy7}, \texttt{diycross7} and \texttt{diyone7}, the litmus tests concrete executor \texttt{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \texttt{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
 
-%TODO: BMC
-One possible approach is to use an efficient implementation of an SMT-solver (a SAT-solver extended by satisfiability modulo theories)~\cite{Porthos17}. % <-- TODO
-This approach allows to capture symbolically the semantics of both the program and the weak memory model while encoding it into a single SMT-formula.
-Most modern SMT-solvers are efficient enough to be able to operate the state space of size~[CITATION-?].
+All aforesaid tools work only with single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
+One of these problems is the \textit{portability} of the program from one hardware architecture to another.
+A program written in a high-level language is then compiled for different hardware.
+Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
+As the result, the program compiled for the platform $X$, can reach states that are unreachable on the platform $Y$, or some states that are reachable on $Y$ can become unreachable on $X$.
+We declare these cases as \textit{portability bugs}.
 
-%In the work of Finnish and German researchers ... names ...
-In the work~\cite{Porthos17}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
+The first tool that performs the WMM-awared portability analysis is \texttt{PORTHOS} introduced in April~2017~\cite{Porthos17}.
+This tool reduces described problem to the bounded model checking (BMC) problem, which can be solved via SMT-solver.
+This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula.
+As most modern SMT-solvers are efficient enough to be able to operate the state space of size ....
+(\textbf{TODO}), the used method can be applicable in solving the real-world problems.
+
+%In the work~\cite{Porthos17}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
 Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a trivial problem of symbolic execution, encoding of the weak memory model is more tedious.
 The reason is that some relations of WMMs are defined as mutually-recursive and need to be linearised in order to be encoded into an equivalent logical formula.
-%"a bounded model checking method that reduces the portability analysis problem to satisfiability modulo theorie" [Porthos17]
+
 %"A portability bug is an execution that is consistent with the target but inconsistent with the source memory model.
@@
 
-The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory models analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
+The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory model-awared analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
 Chapter~...

### 2018-03-22 0_intro.tex
Cleanup wmm intro part. Set up named references. Fix too many commas error. Fix other citing errors.

@@
 %while preserving consistency
-WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may expect.
+They specify how threads interact through shared memory, when a write becomes visible to other threads and what value a read can return. 
+Therefore, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may expect.
 

### 2018-03-22 1_wmm.tex
Cleanup wmm intro part. Set up named references. Fix too many commas error. Fix other citing errors.

@@
 The \textit{weak memory model} is a set of predicative constraints on possible executions of a concurrent program.
-The study of formalisation the weak memory models for different architectures is being rapidly developed over last decade.
-Research of WMM aimes, firstly, to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
+The study of formalisation weak memory models for different architectures has been rapidly developed over last decade.
 
-%The generic framework of program analysis under weak memory models presented in~\ref{alglave2010shared} is discussed in Section~\ref{ch:wmm:event}.
+Firstly, the research of WMM aimes to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
+A generic framework of program analysis with respect to weak memory models was presented in 2010~\cite{alglave2010shared}. Similar approach is used \texttt{PORTHOS} as it is described
+% in Section~\cite{ch:wmm:event}. 
+further in current Chapter.
 
-Secondly, researchers work on extracting the formal hardware memory models from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
-Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, SPARC-TSO for Sparc architecture~\cite{?!?!?!}, much more relaxed memory model ???
-for Power and ARM architectures defined in ???~\cite{sarkar2011understanding}, Alpha (???).
-Moreover, in 2005 <who?Milner> started the work on developing the weak memory model for C++ language, which was introduced in C++11 (?) standard~\cite{???} (//todo: C11 MM ~\cite{batty2011mathematizing}).
-The memory model for Java that is based on the \textit{happens-before} principle was introduced in JVM???
-in ??<year>.
+Secondly, researchers work on extracting the WMMs of hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
+Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding}, for Alpha~\cite{?} and Sparc~\cite{?} memory models were defined in the specification.
+Moreover, most modern high-level programming languages rely on relaxed memory model.
+Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model is based on the relation \textit{strongly happens-before}.
+%batty2011mathematizing
 
 Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to weak memory models.
-In this <area> the notable works <are on> defining the Linux kernel memory model <that is being actively developing these days~\cite{kernel_wmm_1}>.
-Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
+Perhaps, the most notable work in this field is the project for defining the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
+%Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
 
-%wmm as a formal way to define guaranties that a hw, programming language, execution environment provide for programmers.
-%considering wmm as a set of allowed behaviours, the latter wmms are the supersets
-%wmm allows and disallows optimisations: partial sync of memory buffers, out-of-order execution (reordering), <more> => more behaviours that are unallowed in SC.
-%question possible to answer with wmm: which behaviours (in addition to SC) are allowed?
-which new states are allowed?
-Consequently, correctness, absence of data races, deadlocks or portability issues, etc.
-
-%One important question that rises naturally when talking about weak memory models is whether a program is portable from one processor architecture to another. This can be solved ...<ref to PORTHOS>
-
-% more examples of what wmms allow to do?
+%experiments with kernel?
 
@@
 The semantics of a concurrent program is represented by the set of allowed executions.
-An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$ (\textit{data-flow} of a program)~\ref{alglave2010shared}.
-As it was shown in~\ref{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
+An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$ (\textit{data-flow} of a program)~\cite{alglave2010shared}.
+As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
 
@@
 
-The Porthos tool uses the language \texttt{CAT}~\ref{alglave2016syntax} for defining weak memory models.
+The Porthos tool uses the language \texttt{CAT}~\cite{alglave2016syntax} for defining weak memory models.
 

### 2018-04-02 0_intro.tex
Add some words on control-flow encoding.

@@
 For instance, Figure~\ref{simple_wmm_x86} provides simple example of
-unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable in x86 machines (such little examples that illustrate specific behaviours of a WMM are called \textit{litmus tests}).
+unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
 %reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
-This behaviour is caused by write buffers used by all processors in x86 architecture.
-These buffers cache writes to shared variables, so that the writes to shared memory does not become visible by other processes immediately.
-In the example, the write `\texttt{MOV~[x],1}' performed in the process \texttt{P0} stores value~\texttt{1} into the shared variable~\texttt{[x]} in the write buffer of process~\texttt{P0}.
+This state is allowed because in x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not become visible by other processes immediately.
+In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
 Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread.
@@
 Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them.
-The state-of-the-art tool is \texttt{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010.
-It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \texttt{diy7}, \texttt{diycross7} and \texttt{diyone7}, the litmus tests concrete executor \texttt{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \texttt{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
+The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010.
+It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \tool{diy7}, \tool{diycross7} and \tool{diyone7}, the litmus tests concrete executor \tool{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
 
@@
 
-The first tool that performs the WMM-awared portability analysis is \texttt{PORTHOS} introduced in April~2017~\cite{Porthos17}.
+The first tool that performs the WMM-awared portability analysis is \tool{PORTHOS} introduced in April~2017~\cite{Porthos17}.
 This tool reduces described problem to the bounded model checking (BMC) problem, which can be solved via SMT-solver.
@@
 
-Current work aims to improve the proof-of-concept tool \texttt{PORTHOS} firstly introduced in April 2017 in the work~\cite{Porthos17} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
+Current work aims to rework the proof-of-concept tool \tool{PORTHOS} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
+%interprocedural analysis
+We called the new tool \textbf{the \tool{mousquetaires} framework} as now it not only performs the portability analysis, but can serve as the basis for SMT-solver driven weak memory model analysis.
 

### 2018-04-02 1_2_encod.tex
Cleanup wmm intro part. Set up named references. Fix too many commas error. Fix other citing errors.
Add some words on control-flow encoding.

@@
 Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
-$cons_{\mathcal{M_T}}(P) \subseteq cons{\mathcal{M_S}}(P)$
+$cons_{\mathcal{M_T}}(P) \subseteq cons_{\mathcal{M_S}}(P)$
 \end{definition}
@@
 The full SMT formula $\phi$ should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
-The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below.
-However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.
-%$The \phi_{\mathcal{M_T}}$ is an encoding of the derived relations and all assertions of
-%, and φ ¬M S
+If the formula is satisfiable, there exist a portability bug.
+%The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. 
+%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.
 
-\subsection{Encoding of the control-flow constraints}
+\subsection{Encoding the control-flow constraints}
 \label{ch:port:enc:cf}
 
-\subsection{Encoding of the data-flow constraints}
-\label{ch:port:enc:cf}
+The control-flow of a program is represented in the \textit{control-flow graph}.
 
-\subsection{Encoding of the memory model constraints}
+The control-flow graph is a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} described in Section~\ref{ch:impl:xtoy:unrolling}.%TODO: footnote why multiple sinks
+In control-flow graph, there are two types of edges: \textit{primary edges} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative edges} that denote if-false-transitions (pictured with dotted lines).
+Each node on graph can have either one successor (primary) or two successors (primary and alternative); only computation events can serve as a branching point).
+However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.
+
+While working on the \tool{mousquetaires}, we applied some optimisations of the encoding scheme for the control-flow in order to decrease the size of result formula. 
+The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
+This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
+For instance, control-flow of the sequential instruction \m{i_1 := i_2; i_3} is encoded as 
+\m{\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}, 
+and control-flow of the branching instruction \m{i_1 := \{\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3\}} is encoded as 
+\m{\phi_{CF}(\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}.
+
+In contrast, the \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation,
+% (similar to an assembly language)
+that is then encoded into SMT formula.
+The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see further in current Chapter).
+%linearly as it shown in
+
+Let \m{\fx : \mathbb{E} \rightarrow \{0,1\}} be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system). 
+Let \m{\fv : \mathbb{C} \rightarrow \mathbb{N}} be the function that returns the value of the computation event that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \texttt{rf}-relation; see Chapter~?
+for the relations encoding%TODO
+).
+We distinguish the function \m{\fv_p : \mathbb{C_p} \rightarrow \{0,1\}} that evaluates the predicative computation event.
+In the result formula, all symbols $\fx(e)$ and $\fv(e)$ are encoded as boolean variables.
+
+Consider the following possible mutual arrangement of nodes in a control-flow grpah:
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{0.3\textwidth}
+\makebox[\textwidth]{
+  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+  \node[c] (1) [] {$e_1$};
+  \node[c] (2) [below of=1] {$e_2$};
+  \path[->]
+  (1) edge [] node {} (2)
+  ;
+  \end{tikzpicture}
+}
+\caption{Sequential events}
+\label{encode:cf:seq}
+\end{subfigure}
+~
+\begin{subfigure}[b]{0.35\textwidth}
+\makebox[\textwidth]{
+  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+  \node[c] (1) [] {$e_1$};
+  \node[c] (2) [below left of=1] {$e_2$};
+  \node[c] (3) [below right of=1] {$e_3$};
+  \path[->]
+  (1) edge [] node {} (2)
+  (1) edge [dotted] node {} (3)
+  ;
+  \end{tikzpicture}
+  }
+\caption{Conditional branching}
+\label{encode:cf:br}
+\end{subfigure}
+~
+\begin{subfigure}[b]{0.25\textwidth}
+\makebox[\textwidth]{
+  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+  \node[c] (i) [] {$e_i$};
+  \node[c,draw=none] (ii) [right=0.2cm of i] {};
+  \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
+  \node[c,draw=none] (iv) [right=0.2cm of iii] {};
+  \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
+  \node[c] (k) [below of=iii] {$e_k$};
+  \path[->]
+  (i) edge [] node {} (k)
+  (ii) edge [] node {} (k)
+  (iii) edge [dotted] node {} (k)
+  (iv) edge [dotted] node {} (k)
+  (v) edge [] node {} (k)
+  ;
+  \end{tikzpicture}
+}
+\caption{Branch merging}
+\label{encode:cf:merge}
+\end{subfigure}
+\caption{Linear and non-linear cases of control-flow graph}
+\label{encode:cf}
+\end{figure}
+
+For listed cases, below we propose the encoding scheme that uniquely encodes each node of graph and allows to encode partially executed program.
+The Equation~\ref{enc:cf:seq} for encoding the sequential control-flow represented on Figure~\ref{encode:cf:seq} reflects the fact that the event $e_2$ can be executed iff the event $e_1$ was executed.
+The Equation~\ref{enc:cf:br} for encoding the branching control-flow depicted on Figure~\ref{encode:cf:br} allows only following executions: $\{\emptyset, (e_1), (e_1 \rightarrow e_2), (e_1 \rightarrow e3) \}$.
+In encoding~\ref{enc:cf:merge} of the merge-point represented on Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
+
+\begin{align}
+%\phi_{CF_{(e_1;e_2)}}     =
+%\phi_{CF_{(e_1?e_2:e_3)}} =
+\phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
+\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_2) \rightarrow \fv(e_1)] \land \nonumber \\
+				  & [\fx(e_3) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
+				  & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
+\phi_{CF_{merge}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
+\end{align}
+
+\subsection{Encoding the data-flow constraints}
+\label{ch:port:enc:df}
+
+%As it was discussed in ...
+
+% SINGLE-ASSIGNMENT
+
+\subsection{Encoding the memory model constraints}
 \label{ch:port:enc:wmm}

### 2018-04-02 1_wmm.tex
Add some words on control-flow encoding.

@@
 An \textit{execution} (trace, run) of a given program is an ordered set of events.
+The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
 %associated with the instructions of the program
 %the path in the event-graph. 
-An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
+An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
 An execution to be checked on validity is called the \textit{candidate execution}.
@@
 The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}).
-A memory event is specified by its direction with respect to the shared variable, its location~$\texttt{loc}(e_m)$, its processor label~$\texttt{proc}(e_m)$, and a unique event label~$\texttt{id}(e_m)$~\cite{alglave2010shared}.
+A memory event is specified by its direction with respect to the shared variable, its location~$\mathtt{loc}(e_m)$, its processor label~$\mathtt{proc}(e_m)$, and a unique event label~$\mathtt{id}(e_m)$~\cite{alglave2010shared}.
 %\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
 The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
-We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\texttt{load}(r_1 \leftarrow l_2); \ e_m''~=~\texttt{store}(l_1 \leftarrow r_1)$.
+We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\mathtt{load}(r_1 \leftarrow l_2); \ e_m''~=~\mathtt{store}(l_1 \leftarrow r_1)$.
 
 A \textit{computation event} $e_c \in \mathbb{C} \subseteq \mathbb{E}$, represents a low-level assembly computation operation performed solely on local-memory arguments.
-An example of computation event may be the event $e_c = r_3 \leftarrow add(r_1, r_2)$ that writes the sum of values stored in registers $r_1$ and $r_2$ to the register $r_3$.
-The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded to the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
+An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (emulated as a register as well) to the register $r_1$.
+For modelling branching statements, we distinguish the set $\mathbb{C_{p}} \subseteq \mathbb{C}$ of \textit{predicative} computation events, that are evaluated as a boolean value.%TODO: TAUTOLOGY HERE.
+
+%The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
 
 The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}).
-Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the model that restrict the set of allowed behaviours.
-Technically, a fence may either serve as a synchronisation barrier, or flush local memory caches, etc.
+Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours.
+Technically, a fence may be represented either as a synchronisation barrier, or a flush local memory caches, etc.
 
@@
 The basic relation in the event-based program model is the \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of memory events \textit{within single process}, which never relates events from different processes.
-Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\texttt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
+Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
 This relation encodes the control-flow of the program into the event-graph.
@@
 %the event $d$:\texttt{(Rx=0)} reads value \texttt{0} at the shared location \texttt{x} from the initial write event $f$\texttt{(Wx=0)} (the red edge of \texttt{rf}-relation).
+
+// TODO: a bit more relations here
 

### 2018-04-02 2_impl.tex
Forgotten changes for release 14.03.2018.
Add some words on control-flow encoding.

@@
 \chapter{Implementation}
-\label{section:impl}
+\label{ch:impl}
 
@@
 \section{Program Requirements}
-\label{section:impl:requirements}
+\label{ch:impl:requirements}
 
@@
 \section{Program Components}
-\label{section:impl:comp}
+\label{ch:impl:comp}
 Big view
@@
 \section{C11 to YTree parser}
-\label{section:impl:ytree}
+\label{ch:impl:ytree}
 
@@
 \section{YTree to XGraph event converter}
-\label{section:impl:comp:xgraph}
+\label{ch:impl:xtoy}
 
@@
 \subsection{Loop unrolling}
+\label{ch:impl:xtoy:unrolling}
+% If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph.
+Although, in order to be able to encode such a cyclic structure as a boolean formula, this graph needs to be
+% In order to be
+
 The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}).
@@
 \label{fig:loop-unwind}
-\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop}, unwinded up to the bound $k = 6$} 
+\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$} 
 \end{figure}
@@
 \section{XGraph to ZFormula (SMT) encoder}
-\label{section:impl:comp:zformula}
+\label{ch:impl:comp:zformula}
 

### 2018-04-07 1_intro.tex
Cleanup everything.

--- 2018-04-02 0_intro.tex
@@
 Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread.
-More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
+%More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
 
@@
 \end{tabular}
-\caption{Litmus test on memory operations reordering allowed by the x86-TSO weak memory model}
+\caption{Store buffering (SB): a test on write-read reordering allowed under the x86-TSO weak memory model}
 \label{simple_wmm_x86}
@@
 Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them.
-The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010.
-It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \tool{diy7}, \tool{diycross7} and \tool{diyone7}, the litmus tests concrete executor \tool{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
+The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France and University of Cambridge, UK.
+The \tool{diy}\footnote{\url{diy.inria.fr}} is a sofware suite for designing and testing weak memory models.
+It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
+The \tool{diy} consists of several modules: the litmus tests generators \tool{diy7}, \tool{diycross7} and \tool{diyone7}, the litmus tests concrete executor \tool{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
 
-All aforesaid tools work only with single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
+All the \tool{diy} tools work only with single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
 One of these problems is the \textit{portability} of the program from one hardware architecture to another.
@@
 The first tool that performs the WMM-awared portability analysis is \tool{PORTHOS} introduced in April~2017~\cite{Porthos17}.
-This tool reduces described problem to the bounded model checking (BMC) problem, which can be solved via SMT-solver.
-This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula.
-As most modern SMT-solvers are efficient enough to be able to operate the state space of size ....
-(\textbf{TODO}), the used method can be applicable in solving the real-world problems.
+This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
+This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula, augmented by the reachability assertion.
+As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
+% TODO: search space size: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf slide
 

### 2018-04-07 2_wmm.tex
Cleanup everything.

--- 2018-04-02 1_wmm.tex
@@
 The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}).
+A memory event can access at most one shared memory location, high-level instructions that address more than one shared variable must be transformed into a sequence of events.
 A memory event is specified by its direction with respect to the shared variable, its location~$\mathtt{loc}(e_m)$, its processor label~$\mathtt{proc}(e_m)$, and a unique event label~$\mathtt{id}(e_m)$~\cite{alglave2010shared}.
@@
 
-%The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
-
 The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}).
@@
 
-The basic relation in the event-based program model is the \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of memory events \textit{within single process}, which never relates events from different processes.
+%The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
+The \textit{control-flow} of a program is defined by \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of events \textit{from same process}, which never relates events from different processes.
 Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
@@
 
-The data-flow of a program is encoded by the \textit{communication relations}: the \texttt{rf}-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \texttt{co}-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \texttt{fr}-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event.
+The \textit{data-flow} of a program is defined by \textit{communication relations}: the \texttt{rf}-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \texttt{co}-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \texttt{fr}-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \texttt{rf}-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
 %TODO: perephrase last sentence! alglave thesis, p. 36
 
-Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model (the picture is generated by the \texttt{herd7} tool, version 7.47).
+Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \texttt{po}$\cup$\texttt{rf} (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
 %The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
 %the event $d$:\texttt{(Rx=0)} reads value \texttt{0} at the shared location \texttt{x} from the initial write event $f$\texttt{(Wx=0)} (the red edge of \texttt{rf}-relation). 
-
-// TODO: a bit more relations here
+%However, the Power memory model allows such cycles, therefore %TODO: include load buffering example, see presentation pdf
 
@@
 
-The Porthos tool uses the language \texttt{CAT}~\cite{alglave2016syntax} for defining weak memory models.
+One common way to define a weak memory model in a systematic way in to formalise it in \textit{the CAT language} proposed in~\cite{alglave2016syntax}.
+The CAT language allows to axiomatically define new relations, new fences and restrictions over relations.
 
-the event representation
+%// exmaple with reordering
+%// ex. with 
+%// Rev-29 Example 7-6. Stores Are Transitively Visible. %see http://www.cl.cam.ac.uk/~pes20/weakmemory/x86tso-paper.pdf
 
-\section{Some known WMM}
+%There is a barrier instruction \texttt{mfence} that may be used for flushing the buffers into the main memory.
 
-\subsection{x86-TSO}
-\label{ch:wmm:x86}
-// exmaple with reordering
-// ex. with 
-// Rev-29 Example 7-6. Stores Are Transitively Visible. %see http://www.cl.cam.ac.uk/~pes20/weakmemory/x86tso-paper.pdf
+briefly some hw memory models: X86-TSO, Alpha, POWER, ;
 
-There is a barrier instruction \texttt{mfence} that may be used for flushing the buffers into the main memory.
+language memory models: Java, C++;
 
-briefly known hw memory models: X86-TSO, Alpha, POWER, -- ref to Jade;
-language memory models: Java, C++;
 library-level kernel memory model, ref to github with tests
 
-Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}
+%Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}
+
+\section{Some examples of WMM}
+
+//to be done

### 2018-04-07 3_encod.tex
Cleanup everything.

--- 2018-04-02 1_2_encod.tex
@@
-\chapter{Portability of Concurrent Software}
+\chapter{The Portability of Concurrent Software}
 %\chapter{The Concurrent Software Portability Analysis as a Bounded Model Checking Problem}
@@
 
-As it was discussed in Chapter~\ref{ch:intro}, the program may behave differently when compiled for different parallel hardware architectures.
-This can cause the portability bugs, the behaviour allowed under one architecture and forbidden under another.
-The concurrent software portability analysis may be stated as a Bounded Model Checking~(BMC) problem, which in turn can be reduced to the problem of satisfiability modulo theories (SMT)~\cite{Porthos17}.
+As it has been discussed in Chapter~\ref{ch:intro}, the program may behave differently when compiled for different parallel hardware architectures.
+This may cause the portability bugs, the behaviour that is allowed under one architecture and forbidden under another.
+%already said:
+%Although the research of weak memory models has achieved considerable success, most works remain to be rather theoretical that practical and serve mostly as tools for better understanding the concurrent nature of programs.
+% The first tool that tackles the problem 
+%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17}  
+In this Chapter, we describe the general task of analysing the concurrent software portability
+%may be stated 
+as a \textit{bounded reachability} problem, which in turn can be reduced to the satisfiability modulo theories (SMT) problem~\cite{Porthos17}.
 
@@
 
-%[-- TODO: REPHRASE
-One of the first major optimising technique was symbolic model checking with binary decision diagrams (BDDs).
-In this approach, a set of states is represented by a BDD instead of by listing each state individually~\cite{clarke2012model}.
-%--]
+One of the first technique that optimises the search space considerably major was the symbolic model checking with binary decision diagrams (BDDs).
+Instead of by processing each state individually, in this approach the set of states is represented by the BDD, efficient data structure for performing operations on large boolean formulas~\cite{clarke2012model}.
 The BDD representation can be linear of size of variables it encodes if the ordering of variables is optimal, otherwise the size of BDD is exponential.
@@
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
+%As well as the BMC problem, the approach used by the \tool{PORTHOS}
 
@@
 
-\section{The portability as a BMC problem}
+\section{The portability as a bounded reachability problem}
 \label{ch:port:enc}
 
-A program $P$ is called portable from the source weak memory model $\mathcal{M_S}$ to the target memory model $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:
+In general, the BMC problem aims to examine the reachability of the "undesirable" states of a finite-state system.
+Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $Init(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $Trans(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $Bad(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
+Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps is formulated as following:
+%\begin{equation}
+$\mathtt{SAT}( Init(\vec{x_0}) \land Trans(\vec{x}_0, \vec{x}_1) \land \cdots \land Trans(\vec{x}_{k-1}, \vec{x}_k) \land Bad(\vec{x}_k) )$.
+%\end{equation}
+
+The portability probram may also be formulated as a reachability problem, where the undesirable state is the state reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
+However, unlikely the BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:x2y:unrolling}) and the $Trans$ predicate may be stated only for the final state of a program.
+
+As we have said before, a program $P$ is called portable from the source architecture (defined as a memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if \textit{all} executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:
 
@@
 
-It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the BMC problem.
-The full SMT formula $\phi$ should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
+%It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the bounded reachability problem.
+The result SMT formula $\phi$ of the portability problem should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
 If the formula is satisfiable, there exist a portability bug.
@@
 
-The control-flow of a program is represented in the \textit{control-flow graph}.
-
-The control-flow graph is a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} described in Section~\ref{ch:impl:xtoy:unrolling}.%TODO: footnote why multiple sinks
+The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:x2y:unrolling}).%TODO: footnote why multiple sinks
 In control-flow graph, there are two types of edges: \textit{primary edges} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative edges} that denote if-false-transitions (pictured with dotted lines).
@@
 \m{\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}, 
-and control-flow of the branching instruction \m{i_1 := \{\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3\}} is encoded as 
-\m{\phi_{CF}(\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}.
+and control-flow of the branching instruction \m{i_1 := \{c \, ? \, i_2 \, : \, i_3\}} is encoded as 
+\m{\phi_{CF}( \, c \, ?
+\, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)} (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `\texttt{if} $x$ \texttt{then} $y$ \texttt{else} z').
 
@@
 We distinguish the function \m{\fv_p : \mathbb{C_p} \rightarrow \{0,1\}} that evaluates the predicative computation event.
-In the result formula, all symbols $\fx(e)$ and $\fv(e)$ are encoded as boolean variables.
+In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
 
@@
 \phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_2) \rightarrow \fv(e_1)] \land \nonumber \\
-				  & [\fx(e_3) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
-				  & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
+\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_2) \rightarrow \fv(e_1)] \nonumber \\
+				  & \land  [\fx(e_3) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \nonumber \\
+				  & \land \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
 \phi_{CF_{merge}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
@@
 
-%As it was discussed in ...
+While encoding the data-flow constraints, we use the \textit{Static Single-Assignment (SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
+The single-assignment requires that a variable is not assigned once it has been accessed (either read or written).
+That is, on each state of the program a variable (either local or shared) is mapped to its indexed reference, which represents a new boolean variable in the formula.
+On merge-points, the $\phi$-function is emulated as the disjunction of values of the variable, that is computed as the reaching definition.
+For that, we keep track of the map that for each event stores mapping from all declared variables to their SSA-indexes.
+All indexes are globally unique within single program analysis.
 
-% SINGLE-ASSIGNMENT
+As it is implemented in~\cite{Porthos17}, the SSA-indexes of variables are computed in accordance with the following rules: (1) any access to a shared variable (both read and write) increments its SSA-index; (2) only writes to a local variable increment its SSA-index (reads preserve indexes); (3) no access to a constant variable or computed expression changes their SSA-index.
 

### 2018-04-07 5_eval.tex
Cleanup everything.

\chapter{Evaluation}
\label{ch:eval}

\section{Comparison with PORTHOS}

\subsection{Unique Features}
\subsection{Performance}

\section{Comparison with HERD}

\subsection{Unique Features}
\subsection{Performance}

### 2018-04-07 2_wmm.tex
Complete the data-flow encoding.

@@
 %"Semantics and verification under weak memory models have been the subject of study at least since 2007." [PORTHOS]
-The \textit{weak memory model} is a set of predicative constraints on possible executions of a concurrent program.
+A \textit{weak memory model} is a set of predicative constraints over so-called \textit{derived relations}, which defines the set of possible executions of a concurrent program.
+Usually, these constraints include requirements of the reflexivity or acyclicity of derived relations.
 The study of formalisation weak memory models for different architectures has been rapidly developed over last decade.
@@
 Firstly, the research of WMM aimes to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
-A generic framework of program analysis with respect to weak memory models was presented in 2010~\cite{alglave2010shared}. Similar approach is used \texttt{PORTHOS} as it is described
+A generic framework of program analysis with respect to weak memory models was presented in 2010~\cite{alglave2010shared}. Similar approach is used \Porthos as it is described
 % in Section~\cite{ch:wmm:event}. 
@@
 
-The event-based program model represents the directed graph (\textit{event-graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events.
+The event-based program model represents the directed graph (the \textit{event-flow graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events.
 An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program.
@@
 %associated with the instructions of the program
-%the path in the event-graph. 
+%the path in the event-flow graph. 
 An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
@@
 
+In this section, we describe basic derived relations used in memory model-awared program analysis.
+
 %The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
 The \textit{control-flow} of a program is defined by \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of events \textit{from same process}, which never relates events from different processes.
-Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
-This relation encodes the control-flow of the program into the event-graph.
+Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-flow graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
+This relation encodes the control-flow of the program into the event-flow graph.
 %Some new relations may be acquired : dp, po-loc
 
-The \textit{data-flow} of a program is defined by \textit{communication relations}: the \texttt{rf}-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \texttt{co}-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \texttt{fr}-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \texttt{rf}-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
+The \textit{data-flow} of a program is defined by \textit{communication relations}: the \rf-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \co-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \fr-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
 %TODO: perephrase last sentence! alglave thesis, p. 36
 
-Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \texttt{po}$\cup$\texttt{rf} (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
+Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \po$\cup$\rf (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
 %The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
@@
 
-//to be done
+//axioms of TSO wmm
+
+// example of sets: rf, co, ... for the code snipped used before

### 2018-04-07 3_encod.tex
Complete the data-flow encoding.

@@
 
-\subsection{Encoding the control-flow constraints}
+\subsection{The control-flow encoding} %Encoding the control-flow constraints}
 \label{ch:port:enc:cf}
@@
 
-While working on the \tool{mousquetaires}, we applied some optimisations of the encoding scheme for the control-flow in order to decrease the size of result formula. 
-The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
+While working on the \tool{mousquetaires}, we applied some modifications of the encoding scheme for the control-flow.
+The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of C language.
+For that, we compile the recursive abstract syntax tree~(AST) of the parsed C-code to the plain (non-recursive) event-flow graph.
+We show%TODO! to show, actually
+that the new encoding is smaller than the old one used in \Porthos since it does not produces new variables for each high-level statement of the input language.
+%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, control-flow of the sequential instruction \m{i_1 := i_2; i_3} is encoded as 
+For instance, in old scheme the control-flow of the sequential instruction \m{i_1 := i_2; i_3} was encoded as 
 \m{\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}, 
-and control-flow of the branching instruction \m{i_1 := \{c \, ? \, i_2 \, : \, i_3\}} is encoded as 
+and control-flow of the branching instruction \m{i_1 := \{c \, ? \, i_2 \, : \, i_3\}} was encoded as 
 \m{\phi_{CF}( \, c \, ?
 \, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)} (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `\texttt{if} $x$ \texttt{then} $y$ \texttt{else} z').
-
-In contrast, the \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation,
-% (similar to an assembly language)
-that is then encoded into SMT formula.
+In contrast, the new scheme implemented in \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
 The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see further in current Chapter).
@@
 In encoding~\ref{enc:cf:merge} of the merge-point represented on Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
+%TODO: maybe add the overall formula \phi_{CF} ?
 
@@
 \phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_2) \rightarrow \fv(e_1)] \nonumber \\
-				  & \land  [\fx(e_3) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \nonumber \\
-				  & \land \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
+\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
+				  &  [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
+				  & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
 \phi_{CF_{merge}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
@@
 
-\subsection{Encoding the data-flow constraints}
+\subsection{The data-flow encoding} %constraints}
 \label{ch:port:enc:df}
 
-While encoding the data-flow constraints, we use the \textit{Static Single-Assignment (SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
-The single-assignment requires that a variable is not assigned once it has been accessed (either read or written).
+While encoding the data-flow constraints, we use the \textit{Static Single-Assignment~(SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
+The SSA form requires each variable to be assigned only once within entire program.
 That is, on each state of the program a variable (either local or shared) is mapped to its indexed reference, which represents a new boolean variable in the formula.
-On merge-points, the $\phi$-function is emulated as the disjunction of values of the variable, that is computed as the reaching definition.
-For that, we keep track of the map that for each event stores mapping from all declared variables to their SSA-indexes.
-All indexes are globally unique within single program analysis.
+On merge-points, the $\phi$-function is emulated as the disjunction of values of the variable that is computed as the \textit{reaching definition}.
+For that, we carry the map that for each event stores mapping from all declared variables to their SSA-indices.
+All indices are globally unique within single program analysis.
 
-As it is implemented in~\cite{Porthos17}, the SSA-indexes of variables are computed in accordance with the following rules: (1) any access to a shared variable (both read and write) increments its SSA-index; (2) only writes to a local variable increment its SSA-index (reads preserve indexes); (3) no access to a constant variable or computed expression changes their SSA-index.
+As it is implemented in~\cite{Porthos17}, the SSA-indices of variables are computed in accordance with the following rules: (1) any access to a shared variable (both read and write) increments its SSA-index; (2) only writes to a local variable increment its SSA-index (reads preserve indices); (3) no access to a constant variable or computed expression changes their SSA-index.
+These rules determine the following encoding of load, store and computation events within single thread:
+%TODO: maybe add the overall formula \phi_{DF} ?
 
-\subsection{Encoding the memory model constraints}
+\begin{align}
+\phi_{DF_{e = \mathtt{load}(r \leftarrow l)}} = \ & \fx(e) \rightarrow (r_{i+1} = l_{i+1}) \\
+\phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} = \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
+\phi_{DF_{e = \mathtt{compute}(...)}} = \ & \fx(e) \rightarrow \fv(e) % \\
+\end{align}
+
+Comparing to the old \Porthos version, the algorithm of converting the program into the SSA form was also adapted for processing the plain flow-graph data structure.
+That is, for each node we keep track the set of its predecessors, and then, while traversing the graph in topological order, we propagate the SSA-indices for each variable defined so far, and update these indices according the rules listed above.
+
+The time of described algorithm is linear of the size of event-flow graph since it performs only single traverse of the graph.
+Notwithstanding the overhead of storing (or, equivalently but undesirably, computing with%TODO: are you sure that 'with'?
+linear time) the predecessors of each node in order to be able to convert the program into an SSA form, the time of such a tranformation reduced%TODO: How much ????
+comparing to the algorithm implemented in \Porthos, which recomputed SSA-indices recursively for each instruction.%TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
+
+Moreover, the data-flow encoding scheme used in \Porthos faced the problem when the branching statements had different number of assignments of certain variable in their branches.
+//TODO: why?
+why don't we have this problem now?
+
+As it has been stated before, the \rf-relation links the parts%TODO: another word?
+of data-flow stored in equivalence assertions over the SSA-variables. 
+%The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
+The encoding of this linkage left untouch as it is implemented in \Porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
+
+\begin{align}
+\phi_{DF_{mem}}(e_1, e_2) = \ & \mathtt{rf}(e_1, e_2) \rightarrow (l_i = l_j) 
+\end{align}
+
+where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\mathtt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.
+
+\subsection{The memory model encoding} %Encoding the memory model constraints}
 \label{ch:port:enc:wmm}
+
+The encoding of a weak memory model itself does not depend on the analysing program. The problem here...
+
+recursive definitions and Kleene iteration
+
+As it is described in~\cite{Porthos17}, the is based ...
+%~\cite{Porthos17}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"

### 2018-04-07 4_input.tex
Complete the data-flow encoding.

\chapter{The input language}
\label{ch:impl}

the old input language syntax

new one iwth changes highlighted

### 2018-04-12 3_encod.tex
Rename; changes from bus.

@@
 
-Comparing to the old \Porthos version, the algorithm of converting the program into the SSA form was also adapted for processing the plain flow-graph data structure.
+Unlikely the \Porthos tool, the \mousquetaires uses the algorithm of converting the program into the SSA form was also adapted for processing the plain flow-graph data structure.
 That is, for each node we keep track the set of its predecessors, and then, while traversing the graph in topological order, we propagate the SSA-indices for each variable defined so far, and update these indices according the rules listed above.

### 2018-04-12 4_input.tex
add today's released changes.

@@
 
-the old input language syntax
+The \Porthos tool used a small subset of C as an input language as described on Figure~\ref{syntax:porthos}\cite{Porthos17}:
 
-new one iwth changes highlighted
+\begin{figure}[H]
+\begin{lstlisting}[mathescape=true,
+        caption={Syntax of an input language of \Porthos},
+        captionpos=b,
+        label={syntax:porthos},
+        basicstyle=\ttfamily\footnotesize,
+        keywordstyle=\bfseries,
+        literate={<--}{$\leftarrow{}$}{1},%{::=}{???},%:\!:\!=},
+        morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}]
+<prog>  : program <thrd>*
+<thrd>  : thread <tid> <inst>
+<inst>  : <atom> 
+        | <inst> ; <inst>
+        | while <pred> <inst>
+        | if <pred > then <inst> else <inst>
+<atom>  : <reg> <-- <exp> 
+        | <reg> <-- <loc>
+        | <loc> := <reg> 
+        | mfence
+        | sync
+        | lwsync
+        | isync
+\end{lstlisting}
+%\caption={Syntax of an input language of \Porthos}
+%\label={syntax:porthos}
+\end{figure}
+
+// todo: old syntax more precisely
+
+The input language of \Porthos supported recursive definitions of control-flow instructions, which needed to be extended to support arbitrary jumps expressible in C.
+Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment `\texttt{:=}' or local variable assignment `$\mathtt{\rightarrow}$') were defined directly in the grammar for \Porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
+% according to the set of known functions
+(see Chapter~\ref{???}).
+
+//todo: new syntax

### 2018-04-12 5_impl.tex
Complete the data-flow encoding.
Rename; changes from bus.
add today's released changes.

--- 2018-04-02 2_impl.tex
@@
-\chapter{Implementation}
+\chapter{\nohyphens{The~\mousquetaires:~implementation}}
 \label{ch:impl}
 
-This chapter describes the architecture of the tool \texttt{mousquitaires} ...
+This Chapter describes the architecture of the tool \mousquetaires.
 
@@
 
-\section{Program Requirements}
+% why not AST: https://stackoverflow.com/a/13968707/4977823
+
+\section{Requirements}
 \label{ch:impl:requirements}
 
-- stability (tests)
+The main reason for starting the work on re-implementing the \Porthos tool was the need to extend input language to be able to process real C programs in perspective.
 
-- scalability (new features of language, new models, new tasks for a program )
+The target requirements for the new tool was set up as following (in descending order of priority):
 
-- transparency
-
-- efficiency
+\begin{enumerate}
+        \item stability and transparency (KISS principle)
+                \begin{itemize}
+                        \item choice in favour of simplicity and readability at every step;
+                        \item usage of software design patterns if necessary;
+                        \item usage of immutable data structures for intermediate representations;
+                        \item high code coverage of unit and functional tests;
+                \end{itemize}                
+        \item efficiency
+                \begin{itemize}
+                        %\item usage of time- and memory-optimised data structures;
+                        \item keeping the trade-off between execution time and memory consumption;
+                \end{itemize}
+        \item extensibility
+                \begin{itemize}
+                        \item clear modular architecture
+                \end{itemize}
+\end{enumerate}
 
@@
 \label{ch:impl:comp}
-Big view
 
-\section{C11 to YTree parser}
+Generally, \mousquetaires uses following
+
+\section{Parsing the input language as YTree}
 \label{ch:impl:ytree}
@@
 \section{YTree to XGraph event converter}
-\label{ch:impl:xtoy}
+\label{ch:impl:x2y}
 
@@
 \subsection{Loop unrolling}
-\label{ch:impl:xtoy:unrolling}
+\label{ch:impl:x2y:unrolling}
 % If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph.

### 2018-04-12 4_input.tex
add grammar to appendix

@@
 \begin{figure}[H]
+\begin{multicols}{2}
 \begin{lstlisting}[mathescape=true,
         caption={Syntax of an input language of \Porthos},
-        captionpos=b,
+        % captionpos=b,
         label={syntax:porthos},
-        basicstyle=\ttfamily\footnotesize,
-        keywordstyle=\bfseries,
-        literate={<--}{$\leftarrow{}$}{1},%{::=}{???},%:\!:\!=},
-        morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}]
+        % basicstyle=\ttfamily\scriptsize,
+        % keywordstyle=\bfseries\color{blue},
+        %literate={<--}{$\leftarrow{}${}1},%{::=}{???},%:\!:\!=},
+        %morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}
+        %alsoletter={:,|,;},%
+        %morekeywords={:,|,;},%
+        %keywordstyle={\color{red}},% 
+        breaklines=true,
+        % xleftmargin=5mm,framexleftmargin=5mm
+        ]
 <prog>  : program <thrd>*
@@
 %\label={syntax:porthos}
+\end{multicols}
 \end{figure}

### 2018-04-18 4_input.tex
Add the draft on input language.

--- 2018-04-07 4_input.tex
@@
 
-the old input language syntax
+The \Porthos tool used a small subset of C as an input language as described on Figure~\ref{syntax:in_grammar_pts}~\cite{Porthos17}:
 
-new one iwth changes highlighted
+\begin{figure}[H]
+%\begin{multicols}{2}
+\begin{lstlisting}[mathescape=true,
+        caption={Syntax of an input language of \Porthos},
+        label={syntax:in_grammar_pts},
+        literate={<--}{{$\leftarrow$}}1,
+        morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}
+        breaklines=true,
+        ]
+<prog> : program <thrd>*
+       ;
+<thrd> : thread <tid> <inst>
+       ;
+<inst> : <atom> 
+       | <inst> ; <inst>
+       | while <pred> <inst>
+       | if <pred > then <inst> else <inst>
+       ;
+<atom> : <reg> <-- <exp> 
+       | <reg> <-- <loc>
+       | <loc> := <reg> 
+       | mfence
+       | sync
+       | lwsync
+       | isync
+       ;
+\end{lstlisting}
+%\end{multicols}
+\end{figure}
+
+The input language of \Porthos supported recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
+In \mousquetaires, the set of supported instructions is extended in order to support expressions valid in C.%TODO: examples, which  we extended
+ Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment~`\texttt{:=}' or local variable assignment~`$\mathtt{\leftarrow}$') were defined directly in the grammar for \Porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
+% according to the set of known functions
+(see Chapter~\ref{???}).
+
+Both \Porthos and \mousquetaires \ use the parser generator ANTLR~\cite{parr2013definitive}, the powerfull language processing tool.
+The (hand-written) ANTLR grammar used in \Porthos is available at Appendix~\ref{apx:in_grammar_pts}, whereas the \mousquetaires uses the grammar of C language proposed in the C11 standard~\cite{jtc2011sc22} (the ANTLR grammar can be found in the official ANTLR repository on GitHub~\footnote{\url{https://github.com/antlr/grammars-v4}}).

### 2018-04-18 5_impl.tex
Add the draft on input language.

@@
 
-language: java
-
-% why not AST: https://stackoverflow.com/a/13968707/4977823
+The programming language choice for \mousquetaires was made in favour of java in order to reuse some parts of its predecessor \Porthos that was written in java.
+However this language does not show best results in performance benchmarks (comparing to, for example, C++)~\cite{TODO}, the performance cornerstone of \mousquetaires (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \mousquetaires via java API.
+However, considering the perspective of using \mousquetaires as a static analyser for real-world programs, we also must to take into account memory problems at both encoding and solving stages.
 
@@
 The main reason for starting the work on re-implementing the \Porthos tool was the need to extend input language to be able to process real C programs in perspective.
+The existing \Porthos architecture mixes high-level recursive \textit{instructions} (statements of C) with low-level non-recursive \textit{events}, to which the high-level instructions are compiled (see classes of package `\texttt{dartagnan.program}' in the \Porthos project~\footnote{Project web page: \url{http://github.com/hernanponcedeleon/PORTHOS}}).
+Both instructions and events are then encoded into the SMT-formula.
+However, this approach is poorly extensible as to adding support for a high-level control-flow instruction (say, the \texttt{do-while} loop in C) requires making changes in many places of code from parser to encoder.
+Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).
 
-The target requirements for the new tool was set up as following (in descending order of priority):
+Therefore, while planning the architecture of \mousquetaires, we decided to clearly separate high-level recursive code representation as Abstract Syntax Tree (AST) and low-level compiled event-based representation as a flow-graph.
+Such a modular architecture allows to use multiple input-language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \Porthos tool, two syntaxes of litmus tests used by \tool{herd}, and even assembly language for any supported architecture).
 
-\begin{enumerate}
+The target \textbf{requirements} for the new tool was set up as following (in descending order of priority):
+
+\begin{enumerate}[nolistsep]
         \item stability and transparency (KISS principle)
@@
                         \item choice in favour of simplicity and readability at every step;
-                        \item usage of software design patterns if necessary;
-                        \item usage of immutable data structures for intermediate representations;
-                        \item high code coverage of unit and functional tests;
+                        \item using software design patterns if necessary;
+                        \item using immutable data structures for intermediate representations;
+                        \item high code coverage by unit and functional tests;
                 \end{itemize}                
@@
                         %\item usage of time- and memory-optimised data structures;
-                        \item keeping the trade-off between execution time and memory consumption;
+                        \item keeping trade-off between execution time and memory consumption;
                 \end{itemize}
@@
 
-Generally, \mousquetaires uses following
+Generally, \mousquetaires uses following  ...
 
@@
 
-The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}).
+The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{(ENTRY)} (usually for convenience all leaves are connected to the sink node \texttt{(EXIT)}).
 The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).

### 2018-04-18 1_intro.tex
Cleanup intro and wmm chapters.

@@
 Traditionally, studies related to concurrent programming concern on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language.
-Unfortunately,
-when it comes to 
-%in cases of 
-the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
+Unfortunately, when it comes to the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
-For instance, Figure~\ref{simple_wmm_x86} provides simple example of
-unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
+For instance, Figure~\ref{simple_wmm_x86} provides simple example of unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
 %reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
@@
 In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
-Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread.
-%More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
+Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in another thread.
+These problems have lead to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{weak	memory models (WMM)}.
 
@@
 
+Research of weak memory models firstly aimes to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
+The first (and so far the only) such a framework was presented in 2010~\cite{alglave2010shared}.
+%
+Secondly, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
+Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding}, for Alpha~\cite{?} and Sparc~\cite{?} memory models were defined in the specification.
+Moreover, most modern high-level programming languages rely on relaxed memory model.
+Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model is based on the relation \textit{strongly happens-before}.
+%batty2011mathematizing
+Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to WMMs.
+Perhaps, the most notable work in this field is the project for defining the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
+%Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
+
+%experiments with kernel?
+
 The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
@@
 The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France and University of Cambridge, UK.
-The \tool{diy}\footnote{\url{diy.inria.fr}} is a sofware suite for designing and testing weak memory models.
+The \tool{diy}\footnote{Project web page: \url{diy.inria.fr}} is a sofware suite for designing and testing weak memory models.
 It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
-The \tool{diy} consists of several modules: the litmus tests generators \tool{diy7}, \tool{diycross7} and \tool{diyone7}, the litmus tests concrete executor \tool{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.
+The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, \tool{diycross} and \tool{diyone}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
 
@@
 Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, the program compiled for the platform $X$, can reach states that are unreachable on the platform $Y$, or some states that are reachable on $Y$ can become unreachable on $X$.
-We declare these cases as \textit{portability bugs}.
+As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17}.
 

### 2018-04-18 2_wmm.tex
Cleanup intro and wmm chapters.

@@
-\chapter{Weak Memory Models}
+\chapter{Memory Model-Awared Program Analysis}
 \label{ch:wmm}
 
-%"Semantics and verification under weak memory models have been the subject of study at least since 2007." [PORTHOS]
-A \textit{weak memory model} is a set of predicative constraints over so-called \textit{derived relations}, which defines the set of possible executions of a concurrent program.
-Usually, these constraints include requirements of the reflexivity or acyclicity of derived relations.
-The study of formalisation weak memory models for different architectures has been rapidly developed over last decade.
+In general, analysis of concurrent programs with respect to axiomatic memory models require firstly to capture the control-flow of a concurrent program, after which the set of all possible data-flows (called \textit{candidate executions}) is computed as the set of possible communications between threads of the program.
+After that, the \textit{weak memory model} specification constraints this set of candidate executions and decides which execution is valid~\cite{alglave2014herding}.
+More specifically, we use the event-based program representation (see Section~\ref{ch:wmm:model:events}) as a natural way to capture the semantics of a concurrent program.
+Therefore, a candidate execution is defined as a choice of events to be executed.
+The constraints over events imposed by both program and memory model are called \textit{relations} (see Section~\ref{ch:wmm:model:relations})
 
-Firstly, the research of WMM aimes to formalise the weak memory models and provide systematic, sound and complete formal approach of defining WMMs in order to be able to verify systems with respect to them.
-A generic framework of program analysis with respect to weak memory models was presented in 2010~\cite{alglave2010shared}. Similar approach is used \Porthos as it is described
-% in Section~\cite{ch:wmm:event}. 
-further in current Chapter.
-
-Secondly, researchers work on extracting the WMMs of hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
-Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding}, for Alpha~\cite{?} and Sparc~\cite{?} memory models were defined in the specification.
-Moreover, most modern high-level programming languages rely on relaxed memory model.
-Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model is based on the relation \textit{strongly happens-before}.
-%batty2011mathematizing
-
-Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to weak memory models.
-Perhaps, the most notable work in this field is the project for defining the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
-%Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
-
-%experiments with kernel?
+%More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
 
@@
 
-The classical approach to model the concurrent programs is to use the \textit{global time}, a single order of interleavings of all actions happened in different threads.
+The classical approach to model concurrent programs is to use the \textit{global time}, a single order of interleavings of all actions happened in different threads.
 However these models are easy to understand, it may be hard to consider \textit{all} possible states, number of which is exponentially large.
 Another way to do this is to to use non-deterministic computation-centric models defined in~\cite{fri97}, one of which represents the program as the graph of \textit{memory events}.
-The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independant from the order of local computation events.
-These models may be further restricted by constraints of a weak memory model, adding \textit{relations} to the memory events.
+The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
+%These models may be further restricted by constraints of a weak memory model, adding \textit{relations} to the memory events.
 %Perhaps, the most convenient way to model the non-deterministic properties of concurrent programs is to use the program representation based on \textit{memory events}.
@@
 
-The event-based program model represents the directed graph (the \textit{event-flow graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events.
-An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program.
-An \textit{execution} (trace, run) of a given program is an ordered set of events.
-The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
-%associated with the instructions of the program
-%the path in the event-flow graph. 
-An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-An execution to be checked on validity is called the \textit{candidate execution}.
-
+The event-based program model represents the directed graph (the \textit{event-flow graph}), where vertices represent \textit{events} (simple low-level instructions), and edges represent \textit{relations} over the events.
 Below we describe some basic types of events and relations.
@@
 
-The \textit{data-flow} of a program is defined by \textit{communication relations}: the \rf-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write to a read reading its value, the \co-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \fr-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
+The \textit{data-flow} of a program is defined by \textit{communication relations}: the \rf-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write event to the read event that reads its value, the \co-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \fr-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
 %TODO: perephrase last sentence! alglave thesis, p. 36
 
-Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \po$\cup$\rf (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
+\subsection{Executions}
+\label{ch:wmm:model:executions}
+
+The semantics of a concurrent program is represented by the set of allowed executions.
+A \textit{candidate execution} (trace, run) of a program is an ordered set of events defined by \po and \rf relations and set of final writes to a given memory location~\cite{alglave2014herding}.
+%The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
+
+%associated with the instructions of the program
+%the path in the event-flow graph. 
+An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
+An execution to be checked on validity is called the \textit{candidate execution}.
+An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
+As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
+
+The Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \po$\cup$\rf (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
 %The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
@@
 % TODO: re-generate with new name, not 'SO'
-
-\subsection{Executions}
-\label{ch:wmm:model:executions}
-
-The semantics of a concurrent program is represented by the set of allowed executions.
-An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$ (\textit{data-flow} of a program)~\cite{alglave2010shared}.
-As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
 

### 2018-04-18 1_intro.tex
Change the illustration for the litmus-test example.

@@
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
-For instance, Figure~\ref{simple_wmm_x86} provides simple example of unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
+For instance, Figure~\ref{simple_wmm_x86} provides simple example of reachability of the state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
 %reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
@@
 \end{tabular}
-\caption{Store buffering (SB): a test on write-read reordering allowed under the x86-TSO weak memory model}
+\caption{Store buffering (SB): a litmus test on write-read reordering allowed under the x86-TSO weak memory model}
 \label{simple_wmm_x86}
@@
 It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
-The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, \tool{diycross} and \tool{diyone}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
+The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
 

### 2018-04-18 2_wmm.tex
Change the illustration for the litmus-test example.

@@
 As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
-
-The Figure~\ref{simple_wmm_x86_pic} illustrates the candidate execution for the Example~\ref{simple_wmm_x86}, that reaches the state \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)} within x86-TSO memory model as it forbids cycles over \po$\cup$\rf (union) relation (the picture is generated by the \texttt{herd7} tool, version 7.47).
+Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus-test Example~\ref{simple_wmm_x86} (the pictures are generated by the \texttt{herd7} tool, version 7.47).
+Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
+Since each thread performs single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).
+Thus, there are only four possible executions defined by the choice of \rf-relation.
+The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and others.
+However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over \fr$\cup$\po.
 %The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
@@
 
-\includegraphics[width=0.4\textwidth]{img/my/simple_wmm_x86.png}
+%\includegraphics[width=0.4\textwidth]{img/my/simple_wmm_x86.png}
+\begin{figure}
+\centering
+\begin{subfigure}[t]{.28\textwidth}
+  \centering
+  \includegraphics[width=1.2\linewidth]{img/my/sb-example/SB-dot.png}
+  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=1)}}
+  \label{simple_wmm_x86_pic:sub1}
+\end{subfigure}
+\hfill
+\begin{subfigure}[t]{.23\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-2.png}
+  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=0)}}
+  \label{simple_wmm_x86_pic:sub2}
+\end{subfigure}
+\hfill
+\begin{subfigure}[t]{.23\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-3.png}
+  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=1)}}
+  \label{simple_wmm_x86_pic:sub3}
+\end{subfigure}
+\hfill
+\begin{subfigure}[t]{.23\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-4.png}
+  \caption{Final state: \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}}
+  \label{simple_wmm_x86_pic:sub4}
+\end{subfigure}
+\hfill
+\caption{Candidate executions for the litmus test Example~\ref{simple_wmm_x86}}
 \label{simple_wmm_x86_pic}
-% TODO: re-generate with new name, not 'SO'
+\end{figure}
 

### 2018-04-18 3_encod.tex
Change the illustration for the litmus-test example.

@@
 For listed cases, below we propose the encoding scheme that uniquely encodes each node of graph and allows to encode partially executed program.
-The Equation~\ref{enc:cf:seq} for encoding the sequential control-flow represented on Figure~\ref{encode:cf:seq} reflects the fact that the event $e_2$ can be executed iff the event $e_1$ was executed.
-The Equation~\ref{enc:cf:br} for encoding the branching control-flow depicted on Figure~\ref{encode:cf:br} allows only following executions: $\{\emptyset, (e_1), (e_1 \rightarrow e_2), (e_1 \rightarrow e3) \}$.
-In encoding~\ref{enc:cf:merge} of the merge-point represented on Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
+The Equation~\ref{enc:cf:seq} for encoding the sequential control-flow represented in Figure~\ref{encode:cf:seq} reflects the fact that the event $e_2$ can be executed iff the event $e_1$ was executed.
+The Equation~\ref{enc:cf:br} for encoding the branching control-flow depicted in Figure~\ref{encode:cf:br} allows only following executions: $\{\emptyset, (e_1), (e_1 \rightarrow e_2), (e_1 \rightarrow e3) \}$.
+In encoding~\ref{enc:cf:merge} of the merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
 %TODO: maybe add the overall formula \phi_{CF} ?

### 2018-04-18 2_wmm.tex
Cleanup text up to 2.1.2 Relations.

@@
 
-In general, analysis of concurrent programs with respect to axiomatic memory models require firstly to capture the control-flow of a concurrent program, after which the set of all possible data-flows (called \textit{candidate executions}) is computed as the set of possible communications between threads of the program.
-After that, the \textit{weak memory model} specification constraints this set of candidate executions and decides which execution is valid~\cite{alglave2014herding}.
-More specifically, we use the event-based program representation (see Section~\ref{ch:wmm:model:events}) as a natural way to capture the semantics of a concurrent program.
-Therefore, a candidate execution is defined as a choice of events to be executed.
-The constraints over events imposed by both program and memory model are called \textit{relations} (see Section~\ref{ch:wmm:model:relations})
-
-%More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.
+In general, analysis of concurrent programs with respect to axiomatic memory models is performed in several stages.
+Firstly, the control-flow and data-flow of a program is encoded as the set of possible \textit{candidate executions}.
+%possible communications between threads of the program.
+Obtainted model of the program is called an anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
+Thereafter, the anarchic semantics is constrained by the \textit{weak memory model} specification which is a set of axiomatic constraints for filtering out executions inconsistent in particular architecture.
 
@@
 
-The classical approach to model concurrent programs is to use the \textit{global time}, a single order of interleavings of all actions happened in different threads.
-However these models are easy to understand, it may be hard to consider \textit{all} possible states, number of which is exponentially large.
-Another way to do this is to to use non-deterministic computation-centric models defined in~\cite{fri97}, one of which represents the program as the graph of \textit{memory events}.
-The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
-%These models may be further restricted by constraints of a weak memory model, adding \textit{relations} to the memory events.
-%Perhaps, the most convenient way to model the non-deterministic properties of concurrent programs is to use the program representation based on \textit{memory events}.
-%In order to model the non-deterministic properties of concurrent programs
-
-The event-based program model represents the directed graph (the \textit{event-flow graph}), where vertices represent \textit{events} (simple low-level instructions), and edges represent \textit{relations} over the events.
-Below we describe some basic types of events and relations.
+The classical approach to model concurrent programs is to use the \textit{global time}, a single order of interleavings among all events happened in different threads.
+However these models are easy to understand, it may be impossible to process \textit{all} possible states, number of which is exponentially large.
+However, there exist some equivalence classes such that the result of execution different interleavings from single equivalence class is the same (for instance, computations performed by a processor locally do not affect the global state).
+%One way to do this is to use a non-deterministic computation-centric model~\cite{fri97}, such as ...%The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
+One such model is the \textit{event-based} representation of a program, which models a program as a directed graph (the \textit{event-flow graph}).
+The vertices of such a graph represent \textit{events} (simple low-level instructions; see Section~\ref{ch:wmm:model:events}), and edges represent \textit{relations} over the events (see Section~\ref{ch:wmm:model:relations}).
+Thus, a candidate execution constitutes the set of guesses which event (from the set of possible next events) is to be executed.
 
@@
 \label{ch:wmm:model:events}
+
+% TODO: check this subseciton
 
@@
 %\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
+
 The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
-We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\mathtt{load}(r_1 \leftarrow l_2); \ e_m''~=~\mathtt{store}(l_1 \leftarrow r_1)$.
+We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\mathtt{load}(r_1~\leftarrow~l_2); \ e_2~=~\mathtt{store}(l_1~\leftarrow~r_1)$.
+Also, it is important to separate the set of initial write events $\mathbb{IW}~\subset~\mathbb{W}$ that perform initialisation of program variables.
 
 A \textit{computation event} $e_c \in \mathbb{C} \subseteq \mathbb{E}$, represents a low-level assembly computation operation performed solely on local-memory arguments.
-An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (emulated as a register as well) to the register $r_1$.
-For modelling branching statements, we distinguish the set $\mathbb{C_{p}} \subseteq \mathbb{C}$ of \textit{predicative} computation events, that are evaluated as a boolean value.%TODO: TAUTOLOGY HERE.
+An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (which is modelled as a register as well) to the register $r_1$.
+For modelling branching statements, we distinguish the set $\mathbb{C_{p}} \subseteq \mathbb{C}$ of \textit{predicative} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
 
-The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}).
-Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours.
-Technically, a fence may be represented either as a synchronisation barrier, or a flush local memory caches, etc.
+The synchronisation instructions (fences) cause the \textit{barrier events}, that do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours.
+Semantically, a fence may serve as a synchronisation barrier, or as a flush of local memory caches, etc.
 
@@
 
-In this section, we describe basic derived relations used in memory model-awared program analysis.
+%In this section, we describe basic relations used in memory model-awared program analysis.
 
+The relation~$\xrightarrow{r}\,:~\mathbb{E}~\times~\mathbb{E}$ is a binary function over events (set of pairs of events).
+There are two kinds of relations between events: \textit{basic relations} %retrieved from the
+that capture semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification.
+Constraints over relations that are specified by weak memory models are defined as requirements of acyclicity, irreflexivity or emptiness of specific relations~\cite{alglave2016syntax}.
+
+%he read-from relation rf records the correspondance between the reads, the matching writes, and the communicated values on the computation Ξ //From cat.pdf
 %The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
-The \textit{control-flow} of a program is defined by \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of events \textit{from same process}, which never relates events from different processes.
-Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-flow graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$.
-This relation encodes the control-flow of the program into the event-flow graph.
+\begin{itemize}
+	\item
+	The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subset~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
+	For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \xrightarrow{\mathtt{po}} e_2$. 
 	%Some new relations may be acquired : dp, po-loc
 
-The \textit{data-flow} of a program is defined by \textit{communication relations}: the \rf-relation $\subset~\mathbb{W}~\times~\mathbb{R}$ (\textit{read-from} relation) that maps a write event to the read event that reads its value, the \co-relation $\subset~\mathbb{W}~\times~\mathbb{W}$ (\textit{coherence order}, sometimes called \texttt{ws}-relation for \textit{write serialisation}) defines the total order on writes to the same location across all processes, and the \fr-relation $\subset~\mathbb{R}~\times~\mathbb{W}$ (\textit{from-read order}) that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
+	\item
+	The \textit{data-flow} of a program is defined by \textit{communication relations}: the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads its value, the \textit{coherence order}-relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$
+	%(also called \texttt{ws}-relation for \textit{write serialisation})
+	defines the total order on writes to the same location across all processes.
+
+	\item
+	Events from the same process are related by the \textit{scope relation} \sr~$\subset~\mathbb{E}~\times~\mathbb{E}$. 
+	% TODO: note that we do not use the scope tree, just labels
+\end{itemize}
+
+The derived relations can be produced using the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
+
+\begin{itemize}
+	\item \textit{the union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2}
+	\item \textit{the intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2}
+	\item \textit{the difference} of two relations \texttt{r1} and \texttt{r2} is $\mathtt{r1\,\backslash\,r2}$
+	\item \textit{the sequence}
+\footnote{The sequence of two relations \texttt{r1} and \texttt{r2} is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\mathtt{(x,z)} \in \mathtt{r1}$ and $\mathtt{(z,y)} \in \mathtt{r2}$}
+	of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}
+\end{itemize}
+
+%The \textit{from-read order} \fr $\subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
+
 %TODO: perephrase last sentence! alglave thesis, p. 36
@@
 The semantics of a concurrent program is represented by the set of allowed executions.
-A \textit{candidate execution} (trace, run) of a program is an ordered set of events defined by \po and \rf relations and set of final writes to a given memory location~\cite{alglave2014herding}.
+A \textit{candidate execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location~\cite{alglave2014herding}.
 %The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
-
 %associated with the instructions of the program
 %the path in the event-flow graph.
-An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-An execution to be checked on validity is called the \textit{candidate execution}.
+
+%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
+%An execution to be checked on validity is called the \textit{candidate execution}.
 An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
@@
 
+\subsection{Weak Memory Models}
+\label{ch:wmm:wmm}
+
 \section{The CAT language}

### 2018-04-19 1_intro.tex
Some editions. Rename output folder out -> build.

@@
 It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
-The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
+The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, \tool{diycross} and \tool{diyone}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
 

### 2018-04-19 2_wmm.tex
Some editions. Rename output folder out -> build.

@@
 
-The derived relations can be produced using the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
-
-\begin{itemize}
-	\item \textit{the union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2}
-	\item \textit{the intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2}
-	\item \textit{the difference} of two relations \texttt{r1} and \texttt{r2} is $\mathtt{r1\,\backslash\,r2}$
-	\item \textit{the sequence}
-\footnote{The sequence of two relations \texttt{r1} and \texttt{r2} is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\mathtt{(x,z)} \in \mathtt{r1}$ and $\mathtt{(z,y)} \in \mathtt{r2}$}
-	of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}
-\end{itemize}
-
 %The \textit{from-read order} \fr $\subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
-
 %TODO: perephrase last sentence! alglave thesis, p. 36
@@
 
-The semantics of a concurrent program is represented by the set of allowed executions.
-A \textit{candidate execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location~\cite{alglave2014herding}.
+%The semantics of a concurrent program is represented as the set of allowed executions.
+An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
+A \textit{candidate execution} is an execution that is not constrained by the memory model (i.e., the execution for that the validity is not checked).
+As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
 %The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
@@
 %the path in the event-flow graph. 
+%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
 
-%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-%An execution to be checked on validity is called the \textit{candidate execution}.
-An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
-As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program as a whole.
-Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus-test Example~\ref{simple_wmm_x86} (the pictures are generated by the \texttt{herd7} tool, version 7.47).
+%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
+The Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \texttt{herd7} tool, version 7.47).
 Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
@@
 
-\section{The CAT language}
+\section{The \cat language}
 
-One common way to define a weak memory model in a systematic way in to formalise it in \textit{the CAT language} proposed in~\cite{alglave2016syntax}.
-The CAT language allows to axiomatically define new relations, new fences and restrictions over relations.
+Weak memory models can be described in the \cat language~\cite{alglave2016syntax}.
+It is the domain specific language for describing consistency properties of concurrent programs.
+The \cat language combines the expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, expressions and assertions specific for operating with relations and sets of executions.
+% succinctly by constraining an abstraction of parallel program executions into a candidate execution and possibly extending this candidate execution with additional constraints on the execution environment.
+
+%represents the functional language extended by the theory of  allows to axiomatically define derived relations, architecture-specific fences, and constraints over relations.
+
+The definition of new relations  derived relations can be produced using the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
+
+\begin{itemize}
+	\item \textit{the union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2}
+	\item \textit{the intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2}
+	\item \textit{the difference} of two relations \texttt{r1} and \texttt{r2} is $\mathtt{r1\,\backslash\,r2}$
+	\item \textit{the sequence}\footnote{The sequence of two relations \texttt{r1} and \texttt{r2} is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\mathtt{(x,z)} \in \mathtt{r1}$ and $\mathtt{(z,y)} \in \mathtt{r2}$}
+	of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}
+\end{itemize}
 

### 2018-04-19 1_intro.tex
Fix orphography for some files.

@@
 
-Research of weak memory models firstly aimes to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
+Research of weak memory models firstly aims to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
 The first (and so far the only) such a framework was presented in 2010~\cite{alglave2010shared}.
 %
-Secondly, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompletenesses.
+Secondly, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompleteness.
 Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding}, for Alpha~\cite{?} and Sparc~\cite{?} memory models were defined in the specification.
@@
 
-The SC model is considered to be the strong memory model in the sence that it provides strong guarantees regarding the ordering and caused effect of memory operations.
+The SC model is considered to be the strong memory model in the sense that it provides strong guarantees regarding the ordering and caused effect of memory operations.
 Different relaxations of this model lead to the class of \textit{weak memory models~(WMM)}.
@@
 The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France and University of Cambridge, UK.
-The \tool{diy}\footnote{Project web page: \url{diy.inria.fr}} is a sofware suite for designing and testing weak memory models.
+The \tool{diy}\footnote{Project web page: \url{diy.inria.fr}} is a software suite for designing and testing weak memory models.
 It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
@@
 
-The first tool that performs the WMM-awared portability analysis is \tool{PORTHOS} introduced in April~2017~\cite{Porthos17}.
+The first tool that performs the WMM-aware portability analysis is \Porthos introduced in April~2017~\cite{Porthos17}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
@@
 
-Current work aims to rework the proof-of-concept tool \tool{PORTHOS} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
+Current work aims to rework the proof-of-concept tool \Porthos by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
 %interprocedural analysis
@@
 
-The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory model-awared analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
+The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory model-aware analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
 Chapter~...

### 2018-04-19 2_wmm.tex
Fix orphography for some files.

@@
-\chapter{Memory Model-Awared Program Analysis}
+\chapter{Memory Model-Aware Program Analysis}
 \label{ch:wmm}
@@
 %possible communications between threads of the program.
-Obtainted model of the program is called an anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
+Obtained model of the program is called an anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
 Thereafter, the anarchic semantics is constrained by the \textit{weak memory model} specification which is a set of axiomatic constraints for filtering out executions inconsistent in particular architecture.
@@
 
-The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
+The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
 We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\mathtt{load}(r_1~\leftarrow~l_2); \ e_2~=~\mathtt{store}(l_1~\leftarrow~r_1)$.
@@
 
-%In this section, we describe basic relations used in memory model-awared program analysis.
+%In this section, we describe basic relations used in memory model-aware program analysis.
 

### 2018-04-20 1_intro.tex
Cleanup a lot. Fix titles.

@@
 The first (and so far the only) such a framework was presented in 2010~\cite{alglave2010shared}.
-%
-Secondly, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, that are written in natural language and thus suffer from ambiguities and incompleteness.
-Over last decade the memory models have been extracted for most mainstream multiprocessor architectures, such as x86-TSO (\textit{Total Store Order}) model for x86 architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding}, for Alpha~\cite{?} and Sparc~\cite{?} memory models were defined in the specification.
-Moreover, most modern high-level programming languages rely on relaxed memory model.
-Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model is based on the relation \textit{strongly happens-before}.
-%batty2011mathematizing
-Thirdly, important research direction targets the problem of verifying (or at least finding bugs in) existing software systems with respect to WMMs.
-Perhaps, the most notable work in this field is the project for defining the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
-%Distributed databases <also need the wmm, see transaction consistency~\cite{bailis2013highly}>
+In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
+Over last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, and others.
+There are projects for validating hardware architectures wrt. a memory model, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.
+
+Most modern high-level programming languages rely on relaxed memory model as well.
+Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
+Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
+Furthermore, there is a wide range of tools that perform program verification wrt memory models (see~\cite{alglave2013software}, \cite{Porthos17}).
 
@@
 %interprocedural analysis
-We called the new tool \textbf{the \tool{mousquetaires} framework} as now it not only performs the portability analysis, but can serve as the basis for SMT-solver driven weak memory model analysis.
+We called the new tool the~\tool{mousquetaires}~framework as now it not only performs the portability analysis, but can serve as the basis for SMT-solver driven weak memory model analysis.
 

### 2018-04-20 2_wmm.tex
Cleanup a lot. Fix titles.

@@
-\chapter{Memory Model-Aware Program Analysis}
+\chapter{Memory model-aware analysis}
+%\chapter{A generic framework}
 \label{ch:wmm}
@@
 %"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf
-
-The classical approach to model concurrent programs is to use the \textit{global time}, a single order of interleavings among all events happened in different threads.
-However these models are easy to understand, it may be impossible to process \textit{all} possible states, number of which is exponentially large.
-However, there exist some equivalence classes such that the result of execution different interleavings from single equivalence class is the same (for instance, computations performed by a processor locally do not affect the global state).
+The classical approach for modeling concurrent programs is to use the \textit{global time}, a single order of interleavings among all events happened in different threads.
+Although these models are easy to understand, it may be impossible to process \textit{all} possible states, number of which is exponentially large.
+However, there exist equivalence classes such that the result of execution different interleavings from single equivalence class is the same (for instance, computations performed by a processor locally do not affect the global state).
 %One way to do this is to use a non-deterministic computation-centric model~\cite{fri97}, such as ...%The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
-One such model is the \textit{event-based} representation of a program, which models a program as a directed graph (the \textit{event-flow graph}).
-The vertices of such a graph represent \textit{events} (simple low-level instructions; see Section~\ref{ch:wmm:model:events}), and edges represent \textit{relations} over the events (see Section~\ref{ch:wmm:model:relations}).
-Thus, a candidate execution constitutes the set of guesses which event (from the set of possible next events) is to be executed.
+One such model is the \textit{event-based} representation of a program, which models a program as a directed graph of events (the \textit{event-flow graph}).
+The vertices of such a graph represent \textit{events} (independent low-level instructions; see Section~\ref{ch:wmm:model:events}), and edges represent \textit{relations} over the events (see Section~\ref{ch:wmm:model:relations}).
 
@@
 The synchronisation instructions (fences) cause the \textit{barrier events}, that do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours.
-Semantically, a fence may serve as a synchronisation barrier, or as a flush of local memory caches, etc.
+Functionally, a fence may be a synchronisation barrier or a instruction of flushing the local memory caches, etc.
 
@@
 
-%he read-from relation rf records the correspondance between the reads, the matching writes, and the communicated values on the computation Ξ //From cat.pdf
-%The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).
+\vspace{1em}
+The basic relations are the following~\cite{alglave2010shared}:
+
 \begin{itemize}
-	\item
-	The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subset~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
-	For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \xrightarrow{\mathtt{po}} e_2$. 
+    \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subset~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
+	For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relation{po} e_2$.
 	%Some new relations may be acquired : dp, po-loc
 
-	\item
-	The \textit{data-flow} of a program is defined by \textit{communication relations}: the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads its value, the \textit{coherence order}-relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$
-	%(also called \texttt{ws}-relation for \textit{write serialisation})
-	defines the total order on writes to the same location across all processes.
-
-	\item
-	Events from the same process are related by the \textit{scope relation} \sr~$\subset~\mathbb{E}~\times~\mathbb{E}$. 
-	% TODO: note that we do not use the scope tree, just labels
+    \item The \textit{data-flow} of a program is defined by \textit{communication relations}:
+        \begin{itemize}[noitemsep]
+            \item the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads its value;
+            \item the \textit{coherence order} relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes         (also called the \textit{write serialisation} \texttt{ws}-relation);
         \end{itemize}
 
+	\item Events from the same process are related by the \textit{scope relation} \sr~$\subset~\mathbb{E}~\times~\mathbb{E}$.
+    In contrast to the \tool{herd} tool, the \mousquetaires does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple lables that indicate which process has produced certain event.
+\end{itemize}
+
+%\vspace{0.5em}
+Below we enumerate some derived relations~\cite{alglave2010shared}:
+
+\begin{itemize}
+
+    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to writes preceding the write event from which the read reads the value: \\
+    $r \relation{fr} w = (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$.
+
+    \item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on).
+    %$\mathtt{com} = (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
+
+    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes.
+
+    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable:
+    $m_1 \relation{po-loc} m_2 = (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$.
+
+    \item the semantics of fences and barriers specific for different architectures may be defined as derived relations.
+
+    %"Preserved program order defines the set of reorderings guaranteed by the architecture not to occur, based on the types of the accesses." http://delivery.acm.org/10.1145/2750000/2742219/p635-lustig.pdf
+
+\end{itemize}
+
+The work~\cite{alglave2010shared} (Chapter~2) provides definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}.
+Thereafter, weak memory models make asserts over these properties, thus restricting the set of allowed behaviours of the system.
+
+% TODO: some derived relations
 %The \textit{from-read order} \fr $\subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
@@
 
+The \textit{candidate execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
+%the sequence of choices from guesses which event is to be executed next.
 %The semantics of a concurrent program is represented as the set of allowed executions.
-An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
-A \textit{candidate execution} is an execution that is not constrained by the memory model (i.e., the execution for that the validity is not checked).
-As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
+%An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
+%A \textit{candidate execution} is an execution that is not constrained by the memory model (i.e., the execution for that the validity is not checked).
+%As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
 %The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
@@
 %An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-
 %An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
-The Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \texttt{herd7} tool, version 7.47).
+Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \tool{herd7} tool, version~7.47).
 Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
@@
 Thus, there are only four possible executions defined by the choice of \rf-relation.
-The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and others.
-However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over \fr$\cup$\po.
-%The event $b$:\texttt{(Ry=0)} reads value \texttt{0} at the shared location \texttt{y} from the initial write event $e$\texttt{(Wy=0)} (the red edge of \texttt{rf}-relation), consequently, 
-%the event $d$:\texttt{(Rx=0)} reads value \texttt{0} at the shared location \texttt{x} from the initial write event $f$\texttt{(Wx=0)} (the red edge of \texttt{rf}-relation). 
+The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and some others.
+However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over~$\fr\cup\po$.
 %However, the Power memory model allows such cycles, therefore %TODO: include load buffering example, see presentation pdf
 
-%\includegraphics[width=0.4\textwidth]{img/my/simple_wmm_x86.png}
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 \hfill
-\caption{Candidate executions for the litmus test Example~\ref{simple_wmm_x86}}
+\caption{Possible candidate executions for the litmus test Example~\ref{simple_wmm_x86}}
 \label{simple_wmm_x86_pic}
@@
 
-\subsection{Weak Memory Models}
-\label{ch:wmm:wmm}
+\section{The \cat language}
+\label{ch:wmm:cat}
 
-\section{The \cat language}
-
-Weak memory models can be described in the \cat language~\cite{alglave2016syntax}.
-It is the domain specific language for describing consistency properties of concurrent programs.
-The \cat language combines the expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, expressions and assertions specific for operating with relations and sets of executions.
-% succinctly by constraining an abstraction of parallel program executions into a candidate execution and possibly extending this candidate execution with additional constraints on the execution environment.
-
+Weak memory models are defined via the \cat language~\cite{alglave2016syntax}.
+This is a domain specific language for describing consistency properties of concurrent programs.
+The \cat language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, expressions and assertions that are specific for operating with relations and executions.
 %represents the functional language extended by the theory of  allows to axiomatically define derived relations, architecture-specific fences, and constraints over relations.
 
-The definition of new relations  derived relations can be produced using the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
+The derived relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
 
 \begin{itemize}
-	\item \textit{the union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2}
-	\item \textit{the intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2}
-	\item \textit{the difference} of two relations \texttt{r1} and \texttt{r2} is $\mathtt{r1\,\backslash\,r2}$
-	\item \textit{the sequence}\footnote{The sequence of two relations \texttt{r1} and \texttt{r2} is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\mathtt{(x,z)} \in \mathtt{r1}$ and $\mathtt{(z,y)} \in \mathtt{r2}$}
-	of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}
+	\item \textit{the union} of two relations \relname{r1} and \relname{r2} is \relname{r1\,|\,r2}
+	\item \textit{the intersection} of two relations \relname{r1} and \relname{r2} is \relname{r1\,\&\,r2}
+	\item \textit{the difference} of two relations \relname{r1} and \relname{r2} is $\mathtt{r1\,\backslash\,r2}$
+	\item \textit{the sequence}\footnote{The sequence of two relations \relname{r1} and \relname{r2} is defined as the set of pairs \relname{(x,y)} such that there exists an intervening \relname{z}, such that $\relname{(x,z)} \in \relname{r1}$ and $\relname{(z,y)} \in \relname{r2}$}
+	of two relations \relname{r1} and \relname{r2} is \relname{r1;r2}
 \end{itemize}
+
+For instance, Figure~\ref{example:x86-cat} contains part of x86-TSO model~\cite{owens2009better} that asserts acyclicity of communication relation and \relname{po-loc} relation:
+
+\begin{figure}[!h]
+\begin{lstlisting}
+...
+let com = rf | fr | co
+let po-loc = po & loc
+acyclic po-loc | com
+\end{lstlisting}
+\caption{Excerpt from the x86-TSO memory model in \cat language}
+\label{example:x86-cat}
+\end{figure}
 
@@
 
-%There is a barrier instruction \texttt{mfence} that may be used for flushing the buffers into the main memory.
+%There is a barrier instruction \relation{mfence} that may be used for flushing the buffers into the main memory.
 
-briefly some hw memory models: X86-TSO, Alpha, POWER, ;
+%briefly some hw memory models: X86-TSO, Alpha, POWER, ;
 
-language memory models: Java, C++;
+%language memory models: Java, C++;
 
-library-level kernel memory model, ref to github with tests
+%library-level kernel memory model, ref to github with tests
 
@@
 
-\section{Some examples of WMM}
+%\section{Some examples of WMM}
 
-//axioms of TSO wmm
+%//axioms of TSO wmm
 
-// example of sets: rf, co, ... for the code snipped used before
+%// example of sets: rf, co, ... for the code snipped used before

### 2018-04-20 3_encod.tex
Cleanup a lot. Fix titles.

@@
-\chapter{The Portability of Concurrent Software}
-%\chapter{The Concurrent Software Portability Analysis as a Bounded Model Checking Problem}
+%\chapter{SMT-based program analysis}
+%\chapter{Portability of concurrent software}
+\chapter{Portability analysis as a SAT~problem}
 \label{ch:port}
@@
 %may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to the satisfiability modulo theories (SMT) problem~\cite{Porthos17}.
+as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17} (more precisely, to an SMT problem).
 
-\section{The model checking problem}
+\section{Model checking}
 \label{ch:port:mc}
@@
 
-\section{The portability as a bounded reachability problem}
+\section{Portability analysis as a bounded reachability problem}
 \label{ch:port:enc}
 
-In general, the BMC problem aims to examine the reachability of the "undesirable" states of a finite-state system.
+In general, a BMC problem aims to examine the reachability of the "undesirable" states of a finite-state system.
 Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $Init(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $Trans(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $Bad(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
@@
 
-The portability probram may also be formulated as a reachability problem, where the undesirable state is the state reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
+Portability analysis problem may also be stated as a reachability problem, where the undesirable state is the state reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
 However, unlikely the BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:x2y:unrolling}) and the $Trans$ predicate may be stated only for the final state of a program.
@@
 
-Note, that the formulation of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.
+Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.
 
 %It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the bounded reachability problem.
-The result SMT formula $\phi$ of the portability problem should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
+The result SMT formula $\phi$ of the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
 If the formula is satisfiable, there exist a portability bug.
@@
 
-\subsection{The control-flow encoding} %Encoding the control-flow constraints}
+\subsection{Encoding for the control-flow} %Encoding the control-flow constraints}
 \label{ch:port:enc:cf}
@@
 The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:x2y:unrolling}).%TODO: footnote why multiple sinks
-In control-flow graph, there are two types of edges: \textit{primary edges} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative edges} that denote if-false-transitions (pictured with dotted lines).
-Each node on graph can have either one successor (primary) or two successors (primary and alternative); only computation events can serve as a branching point).
+In control-flow graph, there are two types of transitions (edges): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
+Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
 However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.
@@
 For that, we compile the recursive abstract syntax tree~(AST) of the parsed C-code to the plain (non-recursive) event-flow graph.
-We show%TODO! to show, actually
+We show %TODO! to show, actually
 that the new encoding is smaller than the old one used in \Porthos since it does not produces new variables for each high-level statement of the input language.
@@
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, in old scheme the control-flow of the sequential instruction \m{i_1 := i_2; i_3} was encoded as 
-\m{\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}, 
-and control-flow of the branching instruction \m{i_1 := \{c \, ? \, i_2 \, : \, i_3\}} was encoded as 
-\m{\phi_{CF}( \, c \, ?
-\, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)} (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `\texttt{if} $x$ \texttt{then} $y$ \texttt{else} z').
+For instance, \Porthos uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
+and control-flow of the branching instruction $i_1 = (c \, ? \, i_2 \, : \, i_3)$ was encoded as
+$\phi_{CF}( \, c \, ? \, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
+(here we used the notation of C-like ternary operator \texttt{x?y:z} for defining the conditional expression $\mathtt{if} x \mathtt{then} y \mathtt{else} z$).
 In contrast, the new scheme implemented in \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
-The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see further in current Chapter).
-%linearly as it shown in
+The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).
 
-Let \m{\fx : \mathbb{E} \rightarrow \{0,1\}} be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system). 
-Let \m{\fv : \mathbb{C} \rightarrow \mathbb{N}} be the function that returns the value of the computation event that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \texttt{rf}-relation; see Chapter~?
+Let $\fx : \mathbb{E} \rightarrow \{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
+Let $\fv : \mathbb{C} \rightarrow \mathbb{N}$ be the function that returns the value of the computation event that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \relation{rf}-relation; see Chapter~?
 for the relations encoding%TODO
 ).
-We distinguish the function \m{\fv_p : \mathbb{C_p} \rightarrow \{0,1\}} that evaluates the predicative computation event.
+We distinguish the function $\fv_p : \mathbb{C_p} \rightarrow \{0,1\}$ that evaluates the predicative computation event.
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
 
-Consider the following possible mutual arrangement of nodes in a control-flow grpah:
+Consider the following possible mutual arrangement of nodes in a control-flow graph:
 
@@
     \end{subfigure}
+    %
     \caption{Linear and non-linear cases of control-flow graph}
@@
 \begin{align}
-%\phi_{CF_{(e_1;e_2)}}     =
-%\phi_{CF_{(e_1?e_2:e_3)}} =
 \phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
@@
 
-\subsection{The data-flow encoding} %constraints}
+For the sake of encoding correctness, we require all branches to have at least one event.
+Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \texttt{nop}-event as it is shown in Figure~\ref{encode:branching:nop}:
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+        \node[c] at (0,0) (1a) {$e_1$};
+        \node[c] at (0,-1.5) (2a) {$e_2$};
+        \node[c] at (0,-3) (3a) {$e_3$};
+        \node[c,draw=none] at (3,-1.5) (arr) {$\longrightarrow$};
+        \node[c] at (6,0) (1b) {$e_1$};
+        \node[c] at (6,-1.5) (2b) {$e_2$};
+        \node[c,fill=black!20] at (7.5,-1.5) (nop) {$e_{\mbox{\tiny NOP}}$}; %{\tiny{NOP}};
+        \node[c] at (6,-3) (3b) {$e_3$};
+        \path[->]
+        (1a) edge [] node {} (2a)
+        (2a) edge [] node {} (3a)
+        (1a) edge [dotted,bend left=60] node {} (3a)
+        (1b) edge [] node {} (2b)
+        (2b) edge [] node {} (3b)
+        (1b) edge [dotted,bend left] node {} (nop)
+        (nop) edge [dotted,bend left] node {} (3b)
+        ;
+    \end{tikzpicture}
+    \caption{Transformation of the empty-branch nonlinear control-flow}
+    \label{encode:branching:nop}
+\end{figure}
+
+\subsection{Encoding for the data-flow}
 \label{ch:port:enc:df}
 
-While encoding the data-flow constraints, we use the \textit{Static Single-Assignment~(SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
+To encode the data-flow constraints, we use the \textit{static single-assignment (SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
 The SSA form requires each variable to be assigned only once within entire program.
-That is, on each state of the program a variable (either local or shared) is mapped to its indexed reference, which represents a new boolean variable in the formula.
-On merge-points, the $\phi$-function is emulated as the disjunction of values of the variable that is computed as the \textit{reaching definition}.
-For that, we carry the map that for each event stores mapping from all declared variables to their SSA-indices.
-All indices are globally unique within single program analysis.
+In contrast, \Porthos used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
+Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17}.
 
-As it is implemented in~\cite{Porthos17}, the SSA-indices of variables are computed in accordance with the following rules: (1) any access to a shared variable (both read and write) increments its SSA-index; (2) only writes to a local variable increment its SSA-index (reads preserve indices); (3) no access to a constant variable or computed expression changes their SSA-index.
+Following~\cite{Porthos17}, the indexed references of variables are computed in accordance with the following rules:
+(1)~any access to a shared variable (both read and write) increments its SSA-index;
+(2)~only writes to a local variable increment its SSA-index (reads preserve indices);
+(3)~no access to a constant variable or computed expression changes their SSA-index.
 These rules determine the following encoding of load, store and computation events within single thread:
 %TODO: maybe add the overall formula \phi_{DF} ?
-
 \begin{align}
@@
 
-Unlikely the \Porthos tool, the \mousquetaires uses the algorithm of converting the program into the SSA form was also adapted for processing the plain flow-graph data structure.
-That is, for each node we keep track the set of its predecessors, and then, while traversing the graph in topological order, we propagate the SSA-indices for each variable defined so far, and update these indices according the rules listed above.
+To convert the program into SSA form, for each event each variable that is declared so far (either local or shared) is mapped to its indexed reference; this information is stored in the SSA-map "event \textit{to} variable \textit{to} SSA-index".
+%for each event stores the map that each declared variable maps to its SSA-index
+%"$\{$ event : $\{$ variable : index $\}\}$",
+%"$\left\{$ event : $\left\{$ variable : index $\right\} \right\}$",
+The SSA-map is computed iteratively while traversing the event-flow graph in topological order as it is described in Algorithm~\ref{algorithm:ssa-map}:
+
+\begin{algorithm}
+    \caption{Algorithm for computing the SSA-indices}\label{alg:compute-ssa}
+    \algorithmicrequire{The event-flow graph $G = \langle N, E \rangle$ where $V$ is the set of nodes (events), $E$ is the set of control-flow transitions, $e_0$ is the entry node}
+    \algorithmicensure{The SSA-map of the form "$\{$ event : $\{$ variable : index $\}\}$"}
+    \begin{algorithmic}[1]
+        \Function{ComputeSSA}{G}
+            \State {$S \gets$ empty map; $S[e_0] \gets$ empty map}
+            \ForAll {event $e_i \in G.N$ in topological order}
+                \ForAll {predecessor $e_j \in \mathtt{pred}(e_i)$}
+                    \State {$S[e_i] \gets \mathtt{copy}(S[e_j])$}
+                    \ForAll {variable $v_k \in$ set of variables accessed by $e_i$ }
+                        \State {$S[e_i][v_k] \gets \mathtt{max}(S[e_i][v_k], \ S[e_j][v_k])$}
+                        \If {need to update the index of $v_k$} \Comment {cases (1)-(2)}
+                            \State {$S[e_i][v_k] \gets S[e_i][v_k] + 1$}
+                        \EndIf
+                    \EndFor
+                \EndFor
+            \EndFor
+        \EndFunction
+    \end{algorithmic}
+    \label{algorithm:ssa-map}
+\end{algorithm}
 
 The time of described algorithm is linear of the size of event-flow graph since it performs only single traverse of the graph.
-Notwithstanding the overhead of storing (or, equivalently but undesirably, computing with%TODO: are you sure that 'with'?
-linear time) the predecessors of each node in order to be able to convert the program into an SSA form, the time of such a tranformation reduced%TODO: How much ????
-comparing to the algorithm implemented in \Porthos, which recomputed SSA-indices recursively for each instruction.%TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
+%Notwithstanding the overhead of storing (or, equivalently, computing with %TODO: are you sure that 'with'?
+%linear time) the predecessors of each node in order to be able to convert the program into an SSA form, the time of such a transformation reduced %TODO: How much ????
+%comparing to the algorithm implemented in \Porthos, which recomputed SSA-indices recursively for each instruction. %TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
 
-Moreover, the data-flow encoding scheme used in \Porthos faced the problem when the branching statements had different number of assignments of certain variable in their branches.
-//TODO: why?
-why don't we have this problem now?
-
-As it has been stated before, the \rf-relation links the parts%TODO: another word?
+As it has been described before, the \rf-relation links data-flow between events %TODO: another word?
 of data-flow stored in equivalence assertions over the SSA-variables. 
 %The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
-The encoding of this linkage left untouch as it is implemented in \Porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
+The encoding of this linkage left untouched as it is implemented in \Porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
 
@@
 
-\subsection{The memory model encoding} %Encoding the memory model constraints}
+\subsection{Encoding for the memory model} %Encoding the memory model constraints}
 \label{ch:port:enc:wmm}
 
-The encoding of a weak memory model itself does not depend on the analysing program. The problem here...
+%The encoding of a weak memory model itself does not depend on the analysing program. The problem here...
 
@@
 
-As it is described in~\cite{Porthos17}, the is based ...
+As it is described in~\cite{Porthos17}, the is based . . .
 %~\cite{Porthos17}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"

### 2018-04-26 1_intro.tex
Consider Prof. Heljanko's comments.

@@
 \begin{figure}
-\small \ttfamily
+\centering
+\ttfamily
 \begin{tabular}{ |l|l| }
@@
 Most modern high-level programming languages rely on relaxed memory model as well.
-Thus, the memory model for Java that is based on the \textit{happens-before} principle~\cite{lamport1978time} was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{?} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
+Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time}, it was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{iso2012iec} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
 Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.

### 2018-04-26 1_intro.tex
add minor fixes and multiple todo-comments.

@@
 \label{ch:intro}
+
+%TODO: rename mousquetaires: http://www.cprover.org/wmm/musketeer/ !!! (+reference this project: "Don’t sit on the fence".pdf in alglave/ )
+% TODO: see how musketeer deals with recursive calls: http://www.cprover.org/wmm/musketeer/rec.pdf
 
@@
 \end{tabular}
-\caption{Store buffering (SB): a litmus test on write-read reordering allowed under the x86-TSO weak memory model}
+\caption{Store buffering (SB): a litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
 \label{simple_wmm_x86}
@@
 The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France and University of Cambridge, UK.
-The \tool{diy}\footnote{Project web page: \url{diy.inria.fr}} is a software suite for designing and testing weak memory models.
+The \tool{diy}\footnote{Project web site: \url{http://diy.inria.fr/}} is a software suite for designing and testing weak memory models.
 It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
@@
 
-The Chapter~\ref{ch:wmm} gives more detailed description of the weak memory model-aware analysis and provides description of memory models for some common architectures (x86, ARM and POWER, Sparc, ???).
-Chapter~...
+The thesis is organised as following. Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis. Chapter~...
 
 % TODO: Add intro to the bounded reachability analysis using SAT -- ?
+
+% TODO: other tools for bounded model-checking: 
+%Clarke's goto-cc : \cite{clarke2004tool} : http://www.cprover.org/goto-cc/
+% (perhaps, goto-cc is a part of cprover) : cprover https://www.cprover.org/cbmc/doc/manual.pdf
+%cbmc http://www.cprover.org/cbmc/

### 2018-04-26 2_wmm.tex
add minor fixes and multiple todo-comments.

@@
 In general, analysis of concurrent programs with respect to axiomatic memory models is performed in several stages.
-Firstly, the control-flow and data-flow of a program is encoded as the set of possible \textit{candidate executions}.
-%possible communications between threads of the program.
+%TODO: site the 'Weak Memory Models with Matching Axiomatic and Operational Definitions' https://arxiv.org/pdf/1710.04259.pdf
+%TODO: cite also: "DENNIS SHASHA - Efficient and Correct Programs that Share" http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.142.2238&rep=rep1&type=pdf
+Firstly, the control-flow and data-flow of a program is encoded as the set of possible \textit{candidate executions}. %possible communications between threads of the program.
 Obtained model of the program is called an anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
@@
 The vertices of such a graph represent \textit{events} (independent low-level instructions; see Section~\ref{ch:wmm:model:events}), and edges represent \textit{relations} over the events (see Section~\ref{ch:wmm:model:relations}).
+
+% TODO: the event-based representation is a non-deterministic model
+
+% TODO: the semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
+%"To build the aeg , we define a semantics of goto-programs in terms of abstract events, ..."
+% TODO: OR from here: http://www.cprover.org/etaps/esop.pdf "Software Verification for Weak Memory via Program Transformation"
 
@@
 
+% TODO: more fences: "On Power and ARM (see Sec.
+6), we distinguish between control fence (which we write cfence), lightweight fence (lwfence) and full fence (ffence).
+On x86, implementing TSO, there is only one fence, called mfence." (herding cats, p.
+14)
+
 \subsection{Relations}
@@
 The basic relations are the following~\cite{alglave2010shared}:
-
+%TODO: say that po and rf are  defined from the candidate execution:
+%TODO: say that 
 \begin{itemize}
@@
 
-%\vspace{0.5em}
+%TODO: add ppo = characteristics of memory model.
+"Architectures are instances of our model.
+An architecture is a triple of functions (ppo, fences, prop), which specifies the preserved program order ppo, the fences fences and the propagation order prop." (herding cats, p.13)
+
 Below we enumerate some derived relations~\cite{alglave2010shared}:
@@
     \item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on).
+    %TODO: say that po-relation is *static*
     %$\mathtt{com} = (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
@@
 
-The work~\cite{alglave2010shared} (Chapter~2) provides definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}.
+The work~\cite[Chapter~2]{alglave2010shared} provides definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}.
 Thereafter, weak memory models make asserts over these properties, thus restricting the set of allowed behaviours of the system.
@@
 \begin{itemize}
-	\item \textit{the union} of two relations \relname{r1} and \relname{r2} is \relname{r1\,|\,r2}
-	\item \textit{the intersection} of two relations \relname{r1} and \relname{r2} is \relname{r1\,\&\,r2}
-	\item \textit{the difference} of two relations \relname{r1} and \relname{r2} is $\mathtt{r1\,\backslash\,r2}$
-	\item \textit{the sequence}\footnote{The sequence of two relations \relname{r1} and \relname{r2} is defined as the set of pairs \relname{(x,y)} such that there exists an intervening \relname{z}, such that $\relname{(x,z)} \in \relname{r1}$ and $\relname{(z,y)} \in \relname{r2}$}
-	of two relations \relname{r1} and \relname{r2} is \relname{r1;r2}
+	\item \textit{the union} of two relations \rel{r1} and \rel{r2} is \rel{r1\,|\,r2}
+	\item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2}
+	\item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$
+	\item \textit{the sequence}\footnote{The sequence of two relations \rel{r1} and \rel{r2} is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$}
+	of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}
 \end{itemize}
 
-For instance, Figure~\ref{example:x86-cat} contains part of x86-TSO model~\cite{owens2009better} that asserts acyclicity of communication relation and \relname{po-loc} relation:
+For instance, the \fr-relation is defined as follows: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$.
+
+Figure~\ref{example:x86-cat} contains part of x86-TSO model~\cite{owens2009better} that asserts acyclicity of communication relation and \rel{po-loc} relation:
 
@@
 
-%// exmaple with reordering
-%// ex. with
-%// Rev-29 Example 7-6. Stores Are Transitively Visible. %see http://www.cl.cam.ac.uk/~pes20/weakmemory/x86tso-paper.pdf
-
-%There is a barrier instruction \relation{mfence} that may be used for flushing the buffers into the main memory.
-
-%briefly some hw memory models: X86-TSO, Alpha, POWER, ;
-
-%language memory models: Java, C++;
-
-%library-level kernel memory model, ref to github with tests
-
 %Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}
-
-%\section{Some examples of WMM}
-
-%//axioms of TSO wmm
-
-%// example of sets: rf, co, ... for the code snipped used before

### 2018-04-26 3_encod.tex
Consider Prof. Heljanko's comments.
add minor fixes and multiple todo-comments.

@@
 %\chapter{Portability of concurrent software}
-\chapter{Portability analysis as a SAT~problem}
+\chapter{Portability analysis as an SMT~problem}
 \label{ch:port}
@@
 
-\section{Model checking}
+\section{Model checking and reachability analysis}
 \label{ch:port:mc}
 
-The classical model checking algorithms explore the state space of an abstract automata or a transition system in order to find states that violate the specification.
+The model checking is the problem of verifying the system (the model) against the set of constraints (the specification).
+As the state machine model is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
 The general schema of model checking is the following: firstly, the analysing system is being represented as a transition system, a finite directed graph with labeled nodes representing states of the system such that each state corresponds to the unique subset of atomic propositions, that characterise the behavioral properties of each state.
@@
 
-However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables of the system.
+However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables used by the system (its size).
 In case of modeling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
@@
 
-As we have said before, a program $P$ is called portable from the source architecture (defined as a memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if \textit{all} executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:
+Consider the function $cons_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
+Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:
 
@@
 Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.
-
-%It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the bounded reachability problem.
 The result SMT formula $\phi$ of the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
@@
 Let $\fx : \mathbb{E} \rightarrow \{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
-Let $\fv : \mathbb{C} \rightarrow \mathbb{N}$ be the function that returns the value of the computation event that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \relation{rf}-relation; see Chapter~?
-for the relations encoding%TODO
+Let $\fv : \mathbb{C} \rightarrow \mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \relation{rf}-relation; see Chapter~\textbf{?TODO?} for the relations encoding%TODO
 ).
@@
         }
-        \caption{Sequential events}
+        \caption{Sequential events (\texttt{seq})}
         \label{encode:cf:seq}
@@
         }
-        \caption{Conditional branching}
+        \caption{Conditional branching (\texttt{br})}
         \label{encode:cf:br}
@@
         }
-        \caption{Branch merging}
+        \caption{Branch merging (\texttt{mer})}
         \label{encode:cf:merge}
@@
 For listed cases, below we propose the encoding scheme that uniquely encodes each node of graph and allows to encode partially executed program.
-The Equation~\ref{enc:cf:seq} for encoding the sequential control-flow represented in Figure~\ref{encode:cf:seq} reflects the fact that the event $e_2$ can be executed iff the event $e_1$ was executed.
-The Equation~\ref{enc:cf:br} for encoding the branching control-flow depicted in Figure~\ref{encode:cf:br} allows only following executions: $\{\emptyset, (e_1), (e_1 \rightarrow e_2), (e_1 \rightarrow e3) \}$.
+Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
+Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, (e_1), (e_1~\rightarrow~e_2), (e_1 ~\rightarrow~e_3)\}$.
 In encoding~\ref{enc:cf:merge} of the merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
@@
 				  & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
-\phi_{CF_{merge}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
+\phi_{CF_{mer}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
 \end{align}
 
-For the sake of encoding correctness, we require all branches to have at least one event.
+For sake of encoding correctness, we require all branches to have at least one event.
 Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \texttt{nop}-event as it is shown in Figure~\ref{encode:branching:nop}:
@@
 (2)~only writes to a local variable increment its SSA-index (reads preserve indices);
-(3)~no access to a constant variable or computed expression changes their SSA-index.
+(3)~no access to a constant variable or computed (evaluated) expression changes their SSA-index.
 These rules determine the following encoding of load, store and computation events within single thread:
@@
     \phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} = \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
-    \phi_{DF_{e = \mathtt{compute}(...)}} = \ & \fx(e) \rightarrow \fv(e) % \\
+    \phi_{DF_{e = \mathtt{eval}(...)}} = \ & \fx(e) \rightarrow \fv(e) % \\
 \end{align}
@@
 %"$\left\{$ event : $\left\{$ variable : index $\right\} \right\}$",
-The SSA-map is computed iteratively while traversing the event-flow graph in topological order as it is described in Algorithm~\ref{algorithm:ssa-map}:
+The SSA-map is computed iteratively while traversing the event-flow graph in topological order as it is described in Algorithm~\ref{algorithm:ssa-map}.
 
@@
     \begin{algorithmic}[1]
-        \Function{ComputeSSA}{G}
+        \Function{Compute-SSA-map}{G}
             \State {$S \gets$ empty map; $S[e_0] \gets$ empty map}
@@
 
+todo
+%Most bounded model-checking problems are solved via computing the fixpoint for the
+
 %The encoding of a weak memory model itself does not depend on the analysing program. The problem here...
+%recursive definitions and Kleene iteration
 
-recursive definitions and Kleene iteration
-
-As it is described in~\cite{Porthos17}, the is based . . .
+%r 1 ∪r 2 (e 1 , e 2 ) ⇔ r 1 (e 1 , e 2 ) ∨ r 2 (e 1 , e 2 )
+%r 1 ∩r 2 (e 1 , e 2 ) ⇔ r 1 (e 1 , e 2 ) ∧ r 2 (e 1 , e 2 )
+%r 1 \r 2 (e 1 , e 2 ) ⇔ r W
+%r −1 (e 1 , e 2 ) ⇔ r(e 2 , e 1 )
+%1 (e 1 , e 2 ) ∧ ¬r 2 (e 1 , e 2 )
+%r 1 (e 1 , e 3 ) ∧ r 2 (e 3 , e 2 )
+%r ∗ (e 1 , e 2 ) ⇔ r + (e 1 , e 2 ) ∨ (e 1 = e 2 )
+%r 1 ;r 2 (e 1 , e 2 ) ⇔
+%e 3 ∈E
+%r + (e 1 , e 2 ) ⇔ tc ⌈log |E|⌉ (e 1 , e 2 ), where
+%tc 0 (e 1 , e 2 ) ⇔ r(e 1 , e 2 ), and
+%tc i+1 (e 1 , e 2 ) ⇔ r(e 1 , e 2 ) ∨ (tc i (e 1 , e 2 ); tc i (e 1 , e 2 ))
+%
+%As it is described in~\cite{Porthos17}, the is based . . .
 %~\cite{Porthos17}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"

### 2018-04-26 4_input.tex
Change the illustration for the litmus-test example.
add minor fixes and multiple todo-comments.

@@
 
-The \Porthos tool used a small subset of C as an input language as described on Figure~\ref{syntax:in_grammar_pts}~\cite{Porthos17}:
+%TODO: say sth about 100500 litmus-tests for kernel
 
-\begin{figure}[H]
+The \Porthos tool used a small subset of C as an input language as described in Figure~\ref{syntax:in_grammar_pts}~\cite{Porthos17}.
+It supports recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
+In \mousquetaires, the set of supported instructions is extended in order to support expressions valid in C. %TODO: examples, how we extended
+Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment~`\texttt{:=}' or local variable assignment~`$\mathtt{\leftarrow}$') were defined directly in the grammar for \Porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
+% according to the set of known functions
+(see Chapter~\ref{???}).
+
+\begin{figure}%[H]
 %\begin{multicols}{2}
@@
         breaklines=true,
+        basicstyle=\ttfamily\scriptsize,
         ]
@@
 
-The input language of \Porthos supported recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
-In \mousquetaires, the set of supported instructions is extended in order to support expressions valid in C.%TODO: examples, which  we extended
- Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment~`\texttt{:=}' or local variable assignment~`$\mathtt{\leftarrow}$') were defined directly in the grammar for \Porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
-% according to the set of known functions
-(see Chapter~\ref{???}).
-
 Both \Porthos and \mousquetaires \ use the parser generator ANTLR~\cite{parr2013definitive}, the powerfull language processing tool.

### 2018-04-26 5_impl.tex
Cleanup a lot. Fix titles.
add minor fixes and multiple todo-comments.

@@
 
-This Chapter describes the architecture of the tool \mousquetaires.
+% function invocations/calls: inlining with binding
+% contexts: for now, no ctxs
+% comparing to cprover (https://www.cprover.org/cbmc/doc/manual.pdf, page 45, absatz about 'goto') where "the loop is unwound a given number of times" , we SHOULD have 2 modes of unrolling: 1) number of times, and 2) number of high-level instructions executed
+% TODO(code): (https://www.cprover.org/cbmc/doc/manual.pdf, page 45) "The break and continue statements are replaced by equivalent goto statements as described in the ANSI-C standard."
+
+% TODO: arrays: cprover manual page 48
+
+This Chapter describes the architecture of the \mousquetaires framework.
 
@@
 The main reason for starting the work on re-implementing the \Porthos tool was the need to extend input language to be able to process real C programs in perspective.
-The existing \Porthos architecture mixes high-level recursive \textit{instructions} (statements of C) with low-level non-recursive \textit{events}, to which the high-level instructions are compiled (see classes of package `\texttt{dartagnan.program}' in the \Porthos project~\footnote{Project web page: \url{http://github.com/hernanponcedeleon/PORTHOS}}).
+The existing \Porthos architecture mixes high-level recursive \textit{instructions} (statements of C) with low-level non-recursive \textit{events}, to which the high-level instructions are compiled (see classes of package `\texttt{dartagnan.program}' in the \Porthos project~\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}).
 Both instructions and events are then encoded into the SMT-formula.
@@
 \label{fig:loop-unwind}
-\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$} 
+\caption{Example of the flow graph from Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$}
 \end{figure}
@@
 
+% say something about equals() and hashCode()
+
 \section{Optimisations}

### 2018-04-26 1_intro.tex
Change the name of \mousquetaires.

@@
 
-The first tool that performs the WMM-aware portability analysis is \Porthos introduced in April~2017~\cite{Porthos17}.
+The first tool that performs the WMM-aware portability analysis is \porthos introduced in April~2017~\cite{Porthos17}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
@@
 
-Current work aims to rework the proof-of-concept tool \Porthos by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
+Current work aims to rework the proof-of-concept tool \porthos by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
 %interprocedural analysis
-We called the new tool the~\tool{mousquetaires}~framework as now it not only performs the portability analysis, but can serve as the basis for SMT-solver driven weak memory model analysis.
+As the general architecture and almost all components of \porthos have been redesigned, the tool received a new name -- \porthos[2].
+Considering the enhancements of the architecture, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.
 

### 2018-04-26 2_wmm.tex
Change the name of \mousquetaires.

@@
 	\item Events from the same process are related by the \textit{scope relation} \sr~$\subset~\mathbb{E}~\times~\mathbb{E}$.
-    In contrast to the \tool{herd} tool, the \mousquetaires does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple lables that indicate which process has produced certain event.
+    In contrast to the \tool{herd} tool, the \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple lables that indicate which process has produced certain event.
 \end{itemize}

### 2018-04-26 3_encod.tex
Change the name of \mousquetaires.

@@
 We show %TODO! to show, actually
-that the new encoding is smaller than the old one used in \Porthos since it does not produces new variables for each high-level statement of the input language.
+that the new encoding is smaller than the old one used in \porthos since it does not produces new variables for each high-level statement of the input language.
 %The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, \Porthos uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+For instance, \porthos uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
 $\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
@@
 The SSA form requires each variable to be assigned only once within entire program.
-In contrast, \Porthos used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
+In contrast, \porthos used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
 Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17}.
@@
 %linear time) the predecessors of each node in order to be able to convert the program into an SSA form, the time of such a transformation reduced %TODO: How much ????
-%comparing to the algorithm implemented in \Porthos, which recomputed SSA-indices recursively for each instruction. %TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
+%comparing to the algorithm implemented in \porthos, which recomputed SSA-indices recursively for each instruction. %TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
 
@@
 %The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
-The encoding of this linkage left untouched as it is implemented in \Porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
+The encoding of this linkage left untouched as it is implemented in \porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
 

### 2018-04-26 4_input.tex
Change the name of \mousquetaires.

--- 2018-04-12 4_input.tex
@@
 
-The \Porthos tool used a small subset of C as an input language as described on Figure~\ref{syntax:porthos}\cite{Porthos17}:
+%TODO: say sth about 100500 litmus-tests for kernel
 
-\begin{figure}[H]
-\begin{multicols}{2}
+The previous version of \porthos processed the small subset of C as an input language (see grammar in Figure~\ref{syntax:in_grammar_pts})~\cite{Porthos17}.
+It supported recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
+The grammar supported by \porthos[2] was extended in order to support expressions valid in C. %TODO: examples, how imenno extended
+Also, the input language for the first version of \porthos  
+Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment~`\texttt{:=}' or local variable assignment~`$\mathtt{\leftarrow}$') were defined directly in the grammar for \porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
+% according to the set of known functions
+(see Chapter~\ref{???}).
+
+\begin{figure}%[H]
+%\begin{multicols}{2}
 \begin{lstlisting}[mathescape=true,
-        caption={Syntax of an input language of \Porthos},
-        % captionpos=b,
-        label={syntax:porthos},
-        % basicstyle=\ttfamily\scriptsize,
-        % keywordstyle=\bfseries\color{blue},
-        %literate={<--}{$\leftarrow{}${}1},%{::=}{???},%:\!:\!=},
-        %morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}
-        %alsoletter={:,|,;},%
-        %morekeywords={:,|,;},%
-        %keywordstyle={\color{red}},% 
+        caption={Syntax of an input language of \porthos},
+        label={syntax:in_grammar_pts},
+        literate={<--}{{$\leftarrow$}}1,
+        morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}
         breaklines=true,
-        % xleftmargin=5mm,framexleftmargin=5mm
+        basicstyle=\ttfamily\scriptsize,
         ]
 <prog> : program <thrd>*
+       ;
 <thrd> : thread <tid> <inst>
+       ;
 <inst> : <atom> 
@@
        | if <pred > then <inst> else <inst>
+       ;
 <atom> : <reg> <-- <exp> 
@@
        | isync
+       ;
 \end{lstlisting}
-%\caption={Syntax of an input language of \Porthos}
-%\label={syntax:porthos}
-\end{multicols}
+%\end{multicols}
 \end{figure}
 
-// todo: old syntax more precisely
-
-The input language of \Porthos supported recursive definitions of control-flow instructions, which needed to be extended to support arbitrary jumps expressible in C.
-Also, for the purpose of simplicity, all kinds of fence instructions and different types of variable assignments (shared variable assignment `\texttt{:=}' or local variable assignment `$\mathtt{\rightarrow}$') were defined directly in the grammar for \Porthos, whereas the semantics of function invocations need to be determined dynamically during the interpretation
-% according to the set of known functions
-(see Chapter~\ref{???}).
-
-//todo: new syntax
+Both \porthos and \porthos[2] \ use the parser generator ANTLR~\cite{parr2013definitive}, the powerfull language processing tool.
+The (hand-written) ANTLR grammar used in \porthos is available at Appendix~\ref{apx:in_grammar_pts}, whereas the \porthos[2] uses the grammar of C language proposed in the C11 standard~\cite{jtc2011sc22} (the ANTLR grammar can be found in the official ANTLR repository on GitHub~\footnote{\url{https://github.com/antlr/grammars-v4}}).

### 2018-04-26 5_impl.tex
Change the name of \mousquetaires.

@@
-\chapter{\nohyphens{The~\mousquetaires:~implementation}}
+\chapter{\nohyphens{The~\porthos[2]:~implementation}}
 \label{ch:impl}
@@
 
-This Chapter describes the architecture of the \mousquetaires framework.
+This Chapter describes the architecture of the \porthos[2] framework.
 
-The programming language choice for \mousquetaires was made in favour of java in order to reuse some parts of its predecessor \Porthos that was written in java.
-However this language does not show best results in performance benchmarks (comparing to, for example, C++)~\cite{TODO}, the performance cornerstone of \mousquetaires (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \mousquetaires via java API.
-However, considering the perspective of using \mousquetaires as a static analyser for real-world programs, we also must to take into account memory problems at both encoding and solving stages.
+The programming language choice for \porthos[2] was made in favour of java in order to reuse some parts of its predecessor \porthos that was written in java.
+However this language does not show best results in performance benchmarks (comparing to, for example, C++)~\cite{TODO}, the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
+However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, we also must to take into account memory problems at both encoding and solving stages.
 
@@
 
-The main reason for starting the work on re-implementing the \Porthos tool was the need to extend input language to be able to process real C programs in perspective.
-The existing \Porthos architecture mixes high-level recursive \textit{instructions} (statements of C) with low-level non-recursive \textit{events}, to which the high-level instructions are compiled (see classes of package `\texttt{dartagnan.program}' in the \Porthos project~\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}).
+The main reason for starting the work on re-implementing the \porthos tool was the need to extend input language to be able to process real C programs in perspective.
+The existing \porthos architecture mixes high-level recursive \textit{instructions} (statements of C) with low-level non-recursive \textit{events}, to which the high-level instructions are compiled (see classes of package `\texttt{dartagnan.program}' in the \porthos project~\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}).
 Both instructions and events are then encoded into the SMT-formula.
@@
 
-Therefore, while planning the architecture of \mousquetaires, we decided to clearly separate high-level recursive code representation as Abstract Syntax Tree (AST) and low-level compiled event-based representation as a flow-graph.
-Such a modular architecture allows to use multiple input-language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \Porthos tool, two syntaxes of litmus tests used by \tool{herd}, and even assembly language for any supported architecture).
+Therefore, while planning the architecture of \porthos[2], we decided to clearly separate high-level recursive code representation as Abstract Syntax Tree (AST) and low-level compiled event-based representation as a flow-graph.
+Such a modular architecture allows to use multiple input-language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos tool, two syntaxes of litmus tests used by \tool{herd}, and even assembly language for any supported architecture).
 
@@
 
-Generally, \mousquetaires uses following  ...
+Generally, \porthos[2] uses following  ...
 

### 2018-04-26 1_intro.tex
Move input-language section into implementation. Add general scheme picture. Add something to the implementation section.

@@
 
-The first tool that performs the WMM-aware portability analysis is \porthos introduced in April~2017~\cite{Porthos17}.
+The first tool that performs the WMM-aware portability analysis is \porthos~\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} introduced in April~2017~\cite{Porthos17}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.

### 2018-04-26 3_encod.tex
Move input-language section into implementation. Add general scheme picture. Add something to the implementation section.

@@
 
-While working on the \tool{mousquetaires}, we applied some modifications of the encoding scheme for the control-flow.
+While working on the \porthos[2], we applied some modifications of the encoding scheme for the control-flow.
 The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of C language.
@@
 (here we used the notation of C-like ternary operator \texttt{x?y:z} for defining the conditional expression $\mathtt{if} x \mathtt{then} y \mathtt{else} z$).
-In contrast, the new scheme implemented in \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
+In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
 The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).

### 2018-04-26 4_impl.tex
Cleanup everything.
Rename; changes from bus.
Move input-language section into implementation. Add general scheme picture. Add something to the implementation section.

--- 2018-04-02 2_impl.tex
@@
-\chapter{Implementation}
+\chapter{The~\texttt{porthos2}:~implementation}
 \label{ch:impl}
 
-This chapter describes the architecture of the tool \texttt{mousquitaires} ...
+% function invocations/calls: inlining with binding
+% contexts: for now, no ctxs
+% comparing to cprover (https://www.cprover.org/cbmc/doc/manual.pdf, page 45, absatz about 'goto') where "the loop is unwound a given number of times" , we SHOULD have 2 modes of unrolling: 1) number of times, and 2) number of high-level instructions executed
+% TODO(code): (https://www.cprover.org/cbmc/doc/manual.pdf, page 45) "The break and continue statements are replaced by equivalent goto statements as described in the ANSI-C standard."
 
-language: java
+% TODO: arrays: cprover manual page 48
 
-\section{Program Requirements}
-\label{ch:impl:requirements}
+Current Chapter describes the architecture of the \porthos[2] framework.
 
-- stability (tests)
+The programming language choice for \porthos[2] was made in favour of java in order to be able to reuse some parts of its predecessor \porthos written in java.
+However this language does not show best results in performance benchmarks (comparing to C++, for example)~\cite{TODO}, the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
+However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
 
-- scalability (new features of language, new models, new tasks for a program )
+\section{Requirements}
+\label{ch:impl:req}
 
-- transparency
+The main reason for starting the work on re-implementing the \porthos tool was the need for extension the input language and optimisation of the tool so that it is able to process real C programs in perspective.
+In the existing \porthos architecture, the high-level recursive instructions (statements of C) are processed together with low-level non-recursive events abstractions (see classes of package `\texttt{dartagnan.program}' of \porthos) as single AST structure.
+This AST was implemented as a mutable data structure, which is being modified during the stage of SMT-encoding.
+We consider this architecture as hardly manageable (since it is hard to guarantee preserving of the invariants during the program execution) and poorly extensible (since adding support for a new high-level control-flow instruction requires changing multiple components of the program, from parser to encoder).
+%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).
 
-- efficiency
+Therefore, while designing the architecture of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
+Such a modular architecture allows to use multiple input language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
+
+The following list enumerates the target \textit{requirements} for the new tool:
+
+\begin{enumerate}[nolistsep]
+    \item \textit{stability and transparency}
+        \begin{itemize}
+            \item following the principles of simplicity and readability;
+            \item usage of software design patterns if necessary;
+            \item usage of immutable data structures for all data transfer objects~(DTO);
+            \item high code coverage by unit and functional tests;
+        \end{itemize}
+    \item \textit{efficiency}
+        \begin{itemize}[leftmargin=1em]
+            \item keeping the trade-off between execution time and memory usage;
+        \end{itemize}
+    \item \textit{extensibility}
+        \begin{itemize}[leftmargin=1em]
+            \item clear modular architecture
+        \end{itemize}
+\end{enumerate}
 
@@
 \label{ch:impl:comp}
-Big view
 
-\section{C11 to YTree parser}
-\label{ch:impl:ytree}
+The general architecture of \porthos[2] is presented in Figure~\ref{fig:program-components}.
+The program receives as input the program to be analysed (see Section~\ref{ch:impl:input}) and one (the reachability analysis mode) or two (the portability analysis mode) memory models (see Section~\ref{ch:impl:input-parser}).
+Then, the parsed program input (called Y-tree%
+\footnote{In order to avoid confusing different internal representations, we prefix the names of elements of each internal representation with a letter.
+For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
+) is being preprocessed in order to collect information necessary for performing compilation (see Section~\ref{ch:impl:y2x:preproc}), and compiled into the X-graph representation (see Section~\ref{ch:impl:y2x:compil}).
+
+\begin{figure}%[!b]%[H]
+  \centering
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
+  \caption{Main components of \porthos[2]}
+  \label{fig:program-components}
+\end{figure}
+
+\section{The input language}
+\label{ch:impl:input}
+
+%TODO: say sth about 100500 litmus-tests for kernel
+
+The previous version of \porthos processed the small subset of C as an input language (see grammar in Figure~\ref{syntax:in_grammar_pts})~\cite{Porthos17}.
+It supported recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
+The grammar supported by \porthos[2] was extended in order to support expressions valid in C. %TODO: examples, how imenno extended
+Also, the input language of the first version of \porthos syntactically distinguished the type of a memory instruction (shared variable assignments were defined with symbol~`\texttt{:=}', and local variable assignments --- with symbol~`$\mathtt{\leftarrow}$'), whereas \porthos[2] determines the semantics of a function invocation during the preprocessing stage of compilation~(see Chapter~\ref{TODO!!!}).
+
+\begin{figure}%[H]
+%\begin{multicols}{2}
+\begin{lstlisting}[mathescape=true,%
+caption={Syntax of an input language of \porthos version 1},%
+label={syntax:in_grammar_pts},%
+literate={<--}{{$\leftarrow$}}1,%
+morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}%
+breaklines=true,%
+basicstyle=\ttfamily\scriptsize]
+<prog> : program <thrd>*
+       ;
+<thrd> : thread <tid> <inst>
+       ;
+<inst> : <atom>
+       | <inst> ; <inst>
+       | while <pred> <inst>
+       | if <pred > then <inst> else <inst>
+       ;
+<atom> : <reg> <-- <exp>
+       | <reg> <-- <loc>
+       | <loc> := <reg>
+       | mfence
+       | sync
+       | lwsync
+       | isync
+       ;
+\end{lstlisting}
+%\end{multicols}
+\end{figure}
+
+Both \porthos and \porthos[2] \ use the parser generator ANTLR~\cite{parr2013definitive}, a powerful language processing tool.
+The full ANTLR grammar of input language used by previous version of \porthos is available at Appendix~\ref{apx:in_grammar_pts}.
+The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official ANTLR repository on GitHub~\footnote{The repository containing the collection of ANTLR v4 grammars:\\ \url{https://github.com/antlr/grammars-v4}}).
+Current version of \porthos[2] can operate only in the inter-procedural mode, assuming that each function defined in the input file is being executed in a separate thread.
+However, the redesigned architecture of \porthos[2] allow to easily support intra-procedural analysis by inlining function calls.
+
+\subsection{Parsing the input}
+\label{ch:impl:input-parser}
 
@@
 
-\section{YTree to XGraph event converter}
-\label{ch:impl:xtoy}
+\section{Compiling the Y-tree to the X-graph}
+\label{ch:impl:y2x}
 
+\subsection{Preprocessing}
+\label{ch:impl:y2x:preproc}
 - Then, the AST is being interpreted and converted to event-based representation (aliased 'Xrepr' for eXecution representation):
@@
 
+\subsection{Compilation}
+\label{ch:impl:y2x:compil}
+
 \subsection{Loop unrolling}
-\label{ch:impl:xtoy:unrolling}
+\label{ch:impl:x2y:unrolling}
 % If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph.
@@
 \label{fig:loop-unwind}
-\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$} 
+\caption{Example of the flow graph from Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$}
 \end{figure}
+
+\subsection{Input language parser}
+\label{ch:impl:model-parser}
 

### 2018-04-27 1_intro.tex
Fix grammars. Fix spaces after porthos name. Add some drafts to the impl section.

--- 2018-04-26 1_intro.tex
@@
 
-The first tool that performs the WMM-aware portability analysis is \porthos introduced in April~2017~\cite{Porthos17}.
+The first tool that performs the WMM-aware portability analysis is \porthos%
+\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}%
+introduced in April~2017~\cite{Porthos17}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
@@
 
-Current work aims to rework the proof-of-concept tool \porthos by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
+Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
 %interprocedural analysis
-As the general architecture and almost all components of \porthos have been redesigned, the tool received a new name -- \porthos[2].
+As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \porthos[2]%
+\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1 (also addressed as \porthos[1]), whereas the new version of \porthos{} is called \porthos[2].} %
+.
 Considering the enhancements of the architecture, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.

### 2018-04-27 3_encod.tex
Fix grammars. Fix spaces after porthos name. Add some drafts to the impl section.

@@
 We show %TODO! to show, actually
-that the new encoding is smaller than the old one used in \porthos since it does not produces new variables for each high-level statement of the input language.
+that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
 %The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, \porthos uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+For instance, \porthos{} uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
 $\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
@@
 The SSA form requires each variable to be assigned only once within entire program.
-In contrast, \porthos used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
+In contrast, \porthos{} used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
 Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17}.

### 2018-04-27 4_impl.tex
Fix grammars. Fix spaces after porthos name. Add some drafts to the impl section.

@@
 
-The programming language choice for \porthos[2] was made in favour of java in order to be able to reuse some parts of its predecessor \porthos written in java.
+The programming language choice for \porthos[2] was made in favour of java in order to be able to reuse some parts of its predecessor \porthos{} written in java.
 However this language does not show best results in performance benchmarks (comparing to C++, for example)~\cite{TODO}, the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
@@
 
-The main reason for starting the work on re-implementing the \porthos tool was the need for extension the input language and optimisation of the tool so that it is able to process real C programs in perspective.
-In the existing \porthos architecture, the high-level recursive instructions (statements of C) are processed together with low-level non-recursive events abstractions (see classes of package `\texttt{dartagnan.program}' of \porthos) as single AST structure.
+The main reason for starting the work on re-implementing the \porthos{} tool was the need for extension the input language and optimisation of the tool so that it is able to process real C programs in perspective.
+In the existing \porthos{} architecture, the high-level recursive instructions (statements of C) are processed together with low-level non-recursive events abstractions (see classes of package `\texttt{dartagnan.program}' of \porthos) as single AST structure.
 This AST was implemented as a mutable data structure, which is being modified during the stage of SMT-encoding.
@@
 Therefore, while designing the architecture of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
-Such a modular architecture allows to use multiple input language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
+Such a modular architecture allows to use multiple input language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
 
@@
 The general architecture of \porthos[2] is presented in Figure~\ref{fig:program-components}.
-The program receives as input the program to be analysed (see Section~\ref{ch:impl:input}) and one (the reachability analysis mode) or two (the portability analysis mode) memory models (see Section~\ref{ch:impl:input-parser}).
+The program receives as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models (see Section~\ref{ch:impl:input-parser}).
 Then, the parsed program input (called Y-tree%
@@
 For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
-) is being preprocessed in order to collect information necessary for performing compilation (see Section~\ref{ch:impl:y2x:preproc}), and compiled into the X-graph representation (see Section~\ref{ch:impl:y2x:compil}).
+) is being preprocessed in order to collect information necessary for the compilation (see Section~\ref{ch:impl:y2x:precomp}), and compiled into the X-graph representation (see Section~\ref{ch:impl:y2x:compil}).
+If the original program has a loop, the X-graph will have cycles, which need to be unrolled before being encoded into an SMT-formula.
+The unrolling and some other transformations are made by the X-graph transformer (see Section~\ref{}).
 
@@
 
-\section{The input language}
-\label{ch:impl:input}
+%\section{The input language}
+%\label{ch:impl:input}
+\subsection{Parsing the input}
+\label{ch:impl:input-parser}
+
+Both \porthos{} and \porthos[2] \ use the parser generator ANTLR~\cite{parr2013definitive}, a powerful language processing tool.
+The full ANTLR grammar of input language used by previous version of \porthos{} is available at Appendix~\ref{apx:in_grammar_pts}.
+
+% TODO: why we've chosen ANTLR (cite Kvanttt); what kind of parser is ANTLR, why it's good for C and not only for C (unified way to define multiple grammars, large collection, ...) + add references
+
+- Figure~\ref{syntax:in_grammar_pts} represents the informal grammar of the input language processed by \porthos[1] ... %todo: cite \cite{Porthos17}.
+- short characteristics of the old grammar, enumerate its features (wrt the Figure)
+
+- DRAWBACKS: %what parts need to be improved:
+
+    -- the semantics of memory operations and method invocations is encoded directly into the grammar.
+    Taking into account the complexity of developing the accurate grammar for a Turing-complete language,
+
+    -- in old implementation, the logics of parser was written directly in the grammar.
+    Although being fast-to-impelement (1 internal representation less), this is approach has numerous drawbacks s.a.:
+    * hard to read grammar (mixing 2 languages in a single file)
+    * hard to maintain grammar (no debugging with standard utils, no code analysis in the grammar file)
+    * non-extensible (there is no option to just plug-in existing grammar for any language and write a converter from this language's syntax tree to our AST)
+
+    - syntactically only the integers were supported. we:arrays???pointers??enums?structs?
+
+    - no declarations (all shared variables are declared in the init section, local variables are not decalred ever)
+
+    - no arbitrary function calls (need support for the knowledge base after implementing the typisation. cannot be supproted only syntactically)
+
+- Minor drawbacks:
+
+    - restricted sytnax for expressions:
+
+    -- no operator associativity allowed (`1 + 2 * 3` was invalid syntax)
+
+    -- no unary increment/decrement (technically it's hard to implement post-incerment/decrement without interpretation: we don't know when does this expression ends.)
+    Example:
+    `int* x;
+    if (x++ > 0 ) { }`
+    % todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
+
+    -- incorrectly implemented statement in old code: sequence of stmts is separated by ';', but the syntax of C requires ';' as the operator that ends each operation (how to say this? see C standard)
+
+    -- litmus-initialisation  stmt did not allow the non-default values. The new syntax allow any type of declaration statement with initialisation
+
+- this means the full compiler engine that resolves the semantics
+
+%The grammar supported by \porthos[2] was extended in order to support expressions valid in C. %TODO: examples, how imenno extended
+%Also, the input language of the first version of \porthos{} syntactically distinguished the type of a memory instruction (shared variable assignments were defined with symbol~`\texttt{:=}', and local variable assignments --- with symbol~`$\mathtt{\leftarrow}$'), whereas \porthos[2] determines the semantics of a function invocation during the preprocessing stage of compilation~(see Chapter~\ref{TODO!!!}).
+
+% todo: this is the simplified grammar
+
+%TODO: although the grammar is context-free and fail whith <angle brackets>.... it still works fine. Even gcc uses CF grammar? reference to INRIA's parser and paper on discussing grammars for C
 
 %TODO: say sth about 100500 litmus-tests for kernel
-
-The previous version of \porthos processed the small subset of C as an input language (see grammar in Figure~\ref{syntax:in_grammar_pts})~\cite{Porthos17}.
-It supported recursive definitions of control-flow instructions, that were directly encoded into SMT-formula.
-The grammar supported by \porthos[2] was extended in order to support expressions valid in C. %TODO: examples, how imenno extended
-Also, the input language of the first version of \porthos syntactically distinguished the type of a memory instruction (shared variable assignments were defined with symbol~`\texttt{:=}', and local variable assignments --- with symbol~`$\mathtt{\leftarrow}$'), whereas \porthos[2] determines the semantics of a function invocation during the preprocessing stage of compilation~(see Chapter~\ref{TODO!!!}).
 
@@
 \begin{lstlisting}[mathescape=true,%
-caption={Syntax of an input language of \porthos version 1},%
+                  caption={Syntax of an input language of \porthos{} version 1},%
                   label={syntax:in_grammar_pts},%
-literate={<--}{{$\leftarrow$}}1,%
-morekeywords={if,then,else,return,while,program,thread,mfence,sync,lwsync,isync}%
+                  %literate={<--}{{$\leftarrow$}}1,%
+                  morekeywords={if,then,else,return,while,program,thread}%
                   breaklines=true,%
                   basicstyle=\ttfamily\scriptsize]
-<prog> : program <thrd>*
+<prog> : <init> <thrd>* <assert>
        ;
@@
        | while <pred> <inst>
-       | if <pred > then <inst> else <inst>
+       | if <pred> { <inst> } <inst>
        ;
-<atom> : <reg> <-- <exp>
+<atom> : <reg> <-- <expr>
        | <reg> <-- <loc>
        | <loc> := <reg>
-       | mfence
-       | sync
-       | lwsync
-       | isync
+       | 'mfence'
+       | 'sync'
+       | 'lwsync'
+       | 'isync'
+       ;
+<pred> :
+       | true
+       | false
+       | <expr> (and | or) <expr>
+       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
+       ;
+<expr> : [0-9]
+       | <reg>
+       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
        ;
@@
 
-Both \porthos and \porthos[2] \ use the parser generator ANTLR~\cite{parr2013definitive}, a powerful language processing tool.
-The full ANTLR grammar of input language used by previous version of \porthos is available at Appendix~\ref{apx:in_grammar_pts}.
+- NOW: we're using the C grammar from the standard
+
+- for now, we just ignore C macro directives, in future it's planned to support partially the preprocessing before parsing
+
 The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official ANTLR repository on GitHub~\footnote{The repository containing the collection of ANTLR v4 grammars:\\ \url{https://github.com/antlr/grammars-v4}}).
@@
 
-\subsection{Parsing the input}
-\label{ch:impl:input-parser}
+%TODO: say sth about syntax errors handling: still now, what if parser does not accept syntax?
 
-below: mostly mock text.
+- if we don't support , our parser still parses it, and the error is thrown at the moment of converting syntax tree to the AST (Y-tree).
 
-- The language-dependent syntax tree:
-        - for now it's the C subset language which I called 'Cmin'; as a base, I used the C11 grammar from ANTLR github repository, then I simplified it a lot, cutting off many unnecessary C syntax features and making it more convenient for parsing.
-When developing the Cmin language, I kept in mind C elements that are necessary for processing the linux kernel code, though for now not the whole grammar element described in file 'Cmin.g4' are being implemented;
-        - later I am going to add the litmus grammar as well;
-        - in future, it will be not a problem to add any new C-like language;
+- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11->Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
+need ly?)
 
-- The language-independent abstract syntax tree (aliased 'Ytree', where 'Y' resembles branching of the tree):
-        - all tree nodes in my code are prefixed with 'Y', see tentative (yet almost complete) class hierarchy in picture 'YEntity.png';
-        - this AST contains very basic language elements according to the C execution model (statements and expressions);
-        - converting the language-dependent syntax tree to the language-independent syntax tree is performed by Visitor pattern (e.g., for Cmin->Ytree conversion is made by 'CminToYtreeConverterVisitor')
-        - minor changes are performed by converting to ytree representation: desugaring the target code, etc.
+\section{The Y-tree: an AST}
+\label{ch:impl:ytree}
+
+- picture of the Y-hierarchy. everything inherits interface YEntity
+
+- immutable
+
+- AST is untyped (YVariableRef).
+
+- short characteristics with citations from the code (this AST contains very basic language elements according to the C execution model (statements and expressions) )
+
+- minor changes are performed by converting to ytree representation: desugaring the target code, etc. (what else?)
+
+\section{Model parser}
+\label{ch:impl:model}
+
+TODO
 
@@
 
-\subsection{Preprocessing}
-\label{ch:impl:y2x:preproc}
-- Then, the AST is being interpreted and converted to event-based representation (aliased 'Xrepr' for eXecution representation):
-        - more low-level code representation (or high-level assembly);
-        - I try to keep this representation close to the one you described in your papers: basic load \& store events, branching events, fence events;
-        - this representation is being implementing these days, I've just started doing it (see current class hierarchy in the picture 'XEntity.png');
+%todo: some words on necessity of contexts and lack of them. How would we implement them
 
-- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them):
-        - converting to SSA form as one of necessary steps before encoding;
-        - (more? -- I'm not thinking about it yet);
+%todo: rename Interpreter -> Compiler.
+
+%note https://en.wikipedia.org/wiki/Semantics_(computer_science)
+%Operational semantics loosely corresponds to interpretation,
+
+- hierarchy of Compilers (XCompiler is an stateful abstract machine)
+
+- interface that it provides
+
+- dependencies on other modules (memory-manager, etc.)
+
+\subsection{Pre-compilation}
+\label{ch:impl:y2x:precomp}
+
+- collect goto labels (not done yet.)
+
+- determine kind of variables (cannot be done during parsing? don't say this)
+
+- basic typisation
+
+%todo: after all, list which modules (classes, managers) are being initialised during the pre-compilation. I'ts good to have the dependencies graph on moduls (operational semantics:) )
 
@@
 
-\subsection{Loop unrolling}
-\label{ch:impl:x2y:unrolling}
+- more low-level code representation (or high-level assembly); abstract assembly language. refer to the %2_wmm.tex
+
+- X-hierarchie
+
+\subsection{Post-compilation transformations}
+\label{ch:impl:x2y:postcomp}
+
+- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
+
+- converting to SSA form (now: during the encoding. should be: during the post-compilation) (as one of necessary steps before encoding)
+
+- setting up backward edges
+
+- more?
+
+- unrolling: why we cannot encode cyclic structures. reference to the paper (see arXiv version)
 % If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph.
 Although, in order to be able to encode such a cyclic structure as a boolean formula, this graph needs to be
-% In order to be
 
@@
 
-\subsection{Input language parser}
-\label{ch:impl:model-parser}
-
 \section{XGraph to ZFormula (SMT) encoder}

### 2018-04-28 2_wmm.tex
add some todos. add XEntity diagram.

@@
-\chapter{Memory model-aware analysis}
+\chapter{Memory model-aware analysis}%Description of the analysis of concurrent systems wrt weak memory models
 %\chapter{A generic framework}
@@
 Firstly, the control-flow and data-flow of a program is encoded as the set of possible \textit{candidate executions}. %possible communications between threads of the program.
-Obtained model of the program is called an anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
+Obtained model of the program describes the anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
 Thereafter, the anarchic semantics is constrained by the \textit{weak memory model} specification which is a set of axiomatic constraints for filtering out executions inconsistent in particular architecture.
@@
 %One way to do this is to use a non-deterministic computation-centric model~\cite{fri97}, such as ...%The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
+%TODO: nondeterministic
 One such model is the \textit{event-based} representation of a program, which models a program as a directed graph of events (the \textit{event-flow graph}).
@@
 \label{ch:wmm:model:events}
+
+%TODO: event is atomic?
 
@@
 
+%TODO: this is binary relations. ADD UNARY RELS!
+
 The derived relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events

### 2018-04-28 inc/XEventVisitor-cleaned.java

@@ -1,0 +1,28 @@
+public interface XEventVisitor<T> {
+
+    T visit(XEntryEvent event);
+
+    T visit(XExitEvent event);
+
+    T visit(XUnaryComputationEvent event);
+
+    T visit(XBinaryComputationEvent event);
+
+    T visit(XAssertionEvent event);
+
+    T visit(XInitialWriteEvent event);
+
+    T visit(XRegisterMemoryEvent event);
+
+    T visit(XStoreMemoryEvent event);
+
+    T visit(XLoadMemoryEvent event);
+
+    T visit(XMethodCallEvent event);
+    
+    T visit(XJumpEvent event);
+
+    T visit(XNopEvent event);
+
+    T visit(XBarrierEvent event);
+}

### 2018-04-28 4_impl.tex
Set up the new structure of the impl chapter. Rewrite the beginning.

@@
 \label{ch:impl}
+
+The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires extension the input language.
+%process more intricate entities of the C language. 
+%Extension of the input language imply not only ability to process syntactic sugar
+This implies support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+Although far not the entire C language is supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+\footnote{To ensure that, we have merely to look at existing C compilers, for instance, gcc, that uses the hand-written C parser in more than 18.5 thousand lines (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+), we consider the accomplished work as a step towards this.
+%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
+%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
+% Although and optimisation of the tool so that it is able to process real C programs in perspective.
+
+%In current Chapter we discuss the new architecture of \porthos[2].
 
@@
 
-Current Chapter describes the architecture of the \porthos[2] framework.
+%\section{The architecture of \porthos[1]}
+\section{General principles}
+\label{ch:impl:principles}
+
+The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
+Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
+We are inclined to consider this architecture as one that is fast to develop, but hard to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
+%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).
+
+% todo: in old porthos the strategy soft: print info about error. new: exception on any detected violation of any invariant
+
+Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
+Such a modular architecture allows to support multiple input languages by parsing them and converting parsed syntax trees to a simplified AST.
+which, along with all other data-transfer objects (DTO), must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
+The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
+\footnote{Project repository: \url{https://github.com/google/guava/}}%
+).
+
+During the development of \porthos[2] we mainly followed the \textit{KISS principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
+The following list summarises the main rules we followed during the development of \porthos[2]:
+
+%\vspace{1em}
+\begin{enumerate}[nolistsep]
+  \item \textit{robustness:}
+    \begin{enumerate}[label*=\arabic*.]
+      \item usage of immutable data structures for all DTOs;
+      \item interrupting the work on errors;
+      \item modular architecture: each module can be tested independently;
+      \item usage of software design patterns if necessary;
+    \end{enumerate}
+  \item \textit{transparency:}
+    \begin{enumerate}[label*=\arabic*.]
+      \item following the principles of simplicity and readability;
+      \item clear and informative program output;
+    \end{enumerate}
+  \item \textit{efficiency:}
+    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
+      \item keeping the trade-off between execution time and memory usage;
+    \end{enumerate}
+  \item \textit{extensibility:}
+    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
+      \item clear modular architecture
+    \end{enumerate}
+\end{enumerate}
+
+\section{Architecture} %Program Components
+\label{ch:impl:arch}
 
@@
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
-
-\section{Requirements}
-\label{ch:impl:req}
-
-The main reason for starting the work on re-implementing the \porthos{} tool was the need for extension the input language and optimisation of the tool so that it is able to process real C programs in perspective.
-In the existing \porthos{} architecture, the high-level recursive instructions (statements of C) are processed together with low-level non-recursive events abstractions (see classes of package `\texttt{dartagnan.program}' of \porthos) as single AST structure.
-This AST was implemented as a mutable data structure, which is being modified during the stage of SMT-encoding.
-We consider this architecture as hardly manageable (since it is hard to guarantee preserving of the invariants during the program execution) and poorly extensible (since adding support for a new high-level control-flow instruction requires changing multiple components of the program, from parser to encoder).
-%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).
-
-Therefore, while designing the architecture of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
-Such a modular architecture allows to use multiple input language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
-
-The following list enumerates the target \textit{requirements} for the new tool:
-
-\begin{enumerate}[nolistsep]
-    \item \textit{stability and transparency}
-        \begin{itemize}
-            \item following the principles of simplicity and readability;
-            \item usage of software design patterns if necessary;
-            \item usage of immutable data structures for all data transfer objects~(DTO);
-            \item high code coverage by unit and functional tests;
-        \end{itemize}
-    \item \textit{efficiency}
-        \begin{itemize}[leftmargin=1em]
-            \item keeping the trade-off between execution time and memory usage;
-        \end{itemize}
-    \item \textit{extensibility}
-        \begin{itemize}[leftmargin=1em]
-            \item clear modular architecture
-        \end{itemize}
-\end{enumerate}
-
-\section{Program Components}
-\label{ch:impl:comp}
 
@@
   \centering
-  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
+  % TODO: uncomment after a while
+  %\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
   \caption{Main components of \porthos[2]}
@@
 
-%\section{The input language}
-%\label{ch:impl:input}
-\subsection{Parsing the input}
-\label{ch:impl:input-parser}
+\subsection{The program input}
+
+%(for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
+
+\subsection{The Model}
+
+blabla
+
+\subsubsection{Y-tree}
+
+\subsubsection{X-graph}
+
+\subsubsection{W-model}
+
+\subsubsection{Z-formula}
+
+\subsection{The processing units} %Transformation}
+
+\subsubsection{Y-tree parser}
+
+\subsubsection{W-model parser}
+
+\subsubsection{X-graph compiler}
+
+\paragraph{Pre-compiler}
+\paragraph{Compiler}
+\paragraph{Post-compiler}
+
+\subsubsection{The Z-formula encoder}
+
+\subsection{The program output}
+
+%\subsection{Parsing the input}
+%\label{ch:impl:input-parser}
 
@@
     - syntactically only the integers were supported. we:arrays???pointers??enums?structs?
-
     - no declarations (all shared variables are declared in the init section, local variables are not decalred ever)
@@
 
-\section{The Y-tree: an AST}
-\label{ch:impl:ytree}
+%\section{The Y-tree: an AST}
+%\label{ch:impl:ytree}
 
@@
 
-\section{Model parser}
-\label{ch:impl:model}
+%\section{Model parser}
+%\label{ch:impl:model}
 
@@
 \label{ch:impl:y2x}
+
+\begin{figure}%[!b]%[H]
+  \centering
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XEntity.png}
+  \caption{TODO}
+  \label{fig:class-diagrams:XEntity}
+\end{figure}
+
+\lstinputlisting{inc/XEventVisitor-cleaned.java}
 
@@
 
-\subsection{Compilation}
-\label{ch:impl:y2x:compil}
+%\subsection{Compilation}
+%\label{ch:impl:y2x:compil}
 
@@
 
-\subsection{Post-compilation transformations}
-\label{ch:impl:x2y:postcomp}
+%\subsection{Post-compilation}
+%\label{ch:impl:x2y:postcomp}
+
+transformations
 
@@
 
-\section{XGraph to ZFormula (SMT) encoder}
-\label{ch:impl:comp:zformula}
+%\section{XGraph to ZFormula (SMT) encoder}
+%\label{ch:impl:comp:zformula}
 
@@
 
-\section{Optimisations}
-
-... performed on each stage
+%TODO: \section{Optimisations}

### 2018-04-29 2_wmm.tex
cleanup the wmm section a bit.

@@
 
-The relation~$\xrightarrow{r}\,:~\mathbb{E}~\times~\mathbb{E}$ is a binary function over events (set of pairs of events).
+The relation~%\xrightarrow{r}
+$r\,\subseteq~\mathbb{E}~\times~\mathbb{E}$ is a set of pairs of events (a subset of Cartesian product of two sets of events).
 There are two kinds of relations between events: \textit{basic relations} %retrieved from the
@@
 
-    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to writes preceding the write event from which the read reads the value: \\
-    $r \relation{fr} w = (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$.
+    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to write events preceding the write event from which the read event gets the value: \\
+    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$
 
-    \item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on).
+    %\item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on)
+    %TO-DO: formula for dp or exclude
+
+    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
+    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
+    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$
+
+    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes: \\
+    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(w))$ \\
+    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(w))$
+
     %TODO: say that po-relation is *static*
-    %$\mathtt{com} = (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
+    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
+    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$
 
-    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes.
-
-    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable:
-    $m_1 \relation{po-loc} m_2 = (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$.
-
-    \item the semantics of fences and barriers specific for different architectures may be defined as derived relations.
+    \item the semantics of \textit{fences} (memory barriers) specific for different architectures may be defined as derived relations.
 
@@
 
-The work~\cite[Chapter~2]{alglave2010shared} provides definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}.
-Thereafter, weak memory models make asserts over these properties, thus restricting the set of allowed behaviours of the system.
-
-% TODO: some derived relations
-%The \textit{from-read order} \fr $\subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
-%TODO: perephrase last sentence! alglave thesis, p. 36
+The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.
 
@@
 
-The derived relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
+New relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
 

### 2018-04-29 4_impl.tex
Cleanup the ch:impl.

@@
 
-The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires extension the input language.
-%process more intricate entities of the C language. 
-%Extension of the input language imply not only ability to process syntactic sugar
-This implies support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Although far not the entire C language is supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
-\footnote{To ensure that, we have merely to look at existing C compilers, for instance, gcc, that uses the hand-written C parser in more than 18.5 thousand lines (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
+This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+Although far not the entire C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+\footnote{To ensure that, we have merely to look at existing C compilers, for instance, the open-source \texttt{gcc} compiler, that uses a C parser written in more than 18.5 thousand lines (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
 ), we consider the accomplished work as a step towards this.
@@
 
-The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
+The existing implementation of \porthos[1] does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
 
-During the development of \porthos[2] we mainly followed the \textit{KISS principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
+During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
 The following list summarises the main rules we followed during the development of \porthos[2]:
 
-%\vspace{1em}
+\vspace{0.5em}
 \begin{enumerate}[nolistsep]
-  \item \textit{robustness:}
+  \item \textit{Robustness:}
     \begin{enumerate}[label*=\arabic*.]
@@
     \end{enumerate}
-  \item \textit{transparency:}
+  \item \textit{Transparency:}
     \begin{enumerate}[label*=\arabic*.]
@@
     \end{enumerate}
-  \item \textit{efficiency:}
+  \item \textit{Efficiency:}
     \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
@@
     \end{enumerate}
-  \item \textit{extensibility:}
+  \item \textit{Extensibility:}
     \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
-      \item clear modular architecture
+      \item clear modular architecture.
     \end{enumerate}
@@
 
-\section{Architecture} %Program Components
+The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos[1] that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
+Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
+However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
+It is worth noting that, for the reasons of simplicity, the \porthos[2] is not a concurrent program, however, we believe that, due to its modular architecture, it can be easily parallelised on the level of program modules.
+
+\section{Architecture}
 \label{ch:impl:arch}
 
-The programming language choice for \porthos[2] was made in favour of java in order to be able to reuse some parts of its predecessor \porthos{} written in java.
-However this language does not show best results in performance benchmarks (comparing to C++, for example)~\cite{TODO}, the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
-However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
+The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
+The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
+The program input is being parsed (Section~\ref{ch:impl:proc:y-parser}) to a program AST called \texttt{Y-tree}%
+\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
+For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
+(Section~\ref{ch:impl:model:ytree}), and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
 
-The general architecture of \porthos[2] is presented in Figure~\ref{fig:program-components}.
-The program receives as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models (see Section~\ref{ch:impl:input-parser}).
-Then, the parsed program input (called Y-tree%
-\footnote{In order to avoid confusing different internal representations, we prefix the names of elements of each internal representation with a letter.
-For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
-) is being preprocessed in order to collect information necessary for the compilation (see Section~\ref{ch:impl:y2x:precomp}), and compiled into the X-graph representation (see Section~\ref{ch:impl:y2x:compil}).
-If the original program has a loop, the X-graph will have cycles, which need to be unrolled before being encoded into an SMT-formula.
-The unrolling and some other transformations are made by the X-graph transformer (see Section~\ref{}).
+Thereafter, the \texttt{Y-tree} is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-compiler:pre}) in order to collect information necessary for the compilation.
+Then, the \texttt{Y-tree} is being compiled (Section~\ref{ch:impl:proc:x-compiler:comp}) to an \texttt{X-graph} representation (Section~\ref{ch:impl:model:xgraph}), which then undergoes a number of transformations necessary for the encoding to an SMT formula at the post-compilation stage (Section~\ref{ch:impl:proc:x-compiler:post}).
+The post-processed \texttt{X-graph} then "fills in" the memory-model \texttt{W-model}, and both are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \texttt{Z-formula}, which then proceeds to the Z3 solver as an input.
 
@@
   % TODO: uncomment after a while
-  %\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
   \caption{Main components of \porthos[2]}
-  \label{fig:program-components}
+  \label{fig:arch}
 \end{figure}
@@
 \subsection{The program input}
-
+\label{ch:impl:input}
 %(for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
@@
 \subsection{The Model}
-
+\label{ch:impl:model}
 blabla
@@
 \subsubsection{Y-tree}
+\label{ch:impl:model:ytree}
 
 \subsubsection{X-graph}
+\label{ch:impl:model:xgraph}
 
 \subsubsection{W-model}
+\label{ch:impl:model:wmodel}
 
 \subsubsection{Z-formula}
+\label{ch:impl:model:zformula}
 
 \subsection{The processing units} %Transformation}
+\label{ch:impl:proc}
 
 \subsubsection{Y-tree parser}
+\label{ch:impl:proc:y-parser}
 
 \subsubsection{W-model parser}
+\label{ch:impl:proc:w-parser}
 
 \subsubsection{X-graph compiler}
+\label{ch:impl:proc:x-compiler}
 
 \paragraph{Pre-compiler}
+\label{ch:impl:proc:x-compiler:pre}
+
 \paragraph{Compiler}
+\label{ch:impl:proc:x-compiler:comp}
+
 \paragraph{Post-compiler}
+\label{ch:impl:proc:x-compiler:post}
 
 \subsubsection{The Z-formula encoder}
+\label{ch:impl:proc:z-encoder}
 
 \subsection{The program output}
+\label{ch:impl:out}
 

### 2018-04-29 2_wmm.tex
Add input-language section. cleanup a lot.

@@
 An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (which is modelled as a register as well) to the register $r_1$.
-For modelling branching statements, we distinguish the set $\mathbb{C_{p}} \subseteq \mathbb{C}$ of \textit{predicative} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
+For modelling branching statements, we distinguish the set $\mathbb{C_{p}}~\subseteq~\mathbb{C}$ of \textit{predicative} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
 
@@
 
-The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.
+%todo: some words here?
 
@@
 
-The \textit{candidate execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
-%the sequence of choices from guesses which event is to be executed next.
-%The semantics of a concurrent program is represented as the set of allowed executions.
+The semantics of a concurrent program is represented as the set of allowed executions. The 
+The \textit{execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
+It can be interpreted as a sequence of guesses which event is to be executed next.
+The \textit{candidate execution} is an execution that is not yet constrained by a memory model.
+
+%TODO: cite this paper elsewhere
+%As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
+
 %An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
-%A \textit{candidate execution} is an execution that is not constrained by the memory model (i.e., the execution for that the validity is not checked).
-%As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
 %The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
 %associated with the instructions of the program
-%the path in the event-flow graph.
-%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
+%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
 %An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
@@
 
-\section{The \cat language}
+\section{The \cat{} language}
 \label{ch:wmm:cat}
 
-Weak memory models are defined via the \cat language~\cite{alglave2016syntax}.
+Weak memory models are defined via the \cat{} language~\cite{alglave2016syntax}.
 This is a domain specific language for describing consistency properties of concurrent programs.
-The \cat language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, expressions and assertions that are specific for operating with relations and executions.
+The \cat{} language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, operations and assertions that are specific for operating with relations and executions.
 %represents the functional language extended by the theory of  allows to axiomatically define derived relations, architecture-specific fences, and constraints over relations.
+In \cat{}, new relations can be defined via the keyword \texttt{let} and the following operators over relations~\cite{alglave2016syntax}. % 3.1.6 Relations between events
 
@@
 
-New relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
+Below we enumerate pre-defined operators over relations and sets of events:
 
+\begin{enumerate}
+  \item \textit{Unary relations}:
+    \begin{itemize}
+      \item \textit{the complement} of a relation \rel{r} is $\rel{\textasciitilde r}$
+      \item \textit{the transitive closure} of a relation \rel{r} is $\rel{r+}$
+      \item \textit{the reflexive closure} of a relation \rel{r} is $\rel{r?}$
+      \item \textit{the reflexive-transitive closure} of a relation \rel{r} is $\rel{r*}$
+      \item \textit{the inverse} of a relation \rel{r} is $\invrel{r}$
+    \end{itemize}
+  \item \textit{Binary relations:}
   \begin{itemize}
@@
   \end{itemize}
+\end{enumerate}
 
-For instance, the \fr-relation is defined as follows: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$.
+For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$.
+Figure~\ref{example:x86-cat} contains an example of x86-TSO weak memory model definition in \cat{} language that asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}:
 
-Figure~\ref{example:x86-cat} contains part of x86-TSO model~\cite{owens2009better} that asserts acyclicity of communication relation and \rel{po-loc} relation:
+\begin{figure}[h]
+\begin{lstlisting}
+"X86 TSO"
+include "x86fences.cat"
+include "filters.cat"
+include "cos.cat"
 
-\begin{figure}[!h]
-\begin{lstlisting}
-...
+(* Uniproc check *)
 let com = rf | fr | co
-let po-loc = po & loc
 acyclic po-loc | com
+
+(* Atomic *)
+empty rmw & (fre;coe)
+
+(* Global happens-before *)
+#ppo
+let po_ghb = WW(po) | RM(po)
+
+#mfence
+include "x86fences.cat"
+
+#implied barriers
+let poWR = WR(po)
+let i1 = MA(poWR)
+let i2 = AM(poWR)
+let implied = i1 | i2
+
+let com = rfe | fr | co
+let ghb = mfence | implied | po_ghb | com
+show implied
+acyclic ghb as tso
 \end{lstlisting}
-\caption{Excerpt from the x86-TSO memory model in \cat language}
+\caption{The x86-TSO memory model defined in \cat{} language\protect\footnotemark}
 \label{example:x86-cat}
 \end{figure}
+\footnotetext{The \cat{}-memory model for x86-TSO can be found in official repository of \tool{herd} tool: \url{https://github.com/herd/herdtools7}}
+%\begin{figure}[h]
+%\begin{lstlisting}
+%...
+%let com = rf | fr | co
+%let po-loc = po & loc
+%acyclic po-loc | com
+%\end{lstlisting}
+%\caption{Excerpt from the x86-TSO memory model in \cat{} language}
+%\label{example:x86-cat}
+%\end{figure}
 
 %Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}
+
+%The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.

### 2018-04-29 3_encod.tex
add some todos. add XEntity diagram.
Add input-language section. cleanup a lot.

@@
 %\chapter{Portability of concurrent software}
-\chapter{Portability analysis as an SMT~problem}
+\chapter{Portability analysis as an SMT~problem}%Description of the method used by Porthos
 \label{ch:port}
+
+%TODO: the method used by Porthos is complete and efficient, but the implementation sucks. Why? 1), 2), 3). How can it be fixed? 1), 2), 3).
+
+%TODO: more about SMT (= SAT + constraints expressed in background theories. Exapmles of theories that we use and that are implem-ted in Z3).
 
@@
 
+%TODO: overapproximation. loops, Rice's thm. Within the bound, the method is complete.
+
 In general, a BMC problem aims to examine the reachability of the "undesirable" states of a finite-state system.
@@
 %\begin{equation}
-$\mathtt{SAT}( Init(\vec{x_0}) \land Trans(\vec{x}_0, \vec{x}_1) \land \cdots \land Trans(\vec{x}_{k-1}, \vec{x}_k) \land Bad(\vec{x}_k) )$.
+$\mathtt{SAT}( Init(\vec{x_0}) \land Trans(\vec{x}_0, \vec{x}_1) \land ... \land Trans(\vec{x}_{k-1}, \vec{x}_k) \land Bad(\vec{x}_k) )$.
 %\end{equation}
@@
 Portability analysis problem may also be stated as a reachability problem, where the undesirable state is the state reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
-However, unlikely the BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:x2y:unrolling}) and the $Trans$ predicate may be stated only for the final state of a program.
+However, unlikely the BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
 
@@
 Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.
-The result SMT formula $\phi$ of the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$.
+The result SMT formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
 If the formula is satisfiable, there exist a portability bug.
@@
 
-The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:x2y:unrolling}).%TODO: footnote why multiple sinks
+The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:proc:x-transf}).%TODO: footnote why multiple sinks
 In control-flow graph, there are two types of transitions (edges): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
@@
 For that, we compile the recursive abstract syntax tree~(AST) of the parsed C-code to the plain (non-recursive) event-flow graph.
-We show %TODO! to show, actually
+We show %TODO! to show, actually!!!
 that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
@@
 $\phi_{CF}( \, c \, ? \, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
-(here we used the notation of C-like ternary operator \texttt{x?y:z} for defining the conditional expression $\mathtt{if} x \mathtt{then} y \mathtt{else} z$).
+(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\,x\,\mathtt{then}\,y\,\mathtt{else}\,z$').
 In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
@@
 
-Let $\fx : \mathbb{E} \rightarrow \{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
-Let $\fv : \mathbb{C} \rightarrow \mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \relation{rf}-relation; see Chapter~\textbf{?TODO?} for the relations encoding%TODO
-).
-We distinguish the function $\fv_p : \mathbb{C_p} \rightarrow \{0,1\}$ that evaluates the predicative computation event.
+Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
+% TODO: attention to these non-breakable spaces
+Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \rf-relation (see Section~\ref{ch:wmm:model:relations} for details on relations).
+We distinguish the function ${\fv_p\,:\,\mathbb{C_p}\,\rightarrow\,\{0,1\}}$ that evaluates the predicative computation event.
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
@@
 Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
-Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, (e_1), (e_1~\rightarrow~e_2), (e_1 ~\rightarrow~e_3)\}$.
+Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \, (e_1), \, (e_1~\rightarrow~e_2), \, (e_1 ~\rightarrow~e_3)\}$.
 In encoding~\ref{enc:cf:merge} of the merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
@@
 \begin{align}
-\phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
-				  &  [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
+\phi_{CF_{seq}}   = \ & \ \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
+\phi_{CF_{br}}    = \ & \ [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
+                      & \ [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
                       & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
@@
         \node[c] at (6,-1.5) (2b) {$e_2$};
-        \node[c,fill=black!20] at (7.5,-1.5) (nop) {$e_{\mbox{\tiny NOP}}$}; %{\tiny{NOP}};
+        \node[c,fill=black!15] at (7.5,-1.5) (nop) {$e_{\mbox{\tiny NOP}}$};
         \node[c] at (6,-3) (3b) {$e_3$};

### 2018-04-29 4_impl.tex
Add input-language section. cleanup a lot.

@@
 The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
-This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
 Although far not the entire C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
@@
 
+As \porthos[1], the \porthos[2] uses the open-source SMT solver \texttt{Z3} from Microsoft Research~\ref{de2008z3}.
+However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \texttt{Z-formula} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
+%TODO: *should* allow?
+
 The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos[1] that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
-Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver called from \porthos[2] via java API.
+Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver \textit{Z3}%
+\footnote{The Z3 project repository: \url{https://github.com/Z3Prover/z3}} %
+invoked from \porthos[2] via java API.
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
@@
 
-\section{Architecture}
+\section{General architecture}
 \label{ch:impl:arch}
@@
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
-The program input is being parsed (Section~\ref{ch:impl:proc:y-parser}) to a program AST called \texttt{Y-tree}%
+The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \texttt{Y-tree}%
 \footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
-For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
-(Section~\ref{ch:impl:model:ytree}), and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
+For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.} %
+(Section~\ref{ch:impl:model:ytree})
+%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
+, which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
+The \texttt{Y-tree} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \texttt{X-graph} representation (Section~\ref{ch:impl:model:xgraph}).
+The compiled \texttt{X-graph} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \texttt{Z-formula} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
+Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \texttt{X-graph} in order to build the weak memory model \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
+Thereafter, \texttt{W-model} and transformed \texttt{X-graph} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \texttt{Z-formula} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
 
-Thereafter, the \texttt{Y-tree} is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-compiler:pre}) in order to collect information necessary for the compilation.
-Then, the \texttt{Y-tree} is being compiled (Section~\ref{ch:impl:proc:x-compiler:comp}) to an \texttt{X-graph} representation (Section~\ref{ch:impl:model:xgraph}), which then undergoes a number of transformations necessary for the encoding to an SMT formula at the post-compilation stage (Section~\ref{ch:impl:proc:x-compiler:post}).
-The post-processed \texttt{X-graph} then "fills in" the memory-model \texttt{W-model}, and both are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \texttt{Z-formula}, which then proceeds to the Z3 solver as an input.
-
-\begin{figure}%[!b]%[H]
+\begin{figure}[H]
   \centering
-  % TODO: uncomment after a while
-  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
-  \caption{Main components of \porthos[2]}
+  % TODO: remove scale 0.8 if it can fit the whole page
+  \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
+  \caption{General architecture of \porthos[2]}
   \label{fig:arch}
@@
 
-\subsection{The Model}
+Both \porthos[1] and \porthos[2] use the ANTLR parser generator%
+\footnote{The ANTLR project repository: \url{https://github.com/antlr/antlr4}}%
+~\cite{parr2013definitive}, a~powerful language processing tool.
+The ANTLR takes as input the user-defined grammar of the target language in a BNF-like form and produces the LL(*)-parser and optionally some auxiliary classes (such as listeners and visitors for the syntax tree).
+Although this parser may be not as efficient as the hand-written language-optimised parser, it reduces the overhead of implementing the parser significantly.
+Among other advantages ANTLR, it is worth of noting that it has rather large collection of officially supported grammars.
+Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the ANTLR an attractive instrument for solving the parsing problem.
+
+Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full ANTLR grammar is available at Appendix~\ref{apx:in_grammar_pts}.
+
+\begin{figure}[H]
+\begin{lstlisting}[mathescape=true,%
+                  caption={The sketch of the input language grammar used by \oldporthos},%
+                  label={fig:in_grammar_pts},%
+                  morekeywords={if,then,else,return,while,program,thread},%
+                  breaklines=true,%
+                  basicstyle=\ttfamily\scriptsize]
+<prog> : <init> <thrd>* <assert>
+       ;
+<thrd> : thread <tid> <inst>
+       ;
+<inst> : <atom>
+       | <inst> ; <inst>
+       | while <pred> <inst>
+       | if <pred> { <inst> } <inst>
+       ;
+<atom> : <reg> <- <expr>
+       | <reg> <:- <loc>
+       | <loc> := <reg>
+       | <reg> = <loc>.load(<atomic>)
+       | <loc> = <reg>.store(<atomic>)
+       | 'mfence'
+       | 'sync'
+       | 'lwsync'
+       | 'isync'
+       ;
+<pred> :
+       | true
+       | false
+       | <expr> (and | or) <expr>
+       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
+       ;
+<expr> : [0-9]
+       | <reg>
+       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
+       ;
+\end{lstlisting}
+\end{figure}
+
+The input language parser used by \porthos[1] suffered from several disadvantages.
+Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
+Such a combining of two rich languages%
+\footnote{by the term "reach" we mean "expressiveness": at least, the grammar of the language of semantic actions (i.e., java in our case) is Turing-complete.} %
+makes the code hardly understandable, and, therefore, poorly maintainable.
+In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the ANTLR syntax tree to the AST, that is one for all  languages of an input program.
+
+Secondly, the semantics of operations was defined syntactically, whereas processing programs written in most modern languages (including C) requires the semantics resolution based on typification%
+\footnote{for instance, given two functions
+`\lstinline{int foo(int a)}' and `\lstinline{int foo(char a)}', the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
+As the reader can notice from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, the memory operations of different kinds vary syntactically.
+For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the method name.
+In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
+The semantics of the methods also is being resolved during the pre-compilation stage via the \textit{invocation hooks} mechanism (see Section~\ref{ch:impl:proc:x-pre-compiler:hooks}).
+
+Thirdly, the grammar used by \porthos{} had restricted set of allowed operations.
+For example, it allowed only computations over the local variables, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code snippet `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
+Also, only integers were processed by the input language parser of \porthos[1].
+In \porthos[2], we extended support for primitive types supported by the \texttt{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s, floats encoded as \texttt{Real}s, enums encoded as Z3 \texttt{Scalars}).
+%TODO : constant arrays? https://rise4fun.com/Z3/tutorial/guide
+Thus, adding support for more advanced syntactic structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
+
+The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of type \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for statements (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only initialisation of the shared variable and only by default value \lstinline{0}.
+The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
+
+The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars~\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}).
+Currently, \porthos[2] can operate only in the intra-procedural analysis mode (analysing single procedure), assuming that each function defined in the input file is being executed in a separate thread.
+However, the redesigned architecture of \porthos[2] allow to easily support inter-procedural (cross-procedure) analysis by inlining function calls and binding variable contexts.
+Also, Current version of \porthos[2] simply ignores C processor directives, however, in future it is possible to support it.
+
+%TODO: say sth about 100500 litmus-tests for kernel
+
+\subsection{The internal representations}
 \label{ch:impl:model}
-blabla
+
+%As it has already been mentioned, all internal representations constructed by \porthos[2] are \textit{immutable}.
 
@@
 
-\subsubsection{Y-tree parser}
-\label{ch:impl:proc:y-parser}
+\subsubsection{Input program parser}
+\label{ch:impl:proc:inp-prog-parser}
+% todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
 
-\subsubsection{W-model parser}
-\label{ch:impl:proc:w-parser}
+\subsubsection{Input memory model parser}
+\label{ch:impl:proc:inp-mod-parser}
+
+\subsubsection{Y-tree constructor}
+\label{ch:impl:proc:y-constr}
+
+\subsubsection{X-graph pre-compiler}
+\label{ch:impl:proc:x-pre-compiler}
+
+\paragraph{Invocation hooks}
+\label{ch:impl:proc:x-pre-compiler:hooks}
 
@@
 
-\paragraph{Pre-compiler}
-\label{ch:impl:proc:x-compiler:pre}
+\subsubsection{X-graph transformer}
+\label{ch:impl:proc:x-transf}
 
-\paragraph{Compiler}
-\label{ch:impl:proc:x-compiler:comp}
+\subsubsection{W-model constructor}
+\label{ch:impl:proc:w-constr}
 
-\paragraph{Post-compiler}
-\label{ch:impl:proc:x-compiler:post}
+\subsubsection{Z-formula encoder}
+\label{ch:impl:proc:z-encoder}
 
-\subsubsection{The Z-formula encoder}
-\label{ch:impl:proc:z-encoder}
+%todo: mention The SMT-LIB standard http://smtlib.cs.uiowa.edu/papers/smt-lib-reference-v2.6-r2017-07-18.pdf
 
@@
 \label{ch:impl:out}
-
-%\subsection{Parsing the input}
-%\label{ch:impl:input-parser}
-
-Both \porthos{} and \porthos[2] \ use the parser generator ANTLR~\cite{parr2013definitive}, a powerful language processing tool.
-The full ANTLR grammar of input language used by previous version of \porthos{} is available at Appendix~\ref{apx:in_grammar_pts}.
-
-% TODO: why we've chosen ANTLR (cite Kvanttt); what kind of parser is ANTLR, why it's good for C and not only for C (unified way to define multiple grammars, large collection, ...) + add references
-
-- Figure~\ref{syntax:in_grammar_pts} represents the informal grammar of the input language processed by \porthos[1] ... %todo: cite \cite{Porthos17}.
-- short characteristics of the old grammar, enumerate its features (wrt the Figure)
-
-- DRAWBACKS: %what parts need to be improved:
-
-    -- the semantics of memory operations and method invocations is encoded directly into the grammar.
-    Taking into account the complexity of developing the accurate grammar for a Turing-complete language,
-
-    -- in old implementation, the logics of parser was written directly in the grammar.
-    Although being fast-to-impelement (1 internal representation less), this is approach has numerous drawbacks s.a.:
-    * hard to read grammar (mixing 2 languages in a single file)
-    * hard to maintain grammar (no debugging with standard utils, no code analysis in the grammar file)
-    * non-extensible (there is no option to just plug-in existing grammar for any language and write a converter from this language's syntax tree to our AST)
-
-    - syntactically only the integers were supported. we:arrays???pointers??enums?structs?
-    - no declarations (all shared variables are declared in the init section, local variables are not decalred ever)
-
-    - no arbitrary function calls (need support for the knowledge base after implementing the typisation. cannot be supproted only syntactically)
-
-- Minor drawbacks:
-
-    - restricted sytnax for expressions:
-
-    -- no operator associativity allowed (`1 + 2 * 3` was invalid syntax)
-
-    -- no unary increment/decrement (technically it's hard to implement post-incerment/decrement without interpretation: we don't know when does this expression ends.)
-    Example:
-    `int* x;
-    if (x++ > 0 ) { }`
-    % todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
-
-    -- incorrectly implemented statement in old code: sequence of stmts is separated by ';', but the syntax of C requires ';' as the operator that ends each operation (how to say this? see C standard)
-
-    -- litmus-initialisation  stmt did not allow the non-default values. The new syntax allow any type of declaration statement with initialisation
-
-- this means the full compiler engine that resolves the semantics
-
-%The grammar supported by \porthos[2] was extended in order to support expressions valid in C. %TODO: examples, how imenno extended
-%Also, the input language of the first version of \porthos{} syntactically distinguished the type of a memory instruction (shared variable assignments were defined with symbol~`\texttt{:=}', and local variable assignments --- with symbol~`$\mathtt{\leftarrow}$'), whereas \porthos[2] determines the semantics of a function invocation during the preprocessing stage of compilation~(see Chapter~\ref{TODO!!!}).
-
-% todo: this is the simplified grammar
-
-%TODO: although the grammar is context-free and fail whith <angle brackets>.... it still works fine. Even gcc uses CF grammar? reference to INRIA's parser and paper on discussing grammars for C
-
-%TODO: say sth about 100500 litmus-tests for kernel
-
-\begin{figure}%[H]
-%\begin{multicols}{2}
-\begin{lstlisting}[mathescape=true,%
-                  caption={Syntax of an input language of \porthos{} version 1},%
-                  label={syntax:in_grammar_pts},%
-                  %literate={<--}{{$\leftarrow$}}1,%
-                  morekeywords={if,then,else,return,while,program,thread}%
-                  breaklines=true,%
-                  basicstyle=\ttfamily\scriptsize]
-<prog> : <init> <thrd>* <assert>
-       ;
-<thrd> : thread <tid> <inst>
-       ;
-<inst> : <atom>
-       | <inst> ; <inst>
-       | while <pred> <inst>
-       | if <pred> { <inst> } <inst>
-       ;
-<atom> : <reg> <-- <expr>
-       | <reg> <-- <loc>
-       | <loc> := <reg>
-       | 'mfence'
-       | 'sync'
-       | 'lwsync'
-       | 'isync'
-       ;
-<pred> :
-       | true
-       | false
-       | <expr> (and | or) <expr>
-       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
-       ;
-<expr> : [0-9]
-       | <reg>
-       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
-       ;
-\end{lstlisting}
-%\end{multicols}
-\end{figure}
-
-- NOW: we're using the C grammar from the standard
-
-- for now, we just ignore C macro directives, in future it's planned to support partially the preprocessing before parsing
-
-The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official ANTLR repository on GitHub~\footnote{The repository containing the collection of ANTLR v4 grammars:\\ \url{https://github.com/antlr/grammars-v4}}).
-Current version of \porthos[2] can operate only in the inter-procedural mode, assuming that each function defined in the input file is being executed in a separate thread.
-However, the redesigned architecture of \porthos[2] allow to easily support intra-procedural analysis by inlining function calls.
 

### 2018-04-29 4_impl.tex
add support for verbatim listings of linux tree utility.

@@
 As \porthos[1], the \porthos[2] uses the open-source SMT solver \texttt{Z3} from Microsoft Research~\ref{de2008z3}.
-However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \texttt{Z-formula} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
+However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
 %TODO: *should* allow?
@@
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
-The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \texttt{Y-tree}%
+The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree%
 \footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
@@
 (Section~\ref{ch:impl:model:ytree})
-%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
+%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \wmodel (Section~\ref{ch:impl:model:wmodel}).
 , which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
-The \texttt{Y-tree} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \texttt{X-graph} representation (Section~\ref{ch:impl:model:xgraph}).
-The compiled \texttt{X-graph} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \texttt{Z-formula} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
-Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \texttt{X-graph} in order to build the weak memory model \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \texttt{W-model} and transformed \texttt{X-graph} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \texttt{Z-formula} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
+The \ytree then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph representation (Section~\ref{ch:impl:model:xgraph}).
+The compiled \xgraph then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \zformula (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
+Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \xgraph in order to build the weak memory model \wmodel (Section~\ref{ch:impl:model:wmodel}).
+Thereafter, \wmodel and transformed \xgraph are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
 
@@
 \label{ch:impl:model:ytree}
+
+The first internal representation that \porthos[2] uses is the \ytree, which represents rather a high-level AST.
+Any input program (whether it is high-level \texttt{C} code or low-level assembly code) is translated to the \ytree.
+
+\begin{paracol}{2}
+text text text text text text text text text text text text
+
+text text text text text text
+
+text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text
+
+text text text text text text text text text text text text text text text text text text text text text text text text text text
+
+text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text text
+
+\switchcolumn
+
+\VerbatimInput{inc/lst/Ytree-tree.out}
+\end{paracol}
+
+%The syntax tree of the \texttt{C} program is being desugared %examples
 

### 2018-04-30 2_wmm.tex
Change [H] of some pictures.

@@
         \begin{itemize}[noitemsep]
-            \item the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads its value;
-            \item the \textit{coherence order} relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation} \texttt{ws}-relation);
+            \item the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
+            \item the \textit{coherence order} relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
         \end{itemize}
@@
 
-    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to write events preceding the write event from which the read event gets the value: \\
+    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events preceding the write event from which the read event gets its value: \\
     $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$
@@
 Since each thread performs single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).
-Thus, there are only four possible executions defined by the choice of \rf-relation.
-The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and some others.
-However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over~$\fr\cup\po$.
-%However, the Power memory model allows such cycles, therefore %TODO: include load buffering example, see presentation pdf
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering
@@
 \end{figure}
+
+Thus, there are only four possible executions defined by the choice of \rf-relation.
+The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and some others.
+However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over~$\fr\cup\po$.
+%However, the Power memory model allows such cycles, therefore %TODO: include load buffering example, see presentation pdf
 

### 2018-04-30 4_impl.tex
add some text on Y-tree.

--- 2018-04-29 4_impl.tex
@@
 The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
-\footnote{Project repository: \url{https://github.com/google/guava/}}%
+\footnote{Guava project repository: \url{https://github.com/google/guava/}}%
 ).
@@
 
-As \porthos[1], the \porthos[2] uses the open-source SMT solver \texttt{Z3} from Microsoft Research~\ref{de2008z3}.
-However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \texttt{Z-formula} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
+As \porthos[1], the \porthos[2] uses the open-source SMT solver \texttt{Z3} from Microsoft Research~\cite{de2008z3}.
+However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
 %TODO: *should* allow?
@@
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
-The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \texttt{Y-tree}%
+The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
 \footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
@@
 (Section~\ref{ch:impl:model:ytree})
-%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
+%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
 , which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
-The \texttt{Y-tree} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \texttt{X-graph} representation (Section~\ref{ch:impl:model:xgraph}).
-The compiled \texttt{X-graph} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \texttt{Z-formula} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
-Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \texttt{X-graph} in order to build the weak memory model \texttt{W-model} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \texttt{W-model} and transformed \texttt{X-graph} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \texttt{Z-formula} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
+The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
+The compiled \xgraph{} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \zformula{} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
+Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
+Thereafter, \wmodel{} and transformed \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
 
@@
 
-Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full ANTLR grammar is available at Appendix~\ref{apx:in_grammar_pts}.
+Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full ANTLR grammar is available at Appendix~\ref{apx:grammar}.
 
-\begin{figure}[H]
+\begin{figure}[h]%[!hb]
+\centering
 \begin{lstlisting}[mathescape=true,%
@@
 The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of type \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for statements (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
-Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only initialisation of the shared variable and only by default value \lstinline{0}.
-The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
+Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
+tialisation of the shared variable and only by default value \lstinline{0}. The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
 
@@
 \label{ch:impl:model}
-
 %As it has already been mentioned, all internal representations constructed by \porthos[2] are \textit{immutable}.
@@
 \label{ch:impl:model:ytree}
+
+- TODO: everything inherited YEntity + citation
+
+The first internal representation that \porthos[2] uses is the \ytree{}, which represents a rather high-level AST.
+The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to representing the classes file tree only).
+
+%\columnratio{0.7}%\begin{paracol}{2}
+%\switchcolumn%\VerbatimInput{inc/lst/Ytree-tree.out}%\end{paracol}
+The abstract syntax tree \ytree{} is an abstraction level suitable for compiling to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
+As the ANTLR syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule \lstinline{postfixExpression '[' expression ']'}, that is converted to the \texttt{YIndexerExpression}).
+However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite{gries2012science}\textbf{TODO:CHAPTER,PAGE}), we found this level of abstraction as suitable enough for our tasks.
+
+Following the terms of C standard, we distinguish \textit{statements} (actions to be performed) and \textit{expressions} (TODO).
+
+- TODO: all statements inherit \texttt{YStatement} + interface lstlisting
+
+The \porthos[2] processes the following statements: 
+\begin{itemize}[nolistsep]
+  \item \texttt{YBranchingStatement} representing the \texttt{if-then-else} statement,
+  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}-loops, %TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
+  \item \texttt{YJumpStatement} representing unconditional jump (\texttt{goto}-jump to a label and loop-jumps \texttt{break} and \texttt{continue}),
+  \item \texttt{YCompoundStatement} (block statement) representing sequence of \texttt{N} statements grouped into one syntactic unit,
+  \item \texttt{YLinearStatement} representing single expression.
+\end{itemize}
+
+According to the C standard, \textit{"any statement may be preceded by a prefix that declares an identifier as a label name"}.
+However, the \ytree{} statements may have only simple symbolic labels, which have to be resolved at the pre-compilation stage.
+Apart from the listed set of statements, \ytree{} has also the \texttt{YFunctionDefinition} and its litmus-specific inheritor \texttt{YProcessStatement} used in intra-procedural analysis mode.
+The function definition contains the \texttt{YCompoundStatement}-typed body and the signature of type \texttt{YMethodSignature}, which is used in the function resolution during the compilation stage.
+%TODO: pre-compilation?
+The other litmus-specific statements are \texttt{YPreludeStatement} that carries the list of \texttt{YStatement}-typed initial writes, and \texttt{YPostludeStatement} that carries the \texttt{YExpression}-typed binary expression to be asserted by the litmus test.
+
+Similar to statements of \ytree{}, its expressions
+
+pointer operations are modeled as the pointer level of and expression (although it is the property of type, the y-tree is untyped, so elements that represent syntactic expressions shoudl carry tnis information)
+
+Each element of \ytree{} carries the \texttt{OriginLocation} %
+\textbf{(TODO: Rename CodeLocation->OriginLocation in the code!)} %
+instance that contains information about the coordinates of the input text that generated the \ytree{} element.
 
@@
 \label{ch:impl:proc:y-constr}
+
+The ANTLR syntax tree of C
+
+%, comparing to ANTLR syntax tree, does not have 
+- desugar, equiv transform
+- variables distinguished from
 

### 2018-04-30 4_impl.tex
Finish the Y-tree section.

@@
 
-\section{General architecture}
+\section{Architecture}
 \label{ch:impl:arch}
@@
 
-\subsection{The program input}
+\subsection{Program input}
 \label{ch:impl:input}
@@
 
-The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of type \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for statements (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
 Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
@@
 
-The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars~\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}).
+The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars%
+\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}).%
 Currently, \porthos[2] can operate only in the intra-procedural analysis mode (analysing single procedure), assuming that each function defined in the input file is being executed in a separate thread.
@@
 
-- TODO: everything inherited YEntity + citation
-
 The first internal representation that \porthos[2] uses is the \ytree{}, which represents a rather high-level AST.
-The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to representing the classes file tree only).
+The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only).
 
@@
 %\switchcolumn%\VerbatimInput{inc/lst/Ytree-tree.out}%\end{paracol}
-The abstract syntax tree \ytree{} is an abstraction level suitable for compiling to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
+The abstract syntax tree \ytree{} is an abstraction level suitable for compiling the program to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
 As the ANTLR syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule \lstinline{postfixExpression '[' expression ']'}, that is converted to the \texttt{YIndexerExpression}).
-However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite{gries2012science}\textbf{TODO:CHAPTER,PAGE}), we found this level of abstraction as suitable enough for our tasks.
+However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite[Chapter 5]{gries2012science}), we found this level of abstraction suitable enough for our tasks.
 
-Following the terms of C standard, we distinguish \textit{statements} (actions to be performed) and \textit{expressions} (TODO).
+Each \ytree{} element implements the interface \texttt{YEntity} and carries the \texttt{OriginLocation} %
+%\textbf{(TODO: Rename CodeLocation->OriginLocation in the code!)} %TODO <--
+instance that contains information about the coordinates of the input text that generated the \ytree{} element.
 
-- TODO: all statements inherit \texttt{YStatement} + interface lstlisting
+Following the C11 standard~\cite{iso2012iec}, we distinguish a \textit{statement} (\textit{"an action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
 
-The \porthos[2] processes the following statements: 
-\begin{itemize}[nolistsep]
-  \item \texttt{YBranchingStatement} representing the \texttt{if-then-else} statement,
-  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}-loops, %TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
-  \item \texttt{YJumpStatement} representing unconditional jump (\texttt{goto}-jump to a label and loop-jumps \texttt{break} and \texttt{continue}),
-  \item \texttt{YCompoundStatement} (block statement) representing sequence of \texttt{N} statements grouped into one syntactic unit,
-  \item \texttt{YLinearStatement} representing single expression.
+All \ytree{} expressions implement the \texttt{YExpression} interface.
+On the \ytree{} level, the pointer arithmetic is modelled by the integer number \textit{pointer level} of an expression (although in fact this is the property of a type not of an expression, the y-tree is an untyped syntax tree, therefore the elements \ytree{} should carry this property).
+We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is defined also syntactically).
+%\textbf{(TODO:THIS SHOULD BE IN PRE-PROCESSING)} %TODO <--
+
+\vspace{0.5em}
+The \ytree{} expressions are the following:
+
+\begin{itemize}%[nolistsep]
+  \item \texttt{YBinaryExpression} that model the C binary operator (\textit{relative} operator that compares two expressions of any type, \textit{logical} that processes two boolean expressions, and \textit{numerical} that processes two numerical expressions);
+  %\textbf{TODO: rename 'integer'->'numerical' in code} %TODO <--
+
+  \item \texttt{YUnaryExpression} that model the C unary expression (logical negation, numeric prefix and postfix increment and decrement, bitwise complement);
+
+  \item \texttt{YMemberAccessExpression} that has an arbitrary expression of type \texttt{YExpression} as its base expression (it will be resolved during the compilation stage);%TODO: pre-compilation??
+
+  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT solver we use supports the constant-array theory, we can maintain the array type);
+
+  \item \texttt{YAssignmentExpression} that assigns an \texttt{YExpression} to an \texttt{YAtom};
+
+  \item \texttt{YVariableRef} that stores the "reference" to a variable (viz., the name of an object);
+
+  \item \texttt{YLabeledVariableRef} that represents the litmus-specific local variable reference for a certain the process (e.g., `\lstinline{P0:x}' which means the local variable \lstinline{x} of the process \lstinline{P0});
+
+  \item \texttt{YParameter} that represents a typed variable (the type was declared, similarly to the variable definition);
+
+  \item \texttt{YConstant} that represents an untyped non-named constant.
 \end{itemize}
 
+%\vspace{0.5em}
+Similarly to expressions, all \ytree{} statements implement the \texttt{YStatement} interface. The statements are the following: 
+%\textbf{(TODO: extract interface YStatement, create abstract class YStatementBase)} %TODO <--
+
+\begin{itemize}%[nolistsep]
+  \item \texttt{YBranchingStatement} representing the \texttt{if-then-else} statement;
+
+  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}-loops;%TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
+
+  \item \texttt{YJumpStatement} representing unconditional jump (\texttt{goto}-jump to a label and loop-jumps \texttt{break} and \texttt{continue});
+
+  \item \texttt{YCompoundStatement} (block statement) representing sequence of \texttt{N} statements grouped into one syntactic unit;
+
+  \item \texttt{YLinearStatement} representing single expression;
+
+  \item \texttt{YVariableDeclarationStatement} containing the information about the variable type during the variable declaration.
+\end{itemize}
+
+On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as an alias for a type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information of the type for declarations
+
 According to the C standard, \textit{"any statement may be preceded by a prefix that declares an identifier as a label name"}.
-However, the \ytree{} statements may have only simple symbolic labels, which have to be resolved at the pre-compilation stage.
-Apart from the listed set of statements, \ytree{} has also the \texttt{YFunctionDefinition} and its litmus-specific inheritor \texttt{YProcessStatement} used in intra-procedural analysis mode.
+The \ytree{} statements of follow this rule, however they these labels are symbolic, and they need to be resolved at the pre-compilation stage.
+Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific declaration \texttt{YProcessDefinition}
+%\textbf{(TODO: rename YProcessStatement->YProcessDefinition)} %TODO <--
+used in intra-procedural analysis mode.
 The function definition contains the \texttt{YCompoundStatement}-typed body and the signature of type \texttt{YMethodSignature}, which is used in the function resolution during the compilation stage.
 %TODO: pre-compilation?
-The other litmus-specific statements are \texttt{YPreludeStatement} that carries the list of \texttt{YStatement}-typed initial writes, and \texttt{YPostludeStatement} that carries the \texttt{YExpression}-typed binary expression to be asserted by the litmus test.
+The other litmus-specific statements are \texttt{YPreludeDefinition}
+%\textbf{(TODO: rename YPreludeStatement -> YPreludeDefinition)} %TODO <--
+that carries the list of \texttt{YStatement}-typed initial writes, and \texttt{YPostludeDefinition} 
+%\textbf{(TODO: rename YPostludeStatement -> YPostludeDefinition)} %TODO <--
+that carries the \texttt{YExpression}-typed binary expression to be asserted by the litmus test.
 
-Similar to statements of \ytree{}, its expressions
-
-pointer operations are modeled as the pointer level of and expression (although it is the property of type, the y-tree is untyped, so elements that represent syntactic expressions shoudl carry tnis information)
-
-Each element of \ytree{} carries the \texttt{OriginLocation} %
-\textbf{(TODO: Rename CodeLocation->OriginLocation in the code!)} %
-instance that contains information about the coordinates of the input text that generated the \ytree{} element.
+The syntax tree that contains set of definitions (e.g., litmus-initialisations, function definitions, litmus-asserts) is modelled by the class \texttt{YSyntaxTree}.
 
@@
 \label{ch:impl:model:xgraph}
+
+The second internal representation used by \porthos[2] is the event-based flow graph \xgraph{}. This is the
+
+two basic elements of diff nature: XEvent and YMemoryUnit
+
+comp event emulated by both -- optimisation?
 
@@
 
-\subsection{The program output}
+\subsection{Program output}
 \label{ch:impl:out}

### 2018-05-02 4_impl.tex
Add macro for unrolled xgraph. Add some plan + precompilation section.

@@
 The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
+On the picture, rectangles denote processing units (marked with gear sign with a number of the component for referencing it).
+
+\begin{figure}%[H]
+    \centering
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
+  \caption{The general architecture of \porthos[2]}
+  \label{fig:arch}
+\end{figure}
+
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
@@
 The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
-The compiled \xgraph{} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-transf}) necessary for the encoding into a \zformula{} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-transf}).
+The compiled \xgraph{} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-unroll}) necessary for the encoding into a \zformula{} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-unroll}).
 Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
 Thereafter, \wmodel{} and transformed \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
-
-\begin{figure}[H]
-  \centering
-  % TODO: remove scale 0.8 if it can fit the whole page
-  \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
-  \caption{General architecture of \porthos[2]}
-  \label{fig:arch}
-\end{figure}
 
@@
 Currently, \porthos[2] can operate only in the intra-procedural analysis mode (analysing single procedure), assuming that each function defined in the input file is being executed in a separate thread.
+%TODO: describe more intra- and inter-procedural!
 However, the redesigned architecture of \porthos[2] allow to easily support inter-procedural (cross-procedure) analysis by inlining function calls and binding variable contexts.
+%TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs. 
 Also, Current version of \porthos[2] simply ignores C processor directives, however, in future it is possible to support it.
@@
 
-  \item \texttt{YLinearStatement} representing single expression;
+  \item \texttt{YLinearStatement} representing a single expression;
 
@@
 
-The second internal representation used by \porthos[2] is the event-based flow graph \xgraph{}. This is the
+The \ytree{} is compiled into the low-level event-based program representation \xgraph{}.
+The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
+In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
+%.. goto: compilation
 
-two basic elements of diff nature: XEvent and YMemoryUnit
+atomic action?
+
+%However, the \xgraph{}
+%The basic element of \xgraph{} is an event. Each event \texttt{XEvent}
+Firstly, two elements
+%two basic elements of diff nature: XEvent and YMemoryUnit
 
 comp event emulated by both -- optimisation?
+
+secondly, relations edges (special class? no, just map with enum as kind of edge)
+
+CF edge: then-edge (primary), else-edge (alternative)
+
+XEventInfo
+
+hierarchy of memory units
+
+memory events
+
+computation: reduce to method call event
+MethCallE (descr in transformer section)
+
+XBarrierEvent - kinds
+
+controlflow/ %TODO: rename package controlflow->empty (NOT FORGET TO UPDATE FILE TREE LISTING)
+XNopEvent
+XJumpEvent - no kinds, same as nop
+entry
+two kinds of exit
+
+Valid graph:
+- single source, two sinks
+- connected
+- each node has 1 or 2 children
+- branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
+-
+
+//Picture of the CF-graph
+Consider the following code in C:
 
@@
 
+a simple set of recursively defined relations
+
+atoms: sets/basic relations
+
 \subsubsection{Z-formula}
 \label{ch:impl:model:zformula}
+
+%TODO: not implemented at all !
+
+in \porthos[1] : ctx as an argument, calling ctx to create new clause. 
+drawbacks:
+- hard to debug
+- only one solver
+- non-safe way to construct formula (if ctx has changed, runtime excpetion)
+
+in \porthos[2] we created the new abstraction layer \zformula that is translated to
+
+also recursive representation of smt formula
+
+used as an abstrac
+
+implemented as a simple wrapper over the z3 formula
+
+typed: Expr, BoolExpr, ArithExpr, ...
 
@@
 
-\subsubsection{Input program parser}
+\subsubsection{Input memory model parser (1)}
+\label{ch:impl:proc:inp-mod-parser}
+
+ANTLR grammar is extracted from parser used by herd written in OCaml
+
+\subsubsection{Input program parser (2)}
 \label{ch:impl:proc:inp-prog-parser}
@@
 
-\subsubsection{Input memory model parser}
-\label{ch:impl:proc:inp-mod-parser}
+%TODO: maybe merge this and previous subsubsections
 
-\subsubsection{Y-tree constructor}
+todo: move something from the input language chapter here
+
+\subsubsection{W-model constructor (3)}
+\label{ch:impl:proc:w-constr}
+
+same as y-tree constructor
+
+in perspective: work with functional-style definitions
+
+<not implemented yet>
+
+\subsubsection{Y-tree constructor (4)}
 \label{ch:impl:proc:y-constr}
@@
 
-\subsubsection{X-graph pre-compiler}
+- if we don't support , our parser still parses it, and the error is thrown at the moment of converting syntax tree to the AST (Y-tree).
+
+- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11->Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
+need ly?)
+
+\subsubsection{X-graph precompiler (5)}
 \label{ch:impl:proc:x-pre-compiler}
 
-\paragraph{Invocation hooks}
+\paragraph{The macro preprocessor}
+
+typedefs, aliases
+
+not supported yet
+
+\paragraph{Variables kinds determination}
+\label{ch:impl:proc:x-pre-compiler:var}
+
+todo
+
+\paragraph{The label resolution}
+\label{ch:impl:proc:x-pre-compiler:label}
+
+The label resolution is a process of linkage the referenced labels to declared labels.
+
+In C, the labeled statements are declared via the colon-syntax `\texttt{<label> : <statement>}',
+and the labels are referenced by the jump-statement `\texttt{goto <label>}'.
+The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labeled statement.
+%TODO: code: remove temp label generation, it's stupid.
+put null.
+This information is used during compilation (see Section~\ref{ch:impl:proc:x-compiler}).
+
+%TODO: <not implemented yet>
+
+\paragraph{Type analysis}% and function resolution.}
+\label{ch:impl:proc:x-pre-compiler:type}
+
+%\textit{Type analysis} is the process of \textit{converting} the text-based type labels (obtained at variable declarations) to the type instances (modelled by the class \texttt{mousquetaires.languages.common.Type}) and \textit{propagating} them to all expressions.
+The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
+Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
+In \porthos[2], the type is modelled by the class \texttt{mousquetaires.languages.common.Type}.%TODO: + fix class and package name!
+Being carried at Y and X representation levels, the type is converted to a Z-type at the Z-formula encoding stage (see Section~\ref{ch:impl:proc:z-encoder}).
+
+%TODO: describe the Type
+%TODO: maybe rename back to XYType
+%TODO: but definitely get rid of bit32
+
+- X-Y-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+
+- The typification algorithm: propagation rules for each of Y and X
+
+- more about type abstraction: common for all levels. Simple label?
+
+- data-model?
+
+\paragraph{Invocation hooking.}
 \label{ch:impl:proc:x-pre-compiler:hooks}
 
-\subsubsection{X-graph compiler}
+resolve the semantics of a function given its signature (arguments are already typed)
+
+knowledge base
+
+\subsubsection{X-graph compiler (6)}
 \label{ch:impl:proc:x-compiler}
 
-\subsubsection{X-graph transformer}
-\label{ch:impl:proc:x-transf}
+Firstly, the \ytree{} is compiled to the \textit{cyclic control-flow event-based graph} \xgraph[CF].
 
-\subsubsection{W-model constructor}
-\label{ch:impl:proc:w-constr}
+Then: unrolled up to bound $k$ to the \textit{acyclic control-flow event-based graph} \xgraphU[CF].
 
-\subsubsection{Z-formula encoder}
+Then: data-flow analysis to the \textit{acyclic full event-based graph} \xgraphU[CF+DF].
+
+no contexts yet
+
+MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
+reutrn event= assign the return-register + simple jump
+
+\subsubsection{X-graph unroller (7)}
+\label{ch:impl:proc:x-unroll}
+
+simple algorithm
+
+how we attempted, why it didn't work in general case:
+-
+
+why we changed the notion of unrolling bound
+
+\subsubsection{X-graph data-flow constructor (8)}
+\label{ch:impl:proc:x-df}%TODO: ref it in pre-word in this chapter
+
+\subsubsection{Z-formula encoder (9)}
 \label{ch:impl:proc:z-encoder}
+
+Z-type : SMT-specific
+
+\subsubsection{SMT-formula translator (10)}
+\label{ch:impl:proc:smt-transator}%TODO: ref in pre-word
+
+\subsubsection{SMT-formula translator (11)}
+\label{ch:impl:proc:smt-solver}%TODO: ref it in pre-word in this chapter
+%TODO: check that we have ALL references in pre-word in this chapter
 
@@
 
+structure: verdict
+
 %TODO: say sth about syntax errors handling: still now, what if parser does not accept syntax?
 
-- if we don't support , our parser still parses it, and the error is thrown at the moment of converting syntax tree to the AST (Y-tree).
+\subsection{Auxiliary components}
+\label{ch:impl:aux}
 
-- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11->Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
-need ly?)
+\subsubsection{Watchdog timer}
+\label{ch:impl:aux:watchdog}
 
-%\section{The Y-tree: an AST}
-%\label{ch:impl:ytree}
+\subsubsection{}
+\label{ch:impl:aux:watchdog}
 
@@
 TODO
-
-\section{Compiling the Y-tree to the X-graph}
-\label{ch:impl:y2x}
 
@@
 - determine kind of variables (cannot be done during parsing? don't say this)
-
-- basic typisation
 

### 2018-05-02 inc/lst/sample-code.c

@@ -1,0 +1,16 @@
+void P0(int& x, int& y) {
+  int a = 0;
+  // for loop
+  while (x++ > y) {
+    y = y - ++a;
+    // call foo
+    if (a > y) {
+      break;
+    }
+  }
+}
+
+
+int foo(int a, int b) {
+ 
+}

### 2018-05-03 4_impl.tex
Change font of representations names. Add mem-units descrition.

@@
 
-The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
-On the picture, rectangles denote processing units (marked with gear sign with a number of the component for referencing it).
+The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}. On the picture, rectangles denote processing units (marked with gear sign with a unique number of the component).
 
@@
     \centering
+  %draft=true
   \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
@@
 \label{ch:impl:model}
-%As it has already been mentioned, all internal representations constructed by \porthos[2] are \textit{immutable}.
+
+As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
+For keeping the architecture transparent, we build all abstraction levels with interfaces, even if some of them does not add any new functionality.% (as java does not support multiple inheritance for classes (only for interfaces), we
 
@@
 
-The first internal representation that \porthos[2] uses is the \ytree{}, which represents a rather high-level AST.
+The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a rather high-level AST.
 The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only).
@@
 used in intra-procedural analysis mode.
-The function definition contains the \texttt{YCompoundStatement}-typed body and the signature of type \texttt{YMethodSignature}, which is used in the function resolution during the compilation stage.
+%TODO: rename YMethodSignature -> YFunctionSignature
+The function definition contains the \texttt{YCompoundStatement} body and the \texttt{YMethodSignature} signature, which is used in the function resolution during the compilation stage.
 %TODO: pre-compilation?
@@
 %\textbf{(TODO: rename YPreludeStatement -> YPreludeDefinition)} %TODO <--
-that carries the list of \texttt{YStatement}-typed initial writes, and \texttt{YPostludeDefinition} 
+that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeDefinition} 
 %\textbf{(TODO: rename YPostludeStatement -> YPostludeDefinition)} %TODO <--
-that carries the \texttt{YExpression}-typed binary expression to be asserted by the litmus test.
+that carries the \texttt{YExpression} binary expression to be asserted by the litmus test.
 
@@
 
-The \ytree{} is compiled into the low-level event-based program representation \xgraph{}.
+The \ytree{} is compiled into the low-level event-based program representation \textit{\xgraph{}}.
 The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
-In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
-%.. goto: compilation
+%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
+The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
+Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
+The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
+Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.
 
-atomic action?
+\begin{figure}[t]%[!b]%[H]
+  \centering
+  \includegraphics[width=\textwidth,keepaspectratio]{img/my/class-diagrams/XEntity-interfaces.png}
+  \caption{The inheritance tree of main interfaces of \xgraph{} elements}
+  \label{fig:class-diagrams:XEntity-interfaces}
+\end{figure}
 
-%However, the \xgraph{}
-%The basic element of \xgraph{} is an event. Each event \texttt{XEvent}
-Firstly, two elements
-%two basic elements of diff nature: XEvent and YMemoryUnit
+All elements of \xgraph{} implement the interface \texttt{XEntity}.
+There are two main kinds of X-entity: \textit{events} that implement the \texttt{XEvent} interface, and \textit{memory-units} that implement the \texttt{XMemoryUnit} interface.
 
-comp event emulated by both -- optimisation?
+\paragraph{Memory units}
+\label{ch:impl:model:xgraph:mem}
+
+The \textit{memory unit} is a memory cell of an abstract machine executing the code.
+This machine has an infinite number of arbitrary-sized \textit{registers} (local memory units) and \textit{locations} (shared memory units).
+Local memory units inherit the \texttt{XProcessLocalElement} interface, that stores the ID of the owning process.
+Figure~\ref{fig:class-diagrams:XMemoryUnit} represents inheritance hierarchy of memory units.
+
+\begin{figure}[t]%[!b]%[H]
+  \centering
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XMemoryUnit-m.png}
+  \caption{The inheritance tree of \xgraph{} memory units}
+  \label{fig:class-diagrams:XMemoryUnit}
+\end{figure}
+
+Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (l-values may be assigned the new value).
+As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, more detailed discussion goes further in current Section).
+
+Each memory unit has an \texttt{XType} associated with it.
+The X-type is a symbolic representation of the C primitive type%
+\footnote{Here we should note the \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++).
+In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language.
+See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
+that is easily convertable to a SMT type (modelled as \texttt{ZType}).%TODO: add the class ZType
+
+\paragraph{Events}
+\label{ch:impl:model:xgraph:evt}
+
+atomic action
+XEventInfo
+
+%comp event emulated by both -- optimisation?
+Note, the computation-event \texttt{XComputationEvent} implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
+This is a model-level optimisation possible because a computation-event
+%\textit{is} an event (thus it can form a node in \xgraph{}) and it also 
+performs a computation over local-only memory and does not change the value of any memory unit, therefore the \textit{computation} abstraction (the CPU time) can be safely removed from the model, and the computation event can be seen as an atomic operation.
 
@@
 
-XEventInfo
+DF edge: co, rf
 
-hierarchy of memory units
+Kinds of events: (paragraph) <pciture>
 
@@
 
-Valid graph:
+Valid graph: (paragraph)
 - single source, two sinks
@@
 \label{ch:impl:proc}
+
+- numbers -- same as in gears in Figure~\ref{fig:arch}.
 
@@
 %\textit{Type analysis} is the process of \textit{converting} the text-based type labels (obtained at variable declarations) to the type instances (modelled by the class \texttt{mousquetaires.languages.common.Type}) and \textit{propagating} them to all expressions.
+%TODO: remake this paragraph considering that we have already said sth about XType
 The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
 Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
-In \porthos[2], the type is modelled by the class \texttt{mousquetaires.languages.common.Type}.%TODO: + fix class and package name!
+%In \porthos[2], the type is modelled by the class \texttt{XType}.%TODO: + fix class and package name!
 Being carried at Y and X representation levels, the type is converted to a Z-type at the Z-formula encoding stage (see Section~\ref{ch:impl:proc:z-encoder}).
@@
 %TODO: describe the Type
-%TODO: maybe rename back to XYType
+%TODO: maybe rename back to XType
 %TODO: but definitely get rid of bit32

### 2018-05-03 2_wmm.tex
a bit of X.

@@
+%TODO: move this chapter as a section to introduction and add here the new "Problem statement" chapter!!!
+
 \chapter{Memory model-aware analysis}%Description of the analysis of concurrent systems wrt weak memory models
@@
 The classical approach for modeling concurrent programs is to use the \textit{global time}, a single order of interleavings among all events happened in different threads.
-Although these models are easy to understand, it may be impossible to process \textit{all} possible states, number of which is exponentially large.
+Although these models are easy to understand, it may be impossible to treat \textit{all} possible states, number of which is exponentially large.
 However, there exist equivalence classes such that the result of execution different interleavings from single equivalence class is the same (for instance, computations performed by a processor locally do not affect the global state).
@@
 One such model is the \textit{event-based} representation of a program, which models a program as a directed graph of events (the \textit{event-flow graph}).
-The vertices of such a graph represent \textit{events} (independent low-level instructions; see Section~\ref{ch:wmm:model:events}), and edges represent \textit{relations} over the events (see Section~\ref{ch:wmm:model:relations}).
+The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent \textit{basic relations} (see Section~\ref{ch:wmm:model:relations}).
 
@@
 
+The event is an independent low-level atomic operation.
+
 A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory.
+Only memory events change the state of an abstract machine executing the code, since it is completely determined by values stored in its memory.
 Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events. 

### 2018-05-03 4_impl.tex
a bit of X.

@@
   %draft=true
-  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
+  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
   \caption{The general architecture of \porthos[2]}
@@
 \footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
-For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.} %
-(Section~\ref{ch:impl:model:ytree})
-%, and the memory model is being parsed (Section~\ref{ch:impl:proc:w-parser}) to another AST called \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-, which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
+For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
+(Section~\ref{ch:impl:model:ytree}), which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
 The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
@@
 
-  \item \texttt{YVariableRef} that stores the "reference" to a variable (viz., the name of an object);
+  \item \texttt{YVariableRef} that stores the untyped "reference" to a variable (viz., the name only);
 
@@
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/class-diagrams/XEntity-interfaces.png}
-  \caption{The inheritance tree of main interfaces of \xgraph{} elements}
+  \caption{The inheritance tree of interfaces of \xgraph{}}
   \label{fig:class-diagrams:XEntity-interfaces}
@@
 
-Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (l-values may be assigned the new value).
-As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, more detailed discussion goes further in current Section).
+Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (unlike r-values, the l-values may be assigned a new value).
+As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, see more detailed discussion further in current Section).
 
@@
 
+- memory manager: no contexts so far; locals per process + shareds; register/deregister;
+
 \paragraph{Events}
@@
 
-atomic action
-XEventInfo
+The following interfaces capture main types of events (see also Figure~\ref{fig:class-diagrams:XEntity-interfaces}):
 
-%comp event emulated by both -- optimisation?
-Note, the computation-event \texttt{XComputationEvent} implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
-This is a model-level optimisation possible because a computation-event
+\begin{itemize}
+  \item \texttt{XMemoryEvent} \\
+  The memory event defines the transfer of the value from one memory unit to another.
+  There are four types of memory events (the arrow denotes the direction of the data-flow):
+  \begin{enumerate}
+  \item \texttt{XRegisterMemoryEvent}: \\
+    $\texttt{XLocalLvalueMemoryUnit} \leftarrow \texttt{XLocalMemoryUnit}$
+  \item \texttt{XLoadMemoryEvent}: \\
+    $\texttt{XLocalLvalueMemoryUnit} \leftarrow \texttt{XSharedMemoryUnit}$
+  \item \texttt{XStoreMemoryEvent}: \\
+    $\texttt{XSharedLvalueMemoryUnit} \leftarrow \texttt{XLocalMemoryUnit}$
+  \item \texttt{XInitialWriteEvent}: \\
+    $\texttt{XLvalueMemoryUnit} \leftarrow \texttt{XRvalueMemoryUnit}$
+
+  \end{enumerate}%TODO incompleted enumeration
+  %TODO: either say st about getLoc(), getReg(), or remove them from code
+
+  \item \texttt{XComputationEvent} \\
+  - binary and unary \\
+  - implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
+  This is a model-level optimisation possible \textbf{due to} because a computation-event
   %\textit{is} an event (thus it can form a node in \xgraph{}) and it also 
@@
 
-secondly, relations edges (special class? no, just map with enum as kind of edge)
+  \item \texttt{XControlFlowEvent} \\
+    \texttt{XJumpEvent} no computation, no data operation. may be safely removed for optimisation
+    \texttt{XMethodCallEvent} : TODO: meth call should inherit both control-flow and computation.
+    itself - computation result (temp register) + if (method resolved) => context binding + jump to the method body
 
-CF edge: then-edge (primary), else-edge (alternative)
+  \item \texttt{XFenceEvent}
+    \texttt{XBarrierEvent} : mfence, sync, optsync, lwsync, optlwsync, ish, isb, isync
 
-DF edge: co, rf
+  \item \texttt{XFakeEvent}
+    \begin{enumerate}
+    \item \texttt{XEntryEvent}, the source event in the event-flow graph %only event that cannot have incoming edges
+    \item \texttt{XExitEvent} : two kinds (depending on the unrolling boundary)
+    \item \texttt{XNopEvent} : no-operation event, used for correct encoding (picture); same as simple jump to the next event
+    \end{enumerate}
 
-Kinds of events: (paragraph) <pciture>
+\end{itemize}
 
-memory events
+Edges (relations edges (just map with enum as kind of edge)):
+\begin{enumerate}
+\item CF edge: then-edge (primary), else-edge (alternative)
+\item DF edge: co, rf
+\end{enumerate}
 
-computation: reduce to method call event
-MethCallE (descr in transformer section)
+Valid graph:
+\begin{enumerate}
+\item single source, two sinks
+\item connected
+\item each node has 1 or 2 children
+\item branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
+\end{enumerate}
 
-XBarrierEvent - kinds
+consider the following code in C:
 
-controlflow/ %TODO: rename package controlflow->empty (NOT FORGET TO UPDATE FILE TREE LISTING)
-XNopEvent
-XJumpEvent - no kinds, same as nop
-entry
-two kinds of exit
+picture of the CF-graph:
 
-Valid graph: (paragraph)
-- single source, two sinks
-- connected
-- each node has 1 or 2 children
-- branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
--
+XEventInfo:
+Events are specified by the process generated them and a unique label.
+This information is stored by events as the immutable structure \texttt{XEventInfo}.
+%hashcode, uniqueness
 
-//Picture of the CF-graph
-Consider the following code in C:
+%\lstinputlisting{inc/XEventVisitor-cleaned.java}
 
@@
 \label{ch:impl:model:wmodel}
+%TODO: almost not implemented !
+<to be done>
 
-a simple set of recursively defined relations
+a simple set of recursively defined relations over ..
 
 atoms: sets/basic relations
+
+itemize elements
 
@@
 
+%todo: after all, list HERE which modules (classes, managers) are being initialised during the pre-compilation. I'ts good to have the dependencies graph on moduls (operational semantics:) )
+
 - numbers -- same as in gears in Figure~\ref{fig:arch}.
@@
 
-same as y-tree constructor
+visitor of antlr tree: almost direct
+
+todo: grammar of w-model to appendix
 
@@
 
-The ANTLR syntax tree of C
+from ANTLR syntax tree of C
 
-%, comparing to ANTLR syntax tree, does not have 
 - desugar, equiv transform
+
 - variables distinguished from
@@
 
-- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11->Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
+- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11 -> Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
 need ly?)
@@
 \label{ch:impl:proc:x-pre-compiler}
+
+%The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{(ENTRY)} (usually for convenience all leaves are connected to the sink node \texttt{(EXIT)}).
+The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).
+However, the cyclic flow graph cannot be encoded into SMT formula since ...
+%//TODO:REFERENCE.%TODO
+
+%note https://en.wikipedia.org/wiki/Semantics_(computer_science)
+%Operational semantics loosely corresponds to interpretation,
 
@@
 
-todo
+- pointers
+- externs
+- parameters of process-functions
 
@@
 \label{ch:impl:proc:x-pre-compiler:label}
+
+%TODO: <not implemented yet>
 
@@
 
-%TODO: <not implemented yet>
-
 \paragraph{Type analysis}% and function resolution.}
@@
 
-%\textit{Type analysis} is the process of \textit{converting} the text-based type labels (obtained at variable declarations) to the type instances (modelled by the class \texttt{mousquetaires.languages.common.Type}) and \textit{propagating} them to all expressions.
 %TODO: remake this paragraph considering that we have already said sth about XType
@@
 
-- X-Y-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+- X-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
 
-- The typification algorithm: propagation rules for each of Y and X
+- The typification algorithm: propagation rules for each of Y
 
-- more about type abstraction: common for all levels. Simple label?
-
-- data-model?
-
-\paragraph{Invocation hooking.}
+\paragraph{Invocation hooking}
 \label{ch:impl:proc:x-pre-compiler:hooks}
 
-resolve the semantics of a function given its signature (arguments are already typed)
+resolve the semantics of a function given its signature (meth name + params types)
 
-knowledge base
+knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
 
@@
 \label{ch:impl:proc:x-compiler}
+
+- hierarchy of Compilers (XCompiler is an stateful abstract machine)
 
@@
 
-no contexts yet
-
-MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
-reutrn event= assign the return-register + simple jump
+%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
+%reutrn event= assign the return-register + simple jump
 
@@
 \label{ch:impl:proc:x-unroll}
-
-simple algorithm
-
-how we attempted, why it didn't work in general case:
--
-
-why we changed the notion of unrolling bound
-
-\subsubsection{X-graph data-flow constructor (8)}
-\label{ch:impl:proc:x-df}%TODO: ref it in pre-word in this chapter
-
-\subsubsection{Z-formula encoder (9)}
-\label{ch:impl:proc:z-encoder}
-
-Z-type : SMT-specific
-
-\subsubsection{SMT-formula translator (10)}
-\label{ch:impl:proc:smt-transator}%TODO: ref in pre-word
-
-\subsubsection{SMT-formula translator (11)}
-\label{ch:impl:proc:smt-solver}%TODO: ref it in pre-word in this chapter
-%TODO: check that we have ALL references in pre-word in this chapter
-
-%todo: mention The SMT-LIB standard http://smtlib.cs.uiowa.edu/papers/smt-lib-reference-v2.6-r2017-07-18.pdf
-
-\subsection{Program output}
-\label{ch:impl:out}
-
-structure: verdict
-
-%TODO: say sth about syntax errors handling: still now, what if parser does not accept syntax?
-
-\subsection{Auxiliary components}
-\label{ch:impl:aux}
-
-\subsubsection{Watchdog timer}
-\label{ch:impl:aux:watchdog}
-
-\subsubsection{}
-\label{ch:impl:aux:watchdog}
-
-- picture of the Y-hierarchy. everything inherits interface YEntity
-
-- immutable
-
-- AST is untyped (YVariableRef).
-
-- short characteristics with citations from the code (this AST contains very basic language elements according to the C execution model (statements and expressions) )
-
-- minor changes are performed by converting to ytree representation: desugaring the target code, etc. (what else?)
-
-%\section{Model parser}
-%\label{ch:impl:model}
-
-TODO
-
-\begin{figure}%[!b]%[H]
-  \centering
-  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XEntity.png}
-  \caption{TODO}
-  \label{fig:class-diagrams:XEntity}
-\end{figure}
-
-\lstinputlisting{inc/XEventVisitor-cleaned.java}
-
-%todo: some words on necessity of contexts and lack of them. How would we implement them
-
-%todo: rename Interpreter -> Compiler.
-
-%note https://en.wikipedia.org/wiki/Semantics_(computer_science)
-%Operational semantics loosely corresponds to interpretation,
-
-- hierarchy of Compilers (XCompiler is an stateful abstract machine)
-
-- interface that it provides
-
-- dependencies on other modules (memory-manager, etc.)
-
-\subsection{Pre-compilation}
-\label{ch:impl:y2x:precomp}
-
-- collect goto labels (not done yet.)
-
-- determine kind of variables (cannot be done during parsing? don't say this)
-
-%todo: after all, list which modules (classes, managers) are being initialised during the pre-compilation. I'ts good to have the dependencies graph on moduls (operational semantics:) )
-
-%\subsection{Compilation}
-%\label{ch:impl:y2x:compil}
-
-- more low-level code representation (or high-level assembly); abstract assembly language. refer to the %2_wmm.tex
-
-- X-hierarchie
-
-%\subsection{Post-compilation}
-%\label{ch:impl:x2y:postcomp}
-
-transformations
-
-- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
-
-- converting to SSA form (now: during the encoding. should be: during the post-compilation) (as one of necessary steps before encoding)
-
-- setting up backward edges
-
-- more?
 
@@
 
-The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{(ENTRY)} (usually for convenience all leaves are connected to the sink node \texttt{(EXIT)}).
-The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).
-However, the cyclic flow graph cannot be encoded into SMT formula since ...
-//TODO:REFERENCE.%TODO
+simple algorithm
 
-\begin{figure}
+setting up backward edges
+
+how we attempted, why it didn't work in general case:
+
+why we changed the notion of unrolling bound
+
+sth about hashcodes and collections
+
+- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
+
+%TODO: draw (and implement) diff types of exit nodes
+\begin{figure}[h]
+\centering
+\resizebox{\linewidth}{!}{
 \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
@@
 \end{tikzpicture}
+}
 \label{fig:loop-unwind}
@@
 
-%\section{XGraph to ZFormula (SMT) encoder}
-%\label{ch:impl:comp:zformula}
+\subsubsection{X-graph data-flow constructor (8)}
+\label{ch:impl:proc:x-df}%TODO: ref it in pre-word in this chapter
 
-- Then, this modified event-representation is being encoded to SMT formula and sent to the solver.
+- set up co, rf edges => new graph
 
-% say something about equals() and hashCode()
+- computing SSA maps (now: during the encoding. should be: during the post-compilation) (as one of necessary steps before encoding)
+
+\subsubsection{Z-formula encoder (9)}
+\label{ch:impl:proc:z-encoder}
+
+Z-type : SMT-specific (say about smt-lib)
+
+\subsubsection{SMT-formula translator (10)}
+\label{ch:impl:proc:smt-transator}%TODO: ref in pre-word
+
+\subsubsection{SMT-formula translator (11)}
+\label{ch:impl:proc:smt-solver}%TODO: ref it in pre-word in this chapter
+%TODO: check that we have ALL references in pre-word in this chapter
+
+%todo: mention The SMT-LIB standard http://smtlib.cs.uiowa.edu/papers/smt-lib-reference-v2.6-r2017-07-18.pdf
+
+\subsection{Program output}
+\label{ch:impl:out}
+
+structure: verdict
+
+%TODO: say sth about syntax errors handling: still now, what if parser does not accept syntax?
+
+\subsection{Auxiliary components}
+\label{ch:impl:aux}
+
+\subsubsection{Watchdog timer}
+\label{ch:impl:aux:watchdog}
+
+<not implemented>
+
+todo
+
+\subsubsection{Logger}
+\label{ch:impl:aux:logger}
+
+<not implemented>
+
+todo
+
+%todo: some words on necessity of contexts and lack of them. How would we implement them
+
+%todo: rename Interpreter -> Compiler.
 

### 2018-05-04 4_impl.tex
Finish X. Set up listings.

@@
 The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
-The compiled \xgraph{} then undergoes a number of transformations (Section~\ref{ch:impl:proc:x-unroll}) necessary for the encoding into a \zformula{} (Section~\ref{ch:impl:model:zformula}) at the pre-encoding stage (Section~\ref{ch:impl:proc:x-unroll}).
-Then, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the derived relations on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and transformed \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is passed as an input to the SMT-solver.
+The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
+Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
+%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-transator}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
 
@@
 
-\begin{figure}[h]%[!hb]
+\begin{figure}[!h]%[!hb]
 \centering
-\begin{lstlisting}[mathescape=true,%
+\begin{lstlisting}[language=antlr,
+                  basicstyle=\ttfamily\tiny,
                   caption={The sketch of the input language grammar used by \oldporthos},%
-                  label={fig:in_grammar_pts},%
-                  morekeywords={if,then,else,return,while,program,thread},%
-                  breaklines=true,%
-                  basicstyle=\ttfamily\scriptsize]
+                  label={fig:in_grammar_pts}]
 <prog> : <init> <thrd>* <assert>
@@
 The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars%
-\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}).%
+\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}). %
 Currently, \porthos[2] can operate only in the intra-procedural analysis mode (analysing single procedure), assuming that each function defined in the input file is being executed in a separate thread.
@@
 %TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs. 
-Also, Current version of \porthos[2] simply ignores C processor directives, however, in future it is possible to support it.
+Also, current version of \porthos[2] simply ignores C processor directives, however, in future it can be extended to support them.
 
@@
 The abstract syntax tree \ytree{} is an abstraction level suitable for compiling the program to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
-As the ANTLR syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule \lstinline{postfixExpression '[' expression ']'}, that is converted to the \texttt{YIndexerExpression}).
+As the ANTLR syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}", that is converted to the \texttt{YIndexerExpression}).
 However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite[Chapter 5]{gries2012science}), we found this level of abstraction suitable enough for our tasks.
@@
 
-Following the C11 standard~\cite{iso2012iec}, we distinguish a \textit{statement} (\textit{"an action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
+Following the C11 standard~\cite{iso2012iec}, we distinguish a \textit{statement} (\textit{"an~action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
 
@@
 
-\vspace{0.5em}
+\vspace{1em}
 The \ytree{} expressions are the following:
@@
 
-- memory manager: no contexts so far; locals per process + shareds; register/deregister;
+The memory units are crated and stored by the stateful \texttt{XMemoryManager}, which provides interface for accessing memory units during compilation stage.
+The memory manager takes into account the higher priority of local variables over global variables specific for C language.
+The \texttt{XMemoryManager} provides interface for registering shared variables at the precompilation stage, registering local variables dynamically at the compilation stage, and returning memory unit instances for the requests by name.
+\textit{TODO: context stack}%TODO: refactor XMemoryManager!
 
@@
 
-The following interfaces capture main types of events (see also Figure~\ref{fig:class-diagrams:XEntity-interfaces}):
+X-events are the independent atomic operations.
+Each event is specified by the process generated them and a unique event label.
+This information is stored by events in the immutable structure \texttt{XEventInfo}.
 
-\begin{itemize}
-  \item \texttt{XMemoryEvent} \\
+The following interfaces model basic kinds of events (see Figure~\ref{fig:class-diagrams:XEntity-interfaces}):
+%\begin{itemize}[noitemsep]
+\begin{outline}
+  \1 \texttt{XMemoryEvent} \\
   The memory event defines the transfer of the value from one memory unit to another.
   There are four types of memory events (the arrow denotes the direction of the data-flow):
-  \begin{enumerate}
-  \item \texttt{XRegisterMemoryEvent}: \\
-    $\texttt{XLocalLvalueMemoryUnit} \leftarrow \texttt{XLocalMemoryUnit}$
-  \item \texttt{XLoadMemoryEvent}: \\
-    $\texttt{XLocalLvalueMemoryUnit} \leftarrow \texttt{XSharedMemoryUnit}$
-  \item \texttt{XStoreMemoryEvent}: \\
-    $\texttt{XSharedLvalueMemoryUnit} \leftarrow \texttt{XLocalMemoryUnit}$
-  \item \texttt{XInitialWriteEvent}: \\
-    $\texttt{XLvalueMemoryUnit} \leftarrow \texttt{XRvalueMemoryUnit}$
+  %\begin{itemize}[noitemsep]
+  \2 \texttt{XRegisterMemoryEvent}: \\
+    $(\texttt{XLocalLvalueMemoryUnit}) \leftarrow (\texttt{XLocalMemoryUnit})$,
+  \2 \texttt{XLoadMemoryEvent}: \\
+    $(\texttt{XLocalLvalueMemoryUnit}) \leftarrow (\texttt{XSharedMemoryUnit})$,
+  \2 \texttt{XStoreMemoryEvent}: \\
+    $(\texttt{XSharedLvalueMemoryUnit}) \leftarrow (\texttt{XLocalMemoryUnit})$,
+  \2 \texttt{XInitialWriteEvent}: \\
+    $(\texttt{XLvalueMemoryUnit}) \leftarrow (\texttt{XRvalueMemoryUnit})$.
 
-  \end{enumerate}%TODO incompleted enumeration
+  %\end{itemize}%TODO incompleted enumeration
   %TODO: either say st about getLoc(), getReg(), or remove them from code
 
-  \item \texttt{XComputationEvent} \\
-  - binary and unary \\
-  - implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
-  This is a model-level optimisation possible \textbf{due to} because a computation-event
-  %\textit{is} an event (thus it can form a node in \xgraph{}) and it also 
-  performs a computation over local-only memory and does not change the value of any memory unit, therefore the \textit{computation} abstraction (the CPU time) can be safely removed from the model, and the computation event can be seen as an atomic operation.
+  \1 \texttt{XComputationEvent} \\
+    We distinguish two types of computation events:
+      %\begin{itemize}[noitemsep]
+      \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation,
+      \2 \texttt{XBinaryComputationEvent} that encodes numeric operations (such as addition, multiplication, etc.), bit vector operations (such as bit-and, bit-xor, etc.), relative operations (such as greater-then comparison, equality comparison, etc.), and logical operations (such as conjunction and disjunction).
+      %\end{itemize}
+    Since a computation event performs computation over local-only memory and does not change value of any memory unit, the \textit{computation} abstraction (as the CPU time spent for the computation itself) can be safely removed from the model, and the computation event can be seen as a simultaneous operation.
+    Therefore the computation event class implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
 
-  \item \texttt{XControlFlowEvent} \\
-    \texttt{XJumpEvent} no computation, no data operation. may be safely removed for optimisation
-    \texttt{XMethodCallEvent} : TODO: meth call should inherit both control-flow and computation.
-    itself - computation result (temp register) + if (method resolved) => context binding + jump to the method body
+  \1 \texttt{XControlFlowEvent} \\
+    The control-flow event indicates a non-linear jump in the code. We distinguish two kinds of control-flow events: 
+    %\begin{itemize}[noitemsep]
+      \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
+      \2 \texttt{XMethodCallEvent} that models the method call with the \texttt{FASTCALL} calling convention (passing arguments in registers).
+      %TODO: check the diagrams if the XMCE really inherits XLMU
+      The method call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the method call.
+      If the method has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the method body.
+      Each \texttt{return} statement creates the assignment of the method call event on the top of call stack.
+      \\
+      If the method has not been resolved, \textit{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
+    %\end{itemize}
 
-  \item \texttt{XFenceEvent}
-    \texttt{XBarrierEvent} : mfence, sync, optsync, lwsync, optlwsync, ish, isb, isync
+  \1 \texttt{XFenceEvent} \\
+    The fences are implemented as an enumeration \texttt{XBarrierEvent}.
+    Current implementation of \porthos[2] supports all fences supported by \porthos[1]: 
+    \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
+    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each
 
-  \item \texttt{XFakeEvent}
+  \1 \texttt{XFakeEvent} \\
+    The fake events are auxiliary elements of \xgraph{}.
+    %\begin{itemize}[noitemsep]
+    \2 \texttt{XEntryEvent}, the per-process unique source event in the event-flow graph,
+    \2 \texttt{XExitEvent}, the process sink event. There are two kinds of sink events: \textit{complete} and \textit{incomplete} (see description in Section~\ref{ch:impl:proc:x-unroll}),
+    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}).
+   % \end{itemize}
+%\end{itemize}
+\end{outline}
+
+\paragraph{Edges}
+Edges are stored as an immutable hash-map from source to target event.
+The type of edge is stored in the enumeration \texttt{XEdgeKind}:%TODO: IMPLEMENT THIS
+\begin{itemize}[noitemsep]
+  \item control-flow edges: a \textit{primary} edge (epsilon or if-then transition) and an \textit{alternative} edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
+  \item data-flow edges: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
+\end{itemize}
+
+\paragraph{Graph invariants}
+Once being constructed, graph must conform the following requirements:
 \begin{enumerate}
-    \item \texttt{XEntryEvent}, the source event in the event-flow graph %only event that cannot have incoming edges
-    \item \texttt{XExitEvent} : two kinds (depending on the unrolling boundary)
-    \item \texttt{XNopEvent} : no-operation event, used for correct encoding (picture); same as simple jump to the next event
+\item graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
+\item graph must be connected,
+\item each node can have either one or two direct control-flow successors,
+\item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
+%\item each node except the source node must have at least one parent
+\item \textbf{todo: sth about data-flow edges}
 \end{enumerate}
 
-\end{itemize}
-
-Edges (relations edges (just map with enum as kind of edge)):
-\begin{enumerate}
-\item CF edge: then-edge (primary), else-edge (alternative)
-\item DF edge: co, rf
-\end{enumerate}
-
-Valid graph:
-\begin{enumerate}
-\item single source, two sinks
-\item connected
-\item each node has 1 or 2 children
-\item branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
-\end{enumerate}
+\paragraph{Example}
 
@@
 
-XEventInfo:
-Events are specified by the process generated them and a unique label.
-This information is stored by events as the immutable structure \texttt{XEventInfo}.
 %hashcode, uniqueness
@@
 %TODO: almost not implemented !
-<to be done>
 
@@
 
-also recursive representation of smt formula
+recursive representation of smt formula
 
-used as an abstrac
+used as an abstraction btween
 
@@
 
-<not implemented yet>
+%<not implemented yet>
 
@@
 \label{ch:impl:proc:x-unroll}
+
+two kinds of exits
 
@@
 
-<not implemented>
+%<not implemented>
 

### 2018-05-06 4_impl.tex
visitor.

@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-transator}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
 
@@
 In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
-The semantics of the methods also is being resolved during the pre-compilation stage via the \textit{invocation hooks} mechanism (see Section~\ref{ch:impl:proc:x-pre-compiler:hooks}).
+The semantics of the methods also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-pre-compiler:hooks}).
 
@@
 Also, only integers were processed by the input language parser of \porthos[1].
-In \porthos[2], we extended support for primitive types supported by the \texttt{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s, floats encoded as \texttt{Real}s, enums encoded as Z3 \texttt{Scalars}).
+In \porthos[2], we extended support for primitive types supported by the Z3 solver (this apply to 32-bit integers encoded as \texttt{Int}s, floats encoded as \texttt{Real}s, enums encoded as Z3 \texttt{Scalars}).
 %TODO : constant arrays? https://rise4fun.com/Z3/tutorial/guide
@@
 
-As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
+%As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
 For keeping the architecture transparent, we build all abstraction levels with interfaces, even if some of them does not add any new functionality.% (as java does not support multiple inheritance for classes (only for interfaces), we
@@
 
-in \porthos[2] we created the new abstraction layer \zformula that is translated to
+in \porthos[2] we created the new abstraction layer \zformula[] that is translated to
 
@@
 
-- numbers -- same as in gears in Figure~\ref{fig:arch}.
+This section describes program units that construct, transform and analyse internal representations described before.
+% steps of construction of the program representations and transformations applied to them.
+Further in this Section, the subsection titles have a number in parentheses, which correspond to a unique number of the  processing unit being discussed.
+This numbers are pictured within gears on the top-left corner of each processing unit in Figure~\ref{fig:arch}.
+%he numbers in parentheses correspond the numbers of processing units pictured within gears in Figure~\ref{fig:arch}.
 
-\subsubsection{Input memory model parser (1)}
-\label{ch:impl:proc:inp-mod-parser}
+All recursive data structures are processed by units that implement the \textit{visitor} pattern~\cite{palsberg1998essence}.
+This is a behavioural pattern that separates the program logic from the object implementation by specifying handling methods for each element of the object.
+The visitor pattern performs the double-dispatching mechanism: the visiting class defines the method for accepting a visitor, and the visitor itself defines the 
+To visit an instance \lstinline{x} by a visitor \lstinline{v}, the program should call the accept-method of \lstinline{x} and pass \lstinline{v} as an argument.
+In our implementation, the visitor must carry out the continuation of traversing the recursive data structure itself.
+The general structure of the visitor pattern is illustrated by t9he pseudo-java code in Figure~\ref{fig:visitor}.
 
-ANTLR grammar is extracted from parser used by herd written in OCaml
+\begin{figure}[h]
+\centering
+\begin{multicols}{2}
+\begin{lstlisting}[language=java]
+interface Element {
+  ...
+  <T> T accept(Visitor<T> visitor);
+}
 
-\subsubsection{Input program parser (2)}
-\label{ch:impl:proc:inp-prog-parser}
+class AnElement implements Element {
+  ...
+  @Override
+  <T> T accept(Visitor<T> visitor) {
+    return visitor.visit(this);
+  }
+}
+[* \vfill\null \columnbreak *]
+class Visitor<T> {
+  ...
+
+  T visit(AnElement e) {
+    // some visiting logic
+    ...
+    // continue recursively
+    e.getChild().accept(this);
+  }
+
+  T visit(AnOtherElement e) {
+    ...
+  }
+}
+\end{lstlisting}
+\end{multicols}
+\caption{Sketch of the visitor pattern}
+\label{fig:visitor}
+\end{figure}
+
+We consider the visitor pattern as the most natural way for operating the hierarchical data structures such as AST.
+However, we use its double-dispatching capabilities to reduce cost of multiple type casting performed while traversing the elements of a non-recursive data structure.
+The operator instance, instead of having a single method that handles and element of the instance, it extends the visitor interface and splits the handler method into multiple methods, one for each type of element.
+
+For traversing plain (non-recursive) complex data structures, we mostly follow the \texttt{iterator} pattern: the object that 
+the special object Iterator, that has access to elements of the instance, iterates over them and .
+Since in \porthos[] we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
+However, in contrast to the visitor of a recursive data structure, the visitor of a plain data structure should be used with an \textit{iterator} that has access to the elements of the object and passes them to the visitor.
+
 % todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
 
-%TODO: maybe merge this and previous subsubsections
+\subsubsection{Input parsers (1,2)}
+\label{ch:impl:proc:input-parser}
 
-todo: move something from the input language chapter here
+Both input-language and input-model parsers are implemented via ANTLR parser generator.
+Details on the input program language grammar are discussed in Section~\ref{ch:impl:input}.
+The input memory model language \cat is discussed in Section~\ref{ch:wmm:cat}.
+The ANTLR grammar for the \cat language was extracted from the parser used by \tool{herd} tool%
+\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
+written in OCaml.
+%TODO: upload the grammar to github
+%The input memory model
+
+\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}
+
+%todo: maybe: move something from the input language chapter here
 
@@
 
-%The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{(ENTRY)} (usually for convenience all leaves are connected to the sink node \texttt{(EXIT)}).
-The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.).
-However, the cyclic flow graph cannot be encoded into SMT formula since ...
-%//TODO:REFERENCE.%TODO
-
-%note https://en.wikipedia.org/wiki/Semantics_(computer_science)
-%Operational semantics loosely corresponds to interpretation,
-
 \paragraph{The macro preprocessor}
 
-typedefs, aliases
+The preprocessor traverses the \ytree{} and collects information necessary for compiling it into an \xgraph{}.
+
+typedefs
+
+aliases
+
+more macros?
 
@@
 - pointers
+
 - externs
-- parameters of process-functions
+
+- parameters of process-functions (parameters of other funcs are local)
 
@@
 - The typification algorithm: propagation rules for each of Y
+
+- size\_t. ...%https://en.wikipedia.org/wiki/C_data_types#stddef.h
 
@@
 
-\subsubsection{SMT-formula translator (10)}
-\label{ch:impl:proc:smt-transator}%TODO: ref in pre-word
+\subsubsection{SMT-formula converter (10)}
+\label{ch:impl:proc:smt-converter}%TODO: ref in pre-word
 

### 2018-05-06 4_impl.tex
add up to preproc.

@@
     \begin{enumerate}[label*=\arabic*.]
+      \item prohibit under-approximating abstractions;
+      \item preservation the completeness of analysis;
+      \item modular architecture: each module can be tested independently;
+      \item usage of software design patterns where necessary;
       \item usage of immutable data structures for all DTOs;
-      \item interrupting the work on errors;
-      \item modular architecture: each module can be tested independently;
-      \item usage of software design patterns if necessary;
     \end{enumerate}
@@
 
-As \porthos[1], the \porthos[2] uses the open-source SMT solver \texttt{Z3} from Microsoft Research~\cite{de2008z3}.
+The robustness of analysis is the main criterion of \porthos[2] as a verification of a bug-hunting tool.
+This implies modelling the system with only over-approximating abstractions, which, however makes it more vulnerable to the combinatorial growth of the state space size, preserve the completeness of analysis.
+As a robust and transparent tool, the \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
+%The modular architecture of \porthos[2] allow each its module to be tested independently.
+%Usage of immutable data structures for all DTOs helps 
+%immut: preservation of invariants
+%The transparency, efficiency and extensibility are the necessary criteria of any good tool.
+
+As its predecessor, \porthos[2] uses the open-source SMT solver \tool{Z3} by Microsoft Research~\cite{de2008z3}.
 However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
@@
 See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
-that is easily convertable to a SMT type (modelled as \texttt{ZType}).%TODO: add the class ZType
+that is easily convertable to an SMT type (modelled as \texttt{ZType}).%TODO: add the class ZType
 
@@
       %\end{itemize}
-    Since a computation event performs computation over local-only memory and does not change value of any memory unit, the \textit{computation} abstraction (as the CPU time spent for the computation itself) can be safely removed from the model, and the computation event can be seen as a simultaneous operation.
-    Therefore the computation event class implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
+    \newline
+    The computation event class implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
+    This is a model-level optimisation, which is possible because a computation event performs computation over local-only memory and does not change value of any memory unit.
+    Thus, the \textit{computation} abstraction (as the CPU time spent for the computation itself) can be safely removed from the model, and the computation event can be seen as a zero-time operation that produces the \textit{value}.
+    For analysing large expressions this optimisation is sensible because it considers the whole expression as a single computation event, encoded therefore as a single SMT-variable.
+    \newline
+    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to argument `\lstinline{x}' and `\lstinline{y}').
 
@@
       \\
-      If the method has not been resolved, \textit{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
+      If the method has not been resolved, \textbf{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
     %\end{itemize}
@@
 \centering
-\begin{multicols}{2}
+%\begin{multicols*}{2}
+\begin{minipage}[t]{.55\textwidth}
 \begin{lstlisting}[language=java]
 interface Element {
+  <T> T accept(Visitor<T> visitor);
   ...
-  <T> T accept(Visitor<T> visitor);
 }
@@
 class AnElement implements Element {
-  ...
   @Override
@@
   }
+  ...
 }
-[* \vfill\null \columnbreak *]
+\end{lstlisting}
+%[* \vfill\null \columnbreak *]
+\end{minipage}% <---------------- Note the use of "%"
+\begin{minipage}[t]{.45\textwidth}
+\begin{lstlisting}[language=java]
 class Visitor<T> {
-  ...
-
   T visit(AnElement e) {
-    // some visiting logic
+    // visiting logic
     ...
@@
   }
-
   T visit(AnOtherElement e) {
@@
   }
+  ...
 }
 \end{lstlisting}
-\end{multicols}
-\caption{Sketch of the visitor pattern}
+%\end{multicols*}
+\end{minipage}
+\caption{Illustration of the visitor pattern}
 \label{fig:visitor}
@@
 Both input-language and input-model parsers are implemented via ANTLR parser generator.
+
 Details on the input program language grammar are discussed in Section~\ref{ch:impl:input}.
+Currently, \porthos[2] does not consider preprocessor instructions (it ignores them).
+For implementing the inter-procedural mode of \porthos{}, it is crucial to support inclusion of header files (the \texttt{\#include} preprocessor directive).
+Next step would be implementing the support of macros and conditional compilation directives, that are used often in C code.
+As preprocessor statements may appear in arbitrary place of a program, the preprocessor must be a stateful processing unit that reads the token stream and dynamically instrument the program by interpreting directives and expanding macros.
+%The ANTLR grammar for the C preprocessor may be found 
+%more macros (stateful) : https://gcc.gnu.org/onlinedocs/cpp/Macros.html
+
 The input memory model language \cat is discussed in Section~\ref{ch:wmm:cat}.
@@
 %The input memory model 
-
 \textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}
@@
 
+%<not implemented yet>
+
 visitor of antlr tree: almost direct
 
-todo: grammar of w-model to appendix
+todo: file tree of w-model to appendix
 
 in perspective: work with functional-style definitions
-
-%<not implemented yet>
 
@@
 
-\paragraph{The macro preprocessor}
+The precompiler traverses the \ytree{} and collects information necessary for compiling it into an \xgraph{}.
 
-The preprocessor traverses the \ytree{} and collects information necessary for compiling it into an \xgraph{}.
+%TODO: not supported yet
 
-typedefs
+\paragraph{The label resolution}
+\label{ch:impl:proc:x-pre-compiler:label}
 
-aliases
+%TODO: <not implemented yet>
 
-more macros?
+The label resolution is necessary for establishing links to labelled statements. %a process of linkage the text labels referenced in code to the places of their declaration.
+In C, labelled statements are declared via the colon-syntax `\texttt{<label> : <statement>}', and the labels are referenced by the jump-statement `\texttt{goto <label>}'.
+The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labeled statement.
+%TODO: code: remove temp label generation, it's stupid.
+put null.
+This information is used during compilation to set up unconditional jumps.
 
-not supported yet
+\paragraph{Type analysis}
+\label{ch:impl:proc:x-pre-compiler:type}
+
+The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
+Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
+% according the rules of propagation.
+For example,
+Being carried at Y and X representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
+Currently, \porthos[2] operates only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
+Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
+In addition, \porthos[2] supports ... int\_t ....
+
+   %TODO: support this.
+
+%TODO: describe the XType
+%TODO: definitely get rid of bit32
+
+%TODO: <not implemented>
+The type analysis algorithm should consider the type aliases supported by C language (defined by the \texttt{typedef} instruction).
+For resolving the type aliases, the precompiler should make an extra traverse of the \ytree{} before the precompilation stage  and build up a symbol map.
+
+%TODO: <not implemented>
+The type analysis includes the function resolution: for correct resolving the semantics of functions, the compiler needs a mapping of each function invocation to its signature.
+However due to lack of polymorphism and function overloading mechanism inherent to OOP languages the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.
+
+- X-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+
+- The typification algorithm: propagation rules for each of Y
+
+- size\_t. ...%https://en.wikipedia.org/wiki/C_data_types#stddef.h
 
@@
 \label{ch:impl:proc:x-pre-compiler:var}
+
+As it was discussed in Section~\ref{ch:impl:model:ytree}, the \ytree{} encodes all named objects as \texttt{YVariableRef}.
 
@@
 
-\paragraph{The label resolution}
-\label{ch:impl:proc:x-pre-compiler:label}
-
-%TODO: <not implemented yet>
-
-The label resolution is a process of linkage the referenced labels to declared labels.
-
-In C, the labeled statements are declared via the colon-syntax `\texttt{<label> : <statement>}',
-and the labels are referenced by the jump-statement `\texttt{goto <label>}'.
-The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labeled statement.
-%TODO: code: remove temp label generation, it's stupid.
-put null.
-This information is used during compilation (see Section~\ref{ch:impl:proc:x-compiler}).
-
-\paragraph{Type analysis}% and function resolution.}
-\label{ch:impl:proc:x-pre-compiler:type}
-
-%TODO: remake this paragraph considering that we have already said sth about XType
-The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
-Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
-%In \porthos[2], the type is modelled by the class \texttt{XType}.%TODO: + fix class and package name!
-Being carried at Y and X representation levels, the type is converted to a Z-type at the Z-formula encoding stage (see Section~\ref{ch:impl:proc:z-encoder}).
-
-%TODO: describe the Type
-%TODO: maybe rename back to XType
-%TODO: but definitely get rid of bit32
-
-- X-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
-
-- The typification algorithm: propagation rules for each of Y
-
-- size\_t. ...%https://en.wikipedia.org/wiki/C_data_types#stddef.h
-
 \paragraph{Invocation hooking}
@@
 resolve the semantics of a function given its signature (meth name + params types)
+
+mapping
 

### 2018-05-07 1_intro.tex
Add some comment ideas. Cleanup a bit.

@@
 
-Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
+% PROBLEM STATEMENT: MORE CLEAR: 
+% 1) Definition of research scope and goals 
+% 2) "- References throw light on the topic from a variety of perspectives"
+%    "-The thesis combines the cited  works and empirical data consistently and clearly"
+%    "- Thesis demonstrates critIcal evaluation of existing knowledge"
+% 3) Research Methods
+%    "- Appropriate methods have been used in a well-founded manner. Reliability and validity or  trustworthiness of the study/results are evaluated."
+% 4) Results and contribution 
+%    results + justification
+%    Goals and results must match
+%    the results must be based on independent research , not on the references.
+%    "significance for the field of research or industry in question, and contribution to knowledge"
+%    "Cited works must be evaluated critically"
+%    "Conclusions SHOULD NOT BE drawn JUST appropriately from the material"
+
+Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
 %interprocedural analysis

### 2018-05-07 4_impl.tex
Add some comment ideas. Cleanup a bit.

@@
 
-On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as an alias for a type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information of the type for declarations
+On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as a \textit{reference} for the type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information on declaration, including the type itself, type modifiers and qualifiers).
 
@@
 The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
-Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.
+\ Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.
 
@@
     \newline
-    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to argument `\lstinline{x}' and `\lstinline{y}').
+    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to the arguments `\lstinline{x}' and `\lstinline{y}').
 
@@
 \begin{itemize}[noitemsep]
-  \item control-flow edges: a \textit{primary} edge (epsilon or if-then transition) and an \textit{alternative} edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
-  \item data-flow edges: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
+  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
+  \item the \textit{data-flow edges}: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
 \end{itemize}
@@
 \end{enumerate}
+
+\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
+%TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
+todo
 
@@
 
-The precompiler traverses the \ytree{} and collects information necessary for compiling it into an \xgraph{}.
+The precompiler traverses the \ytree{} and collects information necessary for its compilation into an \xgraph{}.
 
@@
 Being carried at Y and X representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
+
+%TODO: YOU REALLY NEED TO THINK MORE ABOUT TYPES ON Y, X, Z LEVELS + reconsider this paragraph (and others about types)
 Currently, \porthos[2] operates only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
 Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
-In addition, \porthos[2] supports ... int\_t ....
-
-   %TODO: support this.
-
-%TODO: describe the XType
-%TODO: definitely get rid of bit32
+In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter}, %TODO: Or XTypeHookManager?
+which can be requested for converting an X-type to a Z-type. %TODO: implement this.
+%TODO: describe the XType (before) + implement this clearly! + definitely get rid of bit32
 
@@
 %TODO: <not implemented>
-The type analysis includes the function resolution: for correct resolving the semantics of functions, the compiler needs a mapping of each function invocation to its signature.
-However due to lack of polymorphism and function overloading mechanism inherent to OOP languages the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.
+The type analysis includes the process of resolving semantics of function invocations.
+The result is stored in a map of function invocation expression to its resolved signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
+However, due to lack of polymorphism and function overloading mechanism inherent to OOP languages, the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.
 
-- X-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+==
 
+The X-type construction: //code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+- on Z level, type is one-pointer-level
 - The typification algorithm: propagation rules for each of Y
+- as vector array as needed.
 
-- size\_t. ...%https://en.wikipedia.org/wiki/C_data_types#stddef.h
+Y: quantifiers+specifiers+typeName
+X: ?.. Z? TODO: rename+rework XType->XZType
+Z: Int, Real, Boolean, BitVector[N]
 
@@
 
-As it was discussed in Section~\ref{ch:impl:model:ytree}, the \ytree{} encodes all named objects as \texttt{YVariableRef}.
+On the compilation stage, once the compiler meets the reference to a variable, it should know whether it is local or global variable.
+In general case,
 
@@
 
-\paragraph{Invocation hooking}
-\label{ch:impl:proc:x-pre-compiler:hooks}
+- EXPORT\_SYMBOL "exports the symbol for dynamic linking"
 
-resolve the semantics of a function given its signature (meth name + params types)
+-EXPORT\_SYMBOL\_GPL %https://lwn.net/Articles/154602/
 
-mapping
-
-knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
+ %https://stackoverflow.com/questions/6670589/how-to-prevent-error-symbol-undeclared-here-despite-export-symbol-in-a-linu
 
@@
 \label{ch:impl:proc:x-compiler}
-
-- hierarchy of Compilers (XCompiler is an stateful abstract machine)
 
@@
 
+converting recursive structure to plain
+
+stacks: context, readyContext, call, 
+  BlockContext (rename): what stores, enum of states
+
+almostReadyContexts: for computation events. Flushed by invocation . When are filled.
+
+setting up edges depending of type of ready-context
+
+stateful. Managers, [graph] builders
+  interfaces of XMemoryManager: listing, paragraph
+
+- hierarchy of Compilers: pic on inheritance (XCompiler is an stateful abstract machine). Restrictions of lude-compilers
+  postlude: assertion: a new event XAssertionEvent
+
+interfaces that offer compilers: listing of interface
+
+algorithm for interpreting
+
 %MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
 %reutrn event= assign the return-register + simple jump
+
+\paragraph{Invocation hooking}
+\label{ch:impl:proc:x-pre-compiler:hooks}
+
+resolve the semantics of a function given its signature (meth name + params types)
+
+mapping - as a result
+
+knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
+
+%TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html
+
+% + RCU %https://www.kernel.org/doc/Documentation/RCU/whatisRCU.txt
 

### 2018-05-07 1_intro.tex
Consider Professor's comments.

@@
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
-For instance, Figure~\ref{simple_wmm_x86} provides simple example of reachability of the state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
+For instance, Figure~\ref{simple_wmm_x86} provides simple example of reachability of the state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' on the x86 architecture (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
 %reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
-This state is allowed because in x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not become visible by other processes immediately.
+This state is allowed because in the x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not immediately become visible by processes running on other cores.
 In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
-Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in another thread.
-These problems have lead to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{weak	memory models (WMM)}.
+Meanwhile, the write cache of the process \texttt{P1} may not have an updated version of the variable~\texttt{[x]}, neither the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in the other thread.
+These problems have lead to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{weak memory models (WMMs)}.
 
@@
 Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
-Furthermore, there is a wide range of tools that perform program verification wrt memory models (see~\cite{alglave2013software}, \cite{Porthos17}).
+Furthermore, there is a wide range of tools that perform program verification wrt. memory models (see~\cite{alglave2013software}, \cite{Porthos17SAS}).
 
@@
 Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17}.
+As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17SAS}.
 
@@
 \footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}%
-introduced in April~2017~\cite{Porthos17}.
+introduced in April~2017~\cite{Porthos17CoRR}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
@@
 
-%In the work~\cite{Porthos17}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
+%In the work~\cite{Porthos17CoRR}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
 Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a trivial problem of symbolic execution, encoding of the weak memory model is more tedious.

### 2018-05-07 3_encod.tex
Change [H] of some pictures.
Consider Professor's comments.

@@
 % The first tool that tackles the problem 
-%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17}  
+%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17SAS}  
 In this Chapter, we describe the general task of analysing the concurrent software portability
 %may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17} (more precisely, to an SMT problem).
+as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17SAS} (more precisely, to an SMT problem).
 
@@
 Consider the function $cons_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
-Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17}:
+Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17SAS}:
 
@@
 
-Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17}.
+Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17CoRR}.
 The result SMT formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
@@
 %The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. 
-%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.
+%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17SAS}.
 
@@
 $\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
-and control-flow of the branching instruction $i_1 = (c \, ? \, i_2 \, : \, i_3)$ was encoded as
-$\phi_{CF}( \, c \, ? \, i_2 \, : \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
-(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\,x\,\mathtt{then}\,y\,\mathtt{else}\,z$').
+and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
+$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
+(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
 In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
@@
 Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \rf-relation (see Section~\ref{ch:wmm:model:relations} for details on relations).
-We distinguish the function ${\fv_p\,:\,\mathbb{C_p}\,\rightarrow\,\{0,1\}}$ that evaluates the predicative computation event.
+We distinguish the function ${\fv_p\;:\;\mathbb{C_p}\;\rightarrow\;\{0,1\}}$ that evaluates the predicative computation event.
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
@@
 
-\begin{figure}[H]
+\begin{figure}[!hb]
     \centering
@@
 Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
-Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \, (e_1), \, (e_1~\rightarrow~e_2), \, (e_1 ~\rightarrow~e_3)\}$.
+Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1~\rightarrow~e_2), \; (e_1 ~\rightarrow~e_3)\}$.
 In encoding~\ref{enc:cf:merge} of the merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 In contrast, \porthos{} used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
-Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17}.
+Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17SAS}.
 
-Following~\cite{Porthos17}, the indexed references of variables are computed in accordance with the following rules:
+Following~\cite{Porthos17CoRR}, the indexed references of variables are computed in accordance with the following rules:
 (1)~any access to a shared variable (both read and write) increments its SSA-index;
@@
 %
-%As it is described in~\cite{Porthos17}, the is based . . .
-%~\cite{Porthos17}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"
+%As it is described in~\cite{Porthos17SAS}, the is based . . .
+%~\cite{Porthos17SAS}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"

### 2018-05-07 4_impl.tex
Consider Professor's comments.

@@
 The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a rather high-level AST.
-The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only).
+The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only, which we tend to retain clearly structured).
 
@@
 
-  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}-loops;%TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
+  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}- loops;%TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
 
@@
 
-The \ytree{} is compiled into the low-level event-based program representation \textit{\xgraph{}}.
+The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
 The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.

### 2018-05-07 99_summary.tex
Consider Professor's comments.

\chapter{Summary}
\label{ch:summary}

enumerate goals in past tense + justification

contribution

### 2018-05-07 1_intro.tex
add a bit up to hooks.

@@
 Unfortunately, when it comes to the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
-The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible.
+The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible~\cite{adve1996shared}.
 For instance, Figure~\ref{simple_wmm_x86} provides simple example of reachability of the state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' on the x86 architecture (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
@@
 Research of weak memory models firstly aims to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
-The first (and so far the only) such a framework was presented in 2010~\cite{alglave2010shared}.
+The first (and so far the only) such a framework was formalised by Jade Alglave in 2010~\cite{alglave2010shared}.
 In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.

### 2018-05-07 3_encod.tex
add a bit up to hooks.

@@
 
-%TODO: more about SMT (= SAT + constraints expressed in background theories. Exapmles of theories that we use and that are implem-ted in Z3).
+%TODO: more about SMT-(= SAT + constraints expressed in background theories. Exapmles of theories that we use and that are implem-ted in Z3).
 
@@
 %may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17SAS} (more precisely, to an SMT problem).
+as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17SAS} (more precisely, to an SMT-problem).
 
@@
 Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17CoRR}.
-The result SMT formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
+The result SMT-formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
 If the formula is satisfiable, there exist a portability bug.

### 2018-05-07 4_impl.tex
add a bit up to hooks.

@@
 
-As its predecessor, \porthos[2] uses the open-source SMT solver \tool{Z3} by Microsoft Research~\cite{de2008z3}.
-However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT solver.
+As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} by Microsoft Research~\cite{de2008z3}.
+However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver.
 %TODO: *should* allow?
@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT logical formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
 
@@
 
-Secondly, the semantics of operations was defined syntactically, whereas processing programs written in most modern languages (including C) requires the semantics resolution based on typification%
-\footnote{for instance, given two functions
-`\lstinline{int foo(int a)}' and `\lstinline{int foo(char a)}', the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
+Secondly, in \porthos[1], the semantics of operations was defined syntactically (in ANTLR grammar), whereas this should be performed by a separate module operating on the AST level%
+\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
+, so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.
+
+%\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
 As the reader can notice from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, the memory operations of different kinds vary syntactically.
-For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the method name.
+For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the function name.
 In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
-The semantics of the methods also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-pre-compiler:hooks}).
+The semantics of the functions also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler:hooking}).
 
@@
 %TODO : constant arrays? https://rise4fun.com/Z3/tutorial/guide
-Thus, adding support for more advanced syntactic structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
+Although Z3 supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of static analysis of pointers and non-constant sized arrays moves the full support of arrays and pointers out of the scope of current thesis.
+%Thus, adding support for more advanced structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
 
-The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1\,+\,2\,*\,3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
 Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
@@
 The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars%
-\footnote{Repository path: \url{https://github.com/antlr/grammars-v4}}). %
-Currently, \porthos[2] can operate only in the intra-procedural analysis mode (analysing single procedure), assuming that each function defined in the input file is being executed in a separate thread.
+\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}).%
+Current version of \porthos[2] does not recognise C processor directives (it ignores them), however, in future it can be extended to support them.
+
+Currently, \porthos[2] can operate only in the intra-procedural analysis mode, assuming that each function defined in the input file is being executed in a separate thread.
 %TODO: describe more intra- and inter-procedural!
-However, the redesigned architecture of \porthos[2] allow to easily support inter-procedural (cross-procedure) analysis by inlining function calls and binding variable contexts.
+However, the redesigned architecture of \porthos[2] allow to easily support the inter-procedural (cross-procedure) analysis by inlining function calls and binding variable contexts.
+Thus, instead of analysing a single code file, the user can specify the whole code project.
+The functions to be executed in parallel can be specified by the user, however, the some concurrent parts of code can be detected automatically at the compilation stage while the semantics of function invocations is being resolved (for instance, the C threads operations from \texttt{thread.h} or \texttt{pthread.h} may be supported).
 %TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs.
-Also, current version of \porthos[2] simply ignores C processor directives, however, in future it can be extended to support them.
 
@@
 
-  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT solver we use supports the constant-array theory, we can maintain the array type);
+  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT-solver we use supports the constant-array theory, we can maintain the array type);
 
@@
 See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
-that is easily convertable to an SMT type (modelled as \texttt{ZType}).%TODO: add the class ZType
+that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType
 
-The memory units are crated and stored by the stateful \texttt{XMemoryManager}, which provides interface for accessing memory units during compilation stage.
-The memory manager takes into account the higher priority of local variables over global variables specific for C language.
-The \texttt{XMemoryManager} provides interface for registering shared variables at the precompilation stage, registering local variables dynamically at the compilation stage, and returning memory unit instances for the requests by name.
-\textit{TODO: context stack}%TODO: refactor XMemoryManager!
+The memory units are created and stored by the \texttt{XMemoryManager}, which provides interface for accessing memory units during compilation stage.
+For more detailed description of memory management see Section~\ref{ch:impl:proc:x-compiler:mem}.
 
@@
     \newline
-    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to the arguments `\lstinline{x}' and `\lstinline{y}').
+    Note, for the purpose of simplifying the X abstraction level, computation events may have been modelled as invocations of bodiless functions (for instance, the operation `\lstinline{x + y}' may be modelled as the invocation of the function `\lstinline{+}' with the arguments `\lstinline{x}' and `\lstinline{y}').
+    However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT-solver.
 
@@
       \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
-      \2 \texttt{XMethodCallEvent} that models the method call with the \texttt{FASTCALL} calling convention (passing arguments in registers).
+      \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
       %TODO: check the diagrams if the XMCE really inherits XLMU
-      The method call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the method call.
-      If the method has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the method body.
-      Each \texttt{return} statement creates the assignment of the method call event on the top of call stack.
+      The function call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the function call.
+      If the function has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the function body.
+      Each \texttt{return} statement creates the assignment of the function call event on the top of call stack.
       \\
-      If the method has not been resolved, \textbf{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
+      If the function has not been resolved, \textbf{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
     %\end{itemize}
@@
 \paragraph{Edges}
+\label{ch:impl:model:xgraph:edges}
 Edges are stored as an immutable hash-map from source to target event.
@@
 \paragraph{Graph invariants}
-Once being constructed, graph must conform the following requirements:
+\label{ch:impl:model:xgraph:invariants}
+Once being constructed, the graph must conform the following requirements:
+
 \begin{enumerate}
@@
 \paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
+\label{ch:impl:model:xgraph:builder}
 %TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
-todo
 
-\paragraph{Example}
+\textbf{todo}
 
-consider the following code in C:
+\paragraph{Graph example}
+\label{ch:impl:model:xgraph:example}
+\textbf{consider the following code in C:}
 
-picture of the CF-graph:
+\textbf{picture of the CF-graph:}
 
@@
 
-recursive representation of smt formula
+recursive representation of SMT-formula
 
@@
 However, we use its double-dispatching capabilities to reduce cost of multiple type casting performed while traversing the elements of a non-recursive data structure.
-The operator instance, instead of having a single method that handles and element of the instance, it extends the visitor interface and splits the handler method into multiple methods, one for each type of element.
+The operator instance, inch:impl:model:xgraph:memstead of having a single method that handles and element of the instance, it extends the visitor interface and splits the handler method into multiple methods, one for each type of element.
 
@@
 
-\subsubsection{X-graph precompiler (5)}
+\subsubsection{X-graph pre-compiler (5)}
 \label{ch:impl:proc:x-pre-compiler}
@@
 The label resolution is necessary for establishing links to labelled statements. %a process of linkage the text labels referenced in code to the places of their declaration.
-In C, labelled statements are declared via the colon-syntax `\texttt{<label> : <statement>}', and the labels are referenced by the jump-statement `\texttt{goto <label>}'.
+In C, labelled statements are declared via the colon-syntax `\texttt{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\texttt{goto\;<label>}'.
 The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labeled statement.
@@
 For example,
-Being carried at Y and X representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
+Being carried at Y- and X- representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
 
 %TODO: YOU REALLY NEED TO THINK MORE ABOUT TYPES ON Y, X, Z LEVELS + reconsider this paragraph (and others about types)
-Currently, \porthos[2] operates only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
+Currently, \porthos[2] handles only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
 Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
@@
 The type analysis algorithm should consider the type aliases supported by C language (defined by the \texttt{typedef} instruction).
-For resolving the type aliases, the precompiler should make an extra traverse of the \ytree{} before the precompilation stage and build up a symbol map.
+For resolving the type aliases, the precompiler should make an extra traverse of the \ytree{} before the pre-compilation stage and build up a symbol map.
 
 %TODO: <not implemented>
-The type analysis includes the process of resolving semantics of function invocations.
-The result is stored in a map of function invocation expression to its resolved signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
+The type analysis also includes the process of resolving semantics of function invocations.
+The result is stored in a map of invocation expression to the resolved function signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
 However, due to lack of polymorphism and function overloading mechanism inherent to OOP languages, the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.
 
-==
+%TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
+%The X-type construction: \textbf{TODO} //code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
+%- on Z level, type is one-pointer-level
+%- The typification algorithm: propagation rules for each of Y
+%- as vector array as needed.
+%%
+%Y: quantifiers+specifiers+typeName
+%X: ?.. Z? TODO: rename+rework XType->XZType
+%Z: Int, Real, Boolean, BitVector[N]
 
-The X-type construction: //code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
-- on Z level, type is one-pointer-level
-- The typification algorithm: propagation rules for each of Y
-- as vector array as needed.
-
-Y: quantifiers+specifiers+typeName
-X: ?.. Z? TODO: rename+rework XType->XZType
-Z: Int, Real, Boolean, BitVector[N]
-
-\paragraph{Variables kinds determination}
+\paragraph{Variable kind analysis}
 \label{ch:impl:proc:x-pre-compiler:var}
 
-On the compilation stage, once the compiler meets the reference to a variable, it should know whether it is local or global variable.
-In general case,
+On the compilation stage, once the compiler meets the reference to a variable, it should know whether it refers to a local or global variable.
+The kinds of variables have to be determined on the pre-compilation stage.
 
-- pointers
-
-- externs
-
-- parameters of process-functions (parameters of other funcs are local)
-
-- EXPORT\_SYMBOL "exports the symbol for dynamic linking"
-
--EXPORT\_SYMBOL\_GPL %https://lwn.net/Articles/154602/
-
- %https://stackoverflow.com/questions/6670589/how-to-prevent-error-symbol-undeclared-here-despite-export-symbol-in-a-linu
+\vspace{1em}
+The following types of variables are detected as \textit{global}:
+%TODO: not implemented
+\begin{itemize}
+\item a variable was declared as a pointer;
+\item a variable whose address was accessed by any process;
+\item a variable that was declared as a parameter of the process function;
+%TODO: say sth about KERNEL ANALYSIS MODE BEFORE !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+\item a variable that was exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking); %https://lwn.net/Articles/154602/
+\end{itemize}
 
@@
 
-Firstly, the \ytree{} is compiled to the \textit{cyclic control-flow event-based graph} \xgraph[CF].
+The X-graph compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
 
-Then: unrolled up to bound $k$ to the \textit{acyclic control-flow event-based graph} \xgraphU[CF].
+The compilation is performed in three stages.
+Firstly, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
+Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
+After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be CF-acyclic (no cycles among control-flow edges).
 
-Then: data-flow analysis to the \textit{acyclic full event-based graph} \xgraphU[CF+DF].
+<COmpiler contains several modules: (todo: remove todo about renaming interpreter to compiler: it's indeed interpreter>
 
-converting recursive structure to plain
+<scheme of who calls whom>
+
+%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
+%reutrn event= assign the return-register + simple jump
+
+\paragraph{Memory manager}
+\label{ch:impl:proc:x-compiler:mem}
+
+The \xgraph{} abstract machine model disposes infinite number of memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
+The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
+However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
+At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
+Following the C standard, local memory units have higher priority over global ones.
+Since the C language allows the variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the block-context stack. %TODO: refactor XMemoryManager!
+
+<some picture ? >
+
+\paragraph{Invocation hook manager}
+\label{ch:impl:proc:x-compiler:hooking}
+
+The invocation hooking module serves as a knowledge base that stores the semantics of functions.
+%TODO: rename HookManager -> XHookManager
+Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in case of program in C language -- only the function name) across all signatures it stores.
+Once the signature matches, the hook manager intercepts the compilation process and interprets the method according to the algorithm it stores.
+%TODO: say sth about the mode
+The hook manager operates multiple invocation hooks depending on the user-defined mode (
+ \texttt{XInvocationHook}
+
+%TODO: rename InvocationHook -> XInvocationHook
+multiple hooks for mult modes
+%, having access to the X-interpreter (see Section~\ref{ch:impl:proc:x-compiler:compilation}),
+
+lambda-hook: knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
+
+todo: kernel hooks
+%TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html
+
+% + RCU %https://www.kernel.org/doc/Documentation/RCU/whatisRCU.txt
+
+\paragraph{Interpreter}
+\label{ch:impl:proc:x-compiler:compilation}
+
+stateful
 
@@
 
-stateful. Managers, [graph] builders
-  interfaces of XMemoryManager: listing, paragraph
-
 - hierarchy of Compilers: pic on inheritance (XCompiler is an stateful abstract machine). Restrictions of lude-compilers
@@
 
-%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
-%reutrn event= assign the return-register + simple jump
-
-\paragraph{Invocation hooking}
-\label{ch:impl:proc:x-pre-compiler:hooks}
-
-resolve the semantics of a function given its signature (meth name + params types)
-
-mapping - as a result
-
-knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
-
-%TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html
-
-% + RCU %https://www.kernel.org/doc/Documentation/RCU/whatisRCU.txt
-
 \subsubsection{X-graph unroller (7)}
 \label{ch:impl:proc:x-unroll}
+
+up to bound $k$
+
+recursion -- may be, if calls are supported
 

### 2018-05-08 1_intro.tex
add some refs to the intro. cleanup a bit.

@@
 Research of weak memory models firstly aims to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
-The first (and so far the only) such a framework was formalised by Jade Alglave in 2010~\cite{alglave2010shared}.
+The first (and so far the only) such a framework was formalised by J.
+Alglave in 2010~\cite{alglave2010shared}.
 In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
@@
 Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
-Furthermore, there is a wide range of tools that perform program verification wrt. memory models (see~\cite{alglave2013software}, \cite{Porthos17SAS}).
+
+Furthermore, there exists a wide range of tools that perform program verification wrt.
+memory models (e.g., \cite{alglave2013software}, the tool \tool{Trencher} for checking programs against the TSO memory model~\cite{bouajjani2013checking}, the tool \porthos{} for portability analysis~\cite{Porthos17SAS})
+, weak memory aware model checking (e.g., the model-checker \tool{blender}~\cite{kuperstein2011partial}, the tool \tool{mmchecker} for model-checking the C\# code~\cite{huynh2006memory})
+, synchronisation synthesis (e.g., the automatic fence insertion tool \tool{musketeer}~\cite{alglave2014don})
+, static instrumentation of concurrent C programs for further analysis wrt.
+weak memory model (e.g., the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc}, the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools for performing memory model-aware analysis~\cite{travkin2016verification})
+, and other.
 

### 2018-05-08 4_impl.tex
add some refs to the intro. cleanup a bit.

@@
 
-The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
+The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
 This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Although far not the entire C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
-\footnote{To ensure that, we have merely to look at existing C compilers, for instance, the open-source \texttt{gcc} compiler, that uses a C parser written in more than 18.5 thousand lines (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+\footnote{To ensure that, we have merely to look at existing C compilers, for instance, the open-source \texttt{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
 ), we consider the accomplished work as a step towards this.
@@
 
-As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} by Microsoft Research~\cite{de2008z3}.
+As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
+\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
+by Microsoft Research~\cite{de2008z3}.
 However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver.
@@
 The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos[1] that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
-Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver \textit{Z3}%
-\footnote{The Z3 project repository: \url{https://github.com/Z3Prover/z3}} %
+Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver%
 invoked from \porthos[2] via java API.
@@
 
-The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}. On the picture, rectangles denote processing units (marked with gear sign with a unique number of the component).
+The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
+The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
 
@@
 For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
-(Section~\ref{ch:impl:model:ytree}), which then is being pre-processed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
+(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
 The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
@@
 
-Both \porthos[1] and \porthos[2] use the ANTLR parser generator%
-\footnote{The ANTLR project repository: \url{https://github.com/antlr/antlr4}}%
+Both \porthos[1] and \porthos[2] use the \tool{ANTLR} parser generator%
+\footnote{The \tool{ANTLR} project repository: \url{https://github.com/antlr/antlr4}}%
 ~\cite{parr2013definitive}, a~powerful language processing tool.
-The ANTLR takes as input the user-defined grammar of the target language in a BNF-like form and produces the LL(*)-parser and optionally some auxiliary classes (such as listeners and visitors for the syntax tree).
+The \tool{ANTLR} takes as input the user-defined grammar of the target language in a BNF-like form and produces the LL(*)-parser and optionally some auxiliary classes (such as listeners and visitors for the syntax tree).
 Although this parser may be not as efficient as the hand-written language-optimised parser, it reduces the overhead of implementing the parser significantly.
-Among other advantages ANTLR, it is worth of noting that it has rather large collection of officially supported grammars.
-Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the ANTLR an attractive instrument for solving the parsing problem.
+Among other advantages \tool{ANTLR}, it is worth of noting that it has rather large collection of officially supported grammars.
+Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the \tool{ANTLR} an attractive instrument for solving the parsing problem.
 
-Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full ANTLR grammar is available at Appendix~\ref{apx:grammar}.
+Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.
 
@@
        | <loc> = <reg>.store(<atomic>)
-       | 'mfence'
-       | 'sync'
-       | 'lwsync'
-       | 'isync'
+       | ('mfence' | sync' | 'lwsync' | 'isync')
        ;
@@
 makes the code hardly understandable, and, therefore, poorly maintainable.
-In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the ANTLR syntax tree to the AST, that is one for all  languages of an input program.
+In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the \tool{ANTLR} syntax tree to the AST, that is one for all  languages of an input program.
 
-Secondly, in \porthos[1], the semantics of operations was defined syntactically (in ANTLR grammar), whereas this should be performed by a separate module operating on the AST level%
+Secondly, in \porthos[1], the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
 \footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
@@
 Also, only integers were processed by the input language parser of \porthos[1].
-In \porthos[2], we extended support for primitive types supported by the Z3 solver (this apply to 32-bit integers encoded as \texttt{Int}s, floats encoded as \texttt{Real}s, enums encoded as Z3 \texttt{Scalars}).
-%TODO : constant arrays? https://rise4fun.com/Z3/tutorial/guide
-Although Z3 supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of static analysis of pointers and non-constant sized arrays moves the full support of arrays and pointers out of the scope of current thesis.
+In \porthos[2], we extended support for primitive types supported by the \tool{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s of \tool{Z3}, floats encoded as \texttt{Real}s, enumerations encoded as \texttt{Scalar}s).
+%TODO : constant arrays? https://rise4fun.com/\tool{Z3}/tutorial/guide
+Although \tool{Z3} supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of static analysis of pointers and non-constant sized arrays moves the full support of arrays and pointers out of the scope of current thesis.
 %Thus, adding support for more advanced structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
 
-The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1\,+\,2\,*\,3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
 Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
@@
 
-The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original ANTLR grammar can be found in the official repository containing the collection of ANTLR v4 grammars%
+The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
 \footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}).%
@@
 The abstract syntax tree \ytree{} is an abstraction level suitable for compiling the program to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
-As the ANTLR syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}", that is converted to the \texttt{YIndexerExpression}).
+As the \tool{ANTLR} syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}", that is converted to the \texttt{YIndexerExpression}).
 However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite[Chapter 5]{gries2012science}), we found this level of abstraction suitable enough for our tasks.
@@
 Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
-The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
+The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
 \ Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.
@@
 \label{ch:impl:model:xgraph:example}
-\textbf{consider the following code in C:}
 
-\textbf{picture of the CF-graph:}
+consider the following code in C:
+
+picture of the CF-graph:
 
@@
 In our implementation, the visitor must carry out the continuation of traversing the recursive data structure itself.
-The general structure of the visitor pattern is illustrated by t9he pseudo-java code in Figure~\ref{fig:visitor}.
+The general structure of the visitor pattern is illustrated by the pseudo-java code in Figure~\ref{fig:visitor}.
 
@@
 However, we use its double-dispatching capabilities to reduce cost of multiple type casting performed while traversing the elements of a non-recursive data structure.
-The operator instance, inch:impl:model:xgraph:memstead of having a single method that handles and element of the instance, it extends the visitor interface and splits the handler method into multiple methods, one for each type of element.
+The operator instance, instead of having a single method that handles and element of the instance, extends the visitor interface and splits the handler method into multiple methods, one for each type of element.
 
-For traversing plain (non-recursive) complex data structures, we mostly follow the \texttt{iterator} pattern: the object that 
-the special object Iterator, that has access to elements of the instance, iterates over them and .
-Since in \porthos[] we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
+For traversing plain (non-recursive) complex data structures, we mostly follow the \texttt{iterator} pattern: the special object Iterator, that has access to elements of the data structure, iterates over them and ...
+. Since in \porthos[] we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
 However, in contrast to the visitor of a recursive data structure, the visitor of a plain data structure should be used with an \textit{iterator} that has access to the elements of the object and passes them to the visitor.
@@
 
-Both input-language and input-model parsers are implemented via ANTLR parser generator.
+Both input-language and input-model parsers are implemented via \tool{ANTLR} parser generator.
 
@@
 As preprocessor statements may appear in arbitrary place of a program, the preprocessor must be a stateful processing unit that reads the token stream and dynamically instrument the program by interpreting directives and expanding macros.
-%The ANTLR grammar for the C preprocessor may be found 
+%The \tool{ANTLR} grammar for the C preprocessor may be found 
 %more macros (stateful) : https://gcc.gnu.org/onlinedocs/cpp/Macros.html
@@
 The input memory model language \cat is discussed in Section~\ref{ch:wmm:cat}.
-The ANTLR grammar for the \cat language was extracted from the parser used by \tool{herd} tool%
+The \tool{ANTLR} grammar for the \cat language was extracted from the parser used by \tool{herd} tool%
 \footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
@@
 
-from ANTLR syntax tree of C
+from \tool{ANTLR} syntax tree of C
 

### 2018-05-09 2_wmm.tex
Add some text, picture on compiler arch. cleanup a bit.

@@
 For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$.
-Figure~\ref{example:x86-cat} contains an example of x86-TSO weak memory model definition in \cat{} language that asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}:
-
-\begin{figure}[h]
-\begin{lstlisting}
-"X86 TSO"
-include "x86fences.cat"
-include "filters.cat"
-include "cos.cat"
-
-(* Uniproc check *)
-let com = rf | fr | co
-acyclic po-loc | com
-
-(* Atomic *)
-empty rmw & (fre;coe)
-
-(* Global happens-before *)
-#ppo
-let po_ghb = WW(po) | RM(po)
-
-#mfence
-include "x86fences.cat"
-
-#implied barriers
-let poWR = WR(po)
-let i1 = MA(poWR)
-let i2 = AM(poWR)
-let implied = i1 | i2
-
-let com = rfe | fr | co
-let ghb = mfence | implied | po_ghb | com
-show implied
-acyclic ghb as tso
-\end{lstlisting}
-\caption{The x86-TSO memory model defined in \cat{} language\protect\footnotemark}
-\label{example:x86-cat}
-\end{figure}
-\footnotetext{The \cat{}-memory model for x86-TSO can be found in official repository of \tool{herd} tool: \url{https://github.com/herd/herdtools7}}
-%\begin{figure}[h]
-%\begin{lstlisting}
-%...
-%let com = rf | fr | co
-%let po-loc = po & loc
-%acyclic po-loc | com
-%\end{lstlisting}
-%\caption{Excerpt from the x86-TSO memory model in \cat{} language}
-%\label{example:x86-cat}
-%\end{figure}
-
-%Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}
-
-%The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.
+As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}.
+This memory model asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}.

### 2018-05-09 4_impl.tex
Add some text, picture on compiler arch. cleanup a bit.

@@
        | <loc> = <reg>.store(<atomic>)
-       | ('mfence' | sync' | 'lwsync' | 'isync')
+       | ('mfence' | 'sync' | 'lwsync' | 'isync')
        ;
@@
 For more detailed description of memory management see Section~\ref{ch:impl:proc:x-compiler:mem}.
+
+Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
+However, all three types of \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
+%TODO: rename to XThread ?
+This simplifies drastically the processing of different types of code blocks as they all are modelled by the same data structure.
+The examples of restrictions on X-graphs are the following: the initialisation block can not have branchings, fence events; the process cannot have assertion events; the assertion block can not have shared-memory events or fence events.
+We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler:compilation}.
 
@@
 
-The input memory model language \cat is discussed in Section~\ref{ch:wmm:cat}.
-The \tool{ANTLR} grammar for the \cat language was extracted from the parser used by \tool{herd} tool%
+The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
+The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
 \footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
@@
 need ly?)
+
+- Y-Sytnax error
 
@@
 
+\begin{figure}[th]
+    \centering
+  %draft=true
+  \includegraphics[max width=\textwidth,keepaspectratio]{img/my/draw.io/X-compiler.png}
+  \caption{Main components of the X-compilation processing unit}
+  \label{fig:compiler}
+\end{figure}
+
 The X-graph compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
@@
 
-<COmpiler contains several modules: (todo: remove todo about renaming interpreter to compiler: it's indeed interpreter>
+The X-compiler is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship the principal components of the X-compiler in the UML language.
+
+The X-interpreter is a stateful processing unit that carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the state of the builder.
+
+X compilation error
 
@@
 
-%todo: rename Interpreter -> Compiler.
-
 %TODO: \section{Optimisations}

### 2018-05-09 inc/lst/XInterpreter.java

@@ -1,0 +1,66 @@
+public interface XInterpreter {
+
+    XProcessId getProcessId();
+    XCyclicProcess getResult();
+    void finishInterpretation();
+
+    // memory manager interface methods:
+    XLocation declareLocation(String name, XType type);
+    XRegister declareRegister(String name, XType type);
+    XRegister declareTempRegister(XType type);
+    XLvalueMemoryUnit declareUnresolvedUnit(String name, boolean isGlobal);
+    XLvalueMemoryUnit getDeclaredUnitOrNull(String name);
+    XRegister getDeclaredRegister(String name, XProcessId processId);
+    XLocalMemoryUnit tryConvertToLocalOrNull(XEntity expression);    
+    XComputationEvent tryEvaluateComputation(XEntity entity);
+    XRegister copyToLocalMemory(XSharedMemoryUnit shared);
+
+    // linear interpretation methods:
+    XEntryEvent emitEntryEvent();
+    XExitEvent emitExitEvent();
+    XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
+    XJumpEvent emitJumpEvent();
+    XNopEvent emitNopEvent();
+    XLocalMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination, XLocalMemoryUnit source);
+    XSharedMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination, XSharedMemoryUnit source);
+    XSharedMemoryEvent emitMemoryEvent(XSharedLvalueMemoryUnit destination, XLocalMemoryUnit source);
+    XAssertionEvent emitAssertionEvent(XBinaryComputationEvent assertion);
+
+    // non-linear interpretation methods:
+    void startBlockDefinition(BlockKind blockKind);
+    void startBlockConditionDefinition();
+    void finishBlockConditionDefinition(XComputationEvent conditionEvent);
+    void startBlockBranchDefinition(BranchKind branchKind);
+    void finishBlockBranchDefinition();
+    void finishNonlinearBlockDefinition();
+    void processJumpStatement(JumpKind kind);
+    XEntity processMethodCall(String methodName, @Nullable XMemoryUnit receiver, XMemoryUnit... arguments);
+
+    // create computation event without processing it:
+    XComputationEvent createComputationEvent(XUnaryOperator operator,
+                                             XLocalMemoryUnit operand);
+    XComputationEvent createComputationEvent(XBinaryOperator operator,
[…]

### 2018-05-09 4_impl.tex
place the interpreter picture.

@@
 Further in this Section, the subsection titles have a number in parentheses, which correspond to a unique number of the  processing unit being discussed.
-This numbers are pictured within gears on the top-left corner of each processing unit in Figure~\ref{fig:arch}.
+These numbers are pictured within the gear in the top-left corner of each processing unit in Figure~\ref{fig:arch}.
 %he numbers in parentheses correspond the numbers of processing units pictured within gears in Figure~\ref{fig:arch}.
+
+the full \xgraph{} construction is performed in three stages.
+First, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
+Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
+After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be \textit{CF-acyclic} (no cycles among control-flow edges).
 
@@
 
-\begin{figure}[th]
+The X-compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
+
+\begin{figure}[!t]%[H]
     \centering
@@
 
-The X-graph compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
-
-The compilation is performed in three stages.
-Firstly, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
-Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
-After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be CF-acyclic (no cycles among control-flow edges).
-
-The X-compiler is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship the principal components of the X-compiler in the UML language.
+The It is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
 

### 2018-05-09 1_intro.tex
Consider Prof. Heljanko's comments for the introduction chapter.

--- 2018-05-07 1_intro.tex
@@
 
+\section{Problem statement}
+\label{ch:intro:problem}
+
 Most modern computer systems contain large parts that operate concurrently.
@@
 
-Traditionally, studies related to concurrent programming concern on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language.
-Unfortunately, when it comes to the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability.
+Traditionally, studies related to concurrent programming focus more on fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language~\cite{ben2006principles}.
+Unfortunately, when it comes to real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their correctness.
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible~\cite{adve1996shared}.
@@
 This state is allowed because in the x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not immediately become visible by processes running on other cores.
-In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
-Meanwhile, the write cache of the process \texttt{P1} may not have an updated version of the variable~\texttt{[x]}, neither the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in the other thread.
-These problems have lead to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{weak memory models (WMMs)}.
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 \hline
-
 \end{tabular}
-\caption{Store buffering (SB): a litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
+\caption{Store buffering (SB): A litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
 \label{simple_wmm_x86}
@@
 
-Research of weak memory models firstly aims to \textit{formalise} develop the formal approach of understanding programs with respect to weak memory models which is systematic, sound and complete.
-The first (and so far the only) such a framework was formalised by Jade Alglave in 2010~\cite{alglave2010shared}.
+In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
+Meanwhile, the write cache of the process \texttt{P1} may not have an updated version of the variable~\texttt{[x]}, neither the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in the other process.
+The problems enumerated above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
+%The automation of the memory model-aware program analysis constitutes the main research interest of the current thesis.
+
+The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
+This memory model, called the \textit{sequential consistency (SC)}, allows only those executions that produce the same result as if the operations had been executed in an interleaved fashion in a single process%
+%
+\footnote{In order to prescind from the implementation details while discussing the memory models theory, we avoid the use of the software-specific term \textit{thread} and the hardware-specific term \textit{processor}.
+Instead, we adhere the terminology of the theory of concurrency employed by Ben-Ari~\cite{ben2006principles} by naming a concurrent piece of code the \textit{process}.}.
+%
+%
+This means that the order of operations executed by a process is strictly defined by the program (the code) it executes.
+The SC model does require the write to a shared variable performed in one process to become visible by all other processes \textit{instantly}.
+This means that each process writes directly to the shared memory, without local buffering.
+Another important requirement of the SC memory model is that it forbids memory operations reordering within a single process (the order is strictly defined by the program).
+
+The SC model is considered to be a \textit{strong memory model} in the sense that it provides firm guarantees regarding the ordering and caused effect of memory operations.
+Different relaxations of this model lead to the class of \textit{weak memory models~(WMMs)} that specify how processes interact through shared memory, when a write becomes visible to other processes and what value a read can return.
+Thus, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may rely on.
+
+\section{Related work}
+\label{ch:intro:related}
+
+Research on weak memory models firstly aims to \textit{formalise} a formal approach of understanding programs with respect to weak memory models which is \textit{systematic}, \textit{sound} and \textit{complete}.
+Perhaps the most well-known framework for weak memory model-aware analysis was formalised by J. Alglave in 2010~\cite{alglave2010shared}.
 In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
-Over last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, and others.
+Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, and others.
 There are projects for validating hardware architectures wrt. a memory model, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.
@@
 Most modern high-level programming languages rely on relaxed memory model as well.
-Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time}, it was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{iso2012iec} has introduced the set of hardware-independent synchronisation fences and atomic operations, whenever the C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
-Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{kernel1}.
-Furthermore, there is a wide range of tools that perform program verification wrt. memory models (see~\cite{alglave2013software}, \cite{Porthos17SAS}).
+Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time}, it was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{iso2012iec} has introduced the set of hardware-independent synchronisation fences and atomic operations.
+The C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
+Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{alglave2018frightening, kernel1}.
 
-%experiments with kernel?
+Furthermore, there exists a wide range of tools that perform memory model-aware analysis.
+\begin{itemize}[noitemsep,topsep=0pt]
 
-The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
-This memory model, called the \textit{sequential consistency (SC)}, allows only those executions (interleavings) that produce the same result as if the operations had been executed by single process.
-This means that the order of operations executed by a process is strictly defined by the program it executes.
-The SC model does requires the write to a shared variable performed in one process to become visible by all other processes not instantly, but simultaneously.
-This means each process communicates to the shared memory directly, without local buffering.
-Another important requirement of SC memory model is that it forbids memory operations reordering within single process (the order is strictly defined by the program).
+\item A state-of-the-art tool is \tool{diy} (\textit{do it yourself}), developed by researchers from INRIA institute, France and University of Cambridge, UK.
+The \tool{diy}%
+\footnote{The \tool{diy} project web site: \url{http://diy.inria.fr/}} %
+is a software suite for designing and testing weak memory models.
+It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
+The \tool{diy} consists of several modules:
+the litmus tests generators \tool{diy}, \tool{diycross} and \tool{diyone},
+the litmus test concrete executor \tool{litmus} that runs tests on a physical machine and collects its behaviours,
+and the weak memory model simulator \tool{herd} that implements reachability analysis for exploring states reachable under the specified~WMM.
 
-The SC model is considered to be the strong memory model in the sense that it provides strong guarantees regarding the ordering and caused effect of memory operations.
-Different relaxations of this model lead to the class of \textit{weak memory models~(WMM)}.
-%while preserving consistency
-They specify how threads interact through shared memory, when a write becomes visible to other threads and what value a read can return. 
-Therefore, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may expect.
+\item There exist tools that perform the weak memory model-aware program verification and model checking.
+The notable examples are
+%the tool \tool{mmchecker} for model-checking the C\# code~\cite{huynh2006memory}
+the stateless model checkers \tool{RCMC}~\cite{kokologiannakis2017effective}, \tool{CHESS}~\cite{musuvathi2008fair} and \tool{Nidhugg}~\cite{abdulla2017stateless},
+the tool \tool{Trencher} for checking programs against the TSO memory model~\cite{bouajjani2013checking}, 
+the tool \porthos{} for portability analysis~\cite{Porthos17a},
 
-Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them.
-The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France and University of Cambridge, UK.
-The \tool{diy}\footnote{Project web site: \url{http://diy.inria.fr/}} is a software suite for designing and testing weak memory models.
-It is firstly released back in 2010, and since that time it remained to be the only tool for testing weak memory models.
-The \tool{diy} consists of several modules: the litmus tests generators \tool{diy}, \tool{diycross} and \tool{diyone}, the litmus tests concrete executor \tool{litmus} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd} that implements reachability analysis for exploring states reachable under specified~WMM.
+\item Some tools tackle the problem of automated synthesis of the synchronisation primitives, such as
+the automatic fence insertion tool \tool{musketeer}~\cite{alglave2014don}, and
+the automatic verification and fence inference tool \tool{blender}~\cite{kuperstein2011partial}.
 
-All the \tool{diy} tools work only with single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
+\item Some other tools perform static instrumentation of concurrent C programs and encode the WMM into the program representation so that it can be model-checked by standard tools.
+The examples are
+the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
+the program transformation tool~\cite{alglave2013software},
+the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.
+\end{itemize}
+
+%Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to them.
+
+All the tools listed above consider only a single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
 One of these problems is the \textit{portability} of the program from one hardware architecture to another.
@@
 Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17SAS}.
+As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17a}.
 
-The first tool that performs the WMM-aware portability analysis is \porthos%
-\footnote{Project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}}%
-introduced in April~2017~\cite{Porthos17CoRR}.
+The very first tool that performs the WMM-aware portability analysis is \porthos%
+\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} %
+introduced in April~2017~\cite{Porthos17b}.
 This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
-This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula, augmented by the reachability assertion.
+This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
 As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
@@
 
-%In the work~\cite{Porthos17CoRR}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
+%In the work~\cite{Porthos17b}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
 Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a trivial problem of symbolic execution, encoding of the weak memory model is more tedious.
@@
 
+\section{Task statement}
+\label{ch:intro:task}
+
 Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
@@
 As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \porthos[2]%
-\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1 (also addressed as \porthos[1]), whereas the new version of \porthos{} is called \porthos[2].} %
+\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1, whereas the new version of \porthos{} is called \porthos[2].} %
 .
@@
 
-The thesis is organised as following. Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis. Chapter~...
+The thesis is organised as following.
+Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis.
+Chapter~\ref{ch:port} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
+Chapter~\ref{ch:impl} delves into the architectural description and implementation details of the \porthos[2] framework.
+\textbf{TODO: other chapters!!!}
 
-% TODO: Add intro to the bounded reachability analysis using SAT -- ?
-
-% TODO: other tools for bounded model-checking: 
 %Clarke's goto-cc : \cite{clarke2004tool} : http://www.cprover.org/goto-cc/

### 2018-05-09 2_wmm.tex
Consider Prof. Heljanko's comments for the introduction chapter.

@@
 %TODO: nondeterministic
-One such model is the \textit{event-based} representation of a program, which models a program as a directed graph of events (the \textit{event-flow graph}).
+One such model is the \textit{event-based} representation of a program, which models the program as a directed graph of events (the \textit{event-flow graph}).
+The event-based model is \textit{non-deterministic}, which
 The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent \textit{basic relations} (see Section~\ref{ch:wmm:model:relations}).
@@
 
-The event is an independent low-level atomic operation.
+%TODO: EVENT -- ATOMIC?
+An event is a fact of executing the low-level primitive operation such as memory access, threads synchronisation, computation, etc.
 
@@
   \begin{itemize}
-    \item \textit{the union} of two relations \rel{r1} and \rel{r2} is \rel{r1\,|\,r2}
-    \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2}
-    \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$
-    \item \textit{the sequence}\footnote{The sequence of two relations \rel{r1} and \rel{r2} is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$}
-    of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}
+    \item \textit{the union} of two relations \rel{r1} and \rel{r2} is \rel{r1\,|\,r2},
+    \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2},
+    \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$,
+    \item \textit{the sequence} of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}, which is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$.
   \end{itemize}

### 2018-05-09 3_encod.tex
Consider Prof. Heljanko's comments for the introduction chapter.

@@
 \label{ch:port}
+
+% TODO: Add intro to the bounded reachability analysis using SAT -- ?
+
+% TODO: other tools for bounded model-checking:
 
@@
 % The first tool that tackles the problem 
-%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17SAS}  
+%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17a}  
 In this Chapter, we describe the general task of analysing the concurrent software portability
 %may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17SAS} (more precisely, to an SMT-problem).
+as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17a} (more precisely, to an SMT-problem).
 
@@
 
-The model checking is the problem of verifying the system (the model) against the set of constraints (the specification).
+The model checking is the problem of verifying the system (the model) against the set of constraints (the specification)~\cite{dkw2008}.
 As the state machine model is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
@@
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
-%As well as the BMC problem, the approach used by the \tool{PORTHOS}
+%As well as the BMC problem, the approach used by the \porthos{}
 
@@
 Consider the function $cons_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
-Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17SAS}:
+Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17a}:
 
@@
 
-Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17CoRR}.
+Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17b}.
 The result SMT-formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
@@
 %The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. 
-%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17SAS}.
+%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17a}.
 
@@
 In contrast, \porthos{} used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
-Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17SAS}.
+Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17a}.
 
-Following~\cite{Porthos17CoRR}, the indexed references of variables are computed in accordance with the following rules:
+Following~\cite{Porthos17b}, the indexed references of variables are computed in accordance with the following rules:
 (1)~any access to a shared variable (both read and write) increments its SSA-index;
@@
 %
-%As it is described in~\cite{Porthos17SAS}, the is based . . .
-%~\cite{Porthos17SAS}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"
+%As it is described in~\cite{Porthos17a}, the is based . . .
+%~\cite{Porthos17a}%, Section "Programs and Memory Models", subsection "Encoding Derived Relations"

### 2018-05-09 4_impl.tex
Consider Prof. Heljanko's comments for the introduction chapter.

@@
-\chapter{The~\texttt{porthos2}:~implementation}
+\chapter{The~\texttt{PorthosC}:~implementation}
 \label{ch:impl}
@@
 
-%\section{The architecture of \porthos[1]}
+%\section{The architecture of \porthos{}}
 \section{General principles}
@@
 
-The existing implementation of \porthos[1] does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
+The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
 
-The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos[1] that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
+The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos{} that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
 Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver%
@@
 
-Both \porthos[1] and \porthos[2] use the \tool{ANTLR} parser generator%
+Both \porthos{} and \porthos[2] use the \tool{ANTLR} parser generator%
 \footnote{The \tool{ANTLR} project repository: \url{https://github.com/antlr/antlr4}}%
@@
 
-The input language parser used by \porthos[1] suffered from several disadvantages.
+The input language parser used by \porthos{} suffered from several disadvantages.
 Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
@@
 
-Secondly, in \porthos[1], the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
+Secondly, in \porthos{}, the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
 \footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
@@
 For example, it allowed only computations over the local variables, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code snippet `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
-Also, only integers were processed by the input language parser of \porthos[1].
+Also, only integers were processed by the input language parser of \porthos{}.
 In \porthos[2], we extended support for primitive types supported by the \tool{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s of \tool{Z3}, floats encoded as \texttt{Real}s, enumerations encoded as \texttt{Scalar}s).
@@
 
-The minor drawbacks of grammar used by \porthos[1] include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
-Also, \porthos[1] supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
+The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
 tialisation of the shared variable and only by default value \lstinline{0}. The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
@@
 
-X-events are the independent atomic operations.
+An event (\texttt{XEvent}) represents the fact of executing the primitive operation, which is independent from other events.
 Each event is specified by the process generated them and a unique event label.
@@
     The fences are implemented as an enumeration \texttt{XBarrierEvent}.
-    Current implementation of \porthos[2] supports all fences supported by \porthos[1]: 
+    Current implementation of \porthos[2] supports all fences supported by \porthos{}: 
     \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
@@
 
-in \porthos[1] : ctx as an argument, calling ctx to create new clause. 
+in \porthos{} : ctx as an argument, calling ctx to create new clause. 
 drawbacks:
@@
 
-The It is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
+The X-compiler is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
 

### 2018-05-09 1_intro.tex
Finish considering Prof. Heljanko's comments, fix some mistakes in first 3 chapters.

@@
 Most modern computer systems contain large parts that operate concurrently.
-Though parallelisation of the system can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
-% Most articles, presentations and books on concurrent programming start with words how hard it is.
+Although the parallelisation of the system can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
 
-Traditionally, studies related to concurrent programming focus more on fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language~\cite{ben2006principles}.
+Traditionally, studies related to concurrent programming focus on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language~\cite{ben2006principles}.
 Unfortunately, when it comes to real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their correctness.
 The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible~\cite{adve1996shared}.
-For instance, Figure~\ref{simple_wmm_x86} provides simple example of reachability of the state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' on the x86 architecture (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
-%reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
-This state is allowed because in the x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not immediately become visible by processes running on other cores.
+
+For example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86}.
+The process~\texttt{P0} writes value~\texttt{1} to the shared variable~\texttt{x} and reads the value of shared variable~\texttt{y}, and the process~\texttt{P1} writes the~\texttt{y} and reads the~\texttt{x}.
+As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
+However, this state is reachable on the x86 architecture as each processor may cache the write to shared memory location into its local write buffer, so that they do not immediately become visible by processes running on other cores.
+This behaviour is known as \textit{store buffering (SB)}.
+The little code examples that illustrate specific behaviour of an execution environment are called \textit{litmus tests}.
 
@@
 \hline
-\multicolumn{2}{|l|}{x86-TSO: allow} \tabularnewline
+\multicolumn{2}{|l|}{x86: allow} \tabularnewline
 \hline
 \end{tabular}
-\caption{Store buffering (SB): A litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
+\caption{Store buffering: A litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
 \label{simple_wmm_x86}
@@
 
-In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
-Meanwhile, the write cache of the process \texttt{P1} may not have an updated version of the variable~\texttt{[x]}, neither the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in the other process.
-The problems enumerated above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
+%In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}.
+%Meanwhile, the write cache of the process \texttt{P1} may not have an updated version of the variable~\texttt{[x]}, neither the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0} even if this variable has been already updated in the other process.
 %The automation of the memory model-aware program analysis constitutes the main research interest of the current thesis.
+
+The issues like one discussed above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
+There are two ways to define a memory model: by specifying its \textit{operational semantics} (in terms of an abstract machine executing the code, as it was done for the SB example above) or by defining its \textit{axiomatic semantics} (in terms of assertions over states of the program formed by sequences of operations).
+Although the former type of memory models may be easier to describe and interpret, existing numerous formal verification tools and methods direct the research towards axiomatic memory models.
 
@@
 This means that the order of operations executed by a process is strictly defined by the program (the code) it executes.
-The SC model does require the write to a shared variable performed in one process to become visible by all other processes \textit{instantly}.
-This means that each process writes directly to the shared memory, without local buffering.
+The SC model does require the write to a shared variable performed in one process to become visible by all other processes \textit{instantly}, each process writes directly to the shared memory, without local buffering.
 Another important requirement of the SC memory model is that it forbids memory operations reordering within a single process (the order is strictly defined by the program).
+Originally, the operational semantics was defined for the SC model, however there exist axiomatic specifications for it~\cite{mansky2015axiomatic}.
 
@@
 Research on weak memory models firstly aims to \textit{formalise} a formal approach of understanding programs with respect to weak memory models which is \textit{systematic}, \textit{sound} and \textit{complete}.
-Perhaps the most well-known framework for weak memory model-aware analysis was formalised by J. Alglave in 2010~\cite{alglave2010shared}.
+One of the most well-known frameworks for weak memory model-aware analysis was formalised by J. Alglave in~2010~\cite{alglave2010shared},.
 In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
-Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architecture formalised in 2009~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, and others.
-There are projects for validating hardware architectures wrt. a memory model, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.
+Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
+%, or Alpha architecture~\cite{alphaWMM}~, and others.
+There are projects for validating hardware architectures wrt. a memory model as well, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.
+
+%TODO: litmus test synthesis 
+%1) http://research.nvidia.com/sites/default/files/pubs/2017-04_Automated-Synthesis-of//ASPLOS_2017_Memory_Model_Verification.pdf
+% 2) kernel litmus tests % <----------------------------- OBIAZATELNO!!!
 
@@
 
-\section{Task statement}
+\section{Task specification}
 \label{ch:intro:task}
 
-Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
+The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
 %interprocedural analysis
-As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \porthos[2]%
-\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1, whereas the new version of \porthos{} is called \porthos[2].} %
-.
-Considering the enhancements of the architecture, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.
+As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \textit{\porthos[2]}\,%
+%
+\footnote{Hereinafter with the name `\porthos{}' we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \porthos[2].}. %
+%
+Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.
 
@@
 Chapter~\ref{ch:port} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
-Chapter~\ref{ch:impl} delves into the architectural description and implementation details of the \porthos[2] framework.
+Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.
 \textbf{TODO: other chapters!!!}

### 2018-05-09 2_wmm.tex
Finish considering Prof. Heljanko's comments, fix some mistakes in first 3 chapters.

@@
-%TODO: move this chapter as a section to introduction and add here the new "Problem statement" chapter!!!
-
-\chapter{Memory model-aware analysis}%Description of the analysis of concurrent systems wrt weak memory models
-%\chapter{A generic framework}
+\chapter{Memory model-aware analysis}
 \label{ch:wmm}
 
-In general, analysis of concurrent programs with respect to axiomatic memory models is performed in several stages.
-%TODO: site the 'Weak Memory Models with Matching Axiomatic and Operational Definitions' https://arxiv.org/pdf/1710.04259.pdf
-%TODO: cite also: "DENNIS SHASHA - Efficient and Correct Programs that Share" http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.142.2238&rep=rep1&type=pdf
-Firstly, the control-flow and data-flow of a program is encoded as the set of possible \textit{candidate executions}. %possible communications between threads of the program.
-Obtained model of the program describes the anarchic semantics, which is a truly parallel semantics with no global time that describes all possible computations with all possible communications~\cite{alglave2016syntax}.
-Thereafter, the anarchic semantics is constrained by the \textit{weak memory model} specification which is a set of axiomatic constraints for filtering out executions inconsistent in particular architecture.
+The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (\textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (\textit{analytic semantics})~\cite{alglave2016syntax}.
+The anarchic semantics of the program is a truly parallel semantics with no global time that describes all possible computations with all possible communications.
+However, the analytic semantics captures the program behaviours on a certain execution environment more precisely.
+%"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf
 
@@
 
-%"The semantics of a program is a set of executions" https://johnwickerson.github.io/papers/memalloy.pdf
-The classical approach for modeling concurrent programs is to use the \textit{global time}, a single order of interleavings among all events happened in different threads.
-Although these models are easy to understand, it may be impossible to treat \textit{all} possible states, number of which is exponentially large.
-However, there exist equivalence classes such that the result of execution different interleavings from single equivalence class is the same (for instance, computations performed by a processor locally do not affect the global state).
-%One way to do this is to use a non-deterministic computation-centric model~\cite{fri97}, such as ...%The idea in this class of models is based on the fact that the behaviour of a concurrent system is defined only by the interleavings of shared-memory operations, while being independent from the order of local computation events.
-%TODO: nondeterministic
-One such model is the \textit{event-based} representation of a program, which models the program as a directed graph of events (the \textit{event-flow graph}).
-The event-based model is \textit{non-deterministic}, which
+The classical approach for analysing concurrent programs is to model it as the set of sequentially consistent programs, obtained by enumerating all possible interleavings.
+These models are deterministic as they include the notion of the \textit{global time}. %, a single order of interleavings among all events happened in different threads.
+Although these models are easy to build and analyse, the number of all possible interleavings grows exponentially (known as the \textit{combinatorial explosion}), which affects the completeness of an analysis method in general case.
+
+One way to fight the combinatorial explosion is to exclude the global time from the model and treat executions from one equivalence class together in a non-deterministic fashion.
+For instance, such an equivalence class can be the set of computations performed by a processor locally that do not affect the global state.
+This idea is used in the \textit{event-based} model, that represents the program as a directed graph of events (the \textit{event-flow graph})~\cite{alglave2010shared}.
 The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent \textit{basic relations} (see Section~\ref{ch:wmm:model:relations}).
+The graph represents the set of executions (sequences of events; see Section~\ref{ch:wmm:model:executions}) defined by the non-deterministic guesses of certain relations on some states.
 
 % TODO: the event-based representation is a non-deterministic model
-
-% TODO: the semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
+%- The semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
 %"To build the aeg , we define a semantics of goto-programs in terms of abstract events, ..."
@@
 \label{ch:wmm:model:events}
-
-%TODO: event is atomic?
-
-% TODO: check this subseciton
 
@@
 We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\mathtt{load}(r_1~\leftarrow~l_2); \ e_2~=~\mathtt{store}(l_1~\leftarrow~r_1)$.
-Also, it is important to separate the set of initial write events $\mathbb{IW}~\subset~\mathbb{W}$ that perform initialisation of program variables.
+Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.
 
@@
 An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (which is modelled as a register as well) to the register $r_1$.
-For modelling branching statements, we distinguish the set $\mathbb{C_{p}}~\subseteq~\mathbb{C}$ of \textit{predicative} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
+For modelling branching statements, we distinguish the set $\mathbb{C_{g}}~\subseteq~\mathbb{C}$ of \textit{guard} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
 
@@
 \begin{itemize}
-    \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subset~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
+    \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subseteq~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
     For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relation{po} e_2$.
@@
         \begin{itemize}[noitemsep]
-            \item the \textit{read-from} relation \rf~$\subset~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
-            \item the \textit{coherence order} relation \co~$\subset~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
+            \item the \textit{read-from} relation \rf~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
+            \item the \textit{coherence order} relation \co~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
         \end{itemize}
 
-    \item Events from the same process are related by the \textit{scope relation} \sr~$\subset~\mathbb{E}~\times~\mathbb{E}$.
-    In contrast to the \tool{herd} tool, the \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple lables that indicate which process has produced certain event.
+    \item Events from the same process are related by the \textit{scope relation} \sr~$\subseteq~\mathbb{E}~\times~\mathbb{E}$.
+    In contrast to the \tool{herd} tool, the \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple labels that indicate which process has produced certain event.
 \end{itemize}
@@
 
-    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events preceding the write event from which the read event gets its value: \\
+    \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
     $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$
@@
 
-    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes: \\
+    \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
     $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(w))$ \\
@@
 
-The semantics of a concurrent program is represented as the set of allowed executions. The 
+The semantics of a concurrent program is represented as the set of allowed executions.
 The \textit{execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
@@
     \begin{itemize}
-      \item \textit{the complement} of a relation \rel{r} is $\rel{\textasciitilde r}$
-      \item \textit{the transitive closure} of a relation \rel{r} is $\rel{r+}$
-      \item \textit{the reflexive closure} of a relation \rel{r} is $\rel{r?}$
-      \item \textit{the reflexive-transitive closure} of a relation \rel{r} is $\rel{r*}$
-      \item \textit{the inverse} of a relation \rel{r} is $\invrel{r}$
+      \item \textit{the complement} of a relation \rel{r} is $\rel{\textasciitilde r}$,
+      \item \textit{the transitive closure} of a relation \rel{r} is $\rel{r+}$,
+      \item \textit{the reflexive closure} of a relation \rel{r} is $\rel{r?}$,
+      \item \textit{the reflexive-transitive closure} of a relation \rel{r} is $\rel{r*}$, and
+      \item \textit{the inverse} of a relation \rel{r} is $\invrel{r}$.
     \end{itemize}
@@
     \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2},
-    \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$,
+    \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$, and
     \item \textit{the sequence} of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}, which is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$.

### 2018-05-09 3_encod.tex
Finish considering Prof. Heljanko's comments, fix some mistakes in first 3 chapters.

@@
-%\chapter{SMT-based program analysis}
-%\chapter{Portability of concurrent software}
 \chapter{Portability analysis as an SMT~problem}%Description of the method used by Porthos
@@
 % The first tool that tackles the problem 
-%The first work that has investigated the practical approach of modeling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17a}  
+%The first work that has investigated the practical approach of modelling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17a}  
 In this Chapter, we describe the general task of analysing the concurrent software portability
@@
 
-The model checking is the problem of verifying the system (the model) against the set of constraints (the specification)~\cite{dkw2008}.
+The model checking is the problem of verifying the system (the model) against a set of constraints (the specification)~\cite{dkw2008}.
 As the state machine model is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
-The general schema of model checking is the following: firstly, the analysing system is being represented as a transition system, a finite directed graph with labeled nodes representing states of the system such that each state corresponds to the unique subset of atomic propositions, that characterise the behavioral properties of each state.
+The general schema of model checking is the following: firstly, the analysing system is being represented as a transition system, a finite directed graph with labelled nodes representing states of the system such that each state corresponds to the unique subset of atomic propositions that characterise the behavioural properties of each state.
 Then, the system constraints are being defined in terms of a modal temporal logic with respect to the atomic propositions.
@@
 However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables used by the system (its size).
-In case of modeling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
+In case of modelling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
 Therefore, the research in model checking over past 40 years was aimed at tackling the state explosion problem, mostly by optimising search space, search strategy or basic data structures of existing algorithms.
 
-One of the first technique that optimises the search space considerably major was the symbolic model checking with binary decision diagrams (BDDs).
-Instead of by processing each state individually, in this approach the set of states is represented by the BDD, efficient data structure for performing operations on large boolean formulas~\cite{clarke2012model}.
+One of the first techniques that optimises the search space considerably is the symbolic model checking with \textit{binary decision diagrams (BDDs)}.
+Instead of processing each state individually, in this approach the set of states is represented by the BDD, efficient data structure for performing operations on large boolean formulas~\cite{clarke2012model}.
 The BDD representation can be linear of size of variables it encodes if the ordering of variables is optimal, otherwise the size of BDD is exponential.
@@
 Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \rf-relation (see Section~\ref{ch:wmm:model:relations} for details on relations).
-We distinguish the function ${\fv_p\;:\;\mathbb{C_p}\;\rightarrow\;\{0,1\}}$ that evaluates the predicative computation event.
+We distinguish the function ${\fv_g\;:\;\mathbb{C_g}\;\rightarrow\;\{0,1\}}$ that evaluates the guard computation event.
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.

### 2018-05-09 4_impl.tex
Finish considering Prof. Heljanko's comments, fix some mistakes in first 3 chapters.

@@
 Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
-\footnote{To ensure that, we have merely to look at existing C compilers, for instance, the open-source \texttt{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
 ), we consider the accomplished work as a step towards this.
@@
 
+- carries the origin (text) -> all the origins can be resolved by the \texttt{LocationService} %todo: rename from CodeCitationService + create 3 instances for each of level.
+Or keep it for Y->text, and store references to the original Y-object by X-objects
+
 \vspace{1em}
@@
 
+- carries the origin (Y)
+
 The following interfaces model basic kinds of events (see Figure~\ref{fig:class-diagrams:XEntity-interfaces}):
@@
 
+todo: rename everywhere assertions <- axioms
+
 \subsubsection{Z-formula}
@@
 typed: Expr, BoolExpr, ArithExpr, ...
+
+carries the origin (X)
 
@@
 
+%TODO: cite again here : alglave2018frightening + other Kernel references
+
 The invocation hooking module serves as a knowledge base that stores the semantics of functions.

### 2018-05-09 2_wmm.tex
Some additions (non-determinism, ...).

@@
 
-% TODO: the event-based representation is a non-deterministic model
+There are three main types of sources of non-determinism in concurrent programs~\cite{musuvathi2008fair}:
+\begin{enumerate}[noitemsep,topsep=0pt]
+\item \textit{input non-determinism}, which is a standard undecidable problem for all static analysis methods: to resolve the user input, system call from the environment, unresolved function calls, etc.;
+\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity;
+\item \textit{memory-model non-determinism}, caused by hardware and compiler relaxations.
+\end{enumerate}
+
+The event-based program model is able to emulate effectively the second and the third types of sources of non-determinism, while the first one can be coped by standard static analysis methods~\cite{landi1992undecidability,SurveySymExec-CSUR18}.
+
 %- The semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9

### 2018-05-09 4_impl.tex
Some additions (non-determinism, ...).

@@
 
-As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
+As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
 \footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
@@
 We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is defined also syntactically).
+Each \ytree{} expression carries the original code coordinates, that can be resolved to text by the service \texttt{LocationService}.
+%todo: rename from CodeCitationService + create 3 instances for each of level.
+Or keep it for Y->text, and store references to the original Y-object by X-objects
+
 %\textbf{(TODO:THIS SHOULD BE IN PRE-PROCESSING)} %TODO <--
-
-- carries the origin (text) -> all the origins can be resolved by the \texttt{LocationService} %todo: rename from CodeCitationService + create 3 instances for each of level.
-Or keep it for Y->text, and store references to the original Y-object by X-objects
 
@@
 The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
-\ Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.
+The UML diagraph on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
+%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring
 
@@
 This information is stored by events in the immutable structure \texttt{XEventInfo}.
-
-- carries the origin (Y)
+Also, each event carries the reference to the Y-instruction that has generated it.
 
@@
     \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
-    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each
+    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each!
 
@@
 
-\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
-\label{ch:impl:model:xgraph:builder}
+%\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
+%\label{ch:impl:model:xgraph:builder}
 %TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
-
-\textbf{todo}
 

### 2018-05-11 1_intro.tex
polish text a bit.

@@
 Most modern computer systems contain large parts that operate concurrently.
-Although the parallelisation of the system can improve its performance drastically, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
+Although the parallelisation of a system can drastically improve its performance, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
 
@@
 Unfortunately, when it comes to real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their correctness.
-The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible~\cite{adve1996shared}.
+The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance of the system~\cite{adve1996shared}.
 
-For example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86}.
-The process~\texttt{P0} writes value~\texttt{1} to the shared variable~\texttt{x} and reads the value of shared variable~\texttt{y}, and the process~\texttt{P1} writes the~\texttt{y} and reads the~\texttt{x}.
-As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
-However, this state is reachable on the x86 architecture as each processor may cache the write to shared memory location into its local write buffer, so that they do not immediately become visible by processes running on other cores.
-This behaviour is known as \textit{store buffering (SB)}.
-The little code examples that illustrate specific behaviour of an execution environment are called \textit{litmus tests}.
+As an example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86}.
+The process~\texttt{P0} writes the value~\texttt{1} to the shared variable~\texttt{x} and reads a value of the shared variable~\texttt{y}, and the process~\texttt{P1} writes the~\texttt{y} and reads the~\texttt{x}.
+Considering all interleavings of the processes, one could expect the final state to be one of the following states:
+\begin{itemize}[noitemsep,topsep=0pt]
+\item \texttt{(0:EAX=0,\,1:EBX=1)}
+\item \texttt{(0:EAX=1,\,1:EBX=0)}
+\item \texttt{(0:EAX=1,\,1:EBX=1)}
+\end{itemize}
+%As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
+However, on the x86 architecture the state `\texttt{(0:EAX=0,\,1:EAX=0)}' is also reachable as the processors may cache the write to shared memory into their local write buffers, so that they do not immediately become visible by processes running on other cores.
+This behaviour is known as \textit{store buffering~(SB)}.
 
@@
 MOV [x],1 & MOV [y],1 \\
-MOV EAX,[y] & MOV EAX,[x] \\
+MOV EAX,[y] & MOV EBX,[x] \\
 \hline
 %\multicolumn{2}{|l|}{locations [x;y;]} \tabularnewline
-\multicolumn{2}{|l|}{exists (0:EAX=0~/\textbackslash~1:EAX=0)} \tabularnewline
+\multicolumn{2}{|l|}{exists (0:EAX=0,~1:EBX=0)} \tabularnewline
 \hline
@@
 
-The issues like one discussed above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
-There are two ways to define a memory model: by specifying its \textit{operational semantics} (in terms of an abstract machine executing the code, as it was done for the SB example above) or by defining its \textit{axiomatic semantics} (in terms of assertions over states of the program formed by sequences of operations).
+Such little examples that illustrate specific behaviour of a concurrent execution environment are called \textit{litmus tests}.
+The formal way to define the semantics of memory operations and synchronisation primitives of a parallel execution environment is to define its \textit{memory model}.
+%The behaviour of a concurrent execution environment is described by a formal memory model
+%The study of concurrency of a specific platform concerns on defining
+%and are widely used to illustrate its behaviour.
+%Properties of an execution environment can be described by a formal \textit{memory model}
+%The issues like one discussed above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
+There are two types of formal memory models.
+Models of the first type characterise the behaviour of the system (its \textit{operational semantics}) in terms of the abstract machine executing the code, as it was done for the SB example above.
+Models of the second type define the \textit{axiomatic semantics} of the system by specifying the set of assertions over states of the program.
+%There are two ways to define a memory model: by specifying its \textit{operational semantics} (in terms of an abstract machine executing the code, as it was done for the SB example above) or by defining its \textit{axiomatic semantics} (in terms of assertions over states of the program formed by sequences of operations).
 Although the former type of memory models may be easier to describe and interpret, existing numerous formal verification tools and methods direct the research towards axiomatic memory models.
@@
 The \tool{diy}%
+%
 \footnote{The \tool{diy} project web site: \url{http://diy.inria.fr/}} %
+%
 is a software suite for designing and testing weak memory models.
@@
 Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled under the platforms $T$ can reach states that are unreachable on the platform $S$, which is a \textit{portability bug} from the source platform $S$ to the target platform $T$~\cite{Porthos17a}.
+As the result, a program compiled under the platforms $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
 
 The very first tool that performs the WMM-aware portability analysis is \porthos%
+%
 \footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} %
+%
 introduced in April~2017~\cite{Porthos17b}.
-This tool reduces described problem to a bounded reachability problem, which can be solved with help of an SMT-solver.
+This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver.
 This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
@@
 %In the work~\cite{Porthos17b}, the SMT-based approach was defined for analysing the portability of a program from one hardware architechture to another, which is defined as ``an execution that is consistent with the target but inconsistent with the source memory model".
-Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a trivial problem of symbolic execution, encoding of the weak memory model is more tedious.
+Although encoding the control-flow and the data-flow of a program into an SMT-formula seems to be a tril problem of symbolic execution, encoding of the weak memory model is more tedious.
 The reason is that some relations of WMMs are defined as mutually-recursive and need to be linearised in order to be encoded into an equivalent logical formula.

### 2018-05-11 2_wmm.tex
polish text a bit.

@@
 This idea is used in the \textit{event-based} model, that represents the program as a directed graph of events (the \textit{event-flow graph})~\cite{alglave2010shared}.
-The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent \textit{basic relations} (see Section~\ref{ch:wmm:model:relations}).
+The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent basic relations (see Section~\ref{ch:wmm:model:relations}).
 The graph represents the set of executions (sequences of events; see Section~\ref{ch:wmm:model:executions}) defined by the non-deterministic guesses of certain relations on some states.
@@
 
-The relation~%\xrightarrow{r}
-$r\,\subseteq~\mathbb{E}~\times~\mathbb{E}$ is a set of pairs of events (a subset of Cartesian product of two sets of events).
-There are two kinds of relations between events: \textit{basic relations} %retrieved from the
-that capture semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification.
+The relation~$r\,\subseteq~\mathbb{E}~\times~\mathbb{E}$ is a set of pairs of events (a subset of Cartesian product of two sets of events).
+There are two kinds of relations between events: \textit{basic relations} that capture the semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification.
 Constraints over relations that are specified by weak memory models are defined as requirements of acyclicity, irreflexivity or emptiness of specific relations~\cite{alglave2016syntax}.
@@
 \begin{enumerate}
-  \item \textit{Unary relations}:
+  \item \textit{Unary operations}:
     \begin{itemize}
@@
     \end{itemize}
-  \item \textit{Binary relations:}
+  \item \textit{Binary operations:}
   \begin{itemize}

### 2018-05-11 3_encod.tex
Some additions (non-determinism, ...).
polish text a bit.

@@
 %may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to a SAT problem~\cite{Porthos17a} (more precisely, to an SMT-problem).
+as a \textit{bounded reachability} problem, which in turn can be reduced to a SMT problem~\cite{Porthos17a}.
+
+%TODO: sound, complete...
 
@@
 As the state machine model is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
-The general schema of model checking is the following: firstly, the analysing system is being represented as a transition system, a finite directed graph with labelled nodes representing states of the system such that each state corresponds to the unique subset of atomic propositions that characterise the behavioural properties of each state.
-Then, the system constraints are being defined in terms of a modal temporal logic with respect to the atomic propositions.
-Commonly, the Linear Temporal Logic~(LTL) or Computational Tree Logic~(CTL), along with their extensions, are used as a specification language due to the expressiveness and verifiability of their statements.
-In the described schema, the model checking problem is reducible to the reachability analysis, an iterative process of a systematic exhaustive search in the state space.
-This approach is called \textit{unbounded model checking (UMC)}.
 
+The general scheme of model checking is the following. 
+The analysing system (the \textit{model}) is represented as a transition system, a directed graph with labelled nodes representing states of the system.
+Each state corresponds to the unique subset of atomic propositions that characterise its behavioural properties.
+Once the model has been constructed, it can be checked for compliance to the \textit{specification}.
+
+Usually, the specification defines temporal constraints over the properties of the system.
+For instance, the specification assert may state that the property \textit{always} holds (the \textit{safety}) or the property will \textit{eventually} hold (the \textit{liveness}).
+Commonly, the \underline{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements.
+
+In the described scheme, the model checking problem is reducible to the reachability analysis, an iterative process of a systematic exhaustive search in the state space.
+This approach is called \textit{Unbounded Model Checking~(UMC)}.
 However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables used by the system (its size).
 In case of modelling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
-Therefore, the research in model checking over past 40 years was aimed at tackling the state explosion problem, mostly by optimising search space, search strategy or basic data structures of existing algorithms.
+Therefore, the research in model checking over past 40 years was fighting the state explosion problem mostly by optimising search space, search strategy or basic data structures of existing algorithms.
 
-One of the first techniques that optimises the search space considerably is the symbolic model checking with \textit{binary decision diagrams (BDDs)}.
-Instead of processing each state individually, in this approach the set of states is represented by the BDD, efficient data structure for performing operations on large boolean formulas~\cite{clarke2012model}.
+One of the first techniques that optimises the search space considerably is the symbolic model checking with \textit{Binary Decision Diagrams~(BDDs)}.
+Instead of processing each state individually, in this approach the set of states is represented by the BDD, a data structure that allow to perform operations on large boolean formulas efficiently~\cite{clarke2012model}.
 The BDD representation can be linear of size of variables it encodes if the ordering of variables is optimal, otherwise the size of BDD is exponential.
@@
 The other idea is to use satisfiability solvers for symbolic exploration of state space~\cite{clarke2001bounded}.
-In this approach, the state space exploration consists of sequence of queries to the SAT-solver, represented as boolean formulas that encode the constraints of the model and the finite path to a state in the corresponding transition system.
+In this approach, the state space exploration consists of the sequence of queries to the SAT-solver, represented as boolean formulas that encode the constraints of the model and the finite path to a state in the corresponding transition system.
 %This approach uses an iterative process of constructing queries to the SAT-solver as a boolean formula which encodes the constraints of the model and the finite path to a state in the corresponding transition system.
-Due to the SAT-solver.
-This technique is called \textit{bounded model checking (BMC)}, because the search process is being repeated up to user-defined bound $k$, which may result to incomplete analysis in general case.
+%Due to the SAT-solver.
+This technique is called \textit{bounded model checking (BMC)} as the search process is being repeated up to the user-defined bound $k$, which may result to incomplete analysis in general case.
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
@@
 
-In general, a BMC problem aims to examine the reachability of the "undesirable" states of a finite-state system.
+In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
 Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $Init(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $Trans(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $Bad(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
-Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps is formulated as following:
+Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
 %\begin{equation}
@@
 
-Portability analysis problem may also be stated as a reachability problem, where the undesirable state is the state reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
-However, unlikely the BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
-
+The portability analysis problem may also be stated as a reachability problem, where the undesirable state is one reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
+%However, unlikely a BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
 Consider the function $cons_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
@@
 The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:proc:x-transf}).%TODO: footnote why multiple sinks
-In control-flow graph, there are two types of transitions (edges): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
+In the control-flow graph, there are two types of transitions (edges): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
 Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
@@
 % TODO: attention to these non-breakable spaces
-Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \rf-relation (see Section~\ref{ch:wmm:model:relations} for details on relations).
-We distinguish the function ${\fv_g\;:\;\mathbb{C_g}\;\rightarrow\;\{0,1\}}$ that evaluates the guard computation event.
+Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it returns the \textit{set} of values determined by the \rf-relation; see Section~\ref{ch:wmm:model:relations} for details on relations).
+We distinguish the function ${\fv_g\;:\;\mathbb{C}_g\;\rightarrow\;\{0,1\}}$ that evaluates the guard computation event.
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
 
-Consider the following possible mutual arrangement of nodes in a control-flow graph:
-
-\begin{figure}[!hb]
+\begin{figure}[b]
   \centering
+  \begin{minipage}{\textwidth}
       \begin{subfigure}[b]{0.3\textwidth}
@@
     %
-    \caption{Linear and non-linear cases of control-flow graph}
+    \caption{Linear and non-linear cases of a control-flow graph}
     \label{encode:cf}
+\end{minipage}
 \end{figure}
 
-For listed cases, below we propose the encoding scheme that uniquely encodes each node of graph and allows to encode partially executed program.
+Consider the possible mutual arrangements of nodes in a control-flow graph presented in Figure~\ref{encode:cf}.
+For these cases, we propose the encoding scheme that uniquely encodes each node of graph and at the same time allows to encode partially executed program.
 Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
-Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1~\rightarrow~e_2), \; (e_1 ~\rightarrow~e_3)\}$.
-In encoding~\ref{enc:cf:merge} of the merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.
+Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1\,\rightarrow\,e_2), \; (e_1\,\rightarrow\,e_3)\}$.
+In Equation~\ref{enc:cf:merge}, which encodes the control-flow of a merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors has been executed, regardless the type of the transition.
 %TODO: maybe add the overall formula \phi_{CF} ?
@@
 
-For sake of encoding correctness, we require all branches to have at least one event.
-Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \texttt{nop}-event as it is shown in Figure~\ref{encode:branching:nop}:
+For sake of correctness of the encoding, we require all branches to have at least one event.
+Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \texttt{nop}-event as it is shown in Figure~\ref{encode:branching:nop}.
 
@@
 
-To encode the data-flow constraints, we use the \textit{static single-assignment (SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
+To encode the data-flow constraints, we use the \textit{Static Single-Assignment~(SSA) form} in order to be able to capture an arbitrary data-flow into a single SMT-formula.
 The SSA form requires each variable to be assigned only once within entire program.
-In contrast, \porthos{} used the dynamic single-assignment (DSA) form, that requires indices to be unique within a branch.
-Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow in merge points~\cite{Porthos17a}.
+In contrast, \porthos{} used the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique within a branch.
+Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow of merge points~\cite{Porthos17a}.
 
 Following~\cite{Porthos17b}, the indexed references of variables are computed in accordance with the following rules:
+%\begin{enumerate}[noitemsep,topsep=0pt]
 (1)~any access to a shared variable (both read and write) increments its SSA-index;
 (2)~only writes to a local variable increment its SSA-index (reads preserve indices);
-(3)~no access to a constant variable or computed (evaluated) expression changes their SSA-index.
+(3)~no access to a constant variable or computed (evaluated) expression changes their %SSA-index.
+%\end{enumerate}
 These rules determine the following encoding of load, store and computation events within single thread:
 %TODO: maybe add the overall formula \phi_{DF} ?
+%
 \begin{align}
@@
     \begin{algorithmic}[1]
-        \Function{Compute-SSA-map}{G}
+        \Function{Compute-SSA-Map}{G}
             \State {$S \gets$ empty map; $S[e_0] \gets$ empty map}
@@
 The encoding of this linkage left untouched as it is implemented in \porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
-
+%
 \begin{align}

### 2018-05-11 4_impl.tex
polish text a bit.

@@
 Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+%
 \footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
-), we consider the accomplished work as a step towards this.
+%
+), we consider the accomplished work as a step towards it.
 %This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
@@
 The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
+%
 \footnote{Guava project repository: \url{https://github.com/google/guava/}}%
+%
 ).
 
-During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
+During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.
 The following list summarises the main rules we followed during the development of \porthos[2]:
@@
 
-The robustness of analysis is the main criterion of \porthos[2] as a verification of a bug-hunting tool.
-This implies modelling the system with only over-approximating abstractions, which, however makes it more vulnerable to the combinatorial growth of the state space size, preserve the completeness of analysis.
-As a robust and transparent tool, the \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
+Robustness of the analysis is the main criterion of \porthos[2] as a verification tool.
+This implies modelling the system with usage of only over-approximating abstractions 
+However it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
+As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
 %The modular architecture of \porthos[2] allow each its module to be tested independently.
@@
 As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
+%
 \footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
+%
 by Microsoft Research~\cite{de2008z3}.
@@
 
-\begin{figure}%[H]
+\begin{figure}
   \centering
@@
 The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
+%
 \footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
 For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
+%
 (Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
@@
 Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the \tool{ANTLR} an attractive instrument for solving the parsing problem.
-
-Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.
 
@@
 
+Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.
 The input language parser used by \porthos{} suffered from several disadvantages.
 Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
-Such a combining of two rich languages%
-\footnote{by the term "reach" we mean "expressiveness": at least, the grammar of the language of semantic actions (i.e., java in our case) is Turing-complete.} %
-makes the code hardly understandable, and, therefore, poorly maintainable.
+Such a combining of two expressive languages makes the code hardly understandable and, therefore, poorly maintainable.
 In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the \tool{ANTLR} syntax tree to the AST, that is one for all languages of an input program.
@@
 Secondly, in \porthos{}, the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
+%
 \footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
+%
 , so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.
@@
 
-The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
+The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form `\lstinline{1 + 2 * 3}' could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
 Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
-tialisation of the shared variable and only by default value \lstinline{0}. The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
+tialisation of the shared variable and only by default value \lstinline{0}. \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
 
-The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
-\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}).%
+\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
+%
+\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}). %
+%
 Current version of \porthos[2] does not recognise C processor directives (it ignores them), however, in future it can be extended to support them.
@@
 The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
-The UML diagraph on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
+The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
 %The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring
@@
 
+%TODO!!! Remove watermark from the pictures!
 \begin{figure}[t]%[!b]%[H]
@@
 The X-type is a symbolic representation of the C primitive type%
-\footnote{Here we should note the \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++).
+%
+\footnote{Here we should note that \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++).
 In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language.
-See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
+See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %TODO!!!!!
+Check that XType is really a simple enumeration
+%
 that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType
@@
 \label{ch:impl:model:xgraph:invariants}
-Once being constructed, the graph must conform the following requirements:
+Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:port:enc:cf}):
 
-\begin{enumerate}
-\item graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
-\item graph must be connected,
-\item each node can have either one or two direct control-flow successors,
+\begin{enumerate}[noitemsep]
+\item the graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
+\item the graph must be connected,
+\item each node of the graph can have either one or two direct control-flow successors,
 \item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
 %\item each node except the source node must have at least one parent
-\item \textbf{todo: sth about data-flow edges}
+\item a \co-edge connects two writes, an \rf-relation edge connects a write and a read,
+\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
+\item all write-event nodes except final write-nodes must have exactly one \co-successor.
+%TODO: \item \textbf{TODO: more?}
 \end{enumerate}
@@
 
+<TODO: finish it later after fixing the code>
+
 %hashcode, uniqueness
@@
 
-a simple set of recursively defined relations over ..
+A \wmodel{} represents a recursively defined AST of relations defined by the memory model and an assertion expression.
+The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
+The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.
 
-atoms: sets/basic relations
+<todo: enumerate elements>
 
-itemize elements
+<carries the origin -- from cat file>
 
-todo: rename everywhere assertions <- axioms
+%todo: rename everywhere assertions <- axioms
 
@@
 
-in \porthos{} : ctx as an argument, calling ctx to create new clause. 
-drawbacks:
-- hard to debug
-- only one solver
-- non-safe way to construct formula (if ctx has changed, runtime excpetion)
+The \zformula{} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
+As all other internal representations, the \zformula{} is implemented as an immutable data structure.
 
-in \porthos[2] we created the new abstraction layer \zformula[] that is translated to
+Following the structure of a Z3-formula, the atomic elements of a \zformula{} are variables, and the expressions are
 
-recursive representation of SMT-formula
+reference to Z3-tutorial AND
 
-used as an abstraction btween
+All expressions and variables of a \zformula{} are typed.
+Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, floating point, <else?>) data type
 
-implemented as a simple wrapper over the z3 formula
-
-typed: Expr, BoolExpr, ArithExpr, ...
+%typed: Expr, BoolExpr, ArithExpr, ...
 
 carries the origin (X)
+
+<todo: enumerate elements>
 
@@
 
-the full \xgraph{} construction is performed in three stages.
+The construction of the \xgraph{} is performed in three stages.
 First, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
@@
 
+Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
+<TODO: SAY MORE: validation of ht einput elements during construction, maybe verif after building>
+
 % todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
@@
 The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
+%
 \footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
+%
 written in OCaml.
@@
 
+builder: filling in + verification
+
 \subsubsection{Y-tree constructor (4)}
@@
 - Y-Sytnax error
+
+The construction of a \zformula{} is performed by the \texttt{ZformulaBuilder}.%TODO: check this class existence
 
@@
 
-The \xgraph{} abstract machine model disposes infinite number of memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
+The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
 The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).

### 2018-05-12 1_intro.tex
Finish up to proc units.

@@
 The examples are
+%TODO: cite the 'wmm-general/Thread-Modular Static Analysis for Relaxed Memory Models.pdf'
 the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
-the program transformation tool~\cite{alglave2013software},
+the tool that performs the sequentialisation of concurrent programs~\cite{alglave2013software},
 the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.

### 2018-05-12 4_impl.tex
Finish up to proc units.

@@
     The control-flow event indicates a non-linear jump in the code. We distinguish two kinds of control-flow events: 
-    %\begin{itemize}[noitemsep]
     \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
@@
     If the function has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the function body.
+    The call stack is bounded by the user-defined parameter; once the stack is full, the interpretation continues without jumping to the body of the invoked function (such cases are properly logged).
+%TODO: call-stack: user-defined parameter
     Each \texttt{return} statement creates the assignment of the function call event on the top of call stack.
+    Note, this approach works for any kind of recursion, which is unrolled up the user-defined boundary as well as explicit loops.
     \\
-      If the function has not been resolved, \textbf{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
-    %\end{itemize}
+    According to the Rice's theorem~\cite{rice1953classes}, in general case there always will be functions unresolved within an analysis launch.
+    If the function has not been resolved, it can be safely assumed to be a no-operation function.
+    In other words, we suppose that the knowledge base of \porthos{} is complete and the tool can resolve all memory-operation functions and fence instructions.
+    Although this can affect the completeness of the analysis, it is safe to make such an assumption as the user can check the log file and manually analyse the semantics of an unresolved call and it to the knowledge base if necessary.
 
@@
   \1 \texttt{XFakeEvent} \\
-    The fake events are auxiliary elements of \xgraph{}.
+    The fake events are the auxiliary elements of \xgraph{}.
     %\begin{itemize}[noitemsep]
@@
 %\item each node except the source node must have at least one parent
-\item a \co-edge connects two writes, an \rf-relation edge connects a write and a read,
+\item a \co{}-edge connects two writes, an \rf{}-relation edge connects a write and a read,
 \item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
@@
 The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.
+For the sake of transparency, each element of a \wmodel{} contains the origin location as the coordinates of the string in the model file. %TODO: implement origin for w-model
 
-<todo: enumerate elements>
-
-<carries the origin -- from cat file>
+<todo: enumerate elements -- after refactoring>
 
@@
 
-Following the structure of a Z3-formula, the atomic elements of a \zformula{} are variables, and the expressions are
+The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
+Generally, a \zformula{} represents the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}. % the semantics of SMT-LIB formulas defined in~\cite[Chapter 5]{smt-lib}.
+Currently, the \zformula{} supports definitions and assertions over \textit{variables} and binary and unary \textit{expressions}.
+However, in future it may be extended to support \textit{function symbols} and \textit{binders} (existential and universal quantification, pattern matching and functional type construction).
 
-reference to Z3-tutorial AND
+All expressions of a \zformula{} are \textit{typed} (or \textit{sorted}, in terms of SMT-LIB standard).
+Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, bitvector, real) expressions.
+The typisation of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-LIB formula.
+For the sake of transparency, each element of a \zformula{} contains the origin location as the reference to the X-element that produced current Z-element. %TODO: implement origin for z-formula
 
-All expressions and variables of a \zformula{} are typed.
-Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, floating point, <else?>) data type
-
-%typed: Expr, BoolExpr, ArithExpr, ...
-
-carries the origin (X)
-
-<todo: enumerate elements>
+<todo: enumerate elements -- after completing>
 
@@
 Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
-<TODO: SAY MORE: validation of ht einput elements during construction, maybe verif after building>
+A builder is a mutable object that contains methods "filling" it by the elements that will constitute the result complex object.
+The methods contain basic validity checks of the elements.
+Once a builder has been built, it cannot be modified.
 
@@
 %The input memory model 
-\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}
+%TODO: add reference to the grammar file in github repository OR include a short version as an appendix}
 
@@
 
-%<not implemented yet>
+%TODO: <not implemented yet>
+The \wmodel{} is constructed by the converter \texttt{Cat2WmodelConverter}, which calls \texttt{Cat2WmodelConverterVisitor}, the stateless visitor of memory-model ANTLR syntax tree.
+%TODO: Implement this!!!
+Currently, the visitor supports only the declarative subset of the \cat{} language, and does not process the functional-style definitions (it can to be supported by implementing the
 
@@
 Following the C standard, local memory units have higher priority over global ones.
-Since the C language allows the variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the block-context stack. %TODO: refactor XMemoryManager!
+Since the C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block block-contexts for local variables.
+%TODO: refactor XMemoryManager!
 
@@
 %TODO: say sth about the mode
-The hook manager operates multiple invocation hooks depending on the user-defined mode (
- \texttt{XInvocationHook}
+The hook manager operates multiple invocation hooks depending on the user-defined mode.
+All invocation hooks implement the interface \texttt{XInvocationHook}.
+Current version of \porthos[2] contains two invocation hooks, the \texttt{} and \texttt{XKernelInvocationHook} that contains ...............................
 
 %TODO: rename InvocationHook -> XInvocationHook
-multiple hooks for mult modes
 %, having access to the X-interpreter (see Section~\ref{ch:impl:proc:x-compiler:compilation}),

### 2018-05-13 3_encod.tex
complete some todos.

@@
-\chapter{Portability analysis as an SMT~problem}%Description of the method used by Porthos
+\chapter{Portability analysis as an SMT~problem}
 \label{ch:port}
@@
 For instance, the specification assert may state that the property \textit{always} holds (the \textit{safety}) or the property will \textit{eventually} hold (the \textit{liveness}).
-Commonly, the \underline{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements.
+Commonly, the \textit{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements.
 
@@
 Note that the definition of portability requirements against \textit{executions} is strong enough, as it implies the portability against \textit{states} (the \textit{state-portability})~\cite{Porthos17b}.
-The result SMT-formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
+The result SMT-formula $\phi$ that encodes the portability problem should contain both encodings of control-flow $\phi_{CF}$ and data-flow $\phi_{DF}$ of the program, and assertions of both memory models: ${\phi \defeq \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$.
 If the formula is satisfiable, there exist a portability bug.
@@
 
-The control-flow of a program is represented in the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} (see Section~\ref{ch:impl:proc:x-transf}).%TODO: footnote why multiple sinks
-In the control-flow graph, there are two types of transitions (edges): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
+The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes. %TODO: footnote why multiple sinks
+In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
 Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
@@
 For instance, \porthos{} uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
-$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
+$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
 and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
-$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
+$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
 (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
@@
 \begin{align}
-\phi_{CF_{seq}}   = \ & \ \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    = \ & \ [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
+\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
+\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
                       & \ [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
                       & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
-\phi_{CF_{mer}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
+\phi_{CF_{mer}} \defeq \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
 \end{align}
@@
 \begin{align}
-    \phi_{DF_{e = \mathtt{load}(r \leftarrow l)}} = \ & \fx(e) \rightarrow (r_{i+1} = l_{i+1}) \\
-    \phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} = \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
-    \phi_{DF_{e = \mathtt{eval}(...)}} = \ & \fx(e) \rightarrow \fv(e) % \\
+    \phi_{DF_{e = \mathtt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \rightarrow (r_{i+1} = l_{i+1}) \\
+    \phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
+    \phi_{DF_{e = \mathtt{eval}(\cdot)}} \defeq \ & \fx(e) \rightarrow \fv(e) % \\
 \end{align}
@@
 \begin{align}
-\phi_{DF_{mem}}(e_1, e_2) = \ & \mathtt{rf}(e_1, e_2) \rightarrow (l_i = l_j) 
+\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \mathtt{rf}(e_1, e_2) \rightarrow (l_i = l_j) 
 \end{align}
@@
 
-todo
+The basic scheme for encoding the memory model is proposed in~\cite{Porthos17a}.
+The encoding consists of two parts: encoding the \textit{derived relations} and encoding the memory model \textit{assertions}.
+
+In an SMT-formula, the relation $x \relation{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
+The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
+\begin{itemize}[noitemsep,topsep=0pt]
+\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$
+\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$
+\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$
+\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$
+\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$
+\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb(E)}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$
+\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), where \\
+\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), and \\ 
+\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$
+\end{itemize}
+
+Note that the \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
+The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.
+ %is used for guessing the execution
+
+The memory model can assert acyclicity, irreflexivity of emptiness of a relation or a set of events.
+As it has been proposed in~\cite{Porthos17a}, encoding the acyclicity assertion uses numerical variable $\Psi_e \in \mathbb{N}$ for each event $e$ in the relation to be asserted: $acyclic(r) \defeq (r(e_1,e_2) \Rightarrow (\Psi_{e_1} < \Psi_{e_2}))$.
+The irreflexivity assertion as $irreflexive(r) \defeq \bigvee\limits_{e_k \in \mathbb(E)}^{} \lnot r(e_k,e_k)$.
+
+%The ... \wmodel{}
+
 %Most bounded model-checking problems are solved via computing the fixpoint for the

### 2018-05-13 4_impl.tex
complete some todos.

@@
 This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+Yet far from all of C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
 %
@@
     If the function has not been resolved, it can be safely assumed to be a no-operation function.
-    In other words, we suppose that the knowledge base of \porthos{} is complete and the tool can resolve all memory-operation functions and fence instructions.
+    In other words, we suppose that the knowledge base of \porthos[2] is complete and the tool can resolve all memory-operation functions and fence instructions.
     Although this can affect the completeness of the analysis, it is safe to make such an assumption as the user can check the log file and manually analyse the semantics of an unresolved call and it to the knowledge base if necessary.
@@
 
-A \wmodel{} represents a recursively defined AST of relations defined by the memory model and an assertion expression.
+A \wmodel{} represents a recursive AST of \textit{computations over relations} and assertion expressions defined by the memory model.
 The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
@@
 %TODO: <not implemented yet>
-The \wmodel{} is constructed by the converter \texttt{Cat2WmodelConverter}, which calls \texttt{Cat2WmodelConverterVisitor}, the stateless visitor of memory-model ANTLR syntax tree.
-%TODO: Implement this!!!
-Currently, the visitor supports only the declarative subset of the \cat{} language, and does not process the functional-style definitions (it can to be supported by implementing the
+The \wmodel{} is constructed by the stateless visitor \texttt{Cat2WmodelConverterVisitor} from the ANTLR syntax tree \texttt{CatParser.MainContext} of the memory-model defined in \cat{} language.
+Currently, the visitor supports the small subset of the \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full OCaml interpreter.
+However, some most commonly used functional-style expressions may be supported by mapping them syntactically to corresponding W-elements by the \wmodel{} constructor.
 
-visitor of antlr tree: almost direct
+Following the principle of transparency, the \wmodel{} constructor aborts its work with the \texttt{NotSupportedException} if received the unrecognised syntax construction (in contrast to the \porthos[1] approach in which the \texttt{null} value was produced in all exceptional states).
 
-todo: file tree of w-model to appendix
-
-in perspective: work with functional-style definitions
-
-builder: filling in + verification
+<todo: file tree of w-model to appendix -- after code fixing>
 
@@
 
-from \tool{ANTLR} syntax tree of C
+The \ytree{} is constructed by the stateless visitor \texttt{C2YtreeConverterVisitor} from the ANTLR syntax tree \texttt{C11Parser.MainContext} of C language.
+As the \wmodel{} constructor, the \ytree{} constructor aborts its work with the \texttt{NotSupportedException} once it meets an unsupported syntax construction.
+%TODO: Rename YConverterException <- YParserException
+A syntax exception \texttt{YConverterException} is thrown if the converted syntax tree contains semantic errors that prevent it to be converted to the \ytree{}.
 
-- desugar, equiv transform
-
-- variables distinguished from
-
-- if we don't support , our parser still parses it, and the error is thrown at the moment of converting syntax tree to the AST (Y-tree).
-
-- So, The language-dependent syntax tree is converted to the AST by the stateless Visitor (e.g., for C11 -> Ytree conversion is made by 'C2YtreeConverterVisitor') + short structure of this visitor (how?..
-need ly?)
-
-- Y-Sytnax error
-
-The construction of a \zformula{} is performed by the \texttt{ZformulaBuilder}.%TODO: check this class existence
+As a \ytree{} constitutes a generic AST, the \ytree{} constructor performs code desugaring (for example, a \texttt{switch}-statement is converted to the equivalent \texttt{if}-statement).
 
@@
 
-The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
+C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
 Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
@@
 For example,
-Being carried at Y- and X- representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
+Being carried at Y- and X- representation levels, the type is converted to a~Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
 
@@
 In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter}, %TODO: Or XTypeHookManager?
-which can be requested for converting an X-type to a Z-type. %TODO: implement this.
+which can be requested for converting an X-type to a~Z-type. %TODO: implement this.
 %TODO: describe the XType (before) + implement this clearly! + definitely get rid of bit32
@@
 Following the C standard, local memory units have higher priority over global ones.
-Since the C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block block-contexts for local variables.
+Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block block-contexts for local variables.
 %TODO: refactor XMemoryManager!

### 2018-05-13 2_wmm.tex
complete some todos.
write up to Interpreter.

@@
   \includegraphics[width=1.2\linewidth]{img/my/sb-example/SB-dot.png}
-  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=1)}}
+  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
   \label{simple_wmm_x86_pic:sub1}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-2.png}
-  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=0)}}
+  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=0)}}
   \label{simple_wmm_x86_pic:sub2}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-3.png}
-  \caption{Final state: \texttt{(0:EAX=1~/\textbackslash~1:EAX=1)}}
+  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
   \label{simple_wmm_x86_pic:sub3}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-4.png}
-  \caption{Final state: \texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}}
+  \caption{Final state: \texttt{(0:EAX=0,~1:EAX=0)}}
   \label{simple_wmm_x86_pic:sub4}
@@
 
-\section{The \cat{} language}
+\section{\cat{} language}
 \label{ch:wmm:cat}
 
-Weak memory models are defined via the \cat{} language~\cite{alglave2016syntax}.
+Weak memory models are defined via \cat{} language~\cite{alglave2016syntax}.
 This is a domain specific language for describing consistency properties of concurrent programs.
-The \cat{} language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, operations and assertions that are specific for operating with relations and executions.
+The language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, operations and assertions that are specific for operating with relations and executions.
 %represents the functional language extended by the theory of  allows to axiomatically define derived relations, architecture-specific fences, and constraints over relations.
@@
     \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2},
-    \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1\,\backslash\,r2}$, and
+      \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1 \relminus r2}$, and
     \item \textit{the sequence} of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}, which is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$.

### 2018-05-13 4_impl.tex
write up to Interpreter.

@@
 The visitor pattern performs the double-dispatching mechanism: the visiting class defines the method for accepting a visitor, and the visitor itself defines the 
+<TODO!!!!>
+
 To visit an instance \lstinline{x} by a visitor \lstinline{v}, the program should call the accept-method of \lstinline{x} and pass \lstinline{v} as an argument.
@@
 The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
-The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
+The \tool{ANTLR} grammar for \cat{} language was extracted from the parser used by \tool{herd} tool%
 %
@@
 %TODO: <not implemented yet>
-The \wmodel{} is constructed by the stateless visitor \texttt{Cat2WmodelConverterVisitor} from the ANTLR syntax tree \texttt{CatParser.MainContext} of the memory-model defined in \cat{} language.
-Currently, the visitor supports the small subset of the \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full OCaml interpreter.
+The \wmodel{} representation is constructed by the stateless visitor \texttt{Cat2WmodelConverterVisitor} from the ANTLR syntax tree \texttt{CatParser.MainContext} of the memory-model defined in \cat{} language.
+Currently, the visitor supports the small subset of \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full OCaml interpreter.
 However, some most commonly used functional-style expressions may be supported by mapping them syntactically to corresponding W-elements by the \wmodel{} constructor.
 
-Following the principle of transparency, the \wmodel{} constructor aborts its work with the \texttt{NotSupportedException} if received the unrecognised syntax construction (in contrast to the \porthos[1] approach in which the \texttt{null} value was produced in all exceptional states).
+Following the principle of transparency, the \wmodel{} constructor aborts its work with the \texttt{NotSupportedException} if met the unrecognised syntax construction (in contrast to the \porthos[1] approach in which the \texttt{null} value was produced in all exceptional states).
 
@@
 Currently, \porthos[2] handles only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
-Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
+The array dimension is stored at the Y-level as an integer number (currently this information is not used as \porthos[2] does not support arrays, although this information can be used at the stage of pre-compilation for converting array elements to separate variables).
+%the array operations are converted to ).
 In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter}, %TODO: Or XTypeHookManager?
@@
 
-\vspace{1em}
-The following types of variables are detected as \textit{global}:
+%\vspace{1em}
+The following types of variables are detected as \textit{global variables}:
 %TODO: not implemented
@@
 The X-compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
+It is a complex processing unit; Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
+
+<TODO: in picture change: xmemorymaager belongs to the program interpreter!>
 
@@
 
-The X-compiler is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
+The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it on Figure.
+It receives as input the \ytree{}, the memory model kind and user settings (for instance, the interpreter mode defining the set of invocation hooks enabled during the analysis run).
+The \texttt{Y2XConverter} creates an instance of the stateless visitor \texttt{Y2XConverterVisitor} that traverses the \ytree{} while invoking the stateful interpreter \texttt{XInterpreter}.
+The \texttt{XInterpreter} carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the filling in the builder.
+The interpreter requests the managers such as \texttt{XMemoryManager}, \texttt{XHookManager}, \texttt{XTypeManager} for additional information.
 
-The X-interpreter is a stateful processing unit that carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the state of the builder.
+<TODO: ADD XTypeManager to the picture!>
 
-X compilation error
+We distinguish three types of processes (each implement the \texttt{XInterpreter} interface):
+\begin{itemize}[noitemsep,topsep=0pt]
+\item the prelude process \texttt{XPreludeInterpreter} (to be executed before all other processes) that allows only declaration of local or shared variables, computations and memory operations;
+\item the postlude process \texttt{XPostludeInterpreter} (to be executed after all other processes) that allows only declaration of local variables, memory operations, computations and assertions;
+\item the process \texttt{XProcessInterpreter} (to be executed in parallel) that allows all X-compiler interface operations except declaration of shared variables %(that should be declared during pre-compilation stage or during compilation of the prelude process)
+and program assertions (declarations of local variables, computations, barriers, unconditional jumps, non-linear branching statements, method calls).
+\end{itemize}
 
-<scheme of who calls whom>
-
-%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
-%reutrn event= assign the return-register + simple jump
+As the prelude process can declare shared variables, its compilation must go first.
+%TODO: think about sorting processes at the pre-compilation stage
+For transparency reasons, the first event of a process must be the \texttt{XEntryEvent} and the last events must be events of type \texttt{XExitEvent}.
 
@@
 
+%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
+%reutrn event= assign the return-register + simple jump
+
 The invocation hooking module serves as a knowledge base that stores the semantics of functions.
@@
 Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in case of program in C language -- only the function name) across all signatures it stores.
-Once the signature matches, the hook manager intercepts the compilation process and interprets the method according to the algorithm it stores.
+If the signature matches, the hook manager intercepts the function call and interprets the hook action instead of invoking the function directly.
 %TODO: say sth about the mode
@@
 All invocation hooks implement the interface \texttt{XInvocationHook}.
-Current version of \porthos[2] contains two invocation hooks, the \texttt{} and \texttt{XKernelInvocationHook} that contains ...............................
+The result of an invocation hook interception is the \texttt{XInvocationHookAction}, a delayed operation implemented on the top of lambda functions of Java.
+The hook action is invoked with actual arguments and returns an arbitrary \texttt{XEntity} as the result of invocation.
 
+Current version of \porthos[2] contains two invocation hooks, the \texttt{XPorthosInvocationHook} for describing the compatibility-mode functions of the input \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
+An invocation hook can model any type of operations (memory, fence, computational, etc.).
+For instance, the \texttt{XPorthosInvocationHook} intercepts calls \texttt{`location.store(atomic, register)'} and processes the store operation with respect to the specified atomic.
+%TODO: more about described functions
 %TODO: rename InvocationHook -> XInvocationHook
-%, having access to the X-interpreter (see Section~\ref{ch:impl:proc:x-compiler:compilation}),
 
-lambda-hook: knowledge base: InterceptionAction(BiFunction<XMemoryUnit, XMemoryUnit[], ? extends XEntity> action)
-
-todo: kernel hooks
 %TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html
@@
 stateful
+
+X compilation error
+
+contexts
+
+computation: how not to collect (in which case we loose)
 

### 2018-05-14 4_impl.tex
write interpreter. update pictures.

@@
 
+Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
+However, all three types of \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
+%TODO: rename to XThread ?
+This simplifies drastically the processing of different types of code blocks as they all are modelled by the same data structure.
+The examples of restrictions on X-graphs are the following: the initialisation block can not have branchings, fence events; the process cannot have assertion events; the assertion block can not have shared-memory events or fence events.
+We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler:compilation}.
+
 \paragraph{Memory units}
@@
 that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType
+As the type of an X-memory unit may not be resolved correctly, the memory units keep this information as a boolean flag.
 
@@
 For more detailed description of memory management see Section~\ref{ch:impl:proc:x-compiler:mem}.
-
-Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
-However, all three types of \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
-%TODO: rename to XThread ?
-This simplifies drastically the processing of different types of code blocks as they all are modelled by the same data structure.
-The examples of restrictions on X-graphs are the following: the initialisation block can not have branchings, fence events; the process cannot have assertion events; the assertion block can not have shared-memory events or fence events.
-We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler:compilation}.
 
@@
     \2 \texttt{XExitEvent}, the process sink event. There are two kinds of sink events: \textit{complete} and \textit{incomplete} (see description in Section~\ref{ch:impl:proc:x-unroll}),
-    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}).
+    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}),
+    \2 \texttt{XAssertionEvent}, the reachability assertion made at the postlude statement of the program.
+    An assertion is modelled as an event for the purpose of encoding the postlude statement as a separate process that is compatible with the encoder.
    % \end{itemize}
@@
 
-<TODO: in picture change: xmemorymaager belongs to the program interpreter!>
-
 \begin{figure}[!t]%[H]
@@
 
-<TODO: ADD XTypeManager to the picture!>
-
-We distinguish three types of processes (each implement the \texttt{XInterpreter} interface):
+We distinguish three types of processes (each implementing the \texttt{XInterpreter} interface):
 \begin{itemize}[noitemsep,topsep=0pt]
@@
 \item the process \texttt{XProcessInterpreter} (to be executed in parallel) that allows all X-compiler interface operations except declaration of shared variables %(that should be declared during pre-compilation stage or during compilation of the prelude process)
-and program assertions (declarations of local variables, computations, barriers, unconditional jumps, non-linear branching statements, method calls).
+and program assertions (i.e., it allows declarations of local variables, computations, barriers, unconditional jumps, non-linear branching statements, and method calls).
 \end{itemize}
@@
 However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
+
+The interface methods exposed by the \texttt{XMemoryManager} are presented in Figure~\ref{fig:mem-manager}.
 At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
-Following the C standard, local memory units have higher priority over global ones.
-Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block block-contexts for local variables.
+Following the C standard, local memory units have higher priority over global ones (the method \texttt{getDeclaredUnitOrNull} returns the first memory unit found, either a global one or a local one, or null if no memory units with requested name have been registered).
+Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block contexts for local variables.
 %TODO: refactor XMemoryManager!
 
-<some picture ? >
+\begin{figure}[h]
+\centering
+\lstinputlisting[language=java]{inc/lst/XMemoryManager.java}
+\caption{X-memory manager public interface}
+\label{fig:mem-manager}
+\end{figure}
 
@@
 An invocation hook can model any type of operations (memory, fence, computational, etc.).
-For instance, the \texttt{XPorthosInvocationHook} intercepts calls \texttt{`location.store(atomic, register)'} and processes the store operation with respect to the specified atomic.
+For instance, the \texttt{XPorthosInvocationHook} intercepts calls \texttt{`location.store(atomic,register)'} and processes the store operation with respect to the specified atomic.
 %TODO: more about described functions
@@
 
-stateful
+%In this paragraph we discuss only the implementation of the process interpreter \texttt{XProcessInterpreter}.
+%This interpreter is used to build the branched event-flow graph with all kinds of jumps, memory operations and fences.
+The X-program interpreter is invoked by the \texttt{Y2XConverterVisitor} which walks down the recursive syntax tree \ytree{}.
+The calls to the X-program interpreter are dispatched to currently maintained X-process interpreter.
+To be able to properly recognise nested statements of the \ytree{}, the X-process interpreter needs to have stacks.
+On any semantic error, the X-interpreter throws an \texttt{XInterpretationError}.
 
-X compilation error
+The interface methods of the \texttt{XInterpreter} are presented in Appendix~\ref{apx:xinterpreter}.
+Note the clear modular independence of the X-interpreter as it has no methods that operate entities of the Y-level (all conversion work with Y-entities is done by the \texttt{Y2XConverterVisitor}).
 
-contexts
+All interface methods can be divided into two groups.
+The first group constitute methods that emit events.
+These methods construct an event object (as events contain the \texttt{XEventInfo} structure that stores information about the owning process, all events must be created by the process constructor, i.e.
+X-interpreter) and change the state of interpreter considering the newly created event.
+Note that the methods \texttt{createComputationEvent} do not emit a computation event but only create one.
+This is performed as optimisation that removes unused computations from the model (otherwise, for instance, the assignment 
+`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- 1 * (2 + 3));}' instead of single event `\texttt{write(register <- 1 * (2 + 3));}').
+If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is not used by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph:evt}).
+The second group of X-interpreter methods consists of the methods for defining non-linear statements (branchings and loops).
+These methods change the state of the interpreter and set up additional non-linear control-flow edges.
+%`\texttt{register = 1 * (location + 2)}' would be compiled to three consequent events `\texttt{load(temp\_reg\_1 <- location); eval(temp\_reg\_1 + 2); write(register <- 1 * (temp\_reg\_1 + 2));}',
 
-computation: how not to collect (in which case we loose)
+As a high-level instruction (Y-level) may be compiled into a sequence of low-level instructions (for example, a computation that involves shared variables should firstly load them into the local memory and then process the computation event over local-only memory), the interpreter must maintain the \textit{stack of contexts} and remember the \textit{previous event} to be able to correctly process nested non-linear statements of C language.
+The context is a data structure that carries the \textit{state} and some additional information in case of processing non-linear statements (e.g., conditional event, first and last then- and else-branch events for binding, etc.).
+Once the new event has been emitted, the interpreter sets up control-flow edges w.r.t.
+state of the context on the top of the stack (the context stack is always non empty: the bottom context is a linear one).
+The context state is an enumeration of the following values:
+\begin{itemize}
+\item \texttt{WaitingAdditionalCommand}: the interpreter is in the state of defining the complex (non-linear) statement and is not able to process any new event (will throw an exception if any) until the state is not changed,
+\item \texttt{WaitingFirstSubBlockEvent}, the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event),
+\item \texttt{WaitingNextLinearEvent}, the standard interpreter state for processing next linear event,
+\item \texttt{Idle}, the interpreter does not set up the edge from the previous event to the new event.
+\end{itemize}
 
-stacks: context, readyContext, call, 
-  BlockContext (rename): what stores, enum of states
+Once the (non-linear) context has been popped out of the context stack (by the methods of the second group that finish non-linear branch definition) and added to the queue of \texttt{almostReadyContexts}.
+The almost-ready-context becomes a ready-context (i.e., it moves to the queue \texttt{readyContexts}) once the non-linear statement definition has finished (the method \texttt{finishNonlinearBlockDefinition}).
 
-almostReadyContexts: for computation events. Flushed by invocation . When are filled.
+Before processing a newly emitted event, the interpreter checks whether the queue \texttt{readyContexts} is not empty.
+If so, it iterates over all ready-contexts and sets up control-flow edges considering that the newly emitted event is an exit-event" of the non-linear context (e.g., for a branching context the interpreter adds primary edges from condition-event to the first-true-branch-event and from the last-true-branch-event to the exit-event, the same is done with alternative edges for the false-branch).
 
-setting up edges depending of type of ready-context
+%The \texttt{almostReadyContexts} queue is necessary for not adding erroneous edges on
 
-- hierarchy of Compilers: pic on inheritance (XCompiler is an stateful abstract machine). Restrictions of lude-compilers
-  postlude: assertion: a new event XAssertionEvent
+For simplicity of the interpretation, we keep the jump events (no-computation events) in the flow-graph, however they can be removed and the edges to and from them can be rewritten if necessary.
 
-interfaces that offer compilers: listing of interface
+%algorithm for interpreting
 
-algorithm for interpreting
+%The method \texttt{tryConvertToLocalOrNull} makes an attempt to convert an arbitrary X-entity to the local memory unit: registers and computation events are casted and returned by the method as-is, shared memory units are copied to a temporary register (\texttt{declareTempRegister}), which is returned by the method.
+%In other cases, the method returns \texttt{null}.
 
@@
 
-up to bound $k$
+The compiled graph needs to be unrolled up to the user-defined bound $k$.
+Comparing to \porthos[1], we have change the meaning of the bound: instead of executing all cycles at most $k$ times, the \porthos[2] interprets the bound as the maximum number of Y-instructions to be executed after the unrolling.
+The meaning of a bound was changed in order to increase predictability of the size of the result SMT-formula.
 
-recursion -- may be, if calls are supported
+Note, recursive function calls create unconditional jumps back in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
 
@@
 
-Z-type : SMT-specific (say about smt-lib)
+todo
 
@@
 
-\subsubsection{SMT-formula translator (11)}
-\label{ch:impl:proc:smt-solver}%TODO: ref it in pre-word in this chapter
+As the \zformula{} follows the SMT-LIB standard, its translation to an SMT-formula constitutes almost one-to-one mapping.
+The translation is performed by the stateless visitor \texttt{Z2SmtTranslator}.
+
+\subsubsection{SMT solver (11)}
+\label{ch:impl:proc:smt-solver}
 %TODO: check that we have ALL references in pre-word in this chapter
-
-%todo: mention The SMT-LIB standard http://smtlib.cs.uiowa.edu/papers/smt-lib-reference-v2.6-r2017-07-18.pdf
+As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT-solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
 
@@
 
-structure: verdict
-
-%TODO: say sth about syntax errors handling: still now, what if parser does not accept syntax?
+The result of execution of \porthos[2] is the verdict modelled by the class \texttt{AppVerdict}.
+This is a structure that contains the result of analysis and auxiliary information such as collected errors and time of execution.
+The app verdict may be rendered to any format convenient for the user.
 
@@
 \label{ch:impl:aux}
+
+There are some ...
 

### 2018-05-14 inc/lst/XInterpreter.java

@@ -1,66 +1,56 @@
 public interface XInterpreter {
+  XProcessId getProcessId();
+  void finish();
+  XCyclicProcess getResult();
+  
+  // linear interpretation methods:
+  XEntryEvent emitEntryEvent();
+  XExitEvent emitExitEvent();
+  XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
+  XJumpEvent emitJumpEvent();
+  XNopEvent emitNopEvent();
+  XAssertionEvent emitAssertionEvent(XBinaryComputationEvent assertion);
+  
+  XLocalMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
+                                    XLocalMemoryUnit source);
+  XSharedMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
+                                     XSharedMemoryUnit source);
+  XSharedMemoryEvent emitMemoryEvent(XSharedLvalueMemoryUnit destination,
+                                     XLocalMemoryUnit source);
 
-    XProcessId getProcessId();
-    XCyclicProcess getResult();
-    void finishInterpretation();
+  // create computation event without processing it:
+  XComputationEvent createComputationEvent(XUnaryOperator operator,
+                                           XLocalMemoryUnit operand);
+  XComputationEvent createComputationEvent(XBinaryOperator operator,
+                                           XLocalMemoryUnit operand1, 
+                                           XLocalMemoryUnit operand2);
 
-    // memory manager interface methods:
-    XLocation declareLocation(String name, XType type);
-    XRegister declareRegister(String name, XType type);
-    XRegister declareTempRegister(XType type);
-    XLvalueMemoryUnit declareUnresolvedUnit(String name, boolean isGlobal);
-    XLvalueMemoryUnit getDeclaredUnitOrNull(String name);
-    XRegister getDeclaredRegister(String name, XProcessId processId);
-    XLocalMemoryUnit tryConvertToLocalOrNull(XEntity expression);    
-    XComputationEvent tryEvaluateComputation(XEntity entity);
-    XRegister copyToLocalMemory(XSharedMemoryUnit shared);
-
-    // linear interpretation methods:
-    XEntryEvent emitEntryEvent();
-    XExitEvent emitExitEvent();
-    XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
-    XJumpEvent emitJumpEvent();
-    XNopEvent emitNopEvent();
[…]

### 2018-05-14 inc/lst/XMemoryManager.java

@@ -1,0 +1,13 @@
+interface XMemoryManager {
+  XLocation declareLocation(String name, XType type);
+  
+  XRegister declareRegister(String name, XType type);
+  
+  XRegister declareTempRegister(XType type);
+  
+  XLvalueMemoryUnit declareUnresolvedUnit(String name, boolean global);
+  
+  XLvalueMemoryUnit getDeclaredUnitOrNull(String name);
+  
+  XRegister getDeclaredRegister(String name, XProcessId processId);
+}

### 2018-05-14 4_impl.tex
write the unrolling.

@@
     \2 \texttt{XEntryEvent}, the per-process unique source event in the event-flow graph,
-    \2 \texttt{XExitEvent}, the process sink event. There are two kinds of sink events: \textit{complete} and \textit{incomplete} (see description in Section~\ref{ch:impl:proc:x-unroll}),
+    \2 \texttt{XExitEvent}, the process sink event,
     \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}),
@@
 \item \texttt{WaitingAdditionalCommand}: the interpreter is in the state of defining the complex (non-linear) statement and is not able to process any new event (will throw an exception if any) until the state is not changed,
-\item \texttt{WaitingFirstSubBlockEvent}, the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event),
-\item \texttt{WaitingNextLinearEvent}, the standard interpreter state for processing next linear event,
-\item \texttt{Idle}, the interpreter does not set up the edge from the previous event to the new event.
+\item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event),
+\item \texttt{WaitingNextLinearEvent}: the standard interpreter state for processing next linear event,
+\item \texttt{Idle}: the interpreter does not set up the edge from the previous event to the new event.
 \end{itemize}
@@
 
-The compiled graph needs to be unrolled up to the user-defined bound $k$.
+In order to be encoded into an SMT-formula, the compiled graph needs to be acyclic~\cite{Porthos17b}.
+To convert it to an acyclic form, we perform the \textit{unrolling} (sometimes called \textit{unwinding}) transfomation: each cycle is unrolled up to the user-defined bound $k$.
 Comparing to \porthos[1], we have change the meaning of the bound: instead of executing all cycles at most $k$ times, the \porthos[2] interprets the bound as the maximum number of Y-instructions to be executed after the unrolling.
 The meaning of a bound was changed in order to increase predictability of the size of the result SMT-formula.
-
 Note, recursive function calls create unconditional jumps back in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
 
-two kinds of exits
-
-- unrolling: why we cannot encode cyclic structures. reference to the paper (see arXiv version)
-% If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph.
-Although, in order to be able to encode such a cyclic structure as a boolean formula, this graph needs to be
-
-simple algorithm
-
-setting up backward edges
-
-how we attempted, why it didn't work in general case:
-
-why we changed the notion of unrolling bound
-
-sth about hashcodes and collections
-
-- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
+Considering the new meaning of the unrolling bound, we create sink events of two types: the \textit{complete} and \textit{incomplete} sink event.
+The first one means that the loop has been executed a whole number of times, and the latter means that the unrolling bound triggered on not-the-last event of the loop.
+The Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, while the $S_{-}$ -- incomplete).
 
 %TODO: draw (and implement) diff types of exit nodes
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 ;
-\node[c] (21) [right=3cm of impl] {$2_1$};
-\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
-\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
-\node[c] (41) [below of=31] {$4_1$};
-\node[c] (12) [below of=21] {$1_2$};
-\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
-\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
-\node[c] (13) [below of=41] {$1_3$};
-\node[c] (14) [below of=22] {$1_4$};
+\node[c] (22) [right=3cm of impl] {$2_2$};
+\node[c] (11) [above right=1cm and 1cm of 22]{$1_1$};
+\node[c] (32) [below right=1cm and 1cm of 11] {$3_2$};
 \node[c] (43) [below of=32] {$4_3$};
-\node[c] (23) [below of=13] {$2_3$};
-\node[c] (33) [below of=13] {$3_3$};
-\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
-\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
-\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
-\node[c] (44) [below of=33] {$4_4$};
-\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
-\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
-\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
-\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
-\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
-\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
-\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
+\node[c] (13) [below of=22] {$1_3$};
+\node[c] (24) [below left=1cm and 1cm of 13] {$2_4$};
+\node[c] (34) [below right=1cm and 1cm of 13] {$3_4$};
+\node[c] (14) [below of=43] {$1_4$};
+\node[c] (15) [below of=24] {$1_5$};
+\node[c] (45) [below of=34] {$4_5$};
+\node[c] (25) [below of=14] {$2_5$};
+\node[c] (35) [below of=14] {$3_5$};
+\node[c] (26) [below left=1cm and 1cm of 15] {$2_6$};
+\node[c] (36) [below right=1cm and 1cm of 15] {$3_6$};
+\node[c] (16) [below right=1cm and 0.2cm of 45] {$1_6$};
+\node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
+\node[rectangle,draw] (6inc) [below left=1cm and 0.6cm of 16] {$S_{-}$};
+\node[rectangle,draw] (6com) [below of=6inc] {$S_{+}$};
+\node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
+\node[] (k2) [right=1.7cm of 32] {$(k = 2)$};
+\node[] (k3) [right=1.7cm of 43] {$(k = 3)$};
+\node[] (k4) [right=1.7cm of 14] {$(k = 4)$};
+\node[] (k5) [right=1.7cm of 35] {$(k = 5)$};
+\node[] (k6) [right=1cm of 46] {$(k = 6)$};
 \path[->]
-(11) edge [] node {} (21)
-(11) edge [] node {} (31)
-(31) edge [] node {} (41)
-(21) edge [] node {} (12)
-(12) edge [bend right=10] node {} (22)
-(12) edge [bend left=10] node {} (32)
-(41) edge [] node {} (13)
-(22) edge [] node {} (14)
+(11) edge [] node {} (22)
+(11) edge [] node {} (32)
 (32) edge [] node {} (43)
-(13) edge [] node {} (23)
-(13) edge [] node {} (33)
-(14) edge [bend right=10] node {} (24)
-(14) edge [bend left=10] node {} (34)
-(43) edge [] node {} (15)
-(33) edge [bend right=10] node {} (15)
-(33) edge [bend left=10] node {} (44)
-(24) edge [bend right=20] node {} (6)
-(34) edge [] node {} (6)
-(15) edge [] node {} (6)
-(44) edge [bend left=20] node {} (6)
+(22) edge [] node {} (13)
+(13) edge [bend right=10] node {} (24)
+(13) edge [bend left=10] node {} (34)
+(43) edge [] node {} (14)
+(24) edge [] node {} (15)
+(34) edge [] node {} (45)
+(14) edge [] node {} (25)
+(14) edge [] node {} (35)
+(15) edge [bend right=10] node {} (26)
+(15) edge [bend left=10] node {} (36)
+(45) edge [] node {} (16)
+(35) edge [bend right=10] node {} (16)
+(35) edge [bend left=10] node {} (46)
+(26) edge [bend right=20] node {} (6com)
+(36) edge [] node {} (6inc)
+(16) edge [] node {} (6inc)
+(46) edge [bend left=20] node {} (6com)
 ;
@@
 }
-\label{fig:loop-unwind}
+\label{fig:loop-unroll}
 \caption{Example of the flow graph from Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$}
 \end{figure}
+
+The unrolling procedure is performed by the \texttt{XFlowGraphUnroller}. %, which extends the \texttt{FlowGraphDfsTraverser}
+For unrolling the flow-graph, we perform the \textit{Deep-First Search (DFS)} while counting the Y-level instructions (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
+Each next event increments the unrolling depth counter (the \textit{event reference-ID}), the integer that is stored by the \texttt{XEvent} thus making it \textit{an event reference} on non-zero values.
+Each \texttt{XEvent} has the method `\texttt{XEvent asNodeRef(int refId)}' that clones the event-receiver with new value of reference-ID.
+The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, setting up hash-code methods properly is a crucial programming task).
+
+During the work on \porthos[2], we tested different schemes of loop unrolling.
+\\<TODO: Say sth about these schemes + pictures>
+
+%- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
 

### 2018-05-14 2_wmm.tex
finish impl tentatively. add draft for eval.

@@
     \begin{itemize}[noitemsep]
-        \item the \textit{read-from} relation \rf~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
-        \item the \textit{coherence order} relation \co~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
+        \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
+        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
     \end{itemize}
 
-    \item Events from the same process are related by the \textit{scope relation} \sr~$\subseteq~\mathbb{E}~\times~\mathbb{E}$.
+    \item Events from the same process are related by the \textit{scope relation} \sr{}~$\subseteq~\mathbb{E}~\times~\mathbb{E}$.
     In contrast to the \tool{herd} tool, the \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple labels that indicate which process has produced certain event.

### 2018-05-14 4_impl.tex
finish impl tentatively. add draft for eval.

@@
 %
-\footnote{Guava project repository: \url{https://github.com/google/guava/}}%
+\footnote{The Guava project repository: \url{https://github.com/google/guava/}}%
 %
@@
 
-As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
+As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
 %
-\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
+\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}} %
 %
@@
 Thirdly, the grammar used by \porthos{} had restricted set of allowed operations.
-For example, it allowed only computations over the local variables, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code snippet `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
+For example, it allowed only computations over the local variables, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
 Also, only integers were processed by the input language parser of \porthos{}.
@@
 %
-\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}). %
+\footnote{ANTLR grammars repository path:~\url{https://github.com/antlr/grammars-v4}}). %
 %
@@
 %In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
-The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
+The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
 Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
@@
     \newline
-    Note, for the purpose of simplifying the X abstraction level, computation events may have been modelled as invocations of bodiless functions (for instance, the operation `\lstinline{x + y}' may be modelled as the invocation of the function `\lstinline{+}' with the arguments `\lstinline{x}' and `\lstinline{y}').
+    Note, for the purpose of simplifying the X abstraction level, computation events may have been modelled as invocations of bodiless functions (for instance, the operation `\lstinline{x + y}' may be modelled as the invocation of the function `\lstinline{+}' with the arguments \lstinline{x} and \lstinline{y}).
     However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT-solver.
@@
   \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
-  \item the \textit{data-flow edges}: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
+  \item the \textit{data-flow edges}: the \co{}-relation edge and the \rf{}-relation edge. %TODO! implement and then describe more!
 \end{itemize}
@@
 
-\paragraph{Graph example}
-\label{ch:impl:model:xgraph:example}
-
-consider the following code in C:
-
-picture of the CF-graph:
-
-<TODO: finish it later after fixing the code>
+%\paragraph{Graph example} %(in 5_eval.tex)
+%\label{ch:impl:model:xgraph:example}
 
@@
 
-A \wmodel{} represents a recursive AST of \textit{computations over relations} and assertion expressions defined by the memory model.
-The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
+The \textit{\wmodel{}} represents a recursive AST of \textit{computations over relations} and assertion expressions defined by the memory model.
+The atomic elements of \wmodel{} are the basic relations (\po{}, \rf{} and \co{}; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
 The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.
@@
 
-The \zformula{} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
+The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
 As all other internal representations, the \zformula{} is implemented as an immutable data structure.
@@
 
-<todo: enumerate elements -- after completing>
+<todo: enumerate elements -- after refactoring>
 
@@
 %
-\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
+\footnote{The \tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
 %
@@
 This is performed as optimisation that removes unused computations from the model (otherwise, for instance, the assignment 
-`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- 1 * (2 + 3));}' instead of single event `\texttt{write(register <- 1 * (2 + 3));}').
-If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is not used by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph:evt}).
+`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
+If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph:evt}).
 The second group of X-interpreter methods consists of the methods for defining non-linear statements (branchings and loops).
@@
 
-For simplicity of the interpretation, we keep the jump events (no-computation events) in the flow-graph, however they can be removed and the edges to and from them can be rewritten if necessary.
+For the sake of simplicity of the interpreter, jump events (no-computation events) are present in the flow-graph, however they can be removed from it with rebinding ingoing and outgoing edges.
 
@@
 The first one means that the loop has been executed a whole number of times, and the latter means that the unrolling bound triggered on not-the-last event of the loop.
-The Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, while the $S_{-}$ -- incomplete).
+Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, while the $S_{-}$ -- incomplete).
 
@@
 }
+\caption{Example of the flow graph unrolling up to bound $k = 6$}
 \label{fig:loop-unroll}
-\caption{Example of the flow graph from Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$}
 \end{figure}
@@
 For unrolling the flow-graph, we perform the \textit{Deep-First Search (DFS)} while counting the Y-level instructions (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
-Each next event increments the unrolling depth counter (the \textit{event reference-ID}), the integer that is stored by the \texttt{XEvent} thus making it \textit{an event reference} on non-zero values.
+Also, during the unrolling each next event increments the \textit{unrolling depth counter} (the \textit{event reference-ID}) -- the integer that is stored by the \texttt{XEvent} thus making it \textit{an event reference} on non-zero values.
 Each \texttt{XEvent} has the method `\texttt{XEvent asNodeRef(int refId)}' that clones the event-receiver with new value of reference-ID.
@@
 
-During the work on \porthos[2], we tested different schemes of loop unrolling.
-\\<TODO: Say sth about these schemes + pictures>
+During the work on \porthos[2], we have tested different schemes of loop unrolling.
+<TODO: Say sth about these schemes + pictures>
 
@@
 
-- set up co, rf edges => new graph
+Once the graph is unrolled, it can be augmented by data-flow edges.
+As it was discussed in Paragraph~\ref{ch:impl:model:xgraph:edges}, the data-flow edges can represent either \rf{}- or \co{}-relation.
+The \rf{}-edges join each write event to all read events that access the same location; the \co{}-edges join write events to the same location.
 
-- computing SSA maps (now: during the encoding. should be: during the post-compilation) (as one of necessary steps before encoding)
+<TODO: why in old implementation we encode the negation of "co" ? same for "rf" above
+\url{https://github.com/hernanponcedeleon/Dat3M/blob/80341bec26021d41cf79eddc943e7cbe6f49322f/dartagnan/src/dartagnan/wmm/Domain.java#L204} >
 
@@
 
-todo
+The \zformula{} encoder transforms the program and the memory models to a single SMT-like representation.
+This includes ... <todo: recall said before>
 
@@
 
-\subsection{Auxiliary components}
-\label{ch:impl:aux}
-
-There are some ...
-
-\subsubsection{Watchdog timer}
-\label{ch:impl:aux:watchdog}
-
-%<not implemented>
-
-todo
-
-\subsubsection{Logger}
-\label{ch:impl:aux:logger}
-
-%<not implemented>
-
-todo
+%\subsection{Auxiliary components}
+%\label{ch:impl:aux}
+%
+%\subsubsection{Watchdog timer}
+%\label{ch:impl:aux:watchdog}
+%
+%%<not implemented>
+%
+%todo
+%
+%\subsubsection{Logger}
+%\label{ch:impl:aux:logger}
+%
+%%<not implemented>
+%
+%todo
 

### 2018-05-14 1_intro.tex
write interpreter. update pictures.
cleanup a bit.

@@
 Most modern computer systems contain large parts that operate concurrently.
-Although the parallelisation of a system can drastically improve its performance, it opens numerous of problems connected to correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
+Although the parallelisation of a system can drastically improve its performance, it opens numerous of problems regarding correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
 
@@
 
-As an example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86}.
+As an example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86} (such little examples that illustrate specific behaviour of a concurrent execution environment are called \textit{litmus tests}).
 The process~\texttt{P0} writes the value~\texttt{1} to the shared variable~\texttt{x} and reads a value of the shared variable~\texttt{y}, and the process~\texttt{P1} writes the~\texttt{y} and reads the~\texttt{x}.
-Considering all interleavings of the processes, one could expect the final state to be one of the following states:
+Considering all interleavings of the processes, one could expect the final state to be one of the following:
 \begin{itemize}[noitemsep,topsep=0pt]
-\item \texttt{(0:EAX=0,\,1:EBX=1)}
-\item \texttt{(0:EAX=1,\,1:EBX=0)}
-\item \texttt{(0:EAX=1,\,1:EBX=1)}
+\item \texttt{(0:EAX=0,\,1:EBX=1)},
+\item \texttt{(0:EAX=1,\,1:EBX=0)},
+\item \texttt{(0:EAX=1,\,1:EBX=1)}.
 \end{itemize}
 %As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
-However, on the x86 architecture the state `\texttt{(0:EAX=0,\,1:EAX=0)}' is also reachable as the processors may cache the write to shared memory into their local write buffers, so that they do not immediately become visible by processes running on other cores.
+However, on the x86 architecture the state `\texttt{(0:EAX=0,\,1:EAX=0)}' is also reachable as the processors may cache the write to shared memory into their local write buffers, so that the write does not immediately become visible by processes running on other cores.
 This behaviour is known as \textit{store buffering~(SB)}.
@@
 
-Such little examples that illustrate specific behaviour of a concurrent execution environment are called \textit{litmus tests}.
 The formal way to define the semantics of memory operations and synchronisation primitives of a parallel execution environment is to define its \textit{memory model}.
@@
 %The issues like one discussed above have led to the need for formalisation of semantics of memory operations within different concurrent architectures defined by \textit{memory models}.
-There are two types of formal memory models.
+There are two main types of formal memory models.
 Models of the first type characterise the behaviour of the system (its \textit{operational semantics}) in terms of the abstract machine executing the code, as it was done for the SB example above.
@@
 %There are two ways to define a memory model: by specifying its \textit{operational semantics} (in terms of an abstract machine executing the code, as it was done for the SB example above) or by defining its \textit{axiomatic semantics} (in terms of assertions over states of the program formed by sequences of operations).
-Although the former type of memory models may be easier to describe and interpret, existing numerous formal verification tools and methods direct the research towards axiomatic memory models.
+Although the former type of memory models may be easier to describe and interpret, existing numerous formal verification tools and methods address the research towards axiomatic memory models.
 
-The first memory model for concurrent systems was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
+The first memory model for a concurrent system was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
 This memory model, called the \textit{sequential consistency (SC)}, allows only those executions that produce the same result as if the operations had been executed in an interleaved fashion in a single process%
@@
 This means that the order of operations executed by a process is strictly defined by the program (the code) it executes.
-The SC model does require the write to a shared variable performed in one process to become visible by all other processes \textit{instantly}, each process writes directly to the shared memory, without local buffering.
-Another important requirement of the SC memory model is that it forbids memory operations reordering within a single process (the order is strictly defined by the program).
+The SC model does require the write to a shared variable performed in one process to become visible by all other processes \textit{instantly} as each process writes directly to the shared memory, without local buffering.
+Another important requirement of the SC memory model is that it forbids reordering of memory operations within a single process (the order is strictly defined by the program).
 Originally, the operational semantics was defined for the SC model, however there exist axiomatic specifications for it~\cite{mansky2015axiomatic}.
 
-The SC model is considered to be a \textit{strong memory model} in the sense that it provides firm guarantees regarding the ordering and caused effect of memory operations.
-Different relaxations of this model lead to the class of \textit{weak memory models~(WMMs)} that specify how processes interact through shared memory, when a write becomes visible to other processes and what value a read can return.
-Thus, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they may rely on.
+The SC model is considered to be a \textit{strong memory model} in the sense that it provides firm guarantees regarding the ordering and effect of memory operations.
+Different relaxations of the SC model lead to \textit{weak memory models~(WMMs)} that specify how processes interact through the shared memory, when a write becomes visible to processes running on other cores, and what value a read operation can get.
+Thus, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they can rely on.
 
@@
 
-Research on weak memory models firstly aims to \textit{formalise} a formal approach of understanding programs with respect to weak memory models which is \textit{systematic}, \textit{sound} and \textit{complete}.
-One of the most well-known frameworks for weak memory model-aware analysis was formalised by J. Alglave in~2010~\cite{alglave2010shared},.
-In addition to developing rather theoretical basis, researchers work on extracting the WMMs for hardware architectures from existing implementations of from their specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
+Research on weak memory models firstly aims to \textit{formalise} a formal approach of understanding programs with respect to weak memory models, which is \textit{systematic}, \textit{sound} and \textit{complete}.
+One of the most well-known frameworks for weak memory model-aware analysis was formalised by J. Alglave in~2010~\cite{alglave2010shared}.
+It is the event-based non-deterministic model without global time (see Section~\ref{ch:wmm:event} for details).
+
+In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
 Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
@@
 The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
+One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{TODO: KERNEL LITMUS TESTS WORK+REPOSITORY}.
 %interprocedural analysis
-As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \textit{\porthos[2]}\,%
+As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name \textit{\porthos[2]}\,%
 %

### 2018-05-14 2_wmm.tex
cleanup a bit.

@@
 An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (which is modelled as a register as well) to the register $r_1$.
-For modelling branching statements, we distinguish the set $\mathbb{C_{g}}~\subseteq~\mathbb{C}$ of \textit{guard} computation events (also called as \textit{branching events}), that are evaluated as a boolean value.
+For modelling branching statements, we define the set $\mathbb{C}_{g}~\subseteq~\mathbb{C}$ of \textit{guard} computation events (also called as \textit{branching events}), that are evaluated to a boolean value.
 

### 2018-05-14 4_impl.tex
cleanup a bit.

@@
 The first one means that the loop has been executed a whole number of times, and the latter means that the unrolling bound triggered on not-the-last event of the loop.
-Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, while the $S_{-}$ -- incomplete).
+Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, and the $S_{-}$ denotes the incomplete sink).
 

### 2018-05-15 1_intro.tex
consider some Professor's comments.

@@
 \end{tabular}
-\caption{Store buffering: A litmus test on write-read reordering allowed under the x86-TSO and forbidden under the SC memory model}
+\caption{Store buffering: A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
 \label{simple_wmm_x86}
@@
 In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
-Over the last decade the memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
+Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
 %, or Alpha architecture~\cite{alphaWMM}~, and others.
@@
 Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled under the platforms $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
+As the result, a program compiled under the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
 
@@
 %
-\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} %
+\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/Dat3M}} %
 %
@@
 The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
-One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{TODO: KERNEL LITMUS TESTS WORK+REPOSITORY}.
+One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
+For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read).
 %interprocedural analysis
+
 As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name \textit{\porthos[2]}\,%
@@
 %
-Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.
+Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static analysis of concurrent programs.
 
@@
 Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.
-\textbf{TODO: other chapters!!!}
-
-%Clarke's goto-cc : \cite{clarke2004tool} : http://www.cprover.org/goto-cc/
-% (perhaps, goto-cc is a part of cprover) : cprover https://www.cprover.org/cbmc/doc/manual.pdf
-%cbmc http://www.cprover.org/cbmc/
+Chapter~\ref{ch:eval} provides an example of the portability analysis via \porthos[2] and presents some evaluation results comparing to other tools. %TODO: check if it's true
+Chapter~\ref{ch:summary} summarises results of the work and proposes future work directions. %TODO: check if it's true

### 2018-05-15 2_wmm.tex
consider some Professor's comments.

@@
 
-The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (\textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (\textit{analytic semantics})~\cite{alglave2016syntax}.
+The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (the \textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (the \textit{analytic semantics})~\cite{alglave2016syntax}.
 The anarchic semantics of the program is a truly parallel semantics with no global time that describes all possible computations with all possible communications.
@@
 \item \textit{input non-determinism}, which is a standard undecidable problem for all static analysis methods: to resolve the user input, system call from the environment, unresolved function calls, etc.;
-\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity;
+\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity; and
 \item \textit{memory-model non-determinism}, caused by hardware and compiler relaxations.
@@
     \begin{itemize}[noitemsep]
-        \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event;
-        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation);
+        \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event; and
+        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation).
     \end{itemize}
@@
     \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
-    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$
+    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$;
 
@@
     %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
-    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$
+    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$;
 
     \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
-    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(w))$ \\
-    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(w))$
+    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(r))$, \\
+    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(r))$;
 
@@
     \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
-    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$
+    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$; and
 
@@
 \hfill
-\caption{Possible candidate executions for the litmus test Example~\ref{simple_wmm_x86}}
+\caption{Candidate executions for the litmus test in Example~\ref{simple_wmm_x86}}
 \label{simple_wmm_x86_pic}
@@
 
-\section{\cat{} language}
+\section{The \cat{} language}
 \label{ch:wmm:cat}

### 2018-05-15 3_encod.tex
write up to Interpreter.
consider some Professor's comments.

@@
 
-While working on the \porthos[2], we applied some modifications of the encoding scheme for the control-flow.
-The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of C language.
-For that, we compile the recursive abstract syntax tree~(AST) of the parsed C-code to the plain (non-recursive) event-flow graph.
+While working on the \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
+The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
+For that, we compile the abstract syntax tree~(AST) of the parsed C-code to the plain event-flow graph.
 We show %TODO! to show, actually!!!
@@
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, \porthos{} uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
 $\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
@@
 \begin{itemize}[noitemsep,topsep=0pt]
-\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$
-\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$
-\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$
-\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$
-\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$
-\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb(E)}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$
-\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), where \\
-\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), and \\ 
-\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$
+\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$;
+\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$;
+\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$;
+\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$;
+\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$;
+\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb{E}}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$; \ and
+\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \\
+\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\ 
+\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$.
 \end{itemize}
 
-Note that the \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
+Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
 The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.

### 2018-05-15 4_impl.tex
consider some Professor's comments.

@@
 The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
-This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Yet far from all of C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
+Yet far from all of C language are supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
 %
@@
     \begin{enumerate}[label*=\arabic*.]
-      \item prohibit under-approximating abstractions;
-      \item preservation the completeness of analysis;
-      \item modular architecture: each module can be tested independently;
-      \item usage of software design patterns where necessary;
+      \item preservation the completeness of analysis,
+      \item modular architecture: each module can be tested independently,
+      \item usage of software design patterns where necessary, and
       \item usage of immutable data structures for all DTOs;
@@
     \begin{enumerate}[label*=\arabic*.]
-      \item following the principles of simplicity and readability;
+      \item following the principles of simplicity and readability, and
       \item clear and informative program output;
@@
 \end{outline}
+
+%TODO: say about IDs from old code
 

### 2018-05-15 2_wmm.tex
Reduce space between letters in notations.

@@
 A memory event can access at most one shared memory location, high-level instructions that address more than one shared variable must be transformed into a sequence of events.
-A memory event is specified by its direction with respect to the shared variable, its location~$\mathtt{loc}(e_m)$, its processor label~$\mathtt{proc}(e_m)$, and a unique event label~$\mathtt{id}(e_m)$~\cite{alglave2010shared}.
+A memory event is specified by its direction with respect to the shared variable, its location~$\texttt{loc}(e_m)$, its processor label~$\texttt{proc}(e_m)$, and a unique event label~$\texttt{id}(e_m)$~\cite{alglave2010shared}.
 %\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
@@
 The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
-We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\mathtt{load}(r_1~\leftarrow~l_2); \ e_2~=~\mathtt{store}(l_1~\leftarrow~r_1)$.
+We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\texttt{load}(r_1~\leftarrow~l_2); \ e_2~=~\texttt{store}(l_1~\leftarrow~r_1)$.
 Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.
@@
     \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subseteq~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
-    For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relation{po} e_2$.
+    For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relarrow{po} e_2$.
     %Some new relations may be acquired : dp, po-loc
@@
     \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
-    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$;
+    $r \relarrow{fr} w ~\triangleq~ (\exists w' . w' \relarrow{rf} r \land w' \relarrow{co} w)$;
 
@@
     \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
-    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
-    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$;
+    %$\texttt{com} \triangleq (\texttt{rf} \cup \texttt{co} \cup \texttt{fr})$
+    $m_1 \relarrow{com} m_2~\triangleq~((m_1\relarrow{rf}m_2) \lor (m_1\relarrow{co}m_2) \lor (m_1\relarrow{fr}m_2))$;
 
     \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
-    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(r))$, \\
-    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(r))$;
+    $w \relarrow{fre} r ~\triangleq~ (w \relarrow{fr} r \land \texttt{proc}(w) = \texttt{proc}(r))$, \\
+    $w \relarrow{fri} r ~\triangleq~ (w \relarrow{fr} r \land \texttt{proc}(w) \neq \texttt{proc}(r))$;
 
@@
     \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
-    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$; and
+    $m_1 \relarrow{po-loc} m_2 ~\triangleq~ (m_1 \relarrow{po} m_2 \land \texttt{loc}(m_1) = \texttt{loc}(m_2))$; and
 
@@
 %An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
+%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\texttt{rf}$ and $\texttt{co}$~\cite{alglave2010shared}.
 Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \tool{herd7} tool, version~7.47).
@@
     \begin{itemize}
-      \item \textit{the complement} of a relation \rel{r} is $\rel{\textasciitilde r}$,
-      \item \textit{the transitive closure} of a relation \rel{r} is $\rel{r+}$,
-      \item \textit{the reflexive closure} of a relation \rel{r} is $\rel{r?}$,
-      \item \textit{the reflexive-transitive closure} of a relation \rel{r} is $\rel{r*}$, and
-      \item \textit{the inverse} of a relation \rel{r} is $\invrel{r}$.
+      \item the \textit{complement} of a relation \texttt{r} is $\texttt{\textasciitilde r}$,
+      \item the \textit{transitive closure} of a relation \texttt{r} is $\texttt{r+}$,
+      \item the \textit{reflexive closure} of a relation \texttt{r} is $\texttt{r?}$,
+      \item the \textit{reflexive-transitive closure} of a relation \texttt{r} is $\texttt{r*}$, and
+      \item the \textit{inverse} of a relation \texttt{r} is $\invrel{r}$.
     \end{itemize}
@@
   \begin{itemize}
-    \item \textit{the union} of two relations \rel{r1} and \rel{r2} is \rel{r1\,|\,r2},
-    \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2},
-      \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1 \relminus r2}$, and
-    \item \textit{the sequence} of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}, which is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$.
+    \item the \textit{union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2},
+    \item the \textit{intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2},
+      \item the \textit{difference} of two relations \texttt{r1} and \texttt{r2} is $\texttt{r1} \relminus \texttt{r2}$, and
+    \item \textit{the sequence} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}, which is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\texttt{(x,z)} \in \texttt{r1}$ and $\texttt{(z,y)} \in \texttt{r2}$.
   \end{itemize}
@@
 
-For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$.
+For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$.
 As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}.
-This memory model asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}.
+This memory model asserts acyclicity of communication relation, \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.

### 2018-05-15 3_encod.tex
Reduce space between letters in notations.

@@
 In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
-Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $Init(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $Trans(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $Bad(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
+Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $\textit{Init}(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $\textit{Trans}(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $\textit{Bad}(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
 Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
 %\begin{equation}
-$\mathtt{SAT}( Init(\vec{x_0}) \land Trans(\vec{x}_0, \vec{x}_1) \land ... \land Trans(\vec{x}_{k-1}, \vec{x}_k) \land Bad(\vec{x}_k) )$.
+$\texttt{SAT}( \textit{Init}(\vec{x_0}) \land \textit{Trans}(\vec{x}_0, \vec{x}_1) \land ... \land \textit{Trans}(\vec{x}_{k-1}, \vec{x}_k) \land \textit{Bad}(\vec{x}_k) )$.
 %\end{equation}
@@
 The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
-For that, we compile the abstract syntax tree~(AST) of the parsed C-code to the plain event-flow graph.
+For that, we compile the \textit{abstract syntax tree~(AST)} of the parsed C-code to the plain event-flow graph.
 We show %TODO! to show, actually!!!
@@
 For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
-$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
+$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
 and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
 $\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
-(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
+(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
 In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
@@
 \begin{align}
-\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
-                      & \ [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
+\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \Rightarrow \fx(e_1) \label{enc:cf:seq} \\
+\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \Rightarrow \fx(e_1)] \land [\fx(e_3) \Rightarrow \fx(e_1)] \land \nonumber \\
+                      & \ [\fx(e_2) \Rightarrow \fv(e_1)] \land [\fx(e_3) \Rightarrow \lnot\fv(e_1)] \land \nonumber \\
                       & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
-\phi_{CF_{mer}} \defeq \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
+\phi_{CF_{mer}} \defeq \ & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \texttt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
 \end{align}
@@
 \begin{align}
-    \phi_{DF_{e = \mathtt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \rightarrow (r_{i+1} = l_{i+1}) \\
-    \phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
-    \phi_{DF_{e = \mathtt{eval}(\cdot)}} \defeq \ & \fx(e) \rightarrow \fv(e) % \\
+    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \Rightarrow (r_{i+1} = l_{i+1}) \\
+    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \Rightarrow (l_{i+1} = r_i) \\
+    \phi_{DF_{e = \texttt{eval}(\cdot)}} \defeq \ & \fx(e) \Rightarrow \fv(e) % \\
 \end{align}
@@
             \ForAll {event $e_i \in G.N$ in topological order}
-                \ForAll {predecessor $e_j \in \mathtt{pred}(e_i)$}
-                    \State {$S[e_i] \gets \mathtt{copy}(S[e_j])$}
+                \ForAll {predecessor $e_j \in \texttt{pred}(e_i)$}
+                    \State {$S[e_i] \gets \texttt{copy}(S[e_j])$}
                     \ForAll {variable $v_k \in$ set of variables accessed by $e_i$ }
-                        \State {$S[e_i][v_k] \gets \mathtt{max}(S[e_i][v_k], \ S[e_j][v_k])$}
+                        \State {$S[e_i][v_k] \gets \texttt{max}(S[e_i][v_k], \ S[e_j][v_k])$}
                         \If {need to update the index of $v_k$} \Comment {cases (1)-(2)}
@@
 \begin{align}
-\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \mathtt{rf}(e_1, e_2) \rightarrow (l_i = l_j) 
+\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \texttt{rf}(e_1, e_2) \rightarrow (l_i = l_j)
 \end{align}
 
-where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\mathtt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.
+where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\texttt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.
 
@@
 
-In an SMT-formula, the relation $x \relation{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
+In an SMT-formula, the relation $x \relarrow{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
 The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
@@
 \item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb{E}}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$; \ and
-\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \\
-\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\ 
-\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$.
+\item $r^{+}(e_1, e_2) \defeq \texttt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \vspace{0.25em} \\
+\texttt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\
+\texttt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\texttt{tc}_{i}(e_1,e_2);\texttt{tc}_{i}(e_1,e_2))$.
 \end{itemize}
+\vspace{1em}
 
-Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
+Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\texttt{r}_\texttt{1} = \texttt{r}_\texttt{2} \cup (\texttt{r}_\texttt{1}; \texttt{r}_\texttt{1})$').
 The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.

### 2018-05-15 3_encod.tex
Consider more Professor's comments.

@@
 %However, unlikely a BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
-Consider the function $cons_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
+Consider the function $\textit{cons}_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
 Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17a}:
@@
 Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
-$cons_{\mathcal{M_T}}(P) \subseteq cons_{\mathcal{M_S}}(P)$
+$\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
 \end{definition}
@@
 $\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
-and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
-$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
+and control-flow of the branching instruction $i_1 = (c\,?\,i_2\!:\!i_3)$ was encoded as
+$\phi_{CF}(c\,?\,i_2\!:\!i_3) \defeq (\textit{cf}_{i_1} \Leftrightarrow (\textit{cf}_{i_2} \lor \textit{cf}_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
 (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
@@
 Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
-Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1\,\rightarrow\,e_2), \; (e_1\,\rightarrow\,e_3)\}$.
+Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1\,\rightarrow\,e_2), \; (e_1\,\Rightarrow\,e_3)\}$.
 In Equation~\ref{enc:cf:merge}, which encodes the control-flow of a merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors has been executed, regardless the type of the transition.
@@
 In contrast, \porthos{} used the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique within a branch.
-Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow of merge points~\cite{Porthos17a}.
+Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in the case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow of merge points~\cite{Porthos17a}.
 
@@
 \begin{align}
-\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \texttt{rf}(e_1, e_2) \rightarrow (l_i = l_j)
+\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)
 \end{align}
@@
 The memory model can assert acyclicity, irreflexivity of emptiness of a relation or a set of events.
-As it has been proposed in~\cite{Porthos17a}, encoding the acyclicity assertion uses numerical variable $\Psi_e \in \mathbb{N}$ for each event $e$ in the relation to be asserted: $acyclic(r) \defeq (r(e_1,e_2) \Rightarrow (\Psi_{e_1} < \Psi_{e_2}))$.
-The irreflexivity assertion as $irreflexive(r) \defeq \bigvee\limits_{e_k \in \mathbb(E)}^{} \lnot r(e_k,e_k)$.
+As it has been proposed in~\cite{Porthos17a}, encoding the acyclicity assertion uses numerical variable $\Psi_e \in \mathbb{N}$ for each event $e$ in the relation to be asserted: $\textit{acyclic}(r) \defeq (r(e_1,e_2) \Rightarrow (\Psi_{e_1} < \Psi_{e_2}))$.
+The irreflexivity assertion as $\textit{irreflexive}(r) \defeq \bigwedge\limits_{e_k \in \mathbb(E)}^{} \lnot r(e_k,e_k)$.
 

### 2018-05-15 4_impl.tex
Consider more Professor's comments.

@@
 This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Yet far from all of C language are supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
-%
-\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+Yet far from all the C language is supported (which, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+%
+\footnote{To ensure this, one merely has to look at existing C compiler implementations, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
 %
@@
 
-The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
+The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos{} tool).
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
 
-% todo: in old porthos the strategy soft: print info about error. new: exception on any detected violation of any invariant
-
-Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
-Such a modular architecture allows to support multiple input languages by parsing them and converting parsed syntax trees to a simplified AST.
-which, along with all other data-transfer objects (DTO), must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
+Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (an AST structure) from the low-level event-based representation (an event-flow graph).
+Such a modular architecture will allow to support multiple input languages%
+%
+\footnote{Apart from the C language, \porthos[2] must be able to analyse litmus tests in different assembly languages and, as a compatibility mode, the input language of \porthos[1].} %
+%
+by parsing them and converting parsed syntax trees to a simplified AST.
+
+All internal representations used by \porthos[2] must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
 The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
@@
 
-During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.
+During the development of \porthos[2], we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.
 The following list summarises the main rules we followed during the development of \porthos[2]:
@@
       \item usage of software design patterns where necessary, and
-      \item usage of immutable data structures for all DTOs;
+      \item usage of immutable data structures for all DTOs.
     \end{enumerate}
@@
       \item following the principles of simplicity and readability, and
-      \item clear and informative program output;
+      \item clear and informative program output.
     \end{enumerate}
@@
     \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
-      \item keeping the trade-off between execution time and memory usage;
+      \item keeping the trade-off between execution time and memory usage.
     \end{enumerate}
@@
 Robustness of the analysis is the main criterion of \porthos[2] as a verification tool.
-This implies modelling the system with usage of only over-approximating abstractions 
-However it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
-As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
+Although it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
+As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected outcome (for instance, if a parser failed to parse the string and the recovery algorithm is not described).
 %The modular architecture of \porthos[2] allow each its module to be tested independently.
@@
 by Microsoft Research~\cite{de2008z3}.
-However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver.
-%TODO: *should* allow?
-
-The programming language choice for \porthos[2] was also made in favour of \textit{java}, firstly, in order to be able to reuse some parts and concepts of \porthos{} that is written in java, and secondly, because the authors find the object-oriented (OOP) concepts of java suitable for modelling languages.
-Although java does not show best results in performance benchmarks (for example, comparing to C++~\cite{hundt2011loop, oaks2014java}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver%
-invoked from \porthos[2] via java API.
+However, unlike its predecessor, the \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT-solver.
+
+%We have decided to implement \porthos[2] in \textit{Java}
+As its predecessor, \porthos[2] is written in \textit{Java},
+%The programming language choice for \porthos[2] was also made in favour of \textit{java},
+firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the object-oriented (OOP) concepts of Java suitable for modelling languages.
+Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver invoked by \porthos[2] via a Java API.
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
@@
 Both \porthos{} and \porthos[2] use the \tool{ANTLR} parser generator%
+%
 \footnote{The \tool{ANTLR} project repository: \url{https://github.com/antlr/antlr4}}%
-~\cite{parr2013definitive}, a~powerful language processing tool.
+%
+~\cite{parr2013definitive}, a powerful language processing tool.
 The \tool{ANTLR} takes as input the user-defined grammar of the target language in a BNF-like form and produces the LL(*)-parser and optionally some auxiliary classes (such as listeners and visitors for the syntax tree).
-Although this parser may be not as efficient as the hand-written language-optimised parser, it reduces the overhead of implementing the parser significantly.
-Among other advantages \tool{ANTLR}, it is worth of noting that it has rather large collection of officially supported grammars.
+Although this parser may not be as efficient as a hand-written language-optimised parser, it reduces the overhead of implementing the parser significantly.
+Among other advantages \tool{ANTLR}, it has a rather large collection of officially supported grammars.
 Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the \tool{ANTLR} an attractive instrument for solving the parsing problem.
 
-\begin{figure}[!h]%[!hb]
+\begin{figure}%[!h]%[!hb]
 \centering
-\begin{lstlisting}[language=antlr,
-                  basicstyle=\ttfamily\tiny,
-                  caption={The sketch of the input language grammar used by \oldporthos},%
-                  label={fig:in_grammar_pts}]
-<prog> : <init> <thrd>* <assert>
+%basicstyle=\ttfamily\tiny,
+\begin{lstlisting}[basicstyle=\rmfamily\scriptsize,%
+  %alsoletter={:,|,;,(,)},
+  %morekeywords={:,|,;,(,)},
+  %keywordstyle=\color{dkblue},
+  literate=%
+  {<}{{$\langle$}}1
+  {>}{{$\rangle$}}1,
+  escapeinside={[*}{*]},]
+<program>
+      :      <initialisation>   <thread>+   <assertion>
       ;
-<thrd> : thread <tid> <inst>
+<thread>   
+      :      thread   <thread-id>   <instruction>
       ;
-<inst> : <atom>
-       | <inst> ; <inst>
-       | while <pred> <inst>
-       | if <pred> { <inst> } <inst>
+<instruction>
+      :      <atom>
+      |      `{'   <instruction>   `}'
+      |      <instruction>   `;'   <instruction>
+      |      `while'   `('   <bool-expr>   `)'   <instruction>
+      |      `if'   <bool-expr>   `{'   <instruction>   `}'   <instruction>
       ;
-<atom> : <reg> <- <expr>
-       | <reg> <:- <loc>
-       | <loc> := <reg>
-       | <reg> = <loc>.load(<atomic>)
-       | <loc> = <reg>.store(<atomic>)
-       | ('mfence' | 'sync' | 'lwsync' | 'isync')
+<atom>   
+      :      <register>   `[*<*]-'   <expression>
+      |      <register>   `[*<*]:-'   <location>
+      |      <location>   `:='   <register>
+      |      <register>   `='   <location>   `.'   `load'      `('   <atomic>   `)'
+      |      <location>   `='   <register>   `.'   `store'   `('   <atomic>   `)'
+      |      (`mfence'   |   `sync'   |   `lwsync'   |   `isync')
       ;
-<pred> :
-       | true
-       | false
-       | <expr> (and | or) <expr>
-       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
+<bool-expr>
+      :      `true'
+      |      `false'
+      |      <expression>   (`and'      |      `or')   <expression>
+      |      <expression>   (`=='   |   `!='   |   `>'   |   `>='   |   `[*<*]'   |   `[*>*]=')   <expression>
       ;
-<expr> : [0-9]
-       | <reg>
-       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
+<expression>   
+      :      [0-9]
+      |      <register>
+      |      <expression>   (`*'   |   `+'   |   `-'   |   `/'   |   `%')   <expression>
       ;
 \end{lstlisting}
+\caption{The sketch of the input language grammar used by \oldporthos}
+\label{fig:in_grammar_pts}
 \end{figure}
 
-Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.
+Figure~\ref{fig:in_grammar_pts} represents the grammar sketch in BNF syntax of the input language used by \porthos[1].
+%(the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}).
+%TODO! Reformulate this.
 The input language parser used by \porthos{} suffered from several disadvantages.
-Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
+Firstly, it contained the parser code inlined directly into the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
 Such a combining of two expressive languages makes the code hardly understandable and, therefore, poorly maintainable.
-In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the \tool{ANTLR} syntax tree to the AST, that is one for all languages of an input program.
-
-Secondly, in \porthos{}, the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
-%
-\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
+In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from converting the \tool{ANTLR} syntax tree to the AST, that is one for all languages of an input program.
+
+Secondly, \porthos{} resolved the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level%
+%
+%\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
 %
 , so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.
-
 %\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
-As the reader can notice from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, the memory operations of different kinds vary syntactically.
-For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the function name.
+As the reader may have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, the memory operations of different kinds vary syntactically as well.
+For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', atomic non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the function name.
+Moreover, only the operator `\lstinline{<-}' could have an expression as the source of data, which means that expressions could be assigned only to registers.
+% (thus, the variable \lstinline{r} was recognised as a local variable as well).
 In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
-The semantics of the functions also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler:hooking}).
-
-Thirdly, the grammar used by \porthos{} had restricted set of allowed operations.
-For example, it allowed only computations over the local variables, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
-Also, only integers were processed by the input language parser of \porthos{}.
+The semantics of the functions also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler}).
+
+Thirdly, the grammar used by \porthos{} allowed only a small set of operations.
+For example, it accepted the computation expressions only over local variables.
+Thus, in the assignment expression `\lstinline{r <- (x + 1);}', the variable \lstinline{x} was parsed as a local variable even though it can be used as a shared variable in other parts of the program.
+In \porthos[2], all shared variables involved into a computation expression are tentatively copied to temporary local variables.
+%Moreover, expressions could be assigned only to local variables only via the assignment operator `\lstinline{<-}' (thus, the variable \lstinline{r} was recognised as a local variable as well).
+%, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
+
+Fourthly, \porthos[1] supported only integer constants and expressions.
+%Also, only integers were processed by the input language parser of \porthos{}.
 In \porthos[2], we extended support for primitive types supported by the \tool{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s of \tool{Z3}, floats encoded as \texttt{Real}s, enumerations encoded as \texttt{Scalar}s).
 %TODO : constant arrays? https://rise4fun.com/\tool{Z3}/tutorial/guide
-Although \tool{Z3} supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of static analysis of pointers and non-constant sized arrays moves the full support of arrays and pointers out of the scope of current thesis.
+Although the \tool{Z3} supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of pointer analysis for the arrays of non-constant size moves the full support of arrays and pointers out of the scope of current thesis.
 %Thus, adding support for more advanced structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
 
-The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form `\lstinline{1 + 2 * 3}' could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator).
-Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
-tialisation of the shared variable and only by default value \lstinline{0}. \porthos[2] supports arbitrary declaration to be performed in initialisation statement.
+Finally, the grammar used by \porthos{} had the following minor drawbacks. %A minor drawback of 
+The operators were implemented as non-associative (expressions of the form `\lstinline{1 + 2 * 3}' could not be parsed).
+Comparing to C statements, that must end with the semicolon punctuator `\lstinline{;}', statements defined in the grammar of \porthos[1] use the semicolon as a separator between statements (the final statement must not end with semicolon).
+The litmus-specific syntax for variables initialisation was used only for declaring the shared variables (all of them were initialised with default value \lstinline{0}), however, this syntax should be used as an initial assignment of both shared and local variables with arbitrary values.
+%initialise them with a value different from the default \lstinline{0} or to initialise a local variable.
+%did not allow to assign an arbitrary value to a variable (the default value \lstinline{0} was used).
+
+%Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only initialisation of the shared variable and only by default value \lstinline{0}.
+%\porthos[2] supports arbitrary declaration to be performed in initialisation statement.
 
@@
 %TODO: describe more intra- and inter-procedural!
-However, the redesigned architecture of \porthos[2] allow to easily support the inter-procedural (cross-procedure) analysis by inlining function calls and binding variable contexts.
-Thus, instead of analysing a single code file, the user can specify the whole code project.
-The functions to be executed in parallel can be specified by the user, however, the some concurrent parts of code can be detected automatically at the compilation stage while the semantics of function invocations is being resolved (for instance, the C threads operations from \texttt{thread.h} or \texttt{pthread.h} may be supported).
+However, the redesigned architecture of \porthos[2] can be easily extended to support the inter-procedural (cross-procedure) analysis that inlines function calls and binds variable contexts.
+%TODO : check in the implementation , THAT WE HAVE NOT IMPLEMENTED THIS FEATURE!!
+Thus, instead of analysing a single source code file, the user can analyse the whole code project.
+In this mode, the functions to be executed in parallel should be specified by the user.
+Also, the tool may be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
+For that, the pre-compiler should recognise functions that commence a new process (such as \texttt{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
+This functionality %implies an additional stage of the pre-compilation and 
+is left beyond current thesis.
+%resolve the semantics of function invocations and remember the functions 
+%function that commence new process or thread 
+%(for instance, the function \texttt{pthread\_create} of the library \texttt{pthread.h} that starts the new thread of the function it received as the first argument).
 %TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs.
@@
 
-The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a rather high-level AST.
-The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only, which we tend to retain clearly structured).
+The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a high-level AST.
+The Apendix~\ref{apx:trees} presents the file tree of the main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only, which we tend to retain clearly structured).
 
@@
 %\switchcolumn%\VerbatimInput{inc/lst/Ytree-tree.out}%\end{paracol}
-The abstract syntax tree \ytree{} is an abstraction level suitable for compiling the program to a low-level representation (in case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
+The abstract syntax tree, \ytree{}, is an abstraction level suitable for compiling the program to a low-level representation (in the case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
 As the \tool{ANTLR} syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}", that is converted to the \texttt{YIndexerExpression}).
@@
 
-  \item \texttt{YParameter} that represents a typed variable (the type was declared, similarly to the variable definition);
+  \item \texttt{YParameter} that represents a typed variable (the type was declared, similarly to the variable definition); and
 
@@
 
-  \item \texttt{YLinearStatement} representing a single expression;
+  \item \texttt{YLinearStatement} representing a single expression; and
 
@@
 Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
-However, all three types of \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
+However, all three types of the \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
 %TODO: rename to XThread ?
@@
   \2 \texttt{XStoreMemoryEvent}: \\
-    $(\texttt{XSharedLvalueMemoryUnit}) \leftarrow (\texttt{XLocalMemoryUnit})$,
+    $(\texttt{XSharedLvalueMemoryUnit}) \leftarrow (\texttt{XLocalMemoryUnit})$, and
   \2 \texttt{XInitialWriteEvent}: \\
@@
       %\begin{itemize}[noitemsep]
-      \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation, 
+      \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation; and 
       \2 \texttt{XBinaryComputationEvent} that encodes numeric operations (such as addition, multiplication, etc.), bit vector operations (such as bit-and, bit-xor, etc.), relative operations (such as greater-then comparison, equality comparison, etc.), and logical operations (such as conjunction and disjunction).
@@
     The control-flow event indicates a non-linear jump in the code. We distinguish two kinds of control-flow events: 
-    \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
+    \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation; and
     \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
@@
     \2 \texttt{XExitEvent}, the process sink event,
-    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}),
+    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}), and
     \2 \texttt{XAssertionEvent}, the reachability assertion made at the postlude statement of the program.
@@
 \begin{itemize}[noitemsep]
-  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
+  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf}; and
   \item the \textit{data-flow edges}: the \co{}-relation edge and the \rf{}-relation edge. %TODO! implement and then describe more!
@@
 %TODO: rename HookManager -> XHookManager
-Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in case of program in C language -- only the function name) across all signatures it stores.
+Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in the case of program in C language -- only the function name) across all signatures it stores.
 If the signature matches, the hook manager intercepts the function call and interprets the hook action instead of invoking the function directly.
@@
 As a high-level instruction (Y-level) may be compiled into a sequence of low-level instructions (for example, a computation that involves shared variables should firstly load them into the local memory and then process the computation event over local-only memory), the interpreter must maintain the \textit{stack of contexts} and remember the \textit{previous event} to be able to correctly process nested non-linear statements of C language.
-The context is a data structure that carries the \textit{state} and some additional information in case of processing non-linear statements (e.g., conditional event, first and last then- and else-branch events for binding, etc.).
+The context is a data structure that carries the \textit{state} and some additional information in the case of processing non-linear statements (e.g., conditional event, first and last then- and else-branch events for binding, etc.).
 Once the new event has been emitted, the interpreter sets up control-flow edges w.r.t.

### 2018-05-17 1_intro.tex
Consider more comments.

@@
 introduced in April~2017~\cite{Porthos17b}.
-This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver.
+This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT solver.
 This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
-As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
+As most modern SMT solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
 % TODO: search space size: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf slide
@@
 Chapter~\ref{ch:wmm} gives a general view on the weak memory model-aware analysis.
-Chapter~\ref{ch:port} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
+Chapter~\ref{ch:enc} examines the portability analysis as an bounded reachability problem that can be encoded into an SMT-formula in order to be solved automatically by an SMT~solver.
 Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.

### 2018-05-17 3_encod.tex
Consider more comments.

@@
 \chapter{Portability analysis as an SMT~problem}
-\label{ch:port}
+\label{ch:enc}
 
@@
 \section{Model checking and reachability analysis}
-\label{ch:port:mc}
+\label{ch:enc:mc}
 
@@
 \section{Portability analysis as a bounded reachability problem}
-\label{ch:port:enc}
+\label{ch:enc:enc}
 
@@
 The portability analysis problem may also be stated as a reachability problem, where the undesirable state is one reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
-%However, unlikely a BMC problem, the portability analysis does not require to call the SMT-solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
+%However, unlikely a BMC problem, the portability analysis does not require to call the SMT solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
 Consider the function $\textit{cons}_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
@@
 \subsection{Encoding for the control-flow} %Encoding the control-flow constraints}
-\label{ch:port:enc:cf}
+\label{ch:enc:enc:cf}
 
@@
 The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
-For that, we compile the \textit{abstract syntax tree~(AST)} of the parsed C-code to the plain event-flow graph.
+For that, we compile the \textit{Abstract Syntax Tree~(AST)} of the parsed C-code to the plain event-flow graph.
 We show %TODO! to show, actually!!!
@@
 In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
-The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).
+The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:enc:df}).
 
@@
             \node[c] (2) [below left of=1] {$e_2$};
+            \node[c] (2) [below of=1] {$e_2$};
             \node[c] (3) [below right of=1] {$e_3$};
@@
 \subsection{Encoding for the data-flow}
-\label{ch:port:enc:df}
+\label{ch:enc:enc:df}
 
@@
 \subsection{Encoding for the memory model} %Encoding the memory model constraints}
-\label{ch:port:enc:wmm}
+\label{ch:enc:enc:wmm}
 

### 2018-05-17 4_impl.tex
Consider more comments.

@@
 
-As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
+As its predecessor, \porthos[2] uses the open-source SMT solver \tool{Z3}%
 %
@@
 by Microsoft Research~\cite{de2008z3}.
-However, unlike its predecessor, the \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT-solver.
+However, unlike its predecessor, the \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT solver.
 
@@
 firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the object-oriented (OOP) concepts of Java suitable for modelling languages.
-Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver invoked by \porthos[2] via a Java API.
+Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT solver invoked by \porthos[2] via a Java API.
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
@@
 The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
-The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
+%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
 
@@
   %draft=true
-  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
+  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
   \caption{The general architecture of \porthos[2]}
@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT solver (Section~\ref{ch:impl:proc:smt-solver}).
 
@@
 
-The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a high-level AST.
+The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a high-level recursively%
+%
+\footnote{Hereinafter we use the term \textit{recursive} data structure (sometimes called \textit{inductive}) to refer a complex data type which can contain elements that contain other elements of the same type.
+For example, an unary expression contains references to the operator and the child expression of the same type.} %
+%
+defined AST.
 The Apendix~\ref{apx:trees} presents the file tree of the main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only, which we tend to retain clearly structured).
@@
 The abstract syntax tree, \ytree{}, is an abstraction level suitable for compiling the program to a low-level representation (in the case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
-As the \tool{ANTLR} syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of C language, it lacks multiple concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}", that is converted to the \texttt{YIndexerExpression}).
-However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite[Chapter 5]{gries2012science}), we found this level of abstraction suitable enough for our tasks.
+In terms of \porthos[1], the \ytree{} is the level of \textit{instructions}.
+%In contrast to \ytree{} representation, the \tool{ANTLR} syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of the C language.
+%However, it lacks some concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}" for an arbitrary postfix %, that is converted to the \texttt{YIndexerExpression}).
+However some details of the syntax might have been abstracted away (for instance, array operations may be emulated by functions invocations, see~\cite[Chapter 5]{gries2012science}), we find this level of abstraction suitable enough for modelling a high-level language.
 
@@
 %\textbf{(TODO: Rename CodeLocation->OriginLocation in the code!)} %TODO <--
-instance that contains information about the coordinates of the input text that generated the \ytree{} element.
+instance that contains information about the coordinates of the input text that has generated the \ytree{} element.
 
@@
 
-  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT-solver we use supports the constant-array theory, we can maintain the array type);
+  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT solver we use supports the constant-array theory, we can maintain the array type);
 
@@
 \label{ch:impl:model:xgraph}
+
+\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}
 
@@
     Note, for the purpose of simplifying the X abstraction level, computation events may have been modelled as invocations of bodiless functions (for instance, the operation `\lstinline{x + y}' may be modelled as the invocation of the function `\lstinline{+}' with the arguments \lstinline{x} and \lstinline{y}).
-    However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT-solver.
+    However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT solver.
 
@@
 \item a \co{}-edge connects two writes, an \rf{}-relation edge connects a write and a read,
-\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
+\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor, and
 \item all write-event nodes except final write-nodes must have exactly one \co-successor.
@@
 
-<todo: enumerate elements -- after refactoring>
+%todo: enumerate elements -- after refactoring>
 
@@
 
-The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
+The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT solvers.
 As all other internal representations, the \zformula{} is implemented as an immutable data structure.
 
-The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
+The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT solver.
 Generally, a \zformula{} represents the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}. % the semantics of SMT-LIB formulas defined in~\cite[Chapter 5]{smt-lib}.
@@
 Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, bitvector, real) expressions.
-The typisation of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-LIB formula.
+The typing of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-LIB formula.
 For the sake of transparency, each element of a \zformula{} contains the origin location as the reference to the X-element that produced current Z-element. %TODO: implement origin for z-formula
 
-<todo: enumerate elements -- after refactoring>
+%todo: enumerate elements -- after refactoring>
 
@@
 % steps of construction of the program representations and transformations applied to them.
-Further in this Section, the subsection titles have a number in parentheses, which correspond to a unique number of the  processing unit being discussed.
-These numbers are pictured within the gear in the top-left corner of each processing unit in Figure~\ref{fig:arch}.
+We assign unique numbers to the processing units, which can checked in Figure~\ref{fig:arch} (within gears depicted in the top-left corner of each processing unit).
+%Further in current Section, the subsection titles have a number in parentheses, which correspond to a unique number of the  processing unit being discussed.
+%These numbers are pictured within the gear in the top-left corner of each processing unit in Figure~\ref{fig:arch}.
 %he numbers in parentheses correspond the numbers of processing units pictured within gears in Figure~\ref{fig:arch}.
@@
 
-All recursive data structures are processed by units that implement the \textit{visitor} pattern~\cite{palsberg1998essence}.
+Most data structures are processed by units that implement the \textit{visitor} pattern~\cite{palsberg1998essence}.
 This is a behavioural pattern that separates the program logic from the object implementation by specifying handling methods for each element of the object.
-The visitor pattern performs the double-dispatching mechanism: the visiting class defines the method for accepting a visitor, and the visitor itself defines the 
-<TODO!!!!>
-
-To visit an instance \lstinline{x} by a visitor \lstinline{v}, the program should call the accept-method of \lstinline{x} and pass \lstinline{v} as an argument.
-In our implementation, the visitor must carry out the continuation of traversing the recursive data structure itself.
 The general structure of the visitor pattern is illustrated by the pseudo-java code in Figure~\ref{fig:visitor}.
+The visitor pattern performs the double-dispatching, a mechanism for decreasing the cohesion between the DTO (the visitee) and the processor class (the visitor): the visitee implements the \textit{accepting} method that gets the visitor as an argument and invokes its \textit{visiting} method with itself as an argument.
+Thus, the method call resolution is performed statically at compile-time without any overhead at run-time.
+%To visit an instance \lstinline{x} by a visitor \lstinline{v}, the program should call the accept-method of \lstinline{x} and pass \lstinline{v} as an argument.
+In our implementation, the visitor (not the visitee) cares about the continuation of traversing its children.
 
@@
 
-For traversing plain (non-recursive) complex data structures, we mostly follow the \texttt{iterator} pattern: the special object Iterator, that has access to elements of the data structure, iterates over them and ...
-. Since in \porthos[] we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
-However, in contrast to the visitor of a recursive data structure, the visitor of a plain data structure should be used with an \textit{iterator} that has access to the elements of the object and passes them to the visitor.
-
-Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
+For traversing graph data structures, we mostly follow the \textit{iterator} pattern: the special object \texttt{Iterator}, that has access to elements of the data structure, iterates over them in some order (for example, in topological order as it is implemented for \xgraph{}).
+%Since in \porthos{} we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
+%However, in contrast to the visitor pattern, the visitor of a plain data structure should be used with an \textit{iterator} that has access to the elements of the object and passes them to the visitor.
+The construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
 A builder is a mutable object that contains methods "filling" it by the elements that will constitute the result complex object.
@@
 
-\subsubsection{Input parsers (1,2)}
+\subsubsection{Input parsers}
 \label{ch:impl:proc:input-parser}
@@
 Both input-language and input-model parsers are implemented via \tool{ANTLR} parser generator.
-
+%In Figure~\ref{fig:arch}, the input parsers are denoted as (1) and (2) processing units.
 Details on the input program language grammar are discussed in Section~\ref{ch:impl:input}.
@@
 
-\subsubsection{W-model constructor (3)}
+\subsubsection{W-model constructor}
 \label{ch:impl:proc:w-constr}
@@
 The \wmodel{} representation is constructed by the stateless visitor \texttt{Cat2WmodelConverterVisitor} from the ANTLR syntax tree \texttt{CatParser.MainContext} of the memory-model defined in \cat{} language.
+%In Figure~\ref{fig:arch}, the \wmodel{} constructor is denoted as the (3) processing unit.
 Currently, the visitor supports the small subset of \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full OCaml interpreter.
@@
 
-<todo: file tree of w-model to appendix -- after code fixing>
-
-\subsubsection{Y-tree constructor (4)}
+%todo: file tree of w-model to appendix -- after code fixing>
+
+\subsubsection{Y-tree constructor}
 \label{ch:impl:proc:y-constr}
 
-The \ytree{} is constructed by the stateless visitor \texttt{C2YtreeConverterVisitor} from the ANTLR syntax tree \texttt{C11Parser.MainContext} of C language.
+The \ytree{} is constructed by the stateless visitor \texttt{C2YtreeConverterVisitor} from the ANTLR syntax tree \texttt{C11Parser.MainContext} of the C language.
 As the \wmodel{} constructor, the \ytree{} constructor aborts its work with the \texttt{NotSupportedException} once it meets an unsupported syntax construction.
@@
 
-As a \ytree{} constitutes a generic AST, the \ytree{} constructor performs code desugaring (for example, a \texttt{switch}-statement is converted to the equivalent \texttt{if}-statement).
-
-\subsubsection{X-graph pre-compiler (5)}
+As a \ytree{} constitutes a generic AST, the \ytree{} constructor expands the syntactic sugar expressions and statements (for example, a \texttt{switch}-statement is converted to the equivalent \texttt{if}-statement).
+
+\subsubsection{X-graph pre-compiler}
 \label{ch:impl:proc:x-pre-compiler}
@@
 \item a variable whose address was accessed by any process;
-\item a variable that was declared as a parameter of the process function;
+\item a variable that was declared as a parameter of the process function; and
 %TODO: say sth about KERNEL ANALYSIS MODE BEFORE !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-\item a variable that was exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking); %https://lwn.net/Articles/154602/
+\item a variable that was exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking).
+%https://lwn.net/Articles/154602/
 \end{itemize}
 
-\subsubsection{X-graph compiler (6)}
+\subsubsection{X-graph compiler}
 \label{ch:impl:proc:x-compiler}
@@
 \item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event),
-\item \texttt{WaitingNextLinearEvent}: the standard interpreter state for processing next linear event,
+\item \texttt{WaitingNextLinearEvent}: the standard interpreter state for processing next linear event, and
 \item \texttt{Idle}: the interpreter does not set up the edge from the previous event to the new event.
@@
 
-\subsubsection{X-graph unroller (7)}
+\subsubsection{X-graph unroller}
 \label{ch:impl:proc:x-unroll}
@@
 
-During the work on \porthos[2], we have tested different schemes of loop unrolling.
-<TODO: Say sth about these schemes + pictures>
+%TODO: Say sth about these schemes + pictures> !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+%During the work on \porthos[2], we have tested different schemes of loop unrolling.
 
@@
 
-\subsubsection{X-graph data-flow constructor (8)}
-\label{ch:impl:proc:x-df}%TODO: ref it in pre-word in this chapter
+\subsubsection{X-graph data-flow constructor}
+\label{ch:impl:proc:x-df}
 
@@
 
-<TODO: why in old implementation we encode the negation of "co" ? same for "rf" above
-\url{https://github.com/hernanponcedeleon/Dat3M/blob/80341bec26021d41cf79eddc943e7cbe6f49322f/dartagnan/src/dartagnan/wmm/Domain.java#L204} >
-
-\subsubsection{Z-formula encoder (9)}
+%TODO: why in old implementation we encode the negation of "co" ? same for "rf" above
+%\url{https://github.com/hernanponcedeleon/Dat3M/blob/80341bec26021d41cf79eddc943e7cbe6f49322f/dartagnan/src/dartagnan/wmm/Domain.java#L204} >
+
+\subsubsection{Z-formula encoder}
 \label{ch:impl:proc:z-encoder}
@@
 The \zformula{} encoder transforms the program and the memory models to a single SMT-like representation.
-This includes ... <todo: recall said before>
-
-\subsubsection{SMT-formula converter (10)}
-\label{ch:impl:proc:smt-converter}%TODO: ref in pre-word
+It follows the encoding scheme described in Chapter~\ref{ch:enc}.
+%TODO: say more!
+
+\subsubsection{SMT-formula converter}
+\label{ch:impl:proc:smt-converter}
 
@@
 
-\subsubsection{SMT solver (11)}
+\subsubsection{SMT solver}
 \label{ch:impl:proc:smt-solver}
 %TODO: check that we have ALL references in pre-word in this chapter
-As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT-solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
+As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
 

### 2018-05-17 3_encod.tex
Fix encoding part.

@@
 
-Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
-% TODO: attention to these non-breakable spaces
-Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it returns the \textit{set} of values determined by the \rf-relation; see Section~\ref{ch:wmm:model:relations} for details on relations).
-We distinguish the function ${\fv_g\;:\;\mathbb{C}_g\;\rightarrow\;\{0,1\}}$ that evaluates the guard computation event.
-In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
+All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
+An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
+A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
+The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
+However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition).
 
-\begin{figure}[b]
+\begin{figure}%[b]
   \centering
@@
         \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-            \node[c] (1) [] {$e_1$};
-            \node[c] (2) [below of=1] {$e_2$};
+            \node[c] (1) [] {$e_i$};
+            \node[c] (2) [below of=1] {$e_{i+1}$};
             \path[->]
-            (1) edge [] node {} (2)
+            (1) edge [] node {$\epsilon$} (2)
             ;
@@
         }
-        \caption{Sequential events (\texttt{seq})}
+        \caption{Sequence of events (seq)}
         \label{encode:cf:seq}
@@
     \begin{subfigure}[b]{0.35\textwidth}
-        \makebox[\textwidth]{
+        \makebox[\textwidth] {
         \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-            \node[c] (1) [] {$e_1$};
-            \node[c] (2) [below left of=1] {$e_2$};
-            \node[c] (2) [below of=1] {$e_2$};
-            \node[c] (3) [below right of=1] {$e_3$};
+            \node[c]           (k)   [] {$e_k$};
+            \node[c,draw=none] (iii) [below of=k] {$...$};
+            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
+            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
+            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
+            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
             \path[->]
-            (1) edge [] node {} (2)
-            (1) edge [dotted] node {} (3)
+            (k) edge [above] node {$g_{i,k}$} (i)
+            (k) edge [] node {} (ii)
+            (k) edge [] node {} (iii)
+            (k) edge [] node {} (iv)
+            (k) edge [above] node {$g_{i+j,k}$} (v)
             ;
@@
         }
-        \caption{Conditional branching (\texttt{br})}
+        \caption{Conditional branching (br)}
         \label{encode:cf:br}
@@
             \path[->]
-            (i) edge [] node {} (k)
-            (ii) edge [] node {} (k)
-            (iii) edge [dotted] node {} (k)
-            (iv) edge [dotted] node {} (k)
-            (v) edge [] node {} (k)
+            (i) edge [above] node {$g_{i,k}$} (k)
+            (ii) edge [above] node {} (k)
+            (iii) edge [] node {} (k)
+            (iv) edge [] node {} (k)
+            (v) edge [] node {} (k) %sloped
             ;
@@
         }
-        \caption{Branch merging (\texttt{mer})}
+        \caption{Branch merging (mer)}
         \label{encode:cf:merge}
@@
     %
-    \caption{Linear and non-linear cases of a control-flow graph}
+    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
     \label{encode:cf}
@@
 
+Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
+%Let $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ be the function that returns the value of the computation event (evaluates it) that will be computed once the event is executed (strictly speaking, it returns the \textit{set} of values determined by the \rf-relation; see Section~\ref{ch:wmm:model:relations} for details on relations).
+%We distinguish the function ${\fv_g\;:\;\mathbb{C}_g\;\rightarrow\;\{0,1\}}$ that evaluates the guard computation event.
+In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
 Consider the possible mutual arrangements of nodes in a control-flow graph presented in Figure~\ref{encode:cf}.
+
+\begin{align}
+\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \label{enc:cf:seq} \\
+\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \land \cdots \land [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \\
+                       \land & [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \land \cdots \land [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
+                       \land & \cdots \nonumber \\
+                       \land & \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \
+                               \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \\ e_n \neq e_m\end{subarray}} 
+                               \lnot [\fx(e_m) \land \fx(e_n)]  \label{enc:cf:br} \\
+\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
+\end{align}
+
 For these cases, we propose the encoding scheme that uniquely encodes each node of graph and at the same time allows to encode partially executed program.
-Equation~\ref{enc:cf:seq} encodes the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
-Equation~\ref{enc:cf:br} encodes the branching control-flow depicted in Figure~\ref{encode:cf:br} by allowing only following executions: $\{\emptyset, \; (e_1), \; (e_1\,\rightarrow\,e_2), \; (e_1\,\Rightarrow\,e_3)\}$.
-In Equation~\ref{enc:cf:merge}, which encodes the control-flow of a merge-point represented in Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors has been executed, regardless the type of the transition.
+%TODO: say about epsilon on (c) picture
+Equation~\ref{enc:cf:seq} shows the encoding for the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
+Equation~\ref{enc:cf:br} shows the encoding for the branching control-flow depicted in Figure~\ref{encode:cf:br}, that considers both transitions and guards.
+ %(similarly to the sequential control-flow)  %$\{\emptyset, \; (e_1), \; (e_1\,\Rightarrow\,e_2), \; (e_1\,\Rightarrow\,e_3)\}$.
+Also, adding negations of pairwise conjunctions over all successors of the branching node, the encoding forbids the execution of two branches simultaneously.
+Equation~\ref{enc:cf:merge} shows the encoding for the control-flow of a merge-point represented in Figure~\ref{encode:cf:merge}: the event $e_k$ is executed if either of its predecessors has been executed, regardless the type of the transition.
+Note that the sequential control-flow is a special case of branching with the only transition guarded by $\epsilon$ (that is encoded as \texttt{true}).
 %TODO: maybe add the overall formula \phi_{CF} ?
 
-\begin{align}
-\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \Rightarrow \fx(e_1) \label{enc:cf:seq} \\
-\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \Rightarrow \fx(e_1)] \land [\fx(e_3) \Rightarrow \fx(e_1)] \land \nonumber \\
-                      & \ [\fx(e_2) \Rightarrow \fv(e_1)] \land [\fx(e_3) \Rightarrow \lnot\fv(e_1)] \land \nonumber \\
-                      & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
-\phi_{CF_{mer}} \defeq \ & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \texttt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
-\end{align}
+For sake of correctness of the encoding, we require all branches to have at least one event.
+Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \textit{nop-event} as it is shown in Figure~\ref{encode:branching:nop}.
 
-For sake of correctness of the encoding, we require all branches to have at least one event.
-Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \texttt{nop}-event as it is shown in Figure~\ref{encode:branching:nop}.
-
-\begin{figure}
+\begin{figure}%[!h]
     \centering
@@
         \node[c] at (6,-1.5) (2b) {$e_2$};
-        \node[c,fill=black!15] at (7.5,-1.5) (nop) {$e_{\mbox{\tiny NOP}}$};
+        \node[c,fill=black!15] at (7.5,-1.5) (nop) {$e_{_{\mbox{\tiny NOP}}}$};
         \node[c] at (6,-3) (3b) {$e_3$};
@@
     \end{tikzpicture}
-    \caption{Transformation of the empty-branch nonlinear control-flow}
+    \caption{Transformation of the forward-jump control-flow}
     \label{encode:branching:nop}
+\end{figure}
+
+As an example of the control-flow encoding, consider the event-flow graph in Figure~\ref{fig:graph-example}, which has two branching points and one merge point with three incoming transitions.
+To illustrate the correctness of the encoding, consider the path $e_1 \rightarrow e_2 \rightarrow e_4$.
+This means, in the formula $\phi_{CF}$, the variables $\fx(e_1)$, $\fx(e_2)$ and $\fx(e_4)$ will have the value $1$, and the variable $\fx(e_3)$ will be assigned to $0$.
+It is easy to check that the $\phi_{CF}$ is satisfiable by the chosen SMT-model (considering the guards $g_{1,2}$ and $g_{2,4}$ that define this path to be evaluated to $1$).
+
+Next, consider the path $e_1 \rightarrow e_3 \rightarrow e_4$, which is not allowed by the control-flow graph.
+The corresponding model $\fx(e_1)=1$, $\fx(e_2)=0$, $\fx(e_3)=1$ and $\fx(e_4)=1$ does not satisfy the formula $\phi_{CF}$.
+The proposed encoding for the control-flow works not only for the whole graph: it can be used for encoding partial control-flow graph.
+For example, the model $\fx(e_1)=1$, $\fx(e_2)=1$, $\fx(e_3)=0$ and $\fx(e_4)=0$, which encodes the path $e_1 \rightarrow e_2$, satisfies $\phi_{CF}$.
+
+\begin{figure}
+\begin{minipage}{.4\textwidth}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+    \node[c] (i)   [] {$e_1$};
+    \node[c] (ii)  [below of=i] {$e_2$};
+    \node[c] (iii) [below of=ii] {$e_3$};
+    \node[c] (iv)  [below of=iii] {$e_4$};
+    \path[->]
+    (i) edge [right] node {$g_{1,2}$} (ii)
+    (ii) edge [left] node {$g_{2,3}$} (iii)
+    (iii) edge [right] node {$\epsilon$} (iv)
+    (i) edge [left,bend right=45] node {$g_{1,4}$} (iv)
+    (ii) edge [right,bend left=55] node {$g_{2,4}$} (iv) %sloped
+    ;
+\end{tikzpicture}
+\end{minipage}
+%
+\begin{minipage}{.6\textwidth}
+\centering
+\begin{align*}
+\phi_{CF} \defeq \ & [\fx(e_2) \Rightarrow \fx(e_1)] \\
+                 \ & \land [\fx(e_3) \Rightarrow \fx(e_2)] \\
+                 \ & \land [\fx(e_4) \Rightarrow (\fx(e_1) \lor \fx(e_2) \lor \fx(e_3))] \\
+                 \ & \land [\fx(e_1) \land \fx(e_2) \Rightarrow g_{1,2}] \\
+                 \ & \land [\fx(e_1) \land \fx(e_4) \Rightarrow g_{1,4}] \\
+                 \ & \land [\fx(e_2) \land \fx(e_3) \Rightarrow g_{2,3}] \\
+                 \ & \land [\fx(e_2) \land \fx(e_4) \Rightarrow g_{2,4}] \\
+                 \ & \land \lnot [\fx(e_2) \land \fx(e_4)] \\
+                 \ & \land \lnot [\fx(e_3) \land \fx(e_4)]
+\end{align*}
+\end{minipage}
+\caption{Example of encoding for the control-flow of the event-flow graph}
+\label{fig:graph-example}
 \end{figure}

### 2018-05-17 4_impl.tex
Consider minor changes.

@@
 
-  \item \texttt{YCompoundStatement} (block statement) representing sequence of \texttt{N} statements grouped into one syntactic unit;
+  \item \texttt{YCompoundStatement} (block statement) representing sequence of \textit{N} statements grouped into one syntactic unit;
 
@@
 The result of execution of \porthos[2] is the verdict modelled by the class \texttt{AppVerdict}.
-This is a structure that contains the result of analysis and auxiliary information such as collected errors and time of execution.
+This is a structure that contains the result of analysis and auxiliary information such as collected errors and time of execution (separately for each stage of computing).
 The app verdict may be rendered to any format convenient for the user.

### 2018-05-17 inc/lst/XInterpreter.java

@@ -33,8 +33,8 @@ public interface XInterpreter {
   void finishBlockBranchDefinition();
   void finishNonlinearBlockDefinition();
   void processJumpStatement(JumpKind kind);
-  XEntity processMethodCall(String methodName,
-                            XMemoryUnit... arguments);
+  XEntity processMethodCall(String methodName, XMemoryUnit... arguments);
+  
   
   // --
   

### 2018-05-18 2_wmm.tex
Consider minor changes.
Consider Professor's comments on abstract.

@@
     \item Events from the same process are related by the \textit{scope relation} \sr{}~$\subseteq~\mathbb{E}~\times~\mathbb{E}$.
-    In contrast to the \tool{herd} tool, the \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple labels that indicate which process has produced certain event.
+    In contrast to the \tool{herd} tool, \porthos[2] does not use hierarchy of scopes (depicted as the scope tree); instead, it uses simple labels that indicate which process has produced certain event.
 \end{itemize}
@@
 For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$.
-As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}.
+As an example of memory model definition in \cat{} language, the x86-TSO model~\cite{herd10tutorial} can be found in Appendix~\ref{apx:x86cat}.
 This memory model asserts acyclicity of communication relation, \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.

### 2018-05-18 3_encod.tex
Consider minor changes.
Consider Professor's comments on abstract.

@@
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
-%As well as the BMC problem, the approach used by the \porthos{}
+%As well as the BMC problem, the approach used by \porthos{}
 
@@
 
-The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with single source and multiple sink nodes. %TODO: footnote why multiple sinks
+The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with a single source and multiple sink nodes. %TODO: footnote why multiple sinks
 In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
@@
 
-While working on the \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
+While working on \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
 The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
@@
 \phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \label{enc:cf:seq} \\
-\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \land \cdots \land [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \\
-                       \land & [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \land \cdots \land [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
-                       \land & \cdots \nonumber \\
-                       \land & \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \
+\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \\
+                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
+                       \land & \ \cdots \nonumber \\
+                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \
                                \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \\ e_n \neq e_m\end{subarray}} 
-                               \lnot [\fx(e_m) \land \fx(e_n)]  \label{enc:cf:br} \\
+                              \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \label{enc:cf:br} \\
 \phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
@@
 Equation~\ref{enc:cf:br} shows the encoding for the branching control-flow depicted in Figure~\ref{encode:cf:br}, that considers both transitions and guards.
- %(similarly to the sequential control-flow)  %$\{\emptyset, \; (e_1), \; (e_1\,\Rightarrow\,e_2), \; (e_1\,\Rightarrow\,e_3)\}$.
 Also, adding negations of pairwise conjunctions over all successors of the branching node, the encoding forbids the execution of two branches simultaneously.
 Equation~\ref{enc:cf:merge} shows the encoding for the control-flow of a merge-point represented in Figure~\ref{encode:cf:merge}: the event $e_k$ is executed if either of its predecessors has been executed, regardless the type of the transition.
-Note that the sequential control-flow is a special case of branching with the only transition guarded by $\epsilon$ (that is encoded as \texttt{true}).
+Note that the sequential control-flow is a special case of branching with the only transition guard $\epsilon$ (that is encoded as \texttt{true}).
 %TODO: maybe add the overall formula \phi_{CF} ?
@@
 \begin{align}
-    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \Rightarrow (r_{i+1} = l_{i+1}) \\
-    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \Rightarrow (l_{i+1} = r_i) \\
-    \phi_{DF_{e = \texttt{eval}(\cdot)}} \defeq \ & \fx(e) \Rightarrow \fv(e) % \\
+    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \\
+    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \\
+    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] % \\
 \end{align}
@@
 
-The time of described algorithm is linear of the size of event-flow graph since it performs only single traverse of the graph.
+The time of described algorithm is linear of the size of event-flow graph since it performs only one graph traverse.
 %Notwithstanding the overhead of storing (or, equivalently, computing with %TODO: are you sure that 'with'?
@@
 \begin{align}
-\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)
+\phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)]
 \end{align}

### 2018-05-18 4_impl.tex
Consider Professor's comments on abstract.

@@
 
-The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos{} tool).
+The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos{} tool).
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
 by Microsoft Research~\cite{de2008z3}.
-However, unlike its predecessor, the \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT solver.
+However, unlike its predecessor, \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT solver.
 
@@
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
-It is worth noting that, for the reasons of simplicity, the \porthos[2] is not a concurrent program, however, we believe that, due to its modular architecture, it can be easily parallelised on the level of program modules.
+It is worth noting that, for the reasons of simplicity, \porthos[2] is not a concurrent program, however, we believe that, due to its modular architecture, it can be easily parallelised on the level of program modules.
 
@@
 This is performed as optimisation that removes unused computations from the model (otherwise, for instance, the assignment 
-`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
+`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
 If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph:evt}).

### 2018-05-18 3_encod.tex
add reference to Unfoldings

@@
 \section{Portability analysis as a bounded reachability problem}
-\label{ch:enc:enc}
+\label{ch:enc:bmc}
 
@@
 \subsection{Encoding for the control-flow} %Encoding the control-flow constraints}
-\label{ch:enc:enc:cf}
+\label{ch:enc:bmc:cf}
 
@@
 While working on \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
-The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
+These changes were motivated by the need to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
 For that, we compile the \textit{Abstract Syntax Tree~(AST)} of the parsed C-code to the plain event-flow graph.
@@
 This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
-For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+
+For instance, \porthos[1] used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
 $\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
@@
 (here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
-In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
-The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:enc:df}).
+In contrast, the new encoding scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
+The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:bmc:df}).
+In general, the new encoding scheme follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} for encoding Petri-nets.
 
@@
 The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
-However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition).
+However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition), see Section~\ref{ch:impl:model:xgraph:edges} for details.
 
@@
 \subsection{Encoding for the data-flow}
-\label{ch:enc:enc:df}
+\label{ch:enc:bmc:df}
 
@@
 \subsection{Encoding for the memory model} %Encoding the memory model constraints}
-\label{ch:enc:enc:wmm}
+\label{ch:enc:bmc:wmm}
 

### 2018-05-18 4_impl.tex
cleanup a bit
add reference to Unfoldings

@@
 
-\section{Architecture}
-\label{ch:impl:arch}
-
-The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
-%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
-
 \begin{figure}
@@
 
+\section{The architecture}
+\label{ch:impl:arch}
+
+The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
+%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
+
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
@@
 
-\begin{figure}%[!h]%[!hb]
+\begin{figure}[t]%[!h]%[!hb]
 \centering
@@
 
-\subsection{The internal representations}
+\subsection{Internal representations}
 \label{ch:impl:model}
@@
 
-\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}
+<TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO!!!!!!!!!!!!!!!!!!!!!!>
+%TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
+\textit{\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}}
 
@@
 The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
-%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring
+Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph:edges}).
 
@@
 \label{ch:impl:model:xgraph:edges}
-Edges are stored as an immutable hash-map from source to target event.
-The type of edge is stored in the enumeration \texttt{XEdgeKind}:%TODO: IMPLEMENT THIS
-\begin{itemize}[noitemsep]
-  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf}; and
-  \item the \textit{data-flow edges}: the \co{}-relation edge and the \rf{}-relation edge. %TODO! implement and then describe more!
-\end{itemize}
+As the graph is represented by an adjacency matrix, its edges are stored in (immutable) hash-maps.
+We distinguish the following kinds of edges:
+
+\begin{outline}%[noitemsep]
+  \1 the \textit{control-flow edges}:
+    \2 the \textit{primary edges}, that denote both $\epsilon$-labelled transitions (in case of linear sequence of events) and conditional transition that evaluates the conditional event (the source of the transition) to the \textit{true}, and
+    \2 the \textit{alternative edges}, that denote conditional transitions for which the conditional event (the source) was evaluated to the \textit{false}; and
+  \1 the \textit{data-flow edges}:
+    \2 the \co{}-relation edges, and 
+    \2 the \rf{}-relation edges. %TODO! implement 
+\end{outline}
 
@@
 \label{ch:impl:model:xgraph:invariants}
-Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:port:enc:cf}):
+Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:enc:bmc:cf}):
 
@@
 
-\subsection{The processing units} %Transformation}
+\subsection{Processing units} %Transformation}
 \label{ch:impl:proc}

### 2018-05-18 1_intro.tex
add reference to Szevczik thesis

@@
 Most modern high-level programming languages rely on relaxed memory model as well.
-Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time}, it was introduced in J2SE 5.0 in 2004~\cite{manson2005java}; the C++11 standard~\cite{iso2012iec} has introduced the set of hardware-independent synchronisation fences and atomic operations.
-The C++17 memory model~\cite{batty2011mathematizing} is based on the relation \textit{strongly happens-before}.
-Weak memory are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{alglave2018frightening, kernel1}.
+Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time} and was introduced in J2SE 5.0 in 2004~\cite{manson2005java}.
+The transformations valid for the Java memory model were discussed in~\cite{sevcik2009program}. 
+A weak memory model for C and C++ is based on the relation \textit{strongly happens-before} and was introduced in the C++17 standard~\cite{batty2011mathematizing}.
+Before that, the C++11 standard~\cite{iso2012iec} proposed the set of hardware-independent synchronisation fences and atomic operations.
+Weak memory models are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{alglave2018frightening, kernel1}.
 

### 2018-05-18 4_impl.tex
redraw unrolling pic.

@@
 The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
-This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
 This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
@@
 \node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
-\node[rectangle,draw] (6inc) [below left=1cm and 0.6cm of 16] {$S_{-}$};
-\node[rectangle,draw] (6com) [below of=6inc] {$S_{+}$};
+\node[rectangle,draw] (6inc) [below left=1cm and 3cm of 16] {$S_{-}$};
+\node[rectangle,draw] (6com) [right=3cm of 6inc] {$S_{+}$};
 \node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
@@
 (35) edge [bend left=10] node {} (46)
-(26) edge [bend right=20] node {} (6com)
-(36) edge [] node {} (6inc)
-(16) edge [] node {} (6inc)
+(26) edge [bend right=10] node {} (6com)
+(36) edge [bend right=5] node {} (6inc)
+(16) edge [bend left=20] node {} (6inc)
 (46) edge [bend left=20] node {} (6com)

### 2018-05-19 slides/slides.tex
add template for slides.

\documentclass{beamer}

\usepackage[english]{babel}
%\usepackage[doublenumbering,twosupervisors]{../lib/aalto-thesis}
%\RequirePackage{aaltologo}
%\RequirePackage{beamercolorthemeAalto}
%\RequirePackage{beamerthemeAalto}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme[school={SCI}]{Aalto}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

%\titlegraphic{\includegraphics[height=1cm,width=2cm]{logo1}}
%\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}

\title{Automated Analysis of Weak Memory Models}

% A subtitle is optional and this may be deleted
%\subtitle{Optional Subtitle}
\author{Artem Yushkovskiy\inst{1,2} \\ 
{\scriptsize MSc Candidate}
\\ \vspace{1em}
{\footnotesize\raggedleft Supervisors: Assoc. Prof. Keijo Heljanko\inst{1} \newline
\hphantom{Supervis} Docent Igor I. Komarov\inst{2} }
}%\small
%{\footnotesize%\raggedleft%does not work
%Supervisor 1: Assoc. Prof. Keijo Heljanko\inst{1} \newline
%Supervisor 2: Docent Igor I. Komarov\inst{2}
%}

\institute[Universities of Somewhere and Elsewhere] % (optional, but mostly needed)
{
  \inst{1}%
  Department of Computer Science, \\
  School of Science, \\
  Aalto University (Espoo, Finland)%
  \and
  \inst{2}%
  Faculty of Information Security \\
  and Computer Technologies, \\
  ITMO University (Saint Petersburg, Russia)%
}

\date{\scriptsize Espoo, Saint Petersburg, 2018}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

%\subject{Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

%\pgfdeclareimage[height=0.5cm]{university-logo}{}
%\logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{First Main Section}

\subsection{First Subsection}

\begin{frame}{First Slide Title}{Optional Subtitle}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{Second Subsection}

% You can reveal the parts of a slide one at a time
% with the \pause command:
\begin{frame}{Second Slide Title}
  \begin{itemize}
  \item {
    First item.
    \pause % The slide will pause after showing the first item
  }
  \item {   
    Second item.
  }
  % You can also specify when the content should appear
  % by using <n->:
  \item<3-> {
    Third item.
  }
  \item<4-> {
    Fourth item.
  }
  % or you can use the \uncover command to reveal general
  % content (not just \items):
  \item<5-> {
    Fifth item. \uncover<6->{Extra text in the fifth item.}
  }
  \end{itemize}
\end{frame}

\section{Second Main Section}

\subsection{Another Subsection}

\begin{frame}{Blocks}
\begin{block}{Block Title}
You can also highlight sections of your presentation in a block, with it's own title
\end{block}
\begin{theorem}
There are separate environments for theorems, examples, definitions and proofs.
\end{theorem}
\begin{example}
Here is an example of an example block.
\end{example}
\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.
\section*{Summary}

\begin{frame}{Summary}
  \begin{itemize}
  \item
    The \alert{first main message} of your talk in one or two lines.
  \item
    The \alert{second main message} of your talk in one or two lines.
  \item
    Perhaps a \alert{third message}, but not more than that.
  \end{itemize}

  \begin{itemize}
  \item
    Outlook
    \begin{itemize}
    \item
      Something you haven't solved.
    \item
      Something else you haven't solved.
    \end{itemize}
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{Bibliography}

  \begin{thebibliography}{10}

  \beamertemplatebookbibitems
  % Start with overview books.

  \bibitem{Author1990}
    A.~Author.
    \newblock {\em Handbook of Everything}.
    \newblock Some Press, 1990.

  \beamertemplatearticlebibitems
  % Followed by interesting articles. Keep the list short.

  \bibitem{Someone2000}
    S.~Someone.
    \newblock On this and that.
    \newblock {\em Journal of This and That}, 2(1):50--100,
    2000.
  \end{thebibliography}
\end{frame}

\end{document}

### 2018-05-19 slides/slides.tex
Add slides for wmm intro.

@@
 
-\usepackage[english]{babel}
-%\usepackage[doublenumbering,twosupervisors]{../lib/aalto-thesis}
+\usepackage[utf8]{inputenc}
+\usepackage[T2A]{fontenc}
+\usepackage[russian,english]{babel}
+%\usepackage{tikz}
+%\usetikzlibrary{tikzmark}
+%\usetikzlibrary{calc}
+
+\usepackage[font=scriptsize]{caption}
+\setlength{\abovecaptionskip}{5pt plus 0pt minus 2pt} % Chosen fairly arbitrarily
+
+\usepackage[backend=biber]{biblatex}
+\addbibresource{../bib/ms-thesis.bib}
+%\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}
+
+\usepackage{array}
+\usepackage{xcolor}
+
+\usepackage{hyphenat} % for using \nohyphens
+%\usepackage{multirow} %Used for tables with merged cells
+%\usepackage{collcell} %pdflatex.exe hangs without this one
+
+\usepackage{booktabs}% http://ctan.org/pkg/booktabs
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+
 %\RequirePackage{aaltologo}
@@
 
-\title{Automated Analysis of Weak Memory Models}
+\title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}
 
@@
 %\subtitle{Optional Subtitle}
-\author{Artem Yushkovskiy\inst{1,2} \\ 
+\author{\textbf{Artem Yushkovskiy}\inst{1,2} \\ 
 {\scriptsize MSc Candidate}
 \\ \vspace{1em}
-{\footnotesize\raggedleft Supervisors: Assoc. Prof. Keijo Heljanko\inst{1} \newline
-\hphantom{Supervis} Docent Igor I. Komarov\inst{2} }
+{\footnotesize\raggedleft Supervisors: \textbf{Assoc. Prof. Keijo Heljanko}\inst{1} \newline
+\hphantom{Supervis} \textbf{Docent Igor I. Komarov}\inst{2} }
 }%\small
@@
 
-\institute[Universities of Somewhere and Elsewhere] % (optional, but mostly needed)
+\institute % (optional, but mostly needed)
 {
@@
   School of Science, \\
-  Aalto University (Espoo, Finland)%
+  \textbf{Aalto University} (Espoo, Finland)%
   \and
@@
   and Computer Technologies, \\
-  ITMO University (Saint Petersburg, Russia)%
+  \textbf{ITMO University} (Saint Petersburg, Russia)%
 }
@@
 
+\usepackage{colortbl}
+\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
+\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
+\definecolor{dkred} {rgb} {0.6, 0.1, 0}
+\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
+\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}
+
 % Let's get started
@@
 \end{frame}
+
+\section{Introduction}
+
+\subsection{Weak memory model-aware analysis}
+
+\begin{frame}{Verification of concurrent software}
+\framesubtitle{Example: Write-write reordering (by compiler)}
+%1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
+%2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
+\begin{table}
+\centering
+\ttfamily
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
+\hline
+\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P & Q \\ \hline
+$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
+$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\end{tabular}
+\end{table}
+
+%\tikzmark{p1} 
+%\begin{tikzpicture}[remember picture, overlay]
+%  \path[draw=red,thick,<->]<4-> (pic cs: p0) to [] (pic cs: p1);
+%\end{tikzpicture}
+
+\addtolength{\tabcolsep}{-3pt}
+
+\def\firstrowcolor{}
+\def\secondrowcolor{}
+\hspace{-20pt}
+\begin{minipage}{.25\textwidth}
+\small
+\onslide<2-> 
+\begin{tabular}{ l r}
+\rowcolor{ltgrey} $\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
+\rowcolor{ltgrey} $\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$ \\
+%\hline
+\pause
+\onslide<3-> 
+$\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
+$\b{p_0}, \r{q_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
+$\r{q_0}, \b{p_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
+$\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
+%\hline
+\pause
+\end{tabular}
+\end{minipage}
+%
+\hspace{5pt}
+\begin{minipage}{.6\textwidth}
+\small
+\onslide<4->
+%\vspace{-1em}
+%"delay the effect of a store past any load from a different location."
+\vspace{-11pt}
+\begin{tabular}{ | l r | l r | l r }
+\rowcolor{ltgrey} $\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
+\rowcolor{ltgrey} $\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
+$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
+$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
+$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
+$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
+\end{tabular}
+\end{minipage}
+
+\end{frame}
+
+\begin{frame}{Verification of concurrent software}
+\framesubtitle{Example: Store buffering (by hardware)}
+\begin{center}
+\scalebox{0.7}{
+\ttfamily
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
+\hline
+\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P & Q \\ \hline
+$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
+$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\end{tabular}
+}
+\end{center}
+
+\vspace{-5pt}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{img/candidate.png}
+\caption{The four candidate executions}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}{Verification of concurrent software}
+\framesubtitle{A weak memory model: Operational semantics} %{Example: Store buffering (hardware)}
+\begin{center}
+\scalebox{0.7}{
+\ttfamily
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
+\hline
+\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P & Q \\ \hline
+$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
+$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\end{tabular}
+}
+\end{center}
+
+\begin{figure}
+\includegraphics[scale=0.37]{img/x86-arch-full.png}
+\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
+%\label{fig:x86-arch}
+\end{figure}
+\end{frame}
+
+%\framesubtitle{A weak memory model: Axiomatic semantics}
+
+\begin{frame}{Verification of concurrent software}
+\framesubtitle{A weak memory model: Axiomatic semantics}
+
+\begin{itemize}
+  \item 
+\end{itemize}
+
+\end{frame}
+
+%\subsection{Актуальность}
+
+\subsection{Problem statement (Цель исследования)}
+
+\subsection{Task specification (Задачи исследования)}
+
+1) 
+2)
+3)
 
@@
   \frametitle<presentation>{Bibliography}
-
-  \begin{thebibliography}{10}
-
-  \beamertemplatebookbibitems
-  % Start with overview books.
-
-  \bibitem{Author1990}
-    A.~Author.
-    \newblock {\em Handbook of Everything}.
-    \newblock Some Press, 1990.
-
-  \beamertemplatearticlebibitems
-  % Followed by interesting articles. Keep the list short.
-
-  \bibitem{Someone2000}
-    S.~Someone.
-    \newblock On this and that.
-    \newblock {\em Journal of This and That}, 2(1):50--100,
-    2000.
-  \end{thebibliography}
+  \printbibliography
 \end{frame}

### 2018-05-19 slides/img/candidate-gen.sh

@@ -1,0 +1,1 @@
+convert +append SB-dot.png SB-dot-*.png candidate.png

### 2018-05-20 slides/slides.tex
Add slides up to eval.

@@
 \usepackage[russian,english]{babel}
-%\usepackage{tikz}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+\usetikzlibrary{calc}
+\usetikzlibrary{arrows}
+\usetikzlibrary{decorations.pathmorphing,decorations.markings}
+\usetikzlibrary{shapes}
+\usetikzlibrary{patterns}
+\usetikzlibrary{automata}
 %\usetikzlibrary{tikzmark}
@@
 
+\tikzset{
+c/.style={
+  circle,solid,
+  inner sep=0pt,
+  %text width=1mm,
+  minimum width=7mm,
+  align=center,
+  draw=black,
+  fill=white
+  }
+}
+
+\newcommand*{\fx}{\ensuremath{\mathbf{x}}}%x}}}
+\newcommand*{\fv}{\ensuremath{\mathbf{v}}} %v}}}
+\newcommand{\defeq}{\ensuremath{=}}%\coloneqq}}%\Leftrightarrow}}%\triangleq}}
+
+\usepackage{multicol}
+
+\usepackage{fancyvrb}%fancy verbatim
+%\makeatletter
+%\newcommand{\verbatimfont}[1]{\renewcommand{\verbatim@font}{\ttfamily\scriptsize#1}}
+%\makeatother
+% redefine \VerbatimInput
+\RecustomVerbatimCommand{\VerbatimInput}{VerbatimInput}{fontsize=\scriptsize}
+
+\usepackage{pmboxdraw}
+\usepackage{newunicodechar}
+\newunicodechar{└}{\textSFii}
+\newunicodechar{├}{\textSFviii}
+\newunicodechar{─}{\textSFx}
+
+\usepackage[export]{adjustbox} % loads also graphicx
+
+\usepackage{listings}
+
+\usepackage{subcaption}
 \usepackage[font=scriptsize]{caption}
@@
 
+\addtolength{\tabcolsep}{-3pt}
+\setlength{\leftmargini}{10pt}
 %\RequirePackage{aaltologo}
@@
 %\usetheme{Marburg}
-\usetheme{Montpellier}
+%\usetheme{Montpellier}
 %\usetheme{PaloAlto}
@@
 %\usetheme{Rochester}
-%\usetheme{Singapore}
+\usetheme{Singapore}
 %\usetheme{Szeged}
 %\usetheme{Warsaw}
+
+%\beamertemplatenavigationsymbolsempty  %no navigation bar
+%\setbeamertemplate{footline}[frame number]
+%\setbeamertemplate{headline}{}
+\addtobeamertemplate{navigation symbols}{}{%
+    \usebeamerfont{footline}%
+    \usebeamercolor[fg]{footline}%
+    \hspace{1em}%
+    \insertframenumber/\inserttotalframenumber
+}
+
+\setbeamertemplate{frametitle}[default][left]
 
@@
   \begin{frame}<beamer>{Outline}
-    \tableofcontents[currentsection,currentsubsection]
+    \tableofcontents[currentsection]%,currentsubsection]
   \end{frame}
@@
 \usepackage{colortbl}
+
 \definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
@@
   \tableofcontents
-  % You might wish to add the option [pausesections]
 \end{frame}
@@
 
-\subsection{Weak memory model-aware analysis}
+\subsection{The problem}
 
-\begin{frame}{Verification of concurrent software}
-\framesubtitle{Example: Write-write reordering (by compiler)}
+\begin{frame}{Problem statement (Цель работы)}
+To rework the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by:
+\begin{itemize}
+\item extending the C-like input language,
+\item revising its architecture and
+\item re-implementing the tool in order to enhance performance, extensibility, reliability and maintainability
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Task specification (Задачи работы)}
+\begin{itemize}
+\item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
+\item Review existing tools for memory model-aware analysis;
+\item Investigate existing architecture of \texttt{Porthos}, its strengths and weaknesses;
+\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
+\end{itemize}
+\end{frame}
+
+\subsection{The background: weak memory model-aware analysis}
+
+\begin{frame}{Verification of Concurrent Software} {Example: Write-write reordering (compiler  relaxations)}
 %1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
 %2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
+
 \begin{table}
@@
 \ttfamily
-\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
+\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
 \hline
-\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
 P & Q \\ \hline
@@
 
-\addtolength{\tabcolsep}{-3pt}
-
-\def\firstrowcolor{}
-\def\secondrowcolor{}
 \hspace{-20pt}
@@
 \onslide<2-> 
-\begin{tabular}{ l r}
-\rowcolor{ltgrey} $\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
-\rowcolor{ltgrey} $\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$ \\
-%\hline
-\pause
+\begin{table}
+\caption*{SC}
+\begin{tabular}{ >{\columncolor{ltgrey}}l >{\columncolor{ltgrey}}r}
+$\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
+$\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$
 \onslide<3-> 
+\\
 $\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
@@
 $\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
-%\hline
+\end{tabular}
+\end{table}
+\end{minipage}
 \pause
-\end{tabular}
-\end{minipage}
 %
@@
 %"delay the effect of a store past any load from a different location."
-\vspace{-11pt}
+%\vspace{-11pt}
+%\vspace{-2pt}
+\begin{table}
+\caption*{TSO}
 \begin{tabular}{ | l r | l r | l r }
-\rowcolor{ltgrey} $\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
-\rowcolor{ltgrey} $\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
+$\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
+$\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
 $\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
@@
 \end{tabular}
+\end{table}
 \end{minipage}
@@
 
-\begin{frame}{Verification of concurrent software}
-\framesubtitle{Example: Store buffering (by hardware)}
+\begin{frame}{Verification of Concurrent Software} {Example: Store buffering (by hardware)}
 \begin{center}
@@
 \ttfamily
-\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
+\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
 \hline
-\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
-P & Q \\ \hline
-$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
-$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
-\hline
-\end{tabular}
-}
-\end{center}
-
-\vspace{-5pt}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{img/candidate.png}
-\caption{The four candidate executions}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}{Verification of concurrent software}
-\framesubtitle{A weak memory model: Operational semantics} %{Example: Store buffering (hardware)}
-\begin{center}
-\scalebox{0.7}{
-\ttfamily
-\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
-\hline
-\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
 P & Q \\ \hline
@@
 
-%\framesubtitle{A weak memory model: Axiomatic semantics}
-
-\begin{frame}{Verification of concurrent software}
-\framesubtitle{A weak memory model: Axiomatic semantics}
+\begin{frame}{Weak Memory Model} {Axiomatic semantics}
 
 \begin{itemize}
-  \item 
+  \addtolength{\tabcolsep}{0pt}
+  \item \textbf{Event} $\in \mathbb{E}$, a low-level primitive operation:
+    \begin{itemize}
+      \item \textit{memory event} $\in \mathbb{M} = \mathbb{R} \cup \mathbb{W}$: access to a local/shared memory, \\
+      \item \textit{computational event} $\in \mathbb{C}$: computation over local memory, and \\
+      \item \textit{barrier event} $\in \mathbb{B}$: synchronisation fences;
+     \end{itemize}
+  \item \textbf{Relation} $\subseteq \mathbb{E} \times \mathbb{E}$: 
+    \begin{itemize}
+    \item \textit{basic relations}:%, computed from the program model:
+      \begin{itemize}
+        \item \textit{program-order} relation $\texttt{po} \subseteq \mathbb{E} \times \mathbb{E}$: (control-flow),
+        \item \textit{read-from} relation $\texttt{rf} \subseteq \mathbb{W} \times \mathbb{R}$: (data-flow), and
+        \item \textit{coherence-order} relation $\texttt{co} \subseteq \mathbb{W} \times \mathbb{W}$: (data-flow);
+      \end{itemize}
+    \item \textit{derived relations}:%, computed from the basic relations using the following operators:
+      \begin{itemize}
+        \item \textit{union} \texttt{r1\,|\,r2},
+        \item \textit{sequence} \texttt{r1} ; \texttt{r2},
+        %\item \textit{inverse} \texttt{r}\^{\texttt{-1}},
+        \item \textit{transitive closure} \texttt{r+},
+        \item $\cdots$;
+      \end{itemize}
+    \end{itemize}
+  %\item \textbf{Candidate execution}, a sequence of guesses of relations;
+  \item \textbf{Assertion} over relations or sets of events: 
+    \begin{itemize}
+    \item \textit{acyclicity}, \textit{irreflexivity} or \textit{emptiness} 
+    \end{itemize}
 \end{itemize}
@@
 
-%\subsection{Актуальность}
+\begin{frame}{Weak Memory Model} {Example: Store buffering (hardware relaxations)}
+%\begin{center}
 
-\subsection{Problem statement (Цель исследования)}
+\vspace{10pt}
+\begin{minipage}{.3\linewidth}
+\hfill
+\end{minipage}
+%
+\noindent\begin{minipage}{.4\linewidth}
+\scalebox{0.7}{
+\ttfamily
+\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
+\hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P & Q \\ \hline
+$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
+$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\end{tabular}
+}
+\end{minipage}
+%
+\raggedleft{
+\noindent\begin{minipage}{.25\linewidth}
+\scriptsize
+SC model: \\
+... \\
+\texttt{fr = (rf\^\,-1; co)} \\
+\color{red}\texttt{acyclic(fr $\cup$ po)}
+\end{minipage}
+}
 
-\subsection{Task specification (Задачи исследования)}
+%\end{center}
 
-1) 
-2)
-3)
+\vspace{-5pt}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{img/candidate-red.png}
+\caption{The four candidate executions allowed under x86-TSO}
+\end{figure}
 
-% Section and subsections will appear in the presentation overview
-% and table of contents.
-\section{First Main Section}
+\end{frame}
 
-\subsection{First Subsection}
+\subsection{Portability analysis}
 
-\begin{frame}{First Slide Title}{Optional Subtitle}
+\begin{frame}{Portability analysis} {The \texttt{Porthos} tool}
+
 \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
+\item Let the function $\textit{cons}_{\mathcal{M}}(P)$ calculate the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
+\end{itemize}
+
+\begin{definition}[Portability~\cite{Porthos17a}]
+Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
+The program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
+$\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
+\end{definition}
+
+\begin{itemize}
+\item Portability as an SMT-based bounded reachability problem: \\
+${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$
+\item $\texttt{SAT}(\phi) \implies $ the portability bug
+\end{itemize}
+
+%end{tabular}
+\end{frame}
+
+\subsection{SMT-encoding}
+
+\begin{frame}{Encoding for the control-flow}% {The \texttt{Porthos v1} encoding scheme}
+\begin{itemize}
+\item In \texttt{Porthos}, the high-level instructions were represented in the SMT-formula by separate variables.
+\item For example, the sequential instruction $i_1 = i_2; i_3$ was encoded as
+$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
 \end{itemize}
@@
 
-\subsection{Second Subsection}
+\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}
 
-% You can reveal the parts of a slide one at a time
-% with the \pause command:
-\begin{frame}{Second Slide Title}
 \begin{itemize}
-  \item {
-    First item.
-    \pause % The slide will pause after showing the first item
+\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
+\item The high-level AST is compiled into the \textit{event-flow graph}, where edges (transitions) are labelled by the \textit{guard} ($\epsilon$ is an empty guard).
+\end{itemize}
+
+\begin{figure}[b]
+  \centering
+  \begin{minipage}{\textwidth}
+      \begin{subfigure}[b]{0.25\textwidth}
+        \makebox[\textwidth]{
+        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+            \node[c] (1) [] {$e_i$};
+            \node[c] (2) [below of=1] {$e_{i+1}$};
+            \path[->]
+            (1) edge [] node {$\epsilon$} (2)
+            ;
+        \end{tikzpicture}
         }
-  \item {   
-    Second item.
+        \caption{The sequence}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.35\textwidth}
+        \makebox[\textwidth] {
+        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+            \node[c]           (k)   [] {$e_k$};
+            \node[c,draw=none] (iii) [below of=k] {$...$};
+            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
+            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
+            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
+            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
+            \path[->]
+            (k) edge [above] node {$g_{i,k}$} (i)
+            (k) edge [] node {} (ii)
+            (k) edge [] node {} (iii)
+            (k) edge [] node {} (iv)
+            (k) edge [above] node {$g_{i+j,k}$} (v)
+            ;
+        \end{tikzpicture}
         }
-  % You can also specify when the content should appear
-  % by using <n->:
-  \item<3-> {
-    Third item.
+        \caption{Conditional branching}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.3\textwidth}
+        \makebox[\textwidth]{
+        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+            \node[c] (i) [] {$e_i$};
+            \node[c,draw=none] (ii) [right=0.2cm of i] {};
+            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
+            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
+            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
+            \node[c] (k) [below of=iii] {$e_k$};
+            \path[->]
+            (i) edge [above] node {$g_{i,k}$} (k)
+            (ii) edge [above] node {} (k)
+            (iii) edge [] node {} (k)
+            (iv) edge [] node {} (k)
+            (v) edge [] node {} (k) %sloped
+            ;
+        \end{tikzpicture}
         }
-  \item<4-> {
-    Fourth item.
-  }
-  % or you can use the \uncover command to reveal general
-  % content (not just \items):
-  \item<5-> {
-    Fifth item. \uncover<6->{Extra text in the fifth item.}
-  }
-  \end{itemize}
+        \caption{Branch merging}
+    \end{subfigure}
+    %
+    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
+    \label{encode:cf}
+\end{minipage}
+\end{figure}
+
 \end{frame}
 
-\section{Second Main Section}
+\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}
 
-\subsection{Another Subsection}
+The control-flow is encoded as following:
+\begin{align}
+\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
+\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
+                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
+                       \land & \ \cdots \nonumber \nonumber\\
+                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
+                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
+\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
+\end{align}
 
-\begin{frame}{Blocks}
-\begin{block}{Block Title}
-You can also highlight sections of your presentation in a block, with it's own title
-\end{block}
-\begin{theorem}
-There are separate environments for theorems, examples, definitions and proofs.
-\end{theorem}
-\begin{example}
-Here is an example of an example block.
-\end{example}
 \end{frame}
 
-% Placing a * after \section means it will not show in the
-% outline or table of contents.
+\begin{frame}{Encoding for the data-flow}
+
+\begin{itemize}
+\item SSA-indices are computed as following:
+\begin{itemize}
+  \item any access to a shared variable (both read and write) increments its SSA-index;
+  \item only writes to a local variable increment its SSA-index (reads preserve indices);
+  \item no access to a constant variable or computed (evaluated) expression changes their SSA-index.
+\end{itemize}
+\end{itemize}
+
+The data-flow of an event is encoded as following:
+\begin{align}
+    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \nonumber \\
+    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \nonumber\\
+    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] \nonumber \\
+    \phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)] \nonumber
+\end{align}
+\end{frame}
+
+\section{Implementation}
+
+%\begin{frame}{General principles}
+%\begin{enumerate}[nolistsep]
+%  \item \textit{Robustness:}
+%    \begin{enumerate}[label*=\arabic*.]
+%      \item preservation the completeness of analysis,
+%      \item modular architecture: each module can be tested independently,
+%      \item usage of software design patterns where necessary, and
+%      \item usage of immutable data structures for all DTOs.
+%    \end{enumerate}
+%  \item \textit{Transparency:}
+%    \begin{enumerate}[label*=\arabic*.]
+%      \item following the principles of simplicity and readability, and
+%      \item clear and informative program output.
+%    \end{enumerate}
+%  \item \textit{Efficiency:}
+%    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
+%      \item keeping the trade-off between execution time and memory usage.
+%    \end{enumerate}
+%  \item \textit{Extensibility:}
+%    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
+%      \item clear modular architecture.
+%    \end{enumerate}
+%\end{enumerate}
+%\end{frame}
+
+\subsection{Architecture}
+
+\begin{frame}{Architecture}%{The general architecture}
+\vspace{-30pt}
+\begin{figure}
+  \centering
+  \includegraphics[height=.99\textheight,keepaspectratio]{../img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
+  %\caption{The general architecture of \texttt{PorthosC}}
+  %\label{fig:arch}
+\end{figure}
+\end{frame}
+
+\begin{frame}{The input language}
+The input language parser used by \texttt{Porthos} suffered from several disadvantages:
+\begin{itemize}
+\item it contained the parser code inlined directly into the grammar (hardly maintainable);
+\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\lstinline{=}', `\lstinline{:=}', `\lstinline{<-}' and `\lstinline{<:-}', each for different kinds of arguments);
+\item restricted syntax for expressions.
+\item In contrast, \texttt{PorthosC} uses the full C language grammar of proposed in the C11 standard~\cite{jtc2011sc22} and the visitor that converts the ANTLR grammar to the AST (\texttt{Y-tree}).
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{The \texttt{X-graph}}
+  \centering
+  \begin{figure}
+    \includegraphics[width=1.1\textwidth,keepaspectratio]{../img/my/class-diagrams/XEntity-interfaces.png}
+    \caption{The inheritance tree of main \texttt{X-graph} interfaces}
+  \end{figure}
+
+\end{frame}
+
+\begin{frame}{The \texttt{X-graph} compiler}
+\begin{figure}
+  \centering
+  \scalebox{0.7}{\includegraphics[height=.99\paperheight]{../img/my/draw.io/X-compiler.png}}
+  \caption{Main components of the X-compilation processing unit}
+\end{figure}
+\end{frame}
+
+\begin{frame}{\texttt{X-graph} unrolling}
+\begin{figure}[t]
+\centering
+\resizebox{.8\linewidth}{!}{
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+\node[c] (1) [] {$1$};
+\node[c] (2) [below left of=1] {$2$};
+\node[c] (3) [below right of=1] {$3$};
+\node[c] (4) [below of=3] {$4$};
+\path[->]
+(1) edge [] node {} (2)
+(2) edge [bend left=50,dotted] node {} (1)
+(1) edge [] node {} (3)
+(3) edge [] node {} (4)
+(4) edge [bend right=80,dotted] node {} (1)
+;
+\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
+;
+\node[c] (22) [right=3cm of impl] {$2_2$};
+\node[c] (11) [above right=1cm and 1cm of 22]{$1_1$};
+\node[c] (32) [below right=1cm and 1cm of 11] {$3_2$};
+\node[c] (43) [below of=32] {$4_3$};
+\node[c] (13) [below of=22] {$1_3$};
+\node[c] (24) [below left=1cm and 1cm of 13] {$2_4$};
+\node[c] (34) [below right=1cm and 1cm of 13] {$3_4$};
+\node[c] (14) [below of=43] {$1_4$};
+\node[c] (15) [below of=24] {$1_5$};
+\node[c] (45) [below of=34] {$4_5$};
+\node[c] (25) [below of=14] {$2_5$};
+\node[c] (35) [below of=14] {$3_5$};
+\node[c] (26) [below left=1cm and 1cm of 15] {$2_6$};
+\node[c] (36) [below right=1cm and 1cm of 15] {$3_6$};
+\node[c] (16) [below right=1cm and 0.2cm of 45] {$1_6$};
+\node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
+\node[rectangle,draw] (6inc) [below left=1cm and 3cm of 16] {$S_{-}$};
+\node[rectangle,draw] (6com) [right=3cm of 6inc] {$S_{+}$};
+\node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
+\node[] (k2) [right=1.7cm of 32] {$(k = 2)$};
+\node[] (k3) [right=1.7cm of 43] {$(k = 3)$};
+\node[] (k4) [right=1.7cm of 14] {$(k = 4)$};
+\node[] (k5) [right=1.7cm of 35] {$(k = 5)$};
+\node[] (k6) [right=1cm of 46] {$(k = 6)$};
+\path[->]
+(11) edge [] node {} (22)
+(11) edge [] node {} (32)
+(32) edge [] node {} (43)
+(22) edge [] node {} (13)
+(13) edge [bend right=10] node {} (24)
+(13) edge [bend left=10] node {} (34)
+(43) edge [] node {} (14)
+(24) edge [] node {} (15)
+(34) edge [] node {} (45)
+(14) edge [] node {} (25)
+(14) edge [] node {} (35)
+(15) edge [bend right=10] node {} (26)
+(15) edge [bend left=10] node {} (36)
+(45) edge [] node {} (16)
+(35) edge [bend right=10] node {} (16)
+(35) edge [bend left=10] node {} (46)
+(26) edge [bend right=10] node {} (6com)
+(36) edge [bend right=5] node {} (6inc)
+(16) edge [bend left=20] node {} (6inc)
+(46) edge [bend left=20] node {} (6com)
+;
+\end{tikzpicture}
+}
+\caption{Example of the flow graph unrolling up to bound $k = 6$}
+\end{figure}
+\end{frame}
+
 \section*{Summary}

### 2018-05-20 2_wmm.tex
Add some comments to the paper.

@@
 There are two kinds of relations between events: \textit{basic relations} that capture the semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification.
-Constraints over relations that are specified by weak memory models are defined as requirements of acyclicity, irreflexivity or emptiness of specific relations~\cite{alglave2016syntax}.
+Constraints over relations that are specified by weak memory models are defined as requirements of \textit{acyclicity}, \textit{irreflexivity} or \textit{emptiness} of specific relations~\cite{alglave2016syntax}.
 
@@
         \item the \textit{read-from} relation \rf{}~$\subseteq~\mathbb{W}~\times~\mathbb{R}$ that maps each write event to the read event that reads the value written by write event; and
-        \item the \textit{coherence order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation).
+        \item the \textit{coherence-order} relation \co{}~$\subseteq~\mathbb{W}~\times~\mathbb{W}$ that defines the total order on writes to the same location across all processes (also called the \textit{write serialisation}, \texttt{ws}-relation).
     \end{itemize}

### 2018-05-20 3_encod.tex
Add some comments to the paper.

@@
 \begin{definition}[Portability]
-Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
+Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
+The program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
 $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
@@
 The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with a single source and multiple sink nodes. %TODO: footnote why multiple sinks
-In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
-Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
-However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.
+The edges of 
+%In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
+All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
+An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
+A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
+The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
+However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition), see Section~\ref{ch:impl:model:xgraph:edges} for details.
+%Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
+%However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.
 
@@
 In general, the new encoding scheme follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} for encoding Petri-nets.
-
-All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
-An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
-A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
-The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
-However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition), see Section~\ref{ch:impl:model:xgraph:edges} for details.
 
@@
 
+\textbf{TODOOOOOOOOOOOOOOOOOOOOOOOOOOOO!!!!!!!!!!!!!!!!!!!!!!! NOPS IN THIS PICTURE!!!!}
+
 \subsection{Encoding for the data-flow}
@@
 Following~\cite{Porthos17b}, the indexed references of variables are computed in accordance with the following rules:
-%\begin{enumerate}[noitemsep,topsep=0pt]
 (1)~any access to a shared variable (both read and write) increments its SSA-index;
 (2)~only writes to a local variable increment its SSA-index (reads preserve indices);
-(3)~no access to a constant variable or computed (evaluated) expression changes their %SSA-index.
-%\end{enumerate}
+(3)~no access to a constant variable or computed (evaluated) expression changes their SSA-index.
 These rules determine the following encoding of load, store and computation events within single thread:

### 2018-05-20 slides/slides.tex
Cleanup slides.

@@
 
+\newcommand{\tool}[1]{\texttt{#1}}
 \newcommand*{\fx}{\ensuremath{\mathbf{x}}}%x}}}
@@
 % the beginning of each subsection:
-\AtBeginSubsection[]
+\AtBeginSection[]
 {
@@
 
-\begin{frame}{Verification of Concurrent Software} {Example: Write-write reordering (compiler  relaxations)}
+\begin{frame}{Verification of concurrent software} {Example: Write-write reordering (compiler  relaxations)}
 %1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
@@
 
-\begin{frame}{Verification of Concurrent Software} {Example: Store buffering (by hardware)}
+\begin{frame}{Verification of concurrent software} {Example: Store buffering (hardware relaxations)}
 \begin{center}
@@
 
-\begin{frame}{Weak Memory Model} {Axiomatic semantics}
+\begin{frame}{The weak memory model} {Axiomatic semantics: The definition}
 
@@
 
-\begin{frame}{Weak Memory Model} {Example: Store buffering (hardware relaxations)}
-%\begin{center}
+\begin{frame}{The weak memory model} {Testing the candidate executions}
+\vspace{10pt}
+\begin{minipage}{.3\linewidth}
+\hfill
+\end{minipage}
+%
+\noindent\begin{minipage}{.4\linewidth}
+\scalebox{0.7}{
+\ttfamily
+\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
+\hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
+P & Q \\ \hline
+$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
+$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\end{tabular}
+}
+\end{minipage}
+%
+\noindent\begin{minipage}{.25\linewidth}
+\end{minipage}
+\vspace{-5pt}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{img/candidate.png}
+\caption{The four candidate executions allowed under x86-TSO}
+\end{figure}
+\end{frame}
 
+\begin{frame}{The weak memory model} {Testing the candidate executions}
 \vspace{10pt}
@@
 }
-
-%\end{center}
-
 \vspace{-5pt}
@@
 
-\subsection{Portability analysis}
+\begin{frame}{Tools for memory model-aware analysis}
+\begin{itemize}
+\item \texttt{diy} tool suite:
+  \begin{itemize}
+  \item \tool{diy}, \tool{diycross} and \tool{diyone}, litmus tests generators,
+  \item \tool{litmus}, a litmus test concrete executor, and 
+  \item \tool{herd}, a weak memory model simulator;
+  \end{itemize}
+%\item the stateless model checkers (\tool{CHESS}~\cite{musuvathi2008fair}, \tool{Nidhugg}~\cite{abdulla2017stateless}); %TODO: description of statelessses
+\item the stateless model checkers (\tool{CHESS}, \tool{Nidhugg}); %TODO: description of statelessses
+\item the tool for automated synthesis of the synchronisation primitives \tool{musketeer};
+\item the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker;
+\item the tool \texttt{Porthos} for analysing the portability of the C programs;
+\item and others.
+\end{itemize}
+\end{frame}
+
+\subsection{Portability Analysis}
 
@@
 
-\begin{frame}{Encoding for the control-flow}% {The \texttt{Porthos v1} encoding scheme}
-\begin{itemize}
-\item In \texttt{Porthos}, the high-level instructions were represented in the SMT-formula by separate variables.
-\item For example, the sequential instruction $i_1 = i_2; i_3$ was encoded as
-$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
-\end{itemize}
-\end{frame}
+%\begin{frame}{Encoding for the control-flow}% {The \texttt{Porthos v1} encoding scheme}
+%\begin{itemize}
+%\item In \texttt{Porthos}, the high-level instructions were represented in the SMT-formula by separate variables.
+%\item For example, the sequential instruction $i_1 = i_2; i_3$ was encoded as
+%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
+%\end{itemize}
+%\end{frame}
 
@@
 \begin{itemize}
-\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
-\item The high-level AST is compiled into the \textit{event-flow graph}, where edges (transitions) are labelled by the \textit{guard} ($\epsilon$ is an empty guard).
+%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
+\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
+$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
+\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
+\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
 \end{itemize}
@@
 
-The control-flow is encoded as following:
+\begin{itemize}
+\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
+\item The control-flow of the program is encoded as following:
+\end{itemize}
+
+\vspace{-20pt}
 \begin{align}
@@
 
-\subsection{Architecture}
-
-\begin{frame}{Architecture}%{The general architecture}
-\vspace{-30pt}
-\begin{figure}
-  \centering
-  \includegraphics[height=.99\textheight,keepaspectratio]{../img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
-  %\caption{The general architecture of \texttt{PorthosC}}
-  %\label{fig:arch}
-\end{figure}
-\end{frame}
+\subsection{The input language and general architecture}
 
@@
 
-\begin{frame}{The \texttt{X-graph}}
+\begin{frame}{Architecture}%{The general architecture}
+\vspace{-30pt}
+\begin{figure}
+  \centering
+  \includegraphics[height=.99\textheight,keepaspectratio]{../img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
+  %\caption{The general architecture of \texttt{PorthosC}}
+  %\label{fig:arch}
+\end{figure}
+\end{frame}
+
+\subsection{The \texttt{X-graph}: structure and construction}
+
+\begin{frame}{The \texttt{X-graph} internal representation}
   \centering
@@
 
+\section{Evaluation}
+
+\begin{frame}{Evaluation}
+\alert{[to be done]}
+\end{frame}
+
 \section*{Summary}
@@
 \begin{itemize}
-  \item
-    The \alert{first main message} of your talk in one or two lines.
-  \item
-    The \alert{second main message} of your talk in one or two lines.
-  \item
-    Perhaps a \alert{third message}, but not more than that.
-  \end{itemize}
-
-  \begin{itemize}
-  \item
-    Outlook
-    \begin{itemize}
-    \item
-      Something you haven't solved.
-    \item
-      Something else you haven't solved.
-    \end{itemize}
+\item The general framework for memory model-aware analysis was implemented in \texttt{PorthosC};
+%\item Review existing tools for memory model-aware analysis;
+\item The input language has been extended; %TODO: how exactly?
+\item The old architecture of \texttt{Porthos} has been analysed and considered while designing the new architecture for \texttt{PorthosC};
+%\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
+\item \alert{to be done: more}
+%\item the new architecture for the tool was designed,
 \end{itemize}

### 2018-05-20 4_impl.tex
Add some comments to the paper.
Add examples of cf graphs to eval section. Cleanup unrolling description.

@@
 
-Thirdly, the grammar used by \porthos{} allowed only a small set of operations.
+Thirdly, the grammar used by \porthos{} allowed only a restricted set of operations.
 For example, it accepted the computation expressions only over local variables.
@@
 %TODO: think about sorting processes at the pre-compilation stage
-For transparency reasons, the first event of a process must be the \texttt{XEntryEvent} and the last events must be events of type \texttt{XExitEvent}.
+The first event of each process must be the \texttt{XEntryEvent} and the last events must be events of type \texttt{XExitEvent}.
 
@@
 \begin{itemize}
-\item \texttt{WaitingAdditionalCommand}: the interpreter is in the state of defining the complex (non-linear) statement and is not able to process any new event (will throw an exception if any) until the state is not changed,
-\item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event),
+\item \texttt{WaitingAdditionalCommand}: the interpreter is in the state of defining the complex (non-linear) statement and is not able to process any new event (will throw an exception if any) until the state is not changed;
+\item \texttt{WaitingFirstConditionEvent}: the first next emitted event will be accepted as the first event of the condition evaluation (later, the loop edges will be set to this event);
+\item \texttt{WaitingLastConditionEvent}: the next event must be of type \texttt{XComputationEvent}; it will be saved as the conditional branching event;
+\item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event);
 \item \texttt{WaitingNextLinearEvent}: the standard interpreter state for processing next linear event, and
@@
 
-Once the (non-linear) context has been popped out of the context stack (by the methods of the second group that finish non-linear branch definition) and added to the queue of \texttt{almostReadyContexts}.
+Each new event emitted by the interpreter is processed considering the state of the context stack:
+the interpreter iterates over the context stack and sets up the edges by the graph builder depending on the state of each stack.
+The state \texttt{WaitingFirstConditionEvent} is necessary for correctly interpreting the branching conditions, which shared variables are involved in.
+If the non-linear statement is a loop statement, the loop back-edges will be set to this event.
+
+Once interpretation of the (non-linear) branch is completed, its non-linear context is being popped out of the context stack (by interpreter method \texttt{startBlockBranchDefinition}) and added to the queue of \texttt{almostReadyContexts}.
 The almost-ready-context becomes a ready-context (i.e., it moves to the queue \texttt{readyContexts}) once the non-linear statement definition has finished (the method \texttt{finishNonlinearBlockDefinition}).
@@
 To convert it to an acyclic form, we perform the \textit{unrolling} (sometimes called \textit{unwinding}) transfomation: each cycle is unrolled up to the user-defined bound $k$.
-Comparing to \porthos[1], we have change the meaning of the bound: instead of executing all cycles at most $k$ times, the \porthos[2] interprets the bound as the maximum number of Y-instructions to be executed after the unrolling.
+Comparing to \porthos[1], we have changed the meaning of the bound: instead of executing all cycles at most $k$ times, the \porthos[2] interprets the bound as the maximum number of events in the trace of the unrolled graph%
+%
+\footnote{The original specification of \porthos[2] stated that the unrolling bound $k$ must be interpreted as the maximum number of instruction in the original code (technically, the number of expressions in the ANTLR syntax tree).
+Nonetheless, current implementation of the \xgraph{} unroller counts the X-level events as it is much simpler to implement (the questions about how to count complex expressions that involve multiple shared variables and method invocations are left open for the future versions of \porthos[2]).
+For now, the information about the element of the ANTLR tree that corresponds to the X-event or Y-tree element is being lost during the transformations.
+Instead, we use the \texttt{LocationService} discussed in Section~\ref{ch:impl:model:ytree} that in perspective may be extended for providing this information along with the text citation of the original code.}
+%
+.
 The meaning of a bound was changed in order to increase predictability of the size of the result SMT-formula.
-Note, recursive function calls create unconditional jumps back in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
-
-Considering the new meaning of the unrolling bound, we create sink events of two types: the \textit{complete} and \textit{incomplete} sink event.
-The first one means that the loop has been executed a whole number of times, and the latter means that the unrolling bound triggered on not-the-last event of the loop.
-Figure~\ref{fig:loop-unroll} illustrates unrolling for the cyclic control-flow graph on the left-hand side (the square node $S_{+}$ denotes the complete sink node, and the $S_{-}$ denotes the incomplete sink).
+Not that recursive function calls create unconditional jumps back in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
+
+Considering the new meaning of the unrolling bound, the graph unrolling may be \textit{complete} (the loop has been executed a whole number of times) or \textit{incomplete} (the unrolling bound has triggered on not-the-last event of the loop), which can be modelled by two types of the sink events.
+The user may need this information for understanding how the \porthos[2] interprets the cyclic program.
+However, current implementation loses this information by having only one sink event.
+Figure~\ref{fig:loop-unroll} illustrates the unrolling for the left-hand side cyclic control-flow graph (the square node $S_{+}$ denotes the complete sink node, and the $S_{-}$ denotes the incomplete sink).
 
@@
 Each \texttt{XEvent} has the method `\texttt{XEvent asNodeRef(int refId)}' that clones the event-receiver with new value of reference-ID.
-The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, setting up hash-code methods properly is a crucial programming task).
-
-%TODO: Say sth about these schemes + pictures> !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-%During the work on \porthos[2], we have tested different schemes of loop unrolling.
+The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
+
+%TODO!!!!!!!!!!! other schemes!\textsl{}
+%During the work on \porthos[2], we have tested several different schemes of loop unrolling.
 

### 2018-05-20 5_eval.tex
Add macro for unrolled xgraph. Add some plan + precompilation section.
Consider Professor's comments.
write interpreter. update pictures.
finish impl tentatively. add draft for eval.
cleanup a bit
Add some comments to the paper.
Add examples of cf graphs to eval section. Cleanup unrolling description.

@@
 
-\section{Comparison with PORTHOS}
+%TODO: resource trebovaniya
 
-\subsection{Unique Features}
+\section{The \texttt{PorthosC} in work}
+
+\subsection{The X-graph compilation}
+
+As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, consider the tiny C test in Figure~\ref{ex:compilation}.
+This function does not perform any useful computation, however it contains several syntactic elements (such as prefix increment or loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
+The control-flow subgraph \xgraph[CF] of the non-unrolled event-flow graph is presented on the right-hand side of the Figure (all graphs generated by \porthos[2] are produced with the help of the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
+
+Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.
+Write events are denoted with the left-directed arrow `\lstinline{<-}', and the functions \lstinline{load} and \lstinline{store} denote the type of the shared memory event.
+The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
+The graph contains a single source event and a single sink event (represented by the grey triangles).
+
+\begin{figure}[!h]
+\begin{minipage}{.4\textwidth}
+\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
+void test(int &x, int &y) {
+    int r;
+    while ((5 + 4/2 * x) 3 
+      == 1) {
+        if (y > 6)
+            continue;
+        else if (++y > 7) {
+            r = r + 10;
+            break;
+        }
+        r = 11;
+    }
+    y = x + 1;
+}
+\end{lstlisting}
+\end{minipage}
+%
+\begin{minipage}{.59\textwidth}
+  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test.png}
+\end{minipage}
+%
+\caption{An example of a C function compiled to the event-flow graph}
+\label{ex:compilation}
+\end{figure}
+
+The arguments \lstinline{x} and \lstinline{y} of the function are passed by reference, therefore they are treated as global variables.
+The local variable declaration `\lstinline{int r;}' produces no events (it is processed by the X-graph pre-compiler that invokes the memory manager to create the new local variable \lstinline{r}).
+
+The first event `\lstinline{load(reg_tmp0 <- x}' loads the value of the global variable \lstinline{x} into the temp register \lstinline{reg_tmp0} in order to satisfy the requirement that all computation must be performed over local variables;
+note that each element of the computational tree `\lstinline{eval(eval(eval(5 + eval(eval(4 / 2) * reg_tmp0)) % 3) == 1)}' is represented by a local memory unit (either \texttt{XComputationEvent} or \texttt{XConstant} or \texttt{XRegister}).
+The node of this computational event in the control-flow graph has two outgoing edges, the primary edge to the `\lstinline{load(reg_tmp1 <- y}', the first event of the then-branch of the while-loop, and the alternative edge to the \lstinline{nop}-event representing the only event of the else-branch.
+
+As is was discussed in Section~\ref{ch:impl:model:xgraph}, all computation events have no impact to the global state of the concurrent system.
+Therefore, the interpreter does not \textit{emit} computational events, but it \textit{creates} them.
+This means, when the \texttt{Y2XConverterVisitor} processes the expression tree `\lstinline{(5 + 4/2 * x) 3 == 1}', it calls the method \texttt{createComputationEvent} of the interpreter that only creates the computation event and does not change its state.
+However, once the converter meets the global variable \lstinline{x}, it calls the interpreter method \texttt{emitMemoryEvent} to copy its value to a temp register; since it meets the global variables involved to the computation before it ends to process the whole computation expression, the values of all these global variables will be copied to temp registers \textit{before} the computation expression is used by any other event.
+Note that if the computation expression has not been used by any other event (for example, as the constant \lstinline{1} in the following C code: `\lstinline{foo(); 1; bar();}'), it is lost from the model (by the term \textit{use} here we mean that the computation event is evaluated as a guard or assigned to another memory unit).
+
+\subsection{The X-graph unrolling}
+
+Once the \xgraph[CF] is constructed, it should be unrolled to an acyclic flow-graph (as it was discussed in Section~\ref{ch:impl:proc:x-unroll}).
+The Figure~\ref{ex:unrolling} shows the control-flow subgraph \xgraphU{CF} of the event-flow graph from the previous example (see Figure~\ref{ex:compilation}) unrolled up to bound $k=16$.
+The labels of events in the picture are augmented by the event reference-id number (the unrolling depth), separated from the event value by the comma.
+
+Note that the graph does not become a tree after removing the sink node.
+Some branches of the graph are merged when the executions have the same event with the same unrolling depth number.
+For example, primary transitions of both events `\lstinline{[r <- 11, 10]}' and `\lstinline{[jump(), 10]}' (produced by executions of the first iteration of the \texttt{while} loop) lead to the same event `\lstinline{[load(reg_tmp0 <- x), 11]}' (the first event of the second loop iteration).
+
+%(said before in 4_impl:unrolling)
+%Also note that comparing to the unrolling bound description given in Section~\ref{ch:impl:proc:x-unroll}, 
+%As the distinction between complete and incomplete sink nodes is not implemented yet,
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test_unrolled.png}
+\caption{An example of an unrolled event-flow graph represented in Figure~\ref{ex:compilation}}
+\label{ex:unrolling}
+\end{figure}
+
+\subsection{Reachability analysis}
+
 \subsection{Performance}
+
+Measure time by \porthos[1], \porthos[2]:
+
+- enc-time absolute values
+
+- rate enc-time/solve-time
+
+Compare sizes of formulas
+
+memory usage -- for both encoding and solving
 
@@
 \subsection{Unique Features}
+
+- (inherited from the first version) : two memory models, smt, ... (?)
+
+- (new) -- ?
+
+- kernel?? TODO: SUPPORT some kernel funcs.
+
 \subsection{Performance}
+
+- perhaps, better performance on a large test

### 2018-05-20 3_encod.tex
Cleanup some old todos.

@@
 The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with a single source and multiple sink nodes. %TODO: footnote why multiple sinks
-The edges of 
+%The edges of 
 %In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
-All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
-An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
+Each control-flow edge contains the label that denotes the \textit{guard}, a predicate determining the transition.
+The empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
 A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
@@
         \node[c] at (6,-1.5) (2b) {$e_2$};
-        \node[c,fill=black!15] at (7.5,-1.5) (nop) {$e_{_{\mbox{\tiny NOP}}}$};
+        \node[c,fill=black!15] at (7.5,-1.5) (nop) {$e_{nop}$};
         \node[c] at (6,-3) (3b) {$e_3$};
@@
 
+Note that the generalised encoding scheme does not require the branching transitions to be mutually-exclusive (for instance, consider two branching transitions from the event $e_2$, both labelled by non-epsilon guards $g_{2,3}$ and $g_{2,4}$).
 Next, consider the path $e_1 \rightarrow e_3 \rightarrow e_4$, which is not allowed by the control-flow graph.
 The corresponding model $\fx(e_1)=1$, $\fx(e_2)=0$, $\fx(e_3)=1$ and $\fx(e_4)=1$ does not satisfy the formula $\phi_{CF}$.
-The proposed encoding for the control-flow works not only for the whole graph: it can be used for encoding partial control-flow graph.
-For example, the model $\fx(e_1)=1$, $\fx(e_2)=1$, $\fx(e_3)=0$ and $\fx(e_4)=0$, which encodes the path $e_1 \rightarrow e_2$, satisfies $\phi_{CF}$.
+The proposed encoding for the control-flow works also for encoding the partial graph.
+For example, the SMT-model $\fx(e_1)=1$, $\fx(e_2)=1$, $\fx(e_3)=0$ and $\fx(e_4)=0$ encodes the path $e_1 \rightarrow e_2$, satisfies $\phi_{CF}$.
 
@@
     \node[c] (ii)  [below of=i] {$e_2$};
+    \node[c,fill=black!15] (lnop) [below left = 0.25cm and 0.7cm of ii] {$e_{nop_{1}}$};
     \node[c] (iii) [below of=ii] {$e_3$};
+    \node[c,fill=black!15] (rnop) [right = 0.4cm of iii] {$e_{nop_{2}}$};
     \node[c] (iv)  [below of=iii] {$e_4$};
     \path[->]
-    (i) edge [right] node {$g_{1,2}$} (ii)
-    (ii) edge [left] node {$g_{2,3}$} (iii)
+    (i) edge [right] node {$\epsilon$} (ii)
+    (i) edge [right,bend right=25] node {$g_{1,4}$} (lnop)
+    (lnop) edge [right,bend right=25] node {$\epsilon$} (iv)
+    (ii) edge [right] node {$g_{2,3}$} (iii)
     (iii) edge [right] node {$\epsilon$} (iv)
-    (i) edge [left,bend right=45] node {$g_{1,4}$} (iv)
-    (ii) edge [right,bend left=55] node {$g_{2,4}$} (iv) %sloped
+    (ii) edge [right,bend left=25] node {$g_{2,4}$} (rnop)
+    (rnop) edge [right,bend left=25] node {$\epsilon$} (iv)
     ;
@@
                  \ & \land [\fx(e_3) \Rightarrow \fx(e_2)] \\
-                 \ & \land [\fx(e_4) \Rightarrow (\fx(e_1) \lor \fx(e_2) \lor \fx(e_3))] \\
-                 \ & \land [\fx(e_1) \land \fx(e_2) \Rightarrow g_{1,2}] \\
-                 \ & \land [\fx(e_1) \land \fx(e_4) \Rightarrow g_{1,4}] \\
-                 \ & \land [\fx(e_2) \land \fx(e_3) \Rightarrow g_{2,3}] \\
-                 \ & \land [\fx(e_2) \land \fx(e_4) \Rightarrow g_{2,4}] \\
-                 \ & \land \lnot [\fx(e_2) \land \fx(e_4)] \\
-                 \ & \land \lnot [\fx(e_3) \land \fx(e_4)]
+                 \ & \land [\fx(e_{nop_{1}}) \Rightarrow \fx(e_1)] \\
+                 \ & \land [\fx(e_{nop_{2}}) \Rightarrow \fx(e_2)] \\
+                 \ & \land [\fx(e_4) \Rightarrow (\fx(e_{nop_{1}}) \lor \fx(e_3) \lor \fx(e_{nop_{2}}))] \\
+                 \ & \land [\fx(e_{nop_{1}}) \land \fx(e_1) \Rightarrow g_{1,4}] \\
+                 \ & \land [(\fx(e_3) \land \fx(e_2)) \Rightarrow g_{2,3}] \\
+                 \ & \land [(\fx(e_{nop_{2}}) \land \fx(e_2)) \Rightarrow g_{2,4}] \\
+                 \ & \land \lnot [\fx(e_2) \land \fx(e_{nop_{1}})] \\
+                 \ & \land \lnot [\fx(e_3) \land \fx(e_{nop_{2}})]
 \end{align*}
@@
 \end{figure}
-
-\textbf{TODOOOOOOOOOOOOOOOOOOOOOOOOOOOO!!!!!!!!!!!!!!!!!!!!!!! NOPS IN THIS PICTURE!!!!}
 

### 2018-05-20 4_impl.tex
Cleanup some old todos.

@@
 
-The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos{} tool).
+The previous version of \porthos{} did not distinguish the event-based program model from the high-level AST, they both were encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
 
+%Internal representations used in \porthos[1]
+
 %As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
@@
 \label{ch:impl:model:xgraph}
-
-<TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO!!!!!!!!!!!!!!!!!!!!!!>
-%TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
-\textit{\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}}
 

### 2018-05-20 slides/slides.tex
Change the slides on encoding.

@@
 
-\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}
+%\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}
+%\begin{itemize}
+%%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
+%\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
+%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
+%\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
+%\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
+%\end{itemize}
+%
+%\begin{figure}[b]
+%  \centering
+%  \begin{minipage}{\textwidth}
+%      \begin{subfigure}[b]{0.25\textwidth}
+%        \makebox[\textwidth]{
+%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+%            \node[c] (1) [] {$e_i$};
+%            \node[c] (2) [below of=1] {$e_{i+1}$};
+%            \path[->]
+%            (1) edge [] node {$\epsilon$} (2)
+%            ;
+%        \end{tikzpicture}
+%        }
+%        \caption{The sequence}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[b]{0.35\textwidth}
+%        \makebox[\textwidth] {
+%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+%            \node[c]           (k)   [] {$e_k$};
+%            \node[c,draw=none] (iii) [below of=k] {$...$};
+%            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
+%            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
+%            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
+%            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
+%            \path[->]
+%            (k) edge [above] node {$g_{i,k}$} (i)
+%            (k) edge [] node {} (ii)
+%            (k) edge [] node {} (iii)
+%            (k) edge [] node {} (iv)
+%            (k) edge [above] node {$g_{i+j,k}$} (v)
+%            ;
+%        \end{tikzpicture}
+%        }
+%        \caption{Conditional branching}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[b]{0.3\textwidth}
+%        \makebox[\textwidth]{
+%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
+%            \node[c] (i) [] {$e_i$};
+%            \node[c,draw=none] (ii) [right=0.2cm of i] {};
+%            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
+%            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
+%            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
+%            \node[c] (k) [below of=iii] {$e_k$};
+%            \path[->]
+%            (i) edge [above] node {$g_{i,k}$} (k)
+%            (ii) edge [above] node {} (k)
+%            (iii) edge [] node {} (k)
+%            (iv) edge [] node {} (k)
+%            (v) edge [] node {} (k) %sloped
+%            ;
+%        \end{tikzpicture}
+%        }
+%        \caption{Branch merging}
+%    \end{subfigure}
+%    %
+%    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
+%    \label{encode:cf}
+%\end{minipage}
+%\end{figure}
+%\end{frame}
 
-\begin{itemize}
-%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
-\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
-$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
-\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
-\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
-\end{itemize}
+%\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}
+%\begin{itemize}
+%\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
+%\item The control-flow of the program is encoded as following:
+%\end{itemize}
+%\vspace{-20pt}
+%\begin{align}
+%\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
+%\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
+%                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
+%                       \land & \ \cdots \nonumber \nonumber\\
+%                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
+%                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
+%\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
+%\end{align}
+%\end{frame}
 
-\begin{figure}[b]
+\begin{frame}{Encoding for the control-flow: An example}
+%Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
+\begin{figure}
+\begin{minipage}{.2\textwidth}
 \centering
-  \begin{minipage}{\textwidth}
-      \begin{subfigure}[b]{0.25\textwidth}
-        \makebox[\textwidth]{
 \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-            \node[c] (1) [] {$e_i$};
-            \node[c] (2) [below of=1] {$e_{i+1}$};
+    \node[c] (i)   [] {$e_1$};
+    \node[c] (ii)  [below of=i] {$e_2$};
+    \node[c,fill=black!15] (lnop) [below left = 0.25cm and 0.7cm of ii] {$e_{nop_{1}}$};
+    \node[c] (iii) [below of=ii] {$e_3$};
+    \node[c,fill=black!15] (rnop) [right = 0.4cm of iii] {$e_{nop_{2}}$};
+    \node[c] (iv)  [below of=iii] {$e_4$};
     \path[->]
-            (1) edge [] node {$\epsilon$} (2)
+    (i) edge [right] node {$\epsilon$} (ii)
+    (i) edge [right,bend right=25] node {$g_{1,4}$} (lnop)
+    (lnop) edge [right,bend right=25] node {$\epsilon$} (iv)
+    (ii) edge [right] node {$g_{2,3}$} (iii)
+    (iii) edge [right] node {$\epsilon$} (iv)
+    (ii) edge [right,bend left=25] node {$g_{2,4}$} (rnop)
+    (rnop) edge [right,bend left=25] node {$\epsilon$} (iv)
     ;
 \end{tikzpicture}
-        }
-        \caption{The sequence}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.35\textwidth}
-        \makebox[\textwidth] {
-        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-            \node[c]           (k)   [] {$e_k$};
-            \node[c,draw=none] (iii) [below of=k] {$...$};
-            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
-            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
-            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
-            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
-            \path[->]
-            (k) edge [above] node {$g_{i,k}$} (i)
-            (k) edge [] node {} (ii)
-            (k) edge [] node {} (iii)
-            (k) edge [] node {} (iv)
-            (k) edge [above] node {$g_{i+j,k}$} (v)
-            ;
-        \end{tikzpicture}
-        }
-        \caption{Conditional branching}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.3\textwidth}
-        \makebox[\textwidth]{
-        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-            \node[c] (i) [] {$e_i$};
-            \node[c,draw=none] (ii) [right=0.2cm of i] {};
-            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
-            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
-            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
-            \node[c] (k) [below of=iii] {$e_k$};
-            \path[->]
-            (i) edge [above] node {$g_{i,k}$} (k)
-            (ii) edge [above] node {} (k)
-            (iii) edge [] node {} (k)
-            (iv) edge [] node {} (k)
-            (v) edge [] node {} (k) %sloped
-            ;
-        \end{tikzpicture}
-        }
-        \caption{Branch merging}
-    \end{subfigure}
+\end{minipage}
 %
-    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
-    \label{encode:cf}
+\hspace{15pt}
+\begin{minipage}{.4\textwidth}
+\centering
+\begin{align*}
+\phi_{CF} \defeq \ & [\fx(e_2) \Rightarrow \fx(e_1)] \\
+                 \ & \land [\fx(e_3) \Rightarrow \fx(e_2)] \\
+                 \ & \land [\fx(e_{nop_{1}}) \Rightarrow \fx(e_1)] \\
+                 \ & \land [\fx(e_{nop_{2}}) \Rightarrow \fx(e_2)] \\
+                 \ & \land [\fx(e_4) \Rightarrow (\fx(e_{nop_{1}}) \lor \fx(e_3) \lor \fx(e_{nop_{2}}))] \\
+                 \ & \land [\fx(e_{nop_{1}}) \land \fx(e_1) \Rightarrow g_{1,4}] \\
+                 \ & \land [(\fx(e_3) \land \fx(e_2)) \Rightarrow g_{2,3}] \\
+                 \ & \land [(\fx(e_{nop_{2}}) \land \fx(e_2)) \Rightarrow g_{2,4}] \\
+                 \ & \land \lnot [\fx(e_2) \land \fx(e_{nop_{1}})] \\
+                 \ & \land \lnot [\fx(e_3) \land \fx(e_{nop_{2}})]
+\end{align*}
 \end{minipage}
+\caption{Example of encoding for the control-flow of the event-flow graph}
 \end{figure}
-
-\end{frame}
-
-\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}
-
-\begin{itemize}
-\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
-\item The control-flow of the program is encoded as following:
-\end{itemize}
-
-\vspace{-20pt}
-\begin{align}
-\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
-\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
-                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
-                       \land & \ \cdots \nonumber \nonumber\\
-                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
-                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
-\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
-\end{align}
-
 \end{frame}
@@
 \begin{frame}{Encoding for the data-flow}
-
 \begin{itemize}
@@
 \end{itemize}
-
 The data-flow of an event is encoded as following:
@@
     \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] \nonumber \\
-    \phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)] \nonumber
 \end{align}
+%    \phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)] \nonumber
 \end{frame}
@@
 \item it contained the parser code inlined directly into the grammar (hardly maintainable);
-\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\lstinline{=}', `\lstinline{:=}', `\lstinline{<-}' and `\lstinline{<:-}', each for different kinds of arguments);
+\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\texttt{=}', `\texttt{:=}', `\texttt{<-}' and `\texttt{<:-}', each for different kinds of arguments);
 \item restricted syntax for expressions.
@@
 \node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
-\node[rectangle,draw] (6inc) [below left=1cm and 3cm of 16] {$S_{-}$};
-\node[rectangle,draw] (6com) [right=3cm of 6inc] {$S_{+}$};
+\node[rectangle,draw] (6) [below left=1cm and 1cm of 16] {$S$};
 \node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
@@
 (35) edge [bend left=10] node {} (46)
-(26) edge [bend right=10] node {} (6com)
-(36) edge [bend right=5] node {} (6inc)
-(16) edge [bend left=20] node {} (6inc)
-(46) edge [bend left=20] node {} (6com)
+(26) edge [bend right=10] node {} (6)
+(36) edge [bend right=5] node {} (6)
+(16) edge [bend left=20] node {} (6)
+(46) edge [bend left=20] node {} (6)
 ;

### 2018-05-21 4_impl.tex
Add some results to eval.

@@
 The previous version of \porthos{} did not distinguish the event-based program model from the high-level AST, they both were encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
+%For 
 Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
@@
     \begin{enumerate}[label*=\arabic*.]
-      \item following the principles of simplicity and readability, and
-      \item clear and informative program output.
+      \item following the principles of simplicity and readability,
+      \item clear and informative program output, and
+      \item following the clear code style.
     \end{enumerate}
@@
 As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected outcome (for instance, if a parser failed to parse the string and the recovery algorithm is not described).
+One of the transparency principles is following the clear code style.
+This mainly means the clear and informative naming of classes, functions and fields.
+It also implies the unified ordering of methods in classes, minimised size of methods code, clear tabulation, etc.
 %The modular architecture of \porthos[2] allow each its module to be tested independently.

### 2018-05-21 5_eval.tex
Cleanup some old todos.
Add some results to eval.

@@
 \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test_unrolled.png}
-\caption{An example of an unrolled event-flow graph represented in Figure~\ref{ex:compilation}}
+\caption{Unrolling of the event-flow graph presented in Figure~\ref{ex:compilation}}
 \label{ex:unrolling}
@@
 
-\subsection{Reachability analysis}
+\section{Performance}
 
-\subsection{Performance}
+\subsection{State reachability analysis}
 
-Measure time by \porthos[1], \porthos[2]:
+As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
 
-- enc-time absolute values
+For checking the reachability of the final state of the last states (sink events) of unrolled graphs compiled for a certain hardware architecture, \porthos[2] encodes both program and hardware memory model constraints as it was discussed in Section~\ref{ch:enc:bmc}.
+%The result formula is then solved by the Z3
 
-- rate enc-time/solve-time
+%TO-DO: fix interface
+%To run \porthos[2] in reachability analysis mode, we use the command `\lstinline{mousquetaires.DartagnanApp -i benchmarks/litmus_C11/Dekker.c -s tso}'
+
+The standard output of the tool looks as following:
+\begin{lstlisting}
+$ java Dartagnan -i benchmarks/C11/Dekker.c -s tso
+Interpreting...
+Unrolling...
+Program encoding...
+Memory model encoding...
+#=59
+Solving...
+{
+  "result": "NonReachable",
+  "interpretationTimer": {
+    "elapsedTimeSec": 0.134
+  },
+  "unrollingTimer": {
+    "elapsedTimeSec": 0.031
+  },
+  "programEncodingTimer": {
+    "elapsedTimeSec": 0.14
+  },
+  "memoryModelEncodingTimer": {
+    "elapsedTimeSec": 2.984
+  },
+  "solvingTimer": {
+    "elapsedTimeSec": 0.037
+  },
+  "errors": []
+}
+\end{lstlisting}
+
+For time benchmarking we ran the tool 5 times and computed the median of the encoding time.
+Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
+The time was measured by the tool itself via the Java method \texttt{System.currentTimeMillis}.
+For \porthos[1] we specify the unrolling bound $k_1 = 2$ (which means that all loops are executed twice), and for \porthos[2] we specify the bound $k_2 = 32$ (which means approximately the same, because
+
+\porthos[2] shows the full encoding time for the Dekker's algorithm 2.699 sec (from which 0.152 sec spent for the program encoding, and other part spent for the memory model encoding).
+In contrast, \porthos[1] shows the encoding time 6.152 sec (from which 0.223 sec spent for the program encoding). %, which is considerably better result.
+Thus, the performance of the encoding stage has been improved \textbf{in 2.7 times}.
+
+As the time spent for the interpretation and unrolling stages is negligible comparing to the encoding time, we conclude that the new architecture implies minor performance overhead comparing to the previous version of the tool.
+The time spent by the SMT solver while considering the SMT formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
+The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
+Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
+We consider this improvement to be the result of technical optimisations applied to \porthos[2].
+The optimisations include:
+\begin{itemize}
+\item \textit{Memoisation} of the frequently requested calls. \\
+    For example, the code of \porthos[1] contains 39 calls that take the subset of a set of events.
+The pattern is the following: `\lstinline{program.getEvents().stream().filter(e -> e instanceof MemEvent || e instanceof Local).collect(Collectors.toSet())}', \\
+    10 of them are performed in a loop over all events (in \texttt{Domain.encode} method of \porthos[2]).
+    These calls were replaced by the lazily initialisable memoising methods of \texttt{XProgram}: \\
+\begin{lstlisting}[language=Java]
+public ImmutableSet<XMemoryEvent> getMemoryEvents() {
+    return memoryEvents != null
+            ? memoryEvents
+            : (memoryEvents = getAllNodesExceptSource(XMemoryEvent.class));
+}
+\end{lstlisting}
+\item Usage of \textit{immutable data structures} (as it was discussed in Chapter~\ref{ch:impl}).
+\end{itemize}
+
+<TODO: compare numbers of clauses of SMT formulas> !!!
+
+<TODO: profile both programs! And say sth about memory usage!>
+
+% state reachability
 

### 2018-05-21 5_eval.tex
Almost complete the eval chapter.

@@
 
-%TODO: resource trebovaniya
-
-\section{The \texttt{PorthosC} in work}
+\section{\texttt{PorthosC} in work}
+\label{ch:eval:show}
 
 \subsection{The X-graph compilation}
+\label{ch:eval:show:compil}
 
@@
 \subsection{The X-graph unrolling}
+\label{ch:eval:show:unrol}
 
@@
 
-\section{Performance}
+\section{Performance evaluation}
+\label{ch:eval:perf}
 
 \subsection{State reachability analysis}
+\label{ch:eval:perf:reach}
 
 As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
-
-For checking the reachability of the final state of the last states (sink events) of unrolled graphs compiled for a certain hardware architecture, \porthos[2] encodes both program and hardware memory model constraints as it was discussed in Section~\ref{ch:enc:bmc}.
+For checking the reachability of the final state of the last states (sink events) of unrolled graphs compiled for a certain hardware architecture, the tool encodes both program and hardware memory model constraints as it was discussed in Section~\ref{ch:enc:bmc}.
+For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
 %The result formula is then solved by the Z3
@@
 
-The standard output of the tool looks as following:
+The standard output of \porthos[2] looks as following:
 \begin{lstlisting}
@@
 Memory model encoding...
-#=59
+#=82(59)
 Solving...
@@
 The time was measured by the tool itself via the Java method \texttt{System.currentTimeMillis}.
-For \porthos[1] we specify the unrolling bound $k_1 = 2$ (which means that all loops are executed twice), and for \porthos[2] we specify the bound $k_2 = 32$ (which means approximately the same, because
+For \porthos[1] we specify the unrolling bound $k_1 = 2$ (which means that all loops are executed twice), and for \porthos[2] we specify the bound $k_2 = 32$ (which means approximately the same, because there are 16 events in one iteration).
 
@@
 In contrast, \porthos[1] shows the encoding time 6.152 sec (from which 0.223 sec spent for the program encoding). %, which is considerably better result.
-Thus, the performance of the encoding stage has been improved \textbf{in 2.7 times}.
+Thus, the performance of the encoding stage has been\textbf{ improved in 2.7 times}.
 
-As the time spent for the interpretation and unrolling stages is negligible comparing to the encoding time, we conclude that the new architecture implies minor performance overhead comparing to the previous version of the tool.
+As the time spent for the interpretation and unrolling stages is negligible comparing to the encoding time, we conclude that the new architecture implies insignificant performance overhead comparing to the previous version of the tool.
 The time spent by the SMT solver while considering the SMT formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
@@
 Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
-We consider this improvement to be the result of technical optimisations applied to \porthos[2].
-The optimisations include:
-\begin{itemize}
-\item \textit{Memoisation} of the frequently requested calls. \\
-    For example, the code of \porthos[1] contains 39 calls that take the subset of a set of events.
-The pattern is the following: `\lstinline{program.getEvents().stream().filter(e -> e instanceof MemEvent || e instanceof Local).collect(Collectors.toSet())}', \\
-    10 of them are performed in a loop over all events (in \texttt{Domain.encode} method of \porthos[2]).
-    These calls were replaced by the lazily initialisable memoising methods of \texttt{XProgram}: \\
+We consider this improvement to be the result of some technical optimisations applied to the encoding stage of \porthos[2].
+The major optimisation is the \textit{memoisation} of frequently requested calls.
+For example, the code of \porthos[1] contains 39 calls that take the subset of a set of events, 10 of them are performed in a loop over all events (in \texttt{Domain.encode} method of \porthos[2]).
+The pattern is the following: `\lstinline{program.getEvents().stream().filter(e -> e instanceof MemEvent || e instanceof Local).collect(Collectors.toSet())}'.
+%TODO: refactor X-program inheritance
+In \porthos[2], these calls were replaced by the lazily initialisable memoisation methods (see \texttt{XProgram}): \\
+
+\vspace{-1em}
 \begin{lstlisting}[language=Java]
@@
 \end{lstlisting}
-\item Usage of \textit{immutable data structures} (as it was discussed in Chapter~\ref{ch:impl}).
-\end{itemize}
+
+%TODO: say how much the \textit{immutable data structures} helped (as it was discussed in Chapter~\ref{ch:impl}).
 
@@
 
-<TODO: profile both programs! And say sth about memory usage!>
+<TODO: profile both programs! And say sth about memory usage! for both encoding and solving>
 
-% state reachability
+\subsection{Portability analysis}
+\label{ch:eval:perf:port}
 
-Compare sizes of formulas
+For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
 
-memory usage -- for both encoding and solving
+!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
 
-\section{Comparison with HERD}
+\begin{lstlisting}
+$ java Porthos -i benchmarks/C11/Dekker.c -s tso -t sc -mode=state
+Program encoding...
+Interpreting...
+Unrolling...
+Unrolling...
+Program encoding...
+Memory model encoding...
+#=82(59)
+#=82(59)
+Solving...
+{
+  "result": "StatePortable",
+  "iterations": 0,
+  "interpretationTimer": {
+    "elapsedTimeSec": 0.213
+  },
+  "unrollingTimer": {
+    "elapsedTimeSec": 0.011
+  },
+  "programEncodingTimer": {
+    "elapsedTimeSec": 0.075
+  },
+  "memoryModelEncodingTimer": {
+    "elapsedTimeSec": 6.733
+  },
+  "solvingTimer": {
+    "elapsedTimeSec": 0.099
+  },
+  "errors": []
+}
+\end{lstlisting}
 
-\subsection{Unique Features}
+Again, \porthos[2] shows better results: 6.733 sec comparing to 12.961 sec spent by \porthos[1].
 
-- (inherited from the first version) : two memory models, smt, ... (?)
+\subsection{New features}
+\label{ch:eval:perf:feat}
 
-- (new) -- ?
+The revised design of \porthos{} tool includes the following new features:
+\begin{itemize}
+\item Partial support for the \textit{Kernel litmus tests}.\\
+\porthos[2] is able to process simple litmus tests that use some macros of the Linux kernel (such as \texttt{READ\_ONCE} and \texttt{WRITE\_ONCE} for atomic load and store with relaxed memory ordering).
+For example, consider the SB litmus test for the Linux kernel (which is similar to the one presented in Figure~\ref{simple_wmm_x86}):
+\begin{lstlisting}
+{ int x = 0; int y = 0;}
 
-- kernel?? TODO: SUPPORT some kernel funcs.
+P0(volatile int* y, volatile int* x) {
+  int r0;
+  WRITE_ONCE(*x,1);
+  r0 = READ_ONCE(*y);
+}
 
-\subsection{Performance}
+P1(volatile int* y, volatile int* x) {
+  int r0;
+  WRITE_ONCE(*y,1);
+  r0 = READ_ONCE(*x);
+}
 
-- perhaps, better performance on a large test
+exists(x == 0 && y == 0)
+\end{lstlisting}
+If checking this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.
+
+\item \textbf{TODO! MORE!}
+\end{itemize}
+
+%\section{Comparison with HERD}

### 2018-05-21 slides/slides.tex
Cleanup slides style.

@@
 \usepackage[russian,english]{babel}
+
 \usepackage{tikz}
@@
   }
+}
+
+\newcommand*{\yellowemph}[1]{%
+  \tikz[baseline=(X.base)] \node[rectangle, fill=yellow, rounded corners, inner sep=0mm] (X) {#1};%
 }
@@
 %\usetheme{Luebeck}
-%\usetheme{Madrid}
+%\usetheme{Madrid}\colorbox{yellow}
 %\usetheme{Malmoe}
@@
 %\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}
-
 \title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}
@@
 \item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
-\item Review existing tools for memory model-aware analysis;
-\item Investigate existing architecture of \texttt{Porthos}, its strengths and weaknesses;
-\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
+\item Review the existing tools for memory model-aware analysis;
+\item Examine the existing architecture of \texttt{Porthos}, its strengths and weaknesses;
+\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
 \end{itemize}
@@
 $p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\hline
+\multicolumn{2}{|c|}{ exists ($r_p=0 \land r_q=0$) } \\
 \hline
@@
 $\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
-$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
-$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
-$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
-$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
+$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
+$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
+$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}}  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
+$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
 \end{tabular}
@@
 
-\begin{frame}{The weak memory model} {Testing the candidate executions}
+\begin{frame}{The weak memory model} {Testing candidate executions}
 \vspace{10pt}
@@
 \hline
+\multicolumn{2}{|c|}{ exists ($r_p=0 \land r_q=0$) } \\
+\hline
 \end{tabular}
@@
 \begin{figure}
-\centering
+%\centering
 \includegraphics[width=\linewidth]{img/candidate.png}
@@
 \end{figure}
+
 \end{frame}
 
-\begin{frame}{The weak memory model} {Testing the candidate executions}
+\begin{frame}{The weak memory model} {Testing candidate executions}
 \vspace{10pt}
@@
 %
-\raggedleft{
 \noindent\begin{minipage}{.25\linewidth}
@@
 \end{minipage}
-}
 \vspace{-5pt}
 \begin{figure}
-\centering
+%\centering
 \includegraphics[width=\linewidth]{img/candidate-red.png}
@@
 \begin{frame}{Evaluation}
+Much better.
+
 \alert{[to be done]}

### 2018-05-22 2_wmm.tex
Move x86 example from appendix to the text.

@@
 For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$.
-As an example of memory model definition in \cat{} language, the x86-TSO model~\cite{herd10tutorial} can be found in Appendix~\ref{apx:x86cat}.
-This memory model asserts acyclicity of communication relation, \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.
+As an example of memory model definition in \cat{} language, Figure~\ref{ex:x86-cat} presents the simplified x86-TSO model~\cite{herd10tutorial}.
+This memory model asserts  acyclicity of the communication relation (the union of \rf-, \fr- and \co- relations), \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.
+
+\begin{figure}[h]
+\begin{lstlisting}
+"X86 TSO"
+include "x86fences.cat"
+include "filters.cat"
+include "cos.cat"
+
+(* Uniproc check *)
+let com = rf | fr | co
+acyclic po-loc | com
+
+(* Atomic *)
+empty rmw & (fre;coe)
+
+(* Global happens-before *)
+#ppo
+let po_ghb = WW(po) | RM(po)
+
+#mfence
+include "x86fences.cat"
+
+#implied barriers
+let poWR = WR(po)
+let i1 = MA(poWR)
+let i2 = AM(poWR)
+let implied = i1 | i2
+
+let com = rfe | fr | co
+let ghb = mfence | implied | po_ghb | com
+show implied
+
+acyclic ghb as tso
+\end{lstlisting}
+\caption{The simplified x86-TSO memory model defined in the \cat{} language~\cite{herd10tutorial}}
+\label{ex:x86-cat}
+\end{figure}

### 2018-05-24 1_intro.tex
Add some comments to the paper.
Rework eval. Add summary. Some cleanup.

@@
 \end{itemize}
+%TODO!!!!!!!!!!!!!!!!!!! cite "mcKenney - Memory Barriers a Hardware View for Software Hackers.pdf" -- a table for multiple architectures
+
 %As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
@@
 %1) http://research.nvidia.com/sites/default/files/pubs/2017-04_Automated-Synthesis-of//ASPLOS_2017_Memory_Model_Verification.pdf
-% 2) kernel litmus tests % <----------------------------- OBIAZATELNO!!!
 
@@
 One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
-For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read).
+For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read) and be easily extensible for defining new functions (have the separate module for the purpose of a knowledge base).
 %interprocedural analysis
 
-As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name \textit{\porthos[2]}\,%
+As the general architecture and almost all components of \porthos{} were to be redesigned, the tool received a new name \textit{\porthos[2]}\,%
 %
-\footnote{Hereinafter with the name `\porthos{}' we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \porthos[2].}. %
+\footnote{Hereinafter with the names \textit{\porthos{}} and \textit{\porthos[1]} we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \textit{\porthos[2]}.}.
+%
 %

### 2018-05-24 3_encod.tex
Rework eval. Add summary. Some cleanup.

@@
 For that, we compile the \textit{Abstract Syntax Tree~(AST)} of the parsed C-code to the plain event-flow graph.
-We show %TODO! to show, actually!!!
-that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
+The new encoding is to be smaller than the old one used in \porthos{} since it does not produce new variables for each high-level statement of the input language.
 %The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
@@
 \begin{align}
-    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \\
-    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \\
-    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] % \\
+    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \label{enc:df:load} \\
+    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \label{enc:df:store} \\
+    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] \label{enc:df:eval} % \\
 \end{align}
 
+In Equation~\ref{enc:df:eval}, the function $\fv:\,\mathbb{C}\,\rightarrow\,\mathbb{R}$ e\textbf{v}aluates the computation event (the value is determined by \co- and \rf- relations).
 To convert the program into SSA form, for each event each variable that is declared so far (either local or shared) is mapped to its indexed reference; this information is stored in the SSA-map "event \textit{to} variable \textit{to} SSA-index".
@@
 
-In an SMT-formula, the relation $x \relarrow{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
-The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
+In the SMT-formula, a relation $x \relarrow{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
+Derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
 \begin{itemize}[noitemsep,topsep=0pt]

### 2018-05-24 4_impl.tex
Cleanup slides style.
Rework eval. Add summary. Some cleanup.

@@
-\chapter{The~\texttt{PorthosC}:~implementation}
+\chapter{PorthosC: The implementation}
 \label{ch:impl}
@@
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
+However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
 The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
@@
   \1 \texttt{XComputationEvent} \\
+  %TODO: say sth about how we not produce thousands of comput events
     We distinguish two types of computation events:
@@
     \2 the \co{}-relation edges, and 
-    \2 the \rf{}-relation edges. %TODO! implement 
+    \2 the \rf{}-relation edges.
 \end{outline}

### 2018-05-24 5_eval.tex
Cleanup slides style.
Rework eval. Add summary. Some cleanup.

--- 2018-05-20 5_eval.tex
@@
 
-%TODO: resource trebovaniya
+\section{Comparison to \porthos[1]}
+\label{ch:eval:show}
 
-\section{The \texttt{PorthosC} in work}
+\subsection{The compilation and unrolling}
+\label{ch:eval:show:compil}
 
-\subsection{The X-graph compilation}
+As an example of the compilation and unrolling processes, consider the simple
 
-As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, consider the tiny C test in Figure~\ref{ex:compilation}.
-This function does not perform any useful computation, however it contains several syntactic elements (such as prefix increment or loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
-The control-flow subgraph \xgraph[CF] of the non-unrolled event-flow graph is presented on the right-hand side of the Figure (all graphs generated by \porthos[2] are produced with the help of the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
-
-Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.
-Write events are denoted with the left-directed arrow `\lstinline{<-}', and the functions \lstinline{load} and \lstinline{store} denote the type of the shared memory event.
-The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
-The graph contains a single source event and a single sink event (represented by the grey triangles).
+%As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, 
+Consider two tiny meaningless function represented in Figure~\ref{ex:test-both-code} (in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
+The functions are state-equivalent (represent the same state machine) and contain three nested \texttt{while}-loops.
+On the example of these functions we will show the differences in compilation and loop unrolling stages between \porthos[1] and \porthos[2].
+%contains several syntactic elements (such as prefix increment and loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
+%In the Figure, the control-flow subgraph \xgraph[CF] of the non-unrolled event-flow graph is presented on the right-hand side (
 
 \begin{figure}[!h]
-\begin{minipage}{.4\textwidth}
+\begin{minipage}{.53\textwidth}
 \begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
-void test(int &x, int &y) {
-    int r;
-    while ((5 + 4/2 * x) 3 
-      == 1) {
-        if (y > 6)
-            continue;
-        else if (++y > 7) {
-            r = r + 10;
-            break;
+void t0(int& x) {
+     int a = 1, c = 1;
+     while (a == 1) {
+         int b = 1;
+         while (b == 1) {
+             while (c == 1) {
+                 x = c;
              }
-        r = 11;
+             x = b;
          }
-    y = x + 1;
+     }
+     x = a;
  }
@@
 %
-\begin{minipage}{.59\textwidth}
-  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test.png}
+\begin{minipage}{.45\textwidth}
+\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
+{ x }
+thread t0 {
+     a <- 1;
+     c <- 1;
+     while (a == 1) {
+         b <- 1;
+         while (b == 1) {
+             while (c == 1) {
+                 x := c
+             };
+             x := b
+         };
+     };
+     x := a
+ }
+\end{lstlisting}
+\end{minipage}
+\caption{Example: A demonstrative cyclic function}
+\label{ex:test-both-code}
+\end{figure}
+
+The functions are processed by \porthos[2] or modified version \porthos[1] that is able to print same \xgraph[CF] as the new tool (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bind by the method processing its parent).
+The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
+%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.
+
+The Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-code} are compiled to.
+The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].
+
+In both pictures, the writes are denoted with the left-directed arrow `\lstinline{<-}', and the functions \lstinline{load} and \lstinline{store} denote the type of the shared memory event.
+The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
+The graphs contains a single source event and a single sink event represented by the grey triangles (actually, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
+For clarity, all branching events, that in current example serve as the conditional events of loops, are highlighed with light-grey colour.
+
+\begin{figure}[!h]
+%
+\begin{minipage}{.49\textwidth}
+  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
 \end{minipage}
 %
-\caption{An example of a C function compiled to the event-flow graph}
-\label{ex:compilation}
+\begin{minipage}{.49\textwidth}
+  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
+\end{minipage}
+%
+\caption{The control-flow graphs of the functions presented in Figure~\ref{ex:test-both-code}}
+\label{ex:test-both-pic}
 \end{figure}
 
-The arguments \lstinline{x} and \lstinline{y} of the function are passed by reference, therefore they are treated as global variables.
-The local variable declaration `\lstinline{int r;}' produces no events (it is processed by the X-graph pre-compiler that invokes the memory manager to create the new local variable \lstinline{r}).
+Two compiled the graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, which are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}.
+However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} look differently as \porthos[2] uses different unrolling algorithm.
+The labels of events in the left-hand side picture (produced by \porthos[2]) are augmented by the unrolling depth number, which is separated from the event label by comma.
 
-The first event `\lstinline{load(reg_tmp0 <- x}' loads the value of the global variable \lstinline{x} into the temp register \lstinline{reg_tmp0} in order to satisfy the requirement that all computation must be performed over local variables;
-note that each element of the computational tree `\lstinline{eval(eval(eval(5 + eval(eval(4 / 2) * reg_tmp0)) % 3) == 1)}' is represented by a local memory unit (either \texttt{XComputationEvent} or \texttt{XConstant} or \texttt{XRegister}).
-The node of this computational event in the control-flow graph has two outgoing edges, the primary edge to the `\lstinline{load(reg_tmp1 <- y}', the first event of the then-branch of the while-loop, and the alternative edge to the \lstinline{nop}-event representing the only event of the else-branch.
+The unrolling algorithm used by \porthos[1] (right-hand side picture) unrolls \textit{all} loops $k$ times (where $k$ is the unrolling bound), and the unrolling algorithm of \porthos[2] unrolls loops so that not more than $k$ events are executed.
+As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times, which makes the old unrolling algorithm not complete).
+As the new unrolling algorithm uses the Deep-First Search, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus is complete.
 
-As is was discussed in Section~\ref{ch:impl:model:xgraph}, all computation events have no impact to the global state of the concurrent system.
-Therefore, the interpreter does not \textit{emit} computational events, but it \textit{creates} them.
-This means, when the \texttt{Y2XConverterVisitor} processes the expression tree `\lstinline{(5 + 4/2 * x) 3 == 1}', it calls the method \texttt{createComputationEvent} of the interpreter that only creates the computation event and does not change its state.
-However, once the converter meets the global variable \lstinline{x}, it calls the interpreter method \texttt{emitMemoryEvent} to copy its value to a temp register; since it meets the global variables involved to the computation before it ends to process the whole computation expression, the values of all these global variables will be copied to temp registers \textit{before} the computation expression is used by any other event.
-Note that if the computation expression has not been used by any other event (for example, as the constant \lstinline{1} in the following C code: `\lstinline{foo(); 1; bar();}'), it is lost from the model (by the term \textit{use} here we mean that the computation event is evaluated as a guard or assigned to another memory unit).
+Note that the unrolled graph produced by \porthos[2] does not necessarily become a tree after removing the sink node.
+Some branches of the graph are merged when the executions have the same event with the same unrolling depth number.
+For example, primary transitions of both events `\lstinline{[b <- 1, 8]}' and `\lstinline{[store(x <- b), 8]}' (produced by executions of the first iteration of the \texttt{while} loop) lead to the same event `\lstinline{[eval(b == 1), 9]}' (the first event of the second iteration of the second loop).
 
-\subsection{The X-graph unrolling}
+\begin{figure}[!h]
+%
+\begin{minipage}{.6\textwidth}
+  \includegraphics[height=.6\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
+\end{minipage}
+%
+\begin{minipage}{.39\textwidth}
+  \includegraphics[height=\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/Porthos/t0_unrolled.png}
+\end{minipage}
+%
+\caption{The unrolled control-flow graphs of the functions presented in Figure~\ref{ex:test-both-code}}
+\label{ex:test-both-pic-unroll}
+\end{figure}
 
-Once the \xgraph[CF] is constructed, it should be unrolled to an acyclic flow-graph (as it was discussed in Section~\ref{ch:impl:proc:x-unroll}).
-The Figure~\ref{ex:unrolling} shows the control-flow subgraph \xgraphU{CF} of the event-flow graph from the previous example (see Figure~\ref{ex:compilation}) unrolled up to bound $k=16$.
-The labels of events in the picture are augmented by the event reference-id number (the unrolling depth), separated from the event value by the comma.
+%The arguments \lstinline{x} and \lstinline{y} of the function are passed by reference, therefore they are treated as global variables.
+%The local variable declaration `\lstinline{int r;}' produces no events (it is processed by the X-graph pre-compiler that invokes the memory manager to create the new local variable \lstinline{r}).
+%The first event `\lstinline{load(reg_tmp0 <- x}' loads the value of the global variable \lstinline{x} into the temp register \lstinline{reg_tmp0} in order to satisfy the requirement that all computation must be performed over local variables;
+%note that each element of the computational tree `\lstinline{eval(eval(eval(5 + eval(eval(4 / 2) * reg_tmp0)) % 3) == 1)}' is represented by a local memory unit (either \texttt{XComputationEvent} or \texttt{XConstant} or \texttt{XRegister}).
+%The node of this computational event in the control-flow graph has two outgoing edges, the primary edge to the `\lstinline{load(reg_tmp1 <- y}', the first event of the then-branch of the while-loop, and the alternative edge to the \lstinline{nop}-event representing the only event of the else-branch.
 
-Note that the graph does not become a tree after removing the sink node.
-Some branches of the graph are merged when the executions have the same event with the same unrolling depth number.
-For example, primary transitions of both events `\lstinline{[r <- 11, 10]}' and `\lstinline{[jump(), 10]}' (produced by executions of the first iteration of the \texttt{while} loop) lead to the same event `\lstinline{[load(reg_tmp0 <- x), 11]}' (the first event of the second loop iteration).
+%As is was discussed in Section~\ref{ch:impl:model:xgraph}, all computation events have no impact to the global state of the concurrent system.
+%Therefore, the interpreter does not \textit{emit} computational events, but it \textit{creates} them.
+%This means, when the \texttt{Y2XConverterVisitor} processes the expression tree `\lstinline{(5 + 4/2 * x) 3 == 1}', it calls the method \texttt{createComputationEvent} of the interpreter that only creates the computation event and does not change its state.
+%However, once the converter meets the global variable \lstinline{x}, it calls the interpreter method \texttt{emitMemoryEvent} to copy its value to a temp register; since it meets the global variables involved to the computation before it ends to process the whole computation expression, the values of all these global variables will be copied to temp registers \textit{before} the computation expression is used by any other event.
+%Note that if the computation expression has not been used by any other event (for example, as the constant \lstinline{1} in the following C code: `\lstinline{foo(); 1; bar();}'), it is lost from the model (by the term \textit{use} here we mean that the computation event is evaluated as a guard or assigned to another memory unit).
 
-%(said before in 4_impl:unrolling)
+%\subsection{The X-graph unrolling}
+%\label{ch:eval:show:unrol}
+
+%Once the \xgraph[CF] is constructed, it should be unrolled to an acyclic flow-graph (as it was discussed in Section~\ref{ch:impl:proc:x-unroll}).
+%The Figure~\ref{ex:unrolling} shows the control-flow subgraph \xgraphU[CF] of the event-flow graph from the previous example (see Figure~\ref{ex:compilation}) unrolled up to bound $k=16$.
+%
+
+%TODO: here (or better in Chap 5 ) say about 'we tried different encoding schemes !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
 %Also note that comparing to the unrolling bound description given in Section~\ref{ch:impl:proc:x-unroll}, 
@@
 
-\begin{figure}[!h]
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test_unrolled.png}
-\caption{An example of an unrolled event-flow graph represented in Figure~\ref{ex:compilation}}
-\label{ex:unrolling}
-\end{figure}
+%\begin{figure}[!h]
+%\includegraphics[width=\textwidth,height=\textheight]{img/my/graphs/test_unrolled.png}
+%\caption{Unrolling of the event-flow graph presented in Figure~\ref{ex:compilation}}
+%\label{ex:unrolling}
+%\end{figure}
 
-\subsection{Reachability analysis}
+\section{Performance evaluation}
+\label{ch:eval:perf}
 
-\subsection{Performance}
+As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
+For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
 
-Measure time by \porthos[1], \porthos[2]:
+\subsection{State reachability analysis}
+\label{ch:eval:perf:reach}
 
-- enc-time absolute values
+%For checking reachability of the final states of the unrolled graphs compiled for a certain hardware architecture, the tool encodes both program and hardware memory model constraints as it was discussed in Section~\ref{ch:enc:bmc}.
+%The result formula is then solved by the Z3
 
-- rate enc-time/solve-time
+%TO-DO: fix interface
+%To run \porthos[2] in reachability analysis mode, we use the command `\lstinline{mousq
 
-Compare sizes of formulas
+%\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
+When running in the state reachability analysis mode, \porthos[2] produces the following output:
 
-memory usage -- for both encoding and solving
+\begin{lstlisting}
+$ java PorthosC --reachability --input benchmarks/C11/Dekker.c \
+      --bound 27 --target TSO
+0.288: Interpretation...
+0.426: Unrolling...
+0.533: Program encoding...
+0.701: Program domain encoding...
+#=193 (107)
+34.187: Memory model encoding...
+36.818: Solving...
+{
+  "result": "NonReachable",
+  "mainTimer": {
+    "elapsedTimeSec": 36.163
+  },
+  "timers": {
+    "Interpretation": {
+      "elapsedTimeSec": 0.132
+    },
+    "ProgramDomainEncoding": {
+      "elapsedTimeSec": 32.486
+    },
+    "Solving": {
+      "elapsedTimeSec": 0.345
+    },
+    "Unrolling": {
+      "elapsedTimeSec": 0.061
+    },
+    "MemoryModelEncoding": {
+      "elapsedTimeSec": 2.631
+    },
+    "ProgramEncoding": {
+      "elapsedTimeSec": 0.168
+    }
+  },
+  "errors": []
+}
+\end{lstlisting}
 
-\section{Comparison with HERD}
+For time benchmarking we ran the tool $5$ times and computed the median of the encoding time.
+Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
+The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.
 
-\subsection{Unique Features}
+%For \porthos[1] we specify the unrolling bound $k_1 = 2$, and for \porthos[2] the bound is $k_2 = 21$ ().
+As the unrolling algorithm has been changed, the number of events after unrolling differs considerably.
+Therefore, the correct performance comparison with \porthos[1] is not manageable as the performance of the tools depends directly on number of events.
+For example, with the bound $k_1=2$ \porthos[1] produces $59$ events (among them, $51$ shared-memory events) with the program domain encoding time 4.608 sec,
+and with the bound $k_1=3$ \porthos[1] produces $95$ events (among them, $107$ shared-memory events) with the program domain encoding time 21.475 sec;
+whereas \porthos[2] with the bound $k_2=17$ produces $57$ ($31$) events with the program domain encoding time 1.734, and with the bound $k_2=27$ it produces $193$ events (among them, $107$ shared-memory events, which equal to the number of shared-memory events produced by \porthos[1] with the bound $k_1=3$) with the time 32.673 sec.
+Therefore, the new unrolling scheme has made the analysis complete but at the same time it reduced the performance considerably.
+%The last result shows that with the bounds $k_1=3$ and $k_2=27$ both tools have executed same number of loops  almost equivalent as they
+%However, the program domain encoding and memory model encoding schemes have not been changed considerably comparing to \porthos[1], therefore the execution time 
+%Although, the performance depends directly on number of events,
 
-- (inherited from the first version) : two memory models, smt, ... (?)
+%\porthos[2] shows the full encoding time for the Dekker's algorithm 2.699 sec (from which 0.152 sec spent for the program encoding, and other part spent for the memory model encoding).
+%In contrast, \porthos[1] shows the encoding time 6.152 sec (from which 0.223 sec spent for the program encoding). %, which is considerably better result.
+%Thus, the performance of the encoding stage has been\textbf{ improved in 2.7 times}.
 
-- (new) -- ?
+<TODO: above as table>
 
-- kernel?? TODO: SUPPORT some kernel funcs.
+Note, that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
+Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
+The time spent by the SMT solver while considering the SMT formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
+The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
+Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
 
-\subsection{Performance}
+%We consider this improvement to be the result of some technical optimisations applied to the encoding stage of \porthos[2].
+%The major optimisation is the \textit{memoisation} of frequently requested calls.
+%For example, the code of \porthos[1] contains 39 calls that take the subset of a set of events, 10 of them are performed in a loop over all events (in \texttt{Domain.encode} method of \porthos[2]).
+%The pattern is the following: `\lstinline{program.getEvents().stream().filter(e -> e instanceof MemEvent || e instanceof Local).collect(Collectors.toSet())}'.
+%%TODO: refactor X-program inheritance
+%In \porthos[2], these calls were replaced by the lazily initialisable memoisation methods (see \texttt{XProgram}): \\
+%
+%\vspace{-1em}
+%\begin{lstlisting}[language=Java]
+%public ImmutableSet<XMemoryEvent> getMemoryEvents() {
+%    return memoryEvents != null
+%        ? memoryEvents
+%        : (memoryEvents = getAllNodesExceptSource(XMemoryEvent.class));
+%}
+%\end{lstlisting}
 
-- perhaps, better performance on a large test
+%TODO: say how much the \textit{immutable data structures} helped (as it was discussed in Chapter~\ref{ch:impl}).
+
+%<TODO: compare numbers of clauses of SMT formulas>
+
+<TODO: profile both programs, say sth about memory usage!>
+
+\subsection{Portability analysis}
+\label{ch:eval:perf:port}
+
+For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
+
+%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
+
+\begin{lstlisting}
+$ java PorthosC --portability --input benchmarks/C11/Dekker.c \
+      --bound 27 --source TSO --target SC --state
+0.359: Interpretation...
+0.364: Unrolling...
+0.590: Unrolling...
+0.623: Program encoding...
+0.847: Program domain encoding...
+#=193 (107)
+#=193 (107)
+66.541: Memory model encoding...
+72.321: Solving...
+{
+  "result": "StatePortable",
+  "iterations": 0,
+  "mainTimer": {
+    "elapsedTimeSec": 73.02
+  },
+  "timers": {
+    "ProgramEncoding": {
+      "elapsedTimeSec": 0.223
+    },
+    "MemoryModelEncoding": {
+      "elapsedTimeSec": 5.779
+    },
+    "Interpretation": {
+      "elapsedTimeSec": 0.248
+    },
+    "Solving": {
+      "elapsedTimeSec": 0.699
+    },
+    "ProgramDomainEncoding": {
+      "elapsedTimeSec": 65.693
+    },
+    "Unrolling": {
+      "elapsedTimeSec": 0.022
+    }
+  },
+  "errors": []
+}
+\end{lstlisting}
+
+As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).
+
+\subsection{New features}
+\label{ch:eval:perf:feat}
+
+%The revised design of \porthos{} tool includes the following new features:
+Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
+Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \texttt{READ\_ONCE} and \texttt{WRITE\_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
+Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler:hooking}.
+
+As an example, consider the SB litmus test for the Linux kernel (which is similar to the one presented in Figure~\ref{simple_wmm_x86}):
+\begin{lstlisting}
+{ int x = 0; int y = 0;}
+
+P0(volatile int* y, volatile int* x) {
+  int r0;
+  WRITE_ONCE(*x,1);
+  r0 = READ_ONCE(*y);
+}
+
+P1(volatile int* y, volatile int* x) {
+  int r0;
+  WRITE_ONCE(*y,1);
+  r0 = READ_ONCE(*x);
+}
+
+exists(x == 0 && y == 0)
+\end{lstlisting}
+
+If verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

### 2018-05-24 99_summary.tex
add a bit up to hooks.
polish text a bit.
write interpreter. update pictures.
finish impl tentatively. add draft for eval.
consider some Professor's comments.
cleanup a bit
Almost complete the eval chapter.
Rework eval. Add summary. Some cleanup.

@@
-\chapter{Summary}
+\chapter{Conclusion}
 \label{ch:summary}
 
-enumerate goals in past tense + justification
+\section{Solved tasks and contributions}
 
-contribution
+During the work on this thesis, we solved the following engineering and scientific tasks:
+
+\begin{enumerate}[label=\Roman*.]
+\item
+Firstly, we studied existing weak memory model-aware analysis frameworks and tools to gather deep understanding of the problem (Sections~\ref{ch:intro:related} and~\ref{ch:enc:mc} and Chapter~\ref{ch:wmm}).
+
+\item
+Then, we explored the existing implementation of the proof-of-concept tool \porthos[1] in order to understand components that need to be redesigned in order to facilitate the extension of the input language.
+
+  \begin{enumerate}[label=\roman*.,leftmargin=\parindent]
+  \item The first stumbling block in the input language extension was the \porthos[1] input language parser, which performs the full semantic analysis of the program.
+  Although this works for a small toy language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
+  For handling that, we implemented the processing units that operate on the \textit{pre-compilation} stage discussed in Section~\ref{ch:impl:proc:x-pre-compiler}.
+
+  \item Secondly, as \porthos[2] determines the \textit{kind (shared or local) of each variable} on the pre-compilation phase, the shared variables can be accessed in arbitrary code location (comparing to \porthos[1] which determines the type of a variable syntactically, depending on the operator of function that uses the variable).
+
+  \item In addition, \porthos[2] supports some \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\texttt{int a, b=2, z;}'), arbitrary expressions (corresponding to the C standard), and others (see Section~\ref{ch:impl:input}).
+
+  \item As the C language supports unconditional \texttt{goto}-jumps that can produce an arbitrary control-flow graph, which cannot be supported solely with the program AST, therefore we needed to compile the AST to the low-level hardware-agnostic program representation \xgraph{}.
+  For that, we implemented the \textit{full C interpreter} discussed in Section~\ref{ch:impl:proc:x-compiler}.
+
+  \item Originally, the control-flow instructions were encoded directly into the SMT-formula~\cite{Porthos17a}.
+  In contrast, \porthos[2] encodes the low-level \xgraph{} representation, which can have arbitrary graph structure.
+  For encoding the \xgraph{} into the SMT-formula, we apply the \textit{new control-flow encoding scheme} that in general follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} (see Section~\ref{ch:enc:bmc:cf}).
+  As the new encoding scheme does not add new variables to every control-flow instruction, the number of variables in the result SMT-formula is expected to be smaller.
+
+  \item For ease the process of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking mechanism}, which allows to modify the compilation process without changing the interpreter.
+  The invocation hooks serve as a knowledge base for the program domain, that are to be extended in future.
+  We illustrate this mechanism with the basic support for Linux kernel litmus tests.
+
+  \item Since the new tool compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from the simple unrolling all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores all possible states that the program can reach within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
+  Although the new algorithm produces considerably more events than the old algorithm and thus the requires more time for the program domain encoding, we claim it to be complete.
+
+  \end{enumerate}
+
+\item 
+As the tests show, the overhead of the new architecture is negligible, therefore we consider the applied architectural changes as acceptable.
+
+\end{enumerate}
+
+Thus, the new tool \porthos[2] with easily extensible knowledge base that allow to support new domain-specific functions can be considered as a generalised framework for SMT-based memory model-aware analysis.
+
+%enumerate goals (from Task specification) in past tense + justification
+
+\section{Limitations and future work}
+
+- improve the encoding: an abstraction over Z
+
+- typing: const-size arrays + structures and so on
+  -only basic datatypes
+
+- hooks : flexible.
+but still in Java, not as separate module (as it requires turing-complete language to be expressive enough) => extension of the knowledge base for supporting new platforms / languages
+
+- new modes: as it was discussed <before>, works only in the litmus test-mode. May work with complete projects. But here: 
+  - memory issues (we store same graph + back edges + memoised queries ==> multiple instances.);
+    - + will need the memory-guarding module
+  - ability to encode partially (once the
+
+-only one mode litmus-test intra proc
+
+- testing with new solvers

### 2018-05-24 5_eval.tex
Write some words on limitations.

@@
 %As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, 
-Consider two tiny meaningless function represented in Figure~\ref{ex:test-both-code} (in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
-The functions are state-equivalent (represent the same state machine) and contain three nested \texttt{while}-loops.
-On the example of these functions we will show the differences in compilation and loop unrolling stages between \porthos[1] and \porthos[2].
+Consider two tiny equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:both-cf}.
+%(in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
+%The functions are state-equivalent (represent the same state machine) and 
+The functions does not compute any useful value, however they contain three nested \texttt{while}-loops and thus can serve as an illustration of the differences in the program compilation and unrolling between \porthos[1] and \porthos[2].
 %contains several syntactic elements (such as prefix increment and loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
@@
 \begin{figure}[!h]
-\begin{minipage}{.53\textwidth}
+%\begin{minipage}{.53\textwidth}
+\begin{subfigure}[b]{.53\textwidth}\centering
 \begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
-void t0(int& x) {
-     int a = 1, c = 1;
+
+void t0(int &x) {
+     int a = 1;
+     int c = 1;
      while (a == 1) {
@@
  }
+
 \end{lstlisting}
-\end{minipage}
-%
-\begin{minipage}{.45\textwidth}
+\caption{An example in the C language}
+\label{ex:both-cf:ptsC}
+\end{subfigure}
+%\begin{minipage}{.45\textwidth}
+\begin{subfigure}[b]{.45\textwidth}\centering
 \begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
@@
 \end{lstlisting}
-\end{minipage}
+\caption{An example in the \porthos[1] input language}
+\label{ex:both-cf:pts1}
+\end{subfigure}
 \caption{Example: A demonstrative cyclic function}
-\label{ex:test-both-code}
+\label{ex:both-cf}
 \end{figure}
@@
 
-The Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-code} are compiled to.
+The Figure~\ref{ex:test-botrepresentedh-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-code} are compiled to.
 The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].
@@
 %
-\begin{minipage}{.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}\centering
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
-\end{minipage}
+  \hfill
+  \caption{The compiled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
+  \label{ex:both-cf:graph:ptsC}
+\end{subfigure}
 %
-\begin{minipage}{.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}\centering
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
-\end{minipage}
+  \hfill
+  \caption{The compiled AST of the function in Figure~\ref{ex:both-cf:pts1}}
+  \label{ex:both-cf:graph:pts1}
+\end{subfigure}
 %
-\caption{The control-flow graphs of the functions presented in Figure~\ref{ex:test-both-code}}
+\caption{The control-flow graphs of the functions represented in Figure~\ref{ex:both-cf}}
 \label{ex:test-both-pic}
@@
 
-Two compiled the graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, which are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}.
-However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} look differently as \porthos[2] uses different unrolling algorithm.
+Two compiled the graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, that are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}, and \texttt{skip}-events in \porthos[1] graph.
+However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} are different as \porthos[2] uses different unrolling algorithm.
 The labels of events in the left-hand side picture (produced by \porthos[2]) are augmented by the unrolling depth number, which is separated from the event label by comma.
@@
 %
-\begin{minipage}{.6\textwidth}
+\begin{subfigure}[b]{.6\textwidth}\centering
   \includegraphics[height=.6\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
-\end{minipage}
+  \hfill
+  \caption{The unrolled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
+  \label{ex:both-cf:graphU:ptsC}
+\end{subfigure}
+\hfill
 %
-\begin{minipage}{.39\textwidth}
-  \includegraphics[height=\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/Porthos/t0_unrolled.png}
-\end{minipage}
+\begin{subfigure}[b]{.3\textwidth}\centering
+  \includegraphics[height=.9\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/Porthos/t0_unrolled.png}
+    \caption{The unrolled AST of the function in Figure~\ref{ex:both-cf:pts1}}
+    \hfill
+  \label{ex:both-cf:graphU:pts1}
+\end{subfigure}
+\hfill
 %
-\caption{The unrolled control-flow graphs of the functions presented in Figure~\ref{ex:test-both-code}}
+\caption{The unrolled control-flow graphs of the functions represented in Figure~\ref{ex:test-both-code}}
 \label{ex:test-both-pic-unroll}
@@
 
-<TODO: above as table>
-
 Note, that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
@@
 Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
+
+<TODO: above as table>
 

### 2018-05-24 99_summary.tex
Write some words on limitations.

@@
 
-  \item For ease the process of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking mechanism}, which allows to modify the compilation process without changing the interpreter.
-  The invocation hooks serve as a knowledge base for the program domain, that are to be extended in future.
+  \item For ease the process of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking}, a flexible mechanism for intercepting the compilation process without changing the interpreter.
+  The invocation hooking mechanism serves as a knowledge base for the program domain, that is to be filled and extended in future.
+  Invocation hooks are defined in Java and thus are flexible, though their extension and modification requires some knowledge of the internals of the tool.
   We illustrate this mechanism with the basic support for Linux kernel litmus tests.
@@
 
-- improve the encoding: an abstraction over Z
+Current implementation of \porthos[2] has the following limitations, that might define the direction of the future work.
+
+\begin{itemize}
+\item The major flaw of \porthos[2] as a tool that performs the complete software analysis is the its sensitivity to the combinatorial explosion.
+%TODO: show it in table , reference here the table
+As it was shown in Section~\ref{ch:eval:perf}, the number of events generated by the program grow rapidly as the user increases the unrolling bound.
+One possible way to reduce the number of states of the program might be applying some traditional techniques for fighting the state explosion problem (such as concrete execution as a part of concolic execution~\cite{majumdar2007hybrid}), however this must be done \textit{with considering the weak memory model}, because otherwise it could possibly lead to the loss of states.
+% cannot be applied to our problem \textit{before} the weak memory model is considered, 
+%One possible enhancement, that could treat but not cure this flaw, could be an extra analysis stage carried before the unrolling, that analyses
+
+%\item As its predecessor, \porthos[2] does not propose any heuristic for finding the optimal unrolling bound.
+
+\item As its predecessor, \porthos[2]
+
+\end{itemize}
+
+%- improve the encoding: an abstraction over Z
+- testing with new solvers
 
@@
 
-- hooks : flexible.
+%- hooks : flexible.
 but still in Java, not as separate module (as it requires turing-complete language to be expressive enough) => extension of the knowledge base for supporting new platforms / languages
@@
 -only one mode litmus-test intra proc
-
-- testing with new solvers

### 2018-05-25 4_impl.tex
Add example on invocation hook.

@@
 
-Current version of \porthos[2] contains two invocation hooks, the \texttt{XPorthosInvocationHook} for describing the compatibility-mode functions of the input \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
+Current version of \porthos[2] contains two invocation hooks, the \texttt{XLegacyInvocationHook} for intercepting the compatibility-mode functions of the \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
+%Note that the same mechanism (with several modifications) can be used for processing assembly language calls.
 An invocation hook can model any type of operations (memory, fence, computational, etc.).
-For instance, the \texttt{XPorthosInvocationHook} intercepts calls \texttt{`location.store(atomic,register)'} and processes the store operation with respect to the specified atomic.
-%TODO: more about described functions
-%TODO: rename InvocationHook -> XInvocationHook
+%For instance, the \texttt{XLegacyInvocationHook} intercepts calls \lstinline{`location.store(atomic, register)'} and processes the store operation with respect to the specified atomic.
+For instance, the \texttt{XKernelInvocationHook} intercepts the call \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces the invocation with its hook action shown in Figure~\ref{code-invoc-hook}.
+%and processes the store operation with respect to the following .
+
+\begin{figure}[!h]
+\begin{lstlisting}
+class XKernelInvocationHook
+       extends XInvocationHookBase 
+       implements XInvocationHook {
+ ...
+ @Override
+ public XInvocationHookAction intercept(String functionName) {
+   switch (functionName) {
+     case "WRITE_ONCE": {
+       return new XInvocationHookAction((receiver, arguments) -> {
+         if (arguments.length != 2 || receiver != null) {
+             return null;  // do not intercept
+         }
+         XMemoryUnit destUnit = arguments[0];
+         if (!(destUnit instanceof XSharedLvalueMemoryUnit)) {
+           throw new XMethodInvocationError(methodName, 
+               "arg 1: not a shared l-value memory unit: " + destUnit);
+         }
+         XSharedLvalueMemoryUnit dest = (XSharedLvalueMemoryUnit) destUnit;
+         XMemoryUnit srcUnit = arguments[1];
+         if (!(srcUnit instanceof XLocalMemoryUnit)) {
+           throw new XMethodInvocationError(methodName, 
+               "arg 2: not a local memory unit: " + srcUnit);
+         }
+         XLocalMemoryUnit src = (XLocalMemoryUnit) srcUnit;
+         return program.emitMemoryEvent(dest, src);
+     });
+   }
+   case "READ_ONCE": {
+     ...
+   }
+ }
+}
+\end{lstlisting}
+\caption{Example of the invocation hook}
+\label{code-invoc-hook}
+\end{figure}
 

### 2018-05-25 1_intro.tex
Finish the summary. Cleanup eval. cleanup others a bit.

@@
 %
-Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static analysis of concurrent programs.
+Considering the enhancements of the architectural design, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the reachability and portability analysis, but serve as a basis for other kinds of static analysis of concurrent programs.
 

### 2018-05-25 4_impl.tex
Finish the summary. Cleanup eval. cleanup others a bit.

@@
 
-\begin{figure}
+Currently, \porthos[2] can operate only in the intra-procedural analysis mode, assuming that each function defined in the input file is being executed in a separate thread.
+%TODO: describe more intra- and inter-procedural!
+However, the redesigned architecture of \porthos[2] can be easily extended to support the inter-procedural (cross-procedure) analysis that inlines function calls and binds variable contexts.
+%TODO : check in the implementation , THAT WE HAVE NOT IMPLEMENTED THIS FEATURE!!
+Thus, instead of analysing a single source code file, the user can analyse the whole code project.
+In this mode, the functions to be executed in parallel should be specified by the user.
+Also, the tool may be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
+For that, the pre-compiler should recognise functions that commence a new process (such as \texttt{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
+This functionality %implies an additional stage of the pre-compilation and 
+is left beyond current thesis.
+%resolve the semantics of function invocations and remember the functions 
+%function that commence new process or thread 
+%(for instance, the function \texttt{pthread\_create} of the library \texttt{pthread.h} that starts the new thread of the function it received as the first argument).
+%TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs.
+
+%TODO: say sth about 100500 litmus-tests for kernel
+
+%TODO: ATTENTION HERE TO THE PICTURE!!!!!!!!!!!!!!!!!!!!!!!!
+\begin{figure}[H]
   \centering
   %draft=true
-  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
+  \includegraphics[height=.83\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
   \caption{The general architecture of \porthos[2]}
@@
 
-Currently, \porthos[2] can operate only in the intra-procedural analysis mode, assuming that each function defined in the input file is being executed in a separate thread.
-%TODO: describe more intra- and inter-procedural!
-However, the redesigned architecture of \porthos[2] can be easily extended to support the inter-procedural (cross-procedure) analysis that inlines function calls and binds variable contexts.
-%TODO : check in the implementation , THAT WE HAVE NOT IMPLEMENTED THIS FEATURE!!
-Thus, instead of analysing a single source code file, the user can analyse the whole code project.
-In this mode, the functions to be executed in parallel should be specified by the user.
-Also, the tool may be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
-For that, the pre-compiler should recognise functions that commence a new process (such as \texttt{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
-This functionality %implies an additional stage of the pre-compilation and 
-is left beyond current thesis.
-%resolve the semantics of function invocations and remember the functions 
-%function that commence new process or thread 
-%(for instance, the function \texttt{pthread\_create} of the library \texttt{pthread.h} that starts the new thread of the function it received as the first argument).
-%TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs.
-
-%TODO: say sth about 100500 litmus-tests for kernel
-
 \subsection{Internal representations}
@@
 \end{lstlisting}
-\caption{Example of the invocation hook}
+\caption{Example of the invocation hook for intercepting the Linux kernel-specific functions}
 \label{code-invoc-hook}

### 2018-05-25 5_eval.tex
Finish the summary. Cleanup eval. cleanup others a bit.

@@
 
-\section{Comparison to \porthos[1]}
+\section{Comparison with \porthos[1]}
 \label{ch:eval:show}
 
-\subsection{The compilation and unrolling}
+\subsection{Compilation and unrolling}
 \label{ch:eval:show:compil}
 
-As an example of the compilation and unrolling processes, consider the simple
-
+%As an example of the compilation and unrolling processes, consider the simple
 %As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, 
-Consider two tiny equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:both-cf}.
+Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
 %(in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
@@
 \caption{Example: A demonstrative cyclic function}
-\label{ex:both-cf}
+\label{ex:test-both-cf}
 \end{figure}
@@
 
-The Figure~\ref{ex:test-botrepresentedh-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-code} are compiled to.
+The Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-cf} are compiled to.
 The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].
@@
 %
-\caption{The control-flow graphs of the functions represented in Figure~\ref{ex:both-cf}}
+\caption{The control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
 \label{ex:test-both-pic}
@@
 %
-\caption{The unrolled control-flow graphs of the functions represented in Figure~\ref{ex:test-both-code}}
+\caption{The unrolled control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
 \label{ex:test-both-pic-unroll}

### 2018-05-25 99_summary.tex
Finish the summary. Cleanup eval. cleanup others a bit.

--- 2018-05-24 99_summary.tex
@@
 \item
-Then, we explored the existing implementation of the proof-of-concept tool \porthos[1] in order to understand components that need to be redesigned in order to facilitate the extension of the input language.
+Then, we explored existing implementation of the portability analysis tool \textit{\porthos{}} and designed the new tool \textit{\porthos[2]} that accepts higher subset of the C language as input, has modular architecture and disposes the extensible knowledge base of the domain-specific functions.
+The key aspects of the architectural design and implementations are the following:
+%explored the existing implementation of the proof-of-concept tool \porthos[1] redesigned the new architecture of \porthos[2] with consider
 
-  \begin{enumerate}[label=\roman*.,leftmargin=\parindent]
+  \begin{enumerate}[leftmargin=\parindent,label=\roman*.]
   \item The first stumbling block in the input language extension was the \porthos[1] input language parser, which performs the full semantic analysis of the program.
@@
 
-  \item Secondly, as \porthos[2] determines the \textit{kind (shared or local) of each variable} on the pre-compilation phase, the shared variables can be accessed in arbitrary code location (comparing to \porthos[1] which determines the type of a variable syntactically, depending on the operator of function that uses the variable).
+  \item Secondly, \porthos[1] determined the \textit{kind of a variable} (shared or local) syntactically, depending on the operator of the function that uses this variable.
+  This strongly restricts the syntax of the input language, therefore, for processing programs in C, \porthos[2] has an additional pre-compilation phase that traverses AST of the program and collects information of all variables accessed in the program (see Section~\ref{ch:impl:proc:x-pre-compiler}).
+  Currently, \porthos[2] considers a variable to be shared if it was declared as a pointer, or its address is accessed at least once by the code of any process, or it was declared as a parameter of the function that defines a process, or it was exported by \texttt{extern} or other keyword.
+  %This approach gives flexibility in detecting shared variables 
+  %  As this approach does not work for rich languages such as C   
+  %as \porthos[2] determines the \textit{kind (shared or local) of each variable} on the pre-compilation phase, the shared variables can be accessed in arbitrary code location (comparing to \porthos[1] which determines the type of a variable syntactically, depending on the operator of function that uses the variable).
 
-  \item In addition, \porthos[2] supports some \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\texttt{int a, b=2, z;}'), arbitrary expressions (corresponding to the C standard), and others (see Section~\ref{ch:impl:input}).
+  \item In addition, \porthos[2] supports some \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\lstinline{int a, b=2, z;}'), arbitrary expressions (corresponding to the C standard), and others (see Section~\ref{ch:impl:input}).
 
@@
 
-  \item For ease the process of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking mechanism}, which allows to modify the compilation process without changing the interpreter.
-  The invocation hooks serve as a knowledge base for the program domain, that are to be extended in future.
+  \item For ease of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking}, a flexible mechanism for intercepting the compilation process without changing the interpreter.
+  The invocation hooking mechanism serves as a knowledge base for the program domain, that is to be filled and extended in future.
+  Invocation hooks are defined in Java and thus are flexible, though their extension and modification requires some knowledge of the internals of the tool.
   We illustrate this mechanism with the basic support for Linux kernel litmus tests.
@@
   \item Since the new tool compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from the simple unrolling all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores all possible states that the program can reach within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
-  Although the new algorithm produces considerably more events than the old algorithm and thus the requires more time for the program domain encoding, we claim it to be complete.
+  Although the new algorithm produces considerably more events than the old one and thus takes more time for analysis,  \porthos[2] uses the new algorithm by default as it is complete (finds \textit{all} possible executions).
 
@@
 
-Thus, the new tool \porthos[2] with easily extensible knowledge base that allow to support new domain-specific functions can be considered as a generalised framework for SMT-based memory model-aware analysis.
+Thus, the new tool \porthos[2] with the extensible program domain knowledge base can be considered as a generalised framework for SMT-based memory model-aware analysis, which can not only perform the reachability and portability analysis, but serve as a basis for other kinds of static analysis of concurrent programs.
 
@@
 
-\section{Limitations and future work}
+\section{Limitations and directions for future work}
 
-- improve the encoding: an abstraction over Z
+Current implementation of \porthos[2] has the following limitations, that might possibly define the direction of the future work.
 
-- typing: const-size arrays + structures and so on
-  -only basic datatypes
+\begin{itemize}[leftmargin=\parindent]
+\item One of the major limitation of \porthos[2] as a software verification tool is its sensitivity to the \textit{combinatorial explosion} of the state space.
+%that performs complete analysis
+%TODO: show it in table , reference here the table
+As it was shown in Section~\ref{ch:eval:perf}, the number of events of the unrolled program grow rapidly as the user increases the unrolling bound.
+One possible way to reduce the number of the program states might be applying some traditional techniques that target the state explosion problem (such as concrete execution as a part of concolic execution~\cite{majumdar2007hybrid} before the unrolling stage).
+However, this must be done carefully and with taking into account the analysing weak memory model, because otherwise it can lead to the loss of states and thus to incorrect analysis results.
+Nonetheless, the small-size litmus test-like programs (containing hundreds of events after the unrolling) remain to be handleable by \porthos[2] within reasonable time.
+% cannot be applied to our problem \textit{before} the weak memory model is considered, 
+%One possible enhancement, that could treat but not cure this flaw, could be an extra analysis stage carried before the unrolling, that analyses
 
-- hooks : flexible.
-but still in Java, not as separate module (as it requires turing-complete language to be expressive enough) => extension of the knowledge base for supporting new platforms / languages
+%\item As its predecessor, \porthos[2] does not propose any heuristic for finding the optimal unrolling bound.
 
-- new modes: as it was discussed <before>, works only in the litmus test-mode. May work with complete projects. But here: 
-  - memory issues (we store same graph + back edges + memoised queries ==> multiple instances.);
-    - + will need the memory-guarding module
-  - ability to encode partially (once the
+\item \textit{Extending the knowledge base} of domain-specific functions may be considered as the main future research direction.
+Thus, to be able to process most Linux kernel litmus tests, \porthos[2] needs to know the semantics of the barrier and memory operation functions that are specific for the Linux kernel in order to be able to compile them into the \xgraph{} and later encode them into the SMT-formula.
+Due to modular architecture of \porthos[2], the extension of the knowledge base can be done by modifying solely the invocation hooking mechanism.
+Note that the flip side of the flexibility of invocation hooks is their complexity: being written in Java, in addition they require from the user some knowledge of general architecture of the tool, the interface methods provided by the interpreter, the class hierarchy of the \xgraph{} elements, etc.
 
--only one mode litmus-test intra proc
+\item Although \porthos[2] has an extended support for primitive data types (integers, reals, booleans), it still is not able to handle the \textit{complex data types}.
+However, as the Z3 solver supports the theory of constant-size arrays, it might be relatively easy to extend the support for constant-size arrays (declared as `\lstinline{int arr[10];}'), enumerations and structures.
+Nevertheless, pointers and variable-size arrays (allocated in the heap by functions \lstinline{malloc}, \lstinline{calloc}, etc.) require a stronger pre-processing analysis before being encoded into an SMT-formula.
 
-- testing with new solvers
+\item Currently, \porthos[2] can operate only in the intra-procedural analysis mode by assuming that each function provided by the user represents a separate thread (so-called litmus test-mode).
+One future direction of the work could be adding the \textit{inter-procedural analysis mode} for scanning a code project instead of a separate file (see the discussion in Section~\ref{ch:impl:principles}).
+Although, processing large programs may require serious memory optimisations (for instance, in storing the full unrolled graph data structure), which can possibly lead to worsening the overall performance.
+In addition, when processing large code projects, \porthos[2] might need the memory-guarding module that tracks the memory consumption of the tool and aborts its work if necessary.
+
+\item As it was mentioned in Section~\ref{ch:enc:bmc:cf} devoted to the encoding for the control-flow of the program, the new encoding scheme used by \porthos[2] allows to analyse partial functions.
+Although the current \porthos[2] interpreter infrastructure does is not configured to perform a partial analysis, it can be easily supported.
+
+\item During the implementation of \porthos[2], we intentionally skipped the time-consuming stage of writing the test suites for the tool.
+Thus, in order to increase the robustness and stability of \porthos[2], we need to cover its code by unit- and functional tests.
+
+\item One might want to compare the performance of \porthos[2] in cases of usage the different SMT-solvers.
+Due to the \zformula{} abstraction layer, new SMT-solvers may be easily added (the general way to support multiple SMT-solvers is to convert the \zformula{} to an SMT-LIB formula~\cite{smt-lib}, that can be passed as input to the SMT-solver run as an external process).
+
+\end{itemize}

### 2018-05-25 1_intro.tex
Consider Professor's comments. Cleanup and extend.

@@
 Although the parallelisation of a system can drastically improve its performance, it opens numerous of problems regarding correctness, robustness and reliability, which makes the concurrent program design one of the most difficult problems of programming~\cite{mckenney2017parallel}.
+In this work, we consider only \textit{Symmetric Multiprocessor (SMP)} parallelism (systems with multiple processors connected to a single shared memory), leaving aside the discussion on distributed concurrency (systems with autonomous nodes that communicate with each other by passing messages through the network).
+%Leaving aside questions on the performance improvement of the system parallelisation, we explore the questions of the \textit{memory consistency},
+We are discussing mostly the questions on the \textit{memory consistency} rather than performance benefits of the concurrency as the inconsistent memory operations can create new program behaviours, unexpected from the programmer's point of view (as well as the loss of expected behaviours), which all can be considered as a security flaw.
 
@@
 However, on the x86 architecture the state `\texttt{(0:EAX=0,\,1:EAX=0)}' is also reachable as the processors may cache the write to shared memory into their local write buffers, so that the write does not immediately become visible by processes running on other cores.
-This behaviour is known as \textit{store buffering~(SB)}.
+This behaviour is known as \textit{Store Buffering~(SB)}.
 
@@
 
-The formal way to define the semantics of memory operations and synchronisation primitives of a parallel execution environment is to define its \textit{memory model}.
+The formal way to define the semantics of memory operations and synchronisation primitives of a parallel \textit{execution environment} (hardware, programming language, compiler, database, operation system, etc.) is to define its \textit{memory model}.
 %The behaviour of a concurrent execution environment is described by a formal memory model
@@
 Models of the first type characterise the behaviour of the system (its \textit{operational semantics}) in terms of the abstract machine executing the code, as it was done for the SB example above.
-Models of the second type define the \textit{axiomatic semantics} of the system by specifying the set of assertions over states of the program.
+Models of the second type define the \textit{axiomatic semantics} of the system by specifying a set of assertions over states of the program.
 %There are two ways to define a memory model: by specifying its \textit{operational semantics} (in terms of an abstract machine executing the code, as it was done for the SB example above) or by defining its \textit{axiomatic semantics} (in terms of assertions over states of the program formed by sequences of operations).
@@
 
-The first memory model for a concurrent system was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
-This memory model, called the \textit{sequential consistency (SC)}, allows only those executions that produce the same result as if the operations had been executed in an interleaved fashion in a single process%
+Chronologically the first memory model for a concurrent system was formulated by Leslie Lamport back in 1979~\cite{lamport1979make}.
+This memory model, called the \textit{Sequential Consistency (SC)}, allows only those executions that produce the same result as if the operations had been executed in an interleaved fashion in a single process%
 %
 \footnote{In order to prescind from the implementation details while discussing the memory models theory, we avoid the use of the software-specific term \textit{thread} and the hardware-specific term \textit{processor}.
-Instead, we adhere the terminology of the theory of concurrency employed by Ben-Ari~\cite{ben2006principles} by naming a concurrent piece of code the \textit{process}.}.
-%
+Instead, we adhere the terminology of the theory of concurrency employed by Ben-Ari~\cite{ben2006principles} by naming a concurrent piece of code the \textit{process}.}. %
 %
@@
 The SC model is considered to be a \textit{strong memory model} in the sense that it provides firm guarantees regarding the ordering and effect of memory operations.
-Different relaxations of the SC model lead to \textit{weak memory models~(WMMs)} that specify how processes interact through the shared memory, when a write becomes visible to processes running on other cores, and what value a read operation can get.
-Thus, WMMs serve as set of guarantees made by designers of execution environment (hardware, programming language, compiler, database, operation system, etc.) to programmers on which behaviours of their concurrent code they can rely on.
+%However, execution environments that provide weakened guarantees (\textit{relaxations} such as memory operations reordering, write buffering, etc.)
+Weakened guarantees (such as memory operations reordering, write buffering, etc.) are called \textit{relaxations} of the memory model.
+The relaxations of the SC model lead to \textit{Weak Memory Models~(WMMs)} that specify how processes interact through the shared memory, when a write becomes visible to processes running on other cores, and what value a read operation can get.
+Thus, WMMs serve as set of guarantees made by designers of an execution environment to programmers on which behaviours of their concurrent code they can rely on.
 
@@
 
-Research on weak memory models firstly aims to \textit{formalise} a formal approach of understanding programs with respect to weak memory models, which is \textit{systematic}, \textit{sound} and \textit{complete}.
+Research on weak memory models firstly aims to \textit{formalise} an approach of understanding programs with respect to weak memory models, which is \textit{systematic}, \textit{sound} and \textit{complete}.
 One of the most well-known frameworks for weak memory model-aware analysis was formalised by J. Alglave in~2010~\cite{alglave2010shared}.
@@
 %, or Alpha architecture~\cite{alphaWMM}~, and others.
-There are projects for validating hardware architectures wrt. a memory model as well, e.g.~\cite{lustig2014pipecheck,lustig2016coatcheck}.
+There are projects for validating hardware architectures wrt. a memory model as well, e.g.,~\cite{lustig2014pipecheck,lustig2016coatcheck}.
 
@@
 
-Most modern high-level programming languages rely on relaxed memory model as well.
+Most modern high-level programming languages rely on relaxed memory models as well.
 Thus, the memory model for Java is based on the \textit{happens-before} principle~\cite{lamport1978time} and was introduced in J2SE 5.0 in 2004~\cite{manson2005java}.
-The transformations valid for the Java memory model were discussed in~\cite{sevcik2009program}. 
-A weak memory model for C and C++ is based on the relation \textit{strongly happens-before} and was introduced in the C++17 standard~\cite{batty2011mathematizing}.
-Before that, the C++11 standard~\cite{iso2012iec} proposed the set of hardware-independent synchronisation fences and atomic operations.
-Weak memory models are being formalised for even more abstract software environments, the notable project in this area is the project on formalising the Linux kernel memory model, which is being actively developing these days~\cite{alglave2018frightening, kernel1}.
+The transformations valid under the Java memory model are discussed in~\cite{sevcik2009program}.
+The weak memory model for C and C++ was defined the C++11 standard~\cite{c11}.
+The standard introduced a set of hardware-independent synchronisation fences and atomic operations, which were formalised in the work~\cite{batty2011mathematizing}.
+%This work influenced the C++11 standard~\cite{c11}, which introduced the set of hardware-independent synchronisation fences and atomic operations.
+The native support for the synchronisation primitives by the programming language, defined in the standard, has replaced the library-based approach for concurrency, therefore the compiler became aware of the concurrent parts of the code.
+%the concurrency libraries and language-independent standards that support concurrency (such as POSIX) that were used to write parallel programs in previous versions of C and C++.
+%is based on the \textit{strong happens-before} relation and was introduced in the C++17 standard~\cite{batty2011mathematizing}.
+Nonetheless, in 2015 some common compiler transformations were shown to be invalid under the C11 memory model~\cite{vafeiadis2015common}.
+%Therefore, the formal memory model of the programming language is an essential part of the compiler infrastructure.
+
+Weak memory models are being formalised for even more abstract software environments.
+The notable project in this area is the project on formalising the Linux kernel memory model, which is under active development nowadays~\cite{alglave2018frightening, kernel1, mckenney2017wg21}.
+This project has also an influence on the C language: the revision P0124R4 of the C standard~\cite{kernel2c} compares the Linux kernel and C11 memory models (including variable access, memory barriers, locking and atomic operations).
 
@@
 
-The very first tool that performs the WMM-aware portability analysis is \porthos%
+The very first tool that performs the WMM-aware portability analysis is \porthos{}%
 %
@@
 Chapter~\ref{ch:impl} delves into the description of architectural solutions and implementation details of the \porthos[2] framework.
-Chapter~\ref{ch:eval} provides an example of the portability analysis via \porthos[2] and presents some evaluation results comparing to other tools. %TODO: check if it's true
-Chapter~\ref{ch:summary} summarises results of the work and proposes future work directions. %TODO: check if it's true
+Chapter~\ref{ch:eval} gives a comparison of key features of \porthos[2] and \porthos[1] by providing examples of the compilation and the unrolling stages; and it provides some performance benchmarks of \porthos[2] in both reachability and portability analysis modes.
+Chapter~\ref{ch:summary} summarises results of the work and proposes possible directions of the future work.

### 2018-05-25 2_wmm.tex
Consider Professor's comments. Cleanup and extend.

@@
 
-The main idea behind the memory model-aware program analysis is that the set of all possible interleavings of the concurrent program (the \textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (the \textit{analytic semantics})~\cite{alglave2016syntax}.
+The main idea behind the memory model-aware program analysis is that the set of all possible executions of the concurrent program (the \textit{anarchic semantics}) can be specified by the axiomatic constraints of the memory model that filter out executions inconsistent in particular architecture (the \textit{analytic semantics})~\cite{alglave2016syntax}.
 The anarchic semantics of the program is a truly parallel semantics with no global time that describes all possible computations with all possible communications.
@@
 
-\section{The event-based program representation}
+\section{Event-based program representation}
 \label{ch:wmm:event}
@@
 \end{lstlisting}
-\caption{The simplified x86-TSO memory model defined in the \cat{} language~\cite{herd10tutorial}}
+\caption{The simplified x86-TSO memory model in the \cat{} language~\cite{herd10tutorial}}
 \label{ex:x86-cat}

### 2018-05-25 3_encod.tex
Consider Professor's comments. Cleanup and extend.

@@
 For instance, the specification assert may state that the property \textit{always} holds (the \textit{safety}) or the property will \textit{eventually} hold (the \textit{liveness}).
-Commonly, the \textit{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements.
+Commonly, the \textit{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements~\cite{clarke1999model}.
 
@@
 %Due to the SAT-solver.
-This technique is called \textit{bounded model checking (BMC)} as the search process is being repeated up to the user-defined bound $k$, which may result to incomplete analysis in general case.
+This technique is called \textit{Bounded Model Checking~(BMC)} as the search process is being repeated up to the user-defined bound $k$, which may result to incomplete analysis in general case.
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
@@
 \end{minipage}
-\caption{Example of encoding for the control-flow of the event-flow graph}
+\caption{Example of encoding for the control-flow of the \xgraph{}}
 \label{fig:graph-example}
@@
 \texttt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\
-\texttt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\texttt{tc}_{i}(e_1,e_2);\texttt{tc}_{i}(e_1,e_2))$.
+\texttt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\texttt{tc}_{i}(e_1,e_3);\texttt{tc}_{i}(e_3,e_2))$.
 \end{itemize}

### 2018-05-25 4_impl.tex
Consider Professor's comments. Cleanup and extend.

@@
 %The programming language choice for \porthos[2] was also made in favour of \textit{java},
-firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the object-oriented (OOP) concepts of Java suitable for modelling languages.
+firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the concepts of Object-Oriented Programming~(OOP) in Java suitable for modelling languages.
 Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT solver invoked by \porthos[2] via a Java API.
@@
 
-\section{The architecture}
+\section{Architecture}
 \label{ch:impl:arch}
@@
 
-\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
+\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{c11}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
 %
@@
 
-Following the C11 standard~\cite{iso2012iec}, we distinguish a \textit{statement} (\textit{"an~action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
+Following the C11 standard~\cite{c11}, we distinguish a \textit{statement} (\textit{"an~action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
 
@@
 %
-that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType
-As the type of an X-memory unit may not be resolved correctly, the memory units keep this information as a boolean flag.
+that is easily convertible to an SMT-type (modelled as \texttt{ZType}). %TODO: add the class ZType
+All X-memory units carry the boolean flag that indicates whether it has been resolved correctly by the X-memory manager.
 
@@
     \\
-    According to the Rice's theorem~\cite{rice1953classes}, in general case there always will be functions unresolved within an analysis launch.
+    %According to the Rice's theorem~\cite{rice1953classes}, in general case there always will be functions unresolved within an analysis launch.
     If the function has not been resolved, it can be safely assumed to be a no-operation function.

### 2018-05-25 99_summary.tex
Consider Professor's comments. Cleanup and extend.

@@
 
-\begin{enumerate}[label=\Roman*.]
+\begin{enumerate}[label=\arabic*.] %[label=\Roman*.]
 \item
@@
 
-  \begin{enumerate}[leftmargin=\parindent,label=\roman*.]
+  \begin{enumerate}[leftmargin=\parindent,label=\alph*.]
   \item The first stumbling block in the input language extension was the \porthos[1] input language parser, which performs the full semantic analysis of the program.
-  Although this works for a small toy language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
+  Although this works for a small subset of the C language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
   For handling that, we implemented the processing units that operate on the \textit{pre-compilation} stage discussed in Section~\ref{ch:impl:proc:x-pre-compiler}.
@@
 
-\item \textit{Extending the knowledge base} of domain-specific functions may be considered as the main future research direction.
+\item \textit{Extending the knowledge base} of domain-specific functions to model synchronisation primitives can be considered as the main future research direction.
 Thus, to be able to process most Linux kernel litmus tests, \porthos[2] needs to know the semantics of the barrier and memory operation functions that are specific for the Linux kernel in order to be able to compile them into the \xgraph{} and later encode them into the SMT-formula.
@@
 
+\item Secondly, the wide range of existing litmus tests may require \porthos[2] to \textit{extend the support for input languages}.
+The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, POWER, Alpha, etc.).
+Note that the programs in low-level languages can be easily compiled to the \xgraph{} representation by existing compiler architecture since low-level assembly-like languages can be considered as the subset of the C language, that restricts the set of non-linear control-flow instructions to the conditional and unconditional jumps (which are supported by the current \xgraph{} compiler of \porthos[2]).
+%Note that adding support for low-level languages requires adding the new \xgraph{} compiler that will compile the low-level syntax tree (with non-conditional jumps as the only )
+
 \item Although \porthos[2] has an extended support for primitive data types (integers, reals, booleans), it still is not able to handle the \textit{complex data types}.
-However, as the Z3 solver supports the theory of constant-size arrays, it might be relatively easy to extend the support for constant-size arrays (declared as `\lstinline{int arr[10];}'), enumerations and structures.
+However, as the \tool{Z3} solver supports the theory of constant-size arrays, it might be relatively easy to extend the support for constant-size arrays (declared as `\lstinline{int arr[10];}'), enumerations and structures.
 Nevertheless, pointers and variable-size arrays (allocated in the heap by functions \lstinline{malloc}, \lstinline{calloc}, etc.) require a stronger pre-processing analysis before being encoded into an SMT-formula.
@@
 
-\item As it was mentioned in Section~\ref{ch:enc:bmc:cf} devoted to the encoding for the control-flow of the program, the new encoding scheme used by \porthos[2] allows to analyse partial functions.
-Although the current \porthos[2] interpreter infrastructure does is not configured to perform a partial analysis, it can be easily supported.
+\item As it was mentioned in Section~\ref{ch:enc:bmc:cf} devoted to the encoding for the control-flow of the program, the new encoding scheme used by \porthos[2] allows to analyse \textit{partial functions}.
+Although the current \porthos[2] interpreter infrastructure is not configured to perform a partial analysis, it can be easily supported.
 
-\item During the implementation of \porthos[2], we intentionally skipped the time-consuming stage of writing the test suites for the tool.
-Thus, in order to increase the robustness and stability of \porthos[2], we need to cover its code by unit- and functional tests.
+\item During the implementation of \porthos[2], we have only done limited \textit{testing of the tool}.
+Thus, in order to increase the stability of \porthos[2], we need to cover its code by unit- and functional tests.
 

### 2018-05-25 inc/lst/XInterpreter.java

@@ -1,4 +1,5 @@
 public interface XInterpreter {
+
   XProcessId getProcessId();
   void finish();
   XCyclicProcess getResult();
@@ -34,23 +35,4 @@ public interface XInterpreter {
   void finishNonlinearBlockDefinition();
   void processJumpStatement(JumpKind kind);
   XEntity processMethodCall(String methodName, XMemoryUnit... arguments);
-  
-  
-  // --
-  
-  enum BlockKind {
-      Sequential,
-      Branching,
-      Loop;
-  }
-  
-  enum BranchKind {
-      Then,
-      Else;
-  }
-  
-  enum JumpKind {
-      Break,
-      Continue;
-  }  
-}  
+}

### 2018-05-26 1_intro.tex
Cleanup finale up to Chap 3.

@@
 %TODO: rename mousquetaires: http://www.cprover.org/wmm/musketeer/ !!! (+reference this project: "Don’t sit on the fence".pdf in alglave/ )
-% TODO: see how musketeer deals with recursive calls: http://www.cprover.org/wmm/musketeer/rec.pdf
 
@@
 
-As an example, consider a shared memory concurrent system that executes two parallel processes as it is described in Figure~\ref{simple_wmm_x86} (such little examples that illustrate specific behaviour of a concurrent execution environment are called \textit{litmus tests}).
-The process~\texttt{P0} writes the value~\texttt{1} to the shared variable~\texttt{x} and reads a value of the shared variable~\texttt{y}, and the process~\texttt{P1} writes the~\texttt{y} and reads the~\texttt{x}.
-Considering all interleavings of the processes, one could expect the final state to be one of the following:
-\begin{itemize}[noitemsep,topsep=0pt]
-\item \texttt{(0:EAX=0,\,1:EBX=1)},
-\item \texttt{(0:EAX=1,\,1:EBX=0)},
-\item \texttt{(0:EAX=1,\,1:EBX=1)}.
-\end{itemize}
-%TODO!!!!!!!!!!!!!!!!!!! cite "mcKenney - Memory Barriers a Hardware View for Software Hackers.pdf" -- a table for multiple architectures
-
-%As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
-However, on the x86 architecture the state `\texttt{(0:EAX=0,\,1:EAX=0)}' is also reachable as the processors may cache the write to shared memory into their local write buffers, so that the write does not immediately become visible by processes running on other cores.
-This behaviour is known as \textit{Store Buffering~(SB)}.
+As an example, consider two x86 assembly programs in Figure~\ref{simple_wmm_x86} (such little code examples that explain specific behaviour of the concurrent execution environment are called \textit{litmus tests}).
+Two programs are the same, except the right-hand program uses the synchronisation instruction \texttt{MFENCE} after the store to the shared memory.
+In both programs, the process~\texttt{P0} writes the value~\texttt{1} to the shared variable~\texttt{x} and reads a value of the shared variable~\texttt{y}, and the process~\texttt{P1} writes to the~\texttt{y} and reads the~\texttt{x}.
 
@@
 \ttfamily
+\begin{subfigure}[b]{.4\textwidth}
+\begin{tabular}[b]{ |l|l| }
+\hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \\ \hline
+\multicolumn{1}{|c|}{P0} & \multicolumn{1}{c|}{P1} \\ \hline
+MOV [x],1 & MOV [y],1 \\
+MOV EAX,[y] & MOV EBX,[x] \\
+& \\
+\hline
+\multicolumn{2}{|l|}{exists (0:EAX=0,\,1:EBX=0)} \tabularnewline
+\hline
+\multicolumn{2}{|l|}{x86-TSO: allowed} \tabularnewline
+\hline
+\end{tabular}
+\caption{Without synchronisation}
+\label{simple_wmm_x86:nosync}
+\end{subfigure}
+\hfill
+%
+\begin{subfigure}[b]{.4\textwidth}
 \begin{tabular}{ |l|l| }
 \hline
-\multicolumn{2}{|l|}{ \{ x=0; y=0; \}} \tabularnewline \hline
-P0 & P1 \\ \hline
+\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \\ \hline
+\multicolumn{1}{|c|}{P0} & \multicolumn{1}{c|}{P1} \\ \hline
 MOV [x],1 & MOV [y],1 \\
+MFENCE & MFENCE \\
 MOV EAX,[y] & MOV EBX,[x] \\
 \hline
-%\multicolumn{2}{|l|}{locations [x;y;]} \tabularnewline
-\multicolumn{2}{|l|}{exists (0:EAX=0,~1:EBX=0)} \tabularnewline
+\multicolumn{2}{|l|}{exists (0:EAX=0,\,1:EBX=0)} \tabularnewline
 \hline
-\multicolumn{2}{|l|}{x86: allow} \tabularnewline
+\multicolumn{2}{|l|}{x86-TSO: forbidden} \tabularnewline
 \hline
 \end{tabular}
-\caption{Store buffering: A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
+\caption{With synchronisation}
+\label{simple_wmm_x86:sync}
+\end{subfigure}
+%
+\caption{Store buffering~(SB): A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
 \label{simple_wmm_x86}
 \end{figure}
+
+Taking into account all possible interleavings of the read and write instructions, both litmus tests allow the following final states:
+\begin{itemize}[noitemsep,topsep=0pt]
+\item \texttt{(0:EAX=0,~1:EBX=1)},
+\item \texttt{(0:EAX=1,~1:EBX=0)},
+\item \texttt{(0:EAX=1,~1:EBX=1)}.
+\end{itemize}
+%As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
+However, in the left-hand (non-synchronised) litmus test, the state `\texttt{(0:EAX=0,~1:EAX=0)}' is also allowed on the x86 architecture as its processors may cache the writes to the shared memory into their local \textit{store buffers} so that the updated value does not immediately become visible by processes running on other cores.
+This behaviour is known as \textit{Store Buffering~(SB)}.
+For preventing such a behaviour, the x86 architecture offers the synchronisation instruction \texttt{MFENCE} that flushes the store buffer of the process.
+In Figure~\ref{simple_wmm_x86:sync}, this instruction is inserted for both processes after the store instructions, thus once it is executed in one process, the other process will read the updated value of this shared memory location.
+The work~\cite{mckenney2010memory} gives a comprehensive review of common hardware architectures from the perspective of the structure of their caches and write buffers, that determine the effect of memory operations, and memory barriers they provide.
 
@@
 %However, execution environments that provide weakened guarantees (\textit{relaxations} such as memory operations reordering, write buffering, etc.)
-Weakened guarantees (such as memory operations reordering, write buffering, etc.) are called \textit{relaxations} of the memory model.
+Weakening of the guarantees (such as memory operations reordering, write buffering, etc.) is called \textit{relaxation} of the memory model.
 The relaxations of the SC model lead to \textit{Weak Memory Models~(WMMs)} that specify how processes interact through the shared memory, when a write becomes visible to processes running on other cores, and what value a read operation can get.
-Thus, WMMs serve as set of guarantees made by designers of an execution environment to programmers on which behaviours of their concurrent code they can rely on.
+Thus, WMMs serve as a set of guarantees made by designers of an execution environment to programmers on which behaviours of their concurrent code they can rely on.
 
@@
 There are projects for validating hardware architectures wrt. a memory model as well, e.g.,~\cite{lustig2014pipecheck,lustig2016coatcheck}.
-
-%TODO: litmus test synthesis 
-%1) http://research.nvidia.com/sites/default/files/pubs/2017-04_Automated-Synthesis-of//ASPLOS_2017_Memory_Model_Verification.pdf
 
@@
 Furthermore, there exists a wide range of tools that perform memory model-aware analysis.
-\begin{itemize}[noitemsep,topsep=0pt]
+
+\begin{itemize}[noitemsep,topsep=0pt,leftmargin=\parindent]
 
@@
 
-\item There exist tools that perform the weak memory model-aware program verification and model checking.
+\item Some tools that perform the weak memory model-aware program verification and model checking.
 The notable examples are
 %the tool \tool{mmchecker} for model-checking the C\# code~\cite{huynh2006memory}
-the stateless model checkers \tool{RCMC}~\cite{kokologiannakis2017effective}, \tool{CHESS}~\cite{musuvathi2008fair} and \tool{Nidhugg}~\cite{abdulla2017stateless},
-the tool \tool{Trencher} for checking programs against the TSO memory model~\cite{bouajjani2013checking}, 
-the tool \porthos{} for portability analysis~\cite{Porthos17a},
+the stateless model checkers \tool{RCMC}~\cite{kokologiannakis2017effective}, \tool{CHESS}~\cite{musuvathi2008fair}, \tool{Nidhugg}~\cite{abdulla2017stateless}, \tool{Trencher}~\cite{bouajjani2013checking}, and others.
 
@@
 
+\item Along with synthesis of synchronisation primitives, some tools were designed to perform the automatic synthesis of litmus tests for the weak memory model, for example \tool{litmustestgen} by the NVidia Research~\cite{lustig2017automated}.
+
+\item The framework \tool{Alloy}~\cite{wickerson2017automatically} can be used for understanding WMMs; it can automatically generate conformance tests between two memory models, distinguish two WMMs, and check monotonicity and compiler mappings of a program.
+
 \item Some other tools perform static instrumentation of concurrent C programs and encode the WMM into the program representation so that it can be model-checked by standard tools.
-The examples are
-%TODO: cite the 'wmm-general/Thread-Modular Static Analysis for Relaxed Memory Models.pdf'
-the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
-the tool that performs the sequentialisation of concurrent programs~\cite{alglave2013software},
-the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.
+The examples are:
+\begin{itemize}[noitemsep,topsep=0pt]
+\item the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
+\item the tool that performs the sequentialisation of concurrent programs~\cite{alglave2013software},
+\item the tool \tool{Weak2SC} generates weak memory model descriptions of the program, which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.
+\end{itemize}
+
 \end{itemize}
@@
 
-All the tools listed above consider only a single memory model, however, in real life we face serious engineering problems involving necessity to model more than one execution environment.
-One of these problems is the \textit{portability} of the program from one hardware architecture to another.
-A program written in a high-level language is then compiled for different hardware.
-Even if all the compiler optimisations were disabled (which is rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
+All the tools listed above consider only a single memory model, however, in real life we face serious engineering problems involving more than one execution environment.
+One of these problems is the \textit{portability analysis} of the program from one hardware architecture to another.
+A program written in a high-level language is compiled for different hardware architectures.
+Even if all the compiler optimisations were disabled (which is a rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
 As the result, a program compiled under the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
@@
 %
-\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/Dat3M}} %
+\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/PORTHOS}} %
 %
-introduced in April~2017~\cite{Porthos17b}.
+introduced in April~2017~\cite{Porthos17a}.
 This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT solver.
@@
 As most modern SMT solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
-% TODO: search space size: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf slide
+% about SAT: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf
 

### 2018-05-26 2_wmm.tex
Cleanup finale up to Chap 3.

@@
 For instance, such an equivalence class can be the set of computations performed by a processor locally that do not affect the global state.
-This idea is used in the \textit{event-based} model, that represents the program as a directed graph of events (the \textit{event-flow graph})~\cite{alglave2010shared}.
+This idea is used in the \textit{event-based} model, that represents the program as a directed graph of events (the \textit{event-flow graph})~\cite{alglave2010shared,alglave2014herding}.
 The vertices of such a graph represent \textit{events} (see Section~\ref{ch:wmm:model:events}), and edges represent basic relations (see Section~\ref{ch:wmm:model:relations}).
@@
 
-%- The semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
-%"To build the aeg , we define a semantics of goto-programs in terms of abstract events, ..."
-% TODO: OR from here: http://www.cprover.org/etaps/esop.pdf "Software Verification for Weak Memory via Program Transformation"
-
 \subsection{Events}
@@
 
-%TODO: EVENT -- ATOMIC?
-An event is a fact of executing the low-level primitive operation such as memory access, threads synchronisation, computation, etc.
+An \textit{event} is a fact of executing the low-level primitive atomic operation such as memory access, threads synchronisation, computation over the local-memory, control-flow jump, etc.
 
@@
 
-The synchronisation instructions (fences) cause the \textit{barrier events}, that do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours.
-Functionally, a fence may be a synchronisation barrier or a instruction of flushing the local memory caches, etc.
-
-% TODO: more fences: "On Power and ARM (see Sec.
-6), we distinguish between control fence (which we write cfence), lightweight fence (lwfence) and full fence (ffence).
-On x86, implementing TSO, there is only one fence, called mfence." (herding cats, p.
-14)
+Synchronisation instructions (fences) cause \textit{barrier events}, which do not perform any computation or memory value transfer, instead, they the set of program behaviours by adding barrier relations to the program model.
+Functionally, a fence may be a synchronisation barrier or a instruction for flushing memory caches into the main memory, etc.
+For instance, the \texttt{mfence} instruction of the x86 assembly flushes the store buffers of the thread, and thus does not allow the \rf-relation (see Section~\ref{ch:wmm:model:relations}) to point to
 
@@
 The basic relations are the following~\cite{alglave2010shared}:
-%TODO: say that po and rf are  defined from the candidate execution:
-%TODO: say that 
 \begin{itemize}
@@
 
-%TODO: add ppo = characteristics of memory model.
+%TO-DO: add ppo = characteristics of memory model.
 "Architectures are instances of our model.
@@
 
-    %TODO: say that po-relation is *static*
+    %TO-DO: say that po-relation is *static*
     \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
@@
 
-%todo: some words here?
-
 \subsection{Executions}
@@
 The semantics of a concurrent program is represented as the set of allowed executions.
-The \textit{execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
+An \textit{execution} is a path in the event-flow graph defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
 It can be interpreted as a sequence of guesses which event is to be executed next.
-The \textit{candidate execution} is an execution that is not yet constrained by a memory model.
-
-%TODO: cite this paper elsewhere
-%As it was shown in~\cite{wickerson2017automatically}, it is enough for memory models to constrain the executions independently instead of constraining the program at a whole.
+A \textit{candidate execution} is an execution that is not yet constrained by a memory model.
 
 %An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
-%The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
-%associated with the instructions of the program
-%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
-%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\texttt{rf}$ and $\texttt{co}$~\cite{alglave2010shared}.
 Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \tool{herd7} tool, version~7.47).
 Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
-Since each thread performs single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).
+Since each thread performs a single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).
 
@@
 Weak memory models are defined via \cat{} language~\cite{alglave2016syntax}.
-This is a domain specific language for describing consistency properties of concurrent programs.
-The language combines expressive power of a functional language (it is inspired by OCaml and adopts its types, first-class functions, pattern matching and other features) with types, operations and assertions that are specific for operating with relations and executions.
-%represents the functional language extended by the theory of  allows to axiomatically define derived relations, architecture-specific fences, and constraints over relations.
-In \cat{}, new relations can be defined via the keyword \texttt{let} and the following operators over relations~\cite{alglave2016syntax}. % 3.1.6 Relations between events
+It is a domain specific language for describing consistency properties of concurrent programs.
+The language combines expressive power of a functional language (being inspired by OCaml, it adopts the OCaml types, first-class functions, pattern matching and some other features) with concepts of memory models (sets of events, relations, operations and assertions over relations).
+In \cat{}, new relations can be defined via the keyword \texttt{let} and the following operators over relations~\cite{alglave2016syntax}.
 
-%TODO: this is binary relations. ADD UNARY RELS!
-
-Below we enumerate pre-defined operators over relations and sets of events:
+Below we enumerate pre-defined operations over relations and sets of events:
 
@@
 For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$.
-As an example of memory model definition in \cat{} language, Figure~\ref{ex:x86-cat} presents the simplified x86-TSO model~\cite{herd10tutorial}.
-This memory model asserts  acyclicity of the communication relation (the union of \rf-, \fr- and \co- relations), \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.
+As an example of memory model definition in \cat{} language, Figure~\ref{ex:x86-cat} presents the excerpt from the x86-TSO memory model~\cite{herd10tutorial}.
+This memory model specification asserts  acyclicity of the communication relation (the union of \rf-, \fr- and \co- relations), \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.
 
-\begin{figure}[h]
+\begin{figure}[H]
 \begin{lstlisting}
-"X86 TSO"
-include "x86fences.cat"
-include "filters.cat"
-include "cos.cat"
+...
+let po_ghb = WW(po) | RM(po)
+let implied = PA(poWR) | WR(po)
+let GHB = mfence | implied | po_ghb | rfe | fr | co
+let com = rf | fr | co
 
-(* Uniproc check *)
-let com = rf | fr | co
+empty atom & (fre;coe)
 acyclic po-loc | com
-
-(* Atomic *)
-empty rmw & (fre;coe)
-
-(* Global happens-before *)
-#ppo
-let po_ghb = WW(po) | RM(po)
-
-#mfence
-include "x86fences.cat"
-
-#implied barriers
-let poWR = WR(po)
-let i1 = MA(poWR)
-let i2 = AM(poWR)
-let implied = i1 | i2
-
-let com = rfe | fr | co
-let ghb = mfence | implied | po_ghb | com
-show implied
-
-acyclic ghb as tso
+acyclic GHB
 \end{lstlisting}
-\caption{The simplified x86-TSO memory model in the \cat{} language~\cite{herd10tutorial}}
+\caption{Excerpt from the x86-TSO memory model in the \cat{} language}
 \label{ex:x86-cat}

### 2018-05-26 3_encod.tex
Cleanup finale up to Chap 3.

@@
-\chapter{Portability analysis as an SMT~problem}
+\chapter{Portability analysis as an SMT problem}
 \label{ch:enc}
 
-% TODO: Add intro to the bounded reachability analysis using SAT -- ?
-
-% TODO: other tools for bounded model-checking:
-
-%TODO: the method used by Porthos is complete and efficient, but the implementation sucks. Why? 1), 2), 3). How can it be fixed? 1), 2), 3).
-
-%TODO: more about SMT-(= SAT + constraints expressed in background theories. Exapmles of theories that we use and that are implem-ted in Z3).
-
-As it has been discussed in Chapter~\ref{ch:intro}, the program may behave differently when compiled for different parallel hardware architectures.
+As it was discussed in Chapter~\ref{ch:intro}, a concurrent program may behave differently when compiled for different hardware architectures.
 This may cause the portability bugs, the behaviour that is allowed under one architecture and forbidden under another. 
-%already said:
-%Although the research of weak memory models has achieved considerable success, most works remain to be rather theoretical that practical and serve mostly as tools for better understanding the concurrent nature of programs.
-% The first tool that tackles the problem 
-%The first work that has investigated the practical approach of modelling and verification the real-world programs with respect to weak memory models was ~\cite{Porthos17a}  
-In this Chapter, we describe the general task of analysing the concurrent software portability
-%may be stated 
-as a \textit{bounded reachability} problem, which in turn can be reduced to a SMT problem~\cite{Porthos17a}.
+This Chapter provides the short introduction into the general model checking and reachability analysis problem, after which it describes of the concurrent software portability analysis stated as a \textit{bounded reachability} problem, which in turn can be reduced to an SMT problem~\cite{Porthos17b}.
 
@@
 
-The model checking is the problem of verifying the system (the model) against a set of constraints (the specification)~\cite{dkw2008}.
-As the state machine model is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
+The \textit{model checking} is the problem of verifying the system (the \textit{model}) against a set of constraints (the \textit{specification})~\cite{dkw2008}.
+As the state machine is the most widespread mathematical model of computation, most classical model checking algorithms explore the state space of a system in order to find states that violate the specification.
 
 The general scheme of model checking is the following.
-The analysing system (the \textit{model}) is represented as a transition system, a directed graph with labelled nodes representing states of the system.
+The analysing system is represented as a transition system, a directed graph with labelled nodes representing states of the system.
 Each state corresponds to the unique subset of atomic propositions that characterise its behavioural properties.
-Once the model has been constructed, it can be checked for compliance to the \textit{specification}.
+Once the model has been constructed, it can be checked for compliance to the specification, a set of constraints.
 
-Usually, the specification defines temporal constraints over the properties of the system.
-For instance, the specification assert may state that the property \textit{always} holds (the \textit{safety}) or the property will \textit{eventually} hold (the \textit{liveness}).
+Usually, the specification defines temporal constraints over the system properties.
+For instance, the specification can assert that the property \textit{always} holds (the \textit{safety} property) or the property will \textit{eventually} hold (the \textit{liveness} property).
 Commonly, the \textit{Linear Temporal Logic~(LTL)} or \textit{Computational Tree Logic~(CTL)} (along with their extensions) is used as a specification language due to the expressiveness and verifiability of their statements~\cite{clarke1999model}.
@@
 However, all model checking techniques are exposed to the \textit{state explosion problem} as the size of the state space grows exponentially with respect to the number of state variables used by the system (its size).
-In case of modelling concurrent systems, this problem becomes much more considerable due to exponential number of possible interleavings of states.
-Therefore, the research in model checking over past 40 years was fighting the state explosion problem mostly by optimising search space, search strategy or basic data structures of existing algorithms.
+In case of modelling concurrent systems, this problem becomes much more serious due to the exponential number of possible interleavings of states.
+Therefore, over past 30 years the research in model checking has been fighting the state explosion problem mostly by optimising search space, search strategy or basic data structures of existing algorithms~\cite{clarke2001progress}.
 
-One of the first techniques that optimises the search space considerably is the symbolic model checking with \textit{Binary Decision Diagrams~(BDDs)}.
+One of the first techniques that optimises the search space considerably is the symbolic model checking with \textit{Binary Decision Diagrams~(BDDs)}~\cite{burch1992symbolic}.
 Instead of processing each state individually, in this approach the set of states is represented by the BDD, a data structure that allow to perform operations on large boolean formulas efficiently~\cite{clarke2012model}.
@@
 %Due to the SAT-solver.
-This technique is called \textit{Bounded Model Checking~(BMC)} as the search process is being repeated up to the user-defined bound $k$, which may result to incomplete analysis in general case.
+This technique is called \textit{Bounded Model Checking~(BMC)} as the search process is being repeated up to the user-defined bound $k$, which may result to \textit{incomplete analysis} in general case (not all the state space has been checked).
 However, there exist numerous techniques for making BMC complete for finite-state systems~(e.g.,~\cite{shtrichman2000tuning}).
-%As well as the BMC problem, the approach used by \porthos{}
 
-%For instance, the idea of grouping states with similar properties into equivalence classes lead to the concept of traces in concurrent systems proposed by A.~Mazurkiewicz in 1986~\cite{mazurkiewicz1986trace}.
+At present time, satisfiability solvers are behind the numerous verification and bounded model checking tools.
+Although, as most modern programming languages are typed, the SAT-formula that encodes the constraints of a program must include non-boolean constraints encoded as boolean variables, which is ineffective, therefore most software analysis and verification tools use \textit{Satisfiability Modulo Theories~(SMT)} -solvers, that integrate first-order reasoning with the background theory reasoning~\cite{kroening2008decision}.
+For example, the Integer theory includes operators `$+$', `$-$', `$\geqslant$' and others).
+Most modern SMT-solvers support integer arithmetic, real arithmetic, fixed-size bit-vectors, arrays, and other theories.
+The direction of the development of SMT-solvers is provided by the \texttt{SMT-LIB} standard~\cite{smt-lib}, which provides definition of the unified language for defining SMT-formulas, and description of most common underlying logics and theories.
+
+In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
+Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $\textit{Init}(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $\textit{Trans}(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $\textit{Bad}(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
+Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
+$\texttt{SAT}( \textit{Init}(\vec{x_0}) \land \textit{Trans}(\vec{x}_0, \vec{x}_1) \land ... \land \textit{Trans}(\vec{x}_{k-1}, \vec{x}_k) \land \textit{Bad}(\vec{x}_k) )$.
 
@@
 
-%TODO: overapproximation. loops, Rice's thm. Within the bound, the method is complete.
-
-In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
-Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $\textit{Init}(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $\textit{Trans}(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $\textit{Bad}(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
-Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
-%\begin{equation}
-$\texttt{SAT}( \textit{Init}(\vec{x_0}) \land \textit{Trans}(\vec{x}_0, \vec{x}_1) \land ... \land \textit{Trans}(\vec{x}_{k-1}, \vec{x}_k) \land \textit{Bad}(\vec{x}_k) )$.
-%\end{equation}
-
-The portability analysis problem may also be stated as a reachability problem, where the undesirable state is one reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
+A portability analysis problem may be stated as a reachability problem, where the undesirable state is one reachable under the target~$\mathcal{M_T}$ memory model and unreachable under the source memory model~$\mathcal{M_S}$.
 %However, unlikely a BMC problem, the portability analysis does not require to call the SMT solver repeatedly, since (imperative) programs may be converted as acyclic state graph (by reducing the loops, see Section~\ref{ch:impl:proc:x-transf}) and the $Trans$ predicate may be stated only for the final state of a program.
-Consider the function $\textit{cons}_{\mathcal{M}}(P)$ calculates the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
-Then, the program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17a}:
+Consider the function $\textit{cons}_{\mathcal{M}}(P)$ which calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$.
+The program $P$ is called portable from the source architecture (memory model) $\mathcal{M_S}$ to the target architecture $\mathcal{M_T}$ if all executions consistent under $\mathcal{M_T}$ are consistent under $\mathcal{M_S}$~\cite{Porthos17a}:
 
@@
 
-The control-flow of a program is represented by the \textit{control-flow graph}, a directed acyclic connected graph with a single source and multiple sink nodes. %TODO: footnote why multiple sinks
-%The edges of 
-%In a control-flow graph, there are two types of edges (transitions): \textit{primary transitions} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative transitions} that denote if-false-transitions (pictured with dotted lines).
-Each control-flow edge contains the label that denotes the \textit{guard}, a predicate determining the transition.
-The empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
-A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
+The control-flow of a program can be represented by the \textit{control-flow graph}, a directed acyclic connected graph with a single source and one or multiple sink nodes.
+Each control-flow edge contains the label that denotes the transition predicate called \textit{guard}.
+The empty guard (a \textit{true} predicate) is denoted as `$\epsilon$'.
+A guard represents a variable or a computational expression over local variables.
+As a guard depends on the data-flow of the program, it is liable to the weak memory model relaxations.
 The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
-However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition), see Section~\ref{ch:impl:model:xgraph:edges} for details.
-%Each node on graph can have either one successor (primary) or two successors (both primary and alternative); only computation events can serve as a branching point).
-%However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.
+However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a computation (the \textit{primary} and \textit{alternative} transitions), see Section~\ref{ch:impl:model:xgraph} for details.
 
@@
 The new encoding is to be smaller than the old one used in \porthos{} since it does not produce new variables for each high-level statement of the input language.
-%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner.
-This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.
 
-For instance, \porthos[1] used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
+For instance, \porthos[1] uses the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ is encoded recursively as %a separate boolean variable
 $\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
-and control-flow of the branching instruction $i_1 = (c\,?\,i_2\!:\!i_3)$ was encoded as
-$\phi_{CF}(c\,?\,i_2\!:\!i_3) \defeq (\textit{cf}_{i_1} \Leftrightarrow (\textit{cf}_{i_2} \lor \textit{cf}_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
-(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
-In contrast, the new encoding scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula.
-The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:bmc:df}).
+and the control-flow of the branching instruction $i_1 = (c\,?\,i_2\!:\!i_3)$ was encoded recursively as
+$\phi_{CF}(c\,?\,i_2\!:\!i_3) \defeq (\textit{cf}_{i_1} \Leftrightarrow (\textit{cf}_{i_2} \lor \textit{cf}_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
+%$\phi_{CF}(c\,?\,i_2\!:\!i_3)
+%(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
+In contrast, the new encoding scheme implemented in \porthos[2] firstly compiles the recursive high-level instructions into the linear low-level event-flow graph, which is then encoded into an SMT-formula.
+A guard in the event-flow graph is a value of conditional variable on the branching event, which is encoded as a data-flow constraint (see Section~\ref{ch:enc:bmc:df}).
 In general, the new encoding scheme follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} for encoding Petri-nets.
@@
     %
-    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
+    \caption{Possible mutual arrangements of events in an event-flow graph} %Linear and non-linear cases of a control-flow graph}
     \label{encode:cf}
@@
 In the result formula, all symbols $\fx(e_i)$ and $\fv(e_i)$ are encoded as boolean variables.
-Consider the possible mutual arrangements of nodes in a control-flow graph presented in Figure~\ref{encode:cf}.
+Consider the possible mutual arrangements of nodes in an event-flow graph presented in Figure~\ref{encode:cf}.
+For these cases, we propose the encoding scheme that uniquely encodes each node of graph and at the same time allows to encode partially executed program:
 
@@
 
-For these cases, we propose the encoding scheme that uniquely encodes each node of graph and at the same time allows to encode partially executed program.
-%TODO: say about epsilon on (c) picture
 Equation~\ref{enc:cf:seq} shows the encoding for the sequential control-flow represented in Figure~\ref{encode:cf:seq} and reflects the fact that the event $e_2$ can be executed iff the event $e_1$ has been executed.
@@
 Note that the sequential control-flow is a special case of branching with the only transition guard $\epsilon$ (that is encoded as \texttt{true}).
-%TODO: maybe add the overall formula \phi_{CF} ?
 
-For sake of correctness of the encoding, we require all branches to have at least one event.
-Thus, for branching statements that do not have any events in one of the branches (such a branch represents a conditional jump forward), we add the synthetic \textit{nop-event} as it is shown in Figure~\ref{encode:branching:nop}.
+For sake of correctness and simplicity of the encoding, we require all branches to have at least one event.
+Thus, for branching statements that do not have any events in one of the branches (such branch represents a conditional jump forward), we add a synthetic \textit{nop-event} as it is shown in Figure~\ref{encode:branching:nop}.
 
-\begin{figure}%[!h]
+\begin{figure}[t]%[!h]
     \centering
@@
 
-As an example of the control-flow encoding, consider the event-flow graph in Figure~\ref{fig:graph-example}, which has two branching points and one merge point with three incoming transitions.
-To illustrate the correctness of the encoding, consider the path $e_1 \rightarrow e_2 \rightarrow e_4$.
-This means, in the formula $\phi_{CF}$, the variables $\fx(e_1)$, $\fx(e_2)$ and $\fx(e_4)$ will have the value $1$, and the variable $\fx(e_3)$ will be assigned to $0$.
-It is easy to check that the $\phi_{CF}$ is satisfiable by the chosen SMT-model (considering the guards $g_{1,2}$ and $g_{2,4}$ that define this path to be evaluated to $1$).
-
-Note that the generalised encoding scheme does not require the branching transitions to be mutually-exclusive (for instance, consider two branching transitions from the event $e_2$, both labelled by non-epsilon guards $g_{2,3}$ and $g_{2,4}$).
-Next, consider the path $e_1 \rightarrow e_3 \rightarrow e_4$, which is not allowed by the control-flow graph.
-The corresponding model $\fx(e_1)=1$, $\fx(e_2)=0$, $\fx(e_3)=1$ and $\fx(e_4)=1$ does not satisfy the formula $\phi_{CF}$.
-The proposed encoding for the control-flow works also for encoding the partial graph.
-For example, the SMT-model $\fx(e_1)=1$, $\fx(e_2)=1$, $\fx(e_3)=0$ and $\fx(e_4)=0$ encodes the path $e_1 \rightarrow e_2$, satisfies $\phi_{CF}$.
-
-\begin{figure}
+\begin{figure}[t]
 \begin{minipage}{.4\textwidth}
@@
 
+%this paragraph should go after the Figure{fig:graph-example} so that the fig is on the top of the same page 
+As an example of the control-flow encoding, consider the event-flow graph in Figure~\ref{fig:graph-example}, which has two branching points and one merge point with three incoming transitions.
+To illustrate the correctness of the encoding, consider the path $e_1 \rightarrow e_2 \rightarrow e_4$.
+This means, in the formula $\phi_{CF}$, the SMT-variables $\fx(e_1)$, $\fx(e_2)$ and $\fx(e_4)$ should be assigned to  $1$, and the variable $\fx(e_3)$ should be assigned to $0$.
+%It is easy to check that the $\phi_{CF}$ is satisfiable by the chosen SMT-model (considering the guards $g_{1,2}$ and $g_{2,4}$ that define this path to be evaluated to $1$).
+
+Note that the generalised encoding scheme does not require the branching transitions to be mutually-exclusive (for instance, consider two branching transitions from the event $e_2$, both labelled by non-epsilon guards $g_{2,3}$ and $g_{2,4}$).
+Next, consider the path $e_1 \rightarrow e_3 \rightarrow e_4$, which is not allowed by the control-flow graph.
+The corresponding model $\fx(e_1)=1$, $\fx(e_2)=0$, $\fx(e_3)=1$ and $\fx(e_4)=1$ does not satisfy the formula $\phi_{CF}$.
+The proposed encoding for the control-flow works also for encoding the partial graph.
+For example, the assignment $\fx(e_1)=1$, $\fx(e_2)=1$, $\fx(e_3)=0$ and $\fx(e_4)=0$, which encodes the path $e_1 \rightarrow e_2$, satisfies $\phi_{CF}$.
+
 \subsection{Encoding for the data-flow}
@@
 The SSA form requires each variable to be assigned only once within entire program.
-In contrast, \porthos{} used the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique within a branch.
+In contrast, \porthos{} used the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique only within a control-flow branch.
 Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in the case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow of merge points~\cite{Porthos17a}.
@@
 (3)~no access to a constant variable or computed (evaluated) expression changes their SSA-index.
-These rules determine the following encoding of load, store and computation events within single thread:
-%TODO: maybe add the overall formula \phi_{DF} ?
+These rules determine the following encoding of load, store and computation events within a single thread:
 %
@@
 
-The time of described algorithm is linear of the size of event-flow graph since it performs only one graph traverse.
-%Notwithstanding the overhead of storing (or, equivalently, computing with %TODO: are you sure that 'with'?
-%linear time) the predecessors of each node in order to be able to convert the program into an SSA form, the time of such a transformation reduced %TODO: How much ????
-%comparing to the algorithm implemented in \porthos, which recomputed SSA-indices recursively for each instruction. %TODO: PROOF OR REMOVE !  Maybe, some words of copying, merging SSA maps ?
+The time of described algorithm is linear of the event-flow graph size as the algorithm performs only a single graph traverse.
 
-As it has been described before, the \rf-relation links data-flow between events %TODO: another word?
-of data-flow stored in equivalence assertions over the SSA-variables. 
+As it has been described before, the \rf-relation links data-flow between events.
+%of data-flow stored in equivalence assertions over the SSA-variables. 
 %The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
-The encoding of this linkage left untouched as it is implemented in \porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
+The encoding of this linkage has been left untouched as it is implemented in \porthos[1]: %for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
 %
@@
 
-where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\texttt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.
+where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\texttt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $\textit{true}$ if $e_1 \relarrow{rf} e_2$ (i.e., if $e_2$ reads the shared variable that was written in $e_1$).
 
-\subsection{Encoding for the memory model} %Encoding the memory model constraints}
+\subsection{Encoding for the memory model}
 \label{ch:enc:bmc:wmm}
 
-The basic scheme for encoding the memory model is proposed in~\cite{Porthos17a}.
-The encoding consists of two parts: encoding the \textit{derived relations} and encoding the memory model \textit{assertions}.
+The basic scheme for encoding the memory model was proposed in~\cite{Porthos17a}.
+The encoding consists of two parts: encoding of the \textit{derived relations} and encoding of the memory model \textit{assertions}.
 

### 2018-05-26 4_impl.tex
Cleanup finale up to Chap 3.

@@
 
-%TODO: ATTENTION HERE TO THE PICTURE!!!!!!!!!!!!!!!!!!!!!!!!
-\begin{figure}[H]
-  \centering
-  %draft=true
-  \includegraphics[height=.83\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
-  \caption{The general architecture of \porthos[2]}
-  \label{fig:arch}
-\end{figure}
-
-\section{Architecture}
-\label{ch:impl:arch}
-
-The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
-%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
-
-The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
-However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
-The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
-%
-\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
-For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
-%
-(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
-The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
-The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
-Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT solver (Section~\ref{ch:impl:proc:smt-solver}).
-
-\subsection{Program input}
+\section{Program input}
 \label{ch:impl:input}
@@
 
+\section{Architecture}
+\label{ch:impl:arch}
+
+The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
+%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
+
+%TODO: ATTENTION HERE TO THE PICTURE!!!!!!!!!!!!!!!!!!!!!!!!
+\begin{figure}[H]
+  \centering
+  \includegraphics[height=.95\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png}
+  \caption{The general architecture of \porthos[2]}
+  \label{fig:arch}
+\end{figure}
+
+The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
+However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
+The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
+%
+\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
+For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
+%
+(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
+The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
+The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
+Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
+%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT solver (Section~\ref{ch:impl:proc:smt-solver}).
+
 \subsection{Internal representations}
@@
 The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
-Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph:edges}).
+Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph}).
 
@@
 
-The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
+The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph}).
 The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
@@
 \begin{lstlisting}
-class XKernelInvocationHook
-       extends XInvocationHookBase 
+class XKernelInvocationHook extends XInvocationHookBase 
        implements XInvocationHook {
@@
          }
+
          XMemoryUnit destUnit = arguments[0];
@@
          }
-         XSharedLvalueMemoryUnit dest = (XSharedLvalueMemoryUnit) destUnit;
+         XSharedLvalueMemoryUnit dest = 
+               (XSharedLvalueMemoryUnit) destUnit;
+
          XMemoryUnit srcUnit = arguments[1];
@@
          XLocalMemoryUnit src = (XLocalMemoryUnit) srcUnit;
+
          return program.emitMemoryEvent(dest, src);
@@
 
+%TODO: say that program is the set of processes , but the implements the same interface !!! it's not clear here.
+
+%TODO: !!! Say that interpreter does not know anyth about Y, all its interface methods are <sth> -> <X>. The only part that know about Y is Y2Xvisitor
+
 %In this paragraph we discuss only the implementation of the process interpreter \texttt{XProcessInterpreter}.
 %This interpreter is used to build the branched event-flow graph with all kinds of jumps, memory operations and fences.
-The X-program interpreter is invoked by the \texttt{Y2XConverterVisitor} which walks down the recursive syntax tree \ytree{}.
+The X-program interpreter is invoked by the \texttt{Y2XConverterVisitor} which walks down the \ytree{} recursively.
 The calls to the X-program interpreter are dispatched to currently maintained X-process interpreter.
@@
 `\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
-If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph:evt}).
+If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph}).
 The second group of X-interpreter methods consists of the methods for defining non-linear statements (branchings and loops).
@@
 
-As a high-level instruction (Y-level) may be compiled into a sequence of low-level instructions (for example, a computation that involves shared variables should firstly load them into the local memory and then process the computation event over local-only memory), the interpreter must maintain the \textit{stack of contexts} and remember the \textit{previous event} to be able to correctly process nested non-linear statements of C language.
+As a high-level (Y-level) instruction may be compiled into a sequence of low-level instructions (for example, a computation that involves shared variables should firstly load them into the local memory and then process the computation event over local-only memory), the interpreter must maintain the \textit{stack of contexts} and remember the \textit{previous event} to be able to correctly process nested non-linear statements of C language.
 The context is a data structure that carries the \textit{state} and some additional information in the case of processing non-linear statements (e.g., conditional event, first and last then- and else-branch events for binding, etc.).
@@
 The context state is an enumeration of the following values:
-\begin{itemize}
+
+\begin{itemize}[noitemsep,topsep=0pt]
 \item \texttt{WaitingAdditionalCommand}: the interpreter is in the state of defining the complex (non-linear) statement and is not able to process any new event (will throw an exception if any) until the state is not changed;
@@
 \item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event);
-\item \texttt{WaitingNextLinearEvent}: the standard interpreter state for processing next linear event, and
+\item \texttt{WaitingNextLinearEvent}: the default interpreter state for processing next linear event, and
 \item \texttt{Idle}: the interpreter does not set up the edge from the previous event to the new event.
@@
 
+As an example of usage the interpreter, consider the code of the \texttt{Y2XConverterVisitor} for processing \texttt{if-then-else} statement of the \ytree{} in Figure~\ref{TODO}.
+
+\begin{figure}
+%TODO!!! refactor the code, move program.tryEvaluateComputation() to Y2Xvisitor or Helper
+\begin{lstlisting}
+public class Ytree2XgraphConverterVisitor implements YtreeVisitor<XEntity> {
+  ...
+
+  @Override
+  public XEvent visit(YBranchingStatement node) {
+    program.startBlockDefinition(XInterpreter.BlockKind.Branching);
+
+    program.startBlockConditionDefinition();
+    XComputationEvent condition = program.tryEvaluateComputation(node.getCondition().accept(this));
+    program.finishBlockConditionDefinition(condition);
+
+    program.startBlockBranchDefinition(XInterpreter.BranchKind.Then);
+    node.getThenBranch().accept(this);
+    program.finishBlockBranchDefinition();
+
+    YStatement elseBranch = node.getElseBranch();
+    program.startBlockBranchDefinition(XInterpreter.BranchKind.Else);
+    if (elseBranch != null) {
+      elseBranch.accept(this);
+    }
+    else {
+      program.emitNopEvent(); // needed for encoding
+    }
+    program.finishBlockBranchDefinition();
+
+    program.finishNonlinearBlockDefinition();
+
+    return null; //statements return null
+    }
+}
+\end{lstlisting}
+\end{figure}
+
 %The \texttt{almostReadyContexts} queue is necessary for not adding erroneous edges on
 
-For the sake of simplicity of the interpreter, jump events (no-computation events) are present in the flow-graph, however they can be removed from it with rebinding ingoing and outgoing edges.
+For the sake of simplicity of the interpretation and encoding, jump events (no-computation events) are present in the flow-graph, however they can be removed from it with rebinding ingoing and outgoing edges.
 
@@
 The meaning of a bound was changed in order to increase predictability of the size of the result SMT-formula.
-Not that recursive function calls create unconditional jumps back in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
+Note that recursive function calls create unconditional back-jumps in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
+The discussion on how to set up the \po-relations in the event-flow graph with recursive function calls can be found in~\cite{musketeer2014recursion}.
 
@@
 Once the graph is unrolled, it can be augmented by data-flow edges.
-As it was discussed in Paragraph~\ref{ch:impl:model:xgraph:edges}, the data-flow edges can represent either \rf{}- or \co{}-relation.
+As it was discussed in Section~\ref{ch:impl:model:xgraph}, the data-flow edges can represent either \rf{}- or \co{}-relation.
 The \rf{}-edges join each write event to all read events that access the same location; the \co{}-edges join write events to the same location.

### 2018-05-26 inc/XEventVisitor-cleaned.java

@@ -1,28 +1,0 @@
-public interface XEventVisitor<T> {
-
-    T visit(XEntryEvent event);
-
-    T visit(XExitEvent event);
-
-    T visit(XUnaryComputationEvent event);
-
-    T visit(XBinaryComputationEvent event);
-
-    T visit(XAssertionEvent event);
-
-    T visit(XInitialWriteEvent event);
-
-    T visit(XRegisterMemoryEvent event);
-
-    T visit(XStoreMemoryEvent event);
-
-    T visit(XLoadMemoryEvent event);
-
-    T visit(XMethodCallEvent event);
-    
-    T visit(XJumpEvent event);
-
-    T visit(XNopEvent event);
-
-    T visit(XBarrierEvent event);
-}

### 2018-05-26 inc/lst/XInterpreter.java

@@ -2,9 +2,9 @@ public interface XInterpreter {
 
   XProcessId getProcessId();
   void finish();
-  XCyclicProcess getResult();
+  XProcess getResult();
   
-  // linear interpretation methods:
+  // emitting new events:
   XEntryEvent emitEntryEvent();
   XExitEvent emitExitEvent();
   XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
@@ -26,7 +26,7 @@ public interface XInterpreter {
                                            XLocalMemoryUnit operand1, 
                                            XLocalMemoryUnit operand2);
 
-  // non-linear interpretation methods:
+  // non-linear statements interpretation:
   void startBlockDefinition(BlockKind blockKind);
   void startBlockConditionDefinition();
   void finishBlockConditionDefinition(XComputationEvent condition);
@@ -35,4 +35,5 @@ public interface XInterpreter {
   void finishNonlinearBlockDefinition();
   void processJumpStatement(JumpKind kind);
   XEntity processMethodCall(String methodName, XMemoryUnit... arguments);
+
 }

### 2018-05-26 inc/lst/sample-code.c

@@ -1,16 +1,0 @@
-void P0(int& x, int& y) {
-  int a = 0;
-  // for loop
-  while (x++ > y) {
-    y = y - ++a;
-    // call foo
-    if (a > y) {
-      break;
-    }
-  }
-}
-
-
-int foo(int a, int b) {
- 
-}

### 2018-05-27 1_intro.tex
Cleanup finale up to Xevent.

@@
 %
-\caption{Store buffering~(SB): A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
+\caption{Store buffering~(SB): A litmus test illustrating the write-read reordering allowed by the \texttt{x86-TSO} memory model}
 \label{simple_wmm_x86}
@@
 One of the most well-known frameworks for weak memory model-aware analysis was formalised by J. Alglave in~2010~\cite{alglave2010shared}.
-It is the event-based non-deterministic model without global time (see Section~\ref{ch:wmm:event} for details).
+It is the event-based non-deterministic model without global time (see Section~\ref{ch:wmm:model} for details).
 
 In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
-Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
+Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as \texttt{x86-TSO} and \texttt{Sparc-TSO} (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
 %, or Alpha architecture~\cite{alphaWMM}~, and others.
@@
 introduced in April~2017~\cite{Porthos17a}.
-This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT solver.
+This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver (see Section~\ref{ch:enc:bmc}).
 This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
-As most modern SMT solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
+As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
 % about SAT: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf

### 2018-05-27 2_wmm.tex
Cleanup finale up to Xevent.

@@
 \section{Event-based program representation}
-\label{ch:wmm:event}
+\label{ch:wmm:model}
 
@@
 Thus, there are only four possible executions defined by the choice of \rf-relation.
-The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models x86-TSO, Power, ARM, and some others.
+The candidate executions pictured in Figures~\ref{simple_wmm_x86_pic:sub1}--\ref{simple_wmm_x86_pic:sub3} are consistent both under strong memory model SC and under relaxed memory models \texttt{x86-TSO}, \texttt{POWER}, \texttt{ARM}, and some others.
 However, the execution shown in Figure~\ref{simple_wmm_x86_pic:sub3} is still consistent under relaxed-memory architectures, but it becomes inconsistent under SC architecture as it forbids cycles over~$\fr\cup\po$.
@@
 For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$.
-As an example of memory model definition in \cat{} language, Figure~\ref{ex:x86-cat} presents the excerpt from the x86-TSO memory model~\cite{herd10tutorial}.
+As an example of memory model definition in \cat{} language, Figure~\ref{ex:x86-cat} presents the excerpt from the \texttt{x86-TSO} memory model~\cite{herd10tutorial}.
 This memory model specification asserts  acyclicity of the communication relation (the union of \rf-, \fr- and \co- relations), \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.
@@
 \end{lstlisting}
-\caption{Excerpt from the x86-TSO memory model in the \cat{} language}
+\caption{Excerpt from the \texttt{x86-TSO} memory model in the \cat{} language}
 \label{ex:x86-cat}

### 2018-05-27 4_impl.tex
Cleanup finale up to Xevent.

@@
 ), we consider the accomplished work as a step towards it.
-%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
-%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
-% Although and optimisation of the tool so that it is able to process real C programs in perspective.
-
-%In current Chapter we discuss the new architecture of \porthos[2].
-
-% function invocations/calls: inlining with binding
-% contexts: for now, no ctxs
-% comparing to cprover (https://www.cprover.org/cbmc/doc/manual.pdf, page 45, absatz about 'goto') where "the loop is unwound a given number of times" , we SHOULD have 2 modes of unrolling: 1) number of times, and 2) number of high-level instructions executed
-% TODO(code): (https://www.cprover.org/cbmc/doc/manual.pdf, page 45) "The break and continue statements are replaced by equivalent goto statements as described in the ANSI-C standard."
-
-% TODO: arrays: cprover manual page 48
-
-%\section{The architecture of \porthos{}}
+
 \section{General principles}
@@
 
-The previous version of \porthos{} did not distinguish the event-based program model from the high-level AST, they both were encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
-%For 
-Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
-We are inclined to consider this architecture as one that is fast to develop, but hard to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
-%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).
-
-Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (an AST structure) from the low-level event-based representation (an event-flow graph).
+The \porthos[1] does not distinguish the event-based program model from the high-level AST, and they both are encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
+Moreover, the syntax tree is implemented in \porthos[1] as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos[1] that recursively compute some properties of the AST and change its state).
+We are inclined to consider the old architecture to be fast to develop, but difficult to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
+
+Therefore, while working on the new design of \porthos[2], we clearly separated the high-level intermediate code representation (the AST structure) from the low-level event-based representation (the event-flow graph).
 Such a modular architecture will allow to support multiple input languages%
 %
-\footnote{Apart from the C language, \porthos[2] must be able to analyse litmus tests in different assembly languages and, as a compatibility mode, the input language of \porthos[1].} %
+\footnote{Apart from the C language, \porthos[2] should be able to analyse litmus tests in different assembly languages and, as a compatibility mode, the input language of \porthos[1].} %
 %
@@
 
-All internal representations used by \porthos[2] must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
-The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
+All data-transfer objects~(DTOs) used by \porthos[2] must be immutable, so that it is possible to guarantee the correctness of the program by controlling preservation of its invariants.
+The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another user-defined immutable object, or an immutable collection provided by the library Guava by Google%
 %
@@
     \begin{enumerate}[label*=\arabic*.]
-      \item preservation the completeness of analysis,
+      \item completeness of the analysis,
       \item modular architecture: each module can be tested independently,
@@
 It also implies the unified ordering of methods in classes, minimised size of methods code, clear tabulation, etc.
-%The modular architecture of \porthos[2] allow each its module to be tested independently.
-%Usage of immutable data structures for all DTOs helps 
-%immut: preservation of invariants
-%The transparency, efficiency and extensibility are the necessary criteria of any good tool.
-
-As its predecessor, \porthos[2] uses the open-source SMT solver \tool{Z3}%
+
+As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
 %
@@
 by Microsoft Research~\cite{de2008z3}.
-However, unlike its predecessor, \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT solver.
-
-%We have decided to implement \porthos[2] in \textit{Java}
-As its predecessor, \porthos[2] is written in \textit{Java},
-%The programming language choice for \porthos[2] was also made in favour of \textit{java},
-firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the concepts of Object-Oriented Programming~(OOP) in Java suitable for modelling languages.
-Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT solver invoked by \porthos[2] via a Java API.
+However, unlike its predecessor, \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT-solver.
+
+As its predecessor, \porthos[2] is written in \textit{Java}, firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the concepts of Object-Oriented Programming~(OOP) inherent in Java suitable for modelling the programming languages.
+Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver invoked by \porthos[2] via a Java API.
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
-It is worth noting that, for the reasons of simplicity, \porthos[2] is not a concurrent program, however, we believe that, due to its modular architecture, it can be easily parallelised on the level of program modules.
-
-Currently, \porthos[2] can operate only in the intra-procedural analysis mode, assuming that each function defined in the input file is being executed in a separate thread.
-%TODO: describe more intra- and inter-procedural!
-However, the redesigned architecture of \porthos[2] can be easily extended to support the inter-procedural (cross-procedure) analysis that inlines function calls and binds variable contexts.
-%TODO : check in the implementation , THAT WE HAVE NOT IMPLEMENTED THIS FEATURE!!
-Thus, instead of analysing a single source code file, the user can analyse the whole code project.
-In this mode, the functions to be executed in parallel should be specified by the user.
-Also, the tool may be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
+It is worth noting that, for the reasons of simplicity, \porthos[2] is not a concurrent program, however, we believe that due to its modular architecture it can be easily parallelised on the level of program modules.
+
+Currently, \porthos[2] can operate only in the intra-procedural analysis mode (not supporting invocations of the user-defined functions), assuming that each function defined in the input file is being executed in a separate thread.
+However, the redesigned architecture of \porthos[2] can be easily extended to support the inter-procedural (cross-function) analysis by inlining user-defined functions calls and binding variable contexts.
+In this mode, instead of analysing a single source code file, the tool can process the whole code project, where the functions to be executed in parallel are specified separately by the user.
+Also, the tool can be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
 For that, the pre-compiler should recognise functions that commence a new process (such as \texttt{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
-This functionality %implies an additional stage of the pre-compilation and 
-is left beyond current thesis.
-%resolve the semantics of function invocations and remember the functions 
-%function that commence new process or thread 
-%(for instance, the function \texttt{pthread\_create} of the library \texttt{pthread.h} that starts the new thread of the function it received as the first argument).
-%TODO: inter: user specifies the repository with code, functions that intended to work in parallel and runs.
-
-%TODO: say sth about 100500 litmus-tests for kernel
+This functionality is left beyond current thesis.
 
@@
 \label{ch:impl:input}
-%(for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).
-
-Both \porthos{} and \porthos[2] use the \tool{ANTLR} parser generator%
+
+Both \porthos[1] and \porthos[2] use the \tool{ANTLR} parser generator%
 %
@@
 %
-~\cite{parr2013definitive}, a powerful language processing tool.
+~\cite{parr2013definitive} for creating the input-language parsers.
 The \tool{ANTLR} takes as input the user-defined grammar of the target language in a BNF-like form and produces the LL(*)-parser and optionally some auxiliary classes (such as listeners and visitors for the syntax tree).
@@
 Among other advantages \tool{ANTLR}, it has a rather large collection of officially supported grammars.
-Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the \tool{ANTLR} an attractive instrument for solving the parsing problem.
+Nonetheless, the intuitive syntax for defining grammars and numerous of tools for debugging grammars make the \tool{ANTLR} an attractive instrument for processing languages.
 
@@
 \end{lstlisting}
-\caption{The sketch of the input language grammar used by \oldporthos}
+\caption{Sketch of the input language grammar used by \oldporthos{}}
 \label{fig:in_grammar_pts}
@@
 Figure~\ref{fig:in_grammar_pts} represents the grammar sketch in BNF syntax of the input language used by \porthos[1].
-%(the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}).
-%TODO! Reformulate this.
-The input language parser used by \porthos{} suffered from several disadvantages.
-Firstly, it contained the parser code inlined directly into the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions).
+The input language parser used by \oldporthos{} suffers from several disadvantages.
+Firstly, it contains the parser code inlined directly into the grammar, so that the grammar serves as a template for the parser code (which is called semantic actions in \tool{ANTLR}).
 Such a combining of two expressive languages makes the code hardly understandable and, therefore, poorly maintainable.
@@
 
-Secondly, \porthos{} resolved the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level%
-%
+Secondly, \porthos{} resolves the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level%
+%
+%TODO: check or remove this
 %\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
@@
 %\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
-As the reader may have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, the memory operations of different kinds vary syntactically as well.
-For example, the assignment of local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation denoted as `\lstinline{<:-}', atomic non-relaxed store operation denoted as `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} are resolved syntactically by matching the function name.
-Moreover, only the operator `\lstinline{<-}' could have an expression as the source of data, which means that expressions could be assigned only to registers.
-% (thus, the variable \lstinline{r} was recognised as a local variable as well).
+As the reader might have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, different kinds of memory operations of of the \oldporthos{} input language vary syntactically as well.
+For example, the assignment of a local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation is denoted by \lstinline{<:-}', atomic non-relaxed store operation is denoted by `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} is resolved syntactically by matching the function name and arguments.
+Moreover, only the operator `\lstinline{<-}' accepts an expression as the source of data, which means that expressions could be assigned only to registers.
 In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
-The semantics of the functions also is being resolved during the pre-compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler}).
-
-Thirdly, the grammar used by \porthos{} allowed only a restricted set of operations.
-For example, it accepted the computation expressions only over local variables.
-Thus, in the assignment expression `\lstinline{r <- (x + 1);}', the variable \lstinline{x} was parsed as a local variable even though it can be used as a shared variable in other parts of the program.
+The semantics of the functions also being resolved during the compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler}).
+
+Thirdly, the grammar used by \oldporthos{} accepts only a restricted set of operations.
+For example, the computation expressions can use only local variables.
+Thus, in the assignment expression `\lstinline{r <- (x + 1);}', the variable \lstinline{x} will be parsed as a local variable even though it can be used as a shared variable in other parts of the program, which may lead to the inconsistency of analysis.
 In \porthos[2], all shared variables involved into a computation expression are tentatively copied to temporary local variables.
-%Moreover, expressions could be assigned only to local variables only via the assignment operator `\lstinline{<-}' (thus, the variable \lstinline{r} was recognised as a local variable as well).
-%, which might lead to inconsistency of the result SMT-formula if the same variable name was used both as a register and as a location, for instance, the in code `\lstinline{x := reg1; reg2 <- (x + 1);}', the first statement interprets the variable \lstinline{x} as a location, while the grammar of second statement requires it to be a register.
-
-Fourthly, \porthos[1] supported only integer constants and expressions.
-%Also, only integers were processed by the input language parser of \porthos{}.
-In \porthos[2], we extended support for primitive types supported by the \tool{Z3} solver (this apply to 32-bit integers encoded as \texttt{Int}s of \tool{Z3}, floats encoded as \texttt{Real}s, enumerations encoded as \texttt{Scalar}s).
-%TODO : constant arrays? https://rise4fun.com/\tool{Z3}/tutorial/guide
+
+Fourthly, \porthos[1] supports only integer constants and expressions.
+In \porthos[2], we extended support for primitive types supported by the \tool{Z3} solver (32-bit integers are encoded as \texttt{Int}s of \tool{Z3}, floats are encoded as \texttt{Real}s).
 Although the \tool{Z3} supports the array theory (characterised by the select-store axioms~\cite{de2011z3}), the complexity of pointer analysis for the arrays of non-constant size moves the full support of arrays and pointers out of the scope of current thesis.
-%Thus, adding support for more advanced structures as pointers, arrays, function definitions and calls was one of the purposes of revising the tool architecture.
-
-Finally, the grammar used by \porthos{} had the following minor drawbacks. %A minor drawback of 
-The operators were implemented as non-associative (expressions of the form `\lstinline{1 + 2 * 3}' could not be parsed).
-Comparing to C statements, that must end with the semicolon punctuator `\lstinline{;}', statements defined in the grammar of \porthos[1] use the semicolon as a separator between statements (the final statement must not end with semicolon).
-The litmus-specific syntax for variables initialisation was used only for declaring the shared variables (all of them were initialised with default value \lstinline{0}), however, this syntax should be used as an initial assignment of both shared and local variables with arbitrary values.
-%initialise them with a value different from the default \lstinline{0} or to initialise a local variable.
-%did not allow to assign an arbitrary value to a variable (the default value \lstinline{0} was used).
-
-%Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only initialisation of the shared variable and only by default value \lstinline{0}.
-%\porthos[2] supports arbitrary declaration to be performed in initialisation statement.
+
+Finally, the grammar used by \porthos{} has the following minor drawbacks. %A minor drawback of 
+The operators are non-associative (expressions of the form `\lstinline{1 + 2 * 3}' can not be parsed).
+Comparing to C statements, which must end with the semicolon punctuator `\lstinline{;}', statements defined in the grammar of \porthos[1] use the semicolon as a separator between statements (the final statement must not end with semicolon).
+The litmus test-specific syntax for variables initialisation is used only for declaring the shared variables (all of them are initialised with default value \lstinline{0}), however, this syntax should be used as an initial assignment of both shared and local variables with arbitrary values.
 
@@
 
-The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}.
-%The rectangles in the picture denote processing units (marked with gear sign with a unique number of the component), and labels of the arrows describe the internal representation carried out during the transformation.
-
-%TODO: ATTENTION HERE TO THE PICTURE!!!!!!!!!!!!!!!!!!!!!!!!
-\begin{figure}[H]
+\begin{figure}%[H]
   \centering
-  \includegraphics[height=.95\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/my/draw.io/general_arch-no_numbering.png}
   \caption{The general architecture of \porthos[2]}
@@
 
+The general processing scheme of old tool \porthos[1] is the following.
+The input program is parsed to the AST, where leaves are the events.
+The instructions of the AST are then encoded into the $\phi_{CF}$ (see Section~\ref{ch:enc:bmc:cf}) and the events are processed by the single function (`\texttt{dartagnan.wmm.Domain.encode}') that loops over all events, checks their properties (the kind of event, the operands of the memory event, the type of the barrier event, etc.) and adds corresponding relations assertions to the formula.
+Encoding all the relations (the \textit{program domain encoding}) in a single function is error-prone and hardly manageable.
+Moreover, this function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
+In \porthos[2], we moved the encoding logic into a separate module \ref{ch:impl:proc:z-encoder} that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.
+
+The high-level architecture of \porthos[2] is presented in Figure~\ref{fig:arch}.
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT solver (Section~\ref{ch:impl:proc:smt-solver}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
 
@@
 
-%Internal representations used in \porthos[1]
-
-%As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
-For keeping the architecture transparent, we build all abstraction levels with interfaces, even if some of them does not add any new functionality.% (as java does not support multiple inheritance for classes (only for interfaces), we
+For keeping the architecture transparent, we build all abstraction levels with interfaces, even if some of them does not add any new functionality.
 
@@
 
-The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a high-level recursively%
-%
-\footnote{Hereinafter we use the term \textit{recursive} data structure (sometimes called \textit{inductive}) to refer a complex data type which can contain elements that contain other elements of the same type.
-For example, an unary expression contains references to the operator and the child expression of the same type.} %
-%
-defined AST.
+The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents an untyped high-level recursively defined AST%
+%
+\footnote{Hereinafter we use the term \textit{recursive} data structure (also called \textit{inductive} data structure) to refer a complex data type that can contain elements of the same type.} %
+%
+.
 The Apendix~\ref{apx:trees} presents the file tree of the main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only, which we tend to retain clearly structured).
 
-%\columnratio{0.7}%\begin{paracol}{2}
-%\switchcolumn%\VerbatimInput{inc/lst/Ytree-tree.out}%\end{paracol}
 The abstract syntax tree, \ytree{}, is an abstraction level suitable for compiling the program to a low-level representation (in the case of processing low-level assembly code, it may be directly converted to the \xgraph{} representation).
 In terms of \porthos[1], the \ytree{} is the level of \textit{instructions}.
-%In contrast to \ytree{} representation, the \tool{ANTLR} syntax tree follows the same structure as the grammar, which is a superset of the real (meaningful) grammar of the C language.
 %However, it lacks some concepts of the language (for example, the syntax of indexer access corresponds the grammar rule "\lstinline{postfixExpression '[' expression ']'}" for an arbitrary postfix %, that is converted to the \texttt{YIndexerExpression}).
@@
 
-Each \ytree{} element implements the interface \texttt{YEntity} and carries the \texttt{OriginLocation} %
-%\textbf{(TODO: Rename CodeLocation->OriginLocation in the code!)} %TODO <--
-instance that contains information about the coordinates of the input text that has generated the \ytree{} element.
+Each \ytree{} element implements the interface \texttt{YEntity} and carries the \texttt{Origin} (original code coordinates) instance that contains information about the coordinates of the input text that has generated the \ytree{} element.
+The origin can be translated to the code citation by the \texttt{CitationService}.
 
 Following the C11 standard~\cite{c11}, we distinguish a \textit{statement} (\textit{"an~action to be performed"}) from an \textit{expression} (\textit{"a sequence of operators and operands that specifies computation of a value, or that designates an object or a function, or that generates side effects, or that performs a combination thereof"}).
-
 All \ytree{} expressions implement the \texttt{YExpression} interface.
-On the \ytree{} level, the pointer arithmetic is modelled by the integer number \textit{pointer level} of an expression (although in fact this is the property of a type not of an expression, the y-tree is an untyped syntax tree, therefore the elements \ytree{} should carry this property).
-We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is defined also syntactically).
-Each \ytree{} expression carries the original code coordinates, that can be resolved to text by the service \texttt{LocationService}.
-%todo: rename from CodeCitationService + create 3 instances for each of level.
-Or keep it for Y->text, and store references to the original Y-object by X-objects
-
-%\textbf{(TODO:THIS SHOULD BE IN PRE-PROCESSING)} %TODO <--
+Although the pointer arithmetic is not fully supported by \porthos[2], the \ytree{} contains information about the \textit{pointer level} of an expression (which is modelled as an integer number). 
+We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is determined syntactically).
 
@@
 \begin{itemize}%[nolistsep]
-  \item \texttt{YBinaryExpression} that model the C binary operator (\textit{relative} operator that compares two expressions of any type, \textit{logical} that processes two boolean expressions, and \textit{numerical} that processes two numerical expressions);
-  %\textbf{TODO: rename 'integer'->'numerical' in code} %TODO <--
+  \item \texttt{YBinaryExpression} that model the C binary operator (\textit{relative} operator that compares two expressions of any type, \textit{logical} that processes two boolean expressions, and \textit{numerical} that processes two integer or real expressions);
 
@@
 
-  \item \texttt{YMemberAccessExpression} that has an arbitrary expression of type \texttt{YExpression} as its base expression (it will be resolved during the compilation stage);%TODO: pre-compilation??
-
-  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT solver we use supports the constant-array theory, we can maintain the array type);
+  \item \texttt{YMemberAccessExpression} that has an arbitrary expression of type \texttt{YExpression} as its base expression (it will be resolved during the compilation stage, see Section~\ref{ch:impl:proc:x-compiler});
+
+  \item \texttt{YIndexerExpression} and \texttt{YInvocationExpression} that as arbitrary expression as its base or arguments (strictly speaking, the indexer expression is an unary-function invocation, but as the SMT-solver we use supports the constant-array theory, we can maintain the array type);
 
@@
 
-  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}- loops;%TODO:RENAME YWhileLoopStatement->YLoopStatement in code!!!)}
+  \item \texttt{YLoopStatement} representing both \texttt{while}- and \texttt{for}- loops;
 
@@
 
-On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as a \textit{reference} for the type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information on declaration, including the type itself, type modifiers and qualifiers).
+On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as a \textit{reference} for the type (since the \ytree{} is not typed, the \texttt{YType} is used for storing the type information on variables or types declarations).
+It consists type modifiers and qualifiers, that will be converted to the X-type by the \texttt{Y2XTypeConverter} during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
 
@@
 Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific declaration \texttt{YProcessDefinition}
-%\textbf{(TODO: rename YProcessStatement->YProcessDefinition)} %TODO <--
 used in intra-procedural analysis mode.
-%TODO: rename YMethodSignature -> YFunctionSignature
-The function definition contains the \texttt{YCompoundStatement} body and the \texttt{YMethodSignature} signature, which is used in the function resolution during the compilation stage.
-%TODO: pre-compilation?
-The other litmus-specific statements are \texttt{YPreludeDefinition}
-%\textbf{(TODO: rename YPreludeStatement -> YPreludeDefinition)} %TODO <--
-that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeDefinition} 
-%\textbf{(TODO: rename YPostludeStatement -> YPostludeDefinition)} %TODO <--
+The function definition contains the \texttt{YCompoundStatement} body and the \texttt{YMethodSignature} signature (method name and types of formal parameters), which is used in the function resolution during the pre-compilation stage.
+The other litmus-specific statements are \texttt{YPreludeStatement}
+that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeStatement} 
 that carries the \texttt{YExpression} binary expression to be asserted by the litmus test.
@@
 The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
-The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
-%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
+The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:model}.
 The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
@@
 
-Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
+Following the format of litmus tests, we distinguish three types of \xgraph{}: one for the parallel processes, one for the litmus-initialisation block and one for the assertion statement.
 However, all three types of the \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph.
-%TODO: rename to XThread ?
 This simplifies drastically the processing of different types of code blocks as they all are modelled by the same data structure.
 The examples of restrictions on X-graphs are the following: the initialisation block can not have branchings, fence events; the process cannot have assertion events; the assertion block can not have shared-memory events or fence events.
-We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler:compilation}.
+We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler}.
 
@@
 
-The \textit{memory unit} is a memory cell of an abstract machine executing the code.
-This machine has an infinite number of arbitrary-sized \textit{registers} (local memory units) and \textit{locations} (shared memory units).
-Local memory units inherit the \texttt{XProcessLocalElement} interface, that stores the ID of the owning process.
+An \textit{X-memory unit} is a memory cell of an abstract machine executing the code (that represents the \xgraph{}).
+The abstract machine has infinite number of arbitrary-sized \textit{registers} (local memory units) and \textit{locations} (shared memory units).
+Local memory units extend the \texttt{XProcessLocalElement} interface, that stores the ID of the owning process.
 Figure~\ref{fig:class-diagrams:XMemoryUnit} represents inheritance hierarchy of memory units.
 
-%TODO!!! Remove watermark from the pictures!
 \begin{figure}[t]%[!b]%[H]
@@
 In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language.
-See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %TODO!!!!!
+See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler}.} %TODO!!!!!
 Check that XType is really a simple enumeration
 %
-that is easily convertible to an SMT-type (modelled as \texttt{ZType}). %TODO: add the class ZType
+that is easily convertible to an SMT-type (modelled by \texttt{ZType}). %TODO: add the class ZType
 All X-memory units carry the boolean flag that indicates whether it has been resolved correctly by the X-memory manager.
@@
 The memory units are created and stored by the \texttt{XMemoryManager}, which provides interface for accessing memory units during compilation stage.
-For more detailed description of memory management see Section~\ref{ch:impl:proc:x-compiler:mem}.
+For more detailed description of memory management see Section~\ref{ch:impl:proc:x-compiler}.
 
@@
 
-An event (\texttt{XEvent}) represents the fact of executing the primitive operation, which is independent from other events.
-Each event is specified by the process generated them and a unique event label.
-This information is stored by events in the immutable structure \texttt{XEventInfo}.
-Also, each event carries the reference to the Y-instruction that has generated it.
+An \textit{X-event} represents the fact of executing the primitive operation, which is independent from other events.
+Each X-event implements the \texttt{XEvent} interface and is carries the information about the process generated them and a unique event label, which is modelled by an immutable structure \texttt{XEventInfo}.
+Each event should carry the reference to the Y-instruction that has generated it (this information can be useful for the unrolling discussed in Section~\ref{ch:impl:proc:x-unroll}, however currently each X-event contains the reference to the original code location, similarly to Y-instructions discussed above).
 
 The following interfaces model basic kinds of events (see Figure~\ref{fig:class-diagrams:XEntity-interfaces}):
-%\begin{itemize}[noitemsep]
+
 \begin{outline}
@@
   There are four types of memory events (the arrow denotes the direction of the data-flow):
-  %\begin{itemize}[noitemsep]
   \2 \texttt{XRegisterMemoryEvent}: \\
@@
 
-  %\end{itemize}%TODO incompleted enumeration
-  %TODO: either say st about getLoc(), getReg(), or remove them from code
+  %TO-DO: either say st about getLoc(), getReg(), or remove them from code
 
   \1 \texttt{XComputationEvent} \\
-  %TODO: say sth about how we not produce thousands of comput events
     We distinguish two types of computation events:
-      %\begin{itemize}[noitemsep]
       \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation; and 
       \2 \texttt{XBinaryComputationEvent} that encodes numeric operations (such as addition, multiplication, etc.), bit vector operations (such as bit-and, bit-xor, etc.), relative operations (such as greater-then comparison, equality comparison, etc.), and logical operations (such as conjunction and disjunction).
-      %\end{itemize}
     \newline
@@
     Note, for the purpose of simplifying the X abstraction level, computation events may have been modelled as invocations of bodiless functions (for instance, the operation `\lstinline{x + y}' may be modelled as the invocation of the function `\lstinline{+}' with the arguments \lstinline{x} and \lstinline{y}).
-    However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT solver.
+    However, current version of \porthos[2] maintains the \texttt{XComputationEvent} as the operators are supported by the SMT-solver.
 
@@
     \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
-    %TODO: check the diagrams if the XMCE really inherits XLMU
     The function call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the function call.
-    If the function has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the function body.
-    The call stack is bounded by the user-defined parameter; once the stack is full, the interpretation continues without jumping to the body of the invoked function (such cases are properly logged).
-%TODO: call-stack: user-defined parameter
-    Each \texttt{return} statement creates the assignment of the function call event on the top of call stack.
-    Note, this approach works for any kind of recursion, which is unrolled up the user-defined boundary as well as explicit loops.
+    Currently, \porthos[2] only supports invocations of the method calls registered in its knowledge base (see invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}).
+    For interpreting known function invocations, \porthos[2] must perform three steps: bind actual arguments to the formal parameters (treated as temporary local memory units), push the called \texttt{XMethodCallEvent} onto the call stack, and the perform the control-flow jump to the function body.
+    The call stack must be bounded by the user-defined parameter; once the stack is full, the interpretation should continue without jumping to the body of the invoked function (such cases must be properly logged).
+    Each \texttt{return} statement must create the assignment of the function call event that is on the top of call stack.
+    Note, this approach will work for recursion as well, and the recursive calls together with loops must be unrolled up the user-defined boundary (see Section~\ref{ch:impl:proc:x-unroll}).
     \\
-    %According to the Rice's theorem~\cite{rice1953classes}, in general case there always will be functions unresolved within an analysis launch.
-    If the function has not been resolved, it can be safely assumed to be a no-operation function.
+    If the function has not been resolved, it can be safely assumed to be a no-operation function (with proper logging).
     In other words, we suppose that the knowledge base of \porthos[2] is complete and the tool can resolve all memory-operation functions and fence instructions.
-    Although this can affect the completeness of the analysis, it is safe to make such an assumption as the user can check the log file and manually analyse the semantics of an unresolved call and it to the knowledge base if necessary.
+    Although this can affect the completeness of the analysis, it is safe to make such an assumption as the user can check the log file and manually analyse the semantics of an unresolved call and add it to the knowledge base if necessary.
 
@@
     Current implementation of \porthos[2] supports all fences supported by \porthos{}: 
-    \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
-    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each!
+    \texttt{mfence}, \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.
 
@@
     The fake events are the auxiliary elements of \xgraph{}.
-    %\begin{itemize}[noitemsep]
     \2 \texttt{XEntryEvent}, the per-process unique source event in the event-flow graph,
@@
     \2 \texttt{XAssertionEvent}, the reachability assertion made at the postlude statement of the program.
-    An assertion is modelled as an event for the purpose of encoding the postlude statement as a separate process that is compatible with the encoder.
-   % \end{itemize}
-%\end{itemize}
+    An assertion is modelled as an event for the purpose of encoding the postlude statement as a separate process that is compatible with the Z-encoder.
+
 \end{outline}
-
-%TODO: say about IDs from old code
 
@@
 \label{ch:impl:model:xgraph:edges}
-As the graph is represented by an adjacency matrix, its edges are stored in (immutable) hash-maps.
+As the graph is represented by an adjacency matrix, its edges are stored in immutable hash-maps.
 We distinguish the following kinds of edges:
 
-\begin{outline}%[noitemsep]
+\begin{outline}
   \1 the \textit{control-flow edges}:
@@
 \label{ch:impl:model:xgraph:invariants}
-Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:enc:bmc:cf}):
+Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:wmm:model:events}):
 
@@
 
-The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT solvers.
+The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
 As all other internal representations, the \zformula{} is implemented as an immutable data structure.
 
-The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT solver.
+The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
 Generally, a \zformula{} represents the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}. % the semantics of SMT-LIB formulas defined in~\cite[Chapter 5]{smt-lib}.
@@
 %TODO: !!! Say that interpreter does not know anyth about Y, all its interface methods are <sth> -> <X>. The only part that know about Y is Y2Xvisitor
+
+  %TODO: say sth about how we not produce thousands of comput events
 
@@
 
-The \zformula{} encoder transforms the program and the memory models to a single SMT-like representation.
-It follows the encoding scheme described in Chapter~\ref{ch:enc}.
+The \zformula{} encoder accepts as input the program and the input memory models.
+It visits the recursive memory model AST and each its derived relation definition (recall the `\texttt{let}' keyword of the \cat{} language) it transforms to the set of SMT-clauses, generated on the basis of the program events with respect to the encoding scheme described in Chapter~\ref{ch:enc}.
+%todo: currenlty not implemented, just converted to the old
 %TODO: say more!
@@
 
-\subsubsection{SMT solver}
+\subsubsection{SMT-solver}
 \label{ch:impl:proc:smt-solver}
 %TODO: check that we have ALL references in pre-word in this chapter
-As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
+As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT-solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
 

### 2018-05-27 5_eval.tex
Cleanup finale up to Chap 3.
Cleanup finale up to Xevent.

@@
 %As an example of the compilation and unrolling processes, consider the simple
-%As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler:compilation}, 
+%As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler}, 
 Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
@@
 
-Note, that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
+Note that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
 Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
-The time spent by the SMT solver while considering the SMT formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
+The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
 The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
@@
 
-%<TODO: compare numbers of clauses of SMT formulas>
+%<TODO: compare numbers of clauses of SMT-formulas>
 
@@
 Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \texttt{READ\_ONCE} and \texttt{WRITE\_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
-Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler:hooking}.
+Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
 

### 2018-05-27 4_impl.tex
Cleanup finale up to unrolling.

@@
 The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
+The \textit{full} list of classes that constitute the \xgraph{} hierarchy can be found in Appendix~\ref{apx:trees}.
 Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph}).
@@
 \paragraph{Memory units}
-\label{ch:impl:model:xgraph:mem}
+%\label{ch:impl:model:xgraph:mem}
 
@@
 \paragraph{Events}
-\label{ch:impl:model:xgraph:evt}
+%\label{ch:impl:model:xgraph:evt}
 
@@
     \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation; and
-    \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
+    \2 \texttt{XFunctionCallEvent} that models function invocations.
     The function call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the function call.
     Currently, \porthos[2] only supports invocations of the method calls registered in its knowledge base (see invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}).
-    For interpreting known function invocations, \porthos[2] must perform three steps: bind actual arguments to the formal parameters (treated as temporary local memory units), push the called \texttt{XMethodCallEvent} onto the call stack, and the perform the control-flow jump to the function body.
+    For interpreting known function invocations, \porthos[2] must perform three steps: bind actual arguments to the formal parameters (treated as temporary registers as in the \texttt{fastcall} calling convention), push the called \texttt{XFunctionCallEvent} onto the call stack, and the perform the control-flow jump to the function body.
     The call stack must be bounded by the user-defined parameter; once the stack is full, the interpretation should continue without jumping to the body of the invoked function (such cases must be properly logged).
@@
 \paragraph{Edges}
-\label{ch:impl:model:xgraph:edges}
+%\label{ch:impl:model:xgraph:edges}
 As the graph is represented by an adjacency matrix, its edges are stored in immutable hash-maps.
@@
 \paragraph{Graph invariants}
-\label{ch:impl:model:xgraph:invariants}
+%\label{ch:impl:model:xgraph:invariants}
 Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:wmm:model:events}):
@@
 \item all write-event nodes except final write-nodes must have exactly one \co-successor.
-%TODO: \item \textbf{TODO: more?}
 \end{enumerate}
@@
 %\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
-%\label{ch:impl:model:xgraph:builder}
-%TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
-
-%\paragraph{Graph example} %(in 5_eval.tex)
-%\label{ch:impl:model:xgraph:example}
-
-%hashcode, uniqueness
-
-%\lstinputlisting{inc/XEventVisitor-cleaned.java}
+%TO-DO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
 
@@
 \label{ch:impl:model:wmodel}
-%TODO: almost not implemented !
 
-The \textit{\wmodel{}} represents a recursive AST of \textit{computations over relations} and assertion expressions defined by the memory model.
+The \textit{\wmodel{}} represents a recursive AST of \textit{computation over relations} and \textit{assertion expressions} defined by the memory model.
 The atomic elements of \wmodel{} are the basic relations (\po{}, \rf{} and \co{}; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
 The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.
-For the sake of transparency, each element of a \wmodel{} contains the origin location as the coordinates of the string in the model file. %TODO: implement origin for w-model
-
-%todo: enumerate elements -- after refactoring>
-
-%todo: rename everywhere assertions <- axioms
+Each element of \wmodel{} implements the interface \texttt{WModel}, and for the sake of transparency contains the origin location in the model file. %TODO: implement origin for w-model
+As the \wmodel{} hierarchy almost completely follows the mathematical structure discussed in Sections~\ref{ch:wmm:cat} and~\ref{ch:enc:bmc:wmm}, we skip the details of its implementation.
 
@@
 
-%TODO: not implemented at all !
+The \textit{\zformula{}} representation is supposed to be a wrapper over an SMT-formula used as an additional abstraction level to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
+However, currently \porthos[2] skips this internal representation and both the program and memory models directly into the SMT-formula via Java API offered by the Z3 solver.
+%As all other internal representations, the \zformula{} should be implemented as an immutable data structure.
 
-The \textit{\zformula{}} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
-As all other internal representations, the \zformula{} is implemented as an immutable data structure.
+The Z-abstraction level should model the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
+Generally, a \zformula{} should represent the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}.
+Currently, the \zformula{} is supposed to represent definitions and assertions over variables and binary and unary expressions.
+However, for the purpose of improving the encoding scheme, it may be extended to support function symbols and binders (existential and universal quantification, pattern matching and functional type construction).
 
-The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
-Generally, a \zformula{} represents the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}. % the semantics of SMT-LIB formulas defined in~\cite[Chapter 5]{smt-lib}.
-Currently, the \zformula{} supports definitions and assertions over \textit{variables} and binary and unary \textit{expressions}.
-However, in future it may be extended to support \textit{function symbols} and \textit{binders} (existential and universal quantification, pattern matching and functional type construction).
-
-All expressions of a \zformula{} are \textit{typed} (or \textit{sorted}, in terms of SMT-LIB standard).
-Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, bitvector, real) expressions.
-The typing of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-LIB formula.
-For the sake of transparency, each element of a \zformula{} contains the origin location as the reference to the X-element that produced current Z-element. %TODO: implement origin for z-formula
-
-%todo: enumerate elements -- after refactoring>
+All expressions of a \zformula{} must be \textit{typed} (or \textit{sorted}, in terms of SMT-LIB standard).
+Basically, the Z-type must support boolean and numeric (integer, bitvector, real) expressions.
+The typing of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-formula.
+For the sake of transparency, each element of a \zformula{} should contains the origin location as the reference to the X-element that produced current Z-element.
 
@@
 
-%todo: after all, list HERE which modules (classes, managers) are being initialised during the pre-compilation. I'ts good to have the dependencies graph on moduls (operational semantics:) )
-
-This section describes program units that construct, transform and analyse internal representations described before.
-% steps of construction of the program representations and transformations applied to them.
-We assign unique numbers to the processing units, which can checked in Figure~\ref{fig:arch} (within gears depicted in the top-left corner of each processing unit).
-%Further in current Section, the subsection titles have a number in parentheses, which correspond to a unique number of the  processing unit being discussed.
-%These numbers are pictured within the gear in the top-left corner of each processing unit in Figure~\ref{fig:arch}.
-%he numbers in parentheses correspond the numbers of processing units pictured within gears in Figure~\ref{fig:arch}.
+This section describes program units that construct, transform and analyse internal representations described above.
+%not anymore %Each processing unit is assigned with a unique number, which can checked in Figure~\ref{fig:arch} (within gears depicted in the top-left corner of each processing unit).
 
 The construction of the \xgraph{} is performed in three stages.
-First, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
+Firstly, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
 Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
-After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be \textit{CF-acyclic} (no cycles among control-flow edges).
+After that, the compiler is able to perform the data-flow analysis over the \xgraphU[CF] and produce the full event-based graph \xgraphU[CF+DF], which remains to be \textit{CF-acyclic} (no cycles among control-flow edges).
 
@@
 Thus, the method call resolution is performed statically at compile-time without any overhead at run-time.
-%To visit an instance \lstinline{x} by a visitor \lstinline{v}, the program should call the accept-method of \lstinline{x} and pass \lstinline{v} as an argument.
 In our implementation, the visitor (not the visitee) cares about the continuation of traversing its children.
@@
 \centering
-%\begin{multicols*}{2}
 \begin{minipage}[t]{.55\textwidth}
@@
 \end{lstlisting}
-%[* \vfill\null \columnbreak *]
-\end{minipage}% <---------------- Note the use of "%"
+\end{minipage}
 \begin{minipage}[t]{.45\textwidth}
@@
 \end{lstlisting}
-%\end{multicols*}
 \end{minipage}
@@
 
-For traversing graph data structures, we mostly follow the \textit{iterator} pattern: the special object \texttt{Iterator}, that has access to elements of the data structure, iterates over them in some order (for example, in topological order as it is implemented for \xgraph{}).
-%Since in \porthos{} we usually do not need to browse an object elements in different order, the iterator creation is performed by the data structure itself.
-%However, in contrast to the visitor pattern, the visitor of a plain data structure should be used with an \textit{iterator} that has access to the elements of the object and passes them to the visitor.
+For traversing plain data structures (such as graphs), we mostly follow the \textit{iterator} pattern: the special object \texttt{Iterator}, that has access to elements of the data structure, iterates over them in some order (for example, in topological order as it is implemented for \xgraph{}).
 The construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
-A builder is a mutable object that contains methods "filling" it by the elements that will constitute the result complex object.
+A builder is a mutable object that contains methods for "filling" it with the elements that will constitute the result complex object.
 The methods contain basic validity checks of the elements.
 Once a builder has been built, it cannot be modified.
-
-% todo: test pre-/post-fix operations //+think how current tool will behave if there's more than one post-operation //and pre-operation
 
@@
 
-Both input-language and input-model parsers are implemented via \tool{ANTLR} parser generator.
-%In Figure~\ref{fig:arch}, the input parsers are denoted as (1) and (2) processing units.
-Details on the input program language grammar are discussed in Section~\ref{ch:impl:input}.
-Currently, \porthos[2] does not consider preprocessor instructions (it ignores them).
+As it was discussed in Section~\ref{ch:impl:input}, both input-language and input-model parsers are implemented via \tool{ANTLR} parser generator.
+Currently, \porthos[2] does not consider C preprocessor instructions (it ignores them).
 For implementing the inter-procedural mode of \porthos{}, it is crucial to support inclusion of header files (the \texttt{\#include} preprocessor directive).
@@
 As preprocessor statements may appear in arbitrary place of a program, the preprocessor must be a stateful processing unit that reads the token stream and dynamically instrument the program by interpreting directives and expanding macros.
-%The \tool{ANTLR} grammar for the C preprocessor may be found 
-%more macros (stateful) : https://gcc.gnu.org/onlinedocs/cpp/Macros.html
 
@@
 %TODO: upload the grammar to github
-%The input memory model 
-%TODO: add reference to the grammar file in github repository OR include a short version as an appendix}
-
-%todo: maybe: move something from the input language chapter here
 
@@
 The \wmodel{} representation is constructed by the stateless visitor \texttt{Cat2WmodelConverterVisitor} from the ANTLR syntax tree \texttt{CatParser.MainContext} of the memory-model defined in \cat{} language.
-%In Figure~\ref{fig:arch}, the \wmodel{} constructor is denoted as the (3) processing unit.
-Currently, the visitor supports the small subset of \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full OCaml interpreter.
-However, some most commonly used functional-style expressions may be supported by mapping them syntactically to corresponding W-elements by the \wmodel{} constructor.
+Currently, the visitor supports the small subset of \cat{} language, which constitutes only non-functional declarative expressions of the language as the support for functional-style expressions requires implementing the full interpreter for OCaml-like language.
+However, some most commonly used functional-style expressions may be supported by mapping them syntactically to corresponding W-elements directly in the \wmodel{} constructor.
 
-Following the principle of transparency, the \wmodel{} constructor aborts its work with the \texttt{NotSupportedException} if met the unrecognised syntax construction (in contrast to the \porthos[1] approach in which the \texttt{null} value was produced in all exceptional states).
+Following the principle of transparency, the \wmodel{} constructor aborts its work with the \texttt{NotSupportedException} if met the unsupported syntax construction (in contrast to the \porthos[1] approach in which the \texttt{null} value was produced in all exceptional states).
 
-%todo: file tree of w-model to appendix -- after code fixing>
+%todo: file tree of w-model to appendix -- after implementing + more discussion here
 
@@
 The \ytree{} is constructed by the stateless visitor \texttt{C2YtreeConverterVisitor} from the ANTLR syntax tree \texttt{C11Parser.MainContext} of the C language.
-As the \wmodel{} constructor, the \ytree{} constructor aborts its work with the \texttt{NotSupportedException} once it meets an unsupported syntax construction.
-%TODO: Rename YConverterException <- YParserException
-A syntax exception \texttt{YConverterException} is thrown if the converted syntax tree contains semantic errors that prevent it to be converted to the \ytree{}.
+The \ytree{} constructor aborts its work with the \texttt{YParserNotImplementedException} once it meets an unsupported syntax construction.
+A syntax exception \texttt{YParserException} is thrown if the converted syntax tree contains semantic errors that prevent it to be converted to the \ytree{}.
 
@@
 
-%TODO: not supported yet
+%TODO: <not implemented yet>
 
 \paragraph{The label resolution}
-\label{ch:impl:proc:x-pre-compiler:label}
+%\label{ch:impl:proc:x-pre-compiler:label}
 
@@
 
-The label resolution is necessary for establishing links to labelled statements. %a process of linkage the text labels referenced in code to the places of their declaration.
+The label resolution is necessary for establishing links to labelled statements.
 In C, labelled statements are declared via the colon-syntax `\texttt{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\texttt{goto\;<label>}'.
-The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labeled statement.
-%TODO: code: remove temp label generation, it's stupid.
-put null.
+The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labelled statement.
 This information is used during compilation to set up unconditional jumps.
@@
 \paragraph{Type analysis}
-\label{ch:impl:proc:x-pre-compiler:type}
+%\label{ch:impl:proc:x-pre-compiler:type}
 
@@
 Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
-% according the rules of propagation.
-For example,
 Being carried at Y- and X- representation levels, the type is converted to a~Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
 
-%TODO: YOU REALLY NEED TO THINK MORE ABOUT TYPES ON Y, X, Z LEVELS + reconsider this paragraph (and others about types)
-Currently, \porthos[2] handles only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
-The array dimension is stored at the Y-level as an integer number (currently this information is not used as \porthos[2] does not support arrays, although this information can be used at the stage of pre-compilation for converting array elements to separate variables).
+Currently, \porthos[2] handles only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which are modelled on the X-level by the enumeration \texttt{XType}.
+%TODO: describe the XType (before) + implement this clearly! + definitely get rid of bit32
+%The array dimension is stored at the Y-level as an integer number (currently this information is not used as \porthos[2] does not support arrays, although this information can be used at the stage of pre-compilation for converting array elements to separate variables).
 %the array operations are converted to ).
-In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter}, %TODO: Or XTypeHookManager?
-which can be requested for converting an X-type to a~Z-type. %TODO: implement this.
-%TODO: describe the XType (before) + implement this clearly! + definitely get rid of bit32
-
-%TODO: <not implemented>
-The type analysis algorithm should consider the type aliases supported by C language (defined by the \texttt{typedef} instruction).
+%In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter},
+%which can be requested for converting an X-type to a~Z-type. %TODO: implement this.
+The type analysis algorithm should consider the type aliases supported by C language (defined by the \texttt{typedef} instruction), which is not implemented yet.
 For resolving the type aliases, the precompiler should make an extra traverse of the \ytree{} before the pre-compilation stage and build up a symbol map.
 
-%TODO: <not implemented>
-The type analysis also includes the process of resolving semantics of function invocations.
-The result is stored in a map of invocation expression to the resolved function signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
-However, due to lack of polymorphism and function overloading mechanism inherent to OOP languages, the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.
-
-%TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
-%The X-type construction: \textbf{TODO} //code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
-%- on Z level, type is one-pointer-level
-%- The typification algorithm: propagation rules for each of Y
-%- as vector array as needed.
-%%
-%Y: quantifiers+specifiers+typeName
-%X: ?.. Z? TODO: rename+rework XType->XZType
-%Z: Int, Real, Boolean, BitVector[N]
+%<not needed below>
+%The type analysis also includes the process of resolving semantics of function invocations.
+%The result is stored in a map of invocation expression to the resolved function signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
+%However, due to lack of polymorphism and function overloading mechanism inherent to OOP languages, the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis processing unit may be used for analysing the C++ code if needed.
 
 \paragraph{Variable kind analysis}
-\label{ch:impl:proc:x-pre-compiler:var}
+%\label{ch:impl:proc:x-pre-compiler:var}
 
@@
 
-%\vspace{1em}
 The following types of variables are detected as \textit{global variables}:
-%TODO: not implemented
-\begin{itemize}
+%TODO: <not implemented yet>
+\begin{itemize}[noitemsep,topsep=0pt]
 \item a variable was declared as a pointer;
 \item a variable whose address was accessed by any process;
-\item a variable that was declared as a parameter of the process function; and
-%TODO: say sth about KERNEL ANALYSIS MODE BEFORE !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-\item a variable that was exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking).
-%https://lwn.net/Articles/154602/
+\item a variable declared as a parameter of the process function; and
+\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking~\cite{argawal16kernel}).
 \end{itemize}
@@
 
-The X-compiler is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
-It is a complex processing unit; Figure~\ref{fig:compiler} illustrates the relationship between the principal components of the X-compiler in the UML language.
+The \textit{X-compiler} is the main component that transforms the recursive \ytree{} data structure to the plain \xgraph{} representation.
+It is a complex processing unit; Figure~\ref{fig:compiler} illustrates the relationship between main components of the X-compiler in the UML language.
 
@@
 The \texttt{XInterpreter} carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the filling in the builder.
-The interpreter requests the managers such as \texttt{XMemoryManager}, \texttt{XHookManager}, \texttt{XTypeManager} for additional information.
+The interpreter has additional modules \texttt{XMemoryManager}, \texttt{XHookManager} and \texttt{XTypeManager} that provide specialised functionality.
 
-We distinguish three types of processes (each implementing the \texttt{XInterpreter} interface):
+We distinguish three types of interpreters (each implementing the \texttt{XInterpreter} interface):
 \begin{itemize}[noitemsep,topsep=0pt]
-\item the prelude process \texttt{XPreludeInterpreter} (to be executed before all other processes) that allows only declaration of local or shared variables, computations and memory operations;
-\item the postlude process \texttt{XPostludeInterpreter} (to be executed after all other processes) that allows only declaration of local variables, memory operations, computations and assertions;
-\item the process \texttt{XProcessInterpreter} (to be executed in parallel) that allows all X-compiler interface operations except declaration of shared variables %(that should be declared during pre-compilation stage or during compilation of the prelude process)
+\item the \textit{program} interpreter \texttt{XProgramInterpreter} that is the compositional element that dispatches calls to currently maintained process interpreter;
+\item the \textit{prelude} process \texttt{XPreludeInterpreter} (to be executed before all other processes) that allows only declaration of local or shared variables, computations and memory operations;
+\item the \textit{postlude} process \texttt{XPostludeInterpreter} (to be executed after all other processes) that allows only declaration of local variables, memory operations, computations and assertions;
+\item the \textit{process} \texttt{XProcessInterpreter} (to be executed in parallel) that allows all X-compiler interface operations except declaration of shared variables
 and program assertions (i.e., it allows declarations of local variables, computations, barriers, unconditional jumps, non-linear branching statements, and method calls).
@@
 As the prelude process can declare shared variables, its compilation must go first.
-%TODO: think about sorting processes at the pre-compilation stage
+%TODO: sort processes at the pre-compilation stage !
 The first event of each process must be the \texttt{XEntryEvent} and the last events must be events of type \texttt{XExitEvent}.
@@
 \paragraph{Memory manager}
-\label{ch:impl:proc:x-compiler:mem}
+%\label{ch:impl:proc:x-compiler:mem}
 
@@
 Following the C standard, local memory units have higher priority over global ones (the method \texttt{getDeclaredUnitOrNull} returns the first memory unit found, either a global one or a local one, or null if no memory units with requested name have been registered).
-Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block contexts for local variables.
-%TODO: refactor XMemoryManager!
+Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} should carry the stack of block contexts for local variables (which is not currently implemented).
 
@@
 
-\paragraph{Invocation hook manager}
-\label{ch:impl:proc:x-compiler:hooking}
-
-%TODO: cite again here : alglave2018frightening + other Kernel references
-
-%MethCallE : bind parameters with values (load them to temp registers) + remember the return-register
-%reutrn event= assign the return-register + simple jump
+\paragraph{Invocation hooking manager}
+%\label{ch:impl:proc:x-compiler:hooking}
 
 The invocation hooking module serves as a knowledge base that stores the semantics of functions.
-%TODO: rename HookManager -> XHookManager
 Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in the case of program in C language -- only the function name) across all signatures it stores.
 If the signature matches, the hook manager intercepts the function call and interprets the hook action instead of invoking the function directly.
-%TODO: say sth about the mode
-The hook manager operates multiple invocation hooks depending on the user-defined mode.
+The hook manager enables only invocation hooks that are compliant with the language of the input program.
 All invocation hooks implement the interface \texttt{XInvocationHook}.
@@
 Current version of \porthos[2] contains two invocation hooks, the \texttt{XLegacyInvocationHook} for intercepting the compatibility-mode functions of the \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
-%Note that the same mechanism (with several modifications) can be used for processing assembly language calls.
+The same mechanism (with minor modifications) can be used for processing assembly language calls.
 An invocation hook can model any type of operations (memory, fence, computational, etc.).
-%For instance, the \texttt{XLegacyInvocationHook} intercepts calls \lstinline{`location.store(atomic, register)'} and processes the store operation with respect to the specified atomic.
-For instance, the \texttt{XKernelInvocationHook} intercepts the call \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces the invocation with its hook action shown in Figure~\ref{code-invoc-hook}.
-%and processes the store operation with respect to the following .
-
-\begin{figure}[!h]
-\begin{lstlisting}
-class XKernelInvocationHook extends XInvocationHookBase 
-       implements XInvocationHook {
- ...
- @Override
- public XInvocationHookAction intercept(String functionName) {
-   switch (functionName) {
-     case "WRITE_ONCE": {
-       return new XInvocationHookAction((receiver, arguments) -> {
-         if (arguments.length != 2 || receiver != null) {
-           return null;  // do not intercept
-         }
-
-         XMemoryUnit destUnit = arguments[0];
-         if (!(destUnit instanceof XSharedLvalueMemoryUnit)) {
-           throw new XMethodInvocationError(methodName, 
-               "arg 1: not a shared l-value memory unit: " + destUnit);
-         }
-         XSharedLvalueMemoryUnit dest = 
-               (XSharedLvalueMemoryUnit) destUnit;
-
-         XMemoryUnit srcUnit = arguments[1];
-         if (!(srcUnit instanceof XLocalMemoryUnit)) {
-           throw new XMethodInvocationError(methodName, 
-               "arg 2: not a local memory unit: " + srcUnit);
-         }
-         XLocalMemoryUnit src = (XLocalMemoryUnit) srcUnit;
-
-         return program.emitMemoryEvent(dest, src);
-     });
-   }
-   case "READ_ONCE": {
-     ...
-   }
- }
-}
-\end{lstlisting}
-\caption{Example of the invocation hook for intercepting the Linux kernel-specific functions}
-\label{code-invoc-hook}
-\end{figure}
-
-%TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html
-
-% + RCU %https://www.kernel.org/doc/Documentation/RCU/whatisRCU.txt
+For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \texttt{READ\_ONCE}, \texttt{WRITE\_ONCE}, and \texttt{ACCESS\_ONCE} may be modelled as volatile \texttt{memory\_order\_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
 
 \paragraph{Interpreter}
-\label{ch:impl:proc:x-compiler:compilation}
+%\label{ch:impl:proc:x-compiler:compilation}
 
-%TODO: say that program is the set of processes , but the implements the same interface !!! it's not clear here.
-
-%TODO: !!! Say that interpreter does not know anyth about Y, all its interface methods are <sth> -> <X>. The only part that know about Y is Y2Xvisitor
-
-  %TODO: say sth about how we not produce thousands of comput events
-
-%In this paragraph we discuss only the implementation of the process interpreter \texttt{XProcessInterpreter}.
-%This interpreter is used to build the branched event-flow graph with all kinds of jumps, memory operations and fences.
-The X-program interpreter is invoked by the \texttt{Y2XConverterVisitor} which walks down the \ytree{} recursively.
+The X-program interpreter is invoked by the \texttt{Y2XConverterVisitor} which recursively walks down the \ytree{}.
 The calls to the X-program interpreter are dispatched to currently maintained X-process interpreter.
@@
 The interface methods of the \texttt{XInterpreter} are presented in Appendix~\ref{apx:xinterpreter}.
-Note the clear modular independence of the X-interpreter as it has no methods that operate entities of the Y-level (all conversion work with Y-entities is done by the \texttt{Y2XConverterVisitor}).
+Note the clear modular independence of the X-interpreter: it knows nothing about \ytree{} or other internal representations, all its interfaces use only X-level elements (all conversion with Y-entities is done by the \texttt{Y2XConverterVisitor}).
 
-All interface methods can be divided into two groups.
+All interface methods of the X-interpreter can be divided into two groups.
 The first group constitute methods that emit events.
@@
 Note that the methods \texttt{createComputationEvent} do not emit a computation event but only create one.
-This is performed as optimisation that removes unused computations from the model (otherwise, for instance, the assignment 
-`\texttt{x = 1 * (2 + 3)}' would be compiled to two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
-If the computation event was not used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph}).
+This is an optimisation that removes unused computations from the model
+(otherwise, for instance, the assignment `\texttt{x = 1 * (2 + 3)}' would be compiled into two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
+Note, if the computation event has not been used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph}).
 The second group of X-interpreter methods consists of the methods for defining non-linear statements (branchings and loops).
 These methods change the state of the interpreter and set up additional non-linear control-flow edges.
-%`\texttt{register = 1 * (location + 2)}' would be compiled to three consequent events `\texttt{load(temp\_reg\_1 <- location); eval(temp\_reg\_1 + 2); write(register <- 1 * (temp\_reg\_1 + 2));}',
 
@@
 The context is a data structure that carries the \textit{state} and some additional information in the case of processing non-linear statements (e.g., conditional event, first and last then- and else-branch events for binding, etc.).
-Once the new event has been emitted, the interpreter sets up control-flow edges w.r.t.
-state of the context on the top of the stack (the context stack is always non empty: the bottom context is a linear one).
+Once the new event has been emitted, the interpreter sets up control-flow edges w.r.t. state of the context on the top of the stack.
+The context stack is always non empty: the bottom context is the linear one that sets up the primary edge from the previous-event (last processed) to the currently processed one and updates the previous-event.
 The context state is an enumeration of the following values:
-
 \begin{itemize}[noitemsep,topsep=0pt]
@@
 \item \texttt{WaitingLastConditionEvent}: the next event must be of type \texttt{XComputationEvent}; it will be saved as the conditional branching event;
-\item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branching statement (later the interpreter will set up jumps to that event);
+\item \texttt{WaitingFirstSubBlockEvent}: the first next emitted event will be accepted as the first event of the branch (later the interpreter will set up jumps to that event);
 \item \texttt{WaitingNextLinearEvent}: the default interpreter state for processing next linear event, and
@@
 If so, it iterates over all ready-contexts and sets up control-flow edges considering that the newly emitted event is an exit-event" of the non-linear context (e.g., for a branching context the interpreter adds primary edges from condition-event to the first-true-branch-event and from the last-true-branch-event to the exit-event, the same is done with alternative edges for the false-branch).
-
-As an example of usage the interpreter, consider the code of the \texttt{Y2XConverterVisitor} for processing \texttt{if-then-else} statement of the \ytree{} in Figure~\ref{TODO}.
-
-\begin{figure}
-%TODO!!! refactor the code, move program.tryEvaluateComputation() to Y2Xvisitor or Helper
-\begin{lstlisting}
-public class Ytree2XgraphConverterVisitor implements YtreeVisitor<XEntity> {
-  ...
-
-  @Override
-  public XEvent visit(YBranchingStatement node) {
-    program.startBlockDefinition(XInterpreter.BlockKind.Branching);
-
-    program.startBlockConditionDefinition();
-    XComputationEvent condition = program.tryEvaluateComputation(node.getCondition().accept(this));
-    program.finishBlockConditionDefinition(condition);
-
-    program.startBlockBranchDefinition(XInterpreter.BranchKind.Then);
-    node.getThenBranch().accept(this);
-    program.finishBlockBranchDefinition();
-
-    YStatement elseBranch = node.getElseBranch();
-    program.startBlockBranchDefinition(XInterpreter.BranchKind.Else);
-    if (elseBranch != null) {
-      elseBranch.accept(this);
-    }
-    else {
-      program.emitNopEvent(); // needed for encoding
-    }
-    program.finishBlockBranchDefinition();
-
-    program.finishNonlinearBlockDefinition();
-
-    return null; //statements return null
-    }
-}
-\end{lstlisting}
-\end{figure}
-
-%The \texttt{almostReadyContexts} queue is necessary for not adding erroneous edges on
-
 For the sake of simplicity of the interpretation and encoding, jump events (no-computation events) are present in the flow-graph, however they can be removed from it with rebinding ingoing and outgoing edges.
-
-%algorithm for interpreting
-
-%The method \texttt{tryConvertToLocalOrNull} makes an attempt to convert an arbitrary X-entity to the local memory unit: registers and computation events are casted and returned by the method as-is, shared memory units are copied to a temporary register (\texttt{declareTempRegister}), which is returned by the method.
-%In other cases, the method returns \texttt{null}.
+As an example of usage the interpreter, consider the code of the \texttt{Y2XConverterVisitor} for processing \texttt{if-then-else} statement of the \ytree{} in Appendix~\ref{apx:y2x}.
 

### 2018-05-27 90_summary.tex
Cleanup finale up to Chap 3.
Cleanup finale up to Xevent.
Cleanup finale up to unrolling.

--- 2018-05-25 99_summary.tex
@@
 \label{ch:summary}
+
+%instead of extending the \textit{grammar} of input language, we extended the \textit{architecture} of the tool so that the full C language can be supported
 
@@
   This strongly restricts the syntax of the input language, therefore, for processing programs in C, \porthos[2] has an additional pre-compilation phase that traverses AST of the program and collects information of all variables accessed in the program (see Section~\ref{ch:impl:proc:x-pre-compiler}).
-  Currently, \porthos[2] considers a variable to be shared if it was declared as a pointer, or its address is accessed at least once by the code of any process, or it was declared as a parameter of the function that defines a process, or it was exported by \texttt{extern} or other keyword.
+  Currently, \porthos[2] considers a variable to be shared if it was declared as a pointer, or its address is accessed at least once by the code of any process, or it was declared as a parameter of the function that defines a process, or it was exported by \texttt{extern} keyword or an exporting function.
   %This approach gives flexibility in detecting shared variables 
@@
 
+\item <parallelised>
+
+\item <C processor directives + typedef > 
+<TODO: a short summary here>
+
+\item <contexts for local mem units in mem manager>
+
 \end{itemize}

### 2018-05-27 4_impl.tex
Cleanup finale: finish impl.

@@
     Note, this approach will work for recursion as well, and the recursive calls together with loops must be unrolled up the user-defined boundary (see Section~\ref{ch:impl:proc:x-unroll}).
+    The discussion on how to set up the \po-relations in the event-flow graph with recursive function calls can be found in~\cite{musketeer2014recursion}.
     \\
@@
 In order to be encoded into an SMT-formula, the compiled graph needs to be acyclic~\cite{Porthos17b}.
-To convert it to an acyclic form, we perform the \textit{unrolling} (sometimes called \textit{unwinding}) transfomation: each cycle is unrolled up to the user-defined bound $k$.
-Comparing to \porthos[1], we have changed the meaning of the bound: instead of executing all cycles at most $k$ times, the \porthos[2] interprets the bound as the maximum number of events in the trace of the unrolled graph%
+For that, we perform the \textit{unrolling} (also called \textit{unwinding}) transformation: all cycles are unrolled so that the result number of events in a trace of the \xgraph{} does not exceed the user-defined bound $k$%
 %
@@
 For now, the information about the element of the ANTLR tree that corresponds to the X-event or Y-tree element is being lost during the transformations.
-Instead, we use the \texttt{LocationService} discussed in Section~\ref{ch:impl:model:ytree} that in perspective may be extended for providing this information along with the text citation of the original code.}
+Instead, we use the \texttt{CitationService} discussed in Section~\ref{ch:impl:model:ytree} that in perspective may be extended for providing this information along with the text citation of the original code.}
 %
 .
+This definition of the unrolling bound differs from one used by \porthos[1], where each cycle was executed $k$ times.
 The meaning of a bound was changed in order to increase predictability of the size of the result SMT-formula.
-Note that recursive function calls create unconditional back-jumps in the event-flow graph, which will also be recognised as a loop and be affected by the unrolling algorithm.
-The discussion on how to set up the \po-relations in the event-flow graph with recursive function calls can be found in~\cite{musketeer2014recursion}.
+Note that recursive function calls create unconditional back-jumps in the event-flow graph, which can also be recognised as a loop and be affected by the unrolling algorithm (however, currently not supported).
 
-Considering the new meaning of the unrolling bound, the graph unrolling may be \textit{complete} (the loop has been executed a whole number of times) or \textit{incomplete} (the unrolling bound has triggered on not-the-last event of the loop), which can be modelled by two types of the sink events.
+Considering the new meaning of the unrolling bound, the execution in the unrolled graph can be \textit{completed} (the loop has been executed a whole number of times) or \textit{uncompleted} (the unrolling bound has triggered on not-the-last event of the loop), which can be modelled by two types of the sink events.
 The user may need this information for understanding how the \porthos[2] interprets the cyclic program.
 However, current implementation loses this information by having only one sink event.
-Figure~\ref{fig:loop-unroll} illustrates the unrolling for the left-hand side cyclic control-flow graph (the square node $S_{+}$ denotes the complete sink node, and the $S_{-}$ denotes the incomplete sink).
+Figure~\ref{fig:loop-unroll} illustrates the unrolling for the left-hand side cyclic control-flow graph (the square node $S_{+}$ denotes the sink node for completed executions, and the $S_{-}$ denotes the sink for uncompleted executions).
 
-%TODO: draw (and implement) diff types of exit nodes
 \begin{figure}[t]
@@
 
-The unrolling procedure is performed by the \texttt{XFlowGraphUnroller}. %, which extends the \texttt{FlowGraphDfsTraverser}
-For unrolling the flow-graph, we perform the \textit{Deep-First Search (DFS)} while counting the Y-level instructions (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
-Also, during the unrolling each next event increments the \textit{unrolling depth counter} (the \textit{event reference-ID}) -- the integer that is stored by the \texttt{XEvent} thus making it \textit{an event reference} on non-zero values.
-Each \texttt{XEvent} has the method `\texttt{XEvent asNodeRef(int refId)}' that clones the event-receiver with new value of reference-ID.
+The unrolling procedure is performed by the \texttt{XFlowGraphUnroller}.
+For unrolling the flow-graph, we perform the \textit{Deep-First Search~(DFS)} while counting the X-events (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
+Back-edges are determined via the depth stack (each next event is pushed onto the stack; if the new event is present in the stack, the last processed edge is a back-edge).
+Also, during the unrolling each next event increments the \textit{unrolling depth counter} (the \textit{event reference-ID}) -- the integer stored by the \texttt{XEvent} thus making it an \textit{event reference} on non-zero values (only events of a non-unrolled graph can have the zero-valued reference-ID).
+As an X-event is immutable, it has the method `\texttt{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
 The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
 
-%TODO!!!!!!!!!!! other schemes!\textsl{}
 %During the work on \porthos[2], we have tested several different schemes of loop unrolling.
-
-%- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them)
 
@@
 The \rf{}-edges join each write event to all read events that access the same location; the \co{}-edges join write events to the same location.
+Currently, \porthos[2] does not yet construct the \xgraph[DF] component of the full \xgraph{}, instead, it uses the function `\texttt{dartagnan.wmm.Domain.encode}' of \porthos[1] to encode data-flow relations and derived relations of the \xgraph[CF].
 
@@
 
-The \zformula{} encoder accepts as input the program and the input memory models.
+The \zformula{} encoder accepts as input the unrolled X-graph and the input memory models.
 It visits the recursive memory model AST and each its derived relation definition (recall the `\texttt{let}' keyword of the \cat{} language) it transforms to the set of SMT-clauses, generated on the basis of the program events with respect to the encoding scheme described in Chapter~\ref{ch:enc}.
-%todo: currenlty not implemented, just converted to the old
-%TODO: say more!
+As the \zformula{} representation is not yet implemented, the \zformula{} encoder encodes the \xgraph{} and \wmodel{} directly into the SMT-formula via Java API of the Z3 solver (the program domain encoding is currently a part of the \zformula{} encoder until the full \xgraph[CF+DF] is implemented).
 
-\subsubsection{SMT-formula converter}
-\label{ch:impl:proc:smt-converter}
-
-As the \zformula{} follows the SMT-LIB standard, its translation to an SMT-formula constitutes almost one-to-one mapping.
-The translation is performed by the stateless visitor \texttt{Z2SmtTranslator}.
-
-\subsubsection{SMT-solver}
-\label{ch:impl:proc:smt-solver}
-%TODO: check that we have ALL references in pre-word in this chapter
-As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT-solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
+%\subsubsection{SMT-formula converter}
+%\label{ch:impl:proc:smt-converter}
+%
+%As the \zformula{} follows the SMT-LIB standard, its translation to an SMT-formula constitutes almost one-to-one mapping.
+%The translation should be performed by the stateless visitor \texttt{Z2SmtTranslator}.
+%
+%
+%\subsubsection{SMT-solver}
+%\label{ch:impl:proc:smt-solver}
+%As it has been mentioned in Section~\ref{ch:impl:principles}, \porthos[2] uses the third-party SMT-solver \tool{Z3}, which offers the Java API for constructing requests to the Z3 solver and interpreting its responses.
 
@@
 
-The result of execution of \porthos[2] is the verdict modelled by the class \texttt{AppVerdict}.
+The result of execution of \porthos[2] is the \textit{verdict} modelled by the class \texttt{AppVerdict}.
 This is a structure that contains the result of analysis and auxiliary information such as collected errors and time of execution (separately for each stage of computing).
 The app verdict may be rendered to any format convenient for the user.
-
-%\subsection{Auxiliary components}
-%\label{ch:impl:aux}
-%
-%\subsubsection{Watchdog timer}
-%\label{ch:impl:aux:watchdog}
-%
-%%<not implemented>
-%
-%todo
-%
-%\subsubsection{Logger}
-%\label{ch:impl:aux:logger}
-%
-%%<not implemented>
-%
-%todo
-
-%todo: some words on necessity of contexts and lack of them. How would we implement them
-
-%TODO: \section{Optimisations}

### 2018-05-29 4_impl.tex
add pictures and stuff.

@@
 
-Secondly, \porthos{} resolves the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level%
+Secondly, \porthos{} resolves the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level, %
 %
@@
 %
-, so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.
+so that it does not require to change the grammar for encoding the semantics of a new function. %adding support for a new function.
 %\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
@@
 Moreover, this function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
-In \porthos[2], we moved the encoding logic into a separate module \ref{ch:impl:proc:z-encoder} that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.
+In \porthos[2], we moved the encoding logic into a separate module (see Section~\ref{ch:impl:proc:z-encoder}) that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.
 
@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula, which then is solved by the SMT-solver.
+Although this general architecture is defined for
 
@@
 The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
-The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
+The UML diagram in Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the class hierarchy%
+%
+\footnote{Hereinafter all class diagrams are generated by IntelliJ IDEA Ultimate.}%
+%
+of main interfaces of the X-abstraction level.
 The \textit{full} list of classes that constitute the \xgraph{} hierarchy can be found in Appendix~\ref{apx:trees}.
@@
 Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (unlike r-values, the l-values may be assigned a new value).
-As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, see more detailed discussion further in current Section).
+As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, see more detailed discussion further in the current Section).
 
@@
 
-The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it on Figure.
+The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it in Figure.
 It receives as input the \ytree{}, the memory model kind and user settings (for instance, the interpreter mode defining the set of invocation hooks enabled during the analysis run).
@@
 Back-edges are determined via the depth stack (each next event is pushed onto the stack; if the new event is present in the stack, the last processed edge is a back-edge).
-Also, during the unrolling each next event increments the \textit{unrolling depth counter} (the \textit{event reference-ID}) -- the integer stored by the \texttt{XEvent} thus making it an \textit{event reference} on non-zero values (only events of a non-unrolled graph can have the zero-valued reference-ID).
+Also, during the unrolling each next event increments the \textit{unrolling depth counter} -- the non-zero integer \textit{event reference-ID} stored by each X-event (only events of a non-unrolled graph can have the zero-valued reference-ID)
 As an X-event is immutable, it has the method `\texttt{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.

### 2018-05-29 5_eval.tex
add pictures and stuff.

@@
 
-\subsection{Compilation and unrolling}
-\label{ch:eval:show:compil}
-
-%As an example of the compilation and unrolling processes, consider the simple
-%As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler}, 
 Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
-%(in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
-%The functions are state-equivalent (represent the same state machine) and 
-The functions does not compute any useful value, however they contain three nested \texttt{while}-loops and thus can serve as an illustration of the differences in the program compilation and unrolling between \porthos[1] and \porthos[2].
-%contains several syntactic elements (such as prefix increment and loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
-%In the Figure, the control-flow subgraph \xgraph[CF] of the non-unrolled event-flow graph is presented on the right-hand side (
+The functions contain three nested \texttt{while}-loops and thus can serve as an illustration of differences in the program compilation and unrolling processes between \porthos[1] and \porthos[2].
 
 \begin{figure}[!h]
-%\begin{minipage}{.53\textwidth}
 \begin{subfigure}[b]{.53\textwidth}\centering
@@
 \end{subfigure}
-%\begin{minipage}{.45\textwidth}
 \begin{subfigure}[b]{.45\textwidth}\centering
@@
 
-The functions are processed by \porthos[2] or modified version \porthos[1] that is able to print same \xgraph[CF] as the new tool (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bind by the method processing its parent).
+The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
 The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
@@
 
-The Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-cf} are compiled to.
+\subsection{Compilation}
+\label{ch:eval:show:compil}
+
+Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-cf} are compiled to.
 The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].
@@
 The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
-The graphs contains a single source event and a single sink event represented by the grey triangles (actually, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
+The graphs contain a single source event and a single sink event represented by the grey triangles (in fact, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
 For clarity, all branching events, that in current example serve as the conditional events of loops, are highlighed with light-grey colour.
@@
 %
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
@@
 %
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
@@
 
-Two compiled the graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, that are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}, and \texttt{skip}-events in \porthos[1] graph.
-However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} are different as \porthos[2] uses different unrolling algorithm.
+Note that two compiled graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, that are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}, and \texttt{skip}-events in \porthos[1] graph.
+However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} are different as the tools use different unrolling algorithms.
 The labels of events in the left-hand side picture (produced by \porthos[2]) are augmented by the unrolling depth number, which is separated from the event label by comma.
+
+\subsection{Unrolling}
+\label{ch:eval:show:unrol}
 
@@
 As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times, which makes the old unrolling algorithm not complete).
-As the new unrolling algorithm uses the Deep-First Search, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus is complete.
+As the new unrolling algorithm is based on the DFS algorithm, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus the program analysis performed on this graph can be complete up to bound~$k$.
 
@@
 
-\begin{figure}[!h]
+\begin{figure}%[!h]
 %
 \begin{subfigure}[b]{.6\textwidth}\centering
-  \includegraphics[height=.6\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
+  \includegraphics[height=.65\textheight,width=1.1\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
   \hfill
@@
 \label{ex:test-both-pic-unroll}
+\end{figure}
+
+Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs on the unrolling bound $k$ for the \texttt{Dekker} test.
+%The left-hand side picture presents the dependency graph for \porthos[2], 
+Both charts are limited above by $420$ events.
+\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
+However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx.
+$70$ states.
+
+\begin{figure}[t]
+\begin{minipage}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-new.png}
+\end{minipage}
+\hfill
+%
+\begin{minipage}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-old.png}
+\end{minipage}
+%
+\caption{Dependency of number of events in the unrolled program on the unrolling bound}
+\label{dep:events}
 \end{figure}
@@
 
-%TODO: here (or better in Chap 5 ) say about 'we tried different encoding schemes !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
 %Also note that comparing to the unrolling bound description given in Section~\ref{ch:impl:proc:x-unroll}, 
@@
 
+For time benchmarking we ran the tool $5$ times and computed the median of the encoding time.
+Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
+The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.
+
 \subsection{State reachability analysis}
@@
 %TO-DO: fix interface
-%To run \porthos[2] in reachability analysis mode, we use the command `\lstinline{mousq
 
 %\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
-When running in the state reachability analysis mode, \porthos[2] produces the following output:
+The sample output of \porthos[2] in the state reachability analysis mode is presented in Appendix~\ref{apx:output}.
+The output includes detailed time measurements for each stage of the analysis, and optionally statistical information about the unrolled program (number of events, transitions, etc.)
 
-\begin{lstlisting}
-$ java PorthosC --reachability --input benchmarks/C11/Dekker.c \
-      --bound 27 --target TSO
-0.288: Interpretation...
-0.426: Unrolling...
-0.533: Program encoding...
-0.701: Program domain encoding...
-#=193 (107)
-34.187: Memory model encoding...
-36.818: Solving...
-{
-  "result": "NonReachable",
-  "mainTimer": {
-    "elapsedTimeSec": 36.163
-  },
-  "timers": {
-    "Interpretation": {
-      "elapsedTimeSec": 0.132
-    },
-    "ProgramDomainEncoding": {
-      "elapsedTimeSec": 32.486
-    },
-    "Solving": {
-      "elapsedTimeSec": 0.345
-    },
-    "Unrolling": {
-      "elapsedTimeSec": 0.061
-    },
-    "MemoryModelEncoding": {
-      "elapsedTimeSec": 2.631
-    },
-    "ProgramEncoding": {
-      "elapsedTimeSec": 0.168
-    }
-  },
-  "errors": []
-}
-\end{lstlisting}
-
-For time benchmarking we ran the tool $5$ times and computed the median of the encoding time.
-Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
-The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.
+\begin{figure}[t]
+\begin{minipage}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-new.png}
+\end{minipage}
+\hfill
+%
+\begin{minipage}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-old.png}
+\end{minipage}
+%
+\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
+\label{dep:time}
+\end{figure}
 
@@
 
+Figure~\ref{dep:time} illustrates the dependency of overall time spent for the program domain encoding on the unrolling bound.
+The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
+
 Note that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
@@
 Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
-
-<TODO: above as table>
 
@@
 
-<TODO: profile both programs, say sth about memory usage!>
+%<TODO: profile both programs, say sth about memory usage!>
 
@@
 For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
-
 %!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
-
-\begin{lstlisting}
-$ java PorthosC --portability --input benchmarks/C11/Dekker.c \
-      --bound 27 --source TSO --target SC --state
-0.359: Interpretation...
-0.364: Unrolling...
-0.590: Unrolling...
-0.623: Program encoding...
-0.847: Program domain encoding...
-#=193 (107)
-#=193 (107)
-66.541: Memory model encoding...
-72.321: Solving...
-{
-  "result": "StatePortable",
-  "iterations": 0,
-  "mainTimer": {
-    "elapsedTimeSec": 73.02
-  },
-  "timers": {
-    "ProgramEncoding": {
-      "elapsedTimeSec": 0.223
-    },
-    "MemoryModelEncoding": {
-      "elapsedTimeSec": 5.779
-    },
-    "Interpretation": {
-      "elapsedTimeSec": 0.248
-    },
-    "Solving": {
-      "elapsedTimeSec": 0.699
-    },
-    "ProgramDomainEncoding": {
-      "elapsedTimeSec": 65.693
-    },
-    "Unrolling": {
-      "elapsedTimeSec": 0.022
-    }
-  },
-  "errors": []
-}
-\end{lstlisting}
-
 As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

### 2018-05-29 90_summary.tex
Cleanup finale: finish impl.
add pictures and stuff.

@@
 \item
-Firstly, we studied existing weak memory model-aware analysis frameworks and tools to gather deep understanding of the problem (Sections~\ref{ch:intro:related} and~\ref{ch:enc:mc} and Chapter~\ref{ch:wmm}).
+We have studied existing weak memory model-aware analysis approaches, tools and frameworks to gather deep understanding of the problem (Sections~\ref{ch:intro:related} and~\ref{ch:enc:mc} and Chapter~\ref{ch:wmm}).
 
 \item
-Then, we explored existing implementation of the portability analysis tool \textit{\porthos{}} and designed the new tool \textit{\porthos[2]} that accepts higher subset of the C language as input, has modular architecture and disposes the extensible knowledge base of the domain-specific functions.
+Then, we explored existing implementation of the portability analysis tool \textit{\porthos{}} and designed the new tool \textit{\porthos[2]} that accepts as input the larger subset of the C language and supports its fundamental concepts by design, has modular architecture and disposes the extensible knowledge base of the domain-specific functions.
 The key aspects of the architectural design and implementations are the following:
-%explored the existing implementation of the proof-of-concept tool \porthos[1] redesigned the new architecture of \porthos[2] with consider
 
   \begin{enumerate}[leftmargin=\parindent,label=\alph*.]
-  \item The first stumbling block in the input language extension was the \porthos[1] input language parser, which performs the full semantic analysis of the program.
-  Although this works for a small subset of the C language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
-  For handling that, we implemented the processing units that operate on the \textit{pre-compilation} stage discussed in Section~\ref{ch:impl:proc:x-pre-compiler}.
+  \item The first stumbling block in extension of the input language was the \porthos[1] program input parser, which performs the full semantic analysis of the program.
+  Although it works for a small subset of the C language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
+  To handle this, we implemented the processing units that analyse the AST of the program on the \textit{pre-compilation} stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
 
-  \item Secondly, \porthos[1] determined the \textit{kind of a variable} (shared or local) syntactically, depending on the operator of the function that uses this variable.
-  This strongly restricts the syntax of the input language, therefore, for processing programs in C, \porthos[2] has an additional pre-compilation phase that traverses AST of the program and collects information of all variables accessed in the program (see Section~\ref{ch:impl:proc:x-pre-compiler}).
+  As an example, the \textit{variable kind analysis} (shared or local) is performed by \porthos[1] syntactically, depending on the operator or the function that uses the variable.
+  In contrast, \porthos[2] traverses the program AST and collects information of all variables accessed in the program during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
   Currently, \porthos[2] considers a variable to be shared if it was declared as a pointer, or its address is accessed at least once by the code of any process, or it was declared as a parameter of the function that defines a process, or it was exported by \texttt{extern} keyword or an exporting function.
-  %This approach gives flexibility in detecting shared variables 
-  %  As this approach does not work for rich languages such as C   
-  %as \porthos[2] determines the \textit{kind (shared or local) of each variable} on the pre-compilation phase, the shared variables can be accessed in arbitrary code location (comparing to \porthos[1] which determines the type of a variable syntactically, depending on the operator of function that uses the variable).
 
-  \item In addition, \porthos[2] supports some \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\lstinline{int a, b=2, z;}'), arbitrary expressions (corresponding to the C standard), and others (see Section~\ref{ch:impl:input}).
+  \item In addition, \porthos[2] supports \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\lstinline{int a, b=2, z;}'), arbitrary expressions allowed by the C11 standard~\cite{c11}, and some others (see Section~\ref{ch:impl:input}).
 
-  \item As the C language supports unconditional \texttt{goto}-jumps that can produce an arbitrary control-flow graph, which cannot be supported solely with the program AST, therefore we needed to compile the AST to the low-level hardware-agnostic program representation \xgraph{}.
+  \item As the C language supports unconditional \texttt{goto}-jumps, the control-flow graph can have an arbitrary structure, which can not be modelled solely by the program AST, therefore \porthos[2] compiles the AST to the low-level hardware-agnostic program representation \xgraph{}.
   For that, we implemented the \textit{full C interpreter} discussed in Section~\ref{ch:impl:proc:x-compiler}.
@@
   \item Originally, the control-flow instructions were encoded directly into the SMT-formula~\cite{Porthos17a}.
-  In contrast, \porthos[2] encodes the low-level \xgraph{} representation, which can have arbitrary graph structure.
-  For encoding the \xgraph{} into the SMT-formula, we apply the \textit{new control-flow encoding scheme} that in general follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} (see Section~\ref{ch:enc:bmc:cf}).
+  In contrast, \porthos[2] encodes the low-level \xgraph{} representation in accordance with the \textit{new control-flow encoding scheme} that in general follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} (see Section~\ref{ch:enc:bmc:cf}).
   As the new encoding scheme does not add new variables to every control-flow instruction, the number of variables in the result SMT-formula is expected to be smaller.
+
+  \item Since \porthos[2] compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from unwinding all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores \textit{all possible executions} that the program can produce within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
+  Although the new unrolling algorithm produces considerably more executions than the old one and thus requires more time for analysis, \porthos[2] uses the new unrolling algorithm by default as it is complete (finds all possible executions).
 
@@
   Invocation hooks are defined in Java and thus are flexible, though their extension and modification requires some knowledge of the internals of the tool.
-  We illustrate this mechanism with the basic support for Linux kernel litmus tests.
-
-  \item Since the new tool compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from the simple unrolling all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores all possible states that the program can reach within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
-  Although the new algorithm produces considerably more events than the old one and thus takes more time for analysis,  \porthos[2] uses the new algorithm by default as it is complete (finds \textit{all} possible executions).
+  We illustrate this mechanism with the basic support for Linux kernel litmus tests (see Sections~\ref{ch:impl:proc:x-compiler} and~\ref{ch:eval:perf:feat}).
 
@@
 \item 
-As the tests show, the overhead of the new architecture is negligible, therefore we consider the applied architectural changes as acceptable.
+As the tests show, the \textit{performance overhead} of the new architecture is negligible (see Section~\ref{ch:eval:perf}), therefore we consider the applied architectural changes as acceptable.
 
 \end{enumerate}
-
-Thus, the new tool \porthos[2] with the extensible program domain knowledge base can be considered as a generalised framework for SMT-based memory model-aware analysis, which can not only perform the reachability and portability analysis, but serve as a basis for other kinds of static analysis of concurrent programs.
-
-%enumerate goals (from Task specification) in past tense + justification
 
@@
 %that performs complete analysis
-%TODO: show it in table , reference here the table
 As it was shown in Section~\ref{ch:eval:perf}, the number of events of the unrolled program grow rapidly as the user increases the unrolling bound.
@@
 
-\item <parallelised>
+%\item <parallelised>
 
-\item <C processor directives + typedef > 
-<TODO: a short summary here>
+%\item <C processor directives + typedef > 
+%<TODO: a short summary here>
 
-\item <contexts for local mem units in mem manager>
+%\item <contexts for local mem units in mem manager>
+
+%\item <as an optimisation, maybe use more comprehensive unrolling algo as in Aho>
+
+%\item In order to be able to assess the A user might want to know
+
+%\item optionally: urnolling algo;
 
 \end{itemize}
+
+The new tool \porthos[2] with the extensible program domain knowledge base can be considered as a generalised framework for SMT-based memory model-aware analysis, which can not only perform the reachability and portability analysis, but also serve as a basis for other kinds of static analysis of concurrent programs.

### 2018-05-30 1_intro.tex
Add example on cfg in cfg to eval. Add two siks to summary.

@@
 Even if all the compiler optimisations were disabled (which is a rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled under the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
+As the result, a program compiled for the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
 

### 2018-05-30 4_impl.tex
Add example on cfg in cfg to eval. Add two siks to summary.

@@
 
-The general processing scheme of old tool \porthos[1] is the following.
+The general processing scheme of the old tool \porthos[1] is the following.
 The input program is parsed to the AST, where leaves are the events.
-The instructions of the AST are then encoded into the $\phi_{CF}$ (see Section~\ref{ch:enc:bmc:cf}) and the events are processed by the single function (`\texttt{dartagnan.wmm.Domain.encode}') that loops over all events, checks their properties (the kind of event, the operands of the memory event, the type of the barrier event, etc.) and adds corresponding relations assertions to the formula.
-Encoding all the relations (the \textit{program domain encoding}) in a single function is error-prone and hardly manageable.
-Moreover, this function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
+%The instructions of the AST are encoded into the $\phi_{CF}$ (see Section~\ref{ch:enc:bmc:cf})
+The control-flow of the program is encoded into an SMT-formula as it was discussed in Section~\ref{ch:enc:bmc:cf}, and the data-flow (\po{}- and \rf{}- relations) and some derived relations are encoded by the single function (`\texttt{dartagnan.wmm.Domain.encode}') that loops over all events, checks their properties (the kind of event, the operands of the memory event, the control-flow properties such as condition level, etc.) and adds the assertions of corresponding relations to the formula.
+Encoding all the relations (the \textit{program domain encoding}) in a single large function is an error-prone and hardly manageable approach.
+Moreover, the events, that constitute the elements of a composite data structure, carry the information about it (such as the condition level, which is the property of the control-flow graph), while it should be carried by the data structure itself.
+%This approach is highly extensible
+In addition, the program domain encoding function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
 In \porthos[2], we moved the encoding logic into a separate module (see Section~\ref{ch:impl:proc:z-encoder}) that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.
@@
 %on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
-Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula, which then is solved by the SMT-solver.
-Although this general architecture is defined for
+Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula, which then is solved by the SMT-solver (note, current implementation of \porthos[2] excludes the \zformula{} as at the moment it is enough for \porthos[2] to use only a single SMT-solver).
+
+%Although the general architecture is designed to separate clearly the internal representations used by the program, we have not yet implemented several modules listed before.
+%For instance, the \xgraph{} is constructed only partially, i.e., only the control-flow graph \xgraph[CF] is built during the program compilation.
+%After this, the graph is unrolled and encoded directly into an SMT-formula, avoiding the \zformula{} abstraction layer as temporarily unnecessary.
+%The data-flow analysis is performed on the stage of SMT-encoding with usage of the SSA-map (see Section~\ref{ch:enc:bmc:df}).
+%Thus, the current implementation of \porthos[2] implies the \xgraphU[CF] to be passed to the program domain encoder of the old code base, which was discussed above.
 
@@
 The \ytree{} statements of follow this rule, however they these labels are symbolic, and they need to be resolved at the pre-compilation stage.
-Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific declaration \texttt{YProcessDefinition}
+Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific definition \texttt{YProcessDefinition}
 used in intra-procedural analysis mode.
 The function definition contains the \texttt{YCompoundStatement} body and the \texttt{YMethodSignature} signature (method name and types of formal parameters), which is used in the function resolution during the pre-compilation stage.
-The other litmus-specific statements are \texttt{YPreludeStatement}
-that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeStatement} 
+The other litmus-specific definitions are \texttt{YPreludeDefinition}
+that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeDefinition} 
 that carries the \texttt{YExpression} binary expression to be asserted by the litmus test.
@@
 \end{outline}
+
+%TODO: check this %Note, as it was above, the data-flow componen \xgraph[DF] of the event-flow graph is not yet constructed
 
@@
 In C, labelled statements are declared via the colon-syntax `\texttt{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\texttt{goto\;<label>}'.
-The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map that points a label to the labelled statement.
+The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map \texttt{JumpsResolver} that points a label to the labelled statement.
 This information is used during compilation to set up unconditional jumps.
@@
 
-The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it in Figure.
+The main class representing the X-compiler is \texttt{Y2XConverter}.
 It receives as input the \ytree{}, the memory model kind and user settings (for instance, the interpreter mode defining the set of invocation hooks enabled during the analysis run).
@@
 
-The invocation hooking module serves as a knowledge base that stores the semantics of functions.
+The invocation hooking module serves as a compiler knowledge base that stores the semantics of high-level functions in terms of low-level event-flow graph elements.
+%Invocation hooks should be used for describing the 
+For example, the C11 sequentially consistent load operation `\lstinline{l.store(memory_order_seq_cst, r)}' whould be compiled for the \texttt{x86} architecture as the write instruction `\lstinline{MOV l, r}' followed by the fence instruction `\lstinline{MFENCE}', whether for the \texttt{POWER} architecture this code would be compiled as the `\lstinline{hwsync}' followed by the store instruction.
+This is called the \textit{compiler mapping}.
+While the \xgraph{} is the abstract (hardware-agnostic) low-level program representation, the compiler mapping is the only component of the compiler that produces differences between two versions of the same program compiled for different architectures.
+
+The function invocation mechanism works as follows.
 Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in the case of program in C language -- only the function name) across all signatures it stores.
@@
 The \rf{}-edges join each write event to all read events that access the same location; the \co{}-edges join write events to the same location.
-Currently, \porthos[2] does not yet construct the \xgraph[DF] component of the full \xgraph{}, instead, it uses the function `\texttt{dartagnan.wmm.Domain.encode}' of \porthos[1] to encode data-flow relations and derived relations of the \xgraph[CF].
+%TODO: check if this is true %Currently, \porthos[2] does not yet construct the \xgraph[DF] component of the full \xgraph{}, instead, it uses the function `\texttt{dartagnan.wmm.Domain.encode}' of \porthos[1] to encode data-flow relations and derived relations of the \xgraph[CF].
 

### 2018-05-30 5_eval.tex
Add example on cfg in cfg to eval. Add two siks to summary.

@@
 The unrolling algorithm used by \porthos[1] (right-hand side picture) unrolls \textit{all} loops $k$ times (where $k$ is the unrolling bound), and the unrolling algorithm of \porthos[2] unrolls loops so that not more than $k$ events are executed.
-As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times, which makes the old unrolling algorithm not complete).
+As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times in a raw, which makes the old unrolling algorithm not complete).
 As the new unrolling algorithm is based on the DFS algorithm, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus the program analysis performed on this graph can be complete up to bound~$k$.
@@
 
+Current section demonstrates some new features of \porthos[2], which is hard or impossible to implement on the top of its predecessor \porthos[1] (see detailed discussion in Chapter~\ref{ch:impl}).
+
+\subsection{Interpretation of a code with an arbitrary control-flow}
+
+The following example in Figure~\ref{ex:cfg-arbitrary} illustrates the interpretation of the program that contains \textit{continue}, \textit{break} and two \texttt{goto} instructions (both forward and backward), that can produce an arbitrary control-flow graph.
+As an addition, the example illustrates the interpretation of complex expressions that involve shared variables (such as the condition of the \texttt{while}-loop in the function \texttt{thread\_0}).
+Note that the \texttt{entry event} of the loop is not necessarily the guard computation event.
+In the case when a computation event involves shared variables, all of them need to be copied to a register before the computation event is emitted.
+The first event that performs such a copy becomes an entry event of the \texttt{while}-loop, which the loop back edge and all \texttt{continue}-jumps point to.
+
+%Note also that if the \texttt{backward} \texttt{goto}-edges (to the already processed event) can be set up at the moment of emitting the jump event, the \texttt{forward} \texttt{goto}-edges can be set up later on
+
+\begin{figure}[!h]
+\begin{lstlisting}
+{
+  int x = 1, *y;
+}
+
+void thread_0(int &x, int &y) {
+  L0: x = 0;
+  int r;
+  while (x * (5 + 4 / 2) % 3 == 1) {
+    if (x != 0) {
+      goto L0; //backward jump
+    }
+    if (y > 6) {
+      continue;
+    }
+    else if (++y > 7) {
+      r = r + 10;
+      break;
+    }
+    else {
+      goto L1; //forward jump
+    }
+    r = 11;
+  }
+  y = x + 1;
+  L1: x = r;
+}
+
+void thread_1() {
+  while (true) {
+    if (x > 7)
+      break;
+  }
+  y = x;
+}
+
+exists (y == x + 1 && thread_0:r > 21)
+
+\end{lstlisting}
+\caption{Example of interpretation of a code with an arbitrary control-flow}
+\label{ex:cfg-arbitrary}
+\end{figure}
+
+The four flow-graphs for the code in Figure~\ref{ex:cfg-arbitrary} are shown in Figure~\ref{ex:cfg-arbitrary:pic}.
+There are two small linear flow-graphs for \textit{prelude} (litmus-specific variables initialisation and shared variables declaration) and \textit{postlude} (assertion) definitions.
+As the initialisation statement assigns value only to one shared variable, its flow-graph has only a single event (although variable declarations change the state of the interpreter while compilation, they do not produce events).
+The postlude has shared variables involved into the assertion, therefore they need to be copied into local registers beforehand.
+Note how the unconditional jump instructions \texttt{goto}, \texttt{continue} and \texttt{break} are compiled in the event-flow graph for processes \texttt{thread\_0} and \texttt{thread\_1}.
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/cfg-arbitrary/collage.jpg}
+\caption{Control-flow components of the event-flow graphs for the code in Figure~\ref{ex:cfg-arbitrary}}
+\label{ex:cfg-arbitrary:pic}
+\end{figure}
+
+\subsection{Extensible compiler mapping}
+
 %The revised design of \porthos{} tool includes the following new features:
@@
 
-If verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.
+When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

### 2018-05-30 90_summary.tex
Add example on cfg in cfg to eval. Add two siks to summary.

@@
 During the work on this thesis, we solved the following engineering and scientific tasks:
+
+%TODO: written a generic hardware-agnostic extensible non-optimising compiler
 
@@
 As it was shown in Section~\ref{ch:eval:perf}, the number of events of the unrolled program grow rapidly as the user increases the unrolling bound.
-One possible way to reduce the number of the program states might be applying some traditional techniques that target the state explosion problem (such as concrete execution as a part of concolic execution~\cite{majumdar2007hybrid} before the unrolling stage).
+One possible way to reduce the number of the program states might be applying some traditional techniques that target the state explosion problem (such as concrete execution as a part of \textit{concolic execution}~\cite{majumdar2007hybrid} before the unrolling stage).
 However, this must be done carefully and with taking into account the analysing weak memory model, because otherwise it can lead to the loss of states and thus to incorrect analysis results.
@@
 
-\item \textit{Extending the knowledge base} of domain-specific functions to model synchronisation primitives can be considered as the main future research direction.
+\item For justifying the correctness of the verification that bases on the result of the input program unrolling, one might want to be aware whether an execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
+This can be modelled with two kinds of sink events (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.
+
+\item \textit{Extending the knowledge base} of domain-specific functions to model synchronisation primitives (the compiler mapping) can be considered as the main future research direction.
 Thus, to be able to process most Linux kernel litmus tests, \porthos[2] needs to know the semantics of the barrier and memory operation functions that are specific for the Linux kernel in order to be able to compile them into the \xgraph{} and later encode them into the SMT-formula.
@@
 
-\item Secondly, the wide range of existing litmus tests may require \porthos[2] to \textit{extend the support for input languages}.
+\item The wide range of existing litmus tests may require \porthos[2] to \textit{support new input languages}.
 The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, POWER, Alpha, etc.).
@@
 \item One might want to compare the performance of \porthos[2] in cases of usage the different SMT-solvers.
-Due to the \zformula{} abstraction layer, new SMT-solvers may be easily added (the general way to support multiple SMT-solvers is to convert the \zformula{} to an SMT-LIB formula~\cite{smt-lib}, that can be passed as input to the SMT-solver run as an external process).
+For that, the \zformula{} abstraction layer will be useful as new SMT-solvers can be easily added (the general way to support multiple SMT-solvers is to convert the \zformula{} to an SMT-LIB formula~\cite{smt-lib}, that can be passed as input to the SMT-solver run as an external process).
 

### 2018-05-30 inc/lst/XInterpreter.java

@@ -9,6 +9,8 @@ public interface XInterpreter {
   XExitEvent emitExitEvent();
   XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
   XJumpEvent emitJumpEvent();
+  XJumpEvent emitJumpEvent(String label);
+  void markNextEventLabel(String label);
   XNopEvent emitNopEvent();
   XAssertionEvent emitAssertionEvent(XBinaryComputationEvent assertion);
   

### 2018-05-30 5_eval.tex
Complete everything. Left only polishing.

@@
 
-Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs on the unrolling bound $k$ for the \texttt{Dekker} test.
-%The left-hand side picture presents the dependency graph for \porthos[2], 
-Both charts are limited above by $420$ events.
-\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
-However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx.
-$70$ states.
+% ---------------------------
 
 \begin{figure}[t]
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-new.png}
-\end{minipage}
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/new-e.png}
+\caption{The graph unrolled by \porthos[2]}
+\label{dep:events:new}
+\end{subfigure}
 \hfill
 %
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-old.png}
-\end{minipage}
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/old-e.png}
+\caption{The graph unrolled by \porthos[1]}
+\label{dep:events:old}
+\end{subfigure}
 %
-\caption{Dependency of number of events in the unrolled program on the unrolling bound}
+\caption{Dependency of number of events in the unrolled Dekker's algorithm on the unrolling bound}
 \label{dep:events}
 \end{figure}
+
+As another example of program liable for analysis consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
+For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
+For performing tests, \porthos[2] was slightly modified in the branch \textit{ay/test-timing}%
+\footnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
+%
+for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script-analyser (commit 48ba7161).
+The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
+\footnote{\porthos[2] repository: \url{https://github.com/ajuszkowski/PorthosC}}. %
+%
+Both \porthos[1] and \porthos[2] are run in the state reachability mode for \texttt{TSO} memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).
+
+Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs of the thread \texttt{t0} on the unrolling bound $k$ for the \texttt{Dekker} test (the left-hand side picture illustrates the case for \porthos[2], and the right-hand side picture illustrates the case for \porthos{}).
+The solid lines denote dependency of number of unrolled events on the unrolling bound, when the dashed lines denote the number of events in non-unrolled graphs.
+%Both charts have limits above $420$ events.
+%\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
+%However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx.
+$70$ states.
+The blue line depicts the dynamics of the overall number of events in the graph (including memory events, nop events, etc.).
+Initially, the thread \texttt{t0} has 22 events.
+As in the unrolling scheme of \porthos[1], each loop is executed exactly $k_1$ times, therefore if $k_1=6$, the loop has been executed at least 6 times by producing around $300$ events.
+Nonetheless, in the new unrolling scheme used by \porthos[2], the same number of events is produced by the unrolling bound $k_2=34$, which contains approximately $1.5$ executions of the $22$-event loop.
+Therefore, according to rough estimates the new encoding scheme produces $6/1.5=4$ times more events than the older one.
+
+The same dynamics can be seen with shared memory events (yellow lines).
+However, the new algorithm produces much more local memory events (green line on the left-hand side picture), which grow exponentially with the unrolling bound, when the number of local memory events produced by the old algorithm (green line in the right-hand side picture) grows relatively slowly.
+This can be explained by the need to copy all shared variables to local ones when performing computation over them.
+On the other hand, the right-hand side picture reveals exponential-like growth of the number of \texttt{nop}- (or \texttt{skip}-) events, that are necessary for implementing the control-flow as an AST, however constitute a redundant part of the model (in the new unrolling scheme, the number of \texttt{nop}-events remains to be relatively constant).
+
+As the new unrolled graph has all possible executions unlike the older one, by default \porthos[2] uses the new unrolling scheme.
+However, in some applications the older unrolling mode may be useful (for example, where the user can use other instruments to prove the limitations on the number of executions of a loop), thus the user should be able to change configure the unrolling algorithm (support for the older algorithm is left to a future work).
 
@@
 
-As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
-For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
-
-For time benchmarking we ran the tool $5$ times and computed the median of the encoding time.
+Performance evaluation has also been performed on the example of the \texttt{Dekker} test.
+%As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
+%For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
+For time benchmarking we ran the tools $5$ times and computed the median of the encoding time.
 Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
@@
 
-%\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
 The sample output of \porthos[2] in the state reachability analysis mode is presented in Appendix~\ref{apx:output}.
@@
 \begin{figure}[t]
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-new.png}
-\end{minipage}
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/new-t.png}
+\caption{Encoding of the graph unrolled by \porthos[2]}
+\label{dep:time:new}
+\end{subfigure}
 \hfill
 %
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-old.png}
-\end{minipage}
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/old-t.png}
+\caption{Encoding of the graph unrolled by \porthos[1]}
+\label{dep:time:old}
+\end{subfigure}
 %
@@
 
-%For \porthos[1] we specify the unrolling bound $k_1 = 2$, and for \porthos[2] the bound is $k_2 = 21$ ().
-As the unrolling algorithm has been changed, the number of events after unrolling differs considerably.
-Therefore, the correct performance comparison with \porthos[1] is not manageable as the performance of the tools depends directly on number of events.
-For example, with the bound $k_1=2$ \porthos[1] produces $59$ events (among them, $51$ shared-memory events) with the program domain encoding time 4.608 sec,
-and with the bound $k_1=3$ \porthos[1] produces $95$ events (among them, $107$ shared-memory events) with the program domain encoding time 21.475 sec;
-whereas \porthos[2] with the bound $k_2=17$ produces $57$ ($31$) events with the program domain encoding time 1.734, and with the bound $k_2=27$ it produces $193$ events (among them, $107$ shared-memory events, which equal to the number of shared-memory events produced by \porthos[1] with the bound $k_1=3$) with the time 32.673 sec.
-Therefore, the new unrolling scheme has made the analysis complete but at the same time it reduced the performance considerably.
+As the unrolling algorithm has been changed, the number of events after unrolling differs considerably, therefore the correct performance comparison with \porthos[1] is not manageable as the performance of the tools depends directly on number of events.
+
+%For example, with the bound $k_1=2$ \porthos[1] produces $59$ events (among them, $51$ shared-memory events) with the program domain encoding time 4.608 sec,
+%and with the bound $k_1=3$ \porthos[1] produces $95$ events (among them, $107$ shared-memory events) with the program domain encoding time 21.475 sec;
+%whereas \porthos[2] with the bound $k_2=17$ produces $57$ ($31$) events with the program domain encoding time 1.734, and with the bound $k_2=27$ it produces $193$ events (among them, $107$ shared-memory events, which equal to the number of shared-memory events produced by \porthos[1] with the bound $k_1=3$) with the time 32.673 sec.
+%Therefore, the new unrolling scheme has made the analysis complete but at the same time it reduced the performance considerably.
+
 %The last result shows that with the bounds $k_1=3$ and $k_2=27$ both tools have executed same number of loops  almost equivalent as they
@@
 
-Figure~\ref{dep:time} illustrates the dependency of overall time spent for the program domain encoding on the unrolling bound.
-The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
+Figure~\ref{dep:time} illustrates the dependency of the time spent for the program domain encoding (for the process \texttt{t0}) on the unrolling bound.
+Since the program domain encoding time is heavily dependent on the number of events (specifically, shared and local memory events as the encoding function contains several multiple nested loops over memory events), the dependency graph follows the results presented in Figure~\ref{dep:events} and resembles exponential relationship between encoding time and unrolling bound.
 
-Note that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
+%The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
+An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part "\texttt{timers}" of the output).
+Note that the time spent for the interpretation (0.222 sec) and unrolling (0.088 sec) stages remains negligible comparing to the full execution time (40.679 sec).
 Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
-The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
-The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
-Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
+%The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] is 0.397 sec, and in case of \porthos[1] the time was 1.291 sec.
+%The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
+%Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
 
@@
 
-For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
+For evaluating the tool working in the portability analysis mode, we used the same files \texttt{Dekker.c} and \texttt{Dekker.pts} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
 %!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!

### 2018-05-30 90_summary.tex
Complete everything. Left only polishing.

@@
   \item Since \porthos[2] compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from unwinding all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores \textit{all possible executions} that the program can produce within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
+  On the other hand, it is shown in Section~\ref{ch:eval:show:unrol} that the implementation of the new unrolling algorithm does not produce exponentially many dummy \texttt{nop}-events as the implementation of the old unrolling algorithm does.
+  Another benefit of the new unrolling scheme is that it is much more configurable (i.e., the number of events grows gradually as the unrolling bound is being increased, see Figure~\ref{dep:events}).
+\\
   Although the new unrolling algorithm produces considerably more executions than the old one and thus requires more time for analysis, \porthos[2] uses the new unrolling algorithm by default as it is complete (finds all possible executions).
@@
 
-\item For justifying the correctness of the verification that bases on the result of the input program unrolling, one might want to be aware whether an execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
+\item For justifying the correctness of the verification that bases on the result of the input program unrolling (via the new unrolling scheme discussed in Section~\ref{ch:impl:proc:x-unroll}), one might want to be aware whether the execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
 This can be modelled with two kinds of sink events (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.

### 2018-05-30 1_intro.tex
Complete everything. Left only polishing.
Add Acknowledgments.

@@
 %Leaving aside questions on the performance improvement of the system parallelisation, we explore the questions of the \textit{memory consistency},
-We are discussing mostly the questions on the \textit{memory consistency} rather than performance benefits of the concurrency as the inconsistent memory operations can create new program behaviours, unexpected from the programmer's point of view (as well as the loss of expected behaviours), which all can be considered as a security flaw.
+This thesis discusses mostly the questions on the \textit{memory consistency} rather than performance benefits of the concurrency since inconsistent memory operations can create new program behaviours, unexpected from the programmer's point of view (as well as the loss of expected behaviours), which all can be considered as a security flaw.
 

### 2018-05-30 1_intro.tex
Final

@@
 As an example, consider two x86 assembly programs in Figure~\ref{simple_wmm_x86} (such little code examples that explain specific behaviour of the concurrent execution environment are called \textit{litmus tests}).
-Two programs are the same, except the right-hand program uses the synchronisation instruction \texttt{MFENCE} after the store to the shared memory.
+Two programs are the same, except the right-hand side program uses the synchronisation instruction \texttt{MFENCE} after the store to the shared memory.
 In both programs, the process~\texttt{P0} writes the value~\texttt{1} to the shared variable~\texttt{x} and reads a value of the shared variable~\texttt{y}, and the process~\texttt{P1} writes to the~\texttt{y} and reads the~\texttt{x}.
@@
 \hline
-\multicolumn{2}{|l|}{exists (0:EAX=0,\,1:EBX=0)} \tabularnewline
+\multicolumn{2}{|l|}{exists (P0:EAX=0,\,P1:EBX=0)} \tabularnewline
 \hline
@@
 %
-\caption{Store buffering~(SB): A litmus test illustrating the write-read reordering allowed by the \texttt{x86-TSO} memory model}
+\caption{Store buffering~(SB): A litmus test illustrating the write-read reordering allowed by the x86-TSO memory model}
 \label{simple_wmm_x86}
@@
 \begin{itemize}[noitemsep,topsep=0pt]
-\item \texttt{(0:EAX=0,~1:EBX=1)},
-\item \texttt{(0:EAX=1,~1:EBX=0)},
-\item \texttt{(0:EAX=1,~1:EBX=1)}.
+\item \texttt{(P0:EAX=0,~P1:EBX=1)},
+\item \texttt{(P0:EAX=1,~P1:EBX=0)},
+\item \texttt{(P0:EAX=1,~P1:EBX=1)}.
 \end{itemize}
-%As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' to be unreachable.
-However, in the left-hand (non-synchronised) litmus test, the state `\texttt{(0:EAX=0,~1:EAX=0)}' is also allowed on the x86 architecture as its processors may cache the writes to the shared memory into their local \textit{store buffers} so that the updated value does not immediately become visible by processes running on other cores.
+%As the program specifies the write to~\texttt{x} to precede the read of~\texttt{y} in process~\texttt{P0}, one could expect the final state `\texttt{(P0:EAX=0~/\textbackslash~P1:EAX=0)}' to be unreachable.
+However, the non-synchronised litmus test in Figure~\ref{simple_wmm_x86:nosync} allows the state `\texttt{(P0:EAX=0,~P1:EAX=0)}' on the x86 architecture as x86 processors may cache the writes to the shared memory into their local \textit{store buffers} so that the updated value does not immediately become visible by processes running on other cores.
 This behaviour is known as \textit{Store Buffering~(SB)}.
@@
 In addition to developing the theoretical basis, researchers work on extracting the memory models for hardware architectures from existing implementations or from the specifications, which are written in natural language and thus suffer from ambiguities and incompleteness.
-Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as \texttt{x86-TSO} and \texttt{Sparc-TSO} (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
+Over the last decade, memory models have been defined for most mainstream multiprocessor architectures, such as x86-TSO and Sparc-TSO (for \textit{Total Store Order}) model for x86 and Sparc architectures~\cite{owens2009better}, much more relaxed memory model for Power and ARM architectures~\cite{alglave2009semantics,sarkar2011understanding, alglave2014herding}, etc.
 %, or Alpha architecture~\cite{alphaWMM}~, and others.

### 2018-05-30 2_wmm.tex
Final

--- 2018-05-26 2_wmm.tex
@@
 The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
-We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\texttt{load}(r_1~\leftarrow~l_2); \ e_2~=~\texttt{store}(l_1~\leftarrow~r_1)$.
+We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1\,=\,{\texttt{load}(r_1\,\leftarrow\,l_2)}; \ e_2\,=\,{\texttt{store}(l_1\,\leftarrow\,r_1)}$.
 Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.
@@
   \includegraphics[width=1.2\linewidth]{img/my/sb-example/SB-dot.png}
-  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
+  \caption{Final state: \texttt{(P0:EAX=1,~P1:EAX=1)}}
   \label{simple_wmm_x86_pic:sub1}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-2.png}
-  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=0)}}
+  \caption{Final state: \texttt{(P0:EAX=1,~P1:EAX=0)}}
   \label{simple_wmm_x86_pic:sub2}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-3.png}
-  \caption{Final state: \texttt{(0:EAX=1,~1:EAX=1)}}
+  \caption{Final state: \texttt{(P0:EAX=1,~P1:EAX=1)}}
   \label{simple_wmm_x86_pic:sub3}
@@
   \includegraphics[width=.9\linewidth]{img/my/sb-example/SB-dot-4.png}
-  \caption{Final state: \texttt{(0:EAX=0,~1:EAX=0)}}
+  \caption{Final state: \texttt{(P0:EAX=0,~P1:EAX=0)}}
   \label{simple_wmm_x86_pic:sub4}

### 2018-05-30 3_encod.tex
Cleanup finale up to Xevent.
Final

@@
 At present time, satisfiability solvers are behind the numerous verification and bounded model checking tools.
-Although, as most modern programming languages are typed, the SAT-formula that encodes the constraints of a program must include non-boolean constraints encoded as boolean variables, which is ineffective, therefore most software analysis and verification tools use \textit{Satisfiability Modulo Theories~(SMT)} -solvers, that integrate first-order reasoning with the background theory reasoning~\cite{kroening2008decision}.
+Although, as most modern programming languages are typed, a SAT-formula that encodes the constraints of a program must include non-boolean constraints encoded as boolean variables, which is ineffective, therefore most software analysis and verification tools use \textit{Satisfiability Modulo Theories~(SMT)} solvers, that integrate first-order reasoning with the background theory reasoning~\cite{kroening2008decision}.
 For example, the Integer theory includes operators `$+$', `$-$', `$\geqslant$' and others).
 Most modern SMT-solvers support integer arithmetic, real arithmetic, fixed-size bit-vectors, arrays, and other theories.
-The direction of the development of SMT-solvers is provided by the \texttt{SMT-LIB} standard~\cite{smt-lib}, which provides definition of the unified language for defining SMT-formulas, and description of most common underlying logics and theories.
+The direction of the development of SMT-solvers is provided by the \textit{SMT-LIB standard}~\cite{smt-lib}, which provides definition of the unified language for defining SMT-formulas, and description of most common underlying logics and theories.
 
@@
 The SSA form requires each variable to be assigned only once within entire program.
-In contrast, \porthos{} used the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique only within a control-flow branch.
+In contrast, \porthos[1] uses the \textit{Dynamic Single-Assignment~(DSA)} form, that requires indices to be unique only within a control-flow branch.
 Although the number of variable references (each of which is encoded as unique SMT-variable) on average is logarithmically less in the case of the DSA form than the SSA form, the result SMT-formula still needs to be complemented by same number of equality assertions when encoding the data-flow of merge points~\cite{Porthos17a}.

### 2018-05-30 4_impl.tex
Final

@@
 ), we consider the accomplished work as a step towards it.
+The repository for \porthos[2] project is \url{https://github.com/ajuszkowski/PorthosC}.
 
@@
 
-As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
+As its predecessor, \porthos[2] is written in \textit{Java}, firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the concepts of Object-Oriented Programming~(OOP) inherent in Java suitable for modelling the programming languages.
+Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver \tool{Z3}%
 %
 \footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}} % 
+by Microsoft Research~\cite{de2008z3}, which is invoked by \porthos[2] via a Java API.
 %
-by Microsoft Research~\cite{de2008z3}.
-However, unlike its predecessor, \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT-solver.
-
-As its predecessor, \porthos[2] is written in \textit{Java}, firstly, in order to be able to reuse some parts and concepts of \porthos{} also written in Java, and secondly, because the authors find the concepts of Object-Oriented Programming~(OOP) inherent in Java suitable for modelling the programming languages.
-Although Java does not show the best results in performance benchmarks (for example, compared to C++~\cite{hundt2011loop}), the performance cornerstone of \porthos[2] (as well as any other SMT-based code analyser) is the phase of solving the SMT-formula, which is left to the third-party SMT-solver invoked by \porthos[2] via a Java API.
+%However, unlike its predecessor, \porthos[2] has an additional abstraction level, \zformula{} (see Section~\ref{ch:impl:model:zformula}), that allows the use any other SMT-solver.
 However, considering the perspective of using \porthos[2] as a static analyser for real-world programs, the memory optimisation problem must also be taken into account during both encoding and solving stages.
-It is worth noting that, for the reasons of simplicity, \porthos[2] is not a concurrent program, however, we believe that due to its modular architecture it can be easily parallelised on the level of program modules.
+For the reasons of simplicity, \porthos[2] is not a concurrent program, however, we believe that due to its modular architecture it can be easily parallelised on the level of program modules if necessary.
 
@@
 Also, the tool can be extended to detect the concurrent parts of the code automatically at the pre-compilation stage.
-For that, the pre-compiler should recognise functions that commence a new process (such as \texttt{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
+For that, the pre-compiler should recognise functions that commence a new process (such as \lstinline{pthread\_create} from \texttt{pthread.h}) and resolve the argument that points to the thread function.
 This functionality is left beyond current thesis.
@@
 As the reader might have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, different kinds of memory operations of of the \oldporthos{} input language vary syntactically as well.
-For example, the assignment of a local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation is denoted by \lstinline{<:-}', atomic non-relaxed store operation is denoted by `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} is resolved syntactically by matching the function name and arguments.
+For example, the assignment of a local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation is denoted by `\lstinline{<:-}', atomic non-relaxed store operation is denoted by `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} is resolved syntactically by matching the function name and arguments.
 Moreover, only the operator `\lstinline{<-}' accepts an expression as the source of data, which means that expressions could be assigned only to registers.
@@
 %This approach is highly extensible
-In addition, the program domain encoding function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
+In addition, the program domain encoding function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the TSO memory model does not contain Power-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
 In \porthos[2], we moved the encoding logic into a separate module (see Section~\ref{ch:impl:proc:z-encoder}) that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.
@@
 The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
-However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
+However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (SC, TSO, PSO, RMO, Alpha, Power, ARM; this feature is inherited from \porthos[1]).
 The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
@@
 The label resolution is necessary for establishing links to labelled statements.
-In C, labelled statements are declared via the colon-syntax `\texttt{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\texttt{goto\;<label>}'.
+In C, labelled statements are declared via the colon-syntax `\lstinline{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\lstinline{goto\;<label>}'.
 The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map \texttt{JumpsResolver} that points a label to the labelled statement.
@@
 \item a variable declared as a parameter of the process function; and
-\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL} also export symbols for dynamic linking~\cite{argawal16kernel}).
+\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \lstinline{EXPORT_SYMBOL} and \lstinline{EXPORT_SYMBOL_GPL} also export symbols for dynamic linking~\cite{argawal16kernel}).
 \end{itemize}
@@
 At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
-Following the C standard, local memory units have higher priority over global ones (the method \texttt{getDeclaredUnitOrNull} returns the first memory unit found, either a global one or a local one, or null if no memory units with requested name have been registered).
+Following the C standard, local memory units have higher priority over global ones (the method \lstinline{getDeclaredUnitOrNull} returns the first memory unit found, either a global one or a local one, or null if no memory units with requested name have been registered).
 Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} should carry the stack of block contexts for local variables (which is not currently implemented).
@@
 %Invocation hooks should be used for describing the 
-For example, the C11 sequentially consistent load operation `\lstinline{l.store(memory_order_seq_cst, r)}' whould be compiled for the \texttt{x86} architecture as the write instruction `\lstinline{MOV l, r}' followed by the fence instruction `\lstinline{MFENCE}', whether for the \texttt{POWER} architecture this code would be compiled as the `\lstinline{hwsync}' followed by the store instruction.
+For example, the C11 sequentially consistent load operation `\lstinline{l.store(memory_order_seq_cst, r)}' whould be compiled for the \texttt{x86} architecture as the write instruction `\lstinline{MOV l, r}' followed by the fence instruction `\lstinline{MFENCE}', whether for the Power architecture this code would be compiled as the `\lstinline{hwsync}' followed by the store instruction.
 This is called the \textit{compiler mapping}.
@@
 An invocation hook can model any type of operations (memory, fence, computational, etc.).
-For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \texttt{READ\_ONCE}, \texttt{WRITE\_ONCE}, and \texttt{ACCESS\_ONCE} may be modelled as volatile \texttt{memory\_order\_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
+For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
+For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
 
@@
 X-interpreter) and change the state of interpreter considering the newly created event.
-Note that the methods \texttt{createComputationEvent} do not emit a computation event but only create one.
+Note that the methods \lstinline{createComputationEvent} do not emit a computation event but only create one.
 This is an optimisation that removes unused computations from the model
-(otherwise, for instance, the assignment `\texttt{x = 1 * (2 + 3)}' would be compiled into two consequent events `\texttt{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\texttt{write(register <- eval(1 * eval(2 + 3)));}').
-Note, if the computation event has not been used at all (for example, in the following C code: `\texttt{foo(); 1; bar();}' the execution event `\texttt{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph}).
+(otherwise, for instance, the assignment `\lstinline{x = 1 * (2 + 3)}' would be compiled into two consequent events `\lstinline{eval(2 + 3); write(register <- eval(1 * eval(2 + 3)));}' instead of a single event `\lstinline{write(register <- eval(1 * eval(2 + 3)));}').
+Note, if the computation event has not been used at all (for example, in the following C code: `\lstinline{foo(); 1; bar();}' the execution event `\lstinline{eval(1)}' is skipped by the model), it is also removed from the event-graph (see justification in Section~\ref{ch:impl:model:xgraph}).
 The second group of X-interpreter methods consists of the methods for defining non-linear statements (branchings and loops).
@@
 Also, during the unrolling each next event increments the \textit{unrolling depth counter} -- the non-zero integer \textit{event reference-ID} stored by each X-event (only events of a non-unrolled graph can have the zero-valued reference-ID)
-As an X-event is immutable, it has the method `\texttt{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
-The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
+As an X-event is immutable, it has the method `\lstinline{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
+The methods \lstinline{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
 

### 2018-05-30 5_eval.tex
Add Acknowledgments.
Final

@@
 
-Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
+Consider two equivalent functions \texttt{t0} in Figure~\ref{ex:test-both-cf}: in the C language (left) and in the \porthos[1] input language (right).
 The functions contain three nested \texttt{while}-loops and thus can serve as an illustration of differences in the program compilation and unrolling processes between \porthos[1] and \porthos[2].
@@
 
-The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
+The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the branching statement \texttt{if-then-else}, the \texttt{while} loop and the \texttt{sequence} statement are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
 The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
@@
 %
-for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script-analyser (commit 48ba7161).
+for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script that analyses and prints the results (commit 48ba7161).
 The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
@@
 %
-Both \porthos[1] and \porthos[2] are run in the state reachability mode for \texttt{TSO} memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).
+Both \porthos[1] and \porthos[2] are run in the state reachability mode for TSO memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).
 
@@
 %The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
-An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part "\texttt{timers}" of the output).
+An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part `\texttt{timers}' of the output).
 Note that the time spent for the interpretation (0.222 sec) and unrolling (0.088 sec) stages remains negligible comparing to the full execution time (40.679 sec).
@@
 %!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
-As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).
+As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 72.744 sec with unrolling bound $k_2=27$ is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).
 
@@
 \end{lstlisting}
-\caption{Example of interpretation of a code with an arbitrary control-flow}
+\caption{Example: A litmus test in C with an arbitrary control-flow}
 \label{ex:cfg-arbitrary}
@@
 
-%The revised design of \porthos{} tool includes the following new features:
-Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
-Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \texttt{READ\_ONCE} and \texttt{WRITE\_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
-Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
-
-As an example, consider the SB litmus test for the Linux kernel (which is similar to the one presented in Figure~\ref{simple_wmm_x86}):
+\begin{figure}[t]
 \begin{lstlisting}
@@
 \end{lstlisting}
+\caption{SB litmus test for the Linux kernel memory model}
+\label{ex:sb:kernel}
+\end{figure}
 
+%The revised design of \porthos{} tool includes the following new features:
+Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
+Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
+Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
+
+As an example, consider the Store Buffering litmus test for the Linux kernel in Figure~\ref{ex:sb:kernel} (which is similar to the one presented in Figure~\ref{simple_wmm_x86}).
 When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

### 2018-05-30 90_summary.tex
Final

@@
 \item The wide range of existing litmus tests may require \porthos[2] to \textit{support new input languages}.
-The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, POWER, Alpha, etc.).
+The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, Power, Alpha, etc.).
 Note that the programs in low-level languages can be easily compiled to the \xgraph{} representation by existing compiler architecture since low-level assembly-like languages can be considered as the subset of the C language, that restricts the set of non-linear control-flow instructions to the conditional and unconditional jumps (which are supported by the current \xgraph{} compiler of \porthos[2]).

### 2018-06-06 2_wmm.tex
Consider final Professor's comments.

@@
 Functionally, a fence may be a synchronisation barrier or a instruction for flushing memory caches into the main memory, etc.
-For instance, the \texttt{mfence} instruction of the x86 assembly flushes the store buffers of the thread, and thus does not allow the \rf-relation (see Section~\ref{ch:wmm:model:relations}) to point to
+For instance, the \texttt{mfence} instruction of the x86 assembly flushes the store buffers of the thread, and thus does not allow the \rf-relation to hold (see Section~\ref{ch:wmm:model:relations}).
 

### 2018-06-06 4_impl.tex
Consider final Professor's comments.

@@
       \item modular architecture: each module can be tested independently,
-      \item usage of software design patterns where necessary, and
-      \item usage of immutable data structures for all DTOs.
+      \item use of software design patterns where necessary, and
+      \item use of immutable data structures for all DTOs.
     \end{enumerate}
@@
 Following the C standard, local memory units have higher priority over global ones (the method \lstinline{getDeclaredUnitOrNull} returns the first memory unit found, either a global one or a local one, or null if no memory units with requested name have been registered).
-Since C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} should carry the stack of block contexts for local variables (which is not currently implemented).
+Since C language allows use of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} should carry the stack of block contexts for local variables (which is not currently implemented).
 
@@
 For the sake of simplicity of the interpretation and encoding, jump events (no-computation events) are present in the flow-graph, however they can be removed from it with rebinding ingoing and outgoing edges.
-As an example of usage the interpreter, consider the code of the \texttt{Y2XConverterVisitor} for processing \texttt{if-then-else} statement of the \ytree{} in Appendix~\ref{apx:y2x}.
+As an example of usage of the interpreter, consider the code of the \texttt{Y2XConverterVisitor} for processing \texttt{if-then-else} statement of the \ytree{} in Appendix~\ref{apx:y2x}.
 
@@
 As an X-event is immutable, it has the method `\lstinline{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
-The methods \lstinline{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
+The methods \lstinline{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the use of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).
 

### 2018-06-06 90_summary.tex
Consider final Professor's comments.

@@
 \item 
-As the tests show, the \textit{performance overhead} of the new architecture is negligible (see Section~\ref{ch:eval:perf}), therefore we consider the applied architectural changes as acceptable.
+As the tests show, the \textit{performance overhead} of the new architecture is negligible (see Section~\ref{ch:eval:perf}), therefore we consider the applied architectural decisions to be acceptable.
 

### 2018-06-06 inc/lst/XInterpreter.java

@@ -13,7 +13,6 @@ public interface XInterpreter {
   void markNextEventLabel(String label);
   XNopEvent emitNopEvent();
   XAssertionEvent emitAssertionEvent(XBinaryComputationEvent assertion);
-  
   XLocalMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
                                     XLocalMemoryUnit source);
   XSharedMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
@@ -21,7 +20,6 @@ public interface XInterpreter {
   XSharedMemoryEvent emitMemoryEvent(XSharedLvalueMemoryUnit destination,
                                      XLocalMemoryUnit source);
 
-  // create computation event without processing it:
   XComputationEvent createComputationEvent(XUnaryOperator operator,
                                            XLocalMemoryUnit operand);
   XComputationEvent createComputationEvent(XBinaryOperator operator,
@@ -37,5 +35,4 @@ public interface XInterpreter {
   void finishNonlinearBlockDefinition();
   void processJumpStatement(JumpKind kind);
   XEntity processMethodCall(String methodName, XMemoryUnit... arguments);
-
 }

### 2018-06-07 5_eval.tex
lots of changes on itmo template.

@@
 The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
-The graphs contain a single source event and a single sink event represented by the grey triangles (in fact, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
+The graphs contain a single source event and a single sink event represented by the dark-grey triangles (in fact, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
 For clarity, all branching events, that in current example serve as the conditional events of loops, are highlighed with light-grey colour.
@@
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
-  \hfill
   \caption{The compiled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
@@
   \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
-  \hfill
   \caption{The compiled AST of the function in Figure~\ref{ex:both-cf:pts1}}
@@
 
-\begin{figure}[!h]
-\begin{lstlisting}
+\begin{figure}%[!h]
+\centering
+\begin{minipage}[t]{.5\textwidth}%for preventing overfull
+\begin{lstlisting}[language=Java]
 {
-  int x = 1, *y;
+  int x = 1, *y; //initialisation
 }
@@
 void thread_0(int &x, int &y) {
-  L0: x = 0;
+  L0: x = 0; //labelled statement
   int r;
@@
     if (y > 6) {
-      continue;
+      continue; //jump to the loop entry
     }
@@
       r = r + 10;
-      break;
+      break; //jump to the loop exit
     }
@@
   y = x + 1;
-  L1: x = r;
+  L1: x = r; //labelled statement
 }
@@
 \end{lstlisting}
+\end{minipage}
 \caption{Example: A litmus test in C with an arbitrary control-flow}
@@
 
-\begin{figure}[t]
+\begin{figure}%[t]
+\centering
+\begin{minipage}[t]{.49\textwidth}%for preventing overfull
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
 \caption{SB litmus test for the Linux kernel memory model}

### 2018-06-07 90_summary.tex
lots of changes on itmo template.

@@
 
-During the work on this thesis, we solved the following engineering and scientific tasks:
+During the work on this thesis, we solved the following research and engineering problems:
 
@@
 
-  \begin{enumerate}[leftmargin=\parindent,label=\alph*.]
+  \begin{enumerate}[label=\alph*.]
   \item The first stumbling block in extension of the input language was the \porthos[1] program input parser, which performs the full semantic analysis of the program.
@@
 
-  As an example, the \textit{variable kind analysis} (shared or local) is performed by \porthos[1] syntactically, depending on the operator or the function that uses the variable.
+  As an example, the \textit{variable kind analysis} (determining whether they are shared or local) is performed by \porthos[1] syntactically, depending on the operator or the function that uses the variable.
   In contrast, \porthos[2] traverses the program AST and collects information of all variables accessed in the program during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
@@
 
+\ifdefined\IsAalto
+\else
+  \vspace*{1em}
+\fi
+
 The new tool \porthos[2] with the extensible program domain knowledge base can be considered as a generalised framework for SMT-based memory model-aware analysis, which can not only perform the reachability and portability analysis, but also serve as a basis for other kinds of static analysis of concurrent programs.

### 2018-06-07 inc/lst/XInterpreter.java
appendices. still have empty pages.

@@ -1,38 +1,0 @@
-public interface XInterpreter {
-
-  XProcessId getProcessId();
-  void finish();
-  XProcess getResult();
-  
-  // emitting new events:
-  XEntryEvent emitEntryEvent();
-  XExitEvent emitExitEvent();
-  XBarrierEvent emitBarrierEvent(XBarrierEvent.Kind kind);
-  XJumpEvent emitJumpEvent();
-  XJumpEvent emitJumpEvent(String label);
-  void markNextEventLabel(String label);
-  XNopEvent emitNopEvent();
-  XAssertionEvent emitAssertionEvent(XBinaryComputationEvent assertion);
-  XLocalMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
-                                    XLocalMemoryUnit source);
-  XSharedMemoryEvent emitMemoryEvent(XLocalLvalueMemoryUnit destination,
-                                     XSharedMemoryUnit source);
-  XSharedMemoryEvent emitMemoryEvent(XSharedLvalueMemoryUnit destination,
-                                     XLocalMemoryUnit source);
-
-  XComputationEvent createComputationEvent(XUnaryOperator operator,
-                                           XLocalMemoryUnit operand);
-  XComputationEvent createComputationEvent(XBinaryOperator operator,
-                                           XLocalMemoryUnit operand1, 
-                                           XLocalMemoryUnit operand2);
-
-  // non-linear statements interpretation:
-  void startBlockDefinition(BlockKind blockKind);
-  void startBlockConditionDefinition();
-  void finishBlockConditionDefinition(XComputationEvent condition);
-  void startBlockBranchDefinition(BranchKind branchKind);
-  void finishBlockBranchDefinition();
-  void finishNonlinearBlockDefinition();
-  void processJumpStatement(JumpKind kind);
-  XEntity processMethodCall(String methodName, XMemoryUnit... arguments);
-}

### 2018-06-07 1_intro.tex
Consider final Professor's comments.
itmo: remove footnotes.

@@
 %
-\footnote{The \porthos{} project web site: \url{http://github.com/hernanponcedeleon/Dat3M}} %
+\footnote{The \porthos{} project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
 %
@@
 Chapter~\ref{ch:eval} gives a comparison of key features of \porthos[2] and \porthos[1] by providing examples of the compilation and the unrolling stages; and it provides some performance benchmarks of \porthos[2] in both reachability and portability analysis modes.
-Chapter~\ref{ch:summary} summarises results of the work and proposes possible directions of the future work.
+Chapter~\ref{ch:summary} summarises results of the work and proposes possible directions for future work.

### 2018-06-07 5_eval.tex
itmo: remove footnotes.

@@
 for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script that analyses and prints the results (commit 48ba7161).
-The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
-\footnote{\porthos[2] repository: \url{https://github.com/ajuszkowski/PorthosC}}. %
-%
+The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] GitHub repository.
 Both \porthos[1] and \porthos[2] are run in the state reachability mode for TSO memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).

### 2018-06-08 1_intro.tex
fix infinite recursion on footnotes.

@@
 %
-\footnote{In order to prescind from the implementation details while discussing the memory models theory, we avoid the use of the software-specific term \textit{thread} and the hardware-specific term \textit{processor}.
+\myfootnote{In order to prescind from the implementation details while discussing the memory models theory, we avoid the use of the software-specific term \textit{thread} and the hardware-specific term \textit{processor}.
 Instead, we adhere the terminology of the theory of concurrency employed by Ben-Ari~\cite{ben2006principles} by naming a concurrent piece of code the \textit{process}.}. %
@@
 %
-\footnote{The \tool{diy} project web site: \url{http://diy.inria.fr/}} %
+\myfootnote{The \tool{diy} project web site: \url{http://diy.inria.fr/}} %
 %
@@
 %
-\footnote{The \porthos{} project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
+\myfootnote{The \porthos{} project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
 %
@@
 %
-\footnote{Hereinafter with the names \textit{\porthos{}} and \textit{\porthos[1]} we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \textit{\porthos[2]}.}.
+\myfootnote{Hereinafter with the names \textit{\porthos{}} and \textit{\porthos[1]} we refer to the tool \porthos{} of version 1, whereas the new implementation of \porthos{} is called \textit{\porthos[2]}.}.
 %

### 2018-06-08 4_impl.tex
lots of changes on itmo template.
itmo: remove footnotes.
fix infinite recursion on footnotes.

@@
 %
-\footnote{To ensure this, one merely has to look at existing C compiler implementations, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+\myfootnote{To ensure this, one merely has to look at existing C compiler implementations, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
 %
 ), we consider the accomplished work as a step towards it.
-The repository for \porthos[2] project is \url{https://github.com/ajuszkowski/PorthosC}.
+The \porthos[2] repository is located at \url{https://github.com/ajuszkowski/PorthosC}.
 
@@
 %
-\footnote{Apart from the C language, \porthos[2] should be able to analyse litmus tests in different assembly languages and, as a compatibility mode, the input language of \porthos[1].} %
+\myfootnote{Apart from the C language, \porthos[2] should be able to analyse litmus tests in different assembly languages and, as a compatibility mode, the input language of \porthos[1].} %
 %
@@
 %
-\footnote{The Guava project repository: \url{https://github.com/google/guava/}}%
+\myfootnote{The Guava project repository: \url{https://github.com/google/guava/}}%
 %
@@
 %
-\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}} % 
+\myfootnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}} % 
 by Microsoft Research~\cite{de2008z3}, which is invoked by \porthos[2] via a Java API.
@@
 %
-\footnote{The \tool{ANTLR} project repository: \url{https://github.com/antlr/antlr4}}%
+\myfootnote{The \tool{ANTLR} project repository: \url{https://github.com/antlr/antlr4}}%
 %
@@
 \centering
+\begin{minipage}[t]{.6\textwidth}%for preventing overfull
 %basicstyle=\ttfamily\tiny,
@@
 \end{lstlisting}
+\end{minipage}
 \caption{Sketch of the input language grammar used by \oldporthos{}}
@@
 %TODO: check or remove this
-%\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
+%\myfootnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
 %
 so that it does not require to change the grammar for encoding the semantics of a new function. %adding support for a new function.
-%\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
+%\myfootnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
 As the reader might have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, different kinds of memory operations of of the \oldporthos{} input language vary syntactically as well.
@@
 %
-\footnote{ANTLR grammars repository path:~\url{https://github.com/antlr/grammars-v4}}). %
+\myfootnote{ANTLR grammars repository path:~\url{https://github.com/antlr/grammars-v4}}). %
 %
@@
 %
-\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
+\myfootnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter.
 For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
@@
 %
-\footnote{Hereinafter we use the term \textit{recursive} data structure (also called \textit{inductive} data structure) to refer a complex data type that can contain elements of the same type.} %
+\myfootnote{Hereinafter we use the term \textit{recursive} data structure (also called \textit{inductive} data structure) to refer a complex data type that can contain elements of the same type.} %
 %
@@
 %
-\footnote{Hereinafter all class diagrams are generated by IntelliJ IDEA Ultimate.}%
+\myfootnote{Hereinafter all class diagrams are generated by IntelliJ IDEA Ultimate.}%
 %
@@
 %
-\footnote{Here we should note that \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++).
+\myfootnote{Here we should note that \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++).
 In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language.
@@
 %
-\footnote{The \tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
+\myfootnote{The \tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
 %
@@
 \centering
+\begin{minipage}[t]{.9\textwidth}%for preventing overfull
 \lstinputlisting[language=java]{inc/lst/XMemoryManager.java}
+\end{minipage}
 \caption{X-memory manager public interface}
@@
 %
-\footnote{The original specification of \porthos[2] stated that the unrolling bound $k$ must be interpreted as the maximum number of instruction in the original code (technically, the number of expressions in the ANTLR syntax tree).
+\myfootnote{The original specification of \porthos[2] stated that the unrolling bound $k$ must be interpreted as the maximum number of instruction in the original code (technically, the number of expressions in the ANTLR syntax tree).
 Nonetheless, current implementation of the \xgraph{} unroller counts the X-level events as it is much simpler to implement (the questions about how to count complex expressions that involve multiple shared variables and method invocations are left open for the future versions of \porthos[2]).

### 2018-06-08 1_intro.tex
cleanup some.

@@
 
+\ifdefined\IsAalto
+\else
+  \vspace*{1em}
+\fi
 \section{Related work}
@@
 Even if all the compiler optimisations were disabled (which is a rare case nowadays), the behaviour of two compiled versions of the same program may differ due to differences between hardware memory models.
-As the result, a program compiled for the platform $\mathcal{T}$ can reach states that are unreachable on the platform $\mathcal{S}$, which is a \textit{portability bug} from the source platform $\mathcal{S}$ to the target platform $\mathcal{T}$~\cite{Porthos17a}.
+As the result, a program compiled for the platform~$\mathcal{T}$ can reach states that are unreachable on the platform~$\mathcal{S}$, which is a \textit{portability bug} from the source platform~$\mathcal{S}$ to the target platform~$\mathcal{T}$~\cite{Porthos17a}.
 

### 2018-06-08 4_impl.tex
cleanup some.

@@
 The label resolution is necessary for establishing links to labelled statements.
-In C, labelled statements are declared via the colon-syntax `\lstinline{<label>\;:\;<statement>}', and the labels are referenced by the jump-statement `\lstinline{goto\;<label>}'.
+In C, labelled statements are declared via the colon-syntax `\lstinline{<label> : <statement>}', and the labels are referenced by the jump-statement `\lstinline{goto <label>}'.
 The label resolution algorithm traverses the \ytree{} and collects all declared labels into a map \texttt{JumpsResolver} that points a label to the labelled statement.

### 2018-06-08 90_summary.tex
cleanup some.

@@
   \item As the C language supports unconditional \texttt{goto}-jumps, the control-flow graph can have an arbitrary structure, which can not be modelled solely by the program AST, therefore \porthos[2] compiles the AST to the low-level hardware-agnostic program representation \xgraph{}.
-  For that, we implemented the \textit{full C interpreter} discussed in Section~\ref{ch:impl:proc:x-compiler}.
+  For that, we implemented the \textit{full C~interpreter} discussed in Section~\ref{ch:impl:proc:x-compiler}.
 
@@
 \item For justifying the correctness of the verification that bases on the result of the input program unrolling (via the new unrolling scheme discussed in Section~\ref{ch:impl:proc:x-unroll}), one might want to be aware whether the execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
-This can be modelled with two kinds of sink events (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.
+This can be modelled with \textit{two kinds of sink events} (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.
 

### 2018-06-08 5_eval.tex
fix infinite recursion on footnotes.
Fix titles spacing and some pics floating.

@@
 For performing tests, \porthos[2] was slightly modified in the branch \textit{ay/test-timing}%
-\footnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
+\myfootnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
 %
@@
 
+\begin{figure}[!h]
+\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/cfg-arbitrary/collage.jpg}
+\caption{Control-flow components of the event-flow graphs for the code in Figure~\ref{ex:cfg-arbitrary}}
+\label{ex:cfg-arbitrary:pic}
+\end{figure}
+
 The four flow-graphs for the code in Figure~\ref{ex:cfg-arbitrary} are shown in Figure~\ref{ex:cfg-arbitrary:pic}.
@@
 
-\begin{figure}[!h]
-\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/cfg-arbitrary/collage.jpg}
-\caption{Control-flow components of the event-flow graphs for the code in Figure~\ref{ex:cfg-arbitrary}}
-\label{ex:cfg-arbitrary:pic}
-\end{figure}
-
 \subsection{Extensible compiler mapping}
 
-\begin{figure}%[t]
+%The revised design of \porthos{} tool includes the following new features:
+Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
+Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
+Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
+
+As an example, consider the Store Buffering litmus test for the Linux kernel in Figure~\ref{ex:sb:kernel} (which is similar to the one presented in Figure~\ref{simple_wmm_x86}).
+When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.
+
+\begin{figure}[h]
 \centering
@@
 
-P0(volatile int* y, volatile int* x) {
+P0(volatile int& y, volatile int& x) {
   int r0;
-  WRITE_ONCE(*x,1);
-  r0 = READ_ONCE(*y);
+  WRITE_ONCE(x,1);
+  r0 = READ_ONCE(y);
 }
 
-P1(volatile int* y, volatile int* x) {
+P1(volatile int& y, volatile int& x) {
   int r0;
-  WRITE_ONCE(*y,1);
-  r0 = READ_ONCE(*x);
+  WRITE_ONCE(y,1);
+  r0 = READ_ONCE(x);
 }
@@
 \end{figure}
-
-%The revised design of \porthos{} tool includes the following new features:
-Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
-Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
-Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
-
-As an example, consider the Store Buffering litmus test for the Linux kernel in Figure~\ref{ex:sb:kernel} (which is similar to the one presented in Figure~\ref{simple_wmm_x86}).
-When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

### 2018-06-08 5_eval.tex
hide hyperlinks. fix appendices empty pages.

@@
 For time benchmarking we ran the tools $5$ times and computed the median of the encoding time.
-Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
+Benchmarking was performed on the Linux machine, 8 Gb RAM, 8 cores Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, running under Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters with maximum heap size 1982 Mb).
 The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.

### 2018-06-08 4_impl.tex
hide hyperlinks. fix appendices empty pages.
Fix some bullets and footnotes for itmo. Update appendix for itmo.

@@
 This implies the support not only for new syntactic structures of the C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
-Yet far from all the C language is supported (which, considering its complexity and numerous pitfalls, goes far beyond current thesis%
+Yet far from all the C language is supported, which, taking into account its complexity and numerous pitfalls, goes far beyond current thesis%
 %
-\myfootnote{To ensure this, one merely has to look at existing C compiler implementations, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
+\myfootnote{To ensure this, one merely has to look at existing C compiler implementations, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code; see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c}.}%
 %
-), we consider the accomplished work as a step towards it.
+, we consider the accomplished work as a step towards it.
 The \porthos[2] repository is located at \url{https://github.com/ajuszkowski/PorthosC}.
@@
 \vspace{0.5em}
-\begin{enumerate}[nolistsep]
+\begin{enumerate}[nolistsep,label=\arabic*.]
   \item \textit{Robustness:}
@@
 
-\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{c11}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
+\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{c11}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements.
+The original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
 %
-\myfootnote{ANTLR grammars repository path:~\url{https://github.com/antlr/grammars-v4}}). %
+\myfootnote{The ANTLR grammars repository path:~\url{https://github.com/antlr/grammars-v4}}. %
 %
@@
 Thus, the method call resolution is performed statically at compile-time without any overhead at run-time.
-In our implementation, the visitor (not the visitee) cares about the continuation of traversing its children.
+%In our implementation, the visitor (not the visitee) cares about the continuation of traversing its children.
 

### 2018-06-10 1_intro.tex
final polishing.

@@
 %
-\myfootnote{The \porthos{} project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
+\myfootnote{The \porthos[1] project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
 %
@@
 One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
-For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read) and be easily extensible for defining new functions (have the separate module for the purpose of a knowledge base).
+For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro `\texttt{READ\_ONCE}' that guarantees memory read) and be easily extensible for defining new functions (have the separate module for the purpose of a knowledge base).
 %interprocedural analysis

### 2018-06-10 2_wmm.tex
final polishing.

@@
 The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
-We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1\,=\,{\texttt{load}(r_1\,\leftarrow\,l_2)}; \ e_2\,=\,{\texttt{store}(l_1\,\leftarrow\,r_1)}$.
+We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events `$e_1\,=\,{\texttt{load}(r_1\,\leftarrow\,l_2)}; \ e_2\,=\,{\texttt{store}(l_1\,\leftarrow\,r_1)}$'.
 Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.
@@
 
+%in Aalto template, the first line stucks on the previous page
+\ifdefined\IsAalto
+  \vspace*{1em}
+\fi
+
 \begin{enumerate}

### 2018-06-10 4_impl.tex
final polishing.

@@
 \item a variable declared as a parameter of the process function; and
-\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \lstinline{EXPORT_SYMBOL} and \lstinline{EXPORT_SYMBOL_GPL} also export symbols for dynamic linking~\cite{argawal16kernel}).
+\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions `\lstinline{EXPORT_SYMBOL}' and `\lstinline{EXPORT_SYMBOL_GPL}' also export symbols for dynamic linking~\cite{argawal16kernel}).
 \end{itemize}
@@
 An invocation hook can model any type of operations (memory, fence, computational, etc.).
-For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
-For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
+For instance, the \texttt{XKernelInvocationHook} intercepts the invocation `\lstinline{WRITE_ONCE(dst_shared, src_local)}' and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions `\lstinline{READ_ONCE}', `\lstinline{WRITE_ONCE}', and `\lstinline{ACCESS_ONCE}' may be modelled as volatile `\lstinline{memory_order_relaxed}' accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
 

### 2018-06-10 5_eval.tex
Fix some bullets and footnotes for itmo. Update appendix for itmo.
final polishing.

@@
 
-The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the branching statement \texttt{if-then-else}, the \texttt{while} loop and the \texttt{sequence} statement are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
+The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, branching statements, while-loops and sequence statements are expanded recursively, and the head and the tails of each complex statement are bound while processing the parent instruction).
 The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
@@
 Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
-Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
+Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as `\lstinline{READ_ONCE}' and `\lstinline{WRITE_ONCE}' for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
 Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
 \centering

### 2018-06-15 slides/slides.tex
Consider Professor's comments. Cleanup and extend.
Improve slides.

--- 2018-05-20 slides/slides.tex
@@
-\documentclass{beamer}
+\documentclass[aspectratio=149]{beamer}
 
@@
 \usepackage[russian,english]{babel}
+
+\usepackage{outlines}
+
+\usepackage{colortbl}
+
+\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
+\definecolor{ltblue}{RGB}{101,147,245}
+\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
+\definecolor{dkred} {rgb} {0.6, 0.1, 0}
+\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
+\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}
+
+\usepackage{xparse}
 \usepackage{tikz}
+\usetikzlibrary{tikzmark}
 \usetikzlibrary{positioning}
@@
 \usetikzlibrary{automata}
-%\usetikzlibrary{tikzmark}
-%\usetikzlibrary{calc}
 
@@
   }
+}
+
+\newcommand*{\yellowemph}[1]{%
+  \tikz[baseline=(X.base)] \node[rectangle, fill=yellow, inner sep=0mm] (X) {#1};%
+}
+
+\newcommand*{\rus}[1]{%
+  %\textcolor{black}{
+  \textit{#1}
+  %}
 }
@@
 
+%\usepackage{../inc/listingset}
 \usepackage{listings}
+\lstset{escapeinside={[*}{*]}}
 
@@
 \addbibresource{../bib/ms-thesis.bib}
-%\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}
 
 \usepackage{array}
-\usepackage{xcolor}
+%\usepackage{xcolor}
 
@@
 %\usetheme{Boadilla}
-%\usetheme{boxes}
+\usetheme{boxes}
 %\usetheme{CambridgeUS}
@@
 %\usetheme{Luebeck}
-%\usetheme{Madrid}
+%\usetheme{Madrid}\colorbox{yellow}
 %\usetheme{Malmoe}
 %\usetheme{Marburg}
-%\usetheme{Montpellier}
+%\usetheme{Montpellier} %!
 %\usetheme{PaloAlto}
@@
 %\usetheme{Rochester}
-\usetheme{Singapore}
+%\usetheme{Singapore}
 %\usetheme{Szeged}
@@
 %\setbeamertemplate{headline}{}
+
 \addtobeamertemplate{navigation symbols}{}{%
@@
     \hspace{1em}%
-    \insertframenumber/\inserttotalframenumber
+    \insertframenumber%/\inserttotalframenumber
 }
 
-\setbeamertemplate{frametitle}[default][left]
+%\setbeamercovered{transparent}
+%\setbeamertemplate{footline}{}
+
+%\setbeamertemplate{page number in head/foot}[framenumber]
+
+%\setbeamertemplate{frametitle}[default][left]
 
@@
 
-\institute % (optional, but mostly needed)
+\institute
 {
   \inst{1}%
+  \textbf{Aalto University} (Espoo, Finland) \\
   Department of Computer Science, \\
-  School of Science, \\
-  \textbf{Aalto University} (Espoo, Finland)%
+  School of Science
   \and
   \inst{2}%
+  \textbf{ITMO University} (Saint Petersburg, Russia) \\
   Faculty of Information Security \\
-  and Computer Technologies, \\
-  \textbf{ITMO University} (Saint Petersburg, Russia)%
+  and Computer Technologies
 }
@@
 % the beginning of each subsection:
-\AtBeginSection[]
-{
-  \begin{frame}<beamer>{Outline}
-    \tableofcontents[currentsection]%,currentsubsection]
-  \end{frame}
-}
-
-\usepackage{colortbl}
-
-\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
-\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
-\definecolor{dkred} {rgb} {0.6, 0.1, 0}
-\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
-\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}
+%\AtBeginSection[]
+%{
+%  \begin{frame}<beamer>{Outline}
+%    \tableofcontents[currentsection]%,currentsubsection]
+%  \end{frame}
+%}
 
@@
 \begin{document}
+%\beamertemplatenavigationsymbolsempty %remove navigation bar
 
@@
 
-\begin{frame}{Outline}
-  \tableofcontents
-\end{frame}
+%\begin{frame}{Outline}
+%  \tableofcontents
+%\end{frame}
 
-\section{Introduction}
+%\section{Introduction}
+%\subsection{The problem}
 
-\subsection{The problem}
-
-\begin{frame}{Problem statement (Цель работы)}
-To rework the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by:
+\begin{frame}{Task statement \rus{(Цель работы)}}
 \begin{itemize}
-\item extending the C-like input language,
-\item revising its architecture and
-\item re-implementing the tool in order to enhance performance, extensibility, reliability and maintainability
+\item Improve the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by extending support for the input language.
 \end{itemize}
@@
 
-\begin{frame}{Task specification (Задачи работы)}
+\begin{frame}{Task specification \rus{(Задачи работы)}}
 \begin{itemize}
@@
 \item Review existing tools for memory model-aware analysis;
-\item Investigate existing architecture of \texttt{Porthos}, its strengths and weaknesses;
-\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
+%\item Examine the architecture of \texttt{Porthos}, its strengths and weaknesses;
+%\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
+\item Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};
+\item Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};
+\item Enhance performance, extensibility, reliability and maintainability of the tool.
 \end{itemize}
@@
 
-\subsection{The background: weak memory model-aware analysis}
+%\subsection{The background: weak memory model-aware analysis}
 
-\begin{frame}{Verification of concurrent software} {Example: Write-write reordering (compiler  relaxations)}
+\begin{frame}{Verification of concurrent software} {Motivating example: Write-write reordering (compiler  relaxations)}
 %1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
 %2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
-
 \begin{table}
@@
 \ttfamily
-\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
 \hline
@@
 P & Q \\ \hline
-$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
-$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+\tikzmark{p0}$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\tikzmark{q0} \\
+\tikzmark{p1}$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\tikzmark{q1} \\
+\hline
+\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
 \hline
@@
 
-%\tikzmark{p1} 
-%\begin{tikzpicture}[remember picture, overlay]
-%  \path[draw=red,thick,<->]<4-> (pic cs: p0) to [] (pic cs: p1);
-%\end{tikzpicture}
-
-\hspace{-20pt}
+\onslide<2->
 \begin{minipage}{.25\textwidth}
 \small
-\onslide<2-> 
 \begin{table}
-\caption*{SC}
-\begin{tabular}{ >{\columncolor{ltgrey}}l >{\columncolor{ltgrey}}r}
+\only<1>{\caption*{~\\~}}
+\only<2>{\caption*{Single thread\\(no concurrency)}}
+\only<3>{\caption*{Sequential Consistency\\(classical concurrency)}}
+\only<4->{\caption*{~\\~}}
+\only<4->{
+  \begin{tikzpicture}[overlay, remember picture]
+    \draw [<->,thick,red,looseness=3] ([xshift=-5pt,yshift=3pt]{pic cs:p0}) to[bend right=90] ([xshift=-5pt]{pic cs:p1});
+    \draw [<->,thick,red,looseness=3] ([xshift=8pt,yshift=3pt]{pic cs:q0}) to[bend left=90] ([xshift=5pt]{pic cs:q1});
+  \end{tikzpicture}%
+}
+%"A multiprocessor is sequentially consistent if the result of any
+%execution is the same as some sequential order, and within any
+%processor, the operations are executed in program order"
+%– Lamport, 1979
+\begin{tabular}{ l r }
 $\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
@@
 \end{minipage}
-\pause
 %
-\hspace{5pt}
+\onslide<4->
 \begin{minipage}{.6\textwidth}
 \small
-\onslide<4->
-%\vspace{-1em}
 %"delay the effect of a store past any load from a different location."
-%\vspace{-11pt}
-%\vspace{-2pt}
 \begin{table}
-\caption*{TSO}
-\begin{tabular}{ | l r | l r | l r }
+%http://events17.linuxfoundation.org/sites/events/files/slides/dbueso-elc2016-membarriers-final.pdf
+%SPARC, x86 (Intel, AMD)
+%Similar to SC, but:
+%‒ Loads may be reordered with writes.
+%Relaxed Models
+% Arbitrary reorder limited only by explicit memorybarrier
+%instructions.
+\caption*{Total Store Order\\(e.g., x86)}
+\begin{tabular}{ | l r | l r | l r } % >{\columncolor{ltgrey}}
 $\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
 $\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
-$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
-$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
-$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & $(\b{0}; \r{0})$ \\
-$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & $(\b{0}; \r{0})$ \\
+$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
+$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
+$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}}  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
+$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
 \end{tabular}
@@
 
-\begin{frame}{Verification of concurrent software} {Example: Store buffering (hardware relaxations)}
+\begin{frame}{Verification of concurrent software} {Motivating example: Store buffering (hardware relaxations)}
 \begin{center}
-\scalebox{0.7}{
+\scalebox{0.75}{
 \ttfamily
-\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
 \hline
@@
 P & Q \\ \hline
-$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
-$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
+$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
+$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
+\hline
+\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
 \hline
@@
 \caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
-%\label{fig:x86-arch}
+\only<2->{
+    \begin{tikzpicture}[overlay]
+    \node at (0.1, 3.1) {\fcolorbox{yellow}{yellow}{
+        \includegraphics[scale=0.2]{img/table.png}} };
+    \end{tikzpicture}
+}
 \end{figure}
+
 \end{frame}
 
-\begin{frame}{The weak memory model} {Axiomatic semantics: The definition}
+\begin{frame}{Weak memory model-aware analysis} {Event-based program representation}
 
@@
 
-\begin{frame}{The weak memory model} {Testing the candidate executions}
+\begin{frame}{Weak memory model-aware analysis} {Testing candidate executions}
 \vspace{10pt}
-\begin{minipage}{.3\linewidth}
+\begin{minipage}{.33\linewidth}
 \hfill
@@
 %
-\noindent\begin{minipage}{.4\linewidth}
-\scalebox{0.7}{
+\noindent\begin{minipage}{.39\linewidth}
+\scalebox{0.65}{
 \ttfamily
-\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
+\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
 \hline
@@
 %
-\noindent\begin{minipage}{.25\linewidth}
-\end{minipage}
-\vspace{-5pt}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{img/candidate.png}
-\caption{The four candidate executions allowed under x86-TSO}
-\end{figure}
-\end{frame}
-
-\begin{frame}{The weak memory model} {Testing the candidate executions}
-\vspace{10pt}
-\begin{minipage}{.3\linewidth}
-\hfill
-\end{minipage}
-%
-\noindent\begin{minipage}{.4\linewidth}
-\scalebox{0.7}{
-\ttfamily
-\begin{tabular}{ |>{\color{dkblue}}c | >{\color{dkred}}c| }
-\hline
-\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
-P & Q \\ \hline
-$p_0$ : $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$   \\
-$p_1$ : $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$ \\
-\hline
-\end{tabular}
-}
-\end{minipage}
-%
-\raggedleft{
-\noindent\begin{minipage}{.25\linewidth}
+\noindent\begin{minipage}{.23\linewidth}
+\only<2->{
 \scriptsize
@@
 \color{red}\texttt{acyclic(fr $\cup$ po)}
+}
 \end{minipage}
+\vspace{-5pt}
+\only<1>{
+\begin{figure}
+%\centering
+\includegraphics[width=\linewidth]{img/candidate.png}
+\caption{The four candidate executions allowed under x86-TSO}
+\end{figure}
 }
-\vspace{-5pt}
+\only<2->{
 \begin{figure}
-\centering
+%\centering
 \includegraphics[width=\linewidth]{img/candidate-red.png}
@@
 \end{figure}
+}
+\only<3>{
+  \begin{tikzpicture}[overlay]
+  \node at (6.1, 4.1) {\fcolorbox{yellow}{yellow}{
+    \includegraphics[scale=0.11]{img/non-portable.png}} };
+  \end{tikzpicture}
+}
 
@@
 
-\begin{frame}{Tools for memory model-aware analysis}
+\begin{frame}<presentation:0>{Tools for memory model-aware analysis}
 \begin{itemize}
@@
 
-\subsection{Portability Analysis}
-
-\begin{frame}{Portability analysis} {The \texttt{Porthos} tool}
-
+\begin{frame}{Portability analysis} {The \tool{Porthos} tool}
+\begin{minipage}{.49\textwidth}
+\only<1->{
+  \begin{figure}
+  \includegraphics[height=0.4\textheight,keepaspectratio]{img/the-problem.png}
+  \caption{The illustration of the portability problem~\cite{Porthos17slides}}
+  \end{figure}
+}
+\end{minipage}
+%
+\begin{minipage}{.5\textwidth}
+\only<2->{
   \begin{itemize}
-\item Let the function $\textit{cons}_{\mathcal{M}}(P)$ calculate the set of executions of program $P$ consistent under the memory model $\mathcal{M}$.
-\end{itemize}
-
     \begin{definition}[Portability~\cite{Porthos17a}]
-Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
-The program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
+    %SAY: Let $\textit{cons}_{\mathcal{M}}(P)$ be a function that calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$
+    %Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
+    The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target one $\mathcal{M_T}$ if
     $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
     \end{definition}
-
-\begin{itemize}
     \item Portability as an SMT-based bounded reachability problem: \\
-${\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}}$
-\item $\texttt{SAT}(\phi) \implies $ the portability bug
+    $\phi = $\only<3>\yellowemph{$\phi_{CF}$}$\land $\only<3>\yellowemph{$\phi_{DF}$}$\land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$
+    \item $\texttt{SAT}(\phi) \Longrightarrow $ portability bug
   \end{itemize}
-
-%end{tabular}
+}
+\end{minipage}
 \end{frame}
 
-\subsection{SMT-encoding}
+\begin{frame}{\tool{PorthosC}: Architecture}{Main components}
+\vspace{-30pt}
+\begin{minipage}{.49\textwidth}
+\begin{itemize}
+\item The enhanced version of \tool{Porthos}, that is able to analyse C programs, has got a new name \textit{PorthosC}.
+\end{itemize}
+\hfill
+\end{minipage}
+%
+\begin{minipage}{.5\textwidth}
+\only<1>{\begin{figure}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering.png}
+\end{figure}}
+\only<2>{\begin{figure}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-labels.png}
+\end{figure}}
+\only<3>{\begin{figure}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-visitors.png}
+\end{figure}}
+\only<4>{\begin{figure}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-compiler.png}
+\end{figure}}
+\only<5>{\begin{figure}
+  \includegraphics[height=.98\textheight,keepaspectratio]{img/arch/general_arch-no_numbering-smt.png}
+\end{figure}}
+\end{minipage}
+\end{frame}
 
-%\begin{frame}{Encoding for the control-flow}% {The \texttt{Porthos v1} encoding scheme}
-%\begin{itemize}
-%\item In \texttt{Porthos}, the high-level instructions were represented in the SMT-formula by separate variables.
-%\item For example, the sequential instruction $i_1 = i_2; i_3$ was encoded as
-%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
-%\end{itemize}
-%\end{frame}
+\begin{frame}{\tool{PorthosC}: Architecture}{Abstract interpretation module}
+  \vspace{-40pt}
+\begin{minipage}{.4\textwidth}
+\begin{itemize}
+\item The abstract interpretation module is an essential part of the compiler of C code into the event-flow graph representation (\texttt{X-graph}).
+\end{itemize}
+\hfill
+\end{minipage}
+%
+\begin{minipage}{.5\textwidth}
+\only<1>{\begin{figure}
+  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler.png}}
+\end{figure}}
+\only<2>{\begin{figure}
+  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-1.png}}
+\end{figure}}
+\only<3>{\begin{figure}
+  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-2.png}}
+\end{figure}}
+\only<4>{\begin{figure}
+  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-3.png}}
+\end{figure}}
+\only<5>{\begin{figure}
+  \scalebox{0.9}{\includegraphics[height=.99\paperheight]{img/compiler/X-compiler-4.png}}
+\end{figure}}
+\end{minipage}
+\end{frame}
 
-%\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}
-%\begin{itemize}
-%%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
-%\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
-%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
-%\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
-%\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
-%\end{itemize}
+\begin{frame}{\tool{Porthos\,v1}: The input language}
+{
+\only<1-2>{ }
+\only<3>{Static syntactic determination of the variables kind}
+\only<4>{Lack of support for functions invocations}
+\only<5>{Lack of support for unconditional jumps}
+}
+{\noindent\centering
+\begin{minipage}{.45\textwidth}
+  \begin{figure}
+  \only<1-2>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-old/example-old.c}}
+    \only<3>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-old/example-old-2.c}}
+    \only<4>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-old/example-old-3.c}}
+    \only<5>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-old/example-old.c}}
+  \caption{A program example in the \tool{Porthos\,v1} input language}
+  \end{figure}
+\end{minipage}
 %
-%\begin{figure}[b]
-%  \centering
-%  \begin{minipage}{\textwidth}
-%      \begin{subfigure}[b]{0.25\textwidth}
-%        \makebox[\textwidth]{
-%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-%            \node[c] (1) [] {$e_i$};
-%            \node[c] (2) [below of=1] {$e_{i+1}$};
-%            \path[->]
-%            (1) edge [] node {$\epsilon$} (2)
-%            ;
-%        \end{tikzpicture}
+\begin{minipage}{.45\textwidth}
+  \begin{figure}
+  \only<2,5>{\lstinputlisting[basicstyle=\ttfamily\Tiny,columns=fullflexible,literate={<}{{$\langle$}}1{>}{{$\rangle$}}1,escapeinside={[*}{*]}]{img/lst/input-grammar/grammar.txt}}
+  \only<3>{\lstinputlisting[basicstyle=\ttfamily\Tiny,columns=fullflexible,literate={<}{{$\langle$}}1{>}{{$\rangle$}}1,escapeinside={[*}{*]}]{img/lst/input-grammar/grammar-2.txt}}
+  \only<4>{\lstinputlisting[basicstyle=\ttfamily\Tiny,columns=fullflexible,literate={<}{{$\langle$}}1{>}{{$\rangle$}}1,escapeinside={[*}{*]}]{img/lst/input-grammar/grammar-3.txt}}
+  \vfill
+  \only<2->{\caption{A sketch of the \tool{Porthos\,v1} input language grammar}}
+  \end{figure}
+\end{minipage}
+}
+\end{frame}
+
+\begin{frame}[label={example-new}]{\tool{PorthosC}: The input language}
+{
+\only<1-2>{ }
+\only<3>{Pre-compilation phase for the syntactic determination of the variables kind}
+\only<4>{Support for functions invocations: The invocation hooking mechanism}
+\only<5>{Support for arbitrary control-flow jumps}
+}
+{\noindent\centering
+\begin{minipage}{.43\textwidth}
+  \begin{figure}
+  \only<1-2>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-new/example-new.c}}
+    \only<3>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-new/example-new-2.c}}
+    \only<4>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-new/example-new-3.c}}
+    \only<5>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-new/example-new-4.c}}
+  \caption{A program example in the \tool{Porthos\,v1} input language}
+  \end{figure}
+\end{minipage}
+%
+\begin{minipage}{.55\textwidth}
+%  \only<1>{\begin{itemize}
+%  \item The grammar used by \texttt{PorthosC} was taken from the C11 standard~\cite{c11}.
+%  \end{itemize}}
+%  %
+%  \only<2>{\begin{itemize}
+%    \item A variable is shared if:
+%      \begin{itemize}
+%      \item it was declared as a \textit{pointer}, or
+%      \item its \textit{address} is accessed at least once by the code of any process, or 
+%      \item it was declared as a \textit{parameter} of the function that defines a process, or 
+%      \item it was exported by the \textit{extern} keyword or an exporting function.
+%      \end{itemize}
+%    \end{itemize}
 %  }
-%        \caption{The sequence}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[b]{0.35\textwidth}
-%        \makebox[\textwidth] {
-%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-%            \node[c]           (k)   [] {$e_k$};
-%            \node[c,draw=none] (iii) [below of=k] {$...$};
-%            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
-%            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
-%            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
-%            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
-%            \path[->]
-%            (k) edge [above] node {$g_{i,k}$} (i)
-%            (k) edge [] node {} (ii)
-%            (k) edge [] node {} (iii)
-%            (k) edge [] node {} (iv)
-%            (k) edge [above] node {$g_{i+j,k}$} (v)
-%            ;
-%        \end{tikzpicture}
+%  \only<3>{\begin{itemize}
+%    \item The invocation hooking mechanism serves as an extensible knowledge base for storing the semantics of functions.
+%    \end{itemize}
 %  }
-%        \caption{Conditional branching}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[b]{0.3\textwidth}
-%        \makebox[\textwidth]{
-%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
-%            \node[c] (i) [] {$e_i$};
-%            \node[c,draw=none] (ii) [right=0.2cm of i] {};
-%            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
-%            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
-%            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
-%            \node[c] (k) [below of=iii] {$e_k$};
-%            \path[->]
-%            (i) edge [above] node {$g_{i,k}$} (k)
-%            (ii) edge [above] node {} (k)
-%            (iii) edge [] node {} (k)
-%            (iv) edge [] node {} (k)
-%            (v) edge [] node {} (k) %sloped
-%            ;
-%        \end{tikzpicture}
+%  \only<4>{\begin{itemize}
+%    \item The labels are being collected during the pre-compilation stage,
+%    \item the jumps are being resolved during the compilation.
+%    \end{itemize}
+%  }  \scalebox{0.75}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0-3.png}}
 %  }
-%        \caption{Branch merging}
-%    \end{subfigure}
-%    %
-%    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
-%    \label{encode:cf}
-%\end{minipage}
-%\end{figure}
-%\end{frame}
+%  \only<5>{
+%    \scalebox{0.75}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0-4.png}}
+%  }
+  \begin{figure}
+  \only<2-3>{
+    \scalebox{0.75}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0.png}}
+  }
+  \only<4>{
+    \scalebox{0.75}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0-4.png}}
+  }
+  \only<5>{
+    \scalebox{0.75}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0-5.png}}
+  }
+  \only<2->{\caption{The compiled event-flow graph}}
+  \end{figure}
+\end{minipage}
+}
+\end{frame}
 
-%\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}
-%\begin{itemize}
-%\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
-%\item The control-flow of the program is encoded as following:
-%\end{itemize}
-%\vspace{-20pt}
-%\begin{align}
-%\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
-%\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
-%                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
-%                       \land & \ \cdots \nonumber \nonumber\\
-%                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
-%                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
-%\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
-%\end{align}
-%\end{frame}
-
-\begin{frame}{Encoding for the control-flow: An example}
-%Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
+\begin{frame}{\texttt{PorthosC}: The new control-flow encoding scheme}
 \begin{figure}
@@
 
-\begin{frame}{Encoding for the data-flow}
+\begin{frame}<presentation:0>{Encoding for the data-flow}
 \begin{itemize}
@@
 
-\section{Implementation}
-
-%\begin{frame}{General principles}
-%\begin{enumerate}[nolistsep]
-%  \item \textit{Robustness:}
-%    \begin{enumerate}[label*=\arabic*.]
-%      \item preservation the completeness of analysis,
-%      \item modular architecture: each module can be tested independently,
-%      \item usage of software design patterns where necessary, and
-%      \item usage of immutable data structures for all DTOs.
-%    \end{enumerate}
-%  \item \textit{Transparency:}
-%    \begin{enumerate}[label*=\arabic*.]
-%      \item following the principles of simplicity and readability, and
-%      \item clear and informative program output.
-%    \end{enumerate}
-%  \item \textit{Efficiency:}
-%    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
-%      \item keeping the trade-off between execution time and memory usage.
-%    \end{enumerate}
-%  \item \textit{Extensibility:}
-%    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
-%      \item clear modular architecture.
-%    \end{enumerate}
-%\end{enumerate}
-%\end{frame}
-
-\subsection{The input language and general architecture}
-
-\begin{frame}{The input language}
-The input language parser used by \texttt{Porthos} suffered from several disadvantages:
-\begin{itemize}
-\item it contained the parser code inlined directly into the grammar (hardly maintainable);
-\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\texttt{=}', `\texttt{:=}', `\texttt{<-}' and `\texttt{<:-}', each for different kinds of arguments);
-\item restricted syntax for expressions.
-\item In contrast, \texttt{PorthosC} uses the full C language grammar of proposed in the C11 standard~\cite{jtc2011sc22} and the visitor that converts the ANTLR grammar to the AST (\texttt{Y-tree}).
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Architecture}%{The general architecture}
-\vspace{-30pt}
-\begin{figure}
-  \centering
-  \includegraphics[height=.99\textheight,keepaspectratio]{../img/my/draw.io/general_arch-no_numbering.png} %{img/my/draw.io/general_arch.png}
-  %\caption{The general architecture of \texttt{PorthosC}}
-  %\label{fig:arch}
-\end{figure}
-\end{frame}
-
-\subsection{The \texttt{X-graph}: structure and construction}
-
-\begin{frame}{The \texttt{X-graph} internal representation}
-  \centering
-  \begin{figure}
-    \includegraphics[width=1.1\textwidth,keepaspectratio]{../img/my/class-diagrams/XEntity-interfaces.png}
-    \caption{The inheritance tree of main \texttt{X-graph} interfaces}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}{The \texttt{X-graph} compiler}
-\begin{figure}
-  \centering
-  \scalebox{0.7}{\includegraphics[height=.99\paperheight]{../img/my/draw.io/X-compiler.png}}
-  \caption{Main components of the X-compilation processing unit}
-\end{figure}
-\end{frame}
-
-\begin{frame}{\texttt{X-graph} unrolling}
+\begin{frame}{\texttt{PorthosC}: The new DFS-based \texttt{X-graph} unrolling scheme}
+\only<1>{
 \begin{figure}[t]
@@
 \end{figure}
+}
+%
+%\only<2>{
+%\begin{figure}
+%  \scalebox{0.8}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/input-new/t0_unrolled.png}}
+%  \caption{The unrolled \texttt{X-graph} of the code used on slide~\ref{example-new}}
+%\end{figure}
+%}
 \end{frame}
 
-\section{Evaluation}
-
-\begin{frame}{Evaluation}
-\alert{[to be done]}
+\begin{frame}{Evaluation}{The new unrolling scheme}
+\begin{minipage}{.23\linewidth}
+\hfill
+\end{minipage}
+%
+\noindent\begin{minipage}{.75\linewidth}
+\begin{figure}
+  \vspace{-20pt}
+  \centering
+  \scalebox{.8}{
+    \colorbox{white}{\includegraphics[height=\textheight,keepaspectratio]{img/lst/unrolled/collage.jpg}}
+  }
+\caption{Illustration of differences in unrolling schemes of \tool{PorthosC}~(left) and \tool{Porthos\,v1}~(right)}
+\end{figure}
+\end{minipage}
 \end{frame}
 
-\section*{Summary}
+\begin{frame}{Evaluation}{Overhead of the new unrolling scheme: Number of events}
+\begin{figure}[t]
+\begin{subfigure}{.47\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/new-e.png}
+\caption{The graph unrolled by \tool{PorthosC}}
+\label{dep:events:new}
+\end{subfigure}
+\hfill
+%
+\begin{subfigure}{.47\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/old-e.png}
+\caption{The graph unrolled by \tool{Porthos\,v1}}
+\label{dep:events:old}
+\end{subfigure}
+%
+\caption{Dependency of unrolled events number on the unrolling bound $k$}
+\label{dep:events}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Evaluation}{Overhead of the new unrolling scheme: Execution time}
+\begin{figure}[t]
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/new-t.png}
+\caption{Program domain encoding time of the graph unrolled by \tool{PorthosC}}
+\label{dep:time:new}
+\end{subfigure}
+\hfill
+%
+\begin{subfigure}{.45\textwidth}
+\includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/old-t.png}
+\caption{Program domain encoding time of the graph unrolled by \tool{Porthos\,v1}}
+\label{dep:time:old}
+\end{subfigure}
+%
+\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
+\label{dep:time}
+\end{figure}
+\end{frame}
+
+%\section*{Summary}
 
@@
 \begin{itemize}
-\item The general framework for memory model-aware analysis was implemented in \texttt{PorthosC};
-%\item Review existing tools for memory model-aware analysis;
-\item The input language has been extended; %TODO: how exactly?
-\item The old architecture of \texttt{Porthos} has been analysed and considered while designing the new architecture for \texttt{PorthosC};
-%\item Design a new architecture for \texttt{PorthosC} that \textit{allow to} easily support the C input language, be robust, transparent, efficient and extensible.
-\item \alert{to be done: more}
-%\item the new architecture for the tool was designed,
+\item The result of the work is the generalised framework for memory model-aware analysis \textbf{\tool{PorthosC}};
+  \begin{itemize}
+  \item The \textit{input language} has been extended to a higher subset of C language (including unconditional control-flow jumps);
+  \item The old \textit{architecture} of \texttt{Porthos} has been revised and redesigned for \texttt{PorthosC};
+  \item The \textit{new unrolling scheme} produces complete state space of the analysing program within the user-defined bound $k$, though the encoding time growth rapidly (exponentially) as the unrolling bound grows.
+  \item The \textit{invocation hooking mechanism} serves as a knowledge base for the program domain, it is to be filled and extended in future.
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Directions for future work}
+\begin{itemize}
+\item Extending the \textit{knowledge base} of domain-specific functions to model synchronisation primitives;
+\item Supporting \textit{new input languages} (e.g., different assembly languages);
+\item Supporting \textit{complex data types} (s.a. arrays, pointers, structures, etc.);
+\item Adding the \textit{inter-procedural analysis mode};
+\item Handling the \textit{state explosion problem} (with standard model-checking techniques adjusted by the information about the weak memory model of the execution environment).
 \end{itemize}

### 2018-06-15 slides/img/lst/control-flow.c

@@ -1,0 +1,30 @@
+{   int x = 1, *y; }
+
+void thread_0(int &x, int &y) {
+    L0: x = 0; //labelled statement
+    int r;
+    while (x * (5 + 4 / 2) % 3 == 1) {
+      if (x != 0) {
+        goto L0; //backward jump
+      }
+      if (y > 6) {
+        continue; //jump to the loop entry
+      }
+      else if (++y > 7) {
+        r = r + 10;
+        break; //jump to the loop exit
+      }
+      else {
+        goto L1; //forward jump
+      }
+      r = 11;
+    }
+    y = x + 1;
+    L1: x = r; //labelled statement
+}
+
+void thread_1() {
+    ...
+}
+
+exists (y == x + 1 && thread_0:r > 21)

### 2018-06-15 slides/img/lst/ekker/Dekker.c

@@ -1,0 +1,24 @@
+{ int flag0 = 0, flag1 = 0, turn = 0; }
+
+void P0() {
+    while (true) {
+      int a = 1, b = 0;
+      flag0.store(memory_order_relaxed, a);
+      f1 = flag1.load(memory_order_relaxed);
+      while (f1 == 1) {
+        t1 = turn.load(memory_order_relaxed);
+        if (t1 != 0) {
+          flag0.store(memory_order_relaxed, b);
+          t1 = turn.load(memory_order_relaxed);
+          while (t1 != 0)
+            t1 = turn.load(memory_order_relaxed);
+          flag0.store(memory_order_relaxed, a);
+        }
+      }
+    }
+}
+
+void P1() {
+    ...
+}
+exists (turn == 10)

### 2018-06-15 slides/img/lst/input-new/example-new-2.c

@@ -1,0 +1,25 @@
+{  [*\yellowemph{int *x}*] = 1; }
+
+[*\yellowemph{extern int z}*];
+
+void thread_0(int &x, [*\yellowemph{int \&y}*]) {
+  L0: x = 0;
+  [*\yellowemph{int r}*];
+  while (x * (5 + 4 / 2) % 3 == 1) {
+    if (x != 0)
+      goto L0;
+    if (y > 6)
+      continue;
+    else if (++y > 7) {
+      r = r + 10;
+      break;
+    }
+    else
+      goto L1;
+    r = 11;
+  }
+  y = x.load(_rx) + 1;
+  L1: x.store(_rx, r);
+}
+
+exists (y == x + 1 && thread_0:r > 21)

### 2018-06-15 slides/img/lst/input-new/example-new-3.c

@@ -1,0 +1,25 @@
+{  int *x = 1; }
+
+extern int z;
+
+void thread_0(int &x, int &y) {
+  L0: x = 0;
+  int r;
+  while (x * (5 + 4 / 2) % 3 == 1) {
+    if (x != 0)
+      goto L0;
+    if (y > 6)
+      continue;
+    else if (++y > 7) {
+      r = r + 10;
+      break;
+    }
+    else
+      goto L1;
+    r = 11;
+  }
+  y = x.[*\yellowemph{load}*](_rx) + 1;
+  L1: x.[*\yellowemph{store}*](_rx, r);
+}
+
+exists (y == x + 1 && thread_0:r > 21)

### 2018-06-15 slides/img/lst/input-new/example-new-4.c

@@ -1,0 +1,25 @@
+{  int *x = 1; }
+
+extern int z;
+
+void thread_0(int &x, int &y) {
+  [*\yellowemph{L0}*]: x = 0;
+  int r;
+  while (x * (5 + 4 / 2) % 3 == 1) {
+    if (x != 0)
+      [*\yellowemph{goto L0}*];
+    if (y > 6)
+      [*\yellowemph{continue}*];
+    else if (++y > 7) {
+      r = r + 10;
+      [*\yellowemph{break}*];
+    }
+    else
+      [*\yellowemph{goto L1}*];
+    r = 11;
+  }
+  y = x.load(_rx) + 1;;
+  [*\yellowemph{L1}*]: x.store(_rx, r);
+}
+
+exists (y == x + 1 && thread_0:r > 21)

### 2018-06-15 slides/img/lst/input-new/example-new-bak.c

@@ -1,0 +1,25 @@
+{ int *x = 0; }
+
+// extern int z;
+
+void t0(int &x, int &y) {
+    L0: y = 3;
+    int r0 = 1;
+    y = r0 * x;
+    while (++r0 > 2) {
+        if (y == 4)
+            break;
+        if (y == 5)
+            continue;
+        if (r0 == 6) {
+        	goto L1;
+        }
+        x.store(_rx, r0);
+        L1: y = x.load(_rx);
+    }
+    if (z == 7) {
+        goto L0;
+    }
+}
+
+exists (x == 9 && (y == 10 || t0:r0 >= 11))

### 2018-06-15 slides/img/lst/input-new/example-new.c

@@ -1,0 +1,25 @@
+{  int *x = 1; }
+
+extern int z;
+
+void thread_0(int &x, int &y) {
+  L0: x = 0;
+  int r;
+  while (x * (5 + 4 / 2) % 3 == 1) {
+    if (x != 0)
+      goto L0;
+    if (y > 6)
+      continue;
+    else if (++y > 7) {
+      r = r + 10;
+      break;
+    }
+    else
+      goto L1;
+    r = 11;
+  }
+  y = x.load(_rx) + 1;
+  L1: x.store(_rx, r);
+}
+
+exists (y == x + 1 && thread_0:r > 21)

### 2018-06-15 slides/img/lst/input-old/example-old-1.c

@@ -1,0 +1,15 @@
+{ x, y }
+
+thread t0 {
+    r0 [*\yellowemph{<-}*] 1;
+    r1 [*\yellowemph{<-}*] (2 + r0) * 3;
+    y [*\yellowemph{:=}*] r2;
+    while (r0 > 4) {
+      r0 [*\yellowemph{<:-}*] x;
+      r1 [*\yellowemph{<-}*] (r0 + 5);
+      x.store(_rx, r1);
+      y [*\yellowemph{=}*] x.load(_rx)
+    };
+}
+
+exists x = 1, y = 2, 0:r1 = 3,

### 2018-06-15 slides/img/lst/input-old/example-old-3.c

@@ -1,0 +1,15 @@
+{ x, y }
+
+thread t0 {
+    r0 <- 1;
+    r1 <- (2 + r0) * 3;
+    y := r2;
+    while (r0 > 4) {
+      r0 <:- x;
+      r1 <- (r0 + 5);
+      x.[*\yellowemph{store}*](_rx, r1);
+      y = x.[*\yellowemph{load}*](_rx)
+    };
+}
+
+exists x = 1, y = 2, 0:r1 = 3,

### 2018-06-15 slides/img/lst/input-old/example-old.c

@@ -1,0 +1,15 @@
+{ x, y }
+
+thread t0 {
+    r0 <- 1;
+    r1 <- (2 + r0) * 3;
+    y := r2;
+    while (r0 > 4) {
+      r0 <:- x;
+      r1 <- (r0 + 5);
+      x.store(_rx, r1);
+      y = x.load(_rx)
+    };
+}
+
+exists x = 1, y = 2, 0:r1 = 3,

### 2018-06-18 slides/slides.tex
Finish slides.

@@
 \documentclass[aspectratio=149]{beamer}
+
+%\def\IsRus{1}
+\newcommand{\double}[2]{\ifdefined\IsRus{#2}\else{#1}\fi}
 
@@
 \usepackage{colortbl}
-
 \definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
@@
 
-\usepackage{fancyvrb}%fancy verbatim
-%\makeatletter
-%\newcommand{\verbatimfont}[1]{\renewcommand{\verbatim@font}{\ttfamily\scriptsize#1}}
-%\makeatother
-% redefine \VerbatimInput
-\RecustomVerbatimCommand{\VerbatimInput}{VerbatimInput}{fontsize=\scriptsize}
-
-\usepackage{pmboxdraw}
-\usepackage{newunicodechar}
-\newunicodechar{└}{\textSFii}
-\newunicodechar{├}{\textSFviii}
-\newunicodechar{─}{\textSFx}
-
 \usepackage[export]{adjustbox} % loads also graphicx
@@
 \addbibresource{../bib/ms-thesis.bib}
+\renewcommand*{\bibfont}{\scriptsize}
 
@@
 %\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}
-\title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}
+\title{\double{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}{Автоматический~анализ~слабых~моделей памяти~параллельного~программирования}}
 
@@
 %\subtitle{Optional Subtitle}
-\author{\textbf{Artem Yushkovskiy}\inst{1,2} \\ 
+\author{
+\double{
+\textbf{Artem Yushkovskiy}\inst{1,2} \\ 
 {\scriptsize MSc Candidate}
@@
 \hphantom{Supervis} \textbf{Docent Igor I. Komarov}\inst{2} }
-}%\small
-%{\footnotesize%\raggedleft%does not work
-%Supervisor 1: Assoc. Prof. Keijo Heljanko\inst{1} \newline
-%Supervisor 2: Docent Igor I. Komarov\inst{2}
-%}
+}
+{\textbf{Юшковский А. В.}\inst{1,2} \\ 
+{\scriptsize магистрант}
+\\ \vspace{1em}
+{\footnotesize\raggedleft Руководители: \textbf{проф. Кеийо Хельянко}\inst{1} \newline
+\hphantom{Руководители} \textbf{доц. Комаров И. И.}\inst{2} }
+}
+}
 
@@
   \inst{1}%
-  \textbf{Aalto University} (Espoo, Finland) \\
+  \double{\textbf{Aalto University} (Espoo, Finland) \\
     Department of Computer Science, \\
     School of Science
+  }
+  {\textbf{Университет Аалто} (Эспоо, Финляндия) \\
+    Факультет Информатики, Школа Наук
+  }
   \and
   \inst{2}%
-  \textbf{ITMO University} (Saint Petersburg, Russia) \\
+  \double{\textbf{ITMO University} (Saint Petersburg, Russia) \\
     Faculty of Information Security \\
@@
   }
+  {\textbf{Университет ИТМО} (Санкт-Петебург, Россия) \\
+    Факультет Безопасности Информационных Технологий
+  }
+}
 
-\date{\scriptsize Espoo, Saint Petersburg, 2018}
+\date{\scriptsize \double{Espoo, Saint Petersburg\\2018}{Эспоо, Санкт Петербург\\2018}}
 % - Either use conference name or its abbreviation.
@@
 
-%\begin{frame}{Outline}
-%  \tableofcontents
-%\end{frame}
-
-%\section{Introduction}
-%\subsection{The problem}
-
-\begin{frame}{Task statement \rus{(Цель работы)}}
+\begin{frame}{\double{Task statement}{Цель работы}}
 \begin{itemize}
-\item Improve the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by extending support for the input language.
+\item \double{Improve the proof-of-concept memory model-aware analysis tool \tool{Porthos}~\cite{Porthos17a} by extending support for the input language.}
+{Усовершенствовать прототип статического анализатора кода  \tool{Porthos}~\cite{Porthos17a}, учитывающий при анализе модель памяти вычислительной среды, путем расширения поддержки входного языка.}
 \end{itemize}
@@
 
-\begin{frame}{Task specification \rus{(Задачи работы)}}
+\begin{frame}{\double{Task specification}{Задачи работы}}
 \begin{itemize}
-\item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
-\item Review existing tools for memory model-aware analysis;
+\item \double{Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};}
+{Изучить существующие подходы к автоматическому проведению анализа кода параллельных программ с учетом модели памяти вычислительной среды~\cite{alglave2010shared};}
+\item \double{Review existing tools for memory model-aware analysis;}
+{Исследовать существующие инструменты анализа кода с учетом модели памяти;}
 %\item Examine the architecture of \texttt{Porthos}, its strengths and weaknesses;
 %\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
-\item Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};
-\item Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};
-\item Enhance performance, extensibility, reliability and maintainability of the tool.
+\item \double{Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};}
+{Разработать инфраструктуру C-компилятора как часть модуля абстрактной интерпретации нового анализатора \tool{PorthosC};}
+\item \double{Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};}
+{Адаптировать схему кодирования произвольного графа потока управления в SMT-формулу;}
+\item \double{Enhance performance, extensibility, reliability and maintainability of the tool.}
+{Сохранить и улучшить функциональные характеристики анализатора (производительность, расширяемость, надежность и поддерживаемость).}
 \end{itemize}
@@
 
-\begin{frame}{Verification of concurrent software} {Motivating example: Write-write reordering (compiler  relaxations)}
+\begin{frame}{\double{Verification of concurrent software}{Верификация параллельных программ}}{\double{Motivating example: Write-write reordering (compiler  relaxations)}{Пример: Перестановка типа запись-запись (оптимизация на уровне компилятора)}}
 %1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
@@
 \only<1>{\caption*{~\\~}}
-\only<2>{\caption*{Single thread\\(no concurrency)}}
-\only<3>{\caption*{Sequential Consistency\\(classical concurrency)}}
+\only<2>{\double{\caption*{Single thread\\(no concurrency)}}{\caption*{Один поток\\(отсутствие параллелизма)}}}
+\only<3>{\double{\caption*{Sequential Consistency\\(classical concurrency)}}{\caption*{Последовательная Согласованность\\(классический параллелизм)}}}
 \only<4->{\caption*{~\\~}}
@@
 %instructions.
-\caption*{Total Store Order\\(e.g., x86)}
+\double{\caption*{Total Store Order\\(e.g., x86)}}{\caption*{Линейный порядок записи\\(напр., x86)}}
 \begin{tabular}{ | l r | l r | l r } % >{\columncolor{ltgrey}}
@@
 
-\begin{frame}{Verification of concurrent software} {Motivating example: Store buffering (hardware relaxations)}
+\begin{frame}{\double{Verification of concurrent software}{Верификация параллельных программ}} {\double{Motivating example: Store buffering (hardware relaxations)}{Пример: Буферизация записи (оптимизация на уровне процессора)}}
 \begin{center}
@@
 \includegraphics[scale=0.37]{img/x86-arch-full.png}
-\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
+\double{\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}}{\caption{Модель абстрактной машины процессора x86~\cite{sewell2010x86}}}
 \only<2->{
@@
 
-\begin{frame}{Weak memory model-aware analysis} {Event-based program representation}
+\begin{frame}{\double{Weak memory model-aware analysis}{Статический анализ с учетом слабой модели памяти}} {\double{Event-based program representation}{Событийная модель кода}}
 
@@
   \addtolength{\tabcolsep}{0pt}
-  \item \textbf{Event} $\in \mathbb{E}$, a low-level primitive operation:
+  \item \textbf{\double{Event}}{\textbf{Событие}} $\in \mathbb{E}$, \double{a low-level primitive operation}{низкоуровневая примитивная операция}:
     \begin{itemize}
-      \item \textit{memory event} $\in \mathbb{M} = \mathbb{R} \cup \mathbb{W}$: access to a local/shared memory, \\
-      \item \textit{computational event} $\in \mathbb{C}$: computation over local memory, and \\
-      \item \textit{barrier event} $\in \mathbb{B}$: synchronisation fences;
+      \item \textit{\double{memory event}{событие памяти}} $\in \mathbb{M} = \mathbb{R} \cup \mathbb{W}$: access to a local/shared memory, \\
+      \item \textit{\double{computational event}{событие вычисления}} $\in \mathbb{C}$: \double{computation over local memory, and}{вычисление в локальной памяти} \\
+      \item \textit{\double{barrier event}{событие синхронизации}} $\in \mathbb{B}$: \double{synchronisation fences}{барьеры синхронизации};
      \end{itemize}
-  \item \textbf{Relation} $\subseteq \mathbb{E} \times \mathbb{E}$: 
+  \item \textbf{\double{Relation}{Отношение}} $\subseteq \mathbb{E} \times \mathbb{E}$: 
     \begin{itemize}
-    \item \textit{basic relations}:%, computed from the program model:
+    \item \textit{\double{basic relations}{базовые отношения}}:%, computed from the program model:
       \begin{itemize}
-        \item \textit{program-order} relation $\texttt{po} \subseteq \mathbb{E} \times \mathbb{E}$: (control-flow),
-        \item \textit{read-from} relation $\texttt{rf} \subseteq \mathbb{W} \times \mathbb{R}$: (data-flow), and
-        \item \textit{coherence-order} relation $\texttt{co} \subseteq \mathbb{W} \times \mathbb{W}$: (data-flow);
+        \item \textit{program-order} \double{relation}{отношение} $\texttt{po} \subseteq \mathbb{E} \times \mathbb{E}$: (\double{control-flow}{поток управления}),
+        \item \textit{read-from} \double{relation}{отношение} $\texttt{rf} \subseteq \mathbb{W} \times \mathbb{R}$: (\double{data-flow, and}{поток данных, и})
+        \item \textit{coherence-order} \double{relation}{отношение} $\texttt{co} \subseteq \mathbb{W} \times \mathbb{W}$: (\double{data-flow}{поток данных});
       \end{itemize}
-    \item \textit{derived relations}:%, computed from the basic relations using the following operators:
+    \item \textit{\double{derived relations}{производные отношения}}:%, computed from the basic relations using the following operators:
       \begin{itemize}
-        \item \textit{union} \texttt{r1\,|\,r2},
-        \item \textit{sequence} \texttt{r1} ; \texttt{r2},
+        \item \textit{\double{union}{объединение}} \texttt{r1\,|\,r2},
+        \item \textit{\double{sequence}{последовательность}} \texttt{r1} ; \texttt{r2},
         %\item \textit{inverse} \texttt{r}\^{\texttt{-1}},
-        \item \textit{transitive closure} \texttt{r+},
+        \item \textit{\double{transitive closure}{транзитивное замыкание}} \texttt{r+},
         \item $\cdots$;
@@
   %\item \textbf{Candidate execution}, a sequence of guesses of relations;
-  \item \textbf{Assertion} over relations or sets of events: 
+  \item \double{\textbf{Assertion} over relations or sets of events}{\textbf{Ограничение} модели памяти, сформулированное в терминах отношений или множеств событий} : 
     \begin{itemize}
-    \item \textit{acyclicity}, \textit{irreflexivity} or \textit{emptiness} 
+    \item \textit{\double{acyclicity}{ацикличность}}, \textit{\double{irreflexivity}{антирефлексивность}} or \textit{\double{emptiness}{отсутствие элементов во множестве}}.
     \end{itemize}
@@
 
-\begin{frame}{Weak memory model-aware analysis} {Testing candidate executions}
+\begin{frame}{\double{Weak memory model-aware analysis}{Статический анализ с учетом слабой модели памяти}} {\double{Testing candidate executions}{Тестирование путей выполнения программы}}
 \vspace{10pt}
@@
 \scriptsize
-SC model: \\
+SC \double{model}{модель}: \\
 ... \\
@@
 \includegraphics[width=\linewidth]{img/candidate.png}
-\caption{The four candidate executions allowed under x86-TSO}
+\double{\caption{The four candidate executions allowed under x86-TSO}}{\caption{Четыре возможных выполнения программы, допустимых моделью x86-TSO}}
 \end{figure}
@@
 \includegraphics[width=\linewidth]{img/candidate-red.png}
-\caption{The four candidate executions allowed under x86-TSO}
+\double{\caption{The four candidate executions allowed under x86-TSO}}{\caption{Четыре возможных выполнения программы, допустимых моделью x86-TSO}}
 \end{figure}
@@
 
-\begin{frame}{Portability analysis} {The \tool{Porthos} tool}
+\begin{frame}{\double{Portability analysis}{Анализ портируемости}} {\double{The \tool{Porthos} tool}{Анализатор \tool{Porthos}}}
 \begin{minipage}{.49\textwidth}
@@
   \includegraphics[height=0.4\textheight,keepaspectratio]{img/the-problem.png}
-  \caption{The illustration of the portability problem~\cite{Porthos17slides}}
+  \caption{\double{The illustration of the portability problem~\cite{Porthos17slides}}{Иллюстрация проблемы определения портируемости программы~\cite{Porthos17slides}}}
   \end{figure}
@@
   \begin{itemize}
-    \begin{definition}[Portability~\cite{Porthos17a}]
+    \begin{definition}[\double{Portability}{Портируемость}~\cite{Porthos17a}]
     %SAY: Let $\textit{cons}_{\mathcal{M}}(P)$ be a function that calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$
     %Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
-    The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target one $\mathcal{M_T}$ if
+    \double{The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target model $\mathcal{M_T}$ if}
+    {Программа $P$ называется портируемой с платформы $\mathcal{M_S}$ на платформу $\mathcal{M_T}$, если}
     $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
     \end{definition}
-    \item Portability as an SMT-based bounded reachability problem: \\
+    \item \double{Portability as an SMT-based bounded reachability problem}{Портируемость как проблема ограниченной достижимости на основе SMT}: \\
     $\phi = $\only<3>\yellowemph{$\phi_{CF}$}$\land $\only<3>\yellowemph{$\phi_{DF}$}$\land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$
-    \item $\texttt{SAT}(\phi) \Longrightarrow $ portability bug
+    \item $\texttt{SAT}(\phi) \Longrightarrow $ \double{portability bug}{ошибка портирования}
   \end{itemize}
@@
 
-\begin{frame}{\tool{PorthosC}: Architecture}{Main components}
+\begin{frame}{\double{\tool{PorthosC}: Architecture}{\tool{PorthosC}: Архитектура}} {\double{Main components}{Основные компоненты}}
 \vspace{-30pt}
@@
 \begin{itemize}
-\item The enhanced version of \tool{Porthos}, that is able to analyse C programs, has got a new name \textit{PorthosC}.
+\item \double{The enhanced version of \tool{Porthos}, that is able to analyse C programs, has got a new name \textit{PorthosC}.}{Усовершенственная версия \tool{Porthos}, способная анализировать программы на языке программирования С, была названа \tool{PorthosC}.}
 \end{itemize}
@@
 
-\begin{frame}{\tool{PorthosC}: Architecture}{Abstract interpretation module}
+\begin{frame}{\double{\tool{PorthosC}: Architecture}{\tool{PorthosC}: Архитектура}} {\double{Abstract interpretation module}{Модуль абстрактной интерпретации}}
   \vspace{-40pt}
@@
 \begin{itemize}
-\item The abstract interpretation module is an essential part of the compiler of C code into the event-flow graph representation (\texttt{X-graph}).
+\item \double{The abstract interpretation module is an essential part of the compiler of C code into the event-flow graph representation}{Модуль абстрактной интерпретации является ключевым компонентом компилятора, преобразующего программы на С в низкоуровневое событийное представление} (\texttt{X-graph}).
 \end{itemize}
@@
 
-\begin{frame}{\tool{Porthos\,v1}: The input language}
+\begin{frame}{\double{\tool{Porthos\,v1}: The input language}{\tool{Porthos\,v1}: Входной язык}}
 {
 \only<1-2>{ }
-\only<3>{Static syntactic determination of the variables kind}
-\only<4>{Lack of support for functions invocations}
-\only<5>{Lack of support for unconditional jumps}
+\only<3>{\double{Static syntactic determination of the variables kind}{Статическое определение вида переменных на синтаксическом уровне}}
+\only<4>{\double{Lack of support for functions invocations}{Отсутствие поддержки вызовов функций}}
+\only<5>{\double{Lack of support for unconditional jumps}{Отсутствие поддержки безусловных переходов}}
 }
 {\noindent\centering
-\begin{minipage}{.45\textwidth}
+\begin{minipage}{.46\textwidth}
   \begin{figure}
@@
     \only<5>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-old/example-old.c}}
-  \caption{A program example in the \tool{Porthos\,v1} input language}
+  \double{\caption{A program example in the \tool{Porthos\,v1} input language}}{\caption{Пример программы на входном языке анализатора \tool{Porthos\,v1}}}
   \end{figure}
@@
 %
-\begin{minipage}{.45\textwidth}
+\begin{minipage}{.47\textwidth}
   \begin{figure}
@@
   \vfill
-  \only<2->{\caption{A sketch of the \tool{Porthos\,v1} input language grammar}}
+  \only<2->{\double{\caption{A sketch of the \tool{Porthos\,v1} input language grammar}}{\caption{Набросок грамматики входного языка анализатора \tool{Porthos\,v1}}}}
   \end{figure}
@@
 
-\begin{frame}[label={example-new}]{\tool{PorthosC}: The input language}
+\begin{frame}[label={example-new}]{\double{\tool{PorthosC}: The input language}{\tool{PorthosC}: Входной язык}}
 {
 \only<1-2>{ }
-\only<3>{Pre-compilation phase for the syntactic determination of the variables kind}
-\only<4>{Support for functions invocations: The invocation hooking mechanism}
-\only<5>{Support for arbitrary control-flow jumps}
+\only<3>{\double{Pre-compilation phase for the syntactic determination of the variables kind}{Стадия предкомпиляции для синтаксического определения вида переменных}}
+\only<4>{\double{Support for functions invocations: The invocation hooking mechanism}{Поддержка вызовов функций: Механизм перехвата вызовов}}
+\only<5>{\double{Support for an arbitrary control-flow}{Поддержка произвольного ветвления графа потока управления}}
 }
 {\noindent\centering
-\begin{minipage}{.43\textwidth}
+\begin{minipage}{.44\textwidth}
   \begin{figure}
@@
     \only<5>{\lstinputlisting[language=Java,basicstyle=\ttfamily\tiny,columns=fullflexible]{img/lst/input-new/example-new-4.c}}
-  \caption{A program example in the \tool{Porthos\,v1} input language}
+  \double{\caption{A program example in C}}{\caption{Пример программы на языке~С}}
   \end{figure}
@@
 %
-\begin{minipage}{.55\textwidth}
+\begin{minipage}{.54\textwidth}
 %  \only<1>{\begin{itemize}
@@
   }
-  \only<2->{\caption{The compiled event-flow graph}}
+  \only<2->{\double{\caption{The compiled event-flow graph}}{\caption{Скомпилированный граф потока событий}}}
   \end{figure}
@@
 
-\begin{frame}{\texttt{PorthosC}: The new control-flow encoding scheme}
+\begin{frame}{\double{\texttt{PorthosC}: The new control-flow encoding scheme}{\texttt{PorthosC}: Новая схема кодирования потока управления}}
 \begin{figure}
@@
 \end{minipage}
-\caption{Example of encoding for the control-flow of the event-flow graph}
+\double{\caption{Example of encoding for the control-flow of the event-flow graph}}{\caption{Пример кодирования потока управления с произвольным ветвлением}}
 \end{figure}
@@
 
-\begin{frame}{\texttt{PorthosC}: The new DFS-based \texttt{X-graph} unrolling scheme}
+\begin{frame}{\double{\texttt{PorthosC}: The new DFS-based \texttt{X-graph} unrolling scheme}{\texttt{PorthosC}: Новая схема развертки циклов, основанная на обходе графа в глубину}}
 \only<1>{
@@
 }
-\caption{Example of the flow graph unrolling up to bound $k = 6$}
+\double{\caption{Example of the flow graph unrolling up to bound $k = 6$}}{\caption{Пример развертки цикла до границы $k=6$}}
 \end{figure}
@@
 
-\begin{frame}{Evaluation}{The new unrolling scheme}
-\begin{minipage}{.23\linewidth}
+\begin{frame}{\double{Evaluation}{Анализ программы}}{\double{The new unrolling scheme}{Новая схема развертки циклов}}
+\begin{minipage}{.3\linewidth}
 \hfill
@@
 %
-\noindent\begin{minipage}{.75\linewidth}
+\noindent\begin{minipage}{.65\linewidth}
 \begin{figure}
@@
   }
-\caption{Illustration of differences in unrolling schemes of \tool{PorthosC}~(left) and \tool{Porthos\,v1}~(right)}
+\double{\caption{Illustration of differences in unrolling schemes of \tool{PorthosC}~(left) and \tool{Porthos\,v1}~(right)}}{\caption{Иллюстрация различий в алгоритмах развертки циклов используемых \tool{PorthosC}~(слева) и \tool{Porthos\,v1}~(справа)}}
 \end{figure}
@@
 
-\begin{frame}{Evaluation}{Overhead of the new unrolling scheme: Number of events}
+\begin{frame}{\double{Evaluation}{Анализ программы}}{\double{Overhead of the new unrolling scheme: Number of events}{Сравнение схем развертки циклов: Количество событий}}
 \begin{figure}[t]
@@
 \includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/new-e.png}
-\caption{The graph unrolled by \tool{PorthosC}}
-\label{dep:events:new}
+\double{\caption{The graph unrolled by \tool{PorthosC}}}{\caption{Граф, развернутый анализатором \tool{PorthosC}}}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/old-e.png}
-\caption{The graph unrolled by \tool{Porthos\,v1}}
-\label{dep:events:old}
+\double{\caption{The graph unrolled by \tool{Porthos\,v1}}}{\caption{Граф, развернутый анализатором \tool{Porthos\,v1}}}
 \end{subfigure}
 %
-\caption{Dependency of unrolled events number on the unrolling bound $k$}
-\label{dep:events}
+\double{\caption{Dependency of unrolled events number on the unrolling bound $k$}}{\caption{Зависимость количества событий от границы развертки циклов $k$}}
 \end{figure}
@@
 
-\begin{frame}{Evaluation}{Overhead of the new unrolling scheme: Execution time}
+\begin{frame}{\double{Evaluation}{Анализ программы}}{\double{Overhead of the new unrolling scheme: Execution time}{Сравнение схем развертки циклов: Время кодирования программы}}
 \begin{figure}[t]
@@
 \includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/new-t.png}
-\caption{Program domain encoding time of the graph unrolled by \tool{PorthosC}}
-\label{dep:time:new}
+\double{\caption{Program domain encoding time of the graph unrolled by \tool{PorthosC}}}{\caption{Время кодирования программы в SMT-формулы, произведенное анализатором \tool{PorthosC}}}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth,keepaspectratio]{../img/my/performance/new/old-t.png}
-\caption{Program domain encoding time of the graph unrolled by \tool{Porthos\,v1}}
-\label{dep:time:old}
+\double{\caption{Program domain encoding time of the graph unrolled by \tool{PorthosC}}}{\caption{Время кодирования программы в SMT-формулы, произведенное анализатором \tool{Porthos\,v1}}}
 \end{subfigure}
 %
-\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
-\label{dep:time}
+\double{\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}}{\caption{Зависимость времени кодирования программы от границы развертки циклов $k$}}
 \end{figure}
@@
 
-\begin{frame}{Summary}
+\begin{frame}{\double{Summary}{Заключение}}
 \begin{itemize}
-\item The result of the work is the generalised framework for memory model-aware analysis \textbf{\tool{PorthosC}};
+\item \double{The result of the work is the generalised framework for memory model-aware analysis \textbf{\tool{PorthosC}}}{Результатом выпускной квалификационной работы является платформа для проведения статического анализа кода с учетом слабых моделей памяти сред выполнения \textbf{\tool{PorthosC}}};
   \begin{itemize}
-  \item The \textit{input language} has been extended to a higher subset of C language (including unconditional control-flow jumps);
-  \item The old \textit{architecture} of \texttt{Porthos} has been revised and redesigned for \texttt{PorthosC};
-  \item The \textit{new unrolling scheme} produces complete state space of the analysing program within the user-defined bound $k$, though the encoding time growth rapidly (exponentially) as the unrolling bound grows.
-  \item The \textit{invocation hooking mechanism} serves as a knowledge base for the program domain, it is to be filled and extended in future.
+  \item \double{The \textit{input language} has been extended to a higher subset of C language (including unconditional control-flow jumps)}{В рамках работы над анализатором была расширена поддержка \textit{входной языка С} (включая поддержку безусловных переходов)};
+  \item \double{The old \textit{architecture} of \texttt{Porthos} has been revised and redesigned for \texttt{PorthosC}}{Была пересмотрена и модифицирована общая \textit{архитектура} анализатора};
+  \item \double{The \textit{new unrolling scheme} produces complete state space of the analysing program within the user-defined bound $k$, though the encoding time growth rapidly (exponentially) as the unrolling bound grows}{Измененная \textit{схема развертки циклов}, работающая на нерекурсивной структуре графа событий, производит \textit{полное множество} выполнений программы, что улучшает качество производимого анализа};
+  \item \double{The \textit{invocation hooking mechanism} is an important part of the abstract interpretation engine, that serves as a knowledge base for the program domain and increases the extensibility of the tool}{\textit{Механизм перехвата вызовов}, являющийся важным элементом модуля абстрактной интерпретации, служит в качестве базы знаний семантики функций и таким образом повышает расширяемость анализатора}.
   \end{itemize}
@@
 
-\begin{frame}{Directions for future work}
+\begin{frame}{\double{Directions for future work}{Направления работы в будущем}}
 \begin{itemize}
-\item Extending the \textit{knowledge base} of domain-specific functions to model synchronisation primitives;
-\item Supporting \textit{new input languages} (e.g., different assembly languages);
-\item Supporting \textit{complex data types} (s.a. arrays, pointers, structures, etc.);
-\item Adding the \textit{inter-procedural analysis mode};
-\item Handling the \textit{state explosion problem} (with standard model-checking techniques adjusted by the information about the weak memory model of the execution environment).
+\item \double{Extending the \textit{knowledge base} of domain-specific functions to model synchronisation primitives}{Расширение \textit{базы знаний} функций для моделирования различных примитивов синхронизации};
+\item \double{Supporting \textit{new input languages} (e.g., different assembly languages)}{Поддержка новых \textit{входных языков} (напр., языков ассемблера различных архитектур)};
+\item \double{Supporting \textit{complex data types} (s.a. arrays, pointers, structures, etc.)}{Поддержка \textit{комплексных структур данных} (таких как массивы, структуры, указатели, и т.д.)};
+\item \double{Adding the \textit{inter-procedural analysis mode}}{Добавление режима \textit{кросс-процедурного анализа}};
+\item \double{Handling the \textit{state explosion problem} (with standard model-checking techniques adjusted by the information about the weak memory model of the execution environment)}{Решение проблемы \textit{комбинаторного взрыва пространства состояний} (с помощью стандартных техник анализа кода, модифицированных с целью учитывания модели памяти среды выполнения)}.
 \end{itemize}
@@
 
+\begin{frame}{\double{Thank you for attention}{Спасибо за внимание}}
+  \centering
+  \includegraphics[height=.8\textheight]{img/thank-you-smirnitsky.jpg}
+\end{frame}
+
 % All of the following is optional and typically not needed. 
-\appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{For Further Reading}
+%\appendix
+%\section<presentation>*{\appendixname}
+%\subsection<presentation>*{For Further Reading}
 
 \begin{frame}[allowframebreaks]
-  \frametitle<presentation>{Bibliography}
+  %\frametitle<presentation>{Bibliography}
+  \frametitle{Bibliography}
   \printbibliography
